### 2016-11-25 amenIntro/amenIntro.Rmd
push up most of intro, waiting on mod results

---
title: "Networks Matter...or some nonsense like that"
output: 
      ioslides_presentation:
            smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Purpose

* Why are we meeting during Thanksgiving weekend...I didn't realize the date and none of you objected :)
* Anyhow...we are here to talk about conducting a meta-analysis of dyadic studies in political science using the Additive and Multiplicative Effects Model for Networks (AMEN)
* What I hope we find from this project is that using standard GLM approaches on dyadic data will produce
      - biased estimates of the effect of independent variables
      - uncalibrated confidence intervals
      - and poor predictive performance.
* Anticipated outputs: 
      - One big meta-analysis piece in which we show the items in the list above to be true on average
      - A number of small replication pieces

## Outline

* **Brief** background discussion of AMEN based on [Hoff's vignette](https://arxiv.org/pdf/1506.08237v1.pdf) and a [joint piece with me and Mike](https://arxiv.org/pdf/1611.00460v1.pdf)
* Example application of AMEN, where we will   
      - Discuss how to structure input data
      - Extract and interpret model results
      - Conduct out-of-sample performance on longitudinal networks
* Discuss project management

## Dyadic data

* Any dataset in which the unit of observation is characterized by a pair of actors
      - MIDs
      - International economic agreements (PTAs, BITs)
      - Alliances (NATO)
      - Trade flows

## What makes dyadic data special

* Formation of a tie between any pair of actors is **not necessarily** independent of other ties
      - This is true for bilateral and multilateral events
* Why?

## Actors are part of a system ... a network

```{r, out.width= "700px",fig.align="center"}
knitr::include_graphics("dyaddata.png")
```

## Flavors of dependencies

* First order dependence
      - Often will find significant heterogeneity in activity levels across nodes
* Second order dependence
      - Reciprocity: Event sent from i $\rightarrow$ j is likely related to an event sent from j $\rightarrow$ i

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \color{red}{\mu} + \color{red}{e_{ij}} \\
      e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\Sigma_{ab}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* $\mu$ baseline measure of network activity, for the purpose of regression we turn this into $\beta^{T}X$
* $e_{ij}$ residual variation that we will use the SRM to decompose

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= \color{red}{a_{i} + b_{j}} + \epsilon_{ij} \\
      \color{red}{\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \}} &\sim N(0,\Sigma_{ab}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* row/sender effect ($a_{i}$) & column/receiver effect ($b_{j}$)
* Modeled jointly to account for correlation in how active an actor is in sending and receiving ties

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\color{red}{\Sigma_{ab}}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \color{red}{\Sigma_{ab}} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$ capture heterogeneity in the row and column means
* $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties)

## What to do: Social Relations Model

$$
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= a_{i} + b_{j} + \color{red}{\epsilon_{ij}} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\Sigma_{ab}) \\ 
      \color{red}{\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\}} &\sim N(0,\color{red}{\Sigma_{\epsilon}}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; & \color{red}{\Sigma_{\epsilon}} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
$$

* $\epsilon_{ij}$ captures the within dyad effect
* Second-order dependencies are described by $\sigma_{\epsilon}^{2}$
* Within dyad correlation, aka reciprocity, represented by $\rho$

## Network level dependencies

```{r, out.width= "700px",fig.align="center"}
knitr::include_graphics("homophStoch.png")
```

* Homophily
* Stochastic equivalence

## Latent variable models for networks

$$
\begin{aligned}
\text{Latent class model} \\
      &\alpha(u_{i}, u_{j}) = m_{u_{i}, u_{j}} \\
      &u_{i} \in \{1, \ldots, K \}, \; i \in \{1,\ldots, n\} \\
      &M \text{ a } K \times K \text{ symmetric matrix} \\
\text{Latent distance model} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = -|\textbf{u}_{i} - \textbf{u}_{j}| \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
\text{Latent factor model} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{T} \Lambda \textbf{u}_{j} \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
      &\Lambda \text{ a } K \times K \text{ diagonal matrix}
\end{aligned}
$$

## Our preferred approach

$$
\begin{aligned}
\text{Latent class model} \\
      &\alpha(u_{i}, u_{j}) = m_{u_{i}, u_{j}} \\
      &u_{i} \in \{1, \ldots, K \}, \; i \in \{1,\ldots, n\} \\
      &M \text{ a } K \times K \text{ symmetric matrix} \\
\text{Latent distance model} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = -|\textbf{u}_{i} - \textbf{u}_{j}| \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
\color{red}{\text{Latent factor model}} \\
      &\alpha(\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{T} \Lambda \textbf{u}_{j} \\
      &\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \\
      &\Lambda \text{ a } K \times K \text{ diagonal matrix}
\end{aligned}
$$

## Putting it all together: AME

$$
\begin{aligned}
      y_{ij} &= g(\theta_{ij}) \\ 
      &\theta_{ij} = \beta^{T} \mathbf{X}_{ij} + e_{ij} \\
      &e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \textbf{u}_{i}^{T} \textbf{D} \textbf{v}_{j} \\
\end{aligned}
$$

* $a_{i} + b_{j}  + \epsilon_{ij}$, are additive random effects and account for sender, receiver, and within-dyad dependence
* multiplicative effects, $\textbf{u}_{i}^{T} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns that are left over in $\theta$ after accounting for any known covariate information

## Benefits of this approach

* At its core, AME is just a GLM with random effects used to ensure that we can treat dyadic observations as conditionally independent
* AME can be used: 
      - on both undirected and directed data, 
      - on longitudinal and static networks,
      - and on a variety of distribution types we commonly encounter in political science (binomial, gaussian, and ordinal).

## How to use it...first installing

```{r,echo=TRUE, message=FALSE, warning=FALSE, suppressPackageStartupMessages=TRUE}
library(devtools)
devtools::install_github('s7minhas/amen')
library(amen)
```

**DON'T USE THE VERSION FROM CRAN**

## Structuring input data

* This is going to be a bit different than what is described in [Hoff's vignette](https://arxiv.org/pdf/1506.08237v1.pdf). 
      - As of this week, AMEN can now be estimated on networks where the actor composition changes over time, and to make that change we switch to using **lists** instead of **arrays**.
* In undirected longitudinal setting three inputs:
      - `Y`: a T length list of n x n adjacency matrices (see Table 2)
      - `Xdyad`: A T length list of n x n x p arrays
      - `Xrow`: A T length list of n x p matrices
* In directed longitudinal setting you can distinguish between:
      - Sender covariates `Xrow`
      - Receiver covariates `Xcol`

## Structuring dependent variable

```{r,echo=FALSE}
load('~/Dropbox/Research/netsMatter/replications/example/inputData/exampleData.rda')
```

```{r, echo=TRUE}
str(yL)
```

## More on dependent variable

```{r, echo=TRUE}
print( yL$'1985'[17:33,17:33]  )
```

## Dyadic covariate(s)

```{r, echo=TRUE}
str(xDyadL)
```

## More on dyadic covariate(s)

```{r, echo=TRUE}
print( round(xDyadL$'1985'[17:33,17:33,1],1)  )
```

## Actor level covariate(s)

```{r, echo=TRUE}
str(xNodeL)
```

## More on actor level covariate(s)

```{r, echo=TRUE}
print( xNodeL$'1985'[17:33,]  )
```

## Running an AME model

```{r, out.width= "600px",fig.align="center"}
knitr::include_graphics("ameRepL.png")
```

## Lets put it all together

* Undirected, `symmetric=TRUE`, binary, `model='bin'`, longitudinal AME model with actor random effects, `nvar=TRUE`, and a two dimensional multiplicative effect, `R=2`.

```{r, echo=TRUE,eval=FALSE}
fit=ame_repL(Y=yL, Xdyad=xDyadL, Xrow=xNodeL, Xcol=NULL,
      nvar=TRUE, R=2,
      model="bin",symmetric=TRUE,
      burn=500,nscan=1000,odens=10, # WILL NEED TO BE LARGER
      plot=FALSE, print=FALSE, 
      seed=6886 # ALWAYS SET A SEED
      )
```

```{r, echo=TRUE}
load('~/Dropbox/Research/netsMatter/replications/example/outputData/model_k2.rda')
```

## The output

```{r, out.width= "600px",fig.align="center"}
knitr::include_graphics("ameOut.png")
```

## Assessing convergence

## Did you set the right K?

## Exogenous covariates

```{r, echo=TRUE}
dim(fit$BETA)

stats = function(x){ c( mu=mean(x), q95  ) }
apply(BETA<)
```

## Out of sample model performance

### 2016-11-25 code/example/buildexampleData.R

@@ -1,0 +1,111 @@
+rm(list=ls())
+library(devtools) ; devtools::install_github('s7minhas/amen')
+library(amen)
+
+##############################
+# load some other data
+data(coldwar)
+
+# code modified from hoff vignette pg 43-44
+Y=sign( coldwar$cc ) ; n=nrow(Y) ; t=dim(Y)[3]
+Y[Y==1] = 0  ; Y[Y==-1] = 1
+unique(c(Y))
+# nodal covariates
+Xn = array(NA,dim=c(n,2,t),dimnames=list(rownames(Y),c('lgdp','polity'),NULL))
+for(x in 1:t){ Xn[,'lgdp',x] = log(coldwar$gdp[,x]) ; Xn[,'polity',x] = coldwar$polity[,x] }
+# dyadic covariates
+Xd=array(dim=c(nrow(Y),nrow(Y),1,t),dimnames=list(rownames(Y),rownames(Y),'ldist',NULL))
+Xd[,,'ldist',] = log(coldwar$distance)
+
+# change actor composition over time
+degYr = apply(Y, c(1,3), sum, na.rm=TRUE)
+noInt = names(rowSums(degYr)[rowSums(degYr)==0])
+noIntLate = names(rowSums(degYr[,5:8])[rowSums(degYr[,5:8])==0])
+noIntEarly = names(rowSums(degYr[,1:4])[rowSums(degYr[,1:4])==0])
+noIntLate = setdiff(noIntLate, noInt)
+noIntEarly = setdiff(noIntEarly, noInt)
+
+actors = rownames(Y)
+noIntNum = which(actors %in% noInt)
+noIntLateNum = which(actors %in% noIntLate)
+noIntEarlyNum = which(actors %in% noIntEarly)
+
+# modify Y matrix
+yL = lapply(1:t, function(x){ Y[,,x] }) ; names(yL) = dimnames(Y)[[3]]
+yL[[1]] = yL[[1]][ actors[-c(noIntNum,noIntEarlyNum[c(3,6,9)])], actors[-c(noIntNum,noIntEarlyNum[c(3,6,9)])] ]
+yL[[2]] = yL[[2]][ actors[-c(noIntNum,noIntEarlyNum[c(3,6)])], actors[-c(noIntNum,noIntEarlyNum[c(3,6)])]  ]
+yL[[3]] = yL[[3]][ actors[-c(noIntNum,noIntEarlyNum[c(3)])], actors[-c(noIntNum,noIntEarlyNum[c(3)])]  ]
[…]

### 2016-11-25 code/mansfield_milner_2012/data.R

@@ -1,0 +1,33 @@
+if(Sys.info()['user']=='s7m' | Sys.info()['user']=='janus829'){
+	source('~/Research/amenRepl/code/mansfield_milner_2012/setup.R') }
+
+# load data
+data = foreign::read.dta(file=paste0(dataPath, 'dyadicdata.dta'))
+
+baseVars <- c('nohs_rat2_onset2', 'polity2_04', 'polconiiiA', 'ptadum_new', 'llntradenew', 'llnGDP_gled08', 'ldGDP_gled08V2', 'larmconflict', 'latopally', 'fmrcol_new', 'lcontig_new', 'lndistance', 'lheg_new', 'pcw89', 'llnGDPratio_g', 'onsetperc2', 'lnewgatt', 'nohs2_ptaonspl_r2', 'nohs2_ptaonsp1_r2', 'nohs2_ptaonsp2_r2', 'nohs2_ptaonsp3_r2', 'region', 'dyadid')
+baseData <- na.omit(data[, baseVars])
+
+# GLM binomial with logit link, region fixed efFects
+baseModel <- glm(
+	nohs_rat2_onset2 ~ polity2_04 + polconiiiA + ptadum_new + llntradenew + 
+	llnGDP_gled08 + ldGDP_gled08V2 + larmconflict + latopally + fmrcol_new + 
+	lcontig_new + lndistance + lheg_new + pcw89 + llnGDPratio_g + onsetperc2 + 
+	lnewgatt + nohs2_ptaonspl_r2 + nohs2_ptaonsp1_r2 + nohs2_ptaonsp2_r2 + 
+	nohs2_ptaonsp3_r2 + I(region), data = baseData, family = binomial(link = 'logit'))
+
+summary(baseModel)
+
+
+# check countries by year
+cntriesYr = lapply(unique(data$year), function(yr){
+	return( unique( data$ccode1[data$year==yr] ) ) })
+
+
+test = data[which(data$year %in% c(min(data$year), max(data$year)) ), ]
+
+
+data(YX_bin_long) 
+
+YX_bin_long$X[1,3,1,1] = NA
+
+fit<-ame_rep(YX_bin_long$Y,YX_bin_long$X,burn=5,nscan=5,odens=1,model="bin")

### 2016-11-25 code/mansfield_milner_2012/setup.R

@@ -1,0 +1,36 @@
+if( Sys.info()['user']=='janus829' | Sys.info()['user']=='s7m' ){
+	dPath='~/Dropbox/Research/amenRepl/'
+	dataPath=paste0(dPath, 'data/mansfield_milner_2012/')
+	graphicsPath=paste0(dPath, 'graphics/mansfield_milner_2012/')
+	resultsPath=paste0(dPath, 'results/mansfield_milner_2012/')
+	gPath='~/Research/amenRepl/'
+}
+
+# install/load libraries
+loadPkg=function(toLoad){
+	for(lib in toLoad){
+	  if(!(lib %in% installed.packages()[,1])){ 
+	    install.packages(lib, repos='http://cran.rstudio.com/') }
+	  suppressMessages( library(lib, character.only=TRUE) )
+	}
+}
+
+# some necessary libs
+loadPkg(c(
+	'foreign',
+	'reshape2', # data management
+	'ggplot2', 'latex2exp', 'Cairo',	# plotting
+	'xtable', # tables
+	'devtools' # loading git packages
+	))
+
+# load amen
+devtools::install_github('s7minhas/amen') ; library(amen)
+
+# Set a theme for gg
+theme_set(theme_bw())
+
+# misc
+char = function(x){ as.character(x) }
+num = function(x){ as.numeric(char(x)) }
+trim = function (x) { gsub("^\\s+|\\s+$", "", x) }

### 2016-11-25 amenIntro/amenIntro.Rmd
init review done

@@
 * Anyhow...we are here to talk about conducting a meta-analysis of dyadic studies in political science using the Additive and Multiplicative Effects Model for Networks (AMEN)
-* What I hope we find from this project is that using standard GLM approaches on dyadic data will produce
+* What I hope we find from this project is that using standard GLM approaches on dyadic data produces
       - biased estimates of the effect of independent variables
@@
       - A number of small replication pieces
+      - New disciples to spread the gospel
 
@@
 
-## Assessing convergence
-
-## Did you set the right K?
-
-## Exogenous covariates
+## First step...did the model converge
 
 ```{r, echo=TRUE}
-dim(fit$BETA)
+fit$BETA[1:3,]
+```
 
-stats = function(x){ c( mu=mean(x), q95  ) }
-apply(BETA<)
+```{r, echo=TRUE, eval=FALSE}
+library(reshape2) ; library(ggplot2)
+
+beta = data.frame(fit$BETA) ; names(beta)[ncol(beta)] = 'distance'
+beta$iter = 1:nrow(beta)*10
+
+ggBeta = reshape2::melt(beta,id='iter')
+ggplot(ggBeta, aes(x=iter,y=value)) + 
+	geom_line() + 
+	facet_wrap(~variable,nrow=2,scales='free_y')
+```
+
+## First step...did the model converge
+
+```{r, echo=FALSE}
+library(reshape2) ; library(ggplot2)
+beta = data.frame(fit$BETA) ; names(beta)[ncol(beta)] = 'distance'
+beta$iter = 1:nrow(beta)*10
+ggBeta = reshape2::melt(beta,id='iter')
+ggplot(ggBeta, aes(x=iter,y=value)) + geom_line() + facet_wrap(~variable,nrow=2,scales='free_y')
+```
+
+## Are we capturing the dependencies in our data?
+
+* Lets find out by drawing from the posterior predictive distribution of our model
+
+```{r, echo=TRUE}
+fit$GOF[1:3,]
+```
+
+```{r,message=FALSE, warning=FALSE, suppressPackageStartupMessages=TRUE,fig.width=8, fig.height=2}
+act = fit$GOF[1,-c(2:3)] ; gof = fit$GOF[-1,-c(2:3)]
+
+ggGof = melt(gof)[,-1]
+ggGof$act = act[match(ggGof$Var2, names(act))]
+
+ggplot(ggGof, aes(x=value)) + geom_histogram() + 
+	facet_wrap(~Var2,nrow=1,scales='free') + 
+	geom_vline(aes(xintercept=act),color='red')
+```
+
+## Parameter estimates
+
+```{r, echo=TRUE}
+library(magrittr) # so it's easier to see
+summStats = function(x){ c(mu=mean(x), sd=sd(x), quantile(x, probs=c(0.025,0.975)))  }
+
+fit$BETA %>% apply(., 2, summStats) %>% round(., 3) %>% t(.)
+```
+* You would then compare these with the corresponding GLM model.
+
+## Visualizing third order dependence
+
+```{r, echo=TRUE, fig.height=5, fig.width=5, fig.align='center'}
+circplot(Y=yL[[length(yL)]], U=fit$U, lcol='white')
 ```
@@
 ## Out of sample model performance
+
+* Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
+* For each $s \in \{1, \ldots, S\}$:
+	- Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
+	- For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
+
+* Code to perform this analysis for the example is shown in `git/netsMatter/code/example/runOutPerf.R`
+* When we get out of sample results we will be evaluating across two metrics:
+	- Receiver Operating Characteristic curves
+	- Precision-Recall curves
+
+## There will be a lot of models to run...
+
+```{r, eval=FALSE, echo=TRUE}
+# run ame in parallel # mcmc params
+imps = 100000
+brn = 50000
+ods = 10
+latDims = 1:4
+seed=6886
+
+# Run amen in parallel
+library(doParallel) ; library(foreach)
+cl=makeCluster(4) ; registerDoParallel(cl)
+foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {
+	ameFit = ame_repL(
+		Y=yL,Xdyad=xDyadL,Xrow=xNodeL,Xcol=NULL, model="bin",symmetric=TRUE,
+		R=latDims[ii], 
+		nscan=imps, seed=seed, burn=brn, odens=ods, 
+		plot=FALSE, print=FALSE) 	
+	save(ameFit, 
+		file='~/Dropbox/netsMatter/replications/example/outputData/model_k',
+			latDims[ii],'.rda')
+}
+stopCluster(cl)
+```
+
+## Code + Data + Conversation
+
+* We will be writing code on Git **ONLY**
+* Input and output data will be stored on Dropbox **ONLY**
+* We can converse on Slack 
+	- As you guys run into issues/questions/concerns we can raise them there and answer them in close to real-time
+
+## Project Management
+
+* Teams of two, each of you will be paired with someone experienced in AME
+	- Dorff + Foster
+	- Gallop + Liu
+	- Ward + Roberts
+	- Me + Tellez 
+* Claim the replications you want to do on Google sheets ... whoever gets there first wins.
+	- Mansfield + Milner is already mine.
+* <span style="color:red">**How many replications should we do?**</span>
+* <span style="color:red">**Timeline?**</span>
+* <span style="color:red">**What else?**</span>

### 2016-11-25 amenIntro/amenIntro.Rmd
some updates

@@
 knitr::opts_chunk$set(echo = FALSE)
+
+if(Sys.info()['user']=='janus829' | Sys.info()['user']=='s7m'){
+	dropPath = '~/Dropbox/Research/netsMatter/replications/example/'
+	gitPath = '~/Research/netsMatter/'
+	dataPath = paste0(dropPath, 'inputData/')
+	resultsPath = paste0(dropPath, 'outputData/') }
 ```
@@
 * This is going to be a bit different than what is described in [Hoff's vignette](https://arxiv.org/pdf/1506.08237v1.pdf). 
-      - As of this week, AMEN can now be estimated on networks where the actor composition changes over time, and to make that change we switch to using **lists** instead of **arrays**.
+      - As of yesterday, AMEN can now be estimated on networks where the actor composition changes over time, and to make that change we switch to using **lists** instead of **arrays**.
+      	+ Hoff still need to give the new code a review.
 * In undirected longitudinal setting three inputs:
@@
       model="bin",symmetric=TRUE,
-      burn=500,nscan=1000,odens=10, # WILL NEED TO BE LARGER
+      burn=500,nscan=1000,odens=10, 
       plot=FALSE, print=FALSE, 
@@
 
-## Out of sample model performance
-
-* Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
-* For each $s \in \{1, \ldots, S\}$:
-	- Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
-	- For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
-
-* Code to perform this analysis for the example is shown in `git/netsMatter/code/example/runOutPerf.R`
-* When we get out of sample results we will be evaluating across two metrics:
-	- Receiver Operating Characteristic curves
-	- Precision-Recall curves
-
 ## There will be a lot of models to run...
@@
 	save(ameFit, 
-		file='~/Dropbox/netsMatter/replications/example/outputData/model_k',
-			latDims[ii],'.rda')
+		file=paste0('~/Dropbox/netsMatter/replications/example/outputData/model_k',
+			latDims[ii],'.rda'))
 }
@@
 ```
+
+## Might need parallelization...
+
+* because we will want to assess how varying k changes GOF
+
+```{r, echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
+load(paste0(resultsPath, 'modelk1.rda')) ; fit1=fit
+load(paste0(resultsPath, 'modelk2.rda')) ; fit2=fit
+load(paste0(resultsPath, 'modelk3.rda')) ; fit3=fit
+load(paste0(resultsPath, 'modelk4.rda')) ; fit4=fit
+
+fits = list(fit1, fit2, fit3, fit4) ; names(fits) = paste0('k=',1:4)
+
+gofSumm = do.call('rbind', lapply(1:length(fits), function(ii){
+	fit = fits[[ii]]
+	act = fit$GOF[1,-c(2:3)]
+	gof = fit$GOF[-1,-c(2:3)]
+
+	ggGof = melt(gof)[,-1]
+	ggGof$act = act[match(ggGof$Var2, names(act))]	
+	ggGof$K = names(fits)[ii]
+	return(ggGof)
+	}) )
+
+ggplot(gofSumm, aes(x=value )) + 
+	geom_histogram() + 
+	geom_vline(aes(xintercept=act), color='red') + 
+	facet_grid(Var2 ~ K) +
+	theme(axis.text.x=element_text(angle=45,hjust=1))
+```
+
+## and to assess performance
+
+* Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
+* For each $s \in \{1, \ldots, S\}$:
+	- Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
+	- For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
+
+* Code to perform this analysis for the example is shown in `git/netsMatter/code/example/runOutPerf.R`
+* When we get out of sample results we will be evaluating across two metrics:
+	- Sensitivity/Specificity & Precision/Recall
+
+```{r, echo=FALSE}
+
+source(paste0(gitPath, 'code/helpers/functions.R'))
+source(paste0(gitPath, 'code/helpers/binPerfHelpers.R'))
+
+load(paste0(resultsPath, 'model_k1_outPerfResults.rda'))
+load(paste0(resultsPath, 'model_k2_outPerfResults.rda'))
+load(paste0(resultsPath, 'model_k3_outPerfResults.rda'))
+load(paste0(resultsPath, 'model_k4_outPerfResults.rda'))
+
+preds = list(do.call('rbind', modsAme1), do.call('rbind', modsAme2), do.call('rbind', modsAme3), do.call('rbind', modsAme4))
+names(preds) = paste0('k=',1:4)
+
+aucSumm = do.call('rbind', 
+	lapply(preds, function(x){ 
+		cbind( 'AUC'=getAUC(x$prob, x$actual), 'AUC (PR)'=auc_pr(x$actual, x$prob) ) 
+		} ) ) ; rownames(aucSumm) = names(preds)
+aucSumm = aucSumm[order(aucSumm[,1],decreasing=TRUE),]
+aucSumm = trim(format(round(aucSumm, 2), nsmall=2))
+print( aucSumm )
+```
+
+## Issues
+
+* Any missingness in covariates inputted into AME automatically get turned into 0 ... this is not good.
+* We can get around this via imputing nodal and dyadic variables within AME
+	- In this vein, I made a modified version of sbgcop that works over nodal and dyadic data...but we still need to incorporate this within the AME MCMC
+	- also *within* because otherwise we'd have to run each model k times over different imputed datasets and use Rubin's rules to combine parameter estimates ...
+problematic especially when it comes to third order deps
+* Computation time ... these models take a notable amount of time to converge
+	- Mike's tower?
+	- EC2 ... though the costs can build
 

### 2016-11-25 amenIntro/amenIntro.Rmd
final updates

@@
 ---
-title: "Networks Matter...or some nonsense like that"
+title: "Amen to replication: What have we learned about networks in International relations "
+author: "Minhas, Dorff, Foster, Gallop, Liu, Roberts, Tellez, Ward"
 output: 
@@
       - One big meta-analysis piece in which we show the items in the list above to be true on average
-      - A number of small replication pieces
+      - <span style="color:red">**After**</span> big project is done, a number of small replication pieces
       - New disciples to spread the gospel
@@
 
-* Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
+* In each year, randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
 * For each $s \in \{1, \ldots, S\}$:
@@
 * <span style="color:red">**How many replications should we do?**</span>
+	- I propose two.
 * <span style="color:red">**Timeline?**</span>
+	- AT LEAST ONE REPLICATION MODEL RUNNING BY JANUARY 1
+	- Paper draft end of January
 * <span style="color:red">**What else?**</span>

### 2016-11-25 amenIntro/amenIntro.Rmd
timeline updates

@@
 problematic especially when it comes to third order deps
-* Computation time ... these models take a notable amount of time to converge
+* <span style="color:red">**Computation time**</span> ... these models take a notable amount of time to converge
 	- Mike's tower?
 	- EC2 ... though the costs can build
-
-## Code + Data + Conversation
-
-* We will be writing code on Git **ONLY**
-* Input and output data will be stored on Dropbox **ONLY**
-* We can converse on Slack 
-	- As you guys run into issues/questions/concerns we can raise them there and answer them in close to real-time
 
@@
 * Claim the replications you want to do on Google sheets ... whoever gets there first wins.
-	- Mansfield + Milner is already mine.
 * <span style="color:red">**How many replications should we do?**</span>
-	- I propose two.
-* <span style="color:red">**Timeline?**</span>
-	- AT LEAST ONE REPLICATION MODEL RUNNING BY JANUARY 1
+	- I propose two per team.
+* <span style="color:red">**Timeline**</span>
+	- Decide on your two replication pieces by December 1
+	- EACH TEAM SHOULD HAVE AT LEAST ONE REPLICATION MODEL RUNNING BY JANUARY 1
 	- Paper draft end of January
 * <span style="color:red">**What else?**</span>
+
+## Code + Data + Conversation
+
+* We will be writing code on Git **ONLY**
+* Input and output data will be stored on Dropbox **ONLY**
+* We can converse on Slack 
+	- As you guys run into issues/questions/concerns we can raise them there and answer them in close to real-time

### 2016-11-25 amenIntro/amenIntro.Rmd
sync up

@@
 ---
-title: "Amen to replication: What have we learned about networks in International relations "
-author: "Minhas, Dorff, Foster, Gallop, Liu, Roberts, Tellez, Ward"
+title: "AMEN to Replication: Networks in IR"
+author: "Minhas, Dorff, Gallop, Foster, Liu, Roberts, Tellez, Ward"
 output: 

### 2017-08-25 political-analysis-draft/draftPA.tex
move manuscript to git

\documentclass[12pt]{amsart} 
      % \newif\ifpreview
      % \previewfalse
% approx iid
\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}\usepackage{natbib}

% Fonts
\usepackage[default,osfigures,scale=0.95]{opensans}
\usepackage[T1]{fontenc}
\usepackage{ae}

% easy commands for number propers
\newcommand{\first}{1$^{\text{st}}$}
\newcommand{\second}{2$^{\text{nd}}$}
\newcommand{\third}{3$^{\text{rd}}$}
\newcommand{\nth}[1]{${#1}^{\text{th}}$}

% Generate some fake text
\usepackage{blindtext}

% add watermark
\usepackage{draftwatermark}
\SetWatermarkText{Draft: Not Ready}
\SetWatermarkScale{.5}
\SetWatermarkColor[gray]{0.9}

% layout control
\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
\usepackage{parallel}
\usepackage{parcolumns}
\usepackage{fancyhdr}
\usepackage[export]{adjustbox}
\usepackage{etoolbox}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{relsize}
\usepackage{mathtools}
\usepackage{bm}
%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}

% restricts float objects to be inserted before end of section
% creates float barriers
\usepackage[section]{placeins}

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{longtable}

% to adapt caption style
\usepackage[font={small},labelfont=bf]{caption}

% footnotes at bottom
%\usepackage[bottom]{footmisc}

% to change enumeration symbols begin{enumerate}[(a)]
\usepackage{enumerate}
\usepackage{paralist}

% to colorize links in document. See color specification below
\usepackage[pdftex,hyperref,x11names]{xcolor}

% for multiple references and insertion of the word "figure" or "table"
% \usepackage{cleveref}

% load the hyper-references package and set document info
\usepackage[pdftex]{hyperref}

% graphics stuff
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage[space]{grffile} % allows us to specify directories that have spaces
\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
\usepackage{tikz}

% sideway figures
\usepackage{rotating}
\usepackage{lscape}

%\usepackage[figuresright]{rotating}
% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
%   {\end{minipage}\end{sidewaysfigure}}

% Spacing
\usepackage[doublespacing]{setspace}

% define clickable links and their colors
\hypersetup{%
	unicode=false,          % non-Latin characters in Acrobat's bookmarks
	pdftoolbar=true,        % show Acrobat's toolbar?
	pdfmenubar=true,        % show Acrobat's menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdfnewwindow=true,%
	%pagebackref=false,%
	pdfauthor={Shahryar Minhas, et alia.},%
	pdftitle={How to Fix IR},%
	colorlinks,%
	citecolor=black,%
	filecolor=black,%
	linkcolor=black,%
	urlcolor=RoyalBlue4
}

% -------------------- title -------------------- %

\title{How to fix IR}
\vspace{\baselineskip}

 \author[Minhas]{Shahryar Minhas}
 \address{Shahryar Minhas: Department of Political Science}
 \curraddr{Michigan State University}
 \email[Corresponding author]{minhassh@msu.edu}

 \author[Dorff]{Cassy L. Dorff}
 \address{Cassy L. Dorff}
 \curraddr{Department of Political Science, University of New Mexico}
 \email{cdorff@unm.edu}

 \author[Foster]{Margaret Foster}
 \address{Margaret Foster}
 \curraddr{Department of Political Science, Duke University}
 \email{margaret.foster@duke.edu}

 \author[Gallop]{Max Gallop}
 \address{Max Gallop}
 \curraddr{Department of Government and Public Policy, University of Strathclyde}
 \email{max.gallop@strath.ac.uk}

  \author[Liu]{Howard Liu}
 \address{Howard Liu}
 \curraddr{Department of Political Science, Duke University}
 \email{hao.liu@duke.edu}

 \author[Tellez]{Juan Tellez}
 \address{Juan Tellez}
 \curraddr{Department of Political Science, Duke University}
 \email{juan.f.tellez@duke.edu}

\author[Ward]{Michael D. Ward}
\address{Michael D. Ward}
\curraddr{Department of Political Science, Duke University}
\email{michael.d.ward@duke.edu}

% \date{\today}

\thanks{Shahryar Minhas and Michael D. Ward acknowledge support from National Science Foundation (NSF) Award 1259266. Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings. We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}

\setlength{\headheight}{15pt}
\setlength{\headsep}{20pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{}}
\chead{\fancyplain{}{How to Fix IR}}
\rhead{\fancyplain{}{\today}}
\rfoot{\fancyplain{}{\thepage}}
% references to graphics
\makeatletter
% let's keep them in the same directory \def\input@path{{}

% easy command for boldface math symbols
\newcommand{\mbs}[1]{\boldsymbol{#1}}

% command for R package font
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

% Add some colors
\definecolor{red1}{RGB}{253,219,199}
\definecolor{red2}{RGB}{244,165,130}
\definecolor{red3}{RGB}{178,24,43}

\definecolor{green1}{RGB}{229,245,224}
\definecolor{green2}{RGB}{161,217,155}
\definecolor{green3}{RGB}{49,163,84}

\definecolor{blue0}{RGB}{255,247,251}
\definecolor{blue1}{RGB}{222,235,247}
\definecolor{blue2}{RGB}{158,202,225}
\definecolor{blue3}{RGB}{49,130,189}
\definecolor{blue4}{RGB}{4,90,141}

\definecolor{purple1}{RGB}{191,211,230}
\definecolor{purple2}{RGB}{140,150,198}
\definecolor{purple3}{RGB}{140,107,177}

\definecolor{brown1}{RGB}{246,232,195}
\definecolor{brown2}{RGB}{223,194,125}
\definecolor{brown3}{RGB}{191,129,45}

% square bracket matrices
\let\bbordermatrix\bordermatrix
\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}

\begin{document}
      % \ifpreview {
      %    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
      %    \pagecolor{personal}
      %    }
      % \fi

\small{\singlespacing{
\begin{abstract}
	International relations scholarship is filled with dyads. This includes dyadic hypotheses and propositions, but especially data.  These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.  However, most empirical studies of such data do not take into account these dependencies. As a result, the studies are often contradictory and produce results that are not compelling. One reason is that the independence required by the typical methods employed does not exist in the data being studied.  We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies. We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach. The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
\end{abstract}
}}

\maketitle\thispagestyle{empty}
\newpage\setcounter{page}{1}

\section{\textbf{Introduction}}

\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM), either OLS or logistic regression, to estimate regression coefficients.  However, we know that extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined. Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases. This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements. It is not just the diagonals of the variance-covariance matrix this affects.

In this article, we introduce a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in the data.  This model has two parts, one of which uses a set of additive random effects capturing \first- and \second-order dependencies in the relational data. The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.  We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a generalized linear model with robust standard errors) to those obtained with the AME framework.

This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches. It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations. Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.  
%We note in closing that the computational requirements of this approach are substantial in terms of computing time. Cloud and parallel methods reduce this burden.

\section{\textbf{Additive and Multiplicative Effect Models for Networks}}

The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages. It is flexible and easy to use for analyzing the kind of relational data often found in social science.  It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.

We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}. Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship. Fortunately, it is easy to empirically examine dependencies in relational data.

The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence. It was introduced in \cite{hoff:2008} and elaborated in \cite{hoff:2015}; the open-source computer package associated with this approach is \cite{amenpkg}. We describe it below.

\subsection*{\textbf{Addressing Dependencies in Dyadic Data}}

Relational, or dyadic, data provide measurements of how sets of actors relate to one another. Typically, pairs of actors are examined. These pairs are called \textit{dyads}. Such data can be arranged as a directed dyadic design in which the unit of analysis is some set of $n$ actors that have been paired together to form a matrix of dyads, i. e., represented as a $n \times n$ matrix, $\mathbf{Y}$, with diagonals typically undefined. The $ij^{th}$ entry defines the relationship of $i$ and $j$ and can be continuous or discrete. It can also be symmetric (known as undirected) or asymmetric (called directed). This is not the stacked dyad structure normally found in social science, but it is equivalent and presents a format in which the interdependencies are exposed and transparent. The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.

One important dependency is nodal dependency, whereby each actor has similar interactions with many of its partners, leading to significant heterogeneity in activity levels across nodes. An important implication of this across-node heterogeneity is within-node homogeneity of ties: the values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, are similar to each other because each has a common sender. If a sender $i$ is more active active in the network than other senders, it will have greater homogeneity of its linkages. More popular targets show greater similarity across linkages in a given column, $\{y_{ji},y_{ki},y_{li}\}$. Actors which are more likely to send ties in a network are also more likely to receive them. As a result the rows and columns in an adjacency matrix may be correlated. Another structural interdependence is reciprocity. This is a second-order, or dyadic dependency whereby values of $y_{ij}$ and $y_{ji}$ are correlated. If the data are simply stacked dyads in some arbitrary order, this kind of dependency is harder to visualize, but exists nonetheless.

The presence of these interdependencies in relational data complicates the practical assumption of observational independence.  Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance. By ignoring these potential interdependencies, we often ignore important features of the problem under study. The study of international relations is founded on the relations among actors. Why ignore the interdependencies that led to the study of IR in the first place?

\subsection{\textbf{Social Relations Model: Additive Part of AME}}

\citet{warner:etal:1979} developed a social relations model (SRM), using ANOVA decomposition, that focuses on such dependencies. SRM provides the error structure for the additive effects component of the AME framework. SRM decomposes the variance of dyadic observations in terms of a) heterogeneity across row means (often called out-degree), b) column means (or, in-degree), c) correlation between row and column means, and d) correlations within dyads.  \citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
\begin{align}
\begin{aligned}
	y_{ij} &= \mu + e_{ij} \\
	e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
	\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
	\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
	\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix},
\label{eqn:srmCov}
\end{aligned}
\end{align}

wherein $\mu$ is the density of a network, and $e_{ij}$ represents residual variation in each cell (or dyad). The residual variation has three separate parts: a)   row/sender effects ($a_{i}$), b)  column/receiver effects ($b_{j}$), and c) within-dyad effects ($\epsilon_{ij}$). The row and column effects are modeled jointly to account for correlation in how active an actor is in sending and receiving ties. Heterogeneity in the row and column means is modeled by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively.  $\sigma_{ab}$ describes whether actors which send a lot of ties also receive a lot. Second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a reciprocity, parameter $\rho$.

The SRM covariance structure described in Equation~\ref{eqn:srmCov} can be incorporated into the systematic component of a GLM framework to produce the social relations regression model (SRRM): $\bm\beta^{\top} \mathbf{X}_{ij} + a_{i} + b_{j} + \epsilon_{ij}$, where $ \bm\beta^{\top} \mathbf{X}_{ij}$ accommodates the inclusion of dyadic, sender, and receiver covariates \citep{hoff:2005}. This is a regression framework with additive, random effects, but it models variances and covariances often seen in relational data. Furthermore, this accommodates a diversity of outcome distributions. In the case of binary data, this can be done by utilizing a random effects logistic (or probit) latent variable representation.

% It has been employed in \citet{hoff:ward:2004, ward:hoff:2007, ward:etal:2007,cao:2009, fletcher:etal:2011, breunig:etal:2012, ward:etal:2012,cao:ward:2014} and is the basis for many recommender systems.

\subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}

The additive part described above does not handle third-order dependency. A third-order dependency is the dependency between triads, not dyads. Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another. Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
For example, a common finding is that democracies are more likely to form trade agreements with one another. Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics. Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations dating back to Sun Tzu. A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another. Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.

Another third-order dependence pattern is stochastic equivalence. Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$. Equivalence implies the existence of groups of nodes in a network with similar relational patterns.  Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}. Latent variables can account for third-order dependence patterns within the context of the SRRM. These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.

A number of latent variable approaches have been developed to represent third-order dependencies in relational data, we focus on the latent factor model (LFM). This expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data. General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.

\begin{align}
\begin{aligned}
	\alpha(&\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{\top} \Lambda \textbf{u}_{j} \\
	&\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \text{, where}\\
	 &\Lambda \equiv K \times K \text{ diagonal matrix}
\label{eqn:latAlpha}
\end{aligned}
\end{align}

Many observed  networks exhibit varying degrees of stochastic equivalence and homophily.  In the latent factor model, each actor has an unobserved vector of characteristics, $\textbf{u}_{i} = \{u_{i,1}, \ldots, u_{i,K} \}$, describing each node's behavior in the network. The probability of a tie between $i$ and $j$ depends on the extent to which $\textbf{u}_{i}$ and $\textbf{u}_{j}$ are ``similar'' and on whether the entries of $\Lambda$ are greater or less than zero. Similarity in the latent factors, $\textbf{u}_{i} \approx \textbf{u}_{j}$, corresponds to stochastic equivalence; the eigenvalue determines whether the network exhibits positive or negative homophily. Accordingly, this approach can represent both homophily and stochastic equivalence.

In directed (asymmetric) networks there are two sets of latent positions, denoted $\textbf{u}$ and $\textbf{v}$, representing the sending and receiving positions separately.  Thus, actors in the network have a vector of latent characteristics describing their behavior as a sender, denoted by $\textbf{u}$, and as a receiver, $\textbf{v}$: $\textbf{u}_{i}, \textbf{v}_{j} \in \mathbb{R}^{K}$. This can alter the probability of an interaction between $ij$ additively: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, where $\textbf{D}$ is a $K \times K$ diagonal matrix.

The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations. Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:

\begin{align}
\begin{aligned}
	y_{ij} &= g(\theta_{ij}) \\
	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
\label{eqn:ame}
\end{aligned}
\end{align}

With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies. In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence. The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information. A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.

\section{Replications}

\subsection{Design}

For the purpose of this study, we choose five prominent studies from the broad field of international relations and international political economy that utilize relation data \citep{mcdonald:2004, reiter:stam:2003, rose:2004, weeks:2012, gibler:2017}. We chose to replicate studies that were written fairly recently, since 2000, and that have been cited over 100 times. Each of these pieces was published in a prominent journal and is well-known in the literature. They all used the standard approach in political science, which is to employ some form of general linearized regression that ignores dyadic interdependencies. Post estimation, standard errors are often adjusted in an attempt to account for clustering of observations.

\begin{table}
\caption{Features of the Studies Replicated. }
\scalebox{.8}{
  \begin{tabular}{ccrcc}
  Study & Model & N. of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \toprule
  Reiter \& Stam (2003) & Logit & $753,456$ & Directed & Robust \\
  McDonald (2004) & Logit &$92,354$& Undirected &Robust\\
  Rose (2004) &OLS&$234,597$&Directed&Robust \\
  Weeks (2012) & Logit & $901,540$ & Directed & Robust \\
 Gibler (2017) &Logit& $650,557$&UnDirected& no \\
  \bottomrule
\end{tabular}
}
\end{table}

We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We examine each of the models using the AME framework described above.  Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.

We also examine the accuracy of the predictions made with each approach. Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.%%%Dont we want to talk more about performance here???? - MG

The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.  We believe that it does, and that the dynamic latent factor model provides a step forward.

\begin{table}[ht]
\centering
\begin{tabular}{r p{4cm} l}
  \hline
 & Central Finding & Does it Replicate in a Network Model? \\ 
  \hline
Reiter and Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
  McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & Fails to Replicate \\ 
  Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
  Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & Fails to Replicate \\ 
  Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & Fails to Replicate \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rl cc}
  \hline
 & Model & AUC (ROC) & AUC(PR) \\ 
  \hline
\multirow{2}{*}{Reiter and Stam (2003)} & GLM & 0.92 & 0.08 \\ 
~  & AME & 0.96 &  0.15 \\ 
\multirow{2}{*}{McDonald (2004)} & GLM & &  \\ 
~  & AME & &  \\ 
\multirow{2}{*}{Rose (2004)} & GLM & &  \\ 
~  & AME & &  \\ 
\multirow{2}{*}{Weeks (2012)} & GLM & &  \\ 
~  & AME & &  \\ 
\multirow{2}{*}{Gibler (2017)} & GLM & &  \\ 
~  & AME & &  \\ 
   \hline
\end{tabular}
\end{table}

\subsection{Replication of Weeks (2012)}

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.  One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. Her time frame is from 1946-1999.  The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.  These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340). She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model. In Weeks's initial models, as detailed in figure \ref{fig:weeksCoef} she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations. When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero. We similarly find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third order effects. Inspection of the multiplicative effects in figure \ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets. In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea). In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China. In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior. So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type. The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects. In terms of out of sample performance, shown in figure \ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.

\begin{figure}
\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
\end{figure}

\begin{figure}
\centering   
\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
\end{figure}

\subsection{Replication of Reiter and Stam (2003)}

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.  They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent. They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.  In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.  They use the MID data, but note that ``We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation. This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force. We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets. \ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures. As is prevalent in these kinds of studies, Reiter and Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.

In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero. In our replication using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero. Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
Our replication using the AME approach therefore cast doubt on Reiter and Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies. Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}. Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitauc} and Figure ~\ref{fig:reitpr}).

\begin{figure}
\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
\caption{\label{fig:reitCoef} Coefficient plot of Reiter and Stam (2003)'s original model (blue) compared to AME model (red). }
\end{figure}

\begin{figure}
\centering   
\subfigure[AUC]{\label{fig:reitauc}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
\caption{Assessments of out-of-sample predictive performance for Reiter and Stam (2003) using ROC curves and PR curves.}
\end{figure}

\newpage
\subsection{Replication of Rose (2004)}

In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations. It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.  Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims. He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades. Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders... (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.

When we compare the results of Rose's original OLS model to our model that accounts for network dependencies, the results are generally similar. As you can see in Figure~\ref{fig:roseCoef}, the main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- remains when we move from an OLS to a Gaussian AME model. The most striking difference between the models is that, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. Excitingly, when we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out of sample -- 3.23 for the OLS model and 1.77 for the AME model.

\begin{figure}
\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
 \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{ABplot_rose_top10.pdf}
 \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
\end{figure}

%\begin{figure}
%\centering   
%\subfigure[Nodal Random Effects]%{\label{fig:rosesend}\includegraphics[width=65mm]{ABplot_rose_top10.pdf}}
%\subfigure[Receiver Random Effects]{\label{fig:rosereceive}\includegraphics[width=65mm]{ABplot_receiver_rose.pdf}}
%\caption{Sender and receiver random effects for AME estimation of Rose (2004).}
%\label{fig:rosesr}
%\end{figure}

\newpage
\subsection{Replication of McDonald (2004)}

\citet{mcdonald:2004} studies whether trade promotes peace between nations. He observes that knowledge about the link between conflict and trade is indeterminate in the field of international relations, noting that competing explanations persist. He calls for more precise empirical tests.  Importantly, \citet{mcdonald:2004} includes the interdependence argument---that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)'' in his overview of underdeveloped hypotheses.

Accordingly, the primary contribution of the study is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade. To do so, \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states. His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences. \citet{mcdonald:2004} measures free trade in two ways. The first  captures the idea that larger protected sectors generate greater societal pressures resulting in  pockets of support for war. This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad. This measure captures the score of the state in the dyad that possesses higher barriers to trade (560). \citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state i's GDP or total dyadic trade divided by state j's GDP.'' (560). Finally, the binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad. \citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables. The models include a spline correction for time-series data as well as robust standard errors clustered on each dyad.

% TO DO: include discussion and analysis of different K's in appendix (for M cD)
\begin{figure}
\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
\end{figure}

\begin{figure}
\centering   
\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
\end{figure}

% TO DO: fix bib

Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.  Or, in other words, if countries {i} and {j} are highly dependent and countries {j} and {k} are also highly dependent, then we are likely to observe high dependency between countries {i} and {k}. Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict. This indicates that conflict is less likely between members of a trade community. Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict. Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and replicated model and Figure~\ref{fig:mcdall} demonstrates the predictive performance of each model. Figure ~\ref{fig:mcdcirc} represents the multiplicative effects of the model. This plot enables us to consider actors with similar sending and receiving patterns resulting from stochastic equivalence and homophily. In the upper arch for senders, there is a clear clustering of United States' allies (the United Kingdom, Netherlands, and Japan). We also observe an African cluster in the top right segment for receivers (Egypt, Niger, and Zambia for example). These countries both trade amongst themselves and have similar conflict behavior -- stochastic equivalence is present among the first group, homophily among the second -- and so once we account for these clusters using the multiplicative effects, the residual effect of trade actually increases the likelihood of conflict.

\begin{figure}
\includegraphics[width=\textwidth]{mcdonaldcirc.pdf}
\caption{\label{fig:mcdcirc} Visualization of multiplicative effects. Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
\end{figure}

\subsection{Replication of Gibler (2017)}

% We chose this article because it is published in 2017, demonstrating that the concern with interdependence continues to the present and because it was published in the so-called flagship journal of political science, \textit{The American Political Science Review.} It examines the link between dyadic capability differences and international conflict in terms of how and when states became independent members of the international system.  It assumes that membership in the international system---defined in typical ways---has a substantial influence on capabilities. In the same way that capabilities don't seem to vary enormously over time, at least comparatively, states become independent in a particular location, clustered with surrounding states. Neither does this change much over time. Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members. This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.

The replications we have undertaken are all from articles over the past fifteen years.  A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.   In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members. This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.

\begin{table}
\begin{center}
\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000$ iterations.  \label{tab:gibme}}
\begin{tabular}{lrr} \toprule
& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
Variable & \texttt{logit} & \texttt{amen} \\ \midrule
Allied & 0.142&  0.023 \\
Joint Democracy &\bf -0.507&  0.045 \\
Peace Years \bf &-0.260& \bf -0.060 \\
Spline 1 &\bf -0.001& \bf 0.00 \\
Spline 2&\bf -0.000& \bf 0.00\\
Spline 3 &-0.000& \bf 0.00\\ 
Contiguity &\bf 2.412& \bf 0.640 \\
Parity &0.075&  -0.013 \\
Parity at entry year&\bf 0.868&  0.002 \\ 
Rivalry &\bf 2.031&\bf 0.721\\ 
Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
\end{tabular}\\
\end{center}
{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017). For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
\end{table}

We re-estimated model 6 from Table 6 (2017,34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).  Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the amen results.  Not only is the value of the this parameter small, but it has a very large relative standard error, over a magnitude larger than the parameter itself ($z= 0.038$). In addition, joint democracy follows the same pattern of importance in the logistic results, but this disappears once interdependencies are modeled.  As might be expected the strong geographic clustering in the original study is about one-quarter as strong in the \texttt{amen} estimations. Similarly, rivalry coefficients are about one-third the size in the \texttt{amen} formulation, but a great deal more precisely measured ($z=18.116$).

Beyond more informative fixed effect coefficients, the \texttt{amen} approach also provides information about the interdependencies that were modeled. The most pertinent of these may be the dyadic and triadic dependencies which are presented in Figure~\ref{fig:gibler:dyadtriad}, which further reveal that the assumption of independence among the dyadic data in this study can be strongly rejected. Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies. Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.

% sd.rowmean sd.colmean   dyad.dep  triad.dep 
% 0.03210221 0.03210221 0.99362646 0.00130249 
\begin{figure}
\caption{\label{fig:gibler:dyadtriad} The left panel is  a plot of the dyadic dependencies in the Gibler model 6; the right panel shows the triad dependencies.  Note that the red line represents the null hypothesis of no dependencies, indicating that the standard logistic approach is far from what is uncovered with this analysis. }
\includegraphics[width=\textwidth]{netplots}
\end{figure}

\section{Conclusion}

It is no longer necessary to assume that the interesting, innate interdependencies in relational data can be ignored. Nor do they have to be approximated with ad hoc, incomplete solutions that purport to control for dependencies by modifying the post-estimation standard errors of the estimated coefficients. Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.

%%What about a summary table with AUCs for the different network models here? - MG
\newpage
% Bib stuff
\clearpage
% \bibliography{master.bib}
\bibliography{/Users/s7m/whistle/master}
\bibliographystyle{authordate1}
\clearpage
\section*{All About Us}
\end{document}

### 2017-08-25 political-analysis-draft/draftPA.tex
editing plot scripts

@@
 
-We also examine the accuracy of the predictions made with each approach.
+We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
+
+% % % % % % % % % % % % % % % % % % 
+By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
+To show that this is the case, we examine whether our approach achieves better predictive performance in an out of sample context than traditional dyadic models.
+To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
+Then for each $s \in \{1,\ldots,k\}$, we:
+
+\begin{enumerate}
+	\item estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$,
+	\item and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters. 
+\end{enumerate}
+
+\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
+
+We set a number of benchmarks for comparison.
+First we compare the AME model to a GLM model using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
+We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
+The lagged dependent variable is the equivalent of saying that conflict and peace are relatively likely to persist between dyads, while the inclusion of a lagged reciprocity term in a GLM framework is a simple way to account for retaliatory strikes.
+
+We utilize three performance criterions to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
+ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
+An issue with an ROC Curve when looking at conflict, is that it is relatively rare at the dyadic level: in most years only 3\% of possible dyads are in conflict with one another.
+If peace is common, even a poor model will have a very low False Positive Rate.
+
+To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
+These examine the trade-offs between the percentage of conflicts a model predicts, and the percentage of predicted conflicts which occur.
+Lastly, we examine separation plots \citep{greenhill:etal:2011}.
+These provide an intuitive visualization of the accuracy of our predictions by juxtaposing a line showing the predicted probability of conflict with whether conflict actually occurs for all cases (where the cases are sorted by the predicted probability and then colored to indicate the outcome).
+Here a perfect model would have all cases where conflict actually exists on the right with a predicted probability of $1$, and would predict $0$ in all other cases.
+All of the models' performance out of sample by these metrics are displayed in figure \ref{fig:roc_ame}.
+The AME model with covariates is the best performing model out of sample in all cases.
+This model outperforms each of the GLM variants by a notable margin.
+% % % % % % % % % % % % % % % % % %
+
 Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.%%%Dont we want to talk more about performance here????
@@
 ~  & AME & &  \\ 
-\multirow{2}{*}{Weeks (2012)} & GLM & &  \\ 
-~  & AME & &  \\ 
+\multirow{2}{*}{Weeks (2012)} & GLM & 0.64 & 0.004  \\ 
+~  & AME & 0.97 &  0.15 \\ 
 \multirow{2}{*}{Gibler (2017)} & GLM & &  \\ 

### 2017-08-25 code/weeks_2012/10a-plotOutSampAUCPRK2.R

@@ -4,12 +4,17 @@
 
 rm(list=ls())
 
-resultsPath = '~/Dropbox/netsMatter/replications/Weeks2012/replication/output/'
-
 ### libraries needed
 
-source('../reiter_stam_2003/helperEx.R')
-source('setup.R')
+if(Sys.info()['user']=='s7m'){
+  source('~/Research/netsMatter/code/weeks_2012/setup.R')
+  source('~/Research/netsMatter/code/reiter_stam_2003/helperEx.R')
+  resultsPath = '~/Dropbox/Research/netsMatter/replications/Weeks2012/replication/output/'
+} else {
+  source('setup.R')
+  source('../reiter_stam_2003/helperEx.R')
+  resultsPath = '~/Dropbox/netsMatter/replications/Weeks2012/replication/output/'
+}
 
 toLoad <- c("RColorBrewer","dplyr" ,"magrittr" ,"ggplot2" ,"stringr"
             ,"gridExtra","Cairo","reshape2","tidyverse","latex2exp")
@@ -28,11 +33,8 @@ loadPkg(toLoad)
 ##(The GLM out samp file is 3.2G!)
 
 load(paste0(resultsPath,'weeksOutPerf.rda'))
-
-load(paste0(resultsPath, 'outsampResults2.rda')); ameOutSamp_k2 <-
-     ameOutSamp_NULL
-
-
+load(paste0(resultsPath, 'outsampResults2.rda'))
+ameOutSamp_k2 <- ameOutSamp_NULL
 
 ## ROC plots
 

### 2017-08-26 political-analysis-draft/draftPA.tex
paper and missing gibler stuff

@@
 		Study & Model & N. of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \toprule
+		Weeks (2012) & Logit & $901,540$ & Directed & Robust \\
 		Reiter \& Stam (2003) & Logit & $753,456$ & Directed & Robust \\
+		Rose (2004) &OLS&$234,597$&Directed&Robust \\
 		McDonald (2004) & Logit &$92,354$& Undirected &Robust\\
-  Rose (2004) &OLS&$234,597$&Directed&Robust \\
-  Weeks (2012) & Logit & $901,540$ & Directed & Robust \\
 		Gibler (2017) &Logit& $650,557$&UnDirected& no \\
@@
 		\hline
-Reiter and Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
+		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & Fails to Replicate \\
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
+		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & Fails to Replicate \\ 
-  Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-  Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & Fails to Replicate \\ 
-  Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & Fails to Replicate \\ 
+		% Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & Fails to Replicate \\ 
 	\hline
@@
 \centering
-\begin{tabular}{rl cc}
-  \hline
- & Model & AUC (ROC) & AUC(PR) \\ 
-  \hline
-\multirow{2}{*}{Reiter and Stam (2003)} & GLM & 0.92 & 0.08 \\ 
-~  & AME & 0.96 &  0.15 \\ 
-\multirow{2}{*}{McDonald (2004)} & GLM & &  \\ 
-~  & AME & &  \\ 
-\multirow{2}{*}{Rose (2004)} & GLM & &  \\ 
-~  & AME & &  \\ 
-\multirow{2}{*}{Weeks (2012)} & GLM & 0.64 & 0.004  \\ 
-~  & AME & 0.97 &  0.15 \\ 
-\multirow{2}{*}{Gibler (2017)} & GLM & &  \\ 
-~  & AME & &  \\ 
-   \hline
+\begin{tabular}{ll cc}
+~ & ~ & GLM & AME \\
+\multirow{2}{*}{Weeks (2012)} & AUC (ROC) & 0.64 & 0.97 \\
+~ & AUC (PR) & 0.004 & 0.15 \\
+\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & 0.96 \\
+~ & AUC (PR) & 0.08 & 0.15 \\
+\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 & 1.99 \\
+~ & RM-D-SE & 2.01 & 1.06 \\
+\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & 0.99 \\
+~ & AUC (PR) & 0.13 & 0.28 \\
+% \multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.92 & 0.96 \\
+% ~ & AUC (PR) & 0.08 & 0.15 \\
 \end{tabular}
@@
 
-\subsection{Replication of Reiter and Stam (2003)}
+\subsection{Replication of Reiter \& Stam (2003)}
 
@@
 \ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in these kinds of studies, Reiter and Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
+As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
 The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
@@
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast doubt on Reiter and Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
+Our replication using the AME approach therefore cast doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}.
@@
 \includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-\caption{\label{fig:reitCoef} Coefficient plot of Reiter and Stam (2003)'s original model (blue) compared to AME model (red). }
+\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
 \end{figure}
@@
 \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
-\caption{Assessments of out-of-sample predictive performance for Reiter and Stam (2003) using ROC curves and PR curves.}
+\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
 \end{figure}

### 2017-08-26 code/McDonald_2004/06-evaloutSampAME.R

@@ -1,12 +1,18 @@
 ## Script to look at the AME xval output
 
-dataPath <- "~/Dropbox/netsMatter/replications/McDonald_2004/data/"
+rm(list=ls())
+if(Sys.info()['user']=='s7m'){
+	dataPath <- "~/Dropbox/Research/netsMatter/replications/McDonald_2004/data/"	
+} else {
+	dataPath <- "~/Dropbox/netsMatter/replications/McDonald_2004/data/"	
+}
+
 
 
 ## GLM
 
 load(paste0(dataPath, "smallMcDGLMStats.rda"))
-GLMoutSamp
+# GLMoutSamp
 
 
 ls()
@@ -16,7 +22,7 @@ smallMcDGLMStats$ROC
 smallMcDGLMStats$PR
 
 ##K0
-
+setwd(dataPath)
 load("outsampResults0.rda")
 
 outFitK0 <- ameOutSamp_NULL

### 2017-08-26 code/gibler_2017/giblerrepl/1_glm.R

@@ -1,0 +1,38 @@
+rm(list=ls())
+
+if(Sys.info()['user']=='s7m'){
+	dataPath = '~/Dropbox/Research/netsMatter/replications/gibler_2017/'	
+	source('~/Research/netsMatter/code/helpers/loadPkg.R')
+	source('~/Research/netsMatter/code/helpers/clusteredSE.R')
+} else {
+	dataPath = '~/Dropbox/netsMatter/replications/gibler_2017/'
+	source('netsMatter/code/gibler_2017/loadPkg.R')
+}
+
+#
+load(paste0(dataPath,"gibler2017_replication.rda"))
+data <- gilberData
+names(data)[which( names(data) %in% paste0('_spline',1:3) )] = paste0('spline',1:3)
+
+# modspec
+# table 6
+# column 6
+ids = c('ccode1', 'ccode2', 'year', 'dyad')
+dv = 'cwmid'
+ivs = c('allied','jointdem', 'cwpceyrs_bkt',
+	'spline1', 'spline2', 'spline3', 'contig', 'parity4',
+	'firstparity', 'riv1' )
+data = na.omit( data[,c(ids, dv, ivs)] )
+
+#
+modForm = paste(ivs, collapse=' + ') %>% paste0(dv, ' ~ ', .) %>% formula()
+
+# model
+mod = glm(modForm, family = binomial(link = 'logit'), data = data)
+
+# clustered SEs
+loadPkg('lmtest')
+modSumm = coeftest(mod, vcov = vcovCluster(mod, cluster = data$dyad))
+
+#
+save(mod, modSumm, file=paste0(dataPath, 'outputData/glmFit.rda'))

### 2017-08-27 political-analysis-draft/draftPA.tex
code stuff, remake mult eff plots

@@
 \centering
-	\begin{tabular}{r p{4cm} l}
+	\begin{tabular}{l| p{7cm} l}
+		& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
+		& &  in a Network Model? \\
+		\hline\hline
+		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & Fails to Replicate \\
 		\hline
-		& Central Finding & Does it Replicate in a Network Model? \\ 
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
 		\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & Fails to Replicate \\
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
 		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
+		\hline
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & Fails to Replicate \\ 
-		% Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & Fails to Replicate \\ 
 		\hline
+		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & Fails to Replicate \\ 
+	\hline\hline
 	\end{tabular}
@@
 \centering
-\begin{tabular}{ll cc}
+\begin{tabular}{l|l cc}
 ~ & ~ & GLM & AME \\
-\multirow{2}{*}{Weeks (2012)} & AUC (ROC) & 0.64 & 0.97 \\
-~ & AUC (PR) & 0.004 & 0.15 \\
+\hline\hline
+\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & 0.97 \\
+~ & Area Under PR Curve (AUC-PR) & 0.004 & 0.15 \\
+\hline
 \multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & 0.96 \\
 ~ & AUC (PR) & 0.08 & 0.15 \\
-\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 & 1.99 \\
-~ & RM-D-SE & 2.01 & 1.06 \\
+\hline
 \multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & 0.99 \\
 ~ & AUC (PR) & 0.13 & 0.28 \\
-% \multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.92 & 0.96 \\
-% ~ & AUC (PR) & 0.08 & 0.15 \\
+\hline
+\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & xx \\
+~ & AUC (PR) & 0.004 & xx \\
+\hline
+\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & 1.99 \\
+~ & Root Median Squared Error (RMDSE) & 2.01 & 1.06 \\
+\hline\hline
 \end{tabular}

### 2017-08-27 code/gibler_2017/giblerrepl/2_glm_outSamp.R

@@ -44,102 +44,94 @@ cores=folds=4
 ################
 
 ################
-glmOutSamp = function(glmForm){
+################
+# divide dataset into folds
+set.seed(seed)
+yListFolds = lapply(yList, function(y){
+  yFold=matrix(sample(1:folds, length(y), replace=TRUE),
+    nrow=nrow(y),ncol=ncol(y), dimnames=dimnames(y))
+  diag(yFold) = NA
+  return(yFold) })
+################
 
-  ################
-  # divide dataset into folds
-  set.seed(seed)
-  yListFolds = lapply(yList, function(y){
-    yFold=matrix(sample(1:folds, length(y), replace=TRUE),
-      nrow=nrow(y),ncol=ncol(y), dimnames=dimnames(y))
-    diag(yFold) = NA
-    return(yFold) })
-  ################
-  
-  ################
-  # run models by fold
-  yCrossValTrain = lapply(1:folds, function(f){
-    yListMiss = lapply(1:length(yList), function(t){
-      foldID = yListFolds[[t]] ; y = yList[[t]]
-      foldID[foldID==f]=NA ; foldID[foldID != f] = 1
-      y=y*foldID
-      return(y) })
-    names(yListMiss) = names(yList)
-    return(yListMiss) }) ; names(yCrossValTrain) = as.character(1:folds)
-  
-  # melt into glm format
-  yCrossValTrain = lapply(yCrossValTrain, function(y){
-    y = melt(y) ; y$id = paste(y$Var1, y$Var2, y$L1, sep='_')
-    xd = melt(xDyadList)
-    xd = tidyr::spread(xd, key = Var3, value = value)
-    xd$id = paste(xd$Var1, xd$Var2, xd$L1, sep='_')
-    glmData = y
-    for(v in ivs){
-      glmData$tmp = xd[match(glmData$id,xd$id),v]
-      names(glmData)[ncol(glmData)] = v }
-    glmData = glmData[which(glmData$Var1!=glmData$Var2),]
-    return(glmData)	
-  })
-  
-  # run glm
-  fitCrossVal = lapply(yCrossValTrain, function(glmData){
-    glmData$value[glmData$value>1] = 1
-    fit = glm(glmForm, data=glmData, family='binomial')	
-    return(fit)
-  })
-  
-  # get preds
[…]

### 2017-08-29 political-analysis-draft/draftPA.tex

@@
 
+Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.%%%Dont we want to talk more about performance here????
+- MG
+
+The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
+We believe that it does, and that the dynamic latent factor model provides a step forward.
+
+\begin{table}[ht]
+\centering
+	\begin{tabular}{l| p{7cm} l}
+	& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
+	& &  in a Network Model? \\
+	\hline\hline
+		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\
+		\hline
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
+		\hline
+		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
+		\hline
+		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
+		\hline
+		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 
+	\hline\hline
+	\end{tabular}
+	\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
+Cases in which the finding is not replicated are highlighted in bold.}
+	\label{tab:modelFindingSumm}
+\end{table}
+
 We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
@@
 
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.%%%Dont we want to talk more about performance here????
-- MG
-
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
-We believe that it does, and that the dynamic latent factor model provides a step forward.
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l| p{7cm} l}
-		& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
-		& &  in a Network Model? \\
-		\hline\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & Fails to Replicate \\
-		\hline
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
-		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-		\hline
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & Fails to Replicate \\ 
-		\hline
-		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & Fails to Replicate \\ 
-	\hline\hline
-	\end{tabular}
-\end{table}
-
 \begin{table}[ht]
@@
 	\hline\hline
-\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & 0.97 \\
-~ & Area Under PR Curve (AUC-PR) & 0.004 & 0.15 \\
+		\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
+		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\
 		\hline
-\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & 0.96 \\
-~ & AUC (PR) & 0.08 & 0.15 \\
+		\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
+		~ & AUC (PR) & 0.08 & \textbf{0.15} \\
 		\hline
-\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & 0.99 \\
-~ & AUC (PR) & 0.13 & 0.28 \\
+		\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
+		~ & AUC (PR) & 0.13 & \textbf{0.28} \\
 		\hline
-\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & xx \\
-~ & AUC (PR) & 0.004 & xx \\
+		\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
+		~ & AUC (PR) & 0.00 & \textbf{0.08} \\
 		\hline
-\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & 1.99 \\
-~ & Root Median Squared Error (RMDSE) & 2.01 & 1.06 \\
+		\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
+		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\
 	\hline\hline
 	\end{tabular}
+	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach.
+Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
+The fifth involved a gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
+Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
+	\label{tab:modelPerfSumm}
 \end{table}

### 2017-08-29 political-analysis-draft/draftPA.tex
switch to simpler homoph sim

@@
 
-% add watermark
-\usepackage{draftwatermark}
-\SetWatermarkText{Draft: Not Ready}
-\SetWatermarkScale{.5}
-\SetWatermarkColor[gray]{0.9}
+% % add watermark
+% \usepackage{draftwatermark}
+% \SetWatermarkText{Draft: Not Ready}
+% \SetWatermarkScale{.5}
+% \SetWatermarkColor[gray]{0.9}
 
@@
 
+\section{\textbf{Limitations of Standard Framework}}
+
+Generalized linear models are typically expressed via a stochastic and systematic component.
+The stochastic component reflects assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function such as the normal, binomial, or Poisson.
+Each dyad in the sample is independently drawn from a particular distribution, given $\theta_{ij}$.
+The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariates, $\mathbf{X}_{ij}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
+A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of the distribution, each of the dyadic observations is conditionally independent.
+
+The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
+After having chosen a set of covariates and specifying a distribution, the joint density function over all dyads using the observations from Table 1 is given:
+
+\vspace{-8mm}
+\begin{align}
+\begin{aligned}
+	P(y_{ij}, y_{ik}, \ldots, y_{lk} | \theta_{ij}, \theta_{ik}, \ldots, \theta_{lk}) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots \times P(y_{lk} | \theta_{lk}) \\
+	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
+\end{aligned}
+\end{align}
+
+\noindent The joint probability is converted into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
+
+The parameters are estimated by maximizing the likelihood.
+The likelihood as defined above is only valid if  $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
+
 \section{\textbf{Additive and Multiplicative Effect Models for Networks}}
@@
 We describe it below.
+
+Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
+
+\begin{align}
+\begin{aligned}
+	y_{ij} &= g(\theta_{ij}) \\
+	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
+	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
+	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
+\label{eqn:ame}
+\end{aligned}
+\end{align}
+
+With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
+In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
+The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
+A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
 
@@
 \citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
+
 \begin{align}
@@
 
-% It has been employed in \citet{hoff:ward:2004, ward:hoff:2007, ward:etal:2007,cao:2009, fletcher:etal:2011, breunig:etal:2012, ward:etal:2012,cao:ward:2014} and is the basis for many recommender systems.
-
 \subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}
@@
 Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
+
 For example, a common finding is that democracies are more likely to form trade agreements with one another.
 Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations dating back to Sun Tzu.
+Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
 A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
@@
 The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
 
-\begin{align}
-\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
-\label{eqn:ame}
-\end{aligned}
-\end{align}
-
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
+\section{Simulation study}
 
@@
 
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.%%%Dont we want to talk more about performance here????
-- MG
+Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
 
@@
 \end{table}
+\FloatBarrier
 

### 2017-08-29 code/gibler_2017/giblerrepl/3_diagPlots.R

@@ -1,0 +1,19 @@
+rm(list = ls())
+
+if(Sys.info()['user']=='s7m'){
+	dropPath = '~/Dropbox/Research/netsMatter/replications/gibler_2017/'
+	funcPath = '~/Research/netsMatter/code/helpers/'
+	resultsPath = paste0(dropPath, 'outputData/')	
+}
+
+# load data and past results
+library(amen)
+load(paste0(dropPath, "amenData_gibler2.rda"))
+load(paste0(resultsPath, 'mdwameFit_k2.rda'))
+
+#
+gofPlot(ameFit$GOF, TRUE)
+paramPlot(ameFit$BETA)
+
+source(paste0(funcPath, 'ameHelpers.R'))
+addEffPlot(ameFit)

### 2017-08-29 code/helpers/ameHelpers.R

@@ -1,0 +1,48 @@
+# AB Effect plot
+getAddEffData = function(fit, row=TRUE, addDegree=FALSE, yList=NULL, orderByDegree=FALSE){
+	if(row){addEffData = data.frame(addEff=fit$APM, stringsAsFactors = FALSE) ; yLabel='Sender Effects'}
+	if(!row){addEffData = data.frame(addEff=fit$BPM, stringsAsFactors = FALSE) ; yLabel='Receiver Effects'}
+	addEffData$actor = rownames(addEffData) ; rownames(addEffData) = NULL
+	if(!orderByDegree){
+		addEffData$actor = factor(addEffData$actor, 
+			levels=addEffData[order(addEffData$addEff),'actor'])
+	}
+	if(addDegree){
+		yArr = listToArray(
+			actors=sort(unique(unlist(lapply(yList,rownames)))), 
+			Y=yList, Xdyad=NULL, Xrow=NULL, Xcol=NULL)$Y
+		if(row){ degree = sort(apply(yArr, 1, mean, na.rm=TRUE)) }
+		if(!row){ degree = sort(apply(yArr, 2, mean, na.rm=TRUE)) }
+		if(orderByDegree){ 
+			addEffData$actor = factor(addEffData$actor, 
+				levels=names(degree) )
+		}
+		addEffData$var = 'Additive Effect'
+		tmp = addEffData ; tmp$addEff = degree[match(tmp$actor,names(degree))] ; tmp$var=' Avg. Degree'
+		addEffData = rbind(addEffData, tmp) ; rm(tmp)
+	}
+	addEffData$max = ifelse(addEffData$addEff>=0,addEffData$addEff,0)
+	addEffData$min = ifelse(addEffData$addEff<0,addEffData$addEff,0) 
+	return(addEffData)
+}
+
+addEffPlot = function(fit, row=TRUE, addDegree=FALSE, yList=NULL, orderByDegree=FALSE, addEffData=NULL){
+	if(is.null(addEffData)){
+		addEffData = getAddEffData(fit, row, addDegree, yList, orderByDegree)
+	}
+	if(row){ yLabel='Sender Effects'}
+	if(!row){ yLabel='Receiver Effects'}		
+	gg = ggplot(addEffData, aes(x=actor, y=addEff)) +
+		geom_point() + geom_linerange(aes(ymax=max,ymin=min)) +
+		ylab(yLabel) + xlab('') + 
[…]

### 2017-08-29 code/simulation/1_ameSim.R

@@ -1,71 +1,54 @@
 rm(list=ls())
-library(statnet)
-library(latentnet)
-library(ergm)
-# install.packages('amen')
-library(amen)
-library(foreach)
-library(doParallel)
 
-### Simulate dataset with intercept val of 1 and dyadvar val of 2
-set.seed(6886)
-n = 30
-sims = 100
-cores = 100
-
-cl = makeCluster(cores) ; registerDoParallel(cl)
-simResults = foreach(i = 1:sims, .packages=c('amen','statnet','latentnet','ergm')) %dopar% {
-	edgeVar = matrix(rnorm(n * n, 0, 1), nrow=n, ncol=n)
-	nw = simulate.formula(
-	  network(n) ~ edges + edgecov(edgeVar) + idegree(1) + odegree(1) + mutual + balance, 
-	  coef=c(-3, .5, .2, .2, .3, .05))
-
-	par(mfrow=c(1,1))
-	plot(nw)
-
-	### run ergm
-	# ergmFit = ergm(nw~edges + edgecov(edgeVar) + balance)
-
-	## run lsm
-	# lsmEuclFit = ergmm(nw ~ euclidean(2) + edgecov(edgeVar))
-	# lsmBilFit = ergmm(nw ~ bilinear(2) + edgecov(edgeVar))
-
-	### run ame
-	nwAdj = as.matrix.network(nw) ; diag(edgeVar) = NA
-	xArr = array(edgeVar, dim=c(dim(edgeVar)[1:2], 1))
-	ameFit_k0 = ame(Y=nwAdj, Xdyad = xArr,
-	                symmetric=FALSE, model='bin',
-	                intercept=TRUE, R=0, 
-	                # rvar=FALSE, cvar=FALSE, dcor=FALSE, nvar=FALSE,
-	                seed=6886, plot=FALSE, print=FALSE
-	                # ,nscan=50000, burn=25000, odens=50
-	                )
-	ameFit_k2 = ame(Y=nwAdj, Xdyad = xArr, 
-	                symmetric=FALSE, model='bin',
-	                intercept=TRUE, R=2,
-	                # rvar=FALSE, cvar=FALSE, dcor=FALSE, nvar=FALSE,
-	                seed=6886, plot=FALSE, print=FALSE
[…]

### 2017-08-29 code/weeks_2012/04-WeeksBasicPosterior.R

@@ -9,7 +9,6 @@ if(Sys.info()['user']== 'margaret'){
     source('~/projects/netsmatter/code/netsMatter/code/weeks_2012/setup.R')
 }
 
-
 loadPkg=function(toLoad){
     for(lib in toLoad){
             if(!(lib %in% installed.packages()[,1])){ 

### 2017-08-31 paper/netsMatter.tex
sim stuff

--- 2017-08-29 political-analysis-draft/draftPA.tex
@@
 
+Here we consider inference for a regression parameter $\beta$ of a linear or generalized linear model for a network in the case where there is an omitted variable.
+The true data-generating models we consider are of the form
+
+$y_{i,j} \sim  \mu + \beta x_i x_j + \gamma w_i w_j + \epsilon_{i,j} $
+
+where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_i \} \in \mathbb R^n$ is a vector of observed node-specific characteristics, and $w = \{ w_i\} \in \mathbb R^n$ is a vector of unobserved node-specific characteristics.
+We compare inference for $\beta$ using three models:
+
+\begin{itemize}
+	\item A naive regression model assuming independent errors; 
+	\item A latent factor model; 
+	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
+\end{itemize}
+
+We make these comparisons in the context of a binary network outcome with a probit regression model.
+
+One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
+
+Alternatively, bias and precision can be summarized by confidence interval coverage and width.
+Coverage should ideally be at the nominal level.
+If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
+
+\begin{figure}
+	\centering
+	\caption{Bias in parameter when homophily is ignored.}
+	\label{fig:ameBias}
+	\includegraphics[width=1\textwidth]{ameSimBias_mu.pdf} \\
+	\includegraphics[width=1\textwidth]{ameSimBias_beta.pdf}
+\end{figure}
+
 \section{Replications}

### 2017-08-31 code/simulation/3_ameCalib.R

@@ -1,0 +1,106 @@
+##############################
+rm(list=ls())
+if(Sys.info()['user']=='s7m'){
+	fPath = '~/Research/netsMatter/code/helpers/'
+	dPath = '~/Dropbox/Research/netsMatter/'
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste('~/Research/netsMatter/paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+toLoad = c(
+	'devtools', 
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+	'latex2exp', 'Cairo'
+	)
+loadPkg(toLoad)
+facet_labeller = function(string){ TeX(string) }
+##############################
+
+##############################
+# params
+NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# load sim results
+load(paste0(simResPath, 'ameSim30.rda')) ; load(paste0(simResPath, 'ameSim50.rda'))
+load(paste0(simResPath, 'deprc/ameSim100.rda'))
+
+#
+modKey = data.frame(dirty=names(ameSim30[[1]]$beta))
+modKey$clean = c('Naive', 'AME', 'Oracle')
+##############################
+
+getCoverage = function(ameSim, model, n, varName, varNum, actual){
+	intSumm = data.frame( do.call('rbind', lapply(ameSim, function(x){
+		quantile(x$beta[[model]][,varNum], probs=c(0.025, 0.975))
+	}) ) )
+	intSumm$coverage = ( intSumm[,1]<actual & actual<intSumm[,2] )
+	intSumm$model = model
+	intSumm$n = n
+	intSumm$varName 
+	return(intSumm)	
+}
+
+
+ameSimCover = lapply(c('naive','ame','oracle'), function(m){
+	getCoverage(ameSim30, model=m, n=30, varName='X1', varNum=2, actual=x1Eff) })
+
+ameSimCover = rbind(
+	getCoverage(ameSim30, model='naive', n=30, varName='X1', varNum=2, actual=x1Eff),
+	getCoverage(ameSim30, model='ame', n=30, varName='X1', varNum=2, actual=x1Eff),
[…]

### 2017-09-01 code/simulation/3_ameCalib.R
sim analysis done

@@ -22,14 +22,14 @@ facet_labeller = function(string){ TeX(string) }
 NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 # load sim results
-load(paste0(simResPath, 'ameSim30.rda')) ; load(paste0(simResPath, 'ameSim50.rda'))
-load(paste0(simResPath, 'deprc/ameSim100.rda'))
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
 
 #
-modKey = data.frame(dirty=names(ameSim30[[1]]$beta))
+modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
 modKey$clean = c('Naive', 'AME', 'Oracle')
 ##############################
 
+##############################
 getCoverage = function(ameSim, model, n, varName, varNum, actual){
 	intSumm = data.frame( do.call('rbind', lapply(ameSim, function(x){
 		quantile(x$beta[[model]][,varNum], probs=c(0.025, 0.975))
@@ -37,70 +37,73 @@ getCoverage = function(ameSim, model, n, varName, varNum, actual){
 	intSumm$coverage = ( intSumm[,1]<actual & actual<intSumm[,2] )
 	intSumm$model = model
 	intSumm$n = n
-	intSumm$varName 
+	intSumm$varName = varName
+	intSumm$actual = actual
 	return(intSumm)	
 }
 
-
-ameSimCover = lapply(c('naive','ame','oracle'), function(m){
-	getCoverage(ameSim30, model=m, n=30, varName='X1', varNum=2, actual=x1Eff) })
-
-ameSimCover = rbind(
-	getCoverage(ameSim30, model='naive', n=30, varName='X1', varNum=2, actual=x1Eff),
-	getCoverage(ameSim30, model='ame', n=30, varName='X1', varNum=2, actual=x1Eff),
-	getCoverage(ameSim30, model='oracle', n=30, varName='X1', varNum=2, actual=x1Eff),
-	)
-
-
-
-FIT0_x1Eff_n30 = do.call('rbind',
-	lapply(ameSim30, function(x){
-		quantile(x$beta$'naive'[,2],probs=c(.025,.975))
-	}) )
-
-FIT1_x1Eff_n30 = do.call('rbind',
-	lapply(ameSim30, function(x){
-		quantile(x$beta$'ame'[,2],probs=c(.025,.975))
-	}) )
-
-FITO_x1Eff_n30 = do.call('rbind',
[…]

### 2017-09-01 code/simulation/4_corUVandMiss.R

@@ -22,12 +22,10 @@ facet_labeller = function(string){ TeX(string) }
 NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 # load sim results
-load(paste0(simResPath, 'ameSim30.rda'))
-load(paste0(simResPath, 'ameSim50.rda'))
-load(paste0(simResPath, 'deprc/ameSim100.rda'))
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
 
 #
-modKey = data.frame(dirty=names(ameSim30[[1]]$beta))
+modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
 modKey$clean = c('Naive', 'AME', 'Oracle')
 ##############################
 
@@ -38,9 +36,55 @@ genMissVar = function(n, imps=NSIM){
 	lapply(1:imps, function(seed){ n=n ; 
 		set.seed(seed) ; xw = matrix(rnorm(n*2),n,2)
 		W = tcrossprod(xw[,2]) ; return(W) }) }
-W30 = genMissVar(30)
-W50 = genMissVar(50)
-W100 = genMissVar(100, imps=NSIM/10)
 
+#
+getCor = function(ameSim, W){
+	cor = lapply(1:length(ameSim), function(i){
+		uv = c(ameSim[[i]]$uv$ame)
+		w = c(W[[i]])
+		cor(uv, w, use='pairwise.complete.obs') })	
+	return(unlist(cor)) }
 
+#
+ameSimCor = rbind(
+	cbind(getCor(ameSim50, genMissVar(50)), n=50),
+	cbind(getCor(ameSim100, genMissVar(100)), n=100)
+	) %>% data.frame()
+ameSimCor$n = paste0('n = ',ameSimCor$n)
+ameSimCor$n = factor(ameSimCor$n, levels=paste0('n = ',c(100,50)))
+
+loadPkg('plyr')
+ameSimCorMeans = ddply(ameSimCor, .(n), summarise, sMedian=median(V1))
+ameSimCorDensity = ddply(ameSimCor, .(n), .fun=function(x){
+	tmp = density(x$V1); x1 = tmp$x; y1 = tmp$y
+	q95 = x1 >= quantile(x$V1,0.025) & x1 <= quantile(x$V1,0.975)
+	q90 = x1 >= quantile(x$V1,0.05) & x1 <= quantile(x$V1,0.95)
[…]

### 2017-09-11 paper/netsMatter.tex
email fix
sim analysis done
more mods

@@
  \curraddr{Department of Political Science, Duke University}
- \email{juan.f.tellez@duke.edu}
+ \email{juan.tellez@duke.edu}
 
@@
 
+\begin{figure}
+	\centering
+	\caption{Coverage in parameter when homophily is ignored.}
+	\label{fig:ameCalib}
+	\includegraphics[width=1\textwidth]{ameSimCover_mu.pdf} \\
+	\includegraphics[width=1\textwidth]{ameSimCover_beta.pdf}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\caption{Correlation between missing variable and multiplicative random effect in AME.}
+	\label{fig:ameCorr}
+	\includegraphics[width=1\textwidth]{ameSimCorr.pdf} \\
+\end{figure}
+
 \section{Replications}
@@
 \caption{Features of the Studies Replicated. }
-\scalebox{.8}{
-	\begin{tabular}{ccrcc}
-		Study & Model & N. of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \toprule
-		Weeks (2012) & Logit & $901,540$ & Directed & Robust \\
-		Reiter \& Stam (2003) & Logit & $753,456$ & Directed & Robust \\
-		Rose (2004) &OLS&$234,597$&Directed&Robust \\
-		McDonald (2004) & Logit &$92,354$& Undirected &Robust\\
-		Gibler (2017) &Logit& $650,557$&UnDirected& no \\
-	\bottomrule
+	\begin{tabular}{lcccccc}
+		 & Model & \# of Actors & Years  & \# of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \hline\hline
+		Weeks (2012) & Logit & & & $901,540$ & Directed & Robust \\
+		Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\
+		McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\		
+		Gibler (2017) & Logit & & & $650,557$ & UnDirected & no \\		
+		Rose (2004) & OLS & & & $234,597$ & Directed & Robust \\
+	\hline\hline
 	\end{tabular}
-}
 \end{table}
@@
 		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-		\hline
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
@@
 		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 		
+		\hline
+		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
 	\hline\hline

### 2017-09-14 paper/netsMatter.tex
reorg

@@
 
-\title{How to fix IR}
+\title{International Relations are a Social Network:\\ A Latent Network Approach for Global Politics}
 \vspace{\baselineskip}
@@
 \lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{How to Fix IR}}
+\chead{\fancyplain{}{How to fix IR}}
 \rhead{\fancyplain{}{\today}}
@@
 
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM), either OLS or logistic regression, to estimate regression coefficients.
-However, we know that extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
+\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.
+However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
 Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
@@
 Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
-%We note in closing that the computational requirements of this approach are substantial in terms of computing time. Cloud and parallel methods reduce this burden.
 
@@
 
-\section{\textbf{Additive and Multiplicative Effect Models for Networks}}
-
-The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
-It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
-Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
-Fortunately, it is easy to empirically examine dependencies in relational data.
-
-The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
-It was introduced in \cite{hoff:2008} and elaborated in \cite{hoff:2015}; the open-source computer package associated with this approach is \cite{amenpkg}.
-We describe it below.
-
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
-\label{eqn:ame}
-\end{aligned}
-\end{align}
-
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
-
-\subsection*{\textbf{Addressing Dependencies in Dyadic Data}}
+\section{\textbf{Dependencies in Dyadic Data}}
 
@@
 Why ignore the interdependencies that led to the study of IR in the first place?
+
+The additive part described above does not handle third-order dependency.
+A third-order dependency is the dependency between triads, not dyads.
+Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
+Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
+
+For example, a common finding is that democracies are more likely to form trade agreements with one another.
+Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
+Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
+A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
+Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
+
+Another third-order dependence pattern is stochastic equivalence.
+Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
+Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
+Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
+Latent variables can account for third-order dependence patterns within the context of the SRRM.
+These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
+
+\section{\textbf{Additive and Multiplicative Effect Models for Networks}}
+
+The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
+It is flexible and easy to use for analyzing the kind of relational data often found in social science.
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
+We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
+Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
+Fortunately, it is easy to empirically examine dependencies in relational data.
+
+The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
+It was introduced in \cite{hoff:2008}; the open-source computer package associated with this approach is \cite{amenpkg}.
+We describe it below.
+
+Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
+
+\begin{align}
+\begin{aligned}
+	y_{ij} &= g(\theta_{ij}) \\
+	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
+	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
+	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
+\label{eqn:ame}
+\end{aligned}
+\end{align}
+
+With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
+In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
+The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
+A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
 
@@
 
-The additive part described above does not handle third-order dependency.
-A third-order dependency is the dependency between triads, not dyads.
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
-
-For example, a common finding is that democracies are more likely to form trade agreements with one another.
-Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
-A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
-Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
-Another third-order dependence pattern is stochastic equivalence.
-Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
-Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
-Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
-Latent variables can account for third-order dependence patterns within the context of the SRRM.
-These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
-
-A number of latent variable approaches have been developed to represent third-order dependencies in relational data, we focus on the latent factor model (LFM).
-This expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
+The latent factor model (LFM) expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
 General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.

### 2017-09-29 paper/1_intro.tex
sync up

\section{\textbf{Introduction}}

\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}. Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.  However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.} Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases. This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements. It is not just the diagonals of the variance-covariance matrix this affects.

In this article, we introduce a Bayesian approach with additive and multiplicative effects for directly modeling relational data to reflect the data generating process that yields interdependencies in relational data. We focus on three types of interdependencies that can complicate dyadic analyses. First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict. Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad, specifically, if a rebel group initiates a conflict with a government, the government will reciprocate that conflictual behavior. We capture these two dependencies, referred to as first- and second-order dependencies within the additive effects portion of the model. The multiplicative effects portion of the model captures dependencies that results from groups of actors clustering together or organizing into communities. Clusters and communities often result in relational data because actors possess some set of shared attributes that affect their probability of interacting with one another. For example, when considering alliance patterns we might reasonably expect that if two actors are already allied to a third that their probability of allying with one another is higher than average.

Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interact- ing with one another. Individuals try to close triads because this is thought to be a more stable or preferable social situation

a specific actor is responsible for sent or received events such

This model has two parts, one an additive effects portion that

, one of which uses a set of additive random effects to capture dependencies arising because a specific actor sent or received an events such as conflict to others.

\first- and \second-order dependencies in the relational data. The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.  We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.

This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches. It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations. Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.

### 2017-09-29 paper/2_model.tex
sync up

--- 2017-09-14 paper/netsMatter.tex
@@
-\documentclass[12pt]{amsart} 
-      % \newif\ifpreview
-      % \previewfalse
-% approx iid
-\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}\usepackage{natbib}
-
-% Fonts
-\usepackage[default,osfigures,scale=0.95]{opensans}
-\usepackage[T1]{fontenc}
-\usepackage{ae}
-
-% easy commands for number propers
-\newcommand{\first}{1$^{\text{st}}$}
-\newcommand{\second}{2$^{\text{nd}}$}
-\newcommand{\third}{3$^{\text{rd}}$}
-\newcommand{\nth}[1]{${#1}^{\text{th}}$}
-
-% Generate some fake text
-\usepackage{blindtext}
-
-% % add watermark
-% \usepackage{draftwatermark}
-% \SetWatermarkText{Draft: Not Ready}
-% \SetWatermarkScale{.5}
-% \SetWatermarkColor[gray]{0.9}
-
-% layout control
-\usepackage{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
-\usepackage{parallel}
-\usepackage{parcolumns}
-\usepackage{fancyhdr}
-\usepackage[export]{adjustbox}
-\usepackage{etoolbox}
-
-% math typesetting
-\usepackage{array}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{relsize}
-\usepackage{mathtools}
-\usepackage{bm}
-%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}
-
-% restricts float objects to be inserted before end of section
-% creates float barriers
-\usepackage[section]{placeins}
-
-% tables
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{longtable}
-
-% to adapt caption style
-\usepackage[font={small},labelfont=bf]{caption}
-
-% footnotes at bottom
-%\usepackage[bottom]{footmisc}
-
-% to change enumeration symbols begin{enumerate}[(a)]
-\usepackage{enumerate}
-\usepackage{paralist}
-
-% to colorize links in document. See color specification below
-\usepackage[pdftex,hyperref,x11names]{xcolor}
-
-% for multiple references and insertion of the word "figure" or "table"
-% \usepackage{cleveref}
-
-% load the hyper-references package and set document info
-\usepackage[pdftex]{hyperref}
-
-% graphics stuff
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage[space]{grffile} % allows us to specify directories that have spaces
-\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
-\usepackage{tikz}
-
-% sideway figures
-\usepackage{rotating}
-\usepackage{lscape}
-
-%\usepackage[figuresright]{rotating}
-% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
-%   {\end{minipage}\end{sidewaysfigure}}
-
-% Spacing
-\usepackage[doublespacing]{setspace}
-
-% define clickable links and their colors
-\hypersetup{%
-	unicode=false,          % non-Latin characters in Acrobat's bookmarks
-	pdftoolbar=true,        % show Acrobat's toolbar?
-	pdfmenubar=true,        % show Acrobat's menu?
-	pdffitwindow=false,     % window fit to page when opened
-	pdfstartview={FitH},    % fits the width of the page to the window
-	pdfnewwindow=true,%
-	%pagebackref=false,%
-	pdfauthor={Shahryar Minhas, et alia.},%
-	pdftitle={How to Fix IR},%
-	colorlinks,%
-	citecolor=black,%
-	filecolor=black,%
-	linkcolor=black,%
-	urlcolor=RoyalBlue4
-}
-
-% -------------------- title -------------------- %
-
-\title{International Relations are a Social Network:\\ A Latent Network Approach for Global Politics}
-\vspace{\baselineskip}
-
- \author[Minhas]{Shahryar Minhas}
- \address{Shahryar Minhas: Department of Political Science}
- \curraddr{Michigan State University}
- \email[Corresponding author]{minhassh@msu.edu}
-
- \author[Dorff]{Cassy L. Dorff}
- \address{Cassy L. Dorff}
- \curraddr{Department of Political Science, University of New Mexico}
- \email{cdorff@unm.edu}
-
- \author[Foster]{Margaret Foster}
- \address{Margaret Foster}
- \curraddr{Department of Political Science, Duke University}
- \email{margaret.foster@duke.edu}
-
- \author[Gallop]{Max Gallop}
- \address{Max Gallop}
- \curraddr{Department of Government and Public Policy, University of Strathclyde}
- \email{max.gallop@strath.ac.uk}
-
-  \author[Liu]{Howard Liu}
- \address{Howard Liu}
- \curraddr{Department of Political Science, Duke University}
- \email{hao.liu@duke.edu}
-
- \author[Tellez]{Juan Tellez}
- \address{Juan Tellez}
- \curraddr{Department of Political Science, Duke University}
- \email{juan.tellez@duke.edu}
-
-\author[Ward]{Michael D. Ward}
-\address{Michael D. Ward}
-\curraddr{Department of Political Science, Duke University}
-\email{michael.d.ward@duke.edu}
-
-% \date{\today}
-
-\thanks{Shahryar Minhas and Michael D.
-Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings.
-We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
-
-\setlength{\headheight}{15pt}
-\setlength{\headsep}{20pt}
-\pagestyle{fancyplain}
-\fancyhf{}
-\lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{How to fix IR}}
-\rhead{\fancyplain{}{\today}}
-\rfoot{\fancyplain{}{\thepage}}
-% references to graphics
-\makeatletter
-% let's keep them in the same directory \def\input@path{{}
-
-% easy command for boldface math symbols
-\newcommand{\mbs}[1]{\boldsymbol{#1}}
-
-% command for R package font
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
-
-% Add some colors
-\definecolor{red1}{RGB}{253,219,199}
-\definecolor{red2}{RGB}{244,165,130}
-\definecolor{red3}{RGB}{178,24,43}
-
-\definecolor{green1}{RGB}{229,245,224}
-\definecolor{green2}{RGB}{161,217,155}
-\definecolor{green3}{RGB}{49,163,84}
-
-\definecolor{blue0}{RGB}{255,247,251}
-\definecolor{blue1}{RGB}{222,235,247}
-\definecolor{blue2}{RGB}{158,202,225}
-\definecolor{blue3}{RGB}{49,130,189}
-\definecolor{blue4}{RGB}{4,90,141}
-
-\definecolor{purple1}{RGB}{191,211,230}
-\definecolor{purple2}{RGB}{140,150,198}
-\definecolor{purple3}{RGB}{140,107,177}
-
-\definecolor{brown1}{RGB}{246,232,195}
-\definecolor{brown2}{RGB}{223,194,125}
-\definecolor{brown3}{RGB}{191,129,45}
-
-% square bracket matrices
-\let\bbordermatrix\bordermatrix
-\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
-\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
-\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}
-
-\begin{document}
-      % \ifpreview {
-      %    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
-      %    \pagecolor{personal}
-      %    }
-      % \fi
-
-\small{\singlespacing{
-\begin{abstract}
-	International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially data.
-These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.
-However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce results that are not compelling.
-One reason is that the independence required by the typical methods employed does not exist in the data being studied.
-We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies.
-We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
-\end{abstract}
-}}
-
-\maketitle\thispagestyle{empty}
-\newpage\setcounter{page}{1}
-
-\section{\textbf{Introduction}}
-
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.
-However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
-It is not just the diagonals of the variance-covariance matrix this affects.
-
-In this article, we introduce a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in the data.
-This model has two parts, one of which uses a set of additive random effects capturing \first- and \second-order dependencies in the relational data.
-The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.
-We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a generalized linear model with robust standard errors) to those obtained with the AME framework.
-
-This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
-
 \section{\textbf{Limitations of Standard Framework}}
@@
 The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
-
-\section{Simulation study}
-
-Here we consider inference for a regression parameter $\beta$ of a linear or generalized linear model for a network in the case where there is an omitted variable.
-The true data-generating models we consider are of the form
-
-$y_{i,j} \sim  \mu + \beta x_i x_j + \gamma w_i w_j + \epsilon_{i,j} $
-
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_i \} \in \mathbb R^n$ is a vector of observed node-specific characteristics, and $w = \{ w_i\} \in \mathbb R^n$ is a vector of unobserved node-specific characteristics.
-We compare inference for $\beta$ using three models:
-
-\begin{itemize}
-	\item A naive regression model assuming independent errors; 
-	\item A latent factor model; 
-	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
-\end{itemize}
-
-We make these comparisons in the context of a binary network outcome with a probit regression model.
-
-One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
-
-Alternatively, bias and precision can be summarized by confidence interval coverage and width.
-Coverage should ideally be at the nominal level.
-If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
-
-\begin{figure}
-	\centering
-	\caption{Bias in parameter when homophily is ignored.}
-	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{ameSimBias_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimBias_beta.pdf}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{Coverage in parameter when homophily is ignored.}
-	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{ameSimCover_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimCover_beta.pdf}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{Correlation between missing variable and multiplicative random effect in AME.}
-	\label{fig:ameCorr}
-	\includegraphics[width=1\textwidth]{ameSimCorr.pdf} \\
-\end{figure}
-
-\section{Replications}
-
-\subsection{Design}
-
-For the purpose of this study, we choose five prominent studies from the broad field of international relations and international political economy that utilize relation data \citep{mcdonald:2004, reiter:stam:2003, rose:2004, weeks:2012, gibler:2017}.
-We chose to replicate studies that were written fairly recently, since 2000, and that have been cited over 100 times.
-Each of these pieces was published in a prominent journal and is well-known in the literature.
-They all used the standard approach in political science, which is to employ some form of general linearized regression that ignores dyadic interdependencies.
-Post estimation, standard errors are often adjusted in an attempt to account for clustering of observations.
-
-\begin{table}
-\caption{Features of the Studies Replicated. }
-	\begin{tabular}{lcccccc}
-		 & Model & \# of Actors & Years  & \# of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \hline\hline
-		Weeks (2012) & Logit & & & $901,540$ & Directed & Robust \\
-		Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\
-		McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\		
-		Gibler (2017) & Logit & & & $650,557$ & UnDirected & no \\		
-		Rose (2004) & OLS & & & $234,597$ & Directed & Robust \\
-	\hline\hline
-	\end{tabular}
-\end{table}
-
-We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
-
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
-We believe that it does, and that the dynamic latent factor model provides a step forward.
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l| p{7cm} l}
-	& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
-	& &  in a Network Model? \\
-	\hline\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\
-		\hline
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
-		\hline
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
-		\hline
-		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 		
-		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
-Cases in which the finding is not replicated are highlighted in bold.}
-	\label{tab:modelFindingSumm}
-\end{table}
-
-We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
-
-% % % % % % % % % % % % % % % % % % 
-By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
-To show that this is the case, we examine whether our approach achieves better predictive performance in an out of sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-Then for each $s \in \{1,\ldots,k\}$, we:
-
-\begin{enumerate}
-	\item estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$,
-	\item and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters. 
-\end{enumerate}
-
-\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-
-We set a number of benchmarks for comparison.
-First we compare the AME model to a GLM model using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
-We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
-The lagged dependent variable is the equivalent of saying that conflict and peace are relatively likely to persist between dyads, while the inclusion of a lagged reciprocity term in a GLM framework is a simple way to account for retaliatory strikes.
-
-We utilize three performance criterions to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
-ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
-An issue with an ROC Curve when looking at conflict, is that it is relatively rare at the dyadic level: in most years only 3\% of possible dyads are in conflict with one another.
-If peace is common, even a poor model will have a very low False Positive Rate.
-
-To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
-These examine the trade-offs between the percentage of conflicts a model predicts, and the percentage of predicted conflicts which occur.
-Lastly, we examine separation plots \citep{greenhill:etal:2011}.
-These provide an intuitive visualization of the accuracy of our predictions by juxtaposing a line showing the predicted probability of conflict with whether conflict actually occurs for all cases (where the cases are sorted by the predicted probability and then colored to indicate the outcome).
-Here a perfect model would have all cases where conflict actually exists on the right with a predicted probability of $1$, and would predict $0$ in all other cases.
-All of the models' performance out of sample by these metrics are displayed in figure \ref{fig:roc_ame}.
-The AME model with covariates is the best performing model out of sample in all cases.
-This model outperforms each of the GLM variants by a notable margin.
-% % % % % % % % % % % % % % % % % %
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l|l cc}
-	~ & ~ & GLM & AME \\
-	\hline\hline
-		\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
-		~ & AUC (PR) & 0.08 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
-		~ & AUC (PR) & 0.13 & \textbf{0.28} \\
-		\hline
-		\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
-		~ & AUC (PR) & 0.00 & \textbf{0.08} \\
-		\hline
-		\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
-		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach.
-Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
-The fifth involved a gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
-Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
-	\label{tab:modelPerfSumm}
-\end{table}
-\FloatBarrier
-
-\subsection{Replication of Weeks (2012)}
-
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
-One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-Her time frame is from 1946-1999.
-The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
-She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
-
-The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
-In Weeks's initial models, as detailed in figure \ref{fig:weeksCoef} she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
-When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
-We similarly find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in figure \ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
-In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
-In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
-So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
-The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in figure \ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Reiter \& Stam (2003)}
-
-\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent.
-They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that ``We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
-This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
-We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
-The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
-Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
-
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
-In our replication using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
-Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitauc} and Figure ~\ref{fig:reitpr}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:reitauc}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
-\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
-\end{figure}
-
-\newpage
-\subsection{Replication of Rose (2004)}
-
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.
-Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
-He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders...
-(2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
-
-When we compare the results of Rose's original OLS model to our model that accounts for network dependencies, the results are generally similar.
-As you can see in Figure~\ref{fig:roseCoef}, the main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- remains when we move from an OLS to a Gaussian AME model.
-The most striking difference between the models is that, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies.
-Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-Excitingly, when we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out of sample -- 3.23 for the OLS model and 1.77 for the AME model.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
- \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{ABplot_rose_top10.pdf}
- \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
-\end{figure}
-
-%\begin{figure}
-%\centering   
-%\subfigure[Nodal Random Effects]%{\label{fig:rosesend}\includegraphics[width=65mm]{ABplot_rose_top10.pdf}}
-%\subfigure[Receiver Random Effects]{\label{fig:rosereceive}\includegraphics[width=65mm]{ABplot_receiver_rose.pdf}}
-%\caption{Sender and receiver random effects for AME estimation of Rose (2004).}
-%\label{fig:rosesr}
-%\end{figure}
-
-\newpage
-\subsection{Replication of McDonald (2004)}
-
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-He observes that knowledge about the link between conflict and trade is indeterminate in the field of international relations, noting that competing explanations persist.
-He calls for more precise empirical tests.
-Importantly, \citet{mcdonald:2004} includes the interdependence argument---that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)'' in his overview of underdeveloped hypotheses.
-
-Accordingly, the primary contribution of the study is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-To do so, \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measures free trade in two ways.
-The first  captures the idea that larger protected sectors generate greater societal pressures resulting in  pockets of support for war.
-This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state i's GDP or total dyadic trade divided by state j's GDP.'' (560).
-Finally, the binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include a spline correction for time-series data as well as robust standard errors clustered on each dyad.
-
-% TO DO: include discussion and analysis of different K's in appendix (for M cD)
-\begin{figure}
-\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-\end{figure}
-
-% TO DO: fix bib
-
-Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries {i} and {j} are highly dependent and countries {j} and {k} are also highly dependent, then we are likely to observe high dependency between countries {i} and {k}.
-Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
-This indicates that conflict is less likely between members of a trade community.
-Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and replicated model and Figure~\ref{fig:mcdall} demonstrates the predictive performance of each model.
-Figure ~\ref{fig:mcdcirc} represents the multiplicative effects of the model.
-This plot enables us to consider actors with similar sending and receiving patterns resulting from stochastic equivalence and homophily.
-In the upper arch for senders, there is a clear clustering of United States' allies (the United Kingdom, Netherlands, and Japan).
-We also observe an African cluster in the top right segment for receivers (Egypt, Niger, and Zambia for example).
-These countries both trade amongst themselves and have similar conflict behavior -- stochastic equivalence is present among the first group, homophily among the second -- and so once we account for these clusters using the multiplicative effects, the residual effect of trade actually increases the likelihood of conflict.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{mcdonaldcirc.pdf}
-\caption{\label{fig:mcdcirc} Visualization of multiplicative effects. Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Gibler (2017)}
-
-% We chose this article because it is published in 2017, demonstrating that the concern with interdependence continues to the present and because it was published in the so-called flagship journal of political science, \textit{The American Political Science Review.} It examines the link between dyadic capability differences and international conflict in terms of how and when states became independent members of the international system.
-It assumes that membership in the international system---defined in typical ways---has a substantial influence on capabilities.
-In the same way that capabilities don't seem to vary enormously over time, at least comparatively, states become independent in a particular location, clustered with surrounding states.
-Neither does this change much over time.
-Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-The replications we have undertaken are all from articles over the past fifteen years.
-A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
-In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-\begin{table}
-\begin{center}
-\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000$ iterations.  \label{tab:gibme}}
-\begin{tabular}{lrr} \toprule
-& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-Allied & 0.142&  0.023 \\
-Joint Democracy &\bf -0.507&  0.045 \\
-Peace Years \bf &-0.260& \bf -0.060 \\
-Spline 1 &\bf -0.001& \bf 0.00 \\
-Spline 2&\bf -0.000& \bf 0.00\\
-Spline 3 &-0.000& \bf 0.00\\ 
-Contiguity &\bf 2.412& \bf 0.640 \\
-Parity &0.075&  -0.013 \\
-Parity at entry year&\bf 0.868&  0.002 \\ 
-Rivalry &\bf 2.031&\bf 0.721\\ 
-Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-\end{tabular}\\
-\end{center}
-{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
-For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
-\end{table}
-
-We re-estimated model 6 from Table 6 (2017,34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
-The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
-Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the amen results.
-Not only is the value of the this parameter small, but it has a very large relative standard error, over a magnitude larger than the parameter itself ($z= 0.038$).
-In addition, joint democracy follows the same pattern of importance in the logistic results, but this disappears once interdependencies are modeled.
-As might be expected the strong geographic clustering in the original study is about one-quarter as strong in the \texttt{amen} estimations.
-Similarly, rivalry coefficients are about one-third the size in the \texttt{amen} formulation, but a great deal more precisely measured ($z=18.116$).
-
-Beyond more informative fixed effect coefficients, the \texttt{amen} approach also provides information about the interdependencies that were modeled.
-The most pertinent of these may be the dyadic and triadic dependencies which are presented in Figure~\ref{fig:gibler:dyadtriad}, which further reveal that the assumption of independence among the dyadic data in this study can be strongly rejected.
-Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies.
-Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.
-
-% sd.rowmean sd.colmean   dyad.dep  triad.dep 
-% 0.03210221 0.03210221 0.99362646 0.00130249 
-\begin{figure}
-\caption{\label{fig:gibler:dyadtriad} The left panel is  a plot of the dyadic dependencies in the Gibler model 6; the right panel shows the triad dependencies.
-Note that the red line represents the null hypothesis of no dependencies, indicating that the standard logistic approach is far from what is uncovered with this analysis.
-}
-\includegraphics[width=\textwidth]{netplots}
-\end{figure}
-
-\section{Conclusion}
-
-It is no longer necessary to assume that the interesting, innate interdependencies in relational data can be ignored.
-Nor do they have to be approximated with ad hoc, incomplete solutions that purport to control for dependencies by modifying the post-estimation standard errors of the estimated coefficients.
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.
-
-%%What about a summary table with AUCs for the different network models here? - MG
-\newpage
-% Bib stuff
-\clearpage
-% \bibliography{master.bib}
-\bibliography{/Users/s7m/whistle/master}
-\bibliographystyle{authordate1}
-\clearpage
-\section*{All About Us}
-\end{document}

### 2017-09-29 paper/3_sim.tex
sync up

--- 2017-09-14 paper/netsMatter.tex
@@
-\documentclass[12pt]{amsart} 
-      % \newif\ifpreview
-      % \previewfalse
-% approx iid
-\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}\usepackage{natbib}
-
-% Fonts
-\usepackage[default,osfigures,scale=0.95]{opensans}
-\usepackage[T1]{fontenc}
-\usepackage{ae}
-
-% easy commands for number propers
-\newcommand{\first}{1$^{\text{st}}$}
-\newcommand{\second}{2$^{\text{nd}}$}
-\newcommand{\third}{3$^{\text{rd}}$}
-\newcommand{\nth}[1]{${#1}^{\text{th}}$}
-
-% Generate some fake text
-\usepackage{blindtext}
-
-% % add watermark
-% \usepackage{draftwatermark}
-% \SetWatermarkText{Draft: Not Ready}
-% \SetWatermarkScale{.5}
-% \SetWatermarkColor[gray]{0.9}
-
-% layout control
-\usepackage{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
-\usepackage{parallel}
-\usepackage{parcolumns}
-\usepackage{fancyhdr}
-\usepackage[export]{adjustbox}
-\usepackage{etoolbox}
-
-% math typesetting
-\usepackage{array}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{relsize}
-\usepackage{mathtools}
-\usepackage{bm}
-%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}
-
-% restricts float objects to be inserted before end of section
-% creates float barriers
-\usepackage[section]{placeins}
-
-% tables
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{longtable}
-
-% to adapt caption style
-\usepackage[font={small},labelfont=bf]{caption}
-
-% footnotes at bottom
-%\usepackage[bottom]{footmisc}
-
-% to change enumeration symbols begin{enumerate}[(a)]
-\usepackage{enumerate}
-\usepackage{paralist}
-
-% to colorize links in document. See color specification below
-\usepackage[pdftex,hyperref,x11names]{xcolor}
-
-% for multiple references and insertion of the word "figure" or "table"
-% \usepackage{cleveref}
-
-% load the hyper-references package and set document info
-\usepackage[pdftex]{hyperref}
-
-% graphics stuff
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage[space]{grffile} % allows us to specify directories that have spaces
-\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
-\usepackage{tikz}
-
-% sideway figures
-\usepackage{rotating}
-\usepackage{lscape}
-
-%\usepackage[figuresright]{rotating}
-% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
-%   {\end{minipage}\end{sidewaysfigure}}
-
-% Spacing
-\usepackage[doublespacing]{setspace}
-
-% define clickable links and their colors
-\hypersetup{%
-	unicode=false,          % non-Latin characters in Acrobat's bookmarks
-	pdftoolbar=true,        % show Acrobat's toolbar?
-	pdfmenubar=true,        % show Acrobat's menu?
-	pdffitwindow=false,     % window fit to page when opened
-	pdfstartview={FitH},    % fits the width of the page to the window
-	pdfnewwindow=true,%
-	%pagebackref=false,%
-	pdfauthor={Shahryar Minhas, et alia.},%
-	pdftitle={How to Fix IR},%
-	colorlinks,%
-	citecolor=black,%
-	filecolor=black,%
-	linkcolor=black,%
-	urlcolor=RoyalBlue4
-}
-
-% -------------------- title -------------------- %
-
-\title{International Relations are a Social Network:\\ A Latent Network Approach for Global Politics}
-\vspace{\baselineskip}
-
- \author[Minhas]{Shahryar Minhas}
- \address{Shahryar Minhas: Department of Political Science}
- \curraddr{Michigan State University}
- \email[Corresponding author]{minhassh@msu.edu}
-
- \author[Dorff]{Cassy L. Dorff}
- \address{Cassy L. Dorff}
- \curraddr{Department of Political Science, University of New Mexico}
- \email{cdorff@unm.edu}
-
- \author[Foster]{Margaret Foster}
- \address{Margaret Foster}
- \curraddr{Department of Political Science, Duke University}
- \email{margaret.foster@duke.edu}
-
- \author[Gallop]{Max Gallop}
- \address{Max Gallop}
- \curraddr{Department of Government and Public Policy, University of Strathclyde}
- \email{max.gallop@strath.ac.uk}
-
-  \author[Liu]{Howard Liu}
- \address{Howard Liu}
- \curraddr{Department of Political Science, Duke University}
- \email{hao.liu@duke.edu}
-
- \author[Tellez]{Juan Tellez}
- \address{Juan Tellez}
- \curraddr{Department of Political Science, Duke University}
- \email{juan.tellez@duke.edu}
-
-\author[Ward]{Michael D. Ward}
-\address{Michael D. Ward}
-\curraddr{Department of Political Science, Duke University}
-\email{michael.d.ward@duke.edu}
-
-% \date{\today}
-
-\thanks{Shahryar Minhas and Michael D.
-Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings.
-We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
-
-\setlength{\headheight}{15pt}
-\setlength{\headsep}{20pt}
-\pagestyle{fancyplain}
-\fancyhf{}
-\lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{How to fix IR}}
-\rhead{\fancyplain{}{\today}}
-\rfoot{\fancyplain{}{\thepage}}
-% references to graphics
-\makeatletter
-% let's keep them in the same directory \def\input@path{{}
-
-% easy command for boldface math symbols
-\newcommand{\mbs}[1]{\boldsymbol{#1}}
-
-% command for R package font
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
-
-% Add some colors
-\definecolor{red1}{RGB}{253,219,199}
-\definecolor{red2}{RGB}{244,165,130}
-\definecolor{red3}{RGB}{178,24,43}
-
-\definecolor{green1}{RGB}{229,245,224}
-\definecolor{green2}{RGB}{161,217,155}
-\definecolor{green3}{RGB}{49,163,84}
-
-\definecolor{blue0}{RGB}{255,247,251}
-\definecolor{blue1}{RGB}{222,235,247}
-\definecolor{blue2}{RGB}{158,202,225}
-\definecolor{blue3}{RGB}{49,130,189}
-\definecolor{blue4}{RGB}{4,90,141}
-
-\definecolor{purple1}{RGB}{191,211,230}
-\definecolor{purple2}{RGB}{140,150,198}
-\definecolor{purple3}{RGB}{140,107,177}
-
-\definecolor{brown1}{RGB}{246,232,195}
-\definecolor{brown2}{RGB}{223,194,125}
-\definecolor{brown3}{RGB}{191,129,45}
-
-% square bracket matrices
-\let\bbordermatrix\bordermatrix
-\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
-\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
-\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}
-
-\begin{document}
-      % \ifpreview {
-      %    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
-      %    \pagecolor{personal}
-      %    }
-      % \fi
-
-\small{\singlespacing{
-\begin{abstract}
-	International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially data.
-These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.
-However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce results that are not compelling.
-One reason is that the independence required by the typical methods employed does not exist in the data being studied.
-We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies.
-We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
-\end{abstract}
-}}
-
-\maketitle\thispagestyle{empty}
-\newpage\setcounter{page}{1}
-
-\section{\textbf{Introduction}}
-
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.
-However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
-It is not just the diagonals of the variance-covariance matrix this affects.
-
-In this article, we introduce a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in the data.
-This model has two parts, one of which uses a set of additive random effects capturing \first- and \second-order dependencies in the relational data.
-The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.
-We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a generalized linear model with robust standard errors) to those obtained with the AME framework.
-
-This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
-
-\section{\textbf{Limitations of Standard Framework}}
-
-Generalized linear models are typically expressed via a stochastic and systematic component.
-The stochastic component reflects assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function such as the normal, binomial, or Poisson.
-Each dyad in the sample is independently drawn from a particular distribution, given $\theta_{ij}$.
-The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariates, $\mathbf{X}_{ij}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
-A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of the distribution, each of the dyadic observations is conditionally independent.
-
-The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
-After having chosen a set of covariates and specifying a distribution, the joint density function over all dyads using the observations from Table 1 is given:
-
-\vspace{-8mm}
-\begin{align}
-\begin{aligned}
-	P(y_{ij}, y_{ik}, \ldots, y_{lk} | \theta_{ij}, \theta_{ik}, \ldots, \theta_{lk}) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots \times P(y_{lk} | \theta_{lk}) \\
-	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
-\end{aligned}
-\end{align}
-
-\noindent The joint probability is converted into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
-
-The parameters are estimated by maximizing the likelihood.
-The likelihood as defined above is only valid if  $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
-
-\section{\textbf{Dependencies in Dyadic Data}}
-
-Relational, or dyadic, data provide measurements of how sets of actors relate to one another.
-Typically, pairs of actors are examined.
-These pairs are called \textit{dyads}.
-Such data can be arranged as a directed dyadic design in which the unit of analysis is some set of $n$ actors that have been paired together to form a matrix of dyads, i.
-e., represented as a $n \times n$ matrix, $\mathbf{Y}$, with diagonals typically undefined.
-The $ij^{th}$ entry defines the relationship of $i$ and $j$ and can be continuous or discrete.
-It can also be symmetric (known as undirected) or asymmetric (called directed).
-This is not the stacked dyad structure normally found in social science, but it is equivalent and presents a format in which the interdependencies are exposed and transparent.
-The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.
-
-One important dependency is nodal dependency, whereby each actor has similar interactions with many of its partners, leading to significant heterogeneity in activity levels across nodes.
-An important implication of this across-node heterogeneity is within-node homogeneity of ties: the values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, are similar to each other because each has a common sender.
-If a sender $i$ is more active active in the network than other senders, it will have greater homogeneity of its linkages.
-More popular targets show greater similarity across linkages in a given column, $\{y_{ji},y_{ki},y_{li}\}$.
-Actors which are more likely to send ties in a network are also more likely to receive them.
-As a result the rows and columns in an adjacency matrix may be correlated.
-Another structural interdependence is reciprocity.
-This is a second-order, or dyadic dependency whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
-If the data are simply stacked dyads in some arbitrary order, this kind of dependency is harder to visualize, but exists nonetheless.
-
-The presence of these interdependencies in relational data complicates the practical assumption of observational independence.
-Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-By ignoring these potential interdependencies, we often ignore important features of the problem under study.
-The study of international relations is founded on the relations among actors.
-Why ignore the interdependencies that led to the study of IR in the first place?
-
-The additive part described above does not handle third-order dependency.
-A third-order dependency is the dependency between triads, not dyads.
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
-
-For example, a common finding is that democracies are more likely to form trade agreements with one another.
-Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
-A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
-Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
-Another third-order dependence pattern is stochastic equivalence.
-Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
-Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
-Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
-Latent variables can account for third-order dependence patterns within the context of the SRRM.
-These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
-
-\section{\textbf{Additive and Multiplicative Effect Models for Networks}}
-
-The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
-It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
-Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
-Fortunately, it is easy to empirically examine dependencies in relational data.
-
-The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
-It was introduced in \cite{hoff:2008}; the open-source computer package associated with this approach is \cite{amenpkg}.
-We describe it below.
-
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
-\label{eqn:ame}
-\end{aligned}
-\end{align}
-
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
-
-\subsection{\textbf{Social Relations Model: Additive Part of AME}}
-
-\citet{warner:etal:1979} developed a social relations model (SRM), using ANOVA decomposition, that focuses on such dependencies.
-SRM provides the error structure for the additive effects component of the AME framework.
-SRM decomposes the variance of dyadic observations in terms of a) heterogeneity across row means (often called out-degree), b) column means (or, in-degree), c) correlation between row and column means, and d) correlations within dyads.
-\citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= \mu + e_{ij} \\
-	e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
-	\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
-	\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
-	\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix},
-\label{eqn:srmCov}
-\end{aligned}
-\end{align}
-
-wherein $\mu$ is the density of a network, and $e_{ij}$ represents residual variation in each cell (or dyad).
-The residual variation has three separate parts: a)   row/sender effects ($a_{i}$), b)  column/receiver effects ($b_{j}$), and c) within-dyad effects ($\epsilon_{ij}$).
-The row and column effects are modeled jointly to account for correlation in how active an actor is in sending and receiving ties.
-Heterogeneity in the row and column means is modeled by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively.
-$\sigma_{ab}$ describes whether actors which send a lot of ties also receive a lot.
-Second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a reciprocity, parameter $\rho$.
-
-The SRM covariance structure described in Equation~\ref{eqn:srmCov} can be incorporated into the systematic component of a GLM framework to produce the social relations regression model (SRRM): $\bm\beta^{\top} \mathbf{X}_{ij} + a_{i} + b_{j} + \epsilon_{ij}$, where $ \bm\beta^{\top} \mathbf{X}_{ij}$ accommodates the inclusion of dyadic, sender, and receiver covariates \citep{hoff:2005}.
-This is a regression framework with additive, random effects, but it models variances and covariances often seen in relational data.
-Furthermore, this accommodates a diversity of outcome distributions.
-In the case of binary data, this can be done by utilizing a random effects logistic (or probit) latent variable representation.
-
-\subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}
-
-The latent factor model (LFM) expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
-General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.
-
-\begin{align}
-\begin{aligned}
-	\alpha(&\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{\top} \Lambda \textbf{u}_{j} \\
-	&\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \text{, where}\\
-	 &\Lambda \equiv K \times K \text{ diagonal matrix}
-\label{eqn:latAlpha}
-\end{aligned}
-\end{align}
-
-Many observed  networks exhibit varying degrees of stochastic equivalence and homophily.
-In the latent factor model, each actor has an unobserved vector of characteristics, $\textbf{u}_{i} = \{u_{i,1}, \ldots, u_{i,K} \}$, describing each node's behavior in the network.
-The probability of a tie between $i$ and $j$ depends on the extent to which $\textbf{u}_{i}$ and $\textbf{u}_{j}$ are ``similar'' and on whether the entries of $\Lambda$ are greater or less than zero.
-Similarity in the latent factors, $\textbf{u}_{i} \approx \textbf{u}_{j}$, corresponds to stochastic equivalence; the eigenvalue determines whether the network exhibits positive or negative homophily.
-Accordingly, this approach can represent both homophily and stochastic equivalence.
-
-In directed (asymmetric) networks there are two sets of latent positions, denoted $\textbf{u}$ and $\textbf{v}$, representing the sending and receiving positions separately.
-Thus, actors in the network have a vector of latent characteristics describing their behavior as a sender, denoted by $\textbf{u}$, and as a receiver, $\textbf{v}$: $\textbf{u}_{i}, \textbf{v}_{j} \in \mathbb{R}^{K}$.
-This can alter the probability of an interaction between $ij$ additively: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, where $\textbf{D}$ is a $K \times K$ diagonal matrix.
-
-The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
-
 \section{Simulation study}
@@
 \end{figure}
-
-\section{Replications}
-
-\subsection{Design}
-
-For the purpose of this study, we choose five prominent studies from the broad field of international relations and international political economy that utilize relation data \citep{mcdonald:2004, reiter:stam:2003, rose:2004, weeks:2012, gibler:2017}.
-We chose to replicate studies that were written fairly recently, since 2000, and that have been cited over 100 times.
-Each of these pieces was published in a prominent journal and is well-known in the literature.
-They all used the standard approach in political science, which is to employ some form of general linearized regression that ignores dyadic interdependencies.
-Post estimation, standard errors are often adjusted in an attempt to account for clustering of observations.
-
-\begin{table}
-\caption{Features of the Studies Replicated. }
-	\begin{tabular}{lcccccc}
-		 & Model & \# of Actors & Years  & \# of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \hline\hline
-		Weeks (2012) & Logit & & & $901,540$ & Directed & Robust \\
-		Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\
-		McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\		
-		Gibler (2017) & Logit & & & $650,557$ & UnDirected & no \\		
-		Rose (2004) & OLS & & & $234,597$ & Directed & Robust \\
-	\hline\hline
-	\end{tabular}
-\end{table}
-
-We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
-
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
-We believe that it does, and that the dynamic latent factor model provides a step forward.
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l| p{7cm} l}
-	& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
-	& &  in a Network Model? \\
-	\hline\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\
-		\hline
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
-		\hline
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
-		\hline
-		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 		
-		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
-Cases in which the finding is not replicated are highlighted in bold.}
-	\label{tab:modelFindingSumm}
-\end{table}
-
-We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
-
-% % % % % % % % % % % % % % % % % % 
-By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
-To show that this is the case, we examine whether our approach achieves better predictive performance in an out of sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-Then for each $s \in \{1,\ldots,k\}$, we:
-
-\begin{enumerate}
-	\item estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$,
-	\item and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters. 
-\end{enumerate}
-
-\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-
-We set a number of benchmarks for comparison.
-First we compare the AME model to a GLM model using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
-We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
-The lagged dependent variable is the equivalent of saying that conflict and peace are relatively likely to persist between dyads, while the inclusion of a lagged reciprocity term in a GLM framework is a simple way to account for retaliatory strikes.
-
-We utilize three performance criterions to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
-ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
-An issue with an ROC Curve when looking at conflict, is that it is relatively rare at the dyadic level: in most years only 3\% of possible dyads are in conflict with one another.
-If peace is common, even a poor model will have a very low False Positive Rate.
-
-To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
-These examine the trade-offs between the percentage of conflicts a model predicts, and the percentage of predicted conflicts which occur.
-Lastly, we examine separation plots \citep{greenhill:etal:2011}.
-These provide an intuitive visualization of the accuracy of our predictions by juxtaposing a line showing the predicted probability of conflict with whether conflict actually occurs for all cases (where the cases are sorted by the predicted probability and then colored to indicate the outcome).
-Here a perfect model would have all cases where conflict actually exists on the right with a predicted probability of $1$, and would predict $0$ in all other cases.
-All of the models' performance out of sample by these metrics are displayed in figure \ref{fig:roc_ame}.
-The AME model with covariates is the best performing model out of sample in all cases.
-This model outperforms each of the GLM variants by a notable margin.
-% % % % % % % % % % % % % % % % % %
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l|l cc}
-	~ & ~ & GLM & AME \\
-	\hline\hline
-		\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
-		~ & AUC (PR) & 0.08 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
-		~ & AUC (PR) & 0.13 & \textbf{0.28} \\
-		\hline
-		\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
-		~ & AUC (PR) & 0.00 & \textbf{0.08} \\
-		\hline
-		\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
-		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach.
-Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
-The fifth involved a gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
-Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
-	\label{tab:modelPerfSumm}
-\end{table}
-\FloatBarrier
-
-\subsection{Replication of Weeks (2012)}
-
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
-One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-Her time frame is from 1946-1999.
-The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
-She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
-
-The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
-In Weeks's initial models, as detailed in figure \ref{fig:weeksCoef} she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
-When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
-We similarly find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in figure \ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
-In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
-In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
-So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
-The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in figure \ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Reiter \& Stam (2003)}
-
-\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent.
-They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that ``We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
-This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
-We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
-The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
-Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
-
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
-In our replication using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
-Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitauc} and Figure ~\ref{fig:reitpr}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:reitauc}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
-\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
-\end{figure}
-
-\newpage
-\subsection{Replication of Rose (2004)}
-
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.
-Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
-He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders...
-(2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
-
-When we compare the results of Rose's original OLS model to our model that accounts for network dependencies, the results are generally similar.
-As you can see in Figure~\ref{fig:roseCoef}, the main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- remains when we move from an OLS to a Gaussian AME model.
-The most striking difference between the models is that, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies.
-Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-Excitingly, when we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out of sample -- 3.23 for the OLS model and 1.77 for the AME model.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
- \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{ABplot_rose_top10.pdf}
- \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
-\end{figure}
-
-%\begin{figure}
-%\centering   
-%\subfigure[Nodal Random Effects]%{\label{fig:rosesend}\includegraphics[width=65mm]{ABplot_rose_top10.pdf}}
-%\subfigure[Receiver Random Effects]{\label{fig:rosereceive}\includegraphics[width=65mm]{ABplot_receiver_rose.pdf}}
-%\caption{Sender and receiver random effects for AME estimation of Rose (2004).}
-%\label{fig:rosesr}
-%\end{figure}
-
-\newpage
-\subsection{Replication of McDonald (2004)}
-
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-He observes that knowledge about the link between conflict and trade is indeterminate in the field of international relations, noting that competing explanations persist.
-He calls for more precise empirical tests.
-Importantly, \citet{mcdonald:2004} includes the interdependence argument---that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)'' in his overview of underdeveloped hypotheses.
-
-Accordingly, the primary contribution of the study is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-To do so, \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measures free trade in two ways.
-The first  captures the idea that larger protected sectors generate greater societal pressures resulting in  pockets of support for war.
-This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state i's GDP or total dyadic trade divided by state j's GDP.'' (560).
-Finally, the binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include a spline correction for time-series data as well as robust standard errors clustered on each dyad.
-
-% TO DO: include discussion and analysis of different K's in appendix (for M cD)
-\begin{figure}
-\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-\end{figure}
-
-% TO DO: fix bib
-
-Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries {i} and {j} are highly dependent and countries {j} and {k} are also highly dependent, then we are likely to observe high dependency between countries {i} and {k}.
-Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
-This indicates that conflict is less likely between members of a trade community.
-Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and replicated model and Figure~\ref{fig:mcdall} demonstrates the predictive performance of each model.
-Figure ~\ref{fig:mcdcirc} represents the multiplicative effects of the model.
-This plot enables us to consider actors with similar sending and receiving patterns resulting from stochastic equivalence and homophily.
-In the upper arch for senders, there is a clear clustering of United States' allies (the United Kingdom, Netherlands, and Japan).
-We also observe an African cluster in the top right segment for receivers (Egypt, Niger, and Zambia for example).
-These countries both trade amongst themselves and have similar conflict behavior -- stochastic equivalence is present among the first group, homophily among the second -- and so once we account for these clusters using the multiplicative effects, the residual effect of trade actually increases the likelihood of conflict.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{mcdonaldcirc.pdf}
-\caption{\label{fig:mcdcirc} Visualization of multiplicative effects. Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Gibler (2017)}
-
-% We chose this article because it is published in 2017, demonstrating that the concern with interdependence continues to the present and because it was published in the so-called flagship journal of political science, \textit{The American Political Science Review.} It examines the link between dyadic capability differences and international conflict in terms of how and when states became independent members of the international system.
-It assumes that membership in the international system---defined in typical ways---has a substantial influence on capabilities.
-In the same way that capabilities don't seem to vary enormously over time, at least comparatively, states become independent in a particular location, clustered with surrounding states.
-Neither does this change much over time.
-Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-The replications we have undertaken are all from articles over the past fifteen years.
-A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
-In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-\begin{table}
-\begin{center}
-\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000$ iterations.  \label{tab:gibme}}
-\begin{tabular}{lrr} \toprule
-& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-Allied & 0.142&  0.023 \\
-Joint Democracy &\bf -0.507&  0.045 \\
-Peace Years \bf &-0.260& \bf -0.060 \\
-Spline 1 &\bf -0.001& \bf 0.00 \\
-Spline 2&\bf -0.000& \bf 0.00\\
-Spline 3 &-0.000& \bf 0.00\\ 
-Contiguity &\bf 2.412& \bf 0.640 \\
-Parity &0.075&  -0.013 \\
-Parity at entry year&\bf 0.868&  0.002 \\ 
-Rivalry &\bf 2.031&\bf 0.721\\ 
-Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-\end{tabular}\\
-\end{center}
-{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
-For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
-\end{table}
-
-We re-estimated model 6 from Table 6 (2017,34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
-The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
-Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the amen results.
-Not only is the value of the this parameter small, but it has a very large relative standard error, over a magnitude larger than the parameter itself ($z= 0.038$).
-In addition, joint democracy follows the same pattern of importance in the logistic results, but this disappears once interdependencies are modeled.
-As might be expected the strong geographic clustering in the original study is about one-quarter as strong in the \texttt{amen} estimations.
-Similarly, rivalry coefficients are about one-third the size in the \texttt{amen} formulation, but a great deal more precisely measured ($z=18.116$).
-
-Beyond more informative fixed effect coefficients, the \texttt{amen} approach also provides information about the interdependencies that were modeled.
-The most pertinent of these may be the dyadic and triadic dependencies which are presented in Figure~\ref{fig:gibler:dyadtriad}, which further reveal that the assumption of independence among the dyadic data in this study can be strongly rejected.
-Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies.
-Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.
-
-% sd.rowmean sd.colmean   dyad.dep  triad.dep 
-% 0.03210221 0.03210221 0.99362646 0.00130249 
-\begin{figure}
-\caption{\label{fig:gibler:dyadtriad} The left panel is  a plot of the dyadic dependencies in the Gibler model 6; the right panel shows the triad dependencies.
-Note that the red line represents the null hypothesis of no dependencies, indicating that the standard logistic approach is far from what is uncovered with this analysis.
-}
-\includegraphics[width=\textwidth]{netplots}
-\end{figure}
-
-\section{Conclusion}
-
-It is no longer necessary to assume that the interesting, innate interdependencies in relational data can be ignored.
-Nor do they have to be approximated with ad hoc, incomplete solutions that purport to control for dependencies by modifying the post-estimation standard errors of the estimated coefficients.
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.
-
-%%What about a summary table with AUCs for the different network models here? - MG
-\newpage
-% Bib stuff
-\clearpage
-% \bibliography{master.bib}
-\bibliography{/Users/s7m/whistle/master}
-\bibliographystyle{authordate1}
-\clearpage
-\section*{All About Us}
-\end{document}

### 2017-09-29 paper/4_app.tex
sync up

--- 2017-09-14 paper/netsMatter.tex
@@
-\documentclass[12pt]{amsart} 
-      % \newif\ifpreview
-      % \previewfalse
-% approx iid
-\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}\usepackage{natbib}
-
-% Fonts
-\usepackage[default,osfigures,scale=0.95]{opensans}
-\usepackage[T1]{fontenc}
-\usepackage{ae}
-
-% easy commands for number propers
-\newcommand{\first}{1$^{\text{st}}$}
-\newcommand{\second}{2$^{\text{nd}}$}
-\newcommand{\third}{3$^{\text{rd}}$}
-\newcommand{\nth}[1]{${#1}^{\text{th}}$}
-
-% Generate some fake text
-\usepackage{blindtext}
-
-% % add watermark
-% \usepackage{draftwatermark}
-% \SetWatermarkText{Draft: Not Ready}
-% \SetWatermarkScale{.5}
-% \SetWatermarkColor[gray]{0.9}
-
-% layout control
-\usepackage{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
-\usepackage{parallel}
-\usepackage{parcolumns}
-\usepackage{fancyhdr}
-\usepackage[export]{adjustbox}
-\usepackage{etoolbox}
-
-% math typesetting
-\usepackage{array}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{relsize}
-\usepackage{mathtools}
-\usepackage{bm}
-%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}
-
-% restricts float objects to be inserted before end of section
-% creates float barriers
-\usepackage[section]{placeins}
-
-% tables
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{longtable}
-
-% to adapt caption style
-\usepackage[font={small},labelfont=bf]{caption}
-
-% footnotes at bottom
-%\usepackage[bottom]{footmisc}
-
-% to change enumeration symbols begin{enumerate}[(a)]
-\usepackage{enumerate}
-\usepackage{paralist}
-
-% to colorize links in document. See color specification below
-\usepackage[pdftex,hyperref,x11names]{xcolor}
-
-% for multiple references and insertion of the word "figure" or "table"
-% \usepackage{cleveref}
-
-% load the hyper-references package and set document info
-\usepackage[pdftex]{hyperref}
-
-% graphics stuff
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage[space]{grffile} % allows us to specify directories that have spaces
-\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
-\usepackage{tikz}
-
-% sideway figures
-\usepackage{rotating}
-\usepackage{lscape}
-
-%\usepackage[figuresright]{rotating}
-% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
-%   {\end{minipage}\end{sidewaysfigure}}
-
-% Spacing
-\usepackage[doublespacing]{setspace}
-
-% define clickable links and their colors
-\hypersetup{%
-	unicode=false,          % non-Latin characters in Acrobat's bookmarks
-	pdftoolbar=true,        % show Acrobat's toolbar?
-	pdfmenubar=true,        % show Acrobat's menu?
-	pdffitwindow=false,     % window fit to page when opened
-	pdfstartview={FitH},    % fits the width of the page to the window
-	pdfnewwindow=true,%
-	%pagebackref=false,%
-	pdfauthor={Shahryar Minhas, et alia.},%
-	pdftitle={How to Fix IR},%
-	colorlinks,%
-	citecolor=black,%
-	filecolor=black,%
-	linkcolor=black,%
-	urlcolor=RoyalBlue4
-}
-
-% -------------------- title -------------------- %
-
-\title{International Relations are a Social Network:\\ A Latent Network Approach for Global Politics}
-\vspace{\baselineskip}
-
- \author[Minhas]{Shahryar Minhas}
- \address{Shahryar Minhas: Department of Political Science}
- \curraddr{Michigan State University}
- \email[Corresponding author]{minhassh@msu.edu}
-
- \author[Dorff]{Cassy L. Dorff}
- \address{Cassy L. Dorff}
- \curraddr{Department of Political Science, University of New Mexico}
- \email{cdorff@unm.edu}
-
- \author[Foster]{Margaret Foster}
- \address{Margaret Foster}
- \curraddr{Department of Political Science, Duke University}
- \email{margaret.foster@duke.edu}
-
- \author[Gallop]{Max Gallop}
- \address{Max Gallop}
- \curraddr{Department of Government and Public Policy, University of Strathclyde}
- \email{max.gallop@strath.ac.uk}
-
-  \author[Liu]{Howard Liu}
- \address{Howard Liu}
- \curraddr{Department of Political Science, Duke University}
- \email{hao.liu@duke.edu}
-
- \author[Tellez]{Juan Tellez}
- \address{Juan Tellez}
- \curraddr{Department of Political Science, Duke University}
- \email{juan.tellez@duke.edu}
-
-\author[Ward]{Michael D. Ward}
-\address{Michael D. Ward}
-\curraddr{Department of Political Science, Duke University}
-\email{michael.d.ward@duke.edu}
-
-% \date{\today}
-
-\thanks{Shahryar Minhas and Michael D.
-Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings.
-We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
-
-\setlength{\headheight}{15pt}
-\setlength{\headsep}{20pt}
-\pagestyle{fancyplain}
-\fancyhf{}
-\lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{How to fix IR}}
-\rhead{\fancyplain{}{\today}}
-\rfoot{\fancyplain{}{\thepage}}
-% references to graphics
-\makeatletter
-% let's keep them in the same directory \def\input@path{{}
-
-% easy command for boldface math symbols
-\newcommand{\mbs}[1]{\boldsymbol{#1}}
-
-% command for R package font
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
-
-% Add some colors
-\definecolor{red1}{RGB}{253,219,199}
-\definecolor{red2}{RGB}{244,165,130}
-\definecolor{red3}{RGB}{178,24,43}
-
-\definecolor{green1}{RGB}{229,245,224}
-\definecolor{green2}{RGB}{161,217,155}
-\definecolor{green3}{RGB}{49,163,84}
-
-\definecolor{blue0}{RGB}{255,247,251}
-\definecolor{blue1}{RGB}{222,235,247}
-\definecolor{blue2}{RGB}{158,202,225}
-\definecolor{blue3}{RGB}{49,130,189}
-\definecolor{blue4}{RGB}{4,90,141}
-
-\definecolor{purple1}{RGB}{191,211,230}
-\definecolor{purple2}{RGB}{140,150,198}
-\definecolor{purple3}{RGB}{140,107,177}
-
-\definecolor{brown1}{RGB}{246,232,195}
-\definecolor{brown2}{RGB}{223,194,125}
-\definecolor{brown3}{RGB}{191,129,45}
-
-% square bracket matrices
-\let\bbordermatrix\bordermatrix
-\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
-\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
-\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}
-
-\begin{document}
-      % \ifpreview {
-      %    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
-      %    \pagecolor{personal}
-      %    }
-      % \fi
-
-\small{\singlespacing{
-\begin{abstract}
-	International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially data.
-These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.
-However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce results that are not compelling.
-One reason is that the independence required by the typical methods employed does not exist in the data being studied.
-We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies.
-We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
-\end{abstract}
-}}
-
-\maketitle\thispagestyle{empty}
-\newpage\setcounter{page}{1}
-
-\section{\textbf{Introduction}}
-
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.
-However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
-It is not just the diagonals of the variance-covariance matrix this affects.
-
-In this article, we introduce a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in the data.
-This model has two parts, one of which uses a set of additive random effects capturing \first- and \second-order dependencies in the relational data.
-The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.
-We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a generalized linear model with robust standard errors) to those obtained with the AME framework.
-
-This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
-
-\section{\textbf{Limitations of Standard Framework}}
-
-Generalized linear models are typically expressed via a stochastic and systematic component.
-The stochastic component reflects assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function such as the normal, binomial, or Poisson.
-Each dyad in the sample is independently drawn from a particular distribution, given $\theta_{ij}$.
-The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariates, $\mathbf{X}_{ij}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
-A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of the distribution, each of the dyadic observations is conditionally independent.
-
-The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
-After having chosen a set of covariates and specifying a distribution, the joint density function over all dyads using the observations from Table 1 is given:
-
-\vspace{-8mm}
-\begin{align}
-\begin{aligned}
-	P(y_{ij}, y_{ik}, \ldots, y_{lk} | \theta_{ij}, \theta_{ik}, \ldots, \theta_{lk}) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots \times P(y_{lk} | \theta_{lk}) \\
-	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
-\end{aligned}
-\end{align}
-
-\noindent The joint probability is converted into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
-
-The parameters are estimated by maximizing the likelihood.
-The likelihood as defined above is only valid if  $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
-
-\section{\textbf{Dependencies in Dyadic Data}}
-
-Relational, or dyadic, data provide measurements of how sets of actors relate to one another.
-Typically, pairs of actors are examined.
-These pairs are called \textit{dyads}.
-Such data can be arranged as a directed dyadic design in which the unit of analysis is some set of $n$ actors that have been paired together to form a matrix of dyads, i.
-e., represented as a $n \times n$ matrix, $\mathbf{Y}$, with diagonals typically undefined.
-The $ij^{th}$ entry defines the relationship of $i$ and $j$ and can be continuous or discrete.
-It can also be symmetric (known as undirected) or asymmetric (called directed).
-This is not the stacked dyad structure normally found in social science, but it is equivalent and presents a format in which the interdependencies are exposed and transparent.
-The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.
-
-One important dependency is nodal dependency, whereby each actor has similar interactions with many of its partners, leading to significant heterogeneity in activity levels across nodes.
-An important implication of this across-node heterogeneity is within-node homogeneity of ties: the values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, are similar to each other because each has a common sender.
-If a sender $i$ is more active active in the network than other senders, it will have greater homogeneity of its linkages.
-More popular targets show greater similarity across linkages in a given column, $\{y_{ji},y_{ki},y_{li}\}$.
-Actors which are more likely to send ties in a network are also more likely to receive them.
-As a result the rows and columns in an adjacency matrix may be correlated.
-Another structural interdependence is reciprocity.
-This is a second-order, or dyadic dependency whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
-If the data are simply stacked dyads in some arbitrary order, this kind of dependency is harder to visualize, but exists nonetheless.
-
-The presence of these interdependencies in relational data complicates the practical assumption of observational independence.
-Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-By ignoring these potential interdependencies, we often ignore important features of the problem under study.
-The study of international relations is founded on the relations among actors.
-Why ignore the interdependencies that led to the study of IR in the first place?
-
-The additive part described above does not handle third-order dependency.
-A third-order dependency is the dependency between triads, not dyads.
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
-
-For example, a common finding is that democracies are more likely to form trade agreements with one another.
-Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
-A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
-Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
-Another third-order dependence pattern is stochastic equivalence.
-Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
-Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
-Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
-Latent variables can account for third-order dependence patterns within the context of the SRRM.
-These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
-
-\section{\textbf{Additive and Multiplicative Effect Models for Networks}}
-
-The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
-It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
-Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
-Fortunately, it is easy to empirically examine dependencies in relational data.
-
-The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
-It was introduced in \cite{hoff:2008}; the open-source computer package associated with this approach is \cite{amenpkg}.
-We describe it below.
-
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
-\label{eqn:ame}
-\end{aligned}
-\end{align}
-
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
-
-\subsection{\textbf{Social Relations Model: Additive Part of AME}}
-
-\citet{warner:etal:1979} developed a social relations model (SRM), using ANOVA decomposition, that focuses on such dependencies.
-SRM provides the error structure for the additive effects component of the AME framework.
-SRM decomposes the variance of dyadic observations in terms of a) heterogeneity across row means (often called out-degree), b) column means (or, in-degree), c) correlation between row and column means, and d) correlations within dyads.
-\citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= \mu + e_{ij} \\
-	e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
-	\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
-	\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
-	\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix},
-\label{eqn:srmCov}
-\end{aligned}
-\end{align}
-
-wherein $\mu$ is the density of a network, and $e_{ij}$ represents residual variation in each cell (or dyad).
-The residual variation has three separate parts: a)   row/sender effects ($a_{i}$), b)  column/receiver effects ($b_{j}$), and c) within-dyad effects ($\epsilon_{ij}$).
-The row and column effects are modeled jointly to account for correlation in how active an actor is in sending and receiving ties.
-Heterogeneity in the row and column means is modeled by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively.
-$\sigma_{ab}$ describes whether actors which send a lot of ties also receive a lot.
-Second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a reciprocity, parameter $\rho$.
-
-The SRM covariance structure described in Equation~\ref{eqn:srmCov} can be incorporated into the systematic component of a GLM framework to produce the social relations regression model (SRRM): $\bm\beta^{\top} \mathbf{X}_{ij} + a_{i} + b_{j} + \epsilon_{ij}$, where $ \bm\beta^{\top} \mathbf{X}_{ij}$ accommodates the inclusion of dyadic, sender, and receiver covariates \citep{hoff:2005}.
-This is a regression framework with additive, random effects, but it models variances and covariances often seen in relational data.
-Furthermore, this accommodates a diversity of outcome distributions.
-In the case of binary data, this can be done by utilizing a random effects logistic (or probit) latent variable representation.
-
-\subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}
-
-The latent factor model (LFM) expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
-General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.
-
-\begin{align}
-\begin{aligned}
-	\alpha(&\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{\top} \Lambda \textbf{u}_{j} \\
-	&\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \text{, where}\\
-	 &\Lambda \equiv K \times K \text{ diagonal matrix}
-\label{eqn:latAlpha}
-\end{aligned}
-\end{align}
-
-Many observed  networks exhibit varying degrees of stochastic equivalence and homophily.
-In the latent factor model, each actor has an unobserved vector of characteristics, $\textbf{u}_{i} = \{u_{i,1}, \ldots, u_{i,K} \}$, describing each node's behavior in the network.
-The probability of a tie between $i$ and $j$ depends on the extent to which $\textbf{u}_{i}$ and $\textbf{u}_{j}$ are ``similar'' and on whether the entries of $\Lambda$ are greater or less than zero.
-Similarity in the latent factors, $\textbf{u}_{i} \approx \textbf{u}_{j}$, corresponds to stochastic equivalence; the eigenvalue determines whether the network exhibits positive or negative homophily.
-Accordingly, this approach can represent both homophily and stochastic equivalence.
-
-In directed (asymmetric) networks there are two sets of latent positions, denoted $\textbf{u}$ and $\textbf{v}$, representing the sending and receiving positions separately.
-Thus, actors in the network have a vector of latent characteristics describing their behavior as a sender, denoted by $\textbf{u}$, and as a receiver, $\textbf{v}$: $\textbf{u}_{i}, \textbf{v}_{j} \in \mathbb{R}^{K}$.
-This can alter the probability of an interaction between $ij$ additively: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, where $\textbf{D}$ is a $K \times K$ diagonal matrix.
-
-The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
-
-\section{Simulation study}
-
-Here we consider inference for a regression parameter $\beta$ of a linear or generalized linear model for a network in the case where there is an omitted variable.
-The true data-generating models we consider are of the form
-
-$y_{i,j} \sim  \mu + \beta x_i x_j + \gamma w_i w_j + \epsilon_{i,j} $
-
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_i \} \in \mathbb R^n$ is a vector of observed node-specific characteristics, and $w = \{ w_i\} \in \mathbb R^n$ is a vector of unobserved node-specific characteristics.
-We compare inference for $\beta$ using three models:
-
-\begin{itemize}
-	\item A naive regression model assuming independent errors; 
-	\item A latent factor model; 
-	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
-\end{itemize}
-
-We make these comparisons in the context of a binary network outcome with a probit regression model.
-
-One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
-
-Alternatively, bias and precision can be summarized by confidence interval coverage and width.
-Coverage should ideally be at the nominal level.
-If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
-
-\begin{figure}
-	\centering
-	\caption{Bias in parameter when homophily is ignored.}
-	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{ameSimBias_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimBias_beta.pdf}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{Coverage in parameter when homophily is ignored.}
-	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{ameSimCover_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimCover_beta.pdf}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{Correlation between missing variable and multiplicative random effect in AME.}
-	\label{fig:ameCorr}
-	\includegraphics[width=1\textwidth]{ameSimCorr.pdf} \\
-\end{figure}
-
 \section{Replications}
@@
 \end{figure}
-
-\section{Conclusion}
-
-It is no longer necessary to assume that the interesting, innate interdependencies in relational data can be ignored.
-Nor do they have to be approximated with ad hoc, incomplete solutions that purport to control for dependencies by modifying the post-estimation standard errors of the estimated coefficients.
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.
-
-%%What about a summary table with AUCs for the different network models here? - MG
-\newpage
-% Bib stuff
-\clearpage
-% \bibliography{master.bib}
-\bibliography{/Users/s7m/whistle/master}
-\bibliographystyle{authordate1}
-\clearpage
-\section*{All About Us}
-\end{document}

### 2017-09-29 paper/5_concl.tex
sync up

--- 2017-08-31 paper/netsMatter.tex
@@
-\documentclass[12pt]{amsart} 
-      % \newif\ifpreview
-      % \previewfalse
-% approx iid
-\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}\usepackage{natbib}
-
-% Fonts
-\usepackage[default,osfigures,scale=0.95]{opensans}
-\usepackage[T1]{fontenc}
-\usepackage{ae}
-
-% easy commands for number propers
-\newcommand{\first}{1$^{\text{st}}$}
-\newcommand{\second}{2$^{\text{nd}}$}
-\newcommand{\third}{3$^{\text{rd}}$}
-\newcommand{\nth}[1]{${#1}^{\text{th}}$}
-
-% Generate some fake text
-\usepackage{blindtext}
-
-% % add watermark
-% \usepackage{draftwatermark}
-% \SetWatermarkText{Draft: Not Ready}
-% \SetWatermarkScale{.5}
-% \SetWatermarkColor[gray]{0.9}
-
-% layout control
-\usepackage{geometry}
-\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
-\usepackage{parallel}
-\usepackage{parcolumns}
-\usepackage{fancyhdr}
-\usepackage[export]{adjustbox}
-\usepackage{etoolbox}
-
-% math typesetting
-\usepackage{array}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{relsize}
-\usepackage{mathtools}
-\usepackage{bm}
-%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}
-
-% restricts float objects to be inserted before end of section
-% creates float barriers
-\usepackage[section]{placeins}
-
-% tables
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{longtable}
-
-% to adapt caption style
-\usepackage[font={small},labelfont=bf]{caption}
-
-% footnotes at bottom
-%\usepackage[bottom]{footmisc}
-
-% to change enumeration symbols begin{enumerate}[(a)]
-\usepackage{enumerate}
-\usepackage{paralist}
-
-% to colorize links in document. See color specification below
-\usepackage[pdftex,hyperref,x11names]{xcolor}
-
-% for multiple references and insertion of the word "figure" or "table"
-% \usepackage{cleveref}
-
-% load the hyper-references package and set document info
-\usepackage[pdftex]{hyperref}
-
-% graphics stuff
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage[space]{grffile} % allows us to specify directories that have spaces
-\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
-\usepackage{tikz}
-
-% sideway figures
-\usepackage{rotating}
-\usepackage{lscape}
-
-%\usepackage[figuresright]{rotating}
-% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
-%   {\end{minipage}\end{sidewaysfigure}}
-
-% Spacing
-\usepackage[doublespacing]{setspace}
-
-% define clickable links and their colors
-\hypersetup{%
-	unicode=false,          % non-Latin characters in Acrobat's bookmarks
-	pdftoolbar=true,        % show Acrobat's toolbar?
-	pdfmenubar=true,        % show Acrobat's menu?
-	pdffitwindow=false,     % window fit to page when opened
-	pdfstartview={FitH},    % fits the width of the page to the window
-	pdfnewwindow=true,%
-	%pagebackref=false,%
-	pdfauthor={Shahryar Minhas, et alia.},%
-	pdftitle={How to Fix IR},%
-	colorlinks,%
-	citecolor=black,%
-	filecolor=black,%
-	linkcolor=black,%
-	urlcolor=RoyalBlue4
-}
-
-% -------------------- title -------------------- %
-
-\title{How to fix IR}
-\vspace{\baselineskip}
-
- \author[Minhas]{Shahryar Minhas}
- \address{Shahryar Minhas: Department of Political Science}
- \curraddr{Michigan State University}
- \email[Corresponding author]{minhassh@msu.edu}
-
- \author[Dorff]{Cassy L. Dorff}
- \address{Cassy L. Dorff}
- \curraddr{Department of Political Science, University of New Mexico}
- \email{cdorff@unm.edu}
-
- \author[Foster]{Margaret Foster}
- \address{Margaret Foster}
- \curraddr{Department of Political Science, Duke University}
- \email{margaret.foster@duke.edu}
-
- \author[Gallop]{Max Gallop}
- \address{Max Gallop}
- \curraddr{Department of Government and Public Policy, University of Strathclyde}
- \email{max.gallop@strath.ac.uk}
-
-  \author[Liu]{Howard Liu}
- \address{Howard Liu}
- \curraddr{Department of Political Science, Duke University}
- \email{hao.liu@duke.edu}
-
- \author[Tellez]{Juan Tellez}
- \address{Juan Tellez}
- \curraddr{Department of Political Science, Duke University}
- \email{juan.f.tellez@duke.edu}
-
-\author[Ward]{Michael D. Ward}
-\address{Michael D. Ward}
-\curraddr{Department of Political Science, Duke University}
-\email{michael.d.ward@duke.edu}
-
-% \date{\today}
-
-\thanks{Shahryar Minhas and Michael D.
-Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings.
-We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
-
-\setlength{\headheight}{15pt}
-\setlength{\headsep}{20pt}
-\pagestyle{fancyplain}
-\fancyhf{}
-\lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{How to Fix IR}}
-\rhead{\fancyplain{}{\today}}
-\rfoot{\fancyplain{}{\thepage}}
-% references to graphics
-\makeatletter
-% let's keep them in the same directory \def\input@path{{}
-
-% easy command for boldface math symbols
-\newcommand{\mbs}[1]{\boldsymbol{#1}}
-
-% command for R package font
-\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
-
-% Add some colors
-\definecolor{red1}{RGB}{253,219,199}
-\definecolor{red2}{RGB}{244,165,130}
-\definecolor{red3}{RGB}{178,24,43}
-
-\definecolor{green1}{RGB}{229,245,224}
-\definecolor{green2}{RGB}{161,217,155}
-\definecolor{green3}{RGB}{49,163,84}
-
-\definecolor{blue0}{RGB}{255,247,251}
-\definecolor{blue1}{RGB}{222,235,247}
-\definecolor{blue2}{RGB}{158,202,225}
-\definecolor{blue3}{RGB}{49,130,189}
-\definecolor{blue4}{RGB}{4,90,141}
-
-\definecolor{purple1}{RGB}{191,211,230}
-\definecolor{purple2}{RGB}{140,150,198}
-\definecolor{purple3}{RGB}{140,107,177}
-
-\definecolor{brown1}{RGB}{246,232,195}
-\definecolor{brown2}{RGB}{223,194,125}
-\definecolor{brown3}{RGB}{191,129,45}
-
-% square bracket matrices
-\let\bbordermatrix\bordermatrix
-\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
-\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
-\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}
-
-\begin{document}
-      % \ifpreview {
-      %    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
-      %    \pagecolor{personal}
-      %    }
-      % \fi
-
-\small{\singlespacing{
-\begin{abstract}
-	International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially data.
-These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.
-However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce results that are not compelling.
-One reason is that the independence required by the typical methods employed does not exist in the data being studied.
-We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies.
-We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
-\end{abstract}
-}}
-
-\maketitle\thispagestyle{empty}
-\newpage\setcounter{page}{1}
-
-\section{\textbf{Introduction}}
-
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM), either OLS or logistic regression, to estimate regression coefficients.
-However, we know that extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
-It is not just the diagonals of the variance-covariance matrix this affects.
-
-In this article, we introduce a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in the data.
-This model has two parts, one of which uses a set of additive random effects capturing \first- and \second-order dependencies in the relational data.
-The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.
-We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a generalized linear model with robust standard errors) to those obtained with the AME framework.
-
-This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
-%We note in closing that the computational requirements of this approach are substantial in terms of computing time. Cloud and parallel methods reduce this burden.
-
-\section{\textbf{Limitations of Standard Framework}}
-
-Generalized linear models are typically expressed via a stochastic and systematic component.
-The stochastic component reflects assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function such as the normal, binomial, or Poisson.
-Each dyad in the sample is independently drawn from a particular distribution, given $\theta_{ij}$.
-The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariates, $\mathbf{X}_{ij}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
-A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of the distribution, each of the dyadic observations is conditionally independent.
-
-The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
-After having chosen a set of covariates and specifying a distribution, the joint density function over all dyads using the observations from Table 1 is given:
-
-\vspace{-8mm}
-\begin{align}
-\begin{aligned}
-	P(y_{ij}, y_{ik}, \ldots, y_{lk} | \theta_{ij}, \theta_{ik}, \ldots, \theta_{lk}) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots \times P(y_{lk} | \theta_{lk}) \\
-	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
-\end{aligned}
-\end{align}
-
-\noindent The joint probability is converted into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
-
-The parameters are estimated by maximizing the likelihood.
-The likelihood as defined above is only valid if  $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
-
-\section{\textbf{Additive and Multiplicative Effect Models for Networks}}
-
-The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
-It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
-Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
-Fortunately, it is easy to empirically examine dependencies in relational data.
-
-The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
-It was introduced in \cite{hoff:2008} and elaborated in \cite{hoff:2015}; the open-source computer package associated with this approach is \cite{amenpkg}.
-We describe it below.
-
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
-\label{eqn:ame}
-\end{aligned}
-\end{align}
-
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
-
-\subsection*{\textbf{Addressing Dependencies in Dyadic Data}}
-
-Relational, or dyadic, data provide measurements of how sets of actors relate to one another.
-Typically, pairs of actors are examined.
-These pairs are called \textit{dyads}.
-Such data can be arranged as a directed dyadic design in which the unit of analysis is some set of $n$ actors that have been paired together to form a matrix of dyads, i.
-e., represented as a $n \times n$ matrix, $\mathbf{Y}$, with diagonals typically undefined.
-The $ij^{th}$ entry defines the relationship of $i$ and $j$ and can be continuous or discrete.
-It can also be symmetric (known as undirected) or asymmetric (called directed).
-This is not the stacked dyad structure normally found in social science, but it is equivalent and presents a format in which the interdependencies are exposed and transparent.
-The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.
-
-One important dependency is nodal dependency, whereby each actor has similar interactions with many of its partners, leading to significant heterogeneity in activity levels across nodes.
-An important implication of this across-node heterogeneity is within-node homogeneity of ties: the values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, are similar to each other because each has a common sender.
-If a sender $i$ is more active active in the network than other senders, it will have greater homogeneity of its linkages.
-More popular targets show greater similarity across linkages in a given column, $\{y_{ji},y_{ki},y_{li}\}$.
-Actors which are more likely to send ties in a network are also more likely to receive them.
-As a result the rows and columns in an adjacency matrix may be correlated.
-Another structural interdependence is reciprocity.
-This is a second-order, or dyadic dependency whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
-If the data are simply stacked dyads in some arbitrary order, this kind of dependency is harder to visualize, but exists nonetheless.
-
-The presence of these interdependencies in relational data complicates the practical assumption of observational independence.
-Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-By ignoring these potential interdependencies, we often ignore important features of the problem under study.
-The study of international relations is founded on the relations among actors.
-Why ignore the interdependencies that led to the study of IR in the first place?
-
-\subsection{\textbf{Social Relations Model: Additive Part of AME}}
-
-\citet{warner:etal:1979} developed a social relations model (SRM), using ANOVA decomposition, that focuses on such dependencies.
-SRM provides the error structure for the additive effects component of the AME framework.
-SRM decomposes the variance of dyadic observations in terms of a) heterogeneity across row means (often called out-degree), b) column means (or, in-degree), c) correlation between row and column means, and d) correlations within dyads.
-\citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= \mu + e_{ij} \\
-	e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
-	\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
-	\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
-	\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix},
-\label{eqn:srmCov}
-\end{aligned}
-\end{align}
-
-wherein $\mu$ is the density of a network, and $e_{ij}$ represents residual variation in each cell (or dyad).
-The residual variation has three separate parts: a)   row/sender effects ($a_{i}$), b)  column/receiver effects ($b_{j}$), and c) within-dyad effects ($\epsilon_{ij}$).
-The row and column effects are modeled jointly to account for correlation in how active an actor is in sending and receiving ties.
-Heterogeneity in the row and column means is modeled by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively.
-$\sigma_{ab}$ describes whether actors which send a lot of ties also receive a lot.
-Second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a reciprocity, parameter $\rho$.
-
-The SRM covariance structure described in Equation~\ref{eqn:srmCov} can be incorporated into the systematic component of a GLM framework to produce the social relations regression model (SRRM): $\bm\beta^{\top} \mathbf{X}_{ij} + a_{i} + b_{j} + \epsilon_{ij}$, where $ \bm\beta^{\top} \mathbf{X}_{ij}$ accommodates the inclusion of dyadic, sender, and receiver covariates \citep{hoff:2005}.
-This is a regression framework with additive, random effects, but it models variances and covariances often seen in relational data.
-Furthermore, this accommodates a diversity of outcome distributions.
-In the case of binary data, this can be done by utilizing a random effects logistic (or probit) latent variable representation.
-
-\subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}
-
-The additive part described above does not handle third-order dependency.
-A third-order dependency is the dependency between triads, not dyads.
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
-
-For example, a common finding is that democracies are more likely to form trade agreements with one another.
-Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
-A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
-Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
-Another third-order dependence pattern is stochastic equivalence.
-Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
-Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
-Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
-Latent variables can account for third-order dependence patterns within the context of the SRRM.
-These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
-
-A number of latent variable approaches have been developed to represent third-order dependencies in relational data, we focus on the latent factor model (LFM).
-This expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
-General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.
-
-\begin{align}
-\begin{aligned}
-	\alpha(&\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{\top} \Lambda \textbf{u}_{j} \\
-	&\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \text{, where}\\
-	 &\Lambda \equiv K \times K \text{ diagonal matrix}
-\label{eqn:latAlpha}
-\end{aligned}
-\end{align}
-
-Many observed  networks exhibit varying degrees of stochastic equivalence and homophily.
-In the latent factor model, each actor has an unobserved vector of characteristics, $\textbf{u}_{i} = \{u_{i,1}, \ldots, u_{i,K} \}$, describing each node's behavior in the network.
-The probability of a tie between $i$ and $j$ depends on the extent to which $\textbf{u}_{i}$ and $\textbf{u}_{j}$ are ``similar'' and on whether the entries of $\Lambda$ are greater or less than zero.
-Similarity in the latent factors, $\textbf{u}_{i} \approx \textbf{u}_{j}$, corresponds to stochastic equivalence; the eigenvalue determines whether the network exhibits positive or negative homophily.
-Accordingly, this approach can represent both homophily and stochastic equivalence.
-
-In directed (asymmetric) networks there are two sets of latent positions, denoted $\textbf{u}$ and $\textbf{v}$, representing the sending and receiving positions separately.
-Thus, actors in the network have a vector of latent characteristics describing their behavior as a sender, denoted by $\textbf{u}$, and as a receiver, $\textbf{v}$: $\textbf{u}_{i}, \textbf{v}_{j} \in \mathbb{R}^{K}$.
-This can alter the probability of an interaction between $ij$ additively: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, where $\textbf{D}$ is a $K \times K$ diagonal matrix.
-
-The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
-
-\section{Simulation study}
-
-Here we consider inference for a regression parameter $\beta$ of a linear or generalized linear model for a network in the case where there is an omitted variable.
-The true data-generating models we consider are of the form
-
-$y_{i,j} \sim  \mu + \beta x_i x_j + \gamma w_i w_j + \epsilon_{i,j} $
-
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_i \} \in \mathbb R^n$ is a vector of observed node-specific characteristics, and $w = \{ w_i\} \in \mathbb R^n$ is a vector of unobserved node-specific characteristics.
-We compare inference for $\beta$ using three models:
-
-\begin{itemize}
-	\item A naive regression model assuming independent errors; 
-	\item A latent factor model; 
-	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
-\end{itemize}
-
-We make these comparisons in the context of a binary network outcome with a probit regression model.
-
-One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
-
-Alternatively, bias and precision can be summarized by confidence interval coverage and width.
-Coverage should ideally be at the nominal level.
-If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
-
-\begin{figure}
-	\centering
-	\caption{Bias in parameter when homophily is ignored.}
-	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{ameSimBias_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimBias_beta.pdf}
-\end{figure}
-
-\section{Replications}
-
-\subsection{Design}
-
-For the purpose of this study, we choose five prominent studies from the broad field of international relations and international political economy that utilize relation data \citep{mcdonald:2004, reiter:stam:2003, rose:2004, weeks:2012, gibler:2017}.
-We chose to replicate studies that were written fairly recently, since 2000, and that have been cited over 100 times.
-Each of these pieces was published in a prominent journal and is well-known in the literature.
-They all used the standard approach in political science, which is to employ some form of general linearized regression that ignores dyadic interdependencies.
-Post estimation, standard errors are often adjusted in an attempt to account for clustering of observations.
-
-\begin{table}
-\caption{Features of the Studies Replicated. }
-\scalebox{.8}{
-	\begin{tabular}{ccrcc}
-		Study & Model & N. of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \toprule
-		Weeks (2012) & Logit & $901,540$ & Directed & Robust \\
-		Reiter \& Stam (2003) & Logit & $753,456$ & Directed & Robust \\
-		Rose (2004) &OLS&$234,597$&Directed&Robust \\
-		McDonald (2004) & Logit &$92,354$& Undirected &Robust\\
-		Gibler (2017) &Logit& $650,557$&UnDirected& no \\
-	\bottomrule
-	\end{tabular}
-}
-\end{table}
-
-We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
-
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
-We believe that it does, and that the dynamic latent factor model provides a step forward.
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l| p{7cm} l}
-	& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
-	& &  in a Network Model? \\
-	\hline\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\
-		\hline
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
-		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-		\hline
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
-		\hline
-		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
-Cases in which the finding is not replicated are highlighted in bold.}
-	\label{tab:modelFindingSumm}
-\end{table}
-
-We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
-
-% % % % % % % % % % % % % % % % % % 
-By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
-To show that this is the case, we examine whether our approach achieves better predictive performance in an out of sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-Then for each $s \in \{1,\ldots,k\}$, we:
-
-\begin{enumerate}
-	\item estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$,
-	\item and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters. 
-\end{enumerate}
-
-\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-
-We set a number of benchmarks for comparison.
-First we compare the AME model to a GLM model using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
-We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
-The lagged dependent variable is the equivalent of saying that conflict and peace are relatively likely to persist between dyads, while the inclusion of a lagged reciprocity term in a GLM framework is a simple way to account for retaliatory strikes.
-
-We utilize three performance criterions to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
-ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
-An issue with an ROC Curve when looking at conflict, is that it is relatively rare at the dyadic level: in most years only 3\% of possible dyads are in conflict with one another.
-If peace is common, even a poor model will have a very low False Positive Rate.
-
-To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
-These examine the trade-offs between the percentage of conflicts a model predicts, and the percentage of predicted conflicts which occur.
-Lastly, we examine separation plots \citep{greenhill:etal:2011}.
-These provide an intuitive visualization of the accuracy of our predictions by juxtaposing a line showing the predicted probability of conflict with whether conflict actually occurs for all cases (where the cases are sorted by the predicted probability and then colored to indicate the outcome).
-Here a perfect model would have all cases where conflict actually exists on the right with a predicted probability of $1$, and would predict $0$ in all other cases.
-All of the models' performance out of sample by these metrics are displayed in figure \ref{fig:roc_ame}.
-The AME model with covariates is the best performing model out of sample in all cases.
-This model outperforms each of the GLM variants by a notable margin.
-% % % % % % % % % % % % % % % % % %
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l|l cc}
-	~ & ~ & GLM & AME \\
-	\hline\hline
-		\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
-		~ & AUC (PR) & 0.08 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
-		~ & AUC (PR) & 0.13 & \textbf{0.28} \\
-		\hline
-		\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
-		~ & AUC (PR) & 0.00 & \textbf{0.08} \\
-		\hline
-		\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
-		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach.
-Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
-The fifth involved a gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
-Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
-	\label{tab:modelPerfSumm}
-\end{table}
-\FloatBarrier
-
-\subsection{Replication of Weeks (2012)}
-
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
-One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-Her time frame is from 1946-1999.
-The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
-She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
-
-The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
-In Weeks's initial models, as detailed in figure \ref{fig:weeksCoef} she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
-When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
-We similarly find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in figure \ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
-In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
-In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
-So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
-The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in figure \ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Reiter \& Stam (2003)}
-
-\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent.
-They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that ``We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
-This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
-We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
-The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
-Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
-
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
-In our replication using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
-Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitauc} and Figure ~\ref{fig:reitpr}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:reitauc}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
-\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
-\end{figure}
-
-\newpage
-\subsection{Replication of Rose (2004)}
-
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.
-Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
-He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders...
-(2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
-
-When we compare the results of Rose's original OLS model to our model that accounts for network dependencies, the results are generally similar.
-As you can see in Figure~\ref{fig:roseCoef}, the main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- remains when we move from an OLS to a Gaussian AME model.
-The most striking difference between the models is that, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies.
-Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-Excitingly, when we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out of sample -- 3.23 for the OLS model and 1.77 for the AME model.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
- \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{ABplot_rose_top10.pdf}
- \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
-\end{figure}
-
-%\begin{figure}
-%\centering   
-%\subfigure[Nodal Random Effects]%{\label{fig:rosesend}\includegraphics[width=65mm]{ABplot_rose_top10.pdf}}
-%\subfigure[Receiver Random Effects]{\label{fig:rosereceive}\includegraphics[width=65mm]{ABplot_receiver_rose.pdf}}
-%\caption{Sender and receiver random effects for AME estimation of Rose (2004).}
-%\label{fig:rosesr}
-%\end{figure}
-
-\newpage
-\subsection{Replication of McDonald (2004)}
-
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-He observes that knowledge about the link between conflict and trade is indeterminate in the field of international relations, noting that competing explanations persist.
-He calls for more precise empirical tests.
-Importantly, \citet{mcdonald:2004} includes the interdependence argument---that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)'' in his overview of underdeveloped hypotheses.
-
-Accordingly, the primary contribution of the study is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-To do so, \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measures free trade in two ways.
-The first  captures the idea that larger protected sectors generate greater societal pressures resulting in  pockets of support for war.
-This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state i's GDP or total dyadic trade divided by state j's GDP.'' (560).
-Finally, the binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include a spline correction for time-series data as well as robust standard errors clustered on each dyad.
-
-% TO DO: include discussion and analysis of different K's in appendix (for M cD)
-\begin{figure}
-\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-\end{figure}
-
-% TO DO: fix bib
-
-Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries {i} and {j} are highly dependent and countries {j} and {k} are also highly dependent, then we are likely to observe high dependency between countries {i} and {k}.
-Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
-This indicates that conflict is less likely between members of a trade community.
-Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and replicated model and Figure~\ref{fig:mcdall} demonstrates the predictive performance of each model.
-Figure ~\ref{fig:mcdcirc} represents the multiplicative effects of the model.
-This plot enables us to consider actors with similar sending and receiving patterns resulting from stochastic equivalence and homophily.
-In the upper arch for senders, there is a clear clustering of United States' allies (the United Kingdom, Netherlands, and Japan).
-We also observe an African cluster in the top right segment for receivers (Egypt, Niger, and Zambia for example).
-These countries both trade amongst themselves and have similar conflict behavior -- stochastic equivalence is present among the first group, homophily among the second -- and so once we account for these clusters using the multiplicative effects, the residual effect of trade actually increases the likelihood of conflict.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{mcdonaldcirc.pdf}
-\caption{\label{fig:mcdcirc} Visualization of multiplicative effects. Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Gibler (2017)}
-
-% We chose this article because it is published in 2017, demonstrating that the concern with interdependence continues to the present and because it was published in the so-called flagship journal of political science, \textit{The American Political Science Review.} It examines the link between dyadic capability differences and international conflict in terms of how and when states became independent members of the international system.
-It assumes that membership in the international system---defined in typical ways---has a substantial influence on capabilities.
-In the same way that capabilities don't seem to vary enormously over time, at least comparatively, states become independent in a particular location, clustered with surrounding states.
-Neither does this change much over time.
-Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-The replications we have undertaken are all from articles over the past fifteen years.
-A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
-In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-\begin{table}
-\begin{center}
-\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000$ iterations.  \label{tab:gibme}}
-\begin{tabular}{lrr} \toprule
-& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-Allied & 0.142&  0.023 \\
-Joint Democracy &\bf -0.507&  0.045 \\
-Peace Years \bf &-0.260& \bf -0.060 \\
-Spline 1 &\bf -0.001& \bf 0.00 \\
-Spline 2&\bf -0.000& \bf 0.00\\
-Spline 3 &-0.000& \bf 0.00\\ 
-Contiguity &\bf 2.412& \bf 0.640 \\
-Parity &0.075&  -0.013 \\
-Parity at entry year&\bf 0.868&  0.002 \\ 
-Rivalry &\bf 2.031&\bf 0.721\\ 
-Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-\end{tabular}\\
-\end{center}
-{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
-For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
-\end{table}
-
-We re-estimated model 6 from Table 6 (2017,34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
-The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
-Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the amen results.
-Not only is the value of the this parameter small, but it has a very large relative standard error, over a magnitude larger than the parameter itself ($z= 0.038$).
-In addition, joint democracy follows the same pattern of importance in the logistic results, but this disappears once interdependencies are modeled.
-As might be expected the strong geographic clustering in the original study is about one-quarter as strong in the \texttt{amen} estimations.
-Similarly, rivalry coefficients are about one-third the size in the \texttt{amen} formulation, but a great deal more precisely measured ($z=18.116$).
-
-Beyond more informative fixed effect coefficients, the \texttt{amen} approach also provides information about the interdependencies that were modeled.
-The most pertinent of these may be the dyadic and triadic dependencies which are presented in Figure~\ref{fig:gibler:dyadtriad}, which further reveal that the assumption of independence among the dyadic data in this study can be strongly rejected.
-Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies.
-Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.
-
-% sd.rowmean sd.colmean   dyad.dep  triad.dep 
-% 0.03210221 0.03210221 0.99362646 0.00130249 
-\begin{figure}
-\caption{\label{fig:gibler:dyadtriad} The left panel is  a plot of the dyadic dependencies in the Gibler model 6; the right panel shows the triad dependencies.
-Note that the red line represents the null hypothesis of no dependencies, indicating that the standard logistic approach is far from what is uncovered with this analysis.
-}
-\includegraphics[width=\textwidth]{netplots}
-\end{figure}
-
 \section{Conclusion}
@@
 Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.
-
-%%What about a summary table with AUCs for the different network models here? - MG
-\newpage
-% Bib stuff
-\clearpage
-% \bibliography{master.bib}
-\bibliography{/Users/s7m/whistle/master}
-\bibliographystyle{authordate1}
-\clearpage
-\section*{All About Us}
-\end{document}

### 2017-09-29 paper/netsMatter.tex
sync up

@@
 
-\section{\textbf{Introduction}}
-
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.}  Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.
-However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when they are highly dependent and intertwined.
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
-It is not just the diagonals of the variance-covariance matrix this affects.
-
-In this article, we introduce a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in the data.
-This model has two parts, one of which uses a set of additive random effects capturing \first- and \second-order dependencies in the relational data.
-The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.
-We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a generalized linear model with robust standard errors) to those obtained with the AME framework.
-
-This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
-
-\section{\textbf{Limitations of Standard Framework}}
-
-Generalized linear models are typically expressed via a stochastic and systematic component.
-The stochastic component reflects assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function such as the normal, binomial, or Poisson.
-Each dyad in the sample is independently drawn from a particular distribution, given $\theta_{ij}$.
-The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariates, $\mathbf{X}_{ij}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
-A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of the distribution, each of the dyadic observations is conditionally independent.
-
-The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
-After having chosen a set of covariates and specifying a distribution, the joint density function over all dyads using the observations from Table 1 is given:
-
-\vspace{-8mm}
-\begin{align}
-\begin{aligned}
-	P(y_{ij}, y_{ik}, \ldots, y_{lk} | \theta_{ij}, \theta_{ik}, \ldots, \theta_{lk}) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots \times P(y_{lk} | \theta_{lk}) \\
-	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
-\end{aligned}
-\end{align}
-
-\noindent The joint probability is converted into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
-
-The parameters are estimated by maximizing the likelihood.
-The likelihood as defined above is only valid if  $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
-
-\section{\textbf{Dependencies in Dyadic Data}}
-
-Relational, or dyadic, data provide measurements of how sets of actors relate to one another.
-Typically, pairs of actors are examined.
-These pairs are called \textit{dyads}.
-Such data can be arranged as a directed dyadic design in which the unit of analysis is some set of $n$ actors that have been paired together to form a matrix of dyads, i.
-e., represented as a $n \times n$ matrix, $\mathbf{Y}$, with diagonals typically undefined.
-The $ij^{th}$ entry defines the relationship of $i$ and $j$ and can be continuous or discrete.
-It can also be symmetric (known as undirected) or asymmetric (called directed).
-This is not the stacked dyad structure normally found in social science, but it is equivalent and presents a format in which the interdependencies are exposed and transparent.
-The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.
-
-One important dependency is nodal dependency, whereby each actor has similar interactions with many of its partners, leading to significant heterogeneity in activity levels across nodes.
-An important implication of this across-node heterogeneity is within-node homogeneity of ties: the values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, are similar to each other because each has a common sender.
-If a sender $i$ is more active active in the network than other senders, it will have greater homogeneity of its linkages.
-More popular targets show greater similarity across linkages in a given column, $\{y_{ji},y_{ki},y_{li}\}$.
-Actors which are more likely to send ties in a network are also more likely to receive them.
-As a result the rows and columns in an adjacency matrix may be correlated.
-Another structural interdependence is reciprocity.
-This is a second-order, or dyadic dependency whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
-If the data are simply stacked dyads in some arbitrary order, this kind of dependency is harder to visualize, but exists nonetheless.
-
-The presence of these interdependencies in relational data complicates the practical assumption of observational independence.
-Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-By ignoring these potential interdependencies, we often ignore important features of the problem under study.
-The study of international relations is founded on the relations among actors.
-Why ignore the interdependencies that led to the study of IR in the first place?
-
-The additive part described above does not handle third-order dependency.
-A third-order dependency is the dependency between triads, not dyads.
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
-
-For example, a common finding is that democracies are more likely to form trade agreements with one another.
-Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
-A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
-Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
-Another third-order dependence pattern is stochastic equivalence.
-Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
-Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
-Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
-Latent variables can account for third-order dependence patterns within the context of the SRRM.
-These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
-
-\section{\textbf{Additive and Multiplicative Effect Models for Networks}}
-
-The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
-It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.''  At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
-Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
-Fortunately, it is easy to empirically examine dependencies in relational data.
-
-The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
-It was introduced in \cite{hoff:2008}; the open-source computer package associated with this approach is \cite{amenpkg}.
-We describe it below.
-
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
-\label{eqn:ame}
-\end{aligned}
-\end{align}
-
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
-
-\subsection{\textbf{Social Relations Model: Additive Part of AME}}
-
-\citet{warner:etal:1979} developed a social relations model (SRM), using ANOVA decomposition, that focuses on such dependencies.
-SRM provides the error structure for the additive effects component of the AME framework.
-SRM decomposes the variance of dyadic observations in terms of a) heterogeneity across row means (often called out-degree), b) column means (or, in-degree), c) correlation between row and column means, and d) correlations within dyads.
-\citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= \mu + e_{ij} \\
-	e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
-	\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
-	\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
-	\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix},
-\label{eqn:srmCov}
-\end{aligned}
-\end{align}
-
-wherein $\mu$ is the density of a network, and $e_{ij}$ represents residual variation in each cell (or dyad).
-The residual variation has three separate parts: a)   row/sender effects ($a_{i}$), b)  column/receiver effects ($b_{j}$), and c) within-dyad effects ($\epsilon_{ij}$).
-The row and column effects are modeled jointly to account for correlation in how active an actor is in sending and receiving ties.
-Heterogeneity in the row and column means is modeled by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively.
-$\sigma_{ab}$ describes whether actors which send a lot of ties also receive a lot.
-Second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a reciprocity, parameter $\rho$.
-
-The SRM covariance structure described in Equation~\ref{eqn:srmCov} can be incorporated into the systematic component of a GLM framework to produce the social relations regression model (SRRM): $\bm\beta^{\top} \mathbf{X}_{ij} + a_{i} + b_{j} + \epsilon_{ij}$, where $ \bm\beta^{\top} \mathbf{X}_{ij}$ accommodates the inclusion of dyadic, sender, and receiver covariates \citep{hoff:2005}.
-This is a regression framework with additive, random effects, but it models variances and covariances often seen in relational data.
-Furthermore, this accommodates a diversity of outcome distributions.
-In the case of binary data, this can be done by utilizing a random effects logistic (or probit) latent variable representation.
-
-\subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}
-
-The latent factor model (LFM) expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
-General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.
-
-\begin{align}
-\begin{aligned}
-	\alpha(&\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{\top} \Lambda \textbf{u}_{j} \\
-	&\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \text{, where}\\
-	 &\Lambda \equiv K \times K \text{ diagonal matrix}
-\label{eqn:latAlpha}
-\end{aligned}
-\end{align}
-
-Many observed  networks exhibit varying degrees of stochastic equivalence and homophily.
-In the latent factor model, each actor has an unobserved vector of characteristics, $\textbf{u}_{i} = \{u_{i,1}, \ldots, u_{i,K} \}$, describing each node's behavior in the network.
-The probability of a tie between $i$ and $j$ depends on the extent to which $\textbf{u}_{i}$ and $\textbf{u}_{j}$ are ``similar'' and on whether the entries of $\Lambda$ are greater or less than zero.
-Similarity in the latent factors, $\textbf{u}_{i} \approx \textbf{u}_{j}$, corresponds to stochastic equivalence; the eigenvalue determines whether the network exhibits positive or negative homophily.
-Accordingly, this approach can represent both homophily and stochastic equivalence.
-
-In directed (asymmetric) networks there are two sets of latent positions, denoted $\textbf{u}$ and $\textbf{v}$, representing the sending and receiving positions separately.
-Thus, actors in the network have a vector of latent characteristics describing their behavior as a sender, denoted by $\textbf{u}$, and as a receiver, $\textbf{v}$: $\textbf{u}_{i}, \textbf{v}_{j} \in \mathbb{R}^{K}$.
-This can alter the probability of an interaction between $ij$ additively: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, where $\textbf{D}$ is a $K \times K$ diagonal matrix.
-
-The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
-
-\section{Simulation study}
-
-Here we consider inference for a regression parameter $\beta$ of a linear or generalized linear model for a network in the case where there is an omitted variable.
-The true data-generating models we consider are of the form
-
-$y_{i,j} \sim  \mu + \beta x_i x_j + \gamma w_i w_j + \epsilon_{i,j} $
-
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_i \} \in \mathbb R^n$ is a vector of observed node-specific characteristics, and $w = \{ w_i\} \in \mathbb R^n$ is a vector of unobserved node-specific characteristics.
-We compare inference for $\beta$ using three models:
-
-\begin{itemize}
-	\item A naive regression model assuming independent errors; 
-	\item A latent factor model; 
-	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
-\end{itemize}
-
-We make these comparisons in the context of a binary network outcome with a probit regression model.
-
-One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
-
-Alternatively, bias and precision can be summarized by confidence interval coverage and width.
-Coverage should ideally be at the nominal level.
-If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
-
-\begin{figure}
-	\centering
-	\caption{Bias in parameter when homophily is ignored.}
-	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{ameSimBias_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimBias_beta.pdf}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{Coverage in parameter when homophily is ignored.}
-	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{ameSimCover_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimCover_beta.pdf}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\caption{Correlation between missing variable and multiplicative random effect in AME.}
-	\label{fig:ameCorr}
-	\includegraphics[width=1\textwidth]{ameSimCorr.pdf} \\
-\end{figure}
-
-\section{Replications}
-
-\subsection{Design}
-
-For the purpose of this study, we choose five prominent studies from the broad field of international relations and international political economy that utilize relation data \citep{mcdonald:2004, reiter:stam:2003, rose:2004, weeks:2012, gibler:2017}.
-We chose to replicate studies that were written fairly recently, since 2000, and that have been cited over 100 times.
-Each of these pieces was published in a prominent journal and is well-known in the literature.
-They all used the standard approach in political science, which is to employ some form of general linearized regression that ignores dyadic interdependencies.
-Post estimation, standard errors are often adjusted in an attempt to account for clustering of observations.
-
-\begin{table}
-\caption{Features of the Studies Replicated. }
-	\begin{tabular}{lcccccc}
-		 & Model & \# of Actors & Years  & \# of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \hline\hline
-		Weeks (2012) & Logit & & & $901,540$ & Directed & Robust \\
-		Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\
-		McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\		
-		Gibler (2017) & Logit & & & $650,557$ & UnDirected & no \\		
-		Rose (2004) & OLS & & & $234,597$ & Directed & Robust \\
-	\hline\hline
-	\end{tabular}
-\end{table}
-
-We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
-
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
-We believe that it does, and that the dynamic latent factor model provides a step forward.
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l| p{7cm} l}
-	& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
-	& &  in a Network Model? \\
-	\hline\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\
-		\hline
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
-		\hline
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
-		\hline
-		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 		
-		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
-Cases in which the finding is not replicated are highlighted in bold.}
-	\label{tab:modelFindingSumm}
-\end{table}
-
-We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
-
-% % % % % % % % % % % % % % % % % % 
-By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
-To show that this is the case, we examine whether our approach achieves better predictive performance in an out of sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-Then for each $s \in \{1,\ldots,k\}$, we:
-
-\begin{enumerate}
-	\item estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$,
-	\item and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters. 
-\end{enumerate}
-
-\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-
-We set a number of benchmarks for comparison.
-First we compare the AME model to a GLM model using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
-We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
-The lagged dependent variable is the equivalent of saying that conflict and peace are relatively likely to persist between dyads, while the inclusion of a lagged reciprocity term in a GLM framework is a simple way to account for retaliatory strikes.
-
-We utilize three performance criterions to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
-ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
-An issue with an ROC Curve when looking at conflict, is that it is relatively rare at the dyadic level: in most years only 3\% of possible dyads are in conflict with one another.
-If peace is common, even a poor model will have a very low False Positive Rate.
-
-To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
-These examine the trade-offs between the percentage of conflicts a model predicts, and the percentage of predicted conflicts which occur.
-Lastly, we examine separation plots \citep{greenhill:etal:2011}.
-These provide an intuitive visualization of the accuracy of our predictions by juxtaposing a line showing the predicted probability of conflict with whether conflict actually occurs for all cases (where the cases are sorted by the predicted probability and then colored to indicate the outcome).
-Here a perfect model would have all cases where conflict actually exists on the right with a predicted probability of $1$, and would predict $0$ in all other cases.
-All of the models' performance out of sample by these metrics are displayed in figure \ref{fig:roc_ame}.
-The AME model with covariates is the best performing model out of sample in all cases.
-This model outperforms each of the GLM variants by a notable margin.
-% % % % % % % % % % % % % % % % % %
-
-\begin{table}[ht]
-\centering
-	\begin{tabular}{l|l cc}
-	~ & ~ & GLM & AME \\
-	\hline\hline
-		\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
-		~ & AUC (PR) & 0.08 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
-		~ & AUC (PR) & 0.13 & \textbf{0.28} \\
-		\hline
-		\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
-		~ & AUC (PR) & 0.00 & \textbf{0.08} \\
-		\hline
-		\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
-		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\
-	\hline\hline
-	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach.
-Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
-The fifth involved a gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
-Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
-	\label{tab:modelPerfSumm}
-\end{table}
-\FloatBarrier
-
-\subsection{Replication of Weeks (2012)}
-
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
-One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-Her time frame is from 1946-1999.
-The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
-She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
-
-The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
-In Weeks's initial models, as detailed in figure \ref{fig:weeksCoef} she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
-When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
-We similarly find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in figure \ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
-In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
-In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
-So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
-The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in figure \ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Reiter \& Stam (2003)}
-
-\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent.
-They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that ``We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
-This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
-We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
-The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
-Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
-
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
-In our replication using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
-Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitauc} and Figure ~\ref{fig:reitpr}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:reitauc}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
-\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
-\end{figure}
-
-\newpage
-\subsection{Replication of Rose (2004)}
-
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.
-Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
-He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders...
-(2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
-
-When we compare the results of Rose's original OLS model to our model that accounts for network dependencies, the results are generally similar.
-As you can see in Figure~\ref{fig:roseCoef}, the main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- remains when we move from an OLS to a Gaussian AME model.
-The most striking difference between the models is that, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies.
-Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-Excitingly, when we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out of sample -- 3.23 for the OLS model and 1.77 for the AME model.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
- \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{ABplot_rose_top10.pdf}
- \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
-\end{figure}
-
-%\begin{figure}
-%\centering   
-%\subfigure[Nodal Random Effects]%{\label{fig:rosesend}\includegraphics[width=65mm]{ABplot_rose_top10.pdf}}
-%\subfigure[Receiver Random Effects]{\label{fig:rosereceive}\includegraphics[width=65mm]{ABplot_receiver_rose.pdf}}
-%\caption{Sender and receiver random effects for AME estimation of Rose (2004).}
-%\label{fig:rosesr}
-%\end{figure}
-
-\newpage
-\subsection{Replication of McDonald (2004)}
-
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-He observes that knowledge about the link between conflict and trade is indeterminate in the field of international relations, noting that competing explanations persist.
-He calls for more precise empirical tests.
-Importantly, \citet{mcdonald:2004} includes the interdependence argument---that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)'' in his overview of underdeveloped hypotheses.
-
-Accordingly, the primary contribution of the study is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-To do so, \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measures free trade in two ways.
-The first  captures the idea that larger protected sectors generate greater societal pressures resulting in  pockets of support for war.
-This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state i's GDP or total dyadic trade divided by state j's GDP.'' (560).
-Finally, the binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include a spline correction for time-series data as well as robust standard errors clustered on each dyad.
-
-% TO DO: include discussion and analysis of different K's in appendix (for M cD)
-\begin{figure}
-\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-\end{figure}
-
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-\end{figure}
-
-% TO DO: fix bib
-
-Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries {i} and {j} are highly dependent and countries {j} and {k} are also highly dependent, then we are likely to observe high dependency between countries {i} and {k}.
-Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
-This indicates that conflict is less likely between members of a trade community.
-Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and replicated model and Figure~\ref{fig:mcdall} demonstrates the predictive performance of each model.
-Figure ~\ref{fig:mcdcirc} represents the multiplicative effects of the model.
-This plot enables us to consider actors with similar sending and receiving patterns resulting from stochastic equivalence and homophily.
-In the upper arch for senders, there is a clear clustering of United States' allies (the United Kingdom, Netherlands, and Japan).
-We also observe an African cluster in the top right segment for receivers (Egypt, Niger, and Zambia for example).
-These countries both trade amongst themselves and have similar conflict behavior -- stochastic equivalence is present among the first group, homophily among the second -- and so once we account for these clusters using the multiplicative effects, the residual effect of trade actually increases the likelihood of conflict.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{mcdonaldcirc.pdf}
-\caption{\label{fig:mcdcirc} Visualization of multiplicative effects. Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\subsection{Replication of Gibler (2017)}
-
-% We chose this article because it is published in 2017, demonstrating that the concern with interdependence continues to the present and because it was published in the so-called flagship journal of political science, \textit{The American Political Science Review.} It examines the link between dyadic capability differences and international conflict in terms of how and when states became independent members of the international system.
-It assumes that membership in the international system---defined in typical ways---has a substantial influence on capabilities.
-In the same way that capabilities don't seem to vary enormously over time, at least comparatively, states become independent in a particular location, clustered with surrounding states.
-Neither does this change much over time.
-Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-The replications we have undertaken are all from articles over the past fifteen years.
-A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
-In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-\begin{table}
-\begin{center}
-\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000$ iterations.  \label{tab:gibme}}
-\begin{tabular}{lrr} \toprule
-& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-Allied & 0.142&  0.023 \\
-Joint Democracy &\bf -0.507&  0.045 \\
-Peace Years \bf &-0.260& \bf -0.060 \\
-Spline 1 &\bf -0.001& \bf 0.00 \\
-Spline 2&\bf -0.000& \bf 0.00\\
-Spline 3 &-0.000& \bf 0.00\\ 
-Contiguity &\bf 2.412& \bf 0.640 \\
-Parity &0.075&  -0.013 \\
-Parity at entry year&\bf 0.868&  0.002 \\ 
-Rivalry &\bf 2.031&\bf 0.721\\ 
-Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-\end{tabular}\\
-\end{center}
-{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
-For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
-\end{table}
-
-We re-estimated model 6 from Table 6 (2017,34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
-The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
-Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the amen results.
-Not only is the value of the this parameter small, but it has a very large relative standard error, over a magnitude larger than the parameter itself ($z= 0.038$).
-In addition, joint democracy follows the same pattern of importance in the logistic results, but this disappears once interdependencies are modeled.
-As might be expected the strong geographic clustering in the original study is about one-quarter as strong in the \texttt{amen} estimations.
-Similarly, rivalry coefficients are about one-third the size in the \texttt{amen} formulation, but a great deal more precisely measured ($z=18.116$).
-
-Beyond more informative fixed effect coefficients, the \texttt{amen} approach also provides information about the interdependencies that were modeled.
-The most pertinent of these may be the dyadic and triadic dependencies which are presented in Figure~\ref{fig:gibler:dyadtriad}, which further reveal that the assumption of independence among the dyadic data in this study can be strongly rejected.
-Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies.
-Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.
-
-% sd.rowmean sd.colmean   dyad.dep  triad.dep 
-% 0.03210221 0.03210221 0.99362646 0.00130249 
-\begin{figure}
-\caption{\label{fig:gibler:dyadtriad} The left panel is  a plot of the dyadic dependencies in the Gibler model 6; the right panel shows the triad dependencies.
-Note that the red line represents the null hypothesis of no dependencies, indicating that the standard logistic approach is far from what is uncovered with this analysis.
-}
-\includegraphics[width=\textwidth]{netplots}
-\end{figure}
-
-\section{Conclusion}
-
-It is no longer necessary to assume that the interesting, innate interdependencies in relational data can be ignored.
-Nor do they have to be approximated with ad hoc, incomplete solutions that purport to control for dependencies by modifying the post-estimation standard errors of the estimated coefficients.
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.
+\input{1_intro}
+\input{2_model}
+\input{3_sim}
+\input{4_app}
+\input{5_concl}
+\input{6_appendix}
 
@@
 \bibliographystyle{authordate1}
-\clearpage
-\section*{All About Us}
+% \clearpage
+% \section*{All About Us}
+
 \end{document}

### 2017-09-29 code/2_codeForPaperGraphics/reiter_stam_posterior.R

@@ -1,0 +1,47 @@
+############################################
+rm(list=ls())
+
+if(Sys.info()['user'] %in% c('s7m', 'janus829')){
+	resultsPath = '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/output/'
+	inputPath= '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/input/'
+	plotPath = '~/Dropbox/Research/netsMatter/replications/0_finalRepFigs/' }
+
+#
+source('~/Research/netsMatter/code/helpers/functions.R')
+source('~/Research/netsMatter/code/helpers/ameHelpers.R')
+
+#
+loadPkg(c('amen','countrycode'))
+############################################
+
+############################################
+#
+load( paste0(resultsPath,'reiterStam_baseModel.rda') )
+load( paste0(resultsPath,'ameFit_k2.rda') )
+############################################
+
+############################################
+# coef summ
+# paramPlot(ameFit$BETA[,1:12])
+# paramPlot(ameFit$BETA[,12:23])
+
+ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+	mu=mean(x),
+	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
+# modSumm
+############################################
+
+############################################
+#
+effdat = getAddEffData(fit = ameFit) ##This function is in helperEx.R
+effdat$actor = countrycode::countrycode(effdat$actor, 'cown', 'country.name')
+effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor'])
+
+## All countries
+addEffPlot(fit = effdat, addEffData = effdat, row = T)
+############################################
+
+############################################
+#
+
+############################################

### 2017-10-01 paper/1_intro.tex
so much to delete

@@
 Most of these studies  use a generalized linear model (GLM) to estimate regression coefficients.
-However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.\footnote{Aronow et alia (2015) provide a solution to simple clustering via covariance estimation but do not address the broader issues of interdependence focused on herein.} Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
+However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
+Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
 This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
@@
 Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad, specifically, if a rebel group initiates a conflict with a government, the government will reciprocate that conflictual behavior.
-We capture these two dependencies, referred to as first- and second-order dependencies within the additive effects portion of the model.
+We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
 The multiplicative effects portion of the model captures dependencies that results from groups of actors clustering together or organizing into communities.
 Clusters and communities often result in relational data because actors possess some set of shared attributes that affect their probability of interacting with one another.
-For example, when considering alliance patterns we might reasonably expect that if two actors are already allied to a third that their probability of allying with one another is higher than average.
 
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interact- ing with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation
+% A common example for why we observe dyads clusters in networks is due to transitivity.
+This is simply the principle that when considering alliance patterns we might reasonably expect that if two actors are already allied to a third that their probability of allying with one another is higher than average.
 
-a specific actor is responsible for sent or received events such
-
-This model has two parts, one an additive effects portion that
-
-, one of which uses a set of additive random effects to capture dependencies arising because a specific actor sent or received an events such as conflict to others.
-
-\first- and \second-order dependencies in the relational data.
-The second part is a multiplicative component that captures \third-order dependencies, i.e., phenomena that leads to triads such as homophily and stochastic equivalence.
 We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
-
 This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
 It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample  predictions (cross-validations) are substantially improved with this approach.
+Finally, we show that out-of-sample predictions are substantially improved with this approach.

### 2017-10-01 paper/2_model.tex
so much to delete

@@
-\section{\textbf{Limitations of Standard Framework}}
-
-Generalized linear models are typically expressed via a stochastic and systematic component.
-The stochastic component reflects assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function such as the normal, binomial, or Poisson.
-Each dyad in the sample is independently drawn from a particular distribution, given $\theta_{ij}$.
-The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariates, $\mathbf{X}_{ij}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
-A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of the distribution, each of the dyadic observations is conditionally independent.
-
-The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
-After having chosen a set of covariates and specifying a distribution, the joint density function over all dyads using the observations from Table 1 is given:
-
-\vspace{-8mm}
-\begin{align}
-\begin{aligned}
-	P(y_{ij}, y_{ik}, \ldots, y_{lk} | \theta_{ij}, \theta_{ik}, \ldots, \theta_{lk}) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots \times P(y_{lk} | \theta_{lk}) \\
-	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
-\end{aligned}
-\end{align}
-
-\noindent The joint probability is converted into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
-
-The parameters are estimated by maximizing the likelihood.
-The likelihood as defined above is only valid if  $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
-
 \section{\textbf{Dependencies in Dyadic Data}}
@@
 The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.
+
+\begin{figure}
+	\includegraphics[width=1\textwidth]{df_adj_net3.png}
+	\caption{Relational data from dyadic designs to networks}
+	\label{fig:ddToNet}
+\end{figure}
 

### 2017-10-01 paper/6_appendix.tex
so much to delete

\clearpage

\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{A.\arabic{table}}
\setcounter{table}{0}
\renewcommand{\thesection}{A.\arabic{section}}
\setcounter{section}{0}

\section*{\textbf{Appendix}}

\subsection*{Additive and Multiplicative Effects Gibbs Sampler}

To estimate, the effects of our exogenous variables and latent attributes we utilize a Bayesian probit model in which we sample from the posterior distribution of the full conditionals until convergence. Specifically, given observed data $\textbf{Y}$ and $\textbf{X}$ -- where $\textbf{X}$ is a design array that includes our sender, receiver, and dyadic covariates -- we estimate our network of binary ties using a probit framework where: $y_{ij,t} = 1(\theta_{ij,t}>0)$ and $\theta_{ij,t} = \bm\beta^{\top}\mathbf{X}_{ij,t} + a_{i} + b_{j} + \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} + \epsilon_{ij}$. The derivation of the full conditionals is described in detail in \citet{hoff:2005} and \citet{hoff:2008}, thus here we only outline the Markov chain Monte Carlo (MCMC) algorithm for the AME model that we utilize in this paper.

\begin{itemize}
 \item Given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows:
 \begin{itemize}
 	\item sample $\bm\theta \; | \;  \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
 	\item sample $\bm\beta \; | \;  \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
 	\item sample $\textbf{a}, \textbf{b} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
	\item sample $\Sigma_{ab} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
 	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
 	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
 	\item For each $k \in K$:
 	\begin{itemize}
 		\item Sample $\textbf{U}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}_{[,-k]}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
 		\item Sample $\textbf{V}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
 		\item Sample $\textbf{D}_{[k,k]}  \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \textbf{D} matrix is absorbed into the calculation for $\textbf{V}$ as we iterate through $K$. }
 	\end{itemize}
 \end{itemize}
\end{itemize}

### 2017-10-05 paper/1_intro.tex
front half

@@
 
-\citet{aronow:etal:2015} claim that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.
+\citet{aronow:etal:2015} estimate that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.
 Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
 However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, in order to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process which remains a threat to inference because of the interdependence of observations or measurements.
-It is not just the diagonals of the variance-covariance matrix this affects.
+Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases.
+This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process that remains a threat to inference because of the interdependence of observations or measurements.
+Namely, it is not just the diagonals of the variance-covariance matrix this affects.
 
-In this article, we introduce a Bayesian approach with additive and multiplicative effects for directly modeling relational data to reflect the data generating process that yields interdependencies in relational data.
+In this article, we discuss a Bayesian approach, the Additive and Multiplicative Effects (AME) model, for directly modeling relational data to reflect the data generating process that yields interdependencies in these types of data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.
@@
 We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-The multiplicative effects portion of the model captures dependencies that results from groups of actors clustering together or organizing into communities.
-Clusters and communities often result in relational data because actors possess some set of shared attributes that affect their probability of interacting with one another.
+The multiplicative effects portion of the model captures dependencies that results from groups of actors clustering together or organizing into communities due to meso-scopic features of networks, such as homophily and stochastic equivalence.
+%Clusters and communities often result in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.
 
-% A common example for why we observe dyads clusters in networks is due to transitivity.
-This is simply the principle that when considering alliance patterns we might reasonably expect that if two actors are already allied to a third that their probability of allying with one another is higher than average.
-
-We apply this model, called the Additive and Multiplicative effects (AME) model, to five prominent studies in the international relations literature, and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
-This latent factor approach is able to better capture \first-, \second-, and \third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and often different, than those found in these literatures, and as such offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that out-of-sample predictions are substantially improved with this approach.
+We begin conducting a simulation study to show how AME can recover unbiased and well calibrated regression coefficients in the presence of network dependencies.
+Then we apply the AME model to five prominent studies in the international relations (IR) literature, and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
+This latent factor approach is able to better capture first, second, and third-order interdependencies than the standard approaches.
+It also produces results that are more precise, and at times different, than those found in these literatures.
+As such this approach offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
+Finally, we show that for each replication our network based approach provides substantively more accurate out of sample predictions than the models used in the original studies.
+Thus the approach that we offer here is one that can be used by scholars in the field to not only generate substantive insights, but also enables us to better model the data generating process behind events of interest in international relations.

### 2017-10-05 paper/2_model.tex
front half

@@
 
-Relational, or dyadic, data provide measurements of how sets of actors relate to one another.
-Typically, pairs of actors are examined.
-These pairs are called \textit{dyads}.
-Such data can be arranged as a directed dyadic design in which the unit of analysis is some set of $n$ actors that have been paired together to form a matrix of dyads, i.
-e., represented as a $n \times n$ matrix, $\mathbf{Y}$, with diagonals typically undefined.
-The $ij^{th}$ entry defines the relationship of $i$ and $j$ and can be continuous or discrete.
-It can also be symmetric (known as undirected) or asymmetric (called directed).
-This is not the stacked dyad structure normally found in social science, but it is equivalent and presents a format in which the interdependencies are exposed and transparent.
-The standard statistical approach to these data uses a linear or generalized linear model (GLM) which assumes that each of the dyadic observations is conditionally independent on observed regressors, even though events and relationships between these actors is most often \textit{inter}dependent.
+In thinking about dyadic data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another.
+Each observation is assumed to be independent of the others.
+Thus, for example, a conflict sent from actor the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
+Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
+As a result, the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others.
+
+\input{dd_to_adjmat}
+\FloatBarrier
+
+To start to move away from this assumption and better understand the dependencies that emerge in relational data it is helpful to shift towards structuring dyadic data in the form of an adjacency matrix as shown in the top right of Table~\ref{tab:netDesign}.
+Here rows designate the senders of an event and columns the receivers.
+The cross-sections in this matrix represent the action that was sent by an actor on the row to one designated in the column.
+Thus $y_{ij}$ designates an action, such as a conflictual event or trade flows, that is sent from actor $i$ to actor $j$.
+
+Using the structure of an adjacency matrix we can visualize the types of first and second order dependencies that complicate the analysis of relational data in traditional GLMs.
+Figure~\ref{fig:adjMatDeps} clarifies the types of dependencies that can manifest in these types of data structures.
+The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that values across a particular row of an adjacency matrix may be more similar to each other than other values in the adjacency matrix because each of these values has a common sender.
+Homogeneity in interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
+Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors who are much more active than others \citep{barabasi:reka:1999}.
+Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor may be more active than others, parameter estimates from standard statistical models will be biased.
 
 \begin{figure}
-	\includegraphics[width=1\textwidth]{df_adj_net3.png}
-	\caption{Relational data from dyadic designs to networks}
-	\label{fig:ddToNet}
+	\begin{tabular}{cc}
+		\texttt{Sender heterogeneity} & \texttt{Receiver Heterogeneity} \\
+		\includegraphics[width=.45\textwidth]{adjRowDep.png} & \includegraphics[width=.45\textwidth]{adjColDep.png} \\
+		\texttt{Sender-Receiver Covariance} & \texttt{Reciprocity} \\
+		\includegraphics[width=.45\textwidth]{adjRowColCovar.png} & \includegraphics[width=.45\textwidth]{adjRecip.png} \\
+	\end{tabular}
+	\caption{Nodal and dyadic dependencies in relational data.}
+	\label{fig:adjMatDeps}
 \end{figure}
 
-One important dependency is nodal dependency, whereby each actor has similar interactions with many of its partners, leading to significant heterogeneity in activity levels across nodes.
-An important implication of this across-node heterogeneity is within-node homogeneity of ties: the values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, are similar to each other because each has a common sender.
-If a sender $i$ is more active active in the network than other senders, it will have greater homogeneity of its linkages.
-More popular targets show greater similarity across linkages in a given column, $\{y_{ji},y_{ki},y_{li}\}$.
-Actors which are more likely to send ties in a network are also more likely to receive them.
-As a result the rows and columns in an adjacency matrix may be correlated.
-Another structural interdependence is reciprocity.
-This is a second-order, or dyadic dependency whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
-If the data are simply stacked dyads in some arbitrary order, this kind of dependency is harder to visualize, but exists nonetheless.
+For similar reasons one also needs to take into account that there is a shared dependence between dyadic observations that share a common receiver.
+The bottom-left panel, illustrates that sender and receiver type dependencies can also blend together.
+Specifically, as actors who are more likely to send ties in a network tend to also be more likely to receive them.
+As a result, the rows and columns in an adjacency matrix are often correlated.
+For example, consider that trade flows both from and to many wealthy, developed countries.
+The bottom-right panel, highlights a second order dependence, specifically, reciprocity.
+This is a dependency occurring within dyads involving the same actors whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
+The concept of reciprocity has deep roots in the study of relations between states \citep{richardson:1960,keohane:1989}.
+The purpose of highlighting each of these dependencies through the set of panels in Figure~\ref{fig:adjMatDeps} is to more easily visualize how each of these dependencies manifest in relational data.
+Alternatively, when we simply simply stack dyads in some arbitrary order, these dependencies are easy to ignore.
 
-The presence of these interdependencies in relational data complicates the practical assumption of observational independence.
+The dependencies discussed so far involve at most dependence between two actors.
+For most relational data, however, dependencies do not simply manifest at this level.
+More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors.
+These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{wasserman:faust:1994,zinnes:1967}.
+In Figure~\ref{fig:thirdDeps} we provide a visualization of a hypothetical relational dataset.
+Here the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
+Each node is colored by the latent group to which they belong.
+
+\begin{figure}[ht]
+	\includegraphics[width=.7\textwidth]{stochEquiv.pdf}
+	\caption{Visualization of network with meso-scopic features.}
+	\label{fig:thirdDeps}
+\end{figure}
+
+Clear from the visualization is that the actors belonging to the same group have a higher likelihood of having an interaction with each other than those from the other group.
+A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another in the 2004 United States Election \citep{adamic:glance:2005}.
+They found that the degree of interaction between right and left leaning blogs was minimal, and that most of these blogs simply linked to those of their own ilk.
+This showcases the types of higher order dependencies that can emerge in relational data.
+First, the fact that interactions was determined by a shared attribute, in this case political ideology, is an example of homophily.
+Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
+The other major type of meso-scopic feature that emerges in relational data is community structure, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
+This concept simply refers to the idea that groups of nodes that act similarly in the network are stochastically equivalent.
+In the example we have laid out above each of the left leaning blogs would be considered stochastically equivalent to one another.
+
+The major implication of the presence of homophily, stochastic equivalence, and the other dependencies we discussed above are that they complicate the practical assumption of observational independence.
 Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
@@
 
-The additive part described above does not handle third-order dependency.
-A third-order dependency is the dependency between triads, not dyads.
-Third-order effects can arise from the presence of some set of shared attributes between actors that affects their probability of interacting with one another.
-Individuals try to close triads because this is thought to be a more stable or preferable social situation \citep{wasserman:faust:1994,zinnes:1967}.
-
-For example, a common finding is that democracies are more likely to form trade agreements with one another.
-Both geographic proximity and a country's political system are examples of homophily, which captures the idea that the relationships between actors with similar characteristics in a network are likely to be stronger than nodes with different characteristics.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which have a long history in international relations.
-A binary network where actors tend to form ties based on shared characteristics  often leads to a network with a high number of ``transitive triads'' in which  sets of actors $\{i,j,k\}$ are each linked to one another.
-Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
-Another third-order dependence pattern is stochastic equivalence.
-Actors $i,j$ are stochastically equivalent if the probability of $i$ relating to every other actor is similar to the probability for $j$.
-Equivalence implies the existence of groups of nodes in a network with similar relational patterns.
-Recent work has found stochastic equivalence to be important \citep{manger:etal:2012}.
-Latent variables can account for third-order dependence patterns within the context of the SRRM.
-These models assume that relationships between nodes are mediated by a small number ($K$) of node-specific unobserved latent variables.
+% Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
 
@@
 We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.'' At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
-Ignoring the assumption of independence is a widely-noted shortcoming of many studies, but it continues to be the standard practice in much scholarship.
-Fortunately, it is easy to empirically examine dependencies in relational data.
 
-The additive and multiplicative effects model (AME) can help to eliminate this color-blindness to interdependence.
-It was introduced in \cite{hoff:2008}; the open-source computer package associated with this approach is \cite{amenpkg}.
-We describe it below.
-
-Parameter interpretation is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
+Further parameter interpretation in the AME framework is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
 
@@
 In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is available in the \pkg{amen} $\sf{R}$ package to estimate this type of generalized linear mixed effects model from Gaussian, binary, ordinal, and other relational data types.
-
-\subsection{\textbf{Social Relations Model: Additive Part of AME}}
-
-\citet{warner:etal:1979} developed a social relations model (SRM), using ANOVA decomposition, that focuses on such dependencies.
-SRM provides the error structure for the additive effects component of the AME framework.
-SRM decomposes the variance of dyadic observations in terms of a) heterogeneity across row means (often called out-degree), b) column means (or, in-degree), c) correlation between row and column means, and d) correlations within dyads.
-\citet{wong:1982,li:loken:2002} provide a random effects representation of the SRM:
-
-\begin{align}
-\begin{aligned}
-	y_{ij} &= \mu + e_{ij} \\
-	e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
-	\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
-	\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
-	\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix},
-\label{eqn:srmCov}
-\end{aligned}
-\end{align}
-
-wherein $\mu$ is the density of a network, and $e_{ij}$ represents residual variation in each cell (or dyad).
-The residual variation has three separate parts: a)   row/sender effects ($a_{i}$), b)  column/receiver effects ($b_{j}$), and c) within-dyad effects ($\epsilon_{ij}$).
-The row and column effects are modeled jointly to account for correlation in how active an actor is in sending and receiving ties.
-Heterogeneity in the row and column means is modeled by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively.
-$\sigma_{ab}$ describes whether actors which send a lot of ties also receive a lot.
-Second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a reciprocity, parameter $\rho$.
-
-The SRM covariance structure described in Equation~\ref{eqn:srmCov} can be incorporated into the systematic component of a GLM framework to produce the social relations regression model (SRRM): $\bm\beta^{\top} \mathbf{X}_{ij} + a_{i} + b_{j} + \epsilon_{ij}$, where $ \bm\beta^{\top} \mathbf{X}_{ij}$ accommodates the inclusion of dyadic, sender, and receiver covariates \citep{hoff:2005}.
-This is a regression framework with additive, random effects, but it models variances and covariances often seen in relational data.
-Furthermore, this accommodates a diversity of outcome distributions.
-In the case of binary data, this can be done by utilizing a random effects logistic (or probit) latent variable representation.
-
-\subsection{\textbf{Latent Factor Model: Multiplicative Part of AME}}
-
-The latent factor model (LFM) expands the SSRM to include an additional term, $y_{ij}$, $\alpha(u_{i}, u_{j})$, that captures latent third-order characteristics of a network---where $u_{i}$ and $u_{j}$ are node-specific latent variables for symmetric data.
-General definitions for how $\alpha(u_{i}, u_{j})$ is defined for these latent variable models are shown in Equation~\ref{eqn:latAlpha}.
-
-\begin{align}
-\begin{aligned}
-	\alpha(&\textbf{u}_{i}, \textbf{u}_{j}) = \textbf{u}_{i}^{\top} \Lambda \textbf{u}_{j} \\
-	&\textbf{u}_{i} \in \mathbb{R}^{K}, \; i \in \{1, \ldots, n \} \text{, where}\\
-	 &\Lambda \equiv K \times K \text{ diagonal matrix}
-\label{eqn:latAlpha}
-\end{aligned}
-\end{align}
-
-Many observed  networks exhibit varying degrees of stochastic equivalence and homophily.
-In the latent factor model, each actor has an unobserved vector of characteristics, $\textbf{u}_{i} = \{u_{i,1}, \ldots, u_{i,K} \}$, describing each node's behavior in the network.
-The probability of a tie between $i$ and $j$ depends on the extent to which $\textbf{u}_{i}$ and $\textbf{u}_{j}$ are ``similar'' and on whether the entries of $\Lambda$ are greater or less than zero.
-Similarity in the latent factors, $\textbf{u}_{i} \approx \textbf{u}_{j}$, corresponds to stochastic equivalence; the eigenvalue determines whether the network exhibits positive or negative homophily.
-Accordingly, this approach can represent both homophily and stochastic equivalence.
-
-In directed (asymmetric) networks there are two sets of latent positions, denoted $\textbf{u}$ and $\textbf{v}$, representing the sending and receiving positions separately.
-Thus, actors in the network have a vector of latent characteristics describing their behavior as a sender, denoted by $\textbf{u}$, and as a receiver, $\textbf{v}$: $\textbf{u}_{i}, \textbf{v}_{j} \in \mathbb{R}^{K}$.
-This can alter the probability of an interaction between $ij$ additively: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, where $\textbf{D}$ is a $K \times K$ diagonal matrix.
-
-The LFM captures a wide range of interdependence patterns that arise in relational data such as often studied in international relations.
+The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.\footnote{This procedure is described in detail in \citet{hoff:2008} and \citet{minhas:etal:2016:arxiv}.
+A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is described in the appendix.}

### 2017-10-05 code/other/homophStochEquivNets.R

@@ -1,0 +1,95 @@
+rm(list=ls())
+library("igraph")
+library('RColorBrewer')
+set.seed(6886)
+
+#### stoch equiv
+# adjMat = matrix(0, nrow=20, ncol=20)
+adjMat = matrix(0, nrow=16, ncol=16)
+high = 1:8
+oth = 9:16
+# med = 5:11
+# low = 12:16
+# oth = 17:20
+
+adjMat[high,high] = rbinom(length(adjMat[high,high]), 1, .4) # high-high
+# adjMat[med,med] = rbinom(length(adjMat[med,med]), 1, .6) # med-med
+# adjMat[low,low] = rbinom(length(adjMat[low,low]), 1, .6) # low-low
+adjMat[oth,oth] = rbinom(length(adjMat[oth,oth]), 1, .4) # low-low
+
+# adjMat[high,med] = rbinom(length(adjMat[high,med]), 1, .2) # high - med
+# adjMat[med,high] = rbinom(length(adjMat[med,high]), 1, .25) # med - high
+
+# adjMat[low,high] = rbinom(length(adjMat[low,high]), 1, .0) # low - high
+# adjMat[high,low] = rbinom(length(adjMat[high,low]), 1, .0) # high - low
+
+# adjMat[med,low] = rbinom(length(adjMat[med,low]), 1, .1) # med - low
+# adjMat[low,med] = rbinom(length(adjMat[low,med]), 1, .3) # low - med
+
+adjMat[oth,high] = rbinom(length(adjMat[oth,high]), 1, .05) # med - high
+adjMat[high,oth] = rbinom(length(adjMat[high,oth]), 1, .05) # med - high
+
+# adjMat[oth,low] = rbinom(length(adjMat[oth,low]), 1, .15) # med - low
+# adjMat[low,oth] = rbinom(length(adjMat[low,oth]), 1, .15) # med - low
+
+# adjMat[oth,med] = rbinom(length(adjMat[oth,med]), 1, 0) # med - med
+# adjMat[med,oth] = rbinom(length(adjMat[med,oth]), 1, 0) # med - med
+
+
+# oops undirected
+# adjMat[lower.tri(adjMat)] = 0
+# adjMat[lower.tri(adjMat)] = t(adjMat)[lower.tri(adjMat)]
+g = graph_from_adjacency_matrix(adjMat, mode='directed', weighted=NULL, diag=FALSE)
+
+# layout
[…]

### 2017-10-05 paper/1_intro.tex
some notes
mdw edits to section 1 and 2.

@@
+% need to do a better job of differentiating this from piece with peter
+%  here we are adding a simulation exercise to show that ame works
+% doing a big replication across five ir papers
+% we're not "introducing" ame
 \section{\textbf{Introduction}}
@@
 
-We begin conducting a simulation study to show how AME can recover unbiased and well calibrated regression coefficients in the presence of network dependencies.
-Then we apply the AME model to five prominent studies in the international relations (IR) literature, and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
-This latent factor approach is able to better capture first, second, and third-order interdependencies than the standard approaches.
-It also produces results that are more precise, and at times different, than those found in these literatures.
-As such this approach offers substantive insights, which are occluded by ignoring the interdependent nature of the relational data that characterizes many studies in the field of international relations.
-Finally, we show that for each replication our network based approach provides substantively more accurate out of sample predictions than the models used in the original studies.
-Thus the approach that we offer here is one that can be used by scholars in the field to not only generate substantive insights, but also enables us to better model the data generating process behind events of interest in international relations.
+We begin by conducting a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
+Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
+The latent factor approach (AME) is able to better capture first, second, and third-order interdependencies than the standard approach.
+It also produces results that are more precise and at times at odds with those found in these studies in particular, and the broader literatures from which they are drawn.
+As such, this approach offers substantive insights which are often occluded by ignoring the interdependent nature of the relational data that characterize many studies in the field of international relations.
+Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
+Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest in international relations.
+It facilitates the concentration on international relations in the field of international relations.

### 2017-10-05 paper/2_model.tex
mdw edits to section 1 and 2.

@@
 
-In thinking about dyadic data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another.
-Each observation is assumed to be independent of the others.
+In working with dyadic data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another.
+This makes sense if each observation is independent of the others.
 Thus, for example, a conflict sent from actor the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
@@
 As a result, the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others.
+We know this assumption to be false in theory and in practice.
 
@@
 
-To start to move away from this assumption and better understand the dependencies that emerge in relational data it is helpful to shift towards structuring dyadic data in the form of an adjacency matrix as shown in the top right of Table~\ref{tab:netDesign}.
-Here rows designate the senders of an event and columns the receivers.
-The cross-sections in this matrix represent the action that was sent by an actor on the row to one designated in the column.
-Thus $y_{ij}$ designates an action, such as a conflictual event or trade flows, that is sent from actor $i$ to actor $j$.
+To start to move away from this assumption and better understand the dependencies that emerge in relational data, it is helpful to shift towards structuring dyadic data in the form of a matrix--often called an adjacency matrix in the context of spatial analysis--shown in the top right of Table~\ref{tab:netDesign}.
+Rows designate the senders of an event and columns the receivers.
+In the context of events, the cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns.
+Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flows, that is sent from actor $i$ to actor $j$.
 
-Using the structure of an adjacency matrix we can visualize the types of first and second order dependencies that complicate the analysis of relational data in traditional GLMs.
-Figure~\ref{fig:adjMatDeps} clarifies the types of dependencies that can manifest in these types of data structures.
+Using the structure of an adjacency matrix, we can visualize the types of first- and second-order dependencies that complicate the analysis of relational data in traditional GLMs.
+Figure~\ref{fig:adjMatDeps}clarifies the types of dependencies that can manifest in these types of data structures.
 The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that values across a particular row of an adjacency matrix may be more similar to each other than other values in the adjacency matrix because each of these values has a common sender.
 Homogeneity in interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
-Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors who are much more active than others \citep{barabasi:reka:1999}.
+Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors which are much more active than others \citep{barabasi:reka:1999}.
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor may be more active than others, parameter estimates from standard statistical models will be biased.
@@
 The purpose of highlighting each of these dependencies through the set of panels in Figure~\ref{fig:adjMatDeps} is to more easily visualize how each of these dependencies manifest in relational data.
-Alternatively, when we simply simply stack dyads in some arbitrary order, these dependencies are easy to ignore.
+Alternatively, when we simply simply stack dyads in some arbitrary order, these dependencies are easy to ignore and virtually impossible to recognize.
 
@@
 For most relational data, however, dependencies do not simply manifest at this level.
-More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors.
+More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors \citep{poast:2010}.
 These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{wasserman:faust:1994,zinnes:1967}.
-In Figure~\ref{fig:thirdDeps} we provide a visualization of a hypothetical relational dataset.
-Here the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
+In Figure~\ref{fig:thirdDeps} we provide a visualization of a hypothetical relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
 Each node is colored by the latent group to which they belong.
@@
 
-Clear from the visualization is that the actors belonging to the same group have a higher likelihood of having an interaction with each other than those from the other group.
-A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another in the 2004 United States Election \citep{adamic:glance:2005}.
-They found that the degree of interaction between right and left leaning blogs was minimal, and that most of these blogs simply linked to those of their own ilk.
-This showcases the types of higher order dependencies that can emerge in relational data.
-First, the fact that interactions was determined by a shared attribute, in this case political ideology, is an example of homophily.
+The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other; interactions across groups are more rare.
+One prominent example of a network with this type of structure was found in the context of the polarization in US politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another in the 2004 United States Election \citep{adamic:glance:2005}.
+They found that the degree of interaction between right and left leaning blogs was minimal; most of these blogs linked to blogs that were similar, not that were different.
+This showcases the types of higher-order dependencies that can emerge in relational data.
+First, the fact that interactions was determined by a shared attribute, in this case political ideology, is an example of what is known as \textit{homophily}.
 Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
-The other major type of meso-scopic feature that emerges in relational data is community structure, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
+The other major type of meso-scopic feature that emerges in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
 This concept simply refers to the idea that groups of nodes that act similarly in the network are stochastically equivalent.
@@
 
-The major implication of the presence of homophily, stochastic equivalence, and the other dependencies we discussed above are that they complicate the practical assumption of observational independence.
+The major implication of the presence of homophily, stochastic equivalence, and the other dependencies we discussed above is that they complicate the practical assumption of observational independence.
 Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
@@
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.'' At a very minimum, it is necessary to examine whether there is interdependence since it challenges substantive arguments as well as statistical modeling \citep{snijders:2011}.
+We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.'' At a very minimum, it is necessary to examine \textit{whether} there is interdependence since it challenges both substantive arguments and statistical modeling \citep{snijders:2011}.
 
-Further parameter interpretation in the AME framework is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} which portrays directed matrix \:
+Further parameter interpretation in the AME framework is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} (shown for a directed matrix, without loss of generality) \:
 
@@
 In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.\footnote{This procedure is described in detail in \citet{hoff:2008} and \citet{minhas:etal:2016:arxiv}.
+The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.\footnote{This procedure is described in detail in \citet{hoff:2008,hoff:2015:arxiv,minhas:etal:2016:arxiv}.
 A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is described in the appendix.}

### 2017-10-06 paper/5_concl.tex
conclusion is now drafted

@@
 
-It is no longer necessary to assume that the interesting, innate interdependencies in relational data can be ignored.
-Nor do they have to be approximated with ad hoc, incomplete solutions that purport to control for dependencies by modifying the post-estimation standard errors of the estimated coefficients.
+International relations is generally about the interactions and dependences among a set of countries or other important actors such as international governmental organizations (IGOs, such as the WTO) and non-governmental organizations (NGOs, such as the Gates Foundation).
+Some approaches focus on only looking a at small number of these actors, but many scholars examine a large number of actors at a time.
+This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.
+One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.
+It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citet{erikson:pinto:2014,aronow:etal:2015}.
+
+At the same time, we know that research designs which focus on the statistical analysis of dyadic data quickly go astray if the dyadic data are assumed to be independent and identically distributed (iid).
+Virtually all of the standard statistical models--ordinary least squares, logistic and probit regressions, to name a few--fail if the data are not iid.
+By definition dyadic data are not iid and thus the standard approaches can not be used cavalierly to analyze these data.
+\citet{signorino:1999} showed why this is true of models of strategic interaction, but it is more broadly true of models that employ dyadic data.
+We show that latent networks can be employed to defeat this vulnerability of dyadic data in the realm of international relations.
+These approaches have been developed for a while, but are not yet widely used in international relations scholarship.
+The statistical model of the latent network captures first-order (example or restatement), second-order (example or restatement), and third-order (example or restatement) dependencies in dyadic data using a familiar regression framework that has been adapted for relational data--such as dyadic data--which are not independent nor identically distributed.
+There is an available, open source computer packages that implements this approach.
+
+To explore this approach in the context of international relations we conduct two broad analyses.
+The first is a simulation of XCLKFJDSLF.
+The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
+These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
+We then reanalyzed each study using the latent network approach which captures that additive and multiplicative aspects of interdependencies among the dyadic data.
+In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.
+In several cases, the new approach overturns the basic findings of the original research.
+This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they only obtain under stringent assumptions that can not possibly be valid.
+This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
+At the same time, the latent factor model we present and demonstrate herein provides a way to easily examine and if necessary defeat these assumptions.
+
+It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.
+Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citep{king:roberts:2014}).
 Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.

### 2017-10-17 paper/1_intro.tex
front half edits

@@
-% need to do a better job of differentiating this from piece with peter
-%  here we are adding a simulation exercise to show that ame works
-% doing a big replication across five ir papers
-% we're not "introducing" ame
 \section{\textbf{Introduction}}
@@
 Namely, it is not just the diagonals of the variance-covariance matrix this affects.
+% \citep{beck:2012, king:roberts:2014}
 
@@
 First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict.
-Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad, specifically, if a rebel group initiates a conflict with a government, the government will reciprocate that conflictual behavior.
+Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; specifically, if a rebel group initiates a conflict with a government, the government will likely reciprocate that conflictual behavior.
 We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-The multiplicative effects portion of the model captures dependencies that results from groups of actors clustering together or organizing into communities due to meso-scopic features of networks, such as homophily and stochastic equivalence.
-%Clusters and communities often result in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.
+The multiplicative effects capture dependencies that results from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence.
+These type of meso-scopic features often arise in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.
 
-We begin by conducting a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
+We begin by providing a brief review of these dependencies and the AME model.
+Next, we move to conducting a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
 Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.

### 2017-10-17 paper/2_model.tex
front half edits

@@
 
-In working with dyadic data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another.
+In working with relational data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another.
 This makes sense if each observation is independent of the others.
-Thus, for example, a conflict sent from actor the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
+Thus, for example, a conflict sent from the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
 Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
 As a result, the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others.
-We know this assumption to be false in theory and in practice.
+We know this assumption to be false in theory and in practice because relational data comes with an explicit structure that in general leads to particular types of dependencies.
+The importance of accounting for underlying structure of our data has been a lesson well understood at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
+As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
 
-\input{dd_to_adjmat}
-\FloatBarrier
-
-To start to move away from this assumption and better understand the dependencies that emerge in relational data, it is helpful to shift towards structuring dyadic data in the form of a matrix--often called an adjacency matrix in the context of spatial analysis--shown in the top right of Table~\ref{tab:netDesign}.
+To uncover the underlying structure that relational data comes with it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often called an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 Using the structure of an adjacency matrix, we can visualize the types of first- and second-order dependencies that complicate the analysis of relational data in traditional GLMs.
-Figure~\ref{fig:adjMatDeps}clarifies the types of dependencies that can manifest in these types of data structures.
-The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that values across a particular row of an adjacency matrix may be more similar to each other than other values in the adjacency matrix because each of these values has a common sender.
+Figure~\ref{fig:adjMatDeps} clarifies the types of dependencies that can manifest in these types of data structures.
+The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that these values may be more similar to each other than other values in the adjacency matrix because each has a common sender.
 Homogeneity in interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
@@
 		\texttt{Sender heterogeneity} & \texttt{Receiver Heterogeneity} \\
-		\includegraphics[width=.45\textwidth]{adjRowDep.png} & \includegraphics[width=.45\textwidth]{adjColDep.png} \\
+		\includegraphics[width=.4\textwidth]{adjRowDep.png} & \includegraphics[width=.4\textwidth]{adjColDep.png} \\
 		\texttt{Sender-Receiver Covariance} & \texttt{Reciprocity} \\
-		\includegraphics[width=.45\textwidth]{adjRowColCovar.png} & \includegraphics[width=.45\textwidth]{adjRecip.png} \\
+		\includegraphics[width=.4\textwidth]{adjRowColCovar.png} & \includegraphics[width=.4\textwidth]{adjRecip.png} \\
 	\end{tabular}
@@
 
-For similar reasons one also needs to take into account that there is a shared dependence between dyadic observations that share a common receiver.
+For similar reasons one also needs to take into account that there is a shared dependence between observations that share a common receiver.
 The bottom-left panel, illustrates that sender and receiver type dependencies can also blend together.
-Specifically, as actors who are more likely to send ties in a network tend to also be more likely to receive them.
+Specifically, actors who are more likely to send ties in a network tend to also be more likely to receive them.
 As a result, the rows and columns in an adjacency matrix are often correlated.
@@
 The concept of reciprocity has deep roots in the study of relations between states \citep{richardson:1960,keohane:1989}.
-The purpose of highlighting each of these dependencies through the set of panels in Figure~\ref{fig:adjMatDeps} is to more easily visualize how each of these dependencies manifest in relational data.
-Alternatively, when we simply simply stack dyads in some arbitrary order, these dependencies are easy to ignore and virtually impossible to recognize.
 
-The dependencies discussed so far involve at most dependence between two actors.
-For most relational data, however, dependencies do not simply manifest at this level.
+For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level.
 More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors \citep{poast:2010}.
-These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{wasserman:faust:1994,zinnes:1967}.
+These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
 In Figure~\ref{fig:thirdDeps} we provide a visualization of a hypothetical relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
-Each node is colored by the latent group to which they belong.
+To highlight third order dependence patterns, nodes with similar latent attributes are colored similarly.
 
 \begin{figure}[ht]
-	\includegraphics[width=.7\textwidth]{stochEquiv.pdf}
+	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
 	\caption{Visualization of network with meso-scopic features.}
@@
 
-The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other; interactions across groups are more rare.
-One prominent example of a network with this type of structure was found in the context of the polarization in US politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another in the 2004 United States Election \citep{adamic:glance:2005}.
-They found that the degree of interaction between right and left leaning blogs was minimal; most of these blogs linked to blogs that were similar, not that were different.
+The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are more rare.
+A prominent example of a network with this type of structure was found in American Politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another in the 2004 United States Election.
+\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs just linked to others that were political similar.
 This showcases the types of higher-order dependencies that can emerge in relational data.
@@
 
-The major implication of the presence of homophily, stochastic equivalence, and the other dependencies we discussed above is that they complicate the practical assumption of observational independence.
-Inferences drawn from models that ignore potential interdependencies between dyadic observations face a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-By ignoring these potential interdependencies, we often ignore important features of the problem under study.
+The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
+Of course, if one is able to specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
+However, it is rarely the case that we can do this even for TSCS data, thus we more than often make certain modeling decisions to account for that structure.
+Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
+Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem under study.
 The study of international relations is founded on the relations among actors.
@@
 
-% Thus, unless we specify a list of exogenous variables that determine this prevalence of triads, the probability of $j$ and $k$ forming a tie is not independent of the ties that already exist between those actors and others.
-
 \section{\textbf{Additive and Multiplicative Effect Models for Networks}}
 
+To account for the dependencies that are prevalent in dyadic data, we turn to the AME model.
 The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
@@
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-We do not know whether interdependence dominates international relations, but as noted by Thoreau in reference to rumors that dairymen on strike were watering down milk: ``\ldots some circumstantial evidence is very strong, as when you find a trout in the milk.'' At a very minimum, it is necessary to examine \textit{whether} there is interdependence since it challenges both substantive arguments and statistical modeling \citep{snijders:2011}.
+\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece details how it contrasts with alternative network based approaches.
+Here we just provide a brief review and then move onto presenting a simulation exercise.
 
-Further parameter interpretation in the AME framework is straightforward because it has a simple regression basis as shown in Equation~\ref{eqn:ame} (shown for a directed matrix, without loss of generality) \:
+The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model for third order dependencies.\footnote{For details on the SRRM see: for details on this model see \citet{li:loken:2002,dorff:minhas:2016}.
+An earlier version of the latent factor model (LFM) used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
+The GBME model is more limited in the types of dependence patterns that it can capture due to differences in the type of matrix decomposition approach utilized.} The AME model is specified as follows:
 
@@
 	\begin{aligned}
-	y_{ij} &= g(\theta_{ij}) \\
-	&\theta_{ij} = \bm\beta^{\top} \mathbf{X}_{ij} + e_{ij} \\
-	&e_{ij} = a_{i} + b_{j}  + \epsilon_{ij} + \alpha(\textbf{u}_{i}, \textbf{v}_{j}) \text{  , where } \\
-	&\qquad \alpha(\textbf{u}_{i}, \textbf{v}_{j}) = \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} = \sum_{k \in K} d_{k} u_{ik} v_{jk}. \\
+		y_{ij} = f(\theta_{ij}) &\text{, where } \\
+		\theta_{ij} =& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} + a_{i} + b_{j}  + \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} + \epsilon_{ij} \\
 	\label{eqn:ame}
@@
 
-With this framework, it is straightforward to model dyadic observations as conditionally independent given $\bm\theta$, where $\bm\theta$ depends on the the unobserved random effects modeled to account for the potential \first, \second, and \third-order dependencies.
-In Equation~\ref{eqn:srmCov}, $a_{i} + b_{j}  + \epsilon_{ij}$ represent the additive random effects in this framework and account for sender, receiver, and within-dyad dependence.
-The multiplicative effects, $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}$, capture higher-order dependence patterns in $\bm\theta$ after accounting for any known covariate information.\footnote{This procedure is described in detail in \citet{hoff:2008,hoff:2015:arxiv,minhas:etal:2016:arxiv}.
-A Bayesian procedure in which parameters are iteratively updated using a Gibbs sampler is described in the appendix.}
+where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver).
+We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$); and then parameters from the SRRM and LFM.
+$a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
+
+\begin{align}
+	\begin{aligned}
+		\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\ 
+		\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
+		\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
+	\label{eqn:srm}
+	\end{aligned}
+\end{align}
+
+The sender and receiver random effects are modeled jointly from a multivariate normal distribution to account for correlation in how active an actor is in sending and receiving ties.
+Heterogeneity in the the sender and receiver effects is captured by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively, and $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties).
+Beyond these first-order dependencies, second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a within dyad correlation, or reciprocity, parameter $\rho$.
+
+The LFM contribution to the AME comes in the multiplicative term: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$.
+This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
+The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third order interdependencies \citep{minhas:etal:2016:arxiv}.
+The representation of third order interdepedencies is accomplished by a process similar to computing the singular value decomposition (SVD) of the observed network.
+When taking the SVD we factorize our observed network into the product of three matrices: $\textbf{U}, \textbf{D}, \text{ and }, \textbf{V}$.
+This provides us with a low-dimensional representation of our original network in terms of third order dependence patterns arising in the network.\footnote{The dimensions of $\textbf{U}$ and $\textbf{V}$ are $n \times K$ and $\textbf{D}$ is a $K \times K$ diagonal matrix.} Specifically, values in $\textbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network, or more simply put how similar actors are in terms of who they are initiating, for example, conflict with.
+$\hat{\textbf{u}}_{i} \approx \hat{\textbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors.
+$\textbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
+The values in $\textbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.} Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
+
+ Specifically, given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
+
+ \begin{itemize}
+ 	\item sample $\bm\theta \; | \;  \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 	\item sample $\bm\beta \; | \;  \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 	\item sample $\textbf{a}, \textbf{b} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+	\item sample $\Sigma_{ab} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
+ 	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
+ 	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
+ 	\item For each $k \in K$:
+ 	\begin{itemize}
+ 		\item Sample $\textbf{U}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}_{[,-k]}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 		\item Sample $\textbf{V}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 		\item Sample $\textbf{D}_{[k,k]}  \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \textbf{D} matrix is absorbed into the calculation for $\textbf{V}$ as we iterate through $K$.
+}
+ 	\end{itemize}
+ \end{itemize}

### 2017-10-17 paper/3_sim.tex
so much to delete
front half edits

@@
 \section{Simulation study}
+
+We use a simulation study to highlight the ability of AME to conduct inference on exogenous parameters in the presence of network dependencies.
+We are particularly interested in whether the LSM can adjust for confounding when the confounding variable is unmeasured.
+In the simulation design, we study how the level of confounding affects the performance of AME relative to the GLM.
+We evaluate performance in terms of bias and inferential error.
+For simplicity and computational efficiency we consider a unidimensional latent space.
 
@@
 	\item A latent factor model; 
-	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
+	\item An ``oracle'' regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
 \end{itemize}

### 2017-10-17 paper/6_appendix.tex
front half edits

@@
 
-\subsection*{Additive and Multiplicative Effects Gibbs Sampler}
+% \subsection*{Additive and Multiplicative Effects Gibbs Sampler}
 
-To estimate, the effects of our exogenous variables and latent attributes we utilize a Bayesian probit model in which we sample from the posterior distribution of the full conditionals until convergence.
-Specifically, given observed data $\textbf{Y}$ and $\textbf{X}$ -- where $\textbf{X}$ is a design array that includes our sender, receiver, and dyadic covariates -- we estimate our network of binary ties using a probit framework where: $y_{ij,t} = 1(\theta_{ij,t}>0)$ and $\theta_{ij,t} = \bm\beta^{\top}\mathbf{X}_{ij,t} + a_{i} + b_{j} + \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} + \epsilon_{ij}$.
-The derivation of the full conditionals is described in detail in \citet{hoff:2005} and \citet{hoff:2008}, thus here we only outline the Markov chain Monte Carlo (MCMC) algorithm for the AME model that we utilize in this paper.
+% To estimate, the effects of our exogenous variables and latent attributes we utilize a Bayesian probit model in which we sample from the posterior distribution of the full conditionals until convergence.
+Specifically, given observed data $\textbf{Y}$ and $\textbf{X}$ -- where $\textbf{X}$ is a design array that includes our sender, receiver, and dyadic covariates -- we estimate our network of binary ties using a probit framework where: $y_{ij} = 1(\theta_{ij}>0)$ and $\theta_{ij} = \bm\beta^{\top}\mathbf{X}_{ij} + a_{i} + b_{j} + \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} + \epsilon_{ij}$.
 
-\begin{itemize}
- \item Given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows:
- \begin{itemize}
- 	\item sample $\bm\theta \; | \;  \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 	\item sample $\bm\beta \; | \;  \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 	\item sample $\textbf{a}, \textbf{b} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-	\item sample $\Sigma_{ab} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
- 	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
- 	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
- 	\item For each $k \in K$:
- 	\begin{itemize}
- 		\item Sample $\textbf{U}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}_{[,-k]}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\textbf{V}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\textbf{D}_{[k,k]}  \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \textbf{D} matrix is absorbed into the calculation for $\textbf{V}$ as we iterate through $K$.
+% \begin{itemize}
+%  \item Given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows:
+%  \begin{itemize}
+%  	\item sample $\bm\theta \; | \;  \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+%  	\item sample $\bm\beta \; | \;  \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+%  	\item sample $\textbf{a}, \textbf{b} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+% 	\item sample $\Sigma_{ab} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
+%  	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
+%  	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
+%  	\item For each $k \in K$:
+%  	\begin{itemize}
+%  		\item Sample $\textbf{U}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}_{[,-k]}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+%  		\item Sample $\textbf{V}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+%  		\item Sample $\textbf{D}_{[k,k]}  \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \textbf{D} matrix is absorbed into the calculation for $\textbf{V}$ as we iterate through $K$.
 }
- 	\end{itemize}
- \end{itemize}
-\end{itemize}
+%  	\end{itemize}
+%  \end{itemize}
+% \end{itemize}

### 2017-10-17 paper/netsMatter.tex
front half edits

@@
       % \previewfalse
+
 % approx iid
-\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}\usepackage{natbib}
+\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}
+\usepackage{natbib}
+\usepackage{titlecaps}
+\Addlcwords{a about above after again against all am an and any are aren't as at be because been before being below between both but by can't cannot could couldn't did didn't do does doesn't doing don't down during each few for from further had hadn't has hasn't have haven't having he he'd he'll he's her here here's hers herself him himself his how how's i i'd i'll i'm i've if in into is isn't it it's its itself let's me more most mustn't my myself no nor not of off on once only or other ought our ours ourselves out over own same shan't she she'd she'll she's should shouldn't so some such than that that's the their theirs them themselves then there there's these they they'd they'll they're they've this those through to too under until up very was wasn't we we'd we'll we're we've were weren't what what's when when's where where's which while who who's whom why why's with won't would wouldn't you you'd you'll you're you've your yours yourself yourselves}
 
@@
 
-%%What about a summary table with AUCs for the different network models here? - MG
 \newpage
@@
 \clearpage
-% \bibliography{master.bib}
 \bibliography{/Users/s7m/whistle/master}
-\bibliographystyle{authordate1}
+% \bibliographystyle{authordate1}
+\bibliographystyle{sm}
 % \clearpage

### 2017-10-17 paper/2_model.tex
sim stuff

@@
 
-To uncover the underlying structure that relational data comes with it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often called an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure that relational data comes with it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often referred to as an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 
-Using the structure of an adjacency matrix, we can visualize the types of first- and second-order dependencies that complicate the analysis of relational data in traditional GLMs.
-Figure~\ref{fig:adjMatDeps} clarifies the types of dependencies that can manifest in these types of data structures.
+Using the structure of an adjacency matrix, we visualize in Figure~\ref{fig:adjMatDeps} the types of first- and second-order dependencies that complicate the analysis of relational data in traditional GLMs.
 The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that these values may be more similar to each other than other values in the adjacency matrix because each has a common sender.
@@
 $\textbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\textbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.} Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
+The values in $\textbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
 
+Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
 Specifically, given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:

### 2017-10-17 paper/3_sim.tex
sim stuff

--- 2017-09-29 paper/3_sim.tex
@@
 
-Here we consider inference for a regression parameter $\beta$ of a linear or generalized linear model for a network in the case where there is an omitted variable.
-The true data-generating models we consider are of the form
+We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
+We are particularly interested in whether AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved network dependencies.
+Specifically, we are concerned with conducting inference on a regression parameter $\beta$ of a linear model for a network in the case where there is an omitted variable.
+Say that the true data-generating process for a particular $Y$ is given by:
 
-$y_{i,j} \sim  \mu + \beta x_i x_j + \gamma w_i w_j + \epsilon_{i,j} $
+$y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j} $
 
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_i \} \in \mathbb R^n$ is a vector of observed node-specific characteristics, and $w = \{ w_i\} \in \mathbb R^n$ is a vector of unobserved node-specific characteristics.
+where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $w = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
 We compare inference for $\beta$ using three models:
@@
 	\item A latent factor model; 
-	\item An "oracle" regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
+	\item An ``oracle'' regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
 \end{itemize}
 
-We make these comparisons in the context of a binary network outcome with a probit regression model.
-
-One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
-
-Alternatively, bias and precision can be summarized by confidence interval coverage and width.
-Coverage should ideally be at the nominal level.
-If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
+We make these comparisons in the context of a linear network outcome with the AME probit regression model described in the previous section.
 
@@
 
+One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
+
 \begin{figure}
@@
 
+Alternatively, bias and precision can be summarized by confidence interval coverage and width.
+Coverage should ideally be at the nominal level.
+If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
+
 \begin{figure}

### 2017-10-17 paper/3_sim.tex
notes

@@
 We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
-We are particularly interested in whether AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved network dependencies.
-Specifically, we are concerned with conducting inference on a regression parameter $\beta$ of a linear model for a network in the case where there is an omitted variable.
+The goal of the simulation is to assess how well AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved network dependencies.
+[link to previous discussion] Specifically, we are concerned with conducting inference on a regression parameter $\beta$ of a linear model for a network in the case where there is an omitted variable.
 Say that the true data-generating process for a particular $Y$ is given by:
 
-$y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j} $
+\begin{align*}
+	y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j} \\
+	\label{eqn:sim}
+\end{align*}
 
@@
 \begin{itemize}
-	\item A naive regression model assuming independent errors; 
-	\item A latent factor model; 
-	\item An ``oracle'' regression model that includes both $x_{i,j}=x_i x_j$ and $w_{i,j}=w_i w_j$. 
+	\item the standard IR approach assuming independent errors; 
+	\item the AME approach outlined in the previous section;
+	\item and an ``oracle'' regression model that includes both $x_{i,j}$ and $w_{i,j}$. 
 \end{itemize}
 
-We make these comparisons in the context of a linear network outcome with the AME probit regression model described in the previous section.
+We run this simulation
 

### 2017-10-17 paper/2_model.tex
more hand holding in model

@@
 
-The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model for third order dependencies.\footnote{For details on the SRRM see: for details on this model see \citet{li:loken:2002,dorff:minhas:2016}.
-An earlier version of the latent factor model (LFM) used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
-The GBME model is more limited in the types of dependence patterns that it can capture due to differences in the type of matrix decomposition approach utilized.} The AME model is specified as follows:
+The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third order dependencies.\footnote{For details on the SRRM see: for details on this model see \citet{li:loken:2002,dorff:minhas:2016}.
+An earlier version of the LFM used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
+The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.} The AME model is specified as follows:
 
@@
 	\begin{aligned}
-		y_{ij} = f(\theta_{ij}) &\text{, where } \\
-		\theta_{ij} =& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} + a_{i} + b_{j}  + \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} + \epsilon_{ij} \\
+		y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
+		\theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} + \text{\qquad(Exogenous parameters)} \\
+		& a_{i} + b_{j} + \epsilon_{ij} + \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
+		& \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
 	\label{eqn:ame}
@@
 where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver).
-We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$); and then parameters from the SRRM and LFM.
-$a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
+We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$);
+
+this is where most people end.
+
+and then parameters from the SRRM and LFM. $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
 
@@
 The LFM contribution to the AME comes in the multiplicative term: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$.
+$K$ denotes the dimensions of the latent space.
 This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
 The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third order interdependencies \citep{minhas:etal:2016:arxiv}.
-The representation of third order interdepedencies is accomplished by a process similar to computing the singular value decomposition (SVD) of the observed network.
+The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network.
 When taking the SVD we factorize our observed network into the product of three matrices: $\textbf{U}, \textbf{D}, \text{ and }, \textbf{V}$.
-This provides us with a low-dimensional representation of our original network in terms of third order dependence patterns arising in the network.\footnote{The dimensions of $\textbf{U}$ and $\textbf{V}$ are $n \times K$ and $\textbf{D}$ is a $K \times K$ diagonal matrix.} Specifically, values in $\textbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network, or more simply put how similar actors are in terms of who they are initiating, for example, conflict with.
+This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\textbf{U}$ and $\textbf{V}$ are $n \times K$ and $\textbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\textbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network, or more simply put how similar actors are in terms of who they are initiating, for example, conflict with.
 $\hat{\textbf{u}}_{i} \approx \hat{\textbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors.
@@
 The values in $\textbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
+
+By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we draw
 

### 2017-10-18 paper/4_app.tex
front half edits
starting to rework the app section.

@@
 
-For the purpose of this study, we choose five prominent studies from the broad field of international relations and international political economy that utilize relation data \citep{mcdonald:2004, reiter:stam:2003, rose:2004, weeks:2012, gibler:2017}.
-We chose to replicate studies that were written fairly recently, since 2000, and that have been cited over 100 times.
+We choose five prominent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+These studies are recent and have been cited over 100 times.
 Each of these pieces was published in a prominent journal and is well-known in the literature.
-They all used the standard approach in political science, which is to employ some form of general linearized regression that ignores dyadic interdependencies.
-Post estimation, standard errors are often adjusted in an attempt to account for clustering of observations.
+Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
+Also part of the gold standard is a post estimation cleanup thought to be a paliative, and each of these studies adjust the posterior standard errors in an attempt to account for the clustering of observations.
 
@@
 	\begin{tabular}{lcccccc}
-		 & Model & \# of Actors & Years  & \# of Dyads & Type of Dyads & Clustered $\sigma_{\hat{\beta}}$ \\ \hline\hline
-		Weeks (2012) & Logit & & & $901,540$ & Directed & Robust \\
-		Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\
-		McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\		
-		Gibler (2017) & Logit & & & $650,557$ & UnDirected & no \\		
-		Rose (2004) & OLS & & & $234,597$ & Directed & Robust \\
-	\hline\hline
+& Model & N. Actors & Years  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
+Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\	McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\
+Rose (2004) & OLS & ? &  ?& $234,597$ & Directed & Robust \\	 
+Weeks (2012) & Logit & ? & ?& $901,540$ & Directed & Robust \\
+Gibler (2017) & Logit & ?& ?& $650,557$ & UnDirected & no \\ \bottomrule
 	\end{tabular}
@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We examine each of the models using the AME framework described above.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.
 Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
 
-Finally, we assess whether there is any substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
+Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
 
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from empirical studies over the past decade or so.
+The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies that use a framework that ignores dependencies.
 We believe that it does, and that the dynamic latent factor model provides a step forward.
+
+WE NEED TO TALK ABOUT THIS TABLE AND WHAT THE CRITERIA WERE FOR THE FINAL COLUMN.
 
@@
 \centering
-	\begin{tabular}{l| p{7cm} l}
-	& \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
-	& &  in a Network Model? \\
-	\hline\hline
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\
-		\hline
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ 
-		\hline
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ 
-		\hline
-		Gibler (2017) & Power Parity at time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ 		
-		\hline
-		Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ 
-	\hline\hline
+\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
+Cases in which the substantive finding is not replicated are highlighted in bold.}
+\begin{tabular}{l p{7cm} l} \toprule
+\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
+	& &  in a Network Model? \\ \toprule
+Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ \midrule
+McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ \midrule
+Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ \midrule
+Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\\midrule
+Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ \bottomrule
 	\end{tabular}
-	\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
-Cases in which the finding is not replicated are highlighted in bold.}
 			\label{tab:modelFindingSumm}
@@
 
-We also examine the accuracy of the predictions made with each approach. Out of sample cross validation strategy.
+We also examine the accuracy of the predictions made with each approach. Out-of-sample cross validation strategy.
 
@@
 By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
-To show that this is the case, we examine whether our approach achieves better predictive performance in an out of sample context than traditional dyadic models.
+To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models.
 To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
@@
 We set a number of benchmarks for comparison.
-First we compare the AME model to a GLM model using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
+First, we compare the AME model to a GLM model (that used in the original study) using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
 We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
@@
 If peace is common, even a poor model will have a very low False Positive Rate.
-
+The ROC curves are best for balanced samples with as many events as non-events.
 To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
@@
 	~ & ~ & GLM & AME \\
-	\hline\hline
+	\toprule
+\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
+		~ & AUC (PR) & 0.08 & \textbf{0.15} \\		\midrule
+\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
+		~ & AUC (PR) & 0.13 & \textbf{0.28} \\		\midrule
+\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
+		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\	\midrule
 \multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
-		~ & AUC (PR) & 0.08 & \textbf{0.15} \\
-		\hline
-		\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
-		~ & AUC (PR) & 0.13 & \textbf{0.28} \\
-		\hline
+		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\		\midrule
 \multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
-		~ & AUC (PR) & 0.00 & \textbf{0.08} \\
-		\hline
-		\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
-		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\
-	\hline\hline
+		~ & AUC (PR) & 0.00 & \textbf{0.08} \\			\bottomrule
 	\end{tabular}
@@
 
+%%% TO HERE
 \subsection{Replication of Weeks (2012)}
@@
 \end{figure}
+\FloatBarrier

### 2017-10-19 paper/2_model.tex
sync up

@@
 Rows designate the senders of an event and columns the receivers.
-In the context of events, the cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns.
+The cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns.
 Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flows, that is sent from actor $i$ to actor $j$.
 
-Using the structure of an adjacency matrix, we visualize in Figure~\ref{fig:adjMatDeps} the types of first- and second-order dependencies that complicate the analysis of relational data in traditional GLMs.
+Using the structure of an adjacency matrix, we visualize in Figure~\ref{fig:adjMatDeps} the types of first- and second-order dependencies that can complicate the analysis of relational data in traditional GLMs.
 The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that these values may be more similar to each other than other values in the adjacency matrix because each has a common sender.
@@
 where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver).
-We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$);
-
-this is where most people end.
-
-and then parameters from the SRRM and LFM. $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
+We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
+Next, to account for the dependencies that emerge in dyadic data and that may complicate inference on the parameter associated with exogenous covariates we add parameters from the SRRM and LFM.
+$a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
 

### 2017-10-19 paper/4_app.tex
working it

@@
 Weeks (2012) & Logit & ? & ?& $901,540$ & Directed & Robust \\
-Gibler (2017) & Logit & ?& ?& $650,557$ & UnDirected & no \\ \bottomrule
+Gibler (2017) & Logit & 21,275 & 193 & $650,557$ & UnDirected & no \\ \bottomrule
 	\end{tabular}
@@
 
-WE NEED TO TALK ABOUT THIS TABLE AND WHAT THE CRITERIA WERE FOR THE FINAL COLUMN.
+Given current practices, most scholars have a single variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models.
+Whether putative control variables, for example, are precisely estimated is less important.
+Generally, interpretations focus on a small set of independent variables.
+In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads--consisting of a democratic country and country ruled by a personalist dictatorship--whether it is the
+democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
+In a regression including fourteen coefficients, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our evaluation we focus on the same question in the re-examination of the Reiter and Stam study to determine whether the model replicates when reestimated within the AME framework.
+Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textbf{fails to replicate} indicates only that the sign and/or significance of the crucial finding in the original study is not also found in the AME estimation.
 
@@
 
-We also examine the accuracy of the predictions made with each approach. Out-of-sample cross validation strategy.
-
-% % % % % % % % % % % % % % % % % % 
+In addition, we also examine the predictions made with each approach using an out-of-sample cross validation strategy.
 By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
@@
 Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
-The fifth involved a gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
+The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
 Cases in which our network based approach outperformed the standard approach are highlighted in bold.}

### 2017-10-19 paper/4_app.tex
finished with Rose

@@
 Generally, interpretations focus on a small set of independent variables.
-In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads--consisting of a democratic country and country ruled by a personalist dictatorship--whether it is the
+In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---whether it is the
 democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
@@
 The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
-In Weeks's initial models, as detailed in figure \ref{fig:weeksCoef} she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
+In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
 When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
@@
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in figure \ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
 In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
@@
 The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in figure \ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
+In terms of out of sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
 
@@
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework, as seen in Figure~\ref{fig:reitCoef}.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitauc} and Figure ~\ref{fig:reitpr}).
+Our replication using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
+Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
+Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitaucpr} and Figure ~\ref{fig:reitpr}).
 
@@
 
-\newpage
 \subsection{Replication of Rose (2004)}
@@
 
-When we compare the results of Rose's original OLS model to our model that accounts for network dependencies, the results are generally similar.
-As you can see in Figure~\ref{fig:roseCoef}, the main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- remains when we move from an OLS to a Gaussian AME model.
-The most striking difference between the models is that, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
+When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
+The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---persists in the AME model.
+However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
 The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
@@
 Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-Excitingly, when we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out of sample -- 3.23 for the OLS model and 1.77 for the AME model.
+When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample---$3.23$ for the OLS model and $1.77$ for the AME model.
+While the basic model replicates with the basic result, there are substantial and substantive improvements gleaned from the AME results.
 
-\begin{figure}
-\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
- \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
+% \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
+%\end{figure}
 

### 2017-10-19 paper/4_app.tex
old McDonald had a typo

@@
 \citet{mcdonald:2004} studies whether trade promotes peace between nations.
-He observes that knowledge about the link between conflict and trade is indeterminate in the field of international relations, noting that competing explanations persist.
-He calls for more precise empirical tests.
-Importantly, \citet{mcdonald:2004} includes the interdependence argument---that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)'' in his overview of underdeveloped hypotheses.
-
-Accordingly, the primary contribution of the study is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-To do so, \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
+The link between conflict and trade is perplexing, with many presistent, yet competing explanations.
+\citet{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)''.
+Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
+\citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
 His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
@@
 This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state i's GDP or total dyadic trade divided by state j's GDP.'' (560).
-Finally, the binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
+\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP'' (560).
+The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
 \citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include a spline correction for time-series data as well as robust standard errors clustered on each dyad.
+The models include splines to correct for temporal dependence, and robust standard errors clustered on each dyad.
 
@@
 
-\begin{figure}
-\centering   
-\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-\end{figure}
+%\begin{figure}
+%\centering   
+%\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
+%\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
+%\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
+%\end{figure}
 
@@
 Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries {i} and {j} are highly dependent and countries {j} and {k} are also highly dependent, then we are likely to observe high dependency between countries {i} and {k}.
+Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
 Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
@@
 Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and replicated model and Figure~\ref{fig:mcdall} demonstrates the predictive performance of each model.
-Figure ~\ref{fig:mcdcirc} represents the multiplicative effects of the model.
-This plot enables us to consider actors with similar sending and receiving patterns resulting from stochastic equivalence and homophily.
-In the upper arch for senders, there is a clear clustering of United States' allies (the United Kingdom, Netherlands, and Japan).
-We also observe an African cluster in the top right segment for receivers (Egypt, Niger, and Zambia for example).
-These countries both trade amongst themselves and have similar conflict behavior -- stochastic equivalence is present among the first group, homophily among the second -- and so once we account for these clusters using the multiplicative effects, the residual effect of trade actually increases the likelihood of conflict.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{mcdonaldcirc.pdf}
-\caption{\label{fig:mcdcirc} Visualization of multiplicative effects. Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
+Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and AME model.
+The most striking thing is that AME finds a positive---not negative as found by McDonald---association between trade dependence and conflict.
+The AME coefficient is not only in the opposite direction, but it is approximately three times more precisely estimated.
+At the same time, the AME has ROC and PR curves (not shown) that dominate the OLS results found in McDonald (2004).
 

### 2017-10-19 paper/4_app.tex
working it

@@
 	\begin{tabular}{lcccccc}
-& Model & N. Actors & Years  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
-Reiter \& Stam (2003) & Logit & & & $753,456$ & Directed & Robust \\	McDonald (2004) & Logit & & & $92,354$ & Undirected & Robust\\
-Rose (2004) & OLS & ? &  ?& $234,597$ & Directed & Robust \\	 
-Weeks (2012) & Logit & ? & ?& $901,540$ & Directed & Robust \\
-Gibler (2017) & Logit & 21,275 & 193 & $650,557$ & UnDirected & no \\ \bottomrule
+& Model &  N. Actors  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
+Reiter \& Stam (2003) & Logit & 193 & $753,456$ & Directed & Robust \\	
+McDonald (2004) & Logit & 198 & $92,354$ & Undirected & Robust\\
+Rose (2004) & OLS & 177 & $234,597$ & Directed & Robust \\	 
+Weeks (2012) & Logit & 197 &  $901,540$ & Directed & Robust \\
+Gibler (2017) & Logit & 193 &    & $650,557$ & UnDirected & none \\ \bottomrule
 	\end{tabular}

### 2017-10-19 paper/4_app.tex
fast getun

@@
-\section{Replications}
+\section{Re-estimation with AME}
 
@@
 \begin{table}
-\caption{Features of the Studies Replicated. }
+\caption{Features of the Studies Re-estimated. }
 	\begin{tabular}{lcccccc}
@@
 
-%%% TO HERE
-\subsection{Replication of Weeks (2012)}
-
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
-One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-Her time frame is from 1946-1999.
-The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
-She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
-
-The replication of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
-In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
-When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
-We similarly find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
-In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
-In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
-So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
-The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
-\end{figure}
-
-\begin{figure}
-\centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
 \subsection{Replication of Reiter \& Stam (2003)}
@@
 In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
-In our replication using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
+In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our replication using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
+Our re-extimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
@@
 \end{figure}
+
+\floatbarrier
+\subsection{Replication of McDonald (2004)}
+
+\citet{mcdonald:2004} studies whether trade promotes peace between nations.
+The link between conflict and trade is perplexing, with many presistent, yet competing explanations.
+\citet{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)''.
+Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
+\citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
+His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
+\citet{mcdonald:2004} measures free trade in two ways.
+The first captures the idea that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
+This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
+This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
+\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP'' (560).
+The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
+\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
+The models include splines to correct for temporal dependence, and robust standard errors clustered on each dyad.
+
+% TO DO: include discussion and analysis of different K's in appendix (for M cD)
+\begin{figure}
+\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
+\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
+\end{figure}
+
+%\begin{figure}
+%\centering   
+%\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
+%\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
+%\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
+%\end{figure}
+
+Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
+Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
+Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
+This indicates that conflict is less likely between members of a trade community.
+Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
+Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and AME model.
+The most striking thing is that AME finds a positive---not negative as found by McDonald---association between trade dependence and conflict.
+The AME coefficient is not only in the opposite direction, but it is approximately three times more precisely estimated.
+At the same time, the AME has ROC and PR curves (not shown) that dominate the OLS results found in McDonald (2004).
 
@@
 
-\newpage
-\subsection{Replication of McDonald (2004)}
+\subsection{Replication of Weeks (2012)}
 
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-The link between conflict and trade is perplexing, with many presistent, yet competing explanations.
-\citet{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)''.
-Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-\citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measures free trade in two ways.
-The first captures the idea that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
-This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP'' (560).
-The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include splines to correct for temporal dependence, and robust standard errors clustered on each dyad.
+\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
+She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
+She adds the nuance that autocratic audiences are not homogeneous.
+When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
+Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
+To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
+These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
+One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
+She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects.
+The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
+Her time frame is from 1946-1999.
+The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
+These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
+She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
 
-% TO DO: include discussion and analysis of different K's in appendix (for M cD)
+The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
+In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
+When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
+We similarly find less pronounced effects for military capabilities.
+One explanation for this divergence is the AME model's ability to account for third order effects.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
+In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
+In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
+So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
+The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
+In terms of out of sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
+
 \begin{figure}
-\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
+	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
+	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
 \end{figure}
 
-%\begin{figure}
-%\centering   
-%\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-%\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-%\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-%\end{figure}
+\begin{figure}
+\centering   
+	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
+\end{figure}
 
-% TO DO: fix bib
-
-Our replication reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
-Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
-This indicates that conflict is less likely between members of a trade community.
-Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and AME model.
-The most striking thing is that AME finds a positive---not negative as found by McDonald---association between trade dependence and conflict.
-The AME coefficient is not only in the opposite direction, but it is approximately three times more precisely estimated.
-At the same time, the AME has ROC and PR curves (not shown) that dominate the OLS results found in McDonald (2004).
+\begin{figure}
+	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
+	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
+\end{figure}
 
@@
 \FloatBarrier
+
+\subsection{Lessons Learned from Re-estimating Five Prominent Studies}

### 2017-10-19 paper/4_app.tex
checking compliation
noch mal ins feur.

@@
 	\begin{tabular}{lcccccc}
-& Model &  N. Actors  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
-Reiter \& Stam (2003) & Logit & 193 & $753,456$ & Directed & Robust \\	
-McDonald (2004) & Logit & 198 & $92,354$ & Undirected & Robust\\
-Rose (2004) & OLS & 177 & $234,597$ & Directed & Robust \\	 
-Weeks (2012) & Logit & 197 &  $901,540$ & Directed & Robust \\
-Gibler (2017) & Logit & 193 &    & $650,557$ & UnDirected & none \\ \bottomrule
+& Model &  Date Range & N. Actors  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
+Reiter \& Stam (2003) &Logit &1945--1995 &  193 & $753,456$ & Directed & Robust \\	
+McDonald (2004) & Logit &1959--2002 & 198 & $92,354$ & Undirected & Robust\\
+Rose (2004) & OLS & 1948--1999 & 177 & $234,597$ & Directed & Robust \\	 
+Weeks (2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
+Gibler (2017) & Logit & 1816--2008 &193 &   $650,557$ & UnDirected & none \\ \bottomrule
 	\end{tabular}
@@
 \begin{tabular}{l p{7cm} l} \toprule
-\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} & Does it Replicate \\ 
+\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
 	& &  in a Network Model? \\ \toprule
-Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & Replicates \\ \midrule
-McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & \textbf{Fails to Replicate} \\ \midrule
-Rose (2004) & WTO Membership Does not Effect Trade & Replicates \\ \midrule
-Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & \textbf{Fails to Replicate} \\\midrule
-Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & \textbf{Fails to Replicate} \\ \bottomrule
+Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Replicates} \\ \midrule
+McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
+Rose (2004) & WTO Membership Does not Effect Trade & {Replicates}\\ \midrule
+Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
+Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate} \\ \bottomrule
 	\end{tabular}
@@
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our re-extimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
+Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitaucpr} and Figure ~\ref{fig:reitpr}).
+Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitaucpr}).
 
-\begin{figure}
-\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-\end{figure}
+%\begin{figure}
+%%\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
+%\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
+%\end{figure}
 
@@
 \centering   
-\subfigure[AUC]{\label{fig:reitauc}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
+\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
 \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
@@
 
-\floatbarrier
 \subsection{Replication of McDonald (2004)}
@@
 % 0.03210221 0.03210221 0.99362646 0.00130249
+
+NEW GRAPHICS TO COME AND BE DESCRIBED.
+
 \begin{figure}

### 2017-10-19 paper/4_app.tex
mostly done in apps section but need a few greaphics and then a few tweaks.
done for the night, I suppose.

@@
 
-We utilize three performance criterions to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
+We utilize three performance criteria to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
 ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
@@
 	\toprule
-\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 & \textbf{0.96} \\
-		~ & AUC (PR) & 0.08 & \textbf{0.15} \\		\midrule
-\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 & \textbf{0.99} \\
-		~ & AUC (PR) & 0.13 & \textbf{0.28} \\		\midrule
-\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error (RMSE) & 3.23 & \textbf{1.99} \\
-		~ & Root Median Squared Error (RMDSE) & 2.01 & \textbf{1.06} \\	\midrule
-\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 & \textbf{0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 & \textbf{0.15} \\		\midrule
-\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 & \textbf{0.91} \\
-		~ & AUC (PR) & 0.00 & \textbf{0.08} \\			\bottomrule
+\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 &   {0.96} \\
+		~ & AUC (PR) & 0.08 &  {0.15} \\		\midrule
+\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 &  {0.99} \\
+		~ & AUC (PR) & 0.13 &  {0.28} \\		\midrule
+\multirow{2}{*}{Rose (2004)} & $ (1/n \sum \varepsilon^2)^{1/2} $ (RMSE) & 3.23 &  {1.99} \\
+		~ & $(\text{median}\quad (\varepsilon^2))^{1/2}$ (RMDSE) & 2.01 &  {1.06} \\	\midrule
+\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 &  {0.97} \\
+		~ & Area Under PR Curve (AUC-PR) & 0.00 &  {0.15} \\		\midrule
+\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 &  {0.91} \\
+		~ & AUC (PR) & 0.00 &  {0.08} \\			\bottomrule
 	\end{tabular}
@@
 \subsection{Lessons Learned from Re-estimating Five Prominent Studies}
+
+First and foremost, many  findings which emerge from models that do not
+take interdependencies into account lose their statistical significance when network effects
+are estimated via AME. This should not be news---since the finding has long been in the theoretical literature, 
+but given the state of current literature in international relations is still pertinent.  Not only
+are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are
+often imprecisely measured, with inflated standard errors.  
+This means that significant testing (for better or worse) is compromised when network effects are ignored.
+
+Second, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights
+emerge from the additional information derived. In particular, there is actual information about the dependencies so that
+clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.  This kind of information
+is absent in standard approaches and add to our ability to explain specific as well as general results.
+
+Third, it is clear that the AME approach dominates the OLS and Logit-based approaches in terms of performance.
+Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).
+In the case of studies
+with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model. It is rare for
+studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.
+
+Fourth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken
+with a grain of salt, until the interdependencies are taken into account.

### 2017-10-20 paper/3_sim.tex
sim stuff

@@
 We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
-The goal of the simulation is to assess how well AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved network dependencies.
-[link to previous discussion] Specifically, we are concerned with conducting inference on a regression parameter $\beta$ of a linear model for a network in the case where there is an omitted variable.
+The goal of the simulation is to assess how well AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
+Specifically, we are concerned with conducting inference on a regression parameter $\beta$ of a linear model for a network in the case where there is an omitted variable.
 Say that the true data-generating process for a particular $Y$ is given by:
@@
 	\item the standard IR approach assuming independent errors; 
-	\item the AME approach outlined in the previous section;
+	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);
 	\item and an ``oracle'' regression model that includes both $x_{i,j}$ and $w_{i,j}$. 
@@
 
-We run this simulation
+To run the simulation we vary the size of the network under study at 50 and 100.
 

### 2017-10-20 paper/4_app.tex
getting there

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{ABplot_rose_top10.pdf}
+\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
  \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
@@
 
-% sd.rowmean sd.colmean   dyad.dep  triad.dep 
-% 0.03210221 0.03210221 0.99362646 0.00130249
-
-NEW GRAPHICS TO COME AND BE DESCRIBED.
-
-\begin{figure}
-\caption{\label{fig:gibler:dyadtriad} The left panel is  a plot of the dyadic dependencies in the Gibler model 6; the right panel shows the triad dependencies.
-Note that the red line represents the null hypothesis of no dependencies, indicating that the standard logistic approach is far from what is uncovered with this analysis.
-}
-\includegraphics[width=\textwidth]{netplots}
-\end{figure}
-\FloatBarrier
+% % SM to MW: hmmm we should talk about the utility of this.
+% see fig gibler_srmvc.pdf in the paper dir
+% \begin{figure}
+% 	\caption{}
+% 	\includegraphics[width=\textwidth]{gibler_srmvc.pdf}
+% 	\label{fig:gibler:dyadtriad}
+% \end{figure}
+% \FloatBarrier
 

### 2017-10-20 code/2_codeForPaperGraphics/gibler_posterior.R

@@ -1,46 +1,39 @@
-############################################
+# setup ###########################################
 rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/gibler_2017/outputData/'
 	plotPath = '~/Dropbox/Research/netsMatter/replications/0_finalRepFigs/' }
 
-#
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-
-#
-loadPkg(c('amen','countrycode'))
 ############################################
 
-############################################
-#
-load(paste0(resultsPath,'McDonald_baseModel.rda'))
+# modData ###########################################
+load(paste0(resultsPath,'mdwameFit_k2_v4.rda'))
 load( paste0(resultsPath,'glmFit.rda') )
 ############################################
 
-############################################
-# coef summ
-# paramPlot(ameFit$BETA[,1:12])
-# paramPlot(ameFit$BETA[,12:23])
-
+# coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
 	mu=mean(x),
 	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
-# modSumm
 ############################################
 
+# plot srm var ###########################################
+# for undirected just do dist of nodal var
+vaViz = ggplot(data.frame(ameFit$VC), aes(x=va)) +
+	geom_density() +
+	geom_vline(aes(xintercept=0), linetype=2, color = "black") +
+	xlab(TeX('Within-Sender Variance ($\\\\sigma_{a]^{2}$)')) +
+	ylab('Density') +
+	theme(
+		axis.ticks=element_blank(), 
+		panel.border=element_blank()
+		)
+ggsave(vaViz, file=paste0(plotPath, 'gibler_srmvc.pdf'), width=7, height=4)
 ############################################
-#
-effdat = getAddEffData(fit = ameFit) ##This function is in helperEx.R
-effdat$actor = countrycode::countrycode(effdat$actor, 'cown', 'country.name')
-effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor'])
 
-## All countries
-addEffPlot(fit = effdat, addEffData = effdat, row = T)
-############################################
-
-############################################
-#
+# plot of first diffs ###########################################
 
 ############################################

### 2017-10-20 code/helpers/functions.R

@@ -9,9 +9,11 @@ loadPkg=function(toLoad){
 
 # some necessary libs
 loadPkg(c(
-	'reshape2',
+	'amen',
+	'reshape2', 'tidyr', 'dplyr', 'stringr',
 	'ggplot2', 'latex2exp', 'Cairo',	# plotting
-	'xtable' # tables
+	'xtable', # tables
+	'countrycode'
 	))
 
 # Set a theme for gg

### 2017-10-20 paper/4_app.tex
okay, more app changes.

@@
 
-% % SM to MW: hmmm we should talk about the utility of this.
+% MDW to SM:  yeah, Rose has so many variables that the network effects aren't that strong.
+% Maybe we should mention this. If you put in a gazillion variables you might be okay.
+% SM to MW: hmmm we should talk about the utility of this.
 % see fig gibler_srmvc.pdf in the paper dir
@@
 Fourth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken
-with a grain of salt, until the interdependencies are taken into account.
+with a grain of salt, and awaits a reassessment using methods that do not rely on an assumption that all the dyads are
+independent and identically distributed.

### 2017-10-20 paper/4_app.tex
okay, good now
won't touch this again until new graphics are updated.  Table too!

@@
 \end{figure}
+\FloatBarrier
 
@@
 \end{figure}
-
+\FloatBarrier
 %\begin{figure}
@@
 \includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
- \caption{Nodal Random Effects for AME estimation of Rose (2004).}\label{fig:roser}
+ \caption{Nodal Random Effects for AME estimation of Rose (2004)}\label{fig:roser}
 \end{figure}
-
+\FloatBarrier
 %\begin{figure}
@@
 The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out of sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
+In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
 
+%. THINK THIS IS GONE 
+%\begin{figure}
+%	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
+%	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
+%\end{figure}
 \begin{figure}
-	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
+	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
+	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
 \end{figure}
@@
 	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
+	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
 \end{figure}

### 2017-10-20 paper/4_app.tex
interim dock
yeah, I know I should fork, but why fix my manners now?
out and synced

@@
 Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
-Also part of the gold standard is a post estimation cleanup thought to be a paliative, and each of these studies adjust the posterior standard errors in an attempt to account for the clustering of observations.
+Also part of the gold standard is a post estimation cleanup thought to be a palliative, and each of these studies adjust the posterior standard errors in an attempt to account for the clustering of observations.
 
@@
 democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
-In a regression including fourteen coefficients, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our evaluation we focus on the same question in the re-examination of the Reiter and Stam study to determine whether the model replicates when reestimated within the AME framework.
+In a regression including fourteen coefficients, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our evaluation we focus on the same question in the re-examination of the Reiter and Stam study to determine whether the model replicates when re-estimated within the AME framework.
 Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textbf{fails to replicate} indicates only that the sign and/or significance of the crucial finding in the original study is not also found in the AME estimation.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not also found in the AME estimation.
 
@@
 	\toprule
-\multirow{2}{*}{Reiter \& Stam (2003)} & AUC (ROC) & 0.92 &   {0.96} \\
-		~ & AUC (PR) & 0.08 &  {0.15} \\		\midrule
-\multirow{2}{*}{McDonald (2004)} & AUC (ROC) & 0.92 &  {0.99} \\
-		~ & AUC (PR) & 0.13 &  {0.28} \\		\midrule
-\multirow{2}{*}{Rose (2004)} & $ (1/n \sum \varepsilon^2)^{1/2} $ (RMSE) & 3.23 &  {1.99} \\
-		~ & $(\text{median}\quad (\varepsilon^2))^{1/2}$ (RMDSE) & 2.01 &  {1.06} \\	\midrule
-\multirow{2}{*}{Weeks (2012)} & Area Under ROC Curve (AUC-ROC) & 0.64 &  {0.97} \\
-		~ & Area Under PR Curve (AUC-PR) & 0.00 &  {0.15} \\		\midrule
-\multirow{2}{*}{Gibler (2017)} & AUC (ROC) & 0.52 &  {0.91} \\
-		~ & AUC (PR) & 0.00 &  {0.08} \\			\bottomrule
+\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
+		~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
+\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
+		~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
+\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error RMSE & 3.23 &  {1.99} \\
+		~ & Root Median Squared Error RMDSE & 2.01 &  {1.06} \\	\midrule
+\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
+		~ & AUC-PR & 0.00 &  {0.15} \\		\midrule
+\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
+		~ & AUC-PR & 0.00 &  {0.08} \\			\bottomrule
 	\end{tabular}
@@
 \citet{mcdonald:2004} studies whether trade promotes peace between nations.
-The link between conflict and trade is perplexing, with many presistent, yet competing explanations.
+The link between conflict and trade is perplexing, with many persistent, yet competing explanations.
 \citet{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)''.
@@
 
-%\begin{figure}
-%\includegraphics[width=\textwidth]{rose_coefs_all_final.pdf}
-% \caption{Coefficient plot of Rose (2004)'s original model (blue) compared to AME model (red).}\label{fig:roseCoef}
-%\end{figure}
-
 \begin{figure}
@@
 \FloatBarrier
-%\begin{figure}
-%\centering   
-%\subfigure[Nodal Random Effects]%{\label{fig:rosesend}\includegraphics[width=65mm]{ABplot_rose_top10.pdf}}
-%\subfigure[Receiver Random Effects]{\label{fig:rosereceive}\includegraphics[width=65mm]{ABplot_receiver_rose.pdf}}
-%\caption{Sender and receiver random effects for AME estimation of Rose (2004).}
-%\label{fig:rosesr}
-%\end{figure}
 
@@
 
-%. THINK THIS IS GONE 
-%\begin{figure}
-%	\includegraphics[width=\textwidth]{weeks_coefs_all.pdf}
-%	\caption{\label{fig:weeksCoef} Coefficient plot of Weeks' (2012) original model (blue) compared to AME model (red). }
-%\end{figure}
 \begin{figure}
@@
 
-% We chose this article because it is published in 2017, demonstrating that the concern with interdependence continues to the present and because it was published in the so-called flagship journal of political science, \textit{The American Political Science Review.} It examines the link between dyadic capability differences and international conflict in terms of how and when states became independent members of the international system.
-It assumes that membership in the international system---defined in typical ways---has a substantial influence on capabilities.
-In the same way that capabilities don't seem to vary enormously over time, at least comparatively, states become independent in a particular location, clustered with surrounding states.
-Neither does this change much over time.
-Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
 The replications we have undertaken are all from articles over the past fifteen years.
@@
 
+THIS TABLE WILL BECOME A FIGURE, and skedaddle.
+
 \begin{table}
 \begin{center}
-\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000$ iterations.  \label{tab:gibme}}
+\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
 \begin{tabular}{lrr} \toprule
@@
 For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
+NEED TO BE FIXED if staying.
 \end{table}
 
-We re-estimated model 6 from Table 6 (2017,34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
+We re-estimated model 6 from Table 6 (2017, 34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
 The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
-Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the amen results.
-Not only is the value of the this parameter small, but it has a very large relative standard error, over a magnitude larger than the parameter itself ($z= 0.038$).
-In addition, joint democracy follows the same pattern of importance in the logistic results, but this disappears once interdependencies are modeled.
-As might be expected the strong geographic clustering in the original study is about one-quarter as strong in the \texttt{amen} estimations.
-Similarly, rivalry coefficients are about one-third the size in the \texttt{amen} formulation, but a great deal more precisely measured ($z=18.116$).
+Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
+Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$).
+In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.
+As might be expected, the strong geographic clustering in the original study is about one-quarter as strong in the AME estimations.
+Similarly, rivalry coefficients are about one-third the size in the AME formulation, but a great deal more precisely measured ($z=18.116$).
 
-Beyond more informative fixed effect coefficients, the \texttt{amen} approach also provides information about the interdependencies that were modeled.
-The most pertinent of these may be the dyadic and triadic dependencies which are presented in Figure~\ref{fig:gibler:dyadtriad}, which further reveal that the assumption of independence among the dyadic data in this study can be strongly rejected.
+We utilized the original and the AME results from Gibler's model 6 in Table 6 (as above) to calculate the expected values for one scenario.
+We focused on the variables measuring rivalry.
+We used mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.
+The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.
+As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ from the GLM in two distinct ways.
+First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably higher when taking interdependencies into account with the AME model.
+These are rare events, so the probabilities are low, but the difference is a factor of $3x$.
+Thus, you get quite substantially different expected values from these two models.
+Second, the GLM model has a standard errors which underestimate the true variability, in part because some of them are assumed to be zero.
+This shows up in the narrow density of predicted outcomes for the GLM model, compared to the density seen with the AME.
+The latter reflects a wider range of potential outcomes, reflecting the incorporation of uncertainty in the AME model.
+
+\begin{figure}
+	\caption{}
+	\includegraphics[width=\textwidth]{gibler_srmvc.pdf}
+ 	\label{fig:gibmargeff}
+ \end{figure}
+
+Beyond more informative fixed-effects coefficients, the AME approach also provides information about the interdependencies that were modeled.
+The assumption of independence among the dyadic data in this study can be strongly rejected.
 Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies.
@@
 
-% MDW to SM:  yeah, Rose has so many variables that the network effects aren't that strong.
-% Maybe we should mention this. If you put in a gazillion variables you might be okay.
-% SM to MW: hmmm we should talk about the utility of this.
-% see fig gibler_srmvc.pdf in the paper dir
-% \begin{figure}
-% 	\caption{}
-% 	\includegraphics[width=\textwidth]{gibler_srmvc.pdf}
-% 	\label{fig:gibler:dyadtriad}
-% \end{figure}
-% \FloatBarrier
-
 \subsection{Lessons Learned from Re-estimating Five Prominent Studies}
 
-First and foremost, many  findings which emerge from models that do not
-take interdependencies into account lose their statistical significance when network effects
-are estimated via AME. This should not be news---since the finding has long been in the theoretical literature, 
-but given the state of current literature in international relations is still pertinent.  Not only
-are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are
-often imprecisely measured, with inflated standard errors.  
+First and foremost, many  findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.
+This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.
+Not only
+are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.
@@
 Second, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights
-emerge from the additional information derived. In particular, there is actual information about the dependencies so that
-clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.  This kind of information
+emerge from the additional information derived.
+In particular, there is actual information about the dependencies so that clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.
+This kind of information
 is absent in standard approaches and add to our ability to explain specific as well as general results.
 
-Third, it is clear that the AME approach dominates the OLS and Logit-based approaches in terms of performance.
+Third, it is evident that the actual results---not the estimated coefficents and their covariances---which are generated by the models differ greatly in expectations.
+This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.
+
+Fourth, it is clear that the AME approach dominates the OLS and Logit-based approaches in terms of performance.
 Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).
 In the case of studies
-with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model. It is rare for
-studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.
+with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.
+It is rare for studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.
 
-Fourth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken
-with a grain of salt, and awaits a reassessment using methods that do not rely on an assumption that all the dyads are
-independent and identically distributed.
+Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, and awaits a reassessment using methods that do not rely on an assumption that all the dyads are independent and identically distributed.

### 2017-10-21 paper/4_app.tex
still deets to finish tomorrow

@@
 \centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=65mm]{weeks_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{weeks_pr_outsamp.pdf}}
+	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
 	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
@@
 	\caption{}
-	\includegraphics[width=\textwidth]{gibler_srmvc.pdf}
+	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
  	\label{fig:gibmargeff}

### 2017-10-21 code/2_codeForPaperGraphics/weeks_posterior.R

@@ -1,47 +1,123 @@
-############################################
+# setup ###########################################
 rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/Weeks2012/replication/output/'
 	inputPath= '~/Dropbox/Research/netsMatter/replications/Weeks2012/replication/input/'
-	plotPath = '~/Dropbox/Research/netsMatter/replications/0_finalRepFigs/' }
+	plotPath = '~/Research/netsMatter/paper/' }
 
-#
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-
-#
-loadPkg(c('amen','countrycode'))
+source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
 ############################################
 
-############################################
-#
+# modData ###########################################
 load( paste0(resultsPath,'model_k2.rda') )
 load(paste0(inputPath,'weeks_baseModel.rda'))
 ############################################
 
-############################################
-# coef summ
-# paramPlot(ameFit$BETA[,1:12])
-# paramPlot(ameFit$BETA[,12:23])
-
+# coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
 	mu=mean(x),
 	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
-# modSumm
 ############################################
 
-############################################
-#
-effdat = getAddEffData(fit = ameFit) ##This function is in helperEx.R
-effdat$actor = countrycode::countrycode(effdat$actor, 'cown', 'country.name')
-effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor'])
-
-## All countries
-addEffPlot(fit = effdat, addEffData = effdat, row = T)
+# plot srm var ###########################################
+plotVC(ameFit$VC, paste0(plotPath, 'weeks_srmvc.pdf'), w=7, h=4)
[…]

### 2017-10-21 code/helpers/binPerfHelpers.R

@@ -41,14 +41,14 @@ rocPlot = function(rocData, type='roc', legPos=c(.56,.25), colorPal = 'Set1', co
     geom_line(lwd=1) +
     ylim(0,1) + 
     scale_linetype_manual(values=linetypes) + 
-    theme_light(base_family="Source Sans Pro") + 
+    # theme_light(base_family="Source Sans Pro") + 
     theme(
   		legend.position=legPos, legend.title=element_blank(),
       legend.background=element_blank(), 
       legend.text.align = 0, legend.text=element_text(size=legText),
       legend.key=element_rect(colour = NA, fill = NA), legend.key.size=unit(legSpace,'lines'),
-      axis.text.x=element_text(family="Source Sans Pro Light"),
-      axis.text.y=element_text(family="Source Sans Pro Light"),    
+      # axis.text.x=element_text(family="Source Sans Pro Light"),
+      # axis.text.y=element_text(family="Source Sans Pro Light"),    
       axis.ticks=element_blank(),    
   		panel.border=element_blank()
 		)
@@ -67,6 +67,43 @@ ggSep = function(actual, proba, color, lty, fPath, save=TRUE){
     geom_line(aes(y = proba, x = seq(length.out = length(actual)), linetype=lty), lwd = 4) + 
     scale_linetype_manual(values=lty) +
     scale_color_manual(values=color) + 
+    scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0), breaks=seq(0,1,.25)) + 
+    theme(
+      legend.position='none', 
+      panel.grid=element_blank(), panel.border=element_rect(colour = "grey13"),
+      axis.ticks=element_blank(),
+      axis.text=element_blank(),
+      axis.title=element_blank()
+      )
+  if(save){ ggsave(tmp, file=fPath, width=12, height=2) } else { return(tmp) }
+}
+
+ggSepBig = function(actual, proba, color, lty, fPath, save=TRUE){
+  sepData = data.frame(actual, proba)
+  sepData = sepData[order(sepData$proba),]
+
+  #
+  sepData$cut = cut(sepData$proba, breaks=quantile(sepData$proba, seq(0, 1, 0.001)) )
+  sepDataSumm = sepData %>% group_by(cut) %>%
[…]

### 2017-10-21 paper/4_app.tex
more clean up.
still needed:

new:
reiter and stam graphics
mcdonald graphics

@@
 Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-
 Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
@@
 The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies that use a framework that ignores dependencies.
-We believe that it does, and that the dynamic latent factor model provides a step forward.
+Based in part our our results as well as the statistical and mathematical characteristics of the
+AME framework, we believe that it does. The dynamic latent factor model provides a step forward in the modeling of international relations.
 
@@
 
+\begin{figure}[!h]
+	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
+	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
+\end{figure}
+\FloatBarrier
+
 The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
-\end{figure}
-
-\begin{figure}
 \centering   
@@
 
-THIS TABLE WILL BECOME A FIGURE, and skedaddle.
-
-\begin{table}
-\begin{center}
-\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
-\begin{tabular}{lrr} \toprule
-& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-Allied & 0.142&  0.023 \\
-Joint Democracy &\bf -0.507&  0.045 \\
-Peace Years \bf &-0.260& \bf -0.060 \\
-Spline 1 &\bf -0.001& \bf 0.00 \\
-Spline 2&\bf -0.000& \bf 0.00\\
-Spline 3 &-0.000& \bf 0.00\\ 
-Contiguity &\bf 2.412& \bf 0.640 \\
-Parity &0.075&  -0.013 \\
-Parity at entry year&\bf 0.868&  0.002 \\ 
-Rivalry &\bf 2.031&\bf 0.721\\ 
-Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-\end{tabular}\\
-\end{center}
-{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
+Insert coefplot here, without splines...\label{fig:giberrepl}
+%
+%\begin{table}
+%\begin{center}
+%\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
+%\begin{tabular}{lrr} \toprule
+%& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
+%Variable & \texttt{logit} & \texttt{amen} \\ \midrule
+%Allied & 0.142&  0.023 \\
+%Joint Democracy &\bf -0.507&  0.045 \\
+%Peace Years \bf &-0.260& \bf -0.060 \\
+%Spline 1 &\bf -0.001& \bf 0.00 \\
+%Spline 2&\bf -0.000& \bf 0.00\\
+%Spline 3 &-0.000& \bf 0.00\\ 
+%Contiguity &\bf 2.412& \bf 0.640 \\
+%Parity &0.075&  -0.013 \\
+%Parity at entry year&\bf 0.868&  0.002 \\ 
+%Rivalry &\bf 2.031&\bf 0.721\\ 
+%Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
+%\end{tabular}\\
+%\end{center}
+%{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
 For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
 NEED TO BE FIXED if staying.
-\end{table}
+%\end{table}
 
-We re-estimated model 6 from Table 6 (2017, 34). The results are presented in Table~\ref{tab:gibme} in order to facilitate explicit comparison.
+We re-estimated model 6 from Table 6 (2017, 34). The results are presented in Figure~\ref{fig:giberrepl} in order to facilitate explicit comparison.
 The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
@@
 
-We utilized the original and the AME results from Gibler's model 6 in Table 6 (as above) to calculate the expected values for one scenario.
+We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario.
 We focused on the variables measuring rivalry.
@@
 \begin{figure}
-	\caption{}
+	\caption{THIS NEEDS TO BE FIXED.  \label{fig:gibmargeff}}
 	\includegraphics[width=\textwidth]{gibler_margeff.pdf}

### 2017-10-21 paper/4_app.tex
more changes

@@
 % TO DO: include discussion and analysis of different K's in appendix (for M cD)
-\begin{figure}
-\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-\end{figure}
-\FloatBarrier
+%\begin{figure}
+%%\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
+%\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
+%\end{figure}
+%\FloatBarrier
 %\begin{figure}
@@
 Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-Figure~\ref{fig:mcdCoefs} shows the coefficients for both the original and AME model.
-The most striking thing is that AME finds a positive---not negative as found by McDonald---association between trade dependence and conflict.
-The AME coefficient is not only in the opposite direction, but it is approximately three times more precisely estimated.
-At the same time, the AME has ROC and PR curves (not shown) that dominate the OLS results found in McDonald (2004).
+The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$).
+%The AME coefficient is not only in the opposite direction, but it is approximately three times more precisely estimated.  
+At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
 

### 2017-10-21 paper/4_app.tex
fix rose out samp

--- 2017-10-21 paper/4_app.tex
@@
 	\centering   
-\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
+	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{reiter_auc_outsamp.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{reiter_pr_outsamp.pdf}}
 	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}

### 2017-10-21 code/2_codeForPaperGraphics/mcdonald_posterior.R

@@ -1,47 +1,48 @@
-############################################
+# setup ###########################################
 rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/McDonald_2004/data/'
 	inputPath= '~/Dropbox/Research/netsMatter/replications/McDonald_2004/data/'
-	plotPath = '~/Dropbox/Research/netsMatter/replications/0_finalRepFigs/' }
+	plotPath = '~/Research/netsMatter/paper/' }
 
-#
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-
-#
-loadPkg(c('amen','countrycode'))
+source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
 ############################################
 
-############################################
-#
+# modData ###########################################
 load(paste0(resultsPath,'McDonald_baseModel.rda'))
 load( paste0(resultsPath,'ameFit_k2.rda') )
 ############################################
 
-############################################
-# coef summ
-# paramPlot(ameFit$BETA[,1:12])
-# paramPlot(ameFit$BETA[,12:23])
-
+# coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
 	mu=mean(x),
 	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
-# modSumm
 ############################################
 
+# outPerf ###########################################
+load(paste0(resultsPath, 'outsampResults2.rda')) # ameOutSamp_NULL
+ameOutSamp=ameOutSamp_NULL ; rm(ameOutSamp_NULL)
+ameOutSamp$outPerf$pred = 1/(1+exp(-ameOutSamp$outPerf$pred))
+load(paste0(resultsPath,'McDGLMPerf.rda')) # glmOutSamp_wFullSpec
+glmOutSamp=glmOutSamp_wFullSpec ; rm(glmOutSamp_wFullSpec)
+
+# org
+predDfs = list(
+	GLM = data.frame(actual=glmOutSamp$outPerf$actual, pred=glmOutSamp$outPerf$pred, model='GLM'),
+	AME = data.frame(actual=ameOutSamp$outPerf$actual, pred=ameOutSamp$outPerf$pred, model='AME') )
+
+# run
+ggPerfCurves(predDfs, 'mcdonald')
 ############################################
-#
+
+# abPlot ###########################################
[…]

### 2017-10-21 code/2_codeForPaperGraphics/reiter_stam_posterior.R

@@ -1,38 +1,28 @@
-############################################
+# setup ###########################################
 rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/output/'
 	inputPath= '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/input/'
-	plotPath = '~/Dropbox/Research/netsMatter/replications/0_finalRepFigs/' }
+	plotPath = '~/Research/netsMatter/paper/' }
 
-#
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-
-#
-loadPkg(c('amen','countrycode'))
+source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
 ############################################
 
-############################################
-#
+# modData ###########################################
 load( paste0(resultsPath,'reiterStam_baseModel.rda') )
 load( paste0(resultsPath,'ameFit_k2.rda') )
 ############################################
 
-############################################
-# coef summ
-# paramPlot(ameFit$BETA[,1:12])
-# paramPlot(ameFit$BETA[,12:23])
-
+# coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
 	mu=mean(x),
 	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
-# modSumm
 ############################################
 
-############################################
-#
+# addeffdata ###########################################
 effdat = getAddEffData(fit = ameFit) ##This function is in helperEx.R
 effdat$actor = countrycode::countrycode(effdat$actor, 'cown', 'country.name')
 effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor'])
@@ -41,7 +31,18 @@ effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor']
 addEffPlot(fit = effdat, addEffData = effdat, row = T)
 ############################################
 
-############################################
-#
+# outPerf ###########################################
+load(paste0(resultsPath, 'ameCrossValResults_k2.rda')) # ameOutSamp_k2
[…]

### 2017-10-21 code/helpers/functions.R

@@ -10,10 +10,10 @@ loadPkg=function(toLoad){
 # some necessary libs
 loadPkg(c(
 	'amen',
-	'reshape2', 'tidyr', 'dplyr', 'stringr',
 	'ggplot2', 'latex2exp', 'Cairo', 'RColorBrewer',	# plotting
 	'xtable', # tables
-	'countrycode'
+	'countrycode',
+	'reshape2', 'tidyr', 'stringr',	'dplyr'
 	))
 
 # Set a theme for gg

### 2017-10-21 code/rose2004/2a_ouSam_amen.R

@@ -1,16 +1,23 @@
 ################
 rm(list=ls())
 
-# workspace
-# if(Sys.info()['user']=='janus829' | Sys.info()['user']=='s7m'){ source('~/Research/conflictEvolution/R/setup.R')  }
-# if(Sys.info()['user']=='cassydorff' | Sys.info()['user']=='cassydorff'){ source('~/ProjectsGit/conflictEvolution/R/setup.R')  }
-# if(Sys.info()['user']=='maxgallop'){ source('~/Documents/conflictEvolution/R/setup.R')  }
-# loadPkg('devtools') ; devtools::install_github('s7minhas/amen') ; library(amen)
+if(Sys.info()['user']=='howardliu'){
+  source('~/netsMatter/code/rose2004/loadPkg.R')
+  pathResults = "/Users/howardliu/Dropbox/netsMatter/replications/rose2004/"
+}
+if(Sys.info()['user']=='s7m'){
+  source('~/Research/netsMatter/code/rose2004/loadPkg.R')
+  pathDrop = '~/Dropbox/Research/netsMatter/'
+  pathResults = paste0(pathDrop, 'replications/rose2004/')
+} 
 ################
 
 # load pkg
-source("clusteredSE.R")
-source("LoadPkg.R")
+loadPkg=function(toLoad){
+    for(lib in toLoad){
+      if(!(lib %in% installed.packages()[,1])){
+        install.packages(lib, repos='http://cran.rstudio.com/') }
+      suppressMessages( library(lib, character.only=TRUE) ) } }
 packs = c('foreign','dplyr', 'ggplot2', 'readr', 'lmtest','multiwayvcov', 'amen')
 loadPkg(packs)
 
@@ -21,18 +28,17 @@ trim = function (x) { gsub("^\\s+|\\s+$", "", x) }
 
 ################
 # load data
-pathResults = "/Users/howardliu/Dropbox/netsMatter/replications/rose2004/"
 load(paste0(pathResults, 'amenData_rose.rda'))
 
-resultsPath = "/Users/howardliu/Dropbox/netsMatter/replications/rose2004/outputData/"
-load(paste0(resultsPath, 'ameFit_k3_v1.rda')); ameFit_k3 = ameFit
[…]

### 2017-10-21 code/rose2004/2b_outSamp_glm.R

@@ -19,7 +19,8 @@ load(paste0(pathResults, 'amenData_rose.rda'))
 
 # crossval params
 seed=6886
-folds=30
+# folds=30
+folds=4
 ################
 base_vars = c('ltrade','bothin', 'onein', 'gsp', 'ldist', 'lrgdp', 'lrgdppc', 'regional',
               'custrict',
@@ -68,26 +69,10 @@ foldVec = foldVec[which(foldVec$Var1!=foldVec$Var2),]
     return(fit)
   })
 
-  # for(f in 1:30){
-  # testData = cbind(int=1,yCrossValTrain[[f]][which(foldVec$value == f),names(coef(fitCrossVal[[f]]))[-1]])
-  # #prob = 1/(1+exp(-as.matrix(testData) %*% coef(fitCrossVal[[f]])))
-  # prob =  as.matrix(testData) %*% coef(fitCrossVal[[f]])
-  # foldIndex = foldVec[which(foldVec$value == f),]
-  # foldIndex = foldIndex[which(!is.na(prob)),]
-  # prob = prob[!is.na(prob)]
-
-  # get actual
-  # actual = c()
-  # for(i in 1:dim(foldIndex)[1]){
-  #   ll = foldIndex$L1[i]
-  #   actual = c(actual, yList[[ll]][which(rownames(yList[[ll]]) == foldIndex$Var1[i]), which(rownames(yList[[ll]]) == foldIndex$Var2[i])])
-  # }
-  # print(length(actual) == length(prob))}
     # get preds
   outPerf = do.call('rbind', lapply(1:folds, function(f){
     # get probs
     testData = cbind(int=1,yCrossValTrain[[f]][which(foldVec$value == f),names(coef(fitCrossVal[[f]]))[-1]])
-    #prob = 1/(1+exp(-as.matrix(testData) %*% coef(fitCrossVal[[f]])))
     prob =  as.matrix(testData) %*% coef(fitCrossVal[[f]])
     foldIndex = foldVec[which(foldVec$value == f),]
     foldIndex = foldIndex[which(!is.na(prob)),]
@@ -101,26 +86,9 @@ foldVec = foldVec[which(foldVec$Var1!=foldVec$Var2),]
     }
     if(length(actual)!=length(prob)){stop('shit went wrong.')}
     res = data.frame(actual=actual, pred=prob, fold=f, stringsAsFactors = FALSE)
[…]

### 2017-10-21 paper/4_app.tex
cleanup, coefs left, waiting on rose out

--- 2017-10-21 paper/4_app.tex
@@
 We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
+Our goal is to ascertain whether the ignored interdependencies -- the non-iid structure of the relational data -- would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
 Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
@@
 The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies that use a framework that ignores dependencies.
-Based in part our our results as well as the statistical and mathematical characteristics of the
-AME framework, we believe that it does. The dynamic latent factor model provides a step forward in the modeling of international relations.
+Based in part our our results as well as the statistical and mathematical characteristics of the AME framework, we believe that it does.
+The dynamic latent factor model provides a step forward in the modeling of international relations.
 
@@
 Generally, interpretations focus on a small set of independent variables.
-In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---whether it is the
-democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
+In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads -- consisting of a democratic country and country ruled by a personalist dictatorship -- whether it is the democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
 In a regression including fourteen coefficients, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our evaluation we focus on the same question in the re-examination of the Reiter and Stam study to determine whether the model replicates when re-estimated within the AME framework.
@@
 This model outperforms each of the GLM variants by a notable margin.
-% % % % % % % % % % % % % % % % % %
 
@@
 
-%\begin{figure}
-%%\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-%\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-%\end{figure}
-
 \begin{figure}
 	\centering   
-\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=65mm]{reiter_auc_outsamp.pdf}}
-\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=65mm]{reiter_pr_outsamp.pdf}}
+	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
 	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
@@
 % TO DO: include discussion and analysis of different K's in appendix (for M cD)
-%\begin{figure}
-%%\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-%\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-%\end{figure}
-%\FloatBarrier
-%\begin{figure}
-%\centering   
-%\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-%\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-%\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-%\end{figure}
 
@@
 The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$).
-%The AME coefficient is not only in the opposite direction, but it is approximately three times more precisely estimated.  
 At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
@@
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders...
-(2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
+Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO -- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
 
 When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
-The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---persists in the AME model.
+The main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- persists in the AME model.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample---$3.23$ for the OLS model and $1.77$ for the AME model.
+When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample -- $3.23$ for the OLS model and $1.77$ for the AME model.
 While the basic model replicates with the basic result, there are substantial and substantive improvements gleaned from the AME results.
@@
 Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
-Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$).
+Not only is the value of this parameter small, but it has a very large relative standard error -- over a magnitude larger than the parameter itself ($z= 0.038$).
 In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.
@@
 As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ from the GLM in two distinct ways.
-First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably higher when taking interdependencies into account with the AME model.
+First, the expected value of the dependent variable -- the probability of the onset of a militarized interstate dispute, is considerably higher when taking interdependencies into account with the AME model.
 These are rare events, so the probabilities are low, but the difference is a factor of $3x$.
@@
 First and foremost, many  findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.
-This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.
+This should not be news -- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.
 Not only
@@
 
-Third, it is evident that the actual results---not the estimated coefficents and their covariances---which are generated by the models differ greatly in expectations.
+Third, it is evident that the actual results -- not the estimated coefficents and their covariances -- which are generated by the models differ greatly in expectations.
 This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.

### 2017-10-21 paper/6_appendix.tex
fix rose out samp
cleanup, coefs left, waiting on rose out

@@
 
-% \subsection*{Additive and Multiplicative Effects Gibbs Sampler}
+\subsection*{Performance Analysis}
 
-% To estimate, the effects of our exogenous variables and latent attributes we utilize a Bayesian probit model in which we sample from the posterior distribution of the full conditionals until convergence.
-Specifically, given observed data $\textbf{Y}$ and $\textbf{X}$ -- where $\textbf{X}$ is a design array that includes our sender, receiver, and dyadic covariates -- we estimate our network of binary ties using a probit framework where: $y_{ij} = 1(\theta_{ij}>0)$ and $\theta_{ij} = \bm\beta^{\top}\mathbf{X}_{ij} + a_{i} + b_{j} + \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j} + \epsilon_{ij}$.
+\subsubsection*{Gibler (2017)}
 
-% \begin{itemize}
-%  \item Given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows:
-%  \begin{itemize}
-%  	\item sample $\bm\theta \; | \;  \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-%  	\item sample $\bm\beta \; | \;  \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-%  	\item sample $\textbf{a}, \textbf{b} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-% 	\item sample $\Sigma_{ab} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
-%  	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
-%  	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
-%  	\item For each $k \in K$:
-%  	\begin{itemize}
-%  		\item Sample $\textbf{U}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}_{[,-k]}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-%  		\item Sample $\textbf{V}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-%  		\item Sample $\textbf{D}_{[k,k]}  \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \textbf{D} matrix is absorbed into the calculation for $\textbf{V}$ as we iterate through $K$.
-}
-%  	\end{itemize}
-%  \end{itemize}
-% \end{itemize}
+\begin{figure}
+	\centering   
+	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves and PR curves.}
+\end{figure}
+\FloatBarrier
+
+\subsubsection*{McDonald (2004)}
+
+\begin{figure}
+	\centering   
+	\subfigure[AUC]{\label{fig:mcdonaldroc}\includegraphics[width=.45\textwidth]{mcdonald_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:mcdonaldpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
+\end{figure}
+\FloatBarrier
+
+\subsubsection*{Reiter \& Stam (2003)}
+
+\begin{figure}
+	\centering   
+	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
+\end{figure}
+\FloatBarrier
+
+\subsubsection*{Rose (2004)}
+
+% \begin{figure}
+% 	\centering   
+% 	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{mcdonald_roc_outSample.pdf}}
+% 	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
+% 	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
+% \end{figure}
+% \FloatBarrier
+
+\subsubsection*{Weeks (2012)}
+
+\begin{figure}
+	\centering   
+	\subfigure[AUC]{\label{fig:weeksroc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:weekspr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
+\end{figure}
+\FloatBarrier

### 2017-10-22 paper/4_app.tex
clean up on aisle 4

--- 2017-10-21 paper/4_app.tex
@@
 		Weeks (2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
-Gibler (2017) & Logit & 1816--2008 &193 &   $650,557$ & UnDirected & none \\ \bottomrule
+		Gibler (2017) & Logit & 1816--2008 &193 &   $650,557$ & Undirected & none \\ \bottomrule
 	\end{tabular}
@@
 Generally, interpretations focus on a small set of independent variables.
-In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---whether it is the
-democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
+In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---whether it is the democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
 In a regression including fourteen coefficients, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our evaluation we focus on the same question in the re-examination of the Reiter and Stam study to determine whether the model replicates when re-estimated within the AME framework.
@@
 
-%\begin{figure}
-%%\includegraphics[width=\textwidth]{reiter_coefs_all.pdf}
-%\caption{\label{fig:reitCoef} Coefficient plot of Reiter \& Stam (2003)'s original model (blue) compared to AME model (red). }
-%\end{figure}
-
 \begin{figure}
 	\centering   
-	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{reiter_auc_outsamp.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{reiter_pr_outsamp.pdf}}
+	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
 	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
@@
 
-% TO DO: include discussion and analysis of different K's in appendix (for M cD)
-%\begin{figure}
-%%\includegraphics[width=\textwidth]{McDonald_coefs.pdf}
-%\caption{\label{fig:mcdCoefs}Coefficient plot of McDonald (2004)'s original model (blue) compared to AME model (red).}
-%\end{figure}
-%\FloatBarrier
-%\begin{figure}
-%\centering   
-%\subfigure[AUC]{\label{fig:mcDauc}\includegraphics[width=65mm]{McDonald_auc_outsamp.pdf}}
-%\subfigure[Precision and Recall]{\label{fig:mcDpr}\includegraphics[width=65mm]{McDonald_pr_outsamp.pdf}}
-%\caption{\label{fig:mcdall} Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-%\end{figure}
-
 Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
@@
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders...
-(2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
+Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
 
@@
 	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Blue represents groups with common sending patterns and red represents groups with common receiving patterns.}
+	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). 
+	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
+	}
 \end{figure}
@@
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+Inspection of the multiplicative effects in
+Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
 In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
@@
 In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
-So, in the GLM, which ignores these third order dependencies, much of these results might have been attributed to regime type.
-The AME model, on the other hand, finds that it is more efficient to attribute this behavior to the multiplicative effects.
-In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third order dependencies cause spurious effects for regime type.
+In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type.
+The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.
+In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
+In terms of performance, the original Weeks' model is slightly better than a coin-toss overall, and fails to indicate more than a couple of events (incorrectly) in almost one million observations.
 
@@
 The assumption of independence among the dyadic data in this study can be strongly rejected.
-Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the \texttt{amen} estimation that explicitly models \first- \second-, and \third-order interdependencies.
+Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the AME estimation that explicitly models \first- \second-, and \third-order interdependencies.
 Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.

### 2017-10-22 paper/5_concl.tex
clean up on aisle 4

@@
 At the same time, we know that research designs which focus on the statistical analysis of dyadic data quickly go astray if the dyadic data are assumed to be independent and identically distributed (iid).
-Virtually all of the standard statistical models--ordinary least squares, logistic and probit regressions, to name a few--fail if the data are not iid.
+Virtually all of the standard statistical models---ordinary least squares, logistic and probit regressions, to name a few---fail if the data are not iid.
 By definition dyadic data are not iid and thus the standard approaches can not be used cavalierly to analyze these data.
@@
 These approaches have been developed for a while, but are not yet widely used in international relations scholarship.
-The statistical model of the latent network captures first-order (example or restatement), second-order (example or restatement), and third-order (example or restatement) dependencies in dyadic data using a familiar regression framework that has been adapted for relational data--such as dyadic data--which are not independent nor identically distributed.
+The statistical model of the latent network captures first-order (example or restatement), second-order (example or restatement), and third-order (example or restatement) dependencies in dyadic data using a familiar regression framework that has been adapted for relational data---such as dyadic data---which are not independent nor identically distributed.
 There is an available, open source computer packages that implements this approach.
@@
 To explore this approach in the context of international relations we conduct two broad analyses.
-The first is a simulation of XCLKFJDSLF.
+The first is a simulation of where the characteristics of the network are known.
+This shows that the AME approach is unbiased in terms of parameter estimation compared with standard approach employed in
+international relations to study dyadic data (i.e., GLM models).
 The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
@@
 We then reanalyzed each study using the latent network approach which captures that additive and multiplicative aspects of interdependencies among the dyadic data.
-In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.
-In several cases, the new approach overturns the basic findings of the original research.
+In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix hold performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
 This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they only obtain under stringent assumptions that can not possibly be valid.
 This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
-At the same time, the latent factor model we present and demonstrate herein provides a way to easily examine and if necessary defeat these assumptions.
+At the same time, the latent factor model provides a way to easily examine and if necessary defeat these assumptions.
 

### 2017-10-22 code/2_codeForPaperGraphics/gibler_posterior.R
stargazer tables

@@ -8,6 +8,8 @@ if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
 source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
+source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
+source('~/Research/netsMatter/code/helpers/clusteredSE.R')
 ############################################
 
 # modData ###########################################
@@ -17,8 +19,35 @@ load( paste0(resultsPath,'glmFit.rda') )
 
 # coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	mu=mean(x),
-	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+
+glmSumm = modSumm[,]
+
+dat = cbind(y=mod$y, mod$data[,names(coef(mod)[-1])], dyad=mod$data$dyad)
+form = paste0('y~',paste(names(coef(mod)[-1]), collapse='+'))
+probMod = glm( form, data=dat, family=binomial(link='probit') )
+loadPkg('lmtest')
+probSumm = coeftest(probMod, vcov = vcovCluster(mod, cluster = dat$dyad))[,]
+
+modList = list(glmSumm, probSumm, ameSumm)
+modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+varKey = data.frame(
+	dirty=names(coef(mod)),
+	clean=c( '(Intercept)', 
+		'Allied', 'Joint Democracy',
[…]

### 2017-10-22 code/2_codeForPaperGraphics/reiter_stam_posterior.R

@@ -9,17 +9,41 @@ if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
 source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
+source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
 ############################################
 
 # modData ###########################################
-load( paste0(resultsPath,'reiterStam_baseModel.rda') )
-load( paste0(resultsPath,'ameFit_k2.rda') )
+load( paste0(resultsPath,'reiter_stam_glmfit.rda') )
+load( paste0(resultsPath,'model_k2_v12.rda') )
 ############################################
 
 # coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	mu=mean(x),
-	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+glmSumm = modSumm[,]
+probSumm = proSumm[,]
+
+modList = list(glmSumm, probSumm, ameSumm)
+modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+
+varKey = data.frame(
+	dirty=names(coef(mod)),
+	clean=c('Intercept', 'Pers/Democ Directed Dyad',
+		'Democ/Pers Directed Dyad',
+        'Personal', 'Military', 'Single', 'Democracy', 
+        'Contiguous', 'Major Power', 'Ally', 'Higher/Lower Power Ratio', 
+        'Economically Advanced', 'Years Since Last Dispute', 'Cubic Spline 1', 
+        'Cubic Spline 2', 'Cubic Spline 3'),
[…]

### 2017-10-22 code/2_codeForPaperGraphics/rose_posterior.R

@@ -7,6 +7,7 @@ if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
+source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
 ############################################
 
 # modData ###########################################
@@ -14,11 +15,29 @@ load(paste0(resultsPath,'outputData/ameFit_k2_v2_imps_50000_intercept.rda'))
 load( paste0(resultsPath,'glmResults_rose.rda') )
 ############################################
 
-# coef summ ###########################################
-# get ame est
+# coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	mu=mean(x),
-	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+lmSumm = baseModelSumm[,]
+
+modList = list(lmSumm, ameSumm)
+modNames = c('LM', 'AME')
+
+varKey = data.frame(
+	dirty=rownames(ameSumm),
+	clean=c('Intercept', 'Both in GATT/WTO', 'One in GATT/WTO', 'GSP',
+		'Log Distance',  'Log Product Real GDP', 'Log Product Real GDPpc',
+		'Regional FTA', 'Currency Union', 'Common language', 'Land Border',
+		'Number Landlocked' ,'Number Islands', 'Log Product Land Area',
+		'Common Colonizer','Currently Colonized', 'Ever Colony','Common Country' ),
+	stringsAsFactors = FALSE )
+
+#
+getCoefTable(varKey, modList, modNames, 'rose', 'Rose (2004)', plotPath, 3)
 ############################################
 
 # outPerf ###########################################

### 2017-10-22 code/2_codeForPaperGraphics/weeks_posterior.R

@@ -3,23 +3,54 @@ rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/Weeks2012/replication/output/'
-	inputPath= '~/Dropbox/Research/netsMatter/replications/Weeks2012/replication/input/'
 	plotPath = '~/Research/netsMatter/paper/' }
 
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')
 source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
+source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
 ############################################
 
 # modData ###########################################
 load( paste0(resultsPath,'model_k2.rda') )
-load(paste0(inputPath,'weeks_baseModel.rda'))
+load(paste0(resultsPath,'weeks_glmfit.rda'))
 ############################################
 
 # coefSumm ###########################################
 ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	mu=mean(x),
-	quantile(x, probs=c(0.025,0.05,0.95,0.975))) }))
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+glmSumm = modSumm[,]
+probSumm = proSumm[,]
+
+modList = list(glmSumm, probSumm, ameSumm)
+modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+
+varKey = data.frame(
+	dirty=names(coef(mod)),
+	clean=c( '(Intercept)', 
+		"Machine", "Junta", "Boss", "Strongman",
+		"Other Type","New/Unstable Regime", "Democracy Target",
+		"Military Capabilities Initiator",
+		"Military Capabilities Target ",
[…]

### 2017-10-22 code/reiter_stam_2003/1 - replication-rs.R

@@ -10,15 +10,30 @@ rm(list=ls())
 packs = c('dplyr', 'ggplot2', 'foreign', 'readr', 'lmtest')
 
 #
-source("LoadPkg.R")
-source("/Users/juanftellez/OneDrive/netsMatter/code/helpers/clusteredSE.R")
+if(Sys.info()['user']=='juanftellez'){
+	source("LoadPkg.R")
+	source("/Users/juanftellez/OneDrive/netsMatter/code/helpers/clusteredSE.R")
 
-#
-loadPkg(packs)
+	#
+	loadPkg(packs)
 
-# read data and add variables
-stam = 
-  read.dta('/Users/juanftellez/Dropbox/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
+	# read data and add variables
+	stam = 
+	  read.dta('/Users/juanftellez/Dropbox/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
+}
+
+if(Sys.info()['user']=='s7m'){
+	source("~/Research/netsMatter/code/helpers/loadPkg.R")
+	source("~/Research/netsMatter/code/helpers/clusteredSE.R")
+
+	#
+	loadPkg(packs)
+
+	# read data and add variables
+	dataPath = '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/input/'
+	resultsPath = '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/output/'
+	stam = read.dta(paste0(dataPath, 'doublereduce.dta'))
+}
 
 # get variables
 base_vars = c('sideaa pdemdtar pdemdin personal military single democ contig majpow ally loglsrat advanced dispyrs dspline1 dspline2 dspline3') %>% 
@@ -32,8 +47,20 @@ mod_dat = select(stam, one_of(base_vars), idyr, statea, stateb) %>%
 form_mod = formula(paste0('sideaa ~ ', paste(base_vars[-1], collapse = '+')))
 
 # model
+if(Sys.info()['user']=='juanftellez'){
 mod1 = glm(form_mod, family = binomial(link = 'logit'), data = mod_dat)
 # clustered SEs
 base_mod1 = round(coeftest(mod1, vcov = vcovCluster(mod1, cluster = mod_dat$idyr)), 3)
 
[…]

### 2017-10-22 code/rose2004/rep_rose.R

@@ -1,15 +1,29 @@
 ## replication Rose 2004 WTO --> trade
-
+rm(list=ls())
 # load pkg
-source("/Users/howardliu/Dropbox/netsMatter/replications/rose2004/clusteredSE.R")
-source("/Users/howardliu/Dropbox/netsMatter/replications/rose2004/LoadPkg.R")
+if(Sys.info()['user']=='howardliu'){
+    source("/Users/howardliu/Dropbox/netsMatter/replications/rose2004/clusteredSE.R")
+    source("/Users/howardliu/Dropbox/netsMatter/replications/rose2004/LoadPkg.R")
 
 packs = c('foreign','dplyr', 'ggplot2', 'readr', 'lmtest','multiwayvcov')
 loadPkg(packs)
 
 # data
 data = read.dta("/Users/howardliu/Dropbox/netsMatter/replications/rose2004/data4web.dta")
-#names(data)
+}
+
+
+if(Sys.info()['user']=='s7m'){
+    source("~/Research/netsMatter/code/helpers/clusteredSE.R")
+    source("~/Research/netsMatter/code/helpers/loadPkg.R")
+
+    packs = c('foreign','dplyr', 'ggplot2', 'readr', 'lmtest','multiwayvcov')
+    loadPkg(packs)
+
+    # data
+    dataPath='~/Dropbox/Research/netsMatter/replications/rose2004/'
+    data = read.dta(paste0(dataPath,'data4web.dta'))
+}
 
 # mod1
 # which var is currency union?

### 2017-10-22 paper/1_intro.tex

@@
 
-We begin by providing a brief review of these dependencies and the AME model.
-Next, we move to conducting a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
+We begin with a brief review of these dependencies and the AME model.
+Next, conduct a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
 Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.

### 2017-10-22 paper/2_model.tex

@@
 
-To uncover the underlying structure that relational data comes with it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often referred to as an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure that relational data comes with it is helpful to shift towards restructuring dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 
-By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we draw
-
+By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw.
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.

### 2017-10-22 paper/4_app.tex
more clean up we are at 99%

@@
 
-We re-estimated model 6 from Table 6 (2017, 34). The results are presented in Figure~\ref{fig:giberrepl} in order to facilitate explicit comparison.
-The results obtained with \texttt{amen} stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
+We re-estimated model 6 from Table 6 (2017, 34). The results are presented in the Appendix, along with each of the models we estimated.
+The results obtained with AME stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
 Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
@@
 \begin{figure}
-	\caption{THIS NEEDS TO BE FIXED.  \label{fig:gibmargeff}}
+	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
 	\includegraphics[width=\textwidth]{gibler_margeff.pdf}

### 2017-10-22 paper/4_app.tex
marg eff fix

--- 2017-10-22 paper/4_app.tex
@@
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
-Finally, our modeling approach outperforms the original model by better, and more accurately, predicting MIDs out-of-sample (Figure~\ref{fig:reitaucpr}).
-
-\begin{figure}
-	\centering   
-	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
-\end{figure}
-\FloatBarrier
 
@@
 The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.
-As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ from the GLM in two distinct ways.
-First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably higher when taking interdependencies into account with the AME model.
-These are rare events, so the probabilities are low, but the difference is a factor of $3x$.
+As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notable.
+First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
+These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
 Thus, you get quite substantially different expected values from these two models.
-Second, the GLM model has a standard errors which underestimate the true variability, in part because some of them are assumed to be zero.
-This shows up in the narrow density of predicted outcomes for the GLM model, compared to the density seen with the AME.
-The latter reflects a wider range of potential outcomes, reflecting the incorporation of uncertainty in the AME model.
 

### 2017-10-22 paper/6_appendix.tex
stargazer tables
more clean up we are at 99%
marg eff fix

@@
 
-\subsection*{Performance Analysis}
+\subsection*{Additional Replication Information}
 
 \subsubsection*{Gibler (2017)}
+
+Coefficient plot and performance comparison.
+
+\input{gibler_coeftable}
+\FloatBarrier
 
@@
 
+\clearpage
 \subsubsection*{McDonald (2004)}
+
+Coefficient plot and performance comparison.
+
+\input{mcdonald_coeftable}
+\FloatBarrier
 
@@
 
+\clearpage
 \subsubsection*{Reiter \& Stam (2003)}
+
+Coefficient plot and performance comparison.
+
+\input{reiter_stam_coeftable}
+\FloatBarrier
 
@@
 
+\clearpage
 \subsubsection*{Rose (2004)}
+
+Coefficient plot and performance comparison.
+
+\input{rose_coeftable}
+\FloatBarrier
 
@@
 % \end{figure}
-% \FloatBarrier
+\FloatBarrier
 
+\clearpage
 \subsubsection*{Weeks (2012)}
 
-\begin{figure}
-	\centering   
-	\subfigure[AUC]{\label{fig:weeksroc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:weekspr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
-\end{figure}
+Coefficient plot and performance comparison.
+
+\input{weeks_coeftable}
 \FloatBarrier
+
+% already show roc-pr in main text
+% \begin{figure}
+% 	\centering   
+% 	\subfigure[AUC]{\label{fig:weeksroc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
+% 	\subfigure[Precision and Recall]{\label{fig:weekspr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
+% 	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
+% \end{figure}
+% \FloatBarrier

### 2017-10-22 code/2_codeForPaperGraphics/gibler_posterior.R

@@ -46,6 +46,8 @@ varKey = data.frame(
 		'Parity at Entry Year', 'Rivalry' ),
 	stringsAsFactors = FALSE )
 
+
+
 #
 getCoefTable(varKey, modList, modNames, 'gibler', 'Gibler (2017)', plotPath, 3)
 ############################################
@@ -108,73 +110,32 @@ scen1 = scen[,seq(1,ncol(scen),2)] ; scen0 = scen[,seq(2,ncol(scen),2)]
 scens = c( 'Allied', 'Joint Democracy', 'Contiguity', 
 	'Rivalry', 'Parity at entry year', 'Parity')
 
-# calc diffs
-getDiff = function(
-	linkType='logit', # logit or probit
-	scenHi, scenLo, scenNames, 
-	beta, modName, type='summStats' # summStats or density
-	){
-	if(linkType=='logit'){
-		predHi = 1/(1+exp(-t(t(scenHi) %*% t(beta))))
-		predLo = 1/(1+exp(-t(t(scenLo) %*% t(beta)))) }
-	if(linkType=='probit'){
-		predHi = pnorm(t(t(scenHi) %*% t(beta)))
-		predLo = pnorm(t(t(scenLo) %*% t(beta))) }
-	predDiff = predHi-predLo
-	colnames(predDiff) = scenNames
-
-	if(type=='summStats'){
-		summPred = matrix(NA,nrow=5, ncol=ncol(predDiff),
-			dimnames=list(
-				c('med','hi95','hi90','lo95','lo90'), colnames(predDiff) ))
-		for(s in colnames(summPred)){
-			summPred['med',s]=median(predDiff[,s])
-			summPred['hi95',s]=quantile(predDiff[,s],.975)
-			summPred['hi90',s]=quantile(predDiff[,s],.95)
-			summPred['lo95',s]=quantile(predDiff[,s],.025)
-			summPred['lo90',s]=quantile(predDiff[,s],.05) }
-	
-		# org and spit
-		summPred = t(summPred) %>% data.frame() %>%
-			mutate(
-				mod=modName,
-				scen=colnames(summPred) )
-		}
-
-	if(type=='density'){
-		summPred = predDiff %>% data.frame() %>%
-			gather(key='scen',value='value') %>%
-			mutate(
-				mod=modName,
-				scenClean=gsub('.',' ', scen, fixed=TRUE)
-				)
-	}
-	return(summPred) }
[…]

### 2017-10-22 paper/5_concl.tex
sdf

@@
 We then reanalyzed each study using the latent network approach which captures that additive and multiplicative aspects of interdependencies among the dyadic data.
-In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix hold performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
+In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
 This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they only obtain under stringent assumptions that can not possibly be valid.

### 2017-10-22 paper/4_app.tex
tiny changes

@@
 
-\subsection{Replication of Reiter \& Stam (2003)}
+\subsection{Re-estimation of Reiter \& Stam (2003)}
 
@@
 In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that ``We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
+They use the MID data, but note that
+\begin{quote}
+We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
 This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
 We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).'' Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
+\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).
+\end{quote} 
+Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
 As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
-The database for this study is constructed using EUGene \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
+The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
 Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
@@
 
-\subsection{Replication of McDonald (2004)}
+\subsection{Re-estimation of McDonald (2004)}
 
@@
 
-\subsection{Replication of Rose (2004)}
+\subsection{Re-estimation of Rose (2004)}
 
@@
 
-\subsection{Replication of Weeks (2012)}
+\subsection{Re-estimation of Weeks (2012)}
 
@@
 
-\subsection{Replication of Gibler (2017)}
+\subsection{Re-estimation of Gibler (2017)}
 
@@
 
-Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, and awaits a reassessment using methods that do not rely on an assumption that all the dyads are independent and identically distributed.
+Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are independent and identically distributed.

### 2017-10-23 paper/1_intro.tex
first cut of sim done

@@
 We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-The multiplicative effects capture dependencies that results from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence.
+The multiplicative effects capture dependencies that result from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence.
 These type of meso-scopic features often arise in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.
@@
 We begin with a brief review of these dependencies and the AME model.
-Next, conduct a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
+Next, we conduct a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
 Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
 The latent factor approach (AME) is able to better capture first, second, and third-order interdependencies than the standard approach.
-It also produces results that are more precise and at times at odds with those found in these studies in particular, and the broader literatures from which they are drawn.
-As such, this approach offers substantive insights which are often occluded by ignoring the interdependent nature of the relational data that characterize many studies in the field of international relations.
+It also produces results that are, at times, at odds with those found in these studies in particular, and the broader literatures from which they are drawn.
+As such, this approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
 Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
-Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest in international relations.
-It facilitates the concentration on international relations in the field of international relations.
+Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
+Most importantly, it facilitates the concentration on the relations aspect of the field of international relations.

### 2017-10-23 paper/2_model.tex
first cut of sim done

--- 2017-10-19 paper/2_model.tex
@@
 We know this assumption to be false in theory and in practice because relational data comes with an explicit structure that in general leads to particular types of dependencies.
-The importance of accounting for underlying structure of our data has been a lesson well understood at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
+The importance of accounting for the underlying structure of our data has been a lesson well understood, at least, when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
 As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
-
-To uncover the underlying structure that relational data comes with it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often referred to as an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often referred to as an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 Homogeneity in interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
-Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors which are much more active than others \citep{barabasi:reka:1999}.
+Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor may be more active than others, parameter estimates from standard statistical models will be biased.
@@
 
-For similar reasons one also needs to take into account that there is a shared dependence between observations that share a common receiver.
+For similar reasons one also needs to take into account the shared dependence between observations that share a common receiver.
 The bottom-left panel, illustrates that sender and receiver type dependencies can also blend together.
@@
 
+% SM: Why are we citing poast????
 For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level.
@@
 The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are more rare.
-A prominent example of a network with this type of structure was found in American Politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another in the 2004 United States Election.
+A prominent example of a network with this type of structure was found in American Politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
 \citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs just linked to others that were political similar.
 This showcases the types of higher-order dependencies that can emerge in relational data.
-First, the fact that interactions was determined by a shared attribute, in this case political ideology, is an example of what is known as \textit{homophily}.
-Homophily can be used to explain the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
+First, the fact that interactions was determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}.
+Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
 The other major type of meso-scopic feature that emerges in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
@@
 Of course, if one is able to specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
-However, it is rarely the case that we can do this even for TSCS data, thus we more than often make certain modeling decisions to account for that structure.
+However, it is rarely the case that we can do this even for TSCS data, thus we more than often make modeling decisions to account for that structure.
 Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
@@
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece details how it contrasts with alternative network based approaches.
-Here we just provide a brief review and then move onto presenting a simulation exercise.
+\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network based approaches.
+Here we just provide a brief review and then move to a simulation exercise.
 
@@
 An earlier version of the LFM used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
-The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.} The AME model is specified as follows:
+The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure underlying the LFM.} The AME model is specified as follows:
 
@@
 We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
-Next, to account for the dependencies that emerge in dyadic data and that may complicate inference on the parameter associated with exogenous covariates we add parameters from the SRRM and LFM.
+Next, to account for the dependencies that emerge in dyadic data and that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM.
 $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
@@
 
-By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we draw
-
+By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw for the exogenous parameters.
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.

### 2017-10-23 paper/3_sim.tex
first cut of sim done

@@
 The goal of the simulation is to assess how well AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Specifically, we are concerned with conducting inference on a regression parameter $\beta$ of a linear model for a network in the case where there is an omitted variable.
+Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable.
 Say that the true data-generating process for a particular $Y$ is given by:
 
-\begin{align*}
-	y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j} \\
+\begin{align}
+	y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
 	\label{eqn:sim}
-\end{align*}
+\end{align}
 
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatix, $x = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $w = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
-We compare inference for $\beta$ using three models:
+where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
+$Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe.
+We compare inference for $\mu$ and $\beta$ using three models -- the latter would be of primarily theoretical concern for applied scholars:
 
@@
 	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);
-	\item and an ``oracle'' regression model that includes both $x_{i,j}$ and $w_{i,j}$. 
+	\item and an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$. 
 \end{itemize}
 
-To run the simulation we vary the size of the network under study at 50 and 100.
+The first model corresponds to the ``naive'' approach in which little is explicitly done to account for latent dependencies in dyadic data.
+In the second model, we account for dependencies in dyadic data using the AME framework described in the previous section.
+For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated in the presence of unobserved dependencies, $W$.
+In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and are able to include it in our specification for $Y$.
+The oracle case provides an important benchmark for the naive and AME approaches.
+
+For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.\footnote{The value of $\gamma$ is also set to 1, which corresponds to a an example where the $W$ character is associated with homophily.} We conduct two sets of simulations, one in which the number of actors in the network is set to 50 and the other at 100.
+In total, we ran 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimated a naive, AME, and oracle model.
+
+We compare the performance of the models first in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias} by depicting the average $\mu$ and $\beta$ estimates from the simulations for the three models.
+The panels in the left show the results for when the number of actors is set to 50 and those on the right for 100; and the top pair of panels represents the estimates for $\mu$ while the bottom pair do the same for $\beta$.
+In each case, we find that the estimates for $\mu$ and $\beta$ produced by the naive approach are notably off from their true values.
+On the other hand, the AME model performs just as well as the oracle case in estimating the true values.
 
@@
 	\centering
-	\caption{Bias in parameter when homophily is ignored.}
+	\caption{Regression parameter estimates for the naive, AME, and oracle models from 1,000 simulations.}
 	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{ameSimBias_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimBias_beta.pdf}
+	\includegraphics[width=1\textwidth]{ameSimBias_all.pdf} \\
 \end{figure}
 
-One way to compare inferences across models is  with bias and variance of the parameter estimates. Both of these quantities are combined to get the MSE.
+Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value actually fell in those intervals.
+The results are summarized in Figure~\ref{fig:ameCalib}, and here again we see that the AME approach performs as well as the oracle while the naive approach performs poorly.
 
@@
 	\centering
-	\caption{Coverage in parameter when homophily is ignored.}
+	\caption{Proportion of times when true value fell in the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
 	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{ameSimCover_mu.pdf} \\
-	\includegraphics[width=1\textwidth]{ameSimCover_beta.pdf}
+	\includegraphics[width=1\textwidth]{ameSimCover_all.pdf} \\
 \end{figure}
 
-Alternatively, bias and precision can be summarized by confidence interval coverage and width.
-Coverage should ideally be at the nominal level.
-If two methods have the same actual coverage rate, the one with the narrower intervals is preferred.
+Last, beyond obtaining less biased and better calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
+In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.
+Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that affects the degree to which they will interact with one another, i.e., $y_{ij}$.
+This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\textbf{U}^{\top} \textbf{D} \textbf{V}$.
+To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50.
+In both cases we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.
 
@@
 	\centering
-	\caption{Correlation between missing variable and multiplicative random effect in AME.}
+	\caption{Distribution of correlation between missing variable and multiplicative random effect in AME across the 1,000 simulations.
+Vertical line through the distribution represents the median value across the simulations.}
 	\label{fig:ameCorr}

### 2017-10-23 code/1_simulation/2_ameBias.R

@@ -35,11 +35,11 @@ getBiasDF = function(ameSim){
 	biasDF = lapply(ameSim, function(x){
 		betaMod=lapply(x$beta, function(z){ apply(z, 2, median)[1:2] })
 		betaMod = betaMod %>% reshape2::melt() %>%
-			mutate(
+			dplyr::mutate(
 				var=rep(c('Intercept','X1'), length(x$beta)),
 				act = rep(c(intEff, x1Eff), length(x$beta)),
 				bias = value - act ) %>%
-			rename(model = L1)
+			dplyr::rename(model = L1)
 		return(betaMod) })
 	return( suppressMessages( reshape2::melt(biasDF,id=names(biasDF[[1]])) ) ) }
 
@@ -52,15 +52,26 @@ ameSimBias = rbind(
 ##############################
 # clean
 ameSimBias = ameSimBias %>% group_by(model, var, n) %>%
-	mutate( mse = mean(bias^2) ) %>% data.frame()
+	dplyr::mutate( mse = mean(bias^2) ) %>% data.frame()
+
+# fix mod labels
 ameSimBias$model = modKey$clean[match(ameSimBias$model, modKey$dirty)]
 ameSimBias$model = factor(ameSimBias$model, levels=modKey$clean)
+modCols = c(Naive='#d6604d', AME='#4393c3', Oracle='#4daf4a')
+
+# fix var labels
 ameSimBias$var[ameSimBias$var=='X1'] = '$\\beta$'
 ameSimBias$var[ameSimBias$var=='Intercept'] = '$\\mu$'
+ameSimBias$var = factor(ameSimBias$var, levels=c("$\\mu$", "$\\beta$"))
 
-ggBiasPlot = function(varName){
-	g=ggplot(filter(ameSimBias, var==paste0('$\\',varName,'$')),
-			aes(x=model, y=value, fill=model,color=model)) +
+# viz
+ggBiasPlot = function(varName, h=4, w=8){
+	if(varName!='all'){
+		g=ggplot(
+			filter(ameSimBias, var==paste0('$\\',varName,'$')),
+			aes(x=model, y=value, fill=model,color=model)) }
+	if(varName=='all'){ g=ggplot(ameSimBias, aes(x=model,y=value,fill=model,color=model)) }
+	g = g +
 		geom_hline(aes(yintercept=act), color='grey60', size=2) +	
 		geom_jitter(alpha=.1) +
[…]

### 2017-10-23 code/1_simulation/3_ameCalib.R

@@ -61,36 +61,52 @@ rm(list=c(
 
 # summ by model
 coverSumm = ameSimCover %>% group_by(model,n, varName) %>%
-	summarise(coverage=mean(coverage)) %>% data.frame()
+	dplyr::summarise(coverage=mean(coverage)) %>% data.frame()
 ##############################
 
 ##############################
-#
+# clean mod labels
 coverSumm$model = modKey$clean[match(coverSumm$model, modKey$dirty)]
 coverSumm$model = factor(coverSumm$model, levels=modKey$clean)
+modCols = c(Naive='#d6604d', AME='#4393c3', Oracle='#4daf4a')
+
+# clean var labels
 coverSumm$varName[coverSumm$varName=='X1'] = '$\\beta$'
 coverSumm$varName[coverSumm$varName=='Intercept'] = '$\\mu$'
+coverSumm$varName = factor(coverSumm$varName, levels=c("$\\mu$", "$\\beta$"))
 
-#
-ggCoverPlot = function(var){
-	g = ggplot( filter(coverSumm, varName==paste0('$\\',var,'$')),
-			aes(x=model, y=coverage, fill=model,color=model)) +
-		geom_hline(aes(yintercept=.95), color='grey60', size=2, alpha=.5) +
+# coverage stat
+coverSumm$covLab = paste0(round(coverSumm$coverage,3)*100,'%')
+coverSumm$covLabY = ifelse(coverSumm$coverage>.9,.95,coverSumm$coverage+.03)
+
+# viz
+ggCoverPlot = function(var,h=3, w=8){
+	if(var!='all'){
+		g=ggplot(
+			filter(coverSumm, varName==paste0('$\\',var,'$')),
+			aes(x=model, y=coverage, fill=model,color=model)) }
+	if(var=='all'){ g=ggplot(coverSumm, aes(x=model, y=coverage, fill=model,color=model)) }
+	g = g +
+		geom_hline(aes(yintercept=.95), color='grey60', size=4, alpha=.5) +
 		geom_linerange(aes(ymin=0, ymax=coverage),size=1.25) + 
 		geom_point(size=2) + 
 		facet_grid(varName~n, scales='free_y',
 			labeller=as_labeller(facet_labeller, default = label_parsed)) + 			
 		xlab('') + 
+		scale_color_manual(values=modCols) +
+		scale_fill_manual(values=modCols) +		
[…]

### 2017-10-23 code/1_simulation/4_corUVandMiss.R

@@ -23,10 +23,6 @@ NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 # load sim results
 for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
-
-#
-modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
-modKey$clean = c('Naive', 'AME', 'Oracle')
 ##############################
 
 ##############################
@@ -63,8 +59,15 @@ ameSimCorDensity = ddply(ameSimCor, .(n), .fun=function(x){
 ##############################
 
 ##############################
-loadPkg('RColorBrewer')
-col = brewer.pal(3, 'Set1')[3]
+# color
+col = '#4393c3'
+
+# cleanup
+ameSimCorMeans$sMedianLab = paste0('Median Correlation: ', round(ameSimCorMeans$sMedian,3))
+ameSimCorMeans$y[ameSimCorMeans$n=='n = 100'] = 45
+ameSimCorMeans$y[ameSimCorMeans$n=='n = 50'] = 20
+
+# viz
 g = ggplot() +
 	geom_line(data=ameSimCorDensity, aes(x=x,y=y)) +
 	geom_ribbon(data=subset(ameSimCorDensity,q95),
@@ -73,6 +76,8 @@ g = ggplot() +
 		aes(x=x,ymax=y),ymin=0,alpha=0.9) + 
 	geom_vline(data=ameSimCorMeans, 
 		aes(xintercept=sMedian),linetype='solid',size=1, color=col) +	
+	geom_text(data=ameSimCorMeans,
+		aes(label=sMedianLab,x=sMedian,y=y), hjust=1.05, size=2.5, fontface='bold', color='black') +
 	facet_grid(n~.) + 	
 	xlab('') + ylab('') +
 	theme(

### 2017-10-23 code/2_codeForPaperGraphics/gibler_posterior.R

@@ -118,15 +118,19 @@ scenDiffs = rbind(
 	getScenDiff(linkType='probit', scen1, scen0, scens, ameFit$BETA, 'AME', type='densityShade') )
 
 ggCols = c(GLM='#d6604d', AME='#4393c3')
-ggLty = c(GLM='dotted', AME='solid')
-scenDiffs = scenDiffs[scenDiffs$scen %in% c('Rivalry'),]
-scenGG = ggplot(data=scenDiffs, aes(color=mod, fill=mod)) +
-	geom_line(data=scenDiffs, aes(x=x,y=y)) +
-	geom_vline(aes(xintercept=mean, color=mod,linetype=mod),size=1) +
-	geom_ribbon(data=subset(scenDiffs,q95), aes(x=x,ymax=y,fill=mod),ymin=0,alpha=0.5) +
-	geom_ribbon(data=subset(scenDiffs,q90), aes(x=x,ymax=y,fill=mod),ymin=0,alpha=0.9) +
+ggLty = c(GLM='dashed', AME='solid')
+scenDiffsSlice = scenDiffs[scenDiffs$scen %in% c('Rivalry'),]
+scenGG = ggplot(data=scenDiffsSlice, aes(color=mod, fill=mod)) +
+	geom_line(data=scenDiffsSlice, aes(x=x,y=y)) +
+	geom_ribbon(data=subset(scenDiffsSlice,q95), aes(x=x,ymax=y,fill=mod),ymin=0,alpha=0.2) +
+	geom_ribbon(data=subset(scenDiffsSlice,q90), aes(x=x,ymax=y,fill=mod),ymin=0,alpha=0.6) +
+	geom_vline(aes(xintercept=mean, color=mod,linetype=mod),size=1.2) +	
 	scale_color_manual(values=ggCols) + scale_fill_manual(values=ggCols) +
 	scale_linetype_manual(values=ggLty) +
+	guides(
+		linetype=guide_legend(override.aes = list(size=.5)),
+		fill=guide_legend(override.aes = list(fill='transparent'))
+		) +
 	xlab('Pr(MID=1 | Rivalry=1) - Pr(MID=1 | Rivalry=0)') +
 	ylab('Density') +
 	facet_wrap(~scen, scales='free', ncol=1) +
@@ -137,5 +141,6 @@ scenGG = ggplot(data=scenDiffs, aes(color=mod, fill=mod)) +
 		strip.text.x = element_text(size = 9, color='white' ),
[…]

### 2017-10-23 paper/4_app.tex
app and concl edits

@@
 Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
-Also part of the gold standard is a post estimation cleanup thought to be a palliative, and each of these studies adjust the posterior standard errors in an attempt to account for the clustering of observations.
+Also each of these studies adjust the posterior standard errors in an attempt to account for the clustering of observations and in our re-estimation of the original model in the paper we follow that strategy as well.
 
 \begin{table}
-\caption{Features of the Studies Re-estimated. }
+\caption{Features of the Studies Re-estimated.}
 	\begin{tabular}{lcccccc}
@@
 We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
+Our goal is to ascertain whether the ignored interdependencies -- the non-iid structure of the relational data -- would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
 Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
 
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies that use a framework that ignores dependencies.
-Based in part our our results as well as the statistical and mathematical characteristics of the AME framework, we believe that it does.
+The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learn about international relations from recent empirical studies using a framework that ignores dependencies.
+Based in part on our results as well as the statistical characteristics of the AME framework, we believe that it does.
 The dynamic latent factor model provides a step forward in the modeling of international relations.
 
+% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
 Given current practices, most scholars have a single variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models.
-Whether putative control variables, for example, are precisely estimated is less important.
 Generally, interpretations focus on a small set of independent variables.
-In the study by Reiter and Stam (2003) the main question is to determine in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---whether it is the democratic country or the personalist country that is associated with a greater likelihood that the dyad is in a dispute.
-In a regression including fourteen coefficients, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our evaluation we focus on the same question in the re-examination of the Reiter and Stam study to determine whether the model replicates when re-estimated within the AME framework.
+In the study by Reiter and Stam (2003) the goal is to determine in mixed dyads -- consisting of a democratic country and country ruled by a personalist dictatorship -- whether it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
+In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
 Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not also found in the AME estimation.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.
 
@@
 
-\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-
-We set a number of benchmarks for comparison.
-First, we compare the AME model to a GLM model (that used in the original study) using the same covariates to show the effect of accounting for network dependencies on predicting conflict.
-We supplement this with an alternative GLM that includes not just these covariates, but also a lagged dependent variable and a lagged reciprocity term.
-The lagged dependent variable is the equivalent of saying that conflict and peace are relatively likely to persist between dyads, while the inclusion of a lagged reciprocity term in a GLM framework is a simple way to account for retaliatory strikes.
-
-We utilize three performance criteria to compare the models: Receiver Operator Characteristic (ROC) curves, Precision Recall (PR) curves, and separation plots.
-ROC curves look at the trade-off between true positive rates and false positive rates at different thresholds of classification.
-An issue with an ROC Curve when looking at conflict, is that it is relatively rare at the dyadic level: in most years only 3\% of possible dyads are in conflict with one another.
-If peace is common, even a poor model will have a very low False Positive Rate.
-The ROC curves are best for balanced samples with as many events as non-events.
-To better assess which models predict the presence of conflict, not just its absence, we look at PR Curves.
-These examine the trade-offs between the percentage of conflicts a model predicts, and the percentage of predicted conflicts which occur.
-Lastly, we examine separation plots \citep{greenhill:etal:2011}.
-These provide an intuitive visualization of the accuracy of our predictions by juxtaposing a line showing the predicted probability of conflict with whether conflict actually occurs for all cases (where the cases are sorted by the predicted probability and then colored to indicate the outcome).
-Here a perfect model would have all cases where conflict actually exists on the right with a predicted probability of $1$, and would predict $0$ in all other cases.
-All of the models' performance out of sample by these metrics are displayed in figure \ref{fig:roc_ame}.
-The AME model with covariates is the best performing model out of sample in all cases.
-This model outperforms each of the GLM variants by a notable margin.
+\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.\footnote{For more details on this type of cross-validation strategy see \citet{minhas:etal:2016:arxiv}.} We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
+For the binary models we just provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
+Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out of sample predictive performance.
 
@@
 		~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
-\multirow{2}{*}{Rose (2004)} & Root Mean Squared Error RMSE & 3.23 &  {1.99} \\
-		~ & Root Median Squared Error RMDSE & 2.01 &  {1.06} \\	\midrule
+\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
+		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
 \multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
@@
 They use the MID data, but note that
+
 \begin{quote}
@@
 \end{quote}
+
 Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
@@
 Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
+Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.\footnote{Further details on each of the regression parameter estimates from the replication can be found in the Appendix.}
 
@@
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
+Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO -- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
 
 When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
-The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---persists in the AME model.
+The main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- persists in the AME model.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in
-Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
 In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
@@
 
-Insert coefplot here, without splines...\label{fig:giberrepl}
-%
-%\begin{table}
-%\begin{center}
-%\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
-%\begin{tabular}{lrr} \toprule
-%& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-%Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-%Allied & 0.142&  0.023 \\
-%Joint Democracy &\bf -0.507&  0.045 \\
-%Peace Years \bf &-0.260& \bf -0.060 \\
-%Spline 1 &\bf -0.001& \bf 0.00 \\
-%Spline 2&\bf -0.000& \bf 0.00\\
-%Spline 3 &-0.000& \bf 0.00\\ 
-%Contiguity &\bf 2.412& \bf 0.640 \\
-%Parity &0.075&  -0.013 \\
-%Parity at entry year&\bf 0.868&  0.002 \\ 
-%Rivalry &\bf 2.031&\bf 0.721\\ 
-%Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-%\end{tabular}\\
-%\end{center}
-%{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
+% SM: we still adding a coefplot here?
+% Insert coefplot here, without splines...\label{fig:giberrepl}
+% %
+% %\begin{table}
+% %\begin{center}
+% %\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
+% %\begin{tabular}{lrr} \toprule
+% %& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
+% %Variable & \texttt{logit} & \texttt{amen} \\ \midrule
+% %Allied & 0.142&  0.023 \\
+% %Joint Democracy &\bf -0.507&  0.045 \\
+% %Peace Years \bf &-0.260& \bf -0.060 \\
+% %Spline 1 &\bf -0.001& \bf 0.00 \\
+% %Spline 2&\bf -0.000& \bf 0.00\\
+% %Spline 3 &-0.000& \bf 0.00\\ 
+% %Contiguity &\bf 2.412& \bf 0.640 \\
+% %Parity &0.075&  -0.013 \\
+% %Parity at entry year&\bf 0.868&  0.002 \\ 
+% %Rivalry &\bf 2.031&\bf 0.721\\ 
+% %Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
+% %\end{tabular}\\
+% %\end{center}
+% %{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
 For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
 NEED TO BE FIXED if staying.
-%\end{table}
+% %\end{table}
 
-We re-estimated model 6 from Table 6 (2017, 34). The results are presented in the Appendix, along with each of the models we estimated.
+We re-estimated model 6 from Table 6 (2017, 34).
+The results are presented in the Appendix.
 The results obtained with AME stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
 Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
-Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$).
+Not only is the value of this parameter small, but it has a very large relative standard error -- over a magnitude larger than the parameter itself ($z= 0.038$).
 In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.
@@
 As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notable.
-First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
+First, the expected value of the dependent variable -- the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
 These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
@@
 
-Beyond more informative fixed-effects coefficients, the AME approach also provides information about the interdependencies that were modeled.
-The assumption of independence among the dyadic data in this study can be strongly rejected.
-Perhaps most importantly, the main substantive conclusions of the 2017 Gibler study do not seem warranted from the perspective of the results obtained with the AME estimation that explicitly models \first- \second-, and \third-order interdependencies.
-Not only do the estimated coefficients tell a different story, but the assumptions under which the original results would hold are shown to be violated by the data.
-
 \subsection{Lessons Learned from Re-estimating Five Prominent Studies}
 
-First and foremost, many  findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.
-This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.
-Not only
-are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
+First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news -- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.
 
-Second, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights
-emerge from the additional information derived.
+Second, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights emerge from the additional information derived.
 In particular, there is actual information about the dependencies so that clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.
-This kind of information
-is absent in standard approaches and add to our ability to explain specific as well as general results.
+This kind of information is absent in standard approaches and add to our ability to explain specific as well as general results.
 
-Third, it is evident that the actual results---not the estimated coefficents and their covariances---which are generated by the models differ greatly in expectations.
+Third, it is evident that the actual results -- not the estimated coefficents and their covariances -- which are generated by the models differ greatly in expectations.
 This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.
@@
 Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).
-In the case of studies
-with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.
+In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.
 It is rare for studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.
 
-Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are independent and identically distributed.
+Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are iid.

### 2017-10-23 paper/5_concl.tex
app and concl edits

@@
 Some approaches focus on only looking a at small number of these actors, but many scholars examine a large number of actors at a time.
-This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.
-One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.
-It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citet{erikson:pinto:2014,aronow:etal:2015}.
+This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citet{erikson:pinto:2014,aronow:etal:2015}.
 
-At the same time, we know that research designs which focus on the statistical analysis of dyadic data quickly go astray if the dyadic data are assumed to be independent and identically distributed (iid).
-Virtually all of the standard statistical models---ordinary least squares, logistic and probit regressions, to name a few---fail if the data are not iid.
+At the same time, we know that research designs focusing on the statistical analysis of dyadic data quickly go astray if the dyadic data are assumed to be iid.
+Virtually all of the standard statistical models---ordinary least squares and logistic regressions, to name a few---fail if the data are not iid.
 By definition dyadic data are not iid and thus the standard approaches can not be used cavalierly to analyze these data.
@@
 The statistical model of the latent network captures first-order (example or restatement), second-order (example or restatement), and third-order (example or restatement) dependencies in dyadic data using a familiar regression framework that has been adapted for relational data---such as dyadic data---which are not independent nor identically distributed.
-There is an available, open source computer packages that implements this approach.
 
-To explore this approach in the context of international relations we conduct two broad analyses.
+To explore this approach in the context of international relations we have presented two broad analyses.
 The first is a simulation of where the characteristics of the network are known.
-This shows that the AME approach is unbiased in terms of parameter estimation compared with standard approach employed in
-international relations to study dyadic data (i.e., GLM models).
+This shows that the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
 The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.

### 2017-10-23 paper/2_model.tex
appendix touchups

@@
 As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
-To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix -- often referred to as an adjacency matrix -- as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix --- often referred to as an adjacency matrix --- as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.

### 2017-10-23 paper/4_app.tex
appendix touchups

@@
 We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies -- the non-iid structure of the relational data -- would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
+Our goal is to ascertain whether the ignored interdependencies --- the non-iid structure of the relational data --- would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
 Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
@@
 Generally, interpretations focus on a small set of independent variables.
-In the study by Reiter and Stam (2003) the goal is to determine in mixed dyads -- consisting of a democratic country and country ruled by a personalist dictatorship -- whether it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
+In the study by Reiter and Stam (2003) the goal is to determine in mixed dyads --- consisting of a democratic country and country ruled by a personalist dictatorship --- whether it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
 In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
@@
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO -- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
+Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO --- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
 
 When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
-The main result of the model -- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables -- persists in the AME model.
+The main result of the model --- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables --- persists in the AME model.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample -- $3.23$ for the OLS model and $1.77$ for the AME model.
+When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
 While the basic model replicates with the basic result, there are substantial and substantive improvements gleaned from the AME results.
@@
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence -- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence --- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
 In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
@@
 Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
-Not only is the value of this parameter small, but it has a very large relative standard error -- over a magnitude larger than the parameter itself ($z= 0.038$).
+Not only is the value of this parameter small, but it has a very large relative standard error --- over a magnitude larger than the parameter itself ($z= 0.038$).
 In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.
@@
 As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notable.
-First, the expected value of the dependent variable -- the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
+First, the expected value of the dependent variable --- the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
 These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
@@
 
-First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news -- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
+First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news --- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.
@@
 
-Third, it is evident that the actual results -- not the estimated coefficents and their covariances -- which are generated by the models differ greatly in expectations.
+Third, it is evident that the actual results --- not the estimated coefficents and their covariances --- which are generated by the models differ greatly in expectations.
 This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.

### 2017-10-23 paper/6_appendix.tex
appendix touchups

@@
 
-\subsubsection*{Gibler (2017)}
+For each of the replications involving a binary dependent variable we provide a table of coefficient estimates that includes the original GLM estimation with a logit link, a GLM estimation with a probit link, and the AME model.
+The GLM estimation with a probit link function is provided so as to ease comparison between the AME model, which is also based on the probit link.
+For Rose (2004) we just provide a table of coefficient estimates that includes the original OLS estimation and the AME model.
 
-Coefficient plot and performance comparison.
+Additionally, for each replication we provide a more detailed visualization illustrating the results of our out of sample performance analysis.
+For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.
 
-\input{gibler_coeftable}
+\clearpage
+\subsubsection*{Reiter \& Stam (2003)}
+
+Additional information for the Reiter \& Stam (2003) re-estimation.
+
+\input{reiter_stam_coeftable}
 \FloatBarrier
@@
 	\centering   
-	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves and PR curves.}
+	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves, PR curves, and separation plots.}
 \end{figure}
 \FloatBarrier
+\clearpage
 
-\clearpage
 \subsubsection*{McDonald (2004)}
 
-Coefficient plot and performance comparison.
+Additional information for the McDonald (2004) re-estimation.
 
@@
 	\subfigure[Precision and Recall]{\label{fig:mcdonaldpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
+	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves, PR curves, and separation plots.}
 \end{figure}
 \FloatBarrier
+\clearpage
 
-\clearpage
-\subsubsection*{Reiter \& Stam (2003)}
+\subsubsection*{Rose (2004)}
 
-Coefficient plot and performance comparison.
+Additional information for the Rose (2004) re-estimation.
 
-\input{reiter_stam_coeftable}
+\input{rose_coeftable}
 \FloatBarrier
@@
 	\centering   
-	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves and PR curves.}
+	\includegraphics[width=1\textwidth]{rose_outSample.pdf}
+	\caption{Assessments of out-of-sample predictive performance for Rose (2004) using root mean squared error (RMSE), root median squared error (RMDSE), and $R^{2}$.}
 \end{figure}
 \FloatBarrier
+\clearpage
 
-\clearpage
-\subsubsection*{Rose (2004)}
-
-Coefficient plot and performance comparison.
-
-\input{rose_coeftable}
-\FloatBarrier
-
-% \begin{figure}
-% 	\centering   
-% 	\subfigure[AUC]{\label{fig:reitaucpr}\includegraphics[width=.45\textwidth]{mcdonald_roc_outSample.pdf}}
-% 	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
-% 	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves and PR curves.}
-% \end{figure}
-\FloatBarrier
-
-\clearpage
 \subsubsection*{Weeks (2012)}
 
-Coefficient plot and performance comparison.
+Additional information for the Weeks (2012) re-estimation.
 
@@
 % 	\subfigure[Precision and Recall]{\label{fig:weekspr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
-% 	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves.}
+% 	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves, PR curves, and separation plots.}
 % \end{figure}
 % \FloatBarrier
+\clearpage
+
+\subsubsection*{Gibler (2017)}
+
+Additional information for the Gibler (2017) re-estimation.
+
+\input{gibler_coeftable}
+\FloatBarrier
+
+\begin{figure}
+	\centering   
+	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
+	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
+	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
+\end{figure}
+\FloatBarrier
+\clearpage
+
+\subsection*{AME Tutorial}
+
+Using the AMEN function requires formatting data into a particular structure. The primary distinction in data formatting is whether the outcome of interest represents a directed or undirected network.
+
+If undirected, the AMEN function has three main inputs:
+
+\begin{itemize}[noitemsep,nolistsep]
+    \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network.
+    \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset. 
+    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
+\end{itemize}
+
+If directed, AMEN further calls for:
+
+\begin{itemize}[noitemsep,nolistsep]
+    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
+    \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset. 
+\end{itemize}
+
+Beyond the data inputs, the AMEN function requires additional specification:
+
+\begin{itemize}
+    \item model: how to model the outcome variable, e.g., `logit'
+    \item symmetric: whether the input network is symmetric
+    \item intercept: whether to estimate an intercept
+    \item nscan: number of iterations of the Markov chain
+    \item burn: burn-in period
+    \item odens: thinning interval
+    \item R: dimension of the multiplicative effect (referred to as $K$ in the paper)
+    \item gof: whether to calculate goodness of fit statistics
+\end{itemize}
+
+There is often little theoretical reason to choose a particular value of $R$ (above).
+One strategy is to estimate models at different values of $R$ and compare goodness of fit statistics across models.
+
+Given the computational intensity needed for parameter estimates to converge, parallelization strategies are recommended to speed up analysis.
+In addition, providing AMEN function with starting values, either dictated by theory, previous research, or previous runs can also help speed up convergence time.
+
+The code below presents an example of an AME model running in parallel across 4 different levels of $R$.
+Note also that the model is using starting values from a previous run, defined in \textit{startVals0}.
+
+\begin{lstlisting}[language=R]
+# running in parallel varying k
+imps = 10000 ; brn = 25000 ; ods = 10 ; latDims = 0:3
+
+# Run amen in parallel
+library(doParallel) ; library(foreach) ; cl=makeCluster(4) ; registerDoParallel(cl)
+foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {
+
+  # load previous model run
+  load(prevModelFiles[ii])
+  # extract start vals
+  startVals0 = ameFit$'startVals'
+  # dump rest
+  rm(ameFit)
+
+  ameFit = ame_repL(
+    Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL, 
+    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii], 
+    nscan=imps, seed=1, burn=brn, odens=ods, 
+    plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0,
+    periodicSave=TRUE )     
+  save(ameFit, file=paste0('model_k', latDims[ii],'_v2.rda') )
+}
+stopCluster(cl)
+\end{lstlisting}

### 2017-10-23 code/2_codeForPaperGraphics/rose_posterior.R

@@ -15,44 +15,89 @@ load(paste0(resultsPath,'outputData/ameFit_k2_v2_imps_50000_intercept.rda'))
 load( paste0(resultsPath,'glmResults_rose.rda') )
 ############################################
 
-# coefSumm ###########################################
-ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-lmSumm = baseModelSumm[,]
+# # coefSumm ###########################################
+# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+# lmSumm = baseModelSumm[,]
 
-modList = list(lmSumm, ameSumm)
-modNames = c('LM', 'AME')
+# modList = list(lmSumm, ameSumm)
+# modNames = c('LM', 'AME')
 
-varKey = data.frame(
-	dirty=rownames(ameSumm),
-	clean=c('Intercept', 'Both in GATT/WTO', 'One in GATT/WTO', 'GSP',
-		'Log Distance',  'Log Product Real GDP', 'Log Product Real GDPpc',
-		'Regional FTA', 'Currency Union', 'Common language', 'Land Border',
-		'Number Landlocked' ,'Number Islands', 'Log Product Land Area',
-		'Common Colonizer','Currently Colonized', 'Ever Colony','Common Country' ),
[…]

### 2017-10-23 paper/3_sim.tex
appendix touchups
share for comments

@@
 $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe.
-We compare inference for $\mu$ and $\beta$ using three models -- the latter would be of primarily theoretical concern for applied scholars:
+We compare inference for $\mu$ and $\beta$ using three models --- the latter would be of primarily theoretical concern for applied scholars:
 
@@
 In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.
-Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that affects the degree to which they will interact with one another, i.e., $y_{ij}$.
+Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$.
 This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\textbf{U}^{\top} \textbf{D} \textbf{V}$.

### 2017-10-24 paper/1_intro.tex
edited manuscript

@@
 Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
-However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
+However, this approach to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
 Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases.
@@
 First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict.
-Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; specifically, if a rebel group initiates a conflict with a government, the government will likely reciprocate that conflictual behavior.
+Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
 We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
@@
 Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
-The latent factor approach (AME) is able to better capture first, second, and third-order interdependencies than the standard approach.
+The latent factor approach (AME) is better able to capture first, second, and third-order interdependencies than the standard approach.
 It also produces results that are, at times, at odds with those found in these studies in particular, and the broader literatures from which they are drawn.
@@
 Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
-Most importantly, it facilitates the concentration on the relations aspect of the field of international relations.
+Most importantly, it facilitates concentration on the relations aspect of the field of international relations.

### 2017-10-24 paper/2_model.tex
edited manuscript

@@
 This makes sense if each observation is independent of the others.
-Thus, for example, a conflict sent from the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
+For example, a conflict sent from the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
 Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
-As a result, the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others.
-We know this assumption to be false in theory and in practice because relational data comes with an explicit structure that in general leads to particular types of dependencies.
+While the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others, we know this assumption to be false in theory and in practice.
+Relational data comes with an explicit structure that in general leads to particular types of dependencies.
 The importance of accounting for the underlying structure of our data has been a lesson well understood, at least, when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
@@
 The cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns.
-Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flows, that is sent from actor $i$ to actor $j$.
+Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flow, that is sent from actor $i$ to actor $j$.
 
@@
 \end{align}
-
 where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver).

### 2017-10-24 paper/3_sim.tex
edited manuscript

@@
 \end{align}
-
 where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
 $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe.
-We compare inference for $\mu$ and $\beta$ using three models --- the latter would be of primarily theoretical concern for applied scholars:
+We compare inference for $\mu$ and $\beta$ -- the latter parameter would be of primarily theoretical concern for applied scholars -- using three models :
 
@@
 Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value actually fell in those intervals.
-The results are summarized in Figure~\ref{fig:ameCalib}, and here again we see that the AME approach performs as well as the oracle while the naive approach performs poorly.
+The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle while the naive approach performs poorly.
 

### 2017-10-24 paper/4_app.tex
edited manuscript

@@
 Generally, interpretations focus on a small set of independent variables.
-In the study by Reiter and Stam (2003) the goal is to determine in mixed dyads --- consisting of a democratic country and country ruled by a personalist dictatorship --- whether it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
+For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads --- consisting of a democratic country and country ruled by a personalist dictatorship --- it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
 In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
@@
 			& &  in a Network Model? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Replicates} \\ \midrule
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
+		\bf{McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}}\\ \midrule
 		Rose (2004) & WTO Membership Does not Effect Trade & {Replicates}\\ \midrule
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
-		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate} \\ \bottomrule
+		\bf{Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate}} \\\midrule
+		\bf{Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate} }\\ \bottomrule
 	\end{tabular}
@@
 To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-Then for each $s \in \{1,\ldots,k\}$, we:
+To evaluate our model, we randomly divide the data  into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.\textcolor{red}{This is inaccurate for directed dyads.
+I think.
+MG} Then for each $s \in \{1,\ldots,k\}$, we:
 
@@
 The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
-Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
+%Cases in which our network based approach outperformed the standard approach are highlighted in bold.
+	}
 	\label{tab:modelPerfSumm}
@@
 It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.
-Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
+Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
@@
 When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
-While the basic model replicates with the basic result, there are substantial and substantive improvements gleaned from the AME results.
+While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.
 
@@
 The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-Her time frame is from 1946-1999.
+%Her time frame is from 1946-1999.
 The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
@@
 When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
-We similarly find less pronounced effects for military capabilities.
+Similarly, we find less pronounced effects for military capabilities.
 One explanation for this divergence is the AME model's ability to account for third order effects.
 Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence --- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
+We also observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
 In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
@@
 In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
-In terms of performance, the original Weeks' model is slightly better than a coin-toss overall, and fails to indicate more than a couple of events (incorrectly) in almost one million observations.
+In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.
 
@@
 
-First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news --- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
+First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news --- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.
@@
 
-Fourth, it is clear that the AME approach dominates the OLS and Logit-based approaches in terms of performance.
+Fourth, it is clear that the AME approach dominates the GLM approaches in terms of performance.
 Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).

### 2017-10-24 paper/5_concl.tex
edited manuscript

@@
 Virtually all of the standard statistical models---ordinary least squares and logistic regressions, to name a few---fail if the data are not iid.
+This fact has been accepted when it comes to temporal dependencies, but adoption of methods to account for network dependencies have seen less progress.
 By definition dyadic data are not iid and thus the standard approaches can not be used cavalierly to analyze these data.
@@
 These approaches have been developed for a while, but are not yet widely used in international relations scholarship.
-The statistical model of the latent network captures first-order (example or restatement), second-order (example or restatement), and third-order (example or restatement) dependencies in dyadic data using a familiar regression framework that has been adapted for relational data---such as dyadic data---which are not independent nor identically distributed.
+The latent network model captures first-order dependencies (the tendencies of certain actors to send and receive more of a certain action), second-order dependencies (the tendency of actors to reciprocate actions), and third-order dependencies (the ability for actions within a dyad to effect a third party) in dyadic data using a familiar regression framework that has been adapted for relational data---such as dyadic data---which are not independent nor identically distributed.
 
@@
 The first is a simulation of where the characteristics of the network are known.
-This shows that the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
+This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
 The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
@@
 This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
-At the same time, the latent factor model provides a way to easily examine and if necessary defeat these assumptions.
 

### 2017-10-24 paper/2_model.tex
small stuff.

@@
 As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
-To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix --- often referred to as an adjacency matrix --- as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 		& a_{i} + b_{j} + \epsilon_{ij} + \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
-		& \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
+		& \mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
 	\label{eqn:ame}
@@
 
-The LFM contribution to the AME comes in the multiplicative term: $\textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$.
+The LFM contribution to the AME comes in the multiplicative term: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$.
 $K$ denotes the dimensions of the latent space.
@@
 The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network.
-When taking the SVD we factorize our observed network into the product of three matrices: $\textbf{U}, \textbf{D}, \text{ and }, \textbf{V}$.
-This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\textbf{U}$ and $\textbf{V}$ are $n \times K$ and $\textbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\textbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network, or more simply put how similar actors are in terms of who they are initiating, for example, conflict with.
-$\hat{\textbf{u}}_{i} \approx \hat{\textbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors.
-$\textbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\textbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
+When taking the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$.
+This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network, or more simply put how similar actors are in terms of who they are initiating, for example, conflict with.
+$\hat{\mathbf{u}}_{i} \approx \hat{\mathbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors.
+$\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
+The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
 
@@
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
-Specifically, given initial values of $\{\bm\beta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
+Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
 
  \begin{itemize}
- 	\item sample $\bm\theta \; | \;  \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 	\item sample $\bm\beta \; | \;  \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 	\item sample $\textbf{a}, \textbf{b} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-	\item sample $\Sigma_{ab} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
+ 	\item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 	\item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 	\item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+	\item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
  	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
- 	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
+ 	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
  	\item For each $k \in K$:
  	\begin{itemize}
- 		\item Sample $\textbf{U}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}_{[,-k]}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\textbf{V}_{[,k]} \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\textbf{D}_{[k,k]}  \; | \; \bm\beta, \textbf{X}, \bm\theta, \textbf{a}, \textbf{b}, \textbf{U}, \textbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \textbf{D} matrix is absorbed into the calculation for $\textbf{V}$ as we iterate through $K$.
+ 		\item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 		\item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+ 		\item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \mathbf{D} matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
 }

### 2017-10-24 paper/3_sim.tex
small stuff.

@@
-\section{Simulation study}
+\section{\textbf{Simulation study}}
 
@@
 $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe.
-We compare inference for $\mu$ and $\beta$ -- the latter parameter would be of primarily theoretical concern for applied scholars -- using three models :
+We compare inference for $\mu$ and $\beta$---the latter parameter would be of primarily theoretical concern for applied scholars---using three models :
 
 \begin{itemize}
-	\item the standard IR approach assuming independent errors; 
+	\item the standard international relations approach assuming independent errors; 
 	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);
@@
 Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$.
-This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\textbf{U}^{\top} \textbf{D} \textbf{V}$.
+This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$.
 To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50.

### 2017-10-24 paper/4_app.tex
small stuff.

@@
-\section{Re-estimation with AME}
+\section{\textbf{Re-estimation with AME}|
 
@@
 We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies --- the non-iid structure of the relational data --- would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
+Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
 Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
@@
 Generally, interpretations focus on a small set of independent variables.
-For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads --- consisting of a democratic country and country ruled by a personalist dictatorship --- it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
+For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
 In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
@@
 For the binary models we just provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
-Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out of sample predictive performance.
+Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
 
@@
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO --- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
+Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
 
 When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
-The main result of the model --- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables --- persists in the AME model.
+The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---persists in the AME model.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence --- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
 We also observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
@@
 The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.
-In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
+In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
 In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.
@@
 Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
-Not only is the value of this parameter small, but it has a very large relative standard error --- over a magnitude larger than the parameter itself ($z= 0.038$).
+Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$).
 In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.
@@
 As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notable.
-First, the expected value of the dependent variable --- the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
+First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
 These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
@@
 
-First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news --- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
+First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.
@@
 
-Third, it is evident that the actual results --- not the estimated coefficents and their covariances --- which are generated by the models differ greatly in expectations.
+Third, it is evident that the actual results---not the estimated coefficents and their covariances---which are generated by the models differ greatly in expectations.
 This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.

### 2017-10-24 paper/6_appendix.tex
small stuff.

@@
 
-Additionally, for each replication we provide a more detailed visualization illustrating the results of our out of sample performance analysis.
+Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis.
 For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.
@@
 
-If directed, AMEN further calls for:
+If directed, AMEN further requires:
 

### 2017-10-24 paper/1_intro.tex
more tiny changes
the coeff tables need to have the st. errors put in math mode $$; OR, we need to take the estimated coefficients out of math mode.

@@
 
-In this article, we discuss a Bayesian approach, the Additive and Multiplicative Effects (AME) model, for directly modeling relational data to reflect the data generating process that yields interdependencies in these types of data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we discuss the Additive and Multiplicative Effects (AME) model, 
+a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2017-10-24 paper/2_model.tex
more tiny changes
the coeff tables need to have the st. errors put in math mode $$; OR, we need to take the estimated coefficients out of math mode.

@@
  		\item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, \mathbf{D} matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
+ 		\item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
 }

### 2017-10-24 paper/4_app.tex
more tiny changes
the coeff tables need to have the st. errors put in math mode $$; OR, we need to take the estimated coefficients out of math mode.

@@
-\section{\textbf{Re-estimation with AME}|
+\section{\textbf{Re-estimation with AME}}
 
@@
 		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
-		\bf{McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}}\\ \midrule
+		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
 		Rose (2004) & WTO Membership Does not Effect Trade & {Replicates}\\ \midrule
-		\bf{Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate}} \\\midrule
-		\bf{Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate} }\\ \bottomrule
+		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
+		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
 	\end{tabular}
@@
 To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the data  into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.\textcolor{red}{This is inaccurate for directed dyads.
-I think.
-MG} Then for each $s \in \{1,\ldots,k\}$, we:
+To evaluate our model, we randomly divide the data  into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
+%\textcolor{red}{This is inaccurate for directed dyads. I think. MG} 
+Then for each $s \in \{1,\ldots,k\}$, we:
 

### 2017-10-24 paper/1_intro.tex
small changes throughout

@@
 
-\citet{aronow:etal:2015} estimate that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.
 Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
-However, this approach to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
+However, this approach to studying dyadic data increases the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
 Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases.
-This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process that remains a threat to inference because of the interdependence of observations or measurements.
+Such approaches may work in limited situations, but are not generally effective because these palliatives do not address the fundamental data generating process that remains a threat to inference because of the interdependence of observations or measurements.
 Namely, it is not just the diagonals of the variance-covariance matrix this affects.
@@
 We begin with a brief review of these dependencies and the AME model.
-Next, we conduct a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
+Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
 Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
 The latent factor approach (AME) is better able to capture first, second, and third-order interdependencies than the standard approach.
-It also produces results that are, at times, at odds with those found in these studies in particular, and the broader literatures from which they are drawn.
+It also produces results that are, at times, at odds with those found in the replicated studies as well as the broader literature from which they are drawn.
 As such, this approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
 Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
-Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
+Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
 Most importantly, it facilitates concentration on the relations aspect of the field of international relations.

### 2017-10-24 paper/2_model.tex
small changes throughout

@@
 \section{\textbf{Dependencies in Dyadic Data}}
-
-In working with relational data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another.
+In working with relational data, scholars typically begin working with relational data by structuring it as a set of dyadic observations stacked on top of one another.
 This makes sense if each observation is independent of the others.
@@
 As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
-To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure of relational data, it is helpful to shift towards restructuring dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
-Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor may be more active than others, parameter estimates from standard statistical models will be biased.
+Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.
 
@@
 A prominent example of a network with this type of structure was found in American Politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
-\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs just linked to others that were political similar.
+\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs linked to others that were politically similar.
 This showcases the types of higher-order dependencies that can emerge in relational data.
-First, the fact that interactions was determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}.
+First, the fact that interactions were determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}.
 Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
@@
 \citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network based approaches.
-Here we just provide a brief review and then move to a simulation exercise.
+Here we provide a brief review and then move to a simulation exercise.
 
@@
 When taking the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$.
-This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network, or more simply put how similar actors are in terms of who they are initiating, for example, conflict with.
+This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network or, for example, how similar actors are in terms of who they initiate conflict with.
 $\hat{\mathbf{u}}_{i} \approx \hat{\mathbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors.

### 2017-10-24 paper/4_app.tex
small changes throughout

@@
 Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
-Also each of these studies adjust the posterior standard errors in an attempt to account for the clustering of observations and in our re-estimation of the original model in the paper we follow that strategy as well.
+In addition, each of these studies adjusts the posterior standard errors in an attempt to account for the clustering of observations.
+We follow the strategy of the original study in our re-estimation of the original models.
 
@@
 Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-Finally, in each study we assess whether there is any new substantive finding that emerges or indeed if any disappear once the interdependencies in the data are modeled.
+Finally, in each study we assess whether any new substantive findings emerge or if any previous findings disappear once the interdependencies in the data are modeled.
 
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learn about international relations from recent empirical studies using a framework that ignores dependencies.
+The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies using a framework that ignores dependencies.
 Based in part on our results as well as the statistical characteristics of the AME framework, we believe that it does.
@@
 For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
-In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
+In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
 Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.
@@
 In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that
+They use the MID data, but note that:
 
@@
 \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
+His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
 \citet{mcdonald:2004} measures free trade in two ways.
@@
 In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.
+It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007,ward:etal:2013}.
 Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
 He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
+Estimating this model using OLS within many differing contexts, his conclusion is that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
 
@@
 
-The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model.
+The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the GLM results and those of the AME Model.
 In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
@@
 The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.
-As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notable.
+As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notably.
 First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
@@
 In particular, there is actual information about the dependencies so that clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.
-This kind of information is absent in standard approaches and add to our ability to explain specific as well as general results.
+This kind of information is absent in standard approaches and adds to our ability to explain specific as well as general results.
 
-Third, it is evident that the actual results---not the estimated coefficents and their covariances---which are generated by the models differ greatly in expectations.
+Third, it is evident that the actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.
 This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.
@@
 In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.
-It is rare for studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.
+It is rare for studies in this area to provide performance statistics, but at the same time at least one of the studies is unable to identify a single case in spite of having almost a million observations.
 

### 2017-10-24 paper/5_concl.tex
small stuff.
small changes throughout

@@
-\section{Conclusion}
+\section{\textbf{Conclusion}}
 
@@
 Some approaches focus on only looking a at small number of these actors, but many scholars examine a large number of actors at a time.
-This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citet{erikson:pinto:2014,aronow:etal:2015}.
+This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
 
@@
 To explore this approach in the context of international relations we have presented two broad analyses.
-The first is a simulation of where the characteristics of the network are known.
+The first is a simulation where the characteristics of the network are known.
 This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
@@
 These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
-We then reanalyzed each study using the latent network approach which captures that additive and multiplicative aspects of interdependencies among the dyadic data.
+We then reanalyzed each study using the latent network approach which captures additive and multiplicative aspects of interdependencies among the dyadic data.
 In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
-This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they only obtain under stringent assumptions that can not possibly be valid.
+This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that can not possibly be valid.
 This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.

### 2017-10-24 paper/netsMatter.tex
starting to rework the app section.
stargazer tables
more clean up we are at 99%
first cut of sim done
appendix touchups
share for comments
more tiny changes
the coeff tables need to have the st. errors put in math mode $$; OR, we need to take the estimated coefficients out of math mode.
small changes throughout

@@
 % Fonts
+% at some point figure out bolding ...
 \usepackage[default,osfigures,scale=0.95]{opensans}
@@
 
-% % add watermark
-% \usepackage{draftwatermark}
-% \SetWatermarkText{Draft: Not Ready}
-% \SetWatermarkScale{.5}
-% \SetWatermarkColor[gray]{0.9}
+% add watermark
+\usepackage{draftwatermark}
+\SetWatermarkText{Version 1.0}
+\SetWatermarkScale{.5}
+\SetWatermarkColor[gray]{0.9}
 
@@
 
+% outside code
+\usepackage{listings}
+\usepackage{enumitem}
+
 % define clickable links and their colors
@@
 
-\title{International Relations are a Social Network:\\ A Latent Network Approach for Global Politics}
+\title{Networks Matter in International Relations:\\A Latent Network Approach to Dyadic Analysis}
 \vspace{\baselineskip}
@@
 \lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{How to fix IR}}
+\chead{\fancyplain{}{Networks Matter in IR}}
 \rhead{\fancyplain{}{\today}}
@@
 However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce results that are not compelling.
-One reason is that the independence required by the typical methods employed does not exist in the data being studied.
-We present a different, regression based method, which constructs a latent network which incorporates \first, \second, and \third-order dependencies.
+As a result, the studies are often contradictory and produce findings that are not compelling.
+One reason is that the independence required by typical methods does not exist in the data.
+We present a different, regression based method, which constructs a latent network and incorporates \first, \second, and \third-order dependencies.
 We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects, and it also dominates standard approaches in terms of out-of-sample cross-validations.
+The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects and it also dominates standard approaches in terms of out-of-sample cross-validations.
 \end{abstract}
@@
 \clearpage
-\bibliography{/Users/s7m/whistle/master}
+%\bibliography{/Users/s7m/whistle/master}
+\bibliography{/Users/mdw/Git/whistle/master}
 % \bibliographystyle{authordate1}

### 2017-10-24 paper/2_model.tex
rock and roll.

@@
 	\begin{tabular}{cc}
-		\texttt{Sender heterogeneity} & \texttt{Receiver Heterogeneity} \\
+		\textsc{Sender heterogeneity} & \textsc{Receiver Heterogeneity} \\
 		\includegraphics[width=.4\textwidth]{adjRowDep.png} & \includegraphics[width=.4\textwidth]{adjColDep.png} \\
-		\texttt{Sender-Receiver Covariance} & \texttt{Reciprocity} \\
+		\textsc{Sender-Receiver Covariance} & \textsc{Reciprocity} \\
 		\includegraphics[width=.4\textwidth]{adjRowColCovar.png} & \includegraphics[width=.4\textwidth]{adjRecip.png} \\
@@
 
-The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third order dependencies.\footnote{For details on the SRRM see: for details on this model see \citet{li:loken:2002,dorff:minhas:2016}.
+The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third order dependencies.
+details on the SRRM see \citet{li:loken:2002,dorff:minhas:2017}.
 An earlier version of the LFM used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
-The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure underlying the LFM.} The AME model is specified as follows:
+The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure underlying the LFM.
+The AME model is specified as follows:
 

### 2017-10-24 paper/3_sim.tex
rock and roll.

@@
 where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
-$Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe.
+$Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
 We compare inference for $\mu$ and $\beta$---the latter parameter would be of primarily theoretical concern for applied scholars---using three models :

### 2017-10-24 paper/4_app.tex
rock and roll.

@@
 			& &  in a Network Model? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicates} \\ \midrule
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
-		Rose (2004) & WTO Membership Does not Effect Trade & {Replicates}\\ \midrule
+		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
 		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
@@
 
+One takeaway from is that many studies make knowledge claims based on differences in statistical significance between covariates.
+These differences may change dramatically when interdependencies are taken into account directly, as they do in this study.
+
 \subsection{Re-estimation of McDonald (2004)}
@@
 Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
-Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
+%Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict. 
 This indicates that conflict is less likely between members of a trade community.
@@
 When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
-The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---persists in the AME model.
+The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
+In the original, this negative linkages is none significant, but in the AME estimation a this negative linkage is more precisely estimated.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-We also observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea).
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
+There is a cluster of ``rogue states,'' who are receivers of conflict, in the top left (Iran, Iraq, Syria, Libya, North Korea).
 In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
@@
 In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type.
+Structural equivalence is present even when accounting for nodal characteristics like regime type.
 The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.
@@
 
-We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario.
-We focused on the variables measuring rivalry.
-We used mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.
+We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario focusing on the variables measuring rivalry.
+We employed mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.
 The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.

### 2017-10-24 paper/5_concl.tex
rock and roll.

@@
 These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
-We then reanalyzed each study using the latent network approach which captures additive and multiplicative aspects of interdependencies among the dyadic data.
+We then re-analyzed each study using the latent network approach which captures additive and multiplicative aspects of interdependencies among the dyadic data.
 In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.

### 2017-10-24 paper/netsMatter.tex
push pdf
rock and roll.

@@
 	pdfauthor={Shahryar Minhas, et alia.},%
-	pdftitle={How to Fix IR},%
+	pdftitle={Networks Matter},%
 	colorlinks,%
@@
 \bibliography{/Users/mdw/Git/whistle/master}
+%\bibliography{/Users/cassydorff/ProjectsGit/master}
 % \bibliographystyle{authordate1}

### 2017-10-25 paper/2_model.tex
typos from MF

@@
 \section{\textbf{Dependencies in Dyadic Data}}
-In working with relational data, scholars typically begin working with relational data by structuring it as a set of dyadic observations stacked on top of one another.
+In working with relational data, scholars typically begin by structuring these data as a set of dyadic observations stacked on top of one another.
 This makes sense if each observation is independent of the others.
@@
 Using the structure of an adjacency matrix, we visualize in Figure~\ref{fig:adjMatDeps} the types of first- and second-order dependencies that can complicate the analysis of relational data in traditional GLMs.
-The adjacency matrix on the top left highlights a particular row of an adjacency matrix, to illustrate that these values may be more similar to each other than other values in the adjacency matrix because each has a common sender.
-Homogeneity in interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
-Thus in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook) we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
+The adjacency matrix on the top left highlights a particular row to illustrate that these values may be more similar to each other than other values because each has a common sender.
+Interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
+Thus, in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook), we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.
@@
 Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem under study.
+Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem.
 The study of international relations is founded on the relations among actors.
@@
 
-The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third order dependencies.
-details on the SRRM see \citet{li:loken:2002,dorff:minhas:2017}.
+The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
+For details on the SRRM see \citet{li:loken:2002,dorff:minhas:2017}.
 An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.

### 2017-10-25 paper/3_sim.tex
typos from MF

@@
 
-Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value actually fell in those intervals.
-The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle while the naive approach performs poorly.
+Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell in those intervals.
+The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the naive approach performs poorly by comparison.
 

### 2017-10-25 paper/4_app.tex
typos from MF

@@
 The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the GLM results and those of the AME Model.
-In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
+In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
 When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.

### 2017-10-27 paper/1_intro.tex
typos from MF
minor cleanup

@@
 We focus on three types of interdependencies that can complicate dyadic analyses.
-First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict.
+First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive events such as conflict.
 Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
@@
 The multiplicative effects capture dependencies that result from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence.
-These type of meso-scopic features often arise in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.
+These type of meso-scopic features often arise in relational data because actors possess some set of latent, shared attributes that affect their probability of interacting with one another.
 

### 2017-10-27 paper/4_app.tex
minor cleanup

@@
 Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
-In addition, each of these studies adjusts the posterior standard errors in an attempt to account for the clustering of observations.
+In addition, almost all of these studies adjusts the standard errors in an attempt to account for the clustering of observations.
 We follow the strategy of the original study in our re-estimation of the original models.
@@
 
-One takeaway from is that many studies make knowledge claims based on differences in statistical significance between covariates.
+One takeaway is that many studies make knowledge claims based on differences in statistical significance between covariates.
 These differences may change dramatically when interdependencies are taken into account directly, as they do in this study.
@@
 The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
-In the original, this negative linkages is none significant, but in the AME estimation a this negative linkage is more precisely estimated.
+In the original, this negative linkages is not significant, but in the AME estimation this negative linkage is more precisely estimated.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 
-First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.
+First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.

### 2017-10-27 paper/netsMatter.tex
font experiment
back to sans
Fixed the font issue
new abstract

@@
 \usepackage[default,osfigures,scale=0.95]{opensans}
+%\usepackage[sfdefault,scaled=.85]{FiraSans}
+\usepackage[T1]{fontenc}
+\usepackage{textcomp}
+\usepackage[varqu,varl]{zi4}% inconsolata typewriter
+\usepackage{amsmath,amsthm}
+\usepackage[cmintegrals]{newtxsf}
 \usepackage[T1]{fontenc}
@@
 
-\title{Networks Matter in International Relations:\\A Latent Network Approach to Dyadic Analysis}
+% \title{Taking Dependencies Seriously:\\A Latent Network Approach to Dyadic Analysis}
+\title{Taking Conditional Independence Seriously:\\A Latent Network Approach to Dyadic Analysis}
 \vspace{\baselineskip}
@@
 	International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially data.
-These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.
-However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce findings that are not compelling.
-One reason is that the independence required by typical methods does not exist in the data.
-We present a different, regression based method, which constructs a latent network and incorporates \first, \second, and \third-order dependencies.
-We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects and it also dominates standard approaches in terms of out-of-sample cross-validations.
+This includes dyadic hypotheses and propositions, but especially dyadic data.
+Yet, much of the work conducted brushes aside the statistical complications that come with analyzing relational data, namely that the observations do not satisfy the conditional independence criterion.
+As a result, these studies often produce results with biased coefficient estimates and poorly calibrated standard errors, factors which have consequences for hypothesis testing.
+We present an alternative regression based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data using a set of additive and multiplicative effects within a Bayesian framework.
+To highlight the utility of this framework for studying relational data, we present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
+Additionally, we replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
+For each study, our approach shows that conventional methods overstate the effect of key variables in the study of interstate conflict, underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance of their variables.
+Further, the new approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
 \end{abstract}

### 2017-10-27 paper/1_intro.tex
modified intro

--- 2017-10-24 paper/1_intro.tex
@@
 Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
-However, this approach to studying dyadic data increases the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent.
+However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
 Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases.
-Such approaches may work in limited situations, but are not generally effective because these palliatives do not address the fundamental data generating process that remains a threat to inference because of the interdependence of observations or measurements.
-Namely, it is not just the diagonals of the variance-covariance matrix this affects.
-% \citep{beck:2012, king:roberts:2014}
+In practice, these palliatves rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
+Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012, king:roberts:2014}.
 
-In this article, we discuss the Additive and Multiplicative Effects (AME) model, 
-a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we discuss the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.
-First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict.
+First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive events such as conflict.
 Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
@@
 We begin with a brief review of these dependencies and the AME model.
-Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies.
-Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework.
-The latent factor approach (AME) is better able to capture first, second, and third-order interdependencies than the standard approach.
-It also produces results that are, at times, at odds with those found in the replicated studies as well as the broader literature from which they are drawn.
-As such, this approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
+Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of the dependencies that arise in dyadic data.
+Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained using the AME framework.
+The comparison reveals that in ignoring observational dependence, the standard approach overestimates the effect of key variables in the literature and underestimates their uncertainty.
+Consequently, the latent factor approach (AME) produces results that, at times, contradict those found in these studies in particular, and the broader literatures from which they are drawn.
+Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
 Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.

### 2017-10-27 paper/1_intro.tex
dump mesoscopic

@@
 We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-The multiplicative effects capture dependencies that result from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence.
-These type of meso-scopic features often arise in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.
+The multiplicative effects capture dependencies that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
 

### 2017-10-27 paper/netsMatter.tex
new(ish) abstract

--- 2017-10-24 paper/netsMatter.tex
@@
 \begin{abstract}
-	International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially data.
-These relational data contain information about the interdependencies of various phenomena, including countries, dyads, and even triads.
-However, most empirical studies of such data do not take into account these dependencies.
-As a result, the studies are often contradictory and produce findings that are not compelling.
-One reason is that the independence required by typical methods does not exist in the data.
-We present a different, regression based method, which constructs a latent network and incorporates \first, \second, and \third-order dependencies.
-We replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-The additive and multiplicative latent factor approach is shown to produce more precise estimates of covariate effects and it also dominates standard approaches in terms of out-of-sample cross-validations.
+	%International relations scholarship is filled with dyads.
+This includes dyadic hypotheses and propositions, but especially dyadic data.
+Yet, much of the work conducted brushes aside the statistical complications that come with analyzing relational data, namely that the observations do not satisfy the conditional independence criterion.
+As a result, these studies often produce results with biased coefficient estimates and poorly calibrated standard errors, factors which have consequences for hypothesis testing.
+We present an alternative regression based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data using a set of additive and multiplicative effects within a Bayesian framework.
+To highlight the utility of this framework for studying relational data, we present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
+Additionally, we replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
+For each study, our approach shows that conventional methods overstate the effect of key variables in the study of interstate conflict, underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance of their variables.
+Further, the new approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
+
+Much international relations scholarship concerns dyads: dyadic hypotheses and propositions, and especially dyadic data.
+Yet, much contemporary research brushes aside the complications that come with analyzing relational data that are interdependent.
+Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
+As a result, many studies often produce results with biased coefficient estimates and poorly calibrated standard errors.
+These biases have profound consequences for evaluating parametric models.
+%parahere?
+We present an alternative, regression-based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data.
+It uses a set of additive and multiplicative effects within a Bayesian regression framework.
+To highlight the utility of this framework for studying relational data, we present a simulation highlighting the ability of this approach to account for the dependencies that frequently characterize relational data such as often utilized in international relations.
+In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to the proposed latent factor approach.
+For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables and the hypotheses they derive from.
+The latent factor approach also dominates the current status quo approach in terms of out-of-sample cross-validations, rendering it more useful in hypothesis testing as well as forecasting.
+In short, it seems able to better capture the data generating process behind outcomes of interest.
+%parahere?
+We summarize the features of this latent factor model, and provide guideance on how to implement it in additional studies.
 \end{abstract}

### 2017-10-27 paper/2_model.tex
modeling section

@@
 \section{\textbf{Dependencies in Dyadic Data}}
+
 In working with relational data, scholars typically begin by structuring these data as a set of dyadic observations stacked on top of one another.
@@
 These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
-In Figure~\ref{fig:thirdDeps} we provide a visualization of a hypothetical relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
+In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
 To highlight third order dependence patterns, nodes with similar latent attributes are colored similarly.
@@
 The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are more rare.
-A prominent example of a network with this type of structure was found in American Politics by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
+A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
 \citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs linked to others that were politically similar.
@@
 		y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
-		\theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} + \text{\qquad(Exogenous parameters)} \\
-		& a_{i} + b_{j} + \epsilon_{ij} + \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
-		& \mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
+		\theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} \text{\qquad(Exogenous parameters)} \\
+		& + a_{i} + b_{j} + \epsilon_{ij} \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
+		& + \mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
 	\label{eqn:ame}
@@
 \end{align}
+
 where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver).

### 2017-10-27 paper/3_sim.tex
sim

@@
 \end{align}
+
 where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
@@
 	\centering
-	\caption{Regression parameter estimates for the naive, AME, and oracle models from 1,000 simulations.}
+	\caption{Regression parameter estimates for the naive, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
 	\label{fig:ameBias}
@@
 This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$.
-To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50.
-In both cases we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.
+To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50.
+Additionally, we calculate the median across the correlations and display that result using a vertical line.
+For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.
 

### 2017-10-27 paper/4_app.tex
app sec changes

@@
 Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
-In addition, almost all of these studies adjusts the standard errors in an attempt to account for the clustering of observations.
+In addition, most of these studies adjust the standard errors in an attempt to account for the clustering of observations.
 We follow the strategy of the original study in our re-estimation of the original models.
@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.
-Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.
-Finally, in each study we assess whether any new substantive findings emerge or if any previous findings disappear once the interdependencies in the data are modeled.
-
-The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies using a framework that ignores dependencies.
-Based in part on our results as well as the statistical characteristics of the AME framework, we believe that it does.
-The dynamic latent factor model provides a step forward in the modeling of international relations.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
 
 % SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
-Given current practices, most scholars have a single variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models.
-Generally, interpretations focus on a small set of independent variables.
+Typically, scholars have a small set of independent variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models.
 For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
@@
 Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
 
@@
 
-In addition, we also examine the predictions made with each approach using an out-of-sample cross validation strategy.
-By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures.
-To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models.
-To evaluate our model, we randomly divide the data  into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
+Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy.
+Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
 %\textcolor{red}{This is inaccurate for directed dyads. I think. MG} 
@@
 
+Next we discuss, each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.
+
 \subsection{Re-estimation of Reiter \& Stam (2003)}
+
+% They use the MID data, but note that:
+% \begin{quote}
+% 	We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
+This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
+We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
+\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).
+% \end{quote}
 
@@
 In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
-They use the MID data, but note that:
-
-\begin{quote}
-	We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
-This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
-We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).
-\end{quote}
-
 Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
@@
 Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.\footnote{Further details on each of the regression parameter estimates from the replication can be found in the Appendix.}
-
-One takeaway is that many studies make knowledge claims based on differences in statistical significance between covariates.
+Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
+An important takeaway here is that many studies make knowledge claims based on differences in statistical significance between covariates.
 These differences may change dramatically when interdependencies are taken into account directly, as they do in this study.
@@
 The link between conflict and trade is perplexing, with many persistent, yet competing explanations.
-\citet{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)''.
+\citet[p.
+547]{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''.
 Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
@@
 This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade (560).
-\citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP'' (560).
+This measure captures the score of the state in the dyad that possesses higher barriers to trade.
+\citet[p.
+560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
 The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
@@
 Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
-%Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict. 
 This indicates that conflict is less likely between members of a trade community.
@@
 At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
+% Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
 
@@
 
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
+\nocite{rose:2004}
+In 2004, Andrew Rose published a study in the \textit{American Economic Review} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
 It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007,ward:etal:2013}.
@@
 The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
-In the original, this negative linkages is not significant, but in the AME estimation this negative linkage is more precisely estimated.
+In the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
 However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.
 One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
@@
 The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-%Her time frame is from 1946-1999.
 The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340).
+These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.
 She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
@@
 	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012).
+Each circle designates a country and the color corresponds to the legend at the center of the visualization.
+Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries.
+The inner ring clusters countries by the similarity of who they receive conflict from.
 	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
@@
 The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the GLM results and those of the AME Model.
-In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations.
-When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero.
+In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone.
+When we examine the AME results, we find that none of these values are distinguishable from zero.
 Similarly, we find less pronounced effects for military capabilities.
 One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see states like the US, the UK, and Israel who often send conflict to similar targets.
-There is a cluster of ``rogue states,'' who are receivers of conflict, in the top left (Iran, Iraq, Syria, Libya, North Korea).
-In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China.
-In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see states like the US, the UK, Germany, Canada and Israel.
+These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries like Iraq, Russia, Syria, North Korea, and China.
+In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries.
+In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category).
 In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type.
@@
 The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.
-In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
+
+% In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
 In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.
-
-\begin{figure}
-\centering   
-	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
-\end{figure}
-\FloatBarrier
 

### 2017-10-27 paper/6_appendix.tex
app sec changes

@@
 Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis.
-For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.
+%For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.
 
@@
 % already show roc-pr in main text
-% \begin{figure}
-% 	\centering   
-% 	\subfigure[AUC]{\label{fig:weeksroc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
-% 	\subfigure[Precision and Recall]{\label{fig:weekspr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
-% 	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves, PR curves, and separation plots.}
-% \end{figure}
-% \FloatBarrier
+\begin{figure}
+\centering   
+  \subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
+  \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
+  \caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
+\end{figure}
+\FloatBarrier
 \clearpage

### 2017-10-27 paper/netsMatter.tex
app sec changes

@@
 
-Much international relations scholarship concerns dyads: dyadic hypotheses and propositions, and especially dyadic data.
-Yet, much contemporary research brushes aside the complications that come with analyzing relational data that are interdependent.
+	Much of international relations scholarship concerns dyads: dyadic hypotheses and propositions, and especially dyadic data.
+Yet, contemporary research brushes aside the complications that come with analyzing relational data that are interdependent.
 Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
@@
 These biases have profound consequences for evaluating parametric models.
-%parahere?
 We present an alternative, regression-based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data.
-It uses a set of additive and multiplicative effects within a Bayesian regression framework.
-To highlight the utility of this framework for studying relational data, we present a simulation highlighting the ability of this approach to account for the dependencies that frequently characterize relational data such as often utilized in international relations.
-In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to the proposed latent factor approach.
-For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables and the hypotheses they derive from.
-The latent factor approach also dominates the current status quo approach in terms of out-of-sample cross-validations, rendering it more useful in hypothesis testing as well as forecasting.
-In short, it seems able to better capture the data generating process behind outcomes of interest.
-%parahere?
-We summarize the features of this latent factor model, and provide guideance on how to implement it in additional studies.
+To highlight the utility of this framework for studying relational data, we first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
+In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to our alternative.
+For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables.
+Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
 \end{abstract}
@@
 \clearpage
-%\bibliography{/Users/s7m/whistle/master}
-\bibliography{/Users/mdw/Git/whistle/master}
+\bibliography{/Users/s7m/whistle/master}
+% \bibliography{/Users/mdw/Git/whistle/master}
 %\bibliography{/Users/cassydorff/ProjectsGit/master}

### 2017-10-27 code/2_codeForPaperGraphics/weeks_posterior.R

@@ -16,61 +16,61 @@ load( paste0(resultsPath,'model_k2.rda') )
 load(paste0(resultsPath,'weeks_glmfit.rda'))
 ############################################
 
-# coefSumm ###########################################
-ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-glmSumm = modSumm[,]
-probSumm = proSumm[,]
+# # coefSumm ###########################################
+# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+# rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+# glmSumm = modSumm[,]
+# probSumm = proSumm[,]
 
-modList = list(glmSumm, probSumm, ameSumm)
-modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+# modList = list(glmSumm, probSumm, ameSumm)
+# modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
 
-varKey = data.frame(
[…]

### 2017-10-27 paper/4_app.tex
tiny detritus removed

@@
 
-The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the GLM results and those of the AME Model.
+The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model.
 In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone.
@@
 Similarly, we find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see states like the US, the UK, Germany, Canada and Israel.
-These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries like Iraq, Russia, Syria, North Korea, and China.
+One explanation for this divergence is the AME model's ability to account for third-order effects.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada and Israel.
+These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
 In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries.
 In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category).
-In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type.
+In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type.
 Structural equivalence is present even when accounting for nodal characteristics like regime type.

### 2017-10-27 paper/4_app.tex
app edits 2

@@
 
-% In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.
-In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.
-
 \subsection{Re-estimation of Gibler (2017)}
@@
 A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
-In addition to this, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
+In addition, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
 This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-% SM: we still adding a coefplot here?
-% Insert coefplot here, without splines...\label{fig:giberrepl}
-% %
-% %\begin{table}
-% %\begin{center}
-% %\caption{Comparison of Gibler (2017) Model 6 results with AME results. This was run over a chain of $10,000 \text{BAZILLION}$ iterations.  \label{tab:gibme}}
-% %\begin{tabular}{lrr} \toprule
-% %& \multicolumn{2}{c}{$\hat{\beta}$ Estimates}\\ \cmidrule{2-3}
-% %Variable & \texttt{logit} & \texttt{amen} \\ \midrule
-% %Allied & 0.142&  0.023 \\
-% %Joint Democracy &\bf -0.507&  0.045 \\
-% %Peace Years \bf &-0.260& \bf -0.060 \\
-% %Spline 1 &\bf -0.001& \bf 0.00 \\
-% %Spline 2&\bf -0.000& \bf 0.00\\
-% %Spline 3 &-0.000& \bf 0.00\\ 
-% %Contiguity &\bf 2.412& \bf 0.640 \\
-% %Parity &0.075&  -0.013 \\
-% %Parity at entry year&\bf 0.868&  0.002 \\ 
-% %Rivalry &\bf 2.031&\bf 0.721\\ 
-% %Constant &\bf -5.526 & \bf  -2.581\\ \bottomrule
-% %\end{tabular}\\
-% %\end{center}
-% %{\bf Note}: {\bf Bold} indicates conventional statistical significance at $p < 0.05$ or less in Gibler (2017).
-For comparative purposes, only, we have employed the same criterion to the results from the \texttt{amen} estimation.
-NEED TO BE FIXED if staying.
-% %\end{table}
 
@@
 
-First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
+First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
 This means that significant testing (for better or worse) is compromised when network effects are ignored.
@@
 
-Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are iid.
+% i thinks that's already pretty explicit
+% Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are iid.

### 2017-10-27 paper/5_concl.tex
app sec changes
v2

--- 2017-10-24 paper/5_concl.tex
@@
 
-International relations is generally about the interactions and dependences among a set of countries or other important actors such as international governmental organizations (IGOs, such as the WTO) and non-governmental organizations (NGOs, such as the Gates Foundation).
-Some approaches focus on only looking a at small number of these actors, but many scholars examine a large number of actors at a time.
+International relations is generally about the interactions and dependencies among a set of countries or other important actors such as international governmental organizations (IGOs).
 This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
@@
 At the same time, we know that research designs focusing on the statistical analysis of dyadic data quickly go astray if the dyadic data are assumed to be iid.
-Virtually all of the standard statistical models---ordinary least squares and logistic regressions, to name a few---fail if the data are not iid.
+Virtually all of the standard statistical models---ordinary least squares and logistic regressions, to name a few---fail if the data are not conditionally independent.
 This fact has been accepted when it comes to temporal dependencies, but adoption of methods to account for network dependencies have seen less progress.
@@
 \citet{signorino:1999} showed why this is true of models of strategic interaction, but it is more broadly true of models that employ dyadic data.
-We show that latent networks can be employed to defeat this vulnerability of dyadic data in the realm of international relations.
-These approaches have been developed for a while, but are not yet widely used in international relations scholarship.
+We show that the AME framework can be employed to account for the statistical issues that arise when studying dyadic data.
+
+% These approaches have been developed for a while, but are not yet widely used in international relations scholarship.
 The latent network model captures first-order dependencies (the tendencies of certain actors to send and receive more of a certain action), second-order dependencies (the tendency of actors to reciprocate actions), and third-order dependencies (the ability for actions within a dyad to effect a third party) in dyadic data using a familiar regression framework that has been adapted for relational data---such as dyadic data---which are not independent nor identically distributed.
@@
 These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
-We then reanalyzed each study using the latent network approach which captures additive and multiplicative aspects of interdependencies among the dyadic data.
+We then re-analyzed each study using the AME model.
 In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
@@
 Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citep{king:roberts:2014}).
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences and better out-of-sample predictive performance, along with new substantive insights.
+Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences, better out-of-sample predictive performance, and new substantive insights.
+
+% We also examine each of the models using the AME framework described above.
+Our goal is to ascertain whether accounting for the statistical complications that arise in dyadic data result in different model estimates when addressed in an AME framework, and more importantly we highlight the substantive opportunities presented by our alternative approach.
+
+% Based in part on our results as well as the statistical characteristics of the AME framework, we believe that it does.
+The dynamic latent factor model provides a step forward in the modeling of international relations.

### 2017-10-28 paper/4_app.tex
booboo

@@
 			& &  in a Network Model? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicates} \\ \midrule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicate} \\ \midrule
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule

### 2017-10-28 paper/netsMatter.tex
v2
600 grit sand paper
not sure if title is right.

@@
 
-% add watermark
-\usepackage{draftwatermark}
-\SetWatermarkText{Version 1.0}
-\SetWatermarkScale{.5}
-\SetWatermarkColor[gray]{0.9}
+% % add watermark
+% \usepackage{draftwatermark}
+% \SetWatermarkText{Version 1.0}
+% \SetWatermarkScale{.5}
+% \SetWatermarkColor[gray]{0.9}
 
@@
 % \title{Taking Dependencies Seriously:\\A Latent Network Approach to Dyadic Analysis}
-\title{Taking Conditional Independence Seriously:\\A Latent Network Approach to Dyadic Analysis}
+\title{Taking Dyads Seriously:\\A Latent Network Approach to Dyadic Analysis}
 \vspace{\baselineskip}
@@
 
-	Much of international relations scholarship concerns dyads: dyadic hypotheses and propositions, and especially dyadic data.
+	Much international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
 Yet, contemporary research brushes aside the complications that come with analyzing relational data that are interdependent.
@@
 \clearpage
-\bibliography{/Users/s7m/whistle/master}
-% \bibliography{/Users/mdw/Git/whistle/master}
+%\bibliography{/Users/s7m/whistle/master}
+\bibliography{/Users/mdw/Git/whistle/master}
 %\bibliography{/Users/cassydorff/ProjectsGit/master}

### 2017-10-28 paper/1_intro.tex
proofread manuscript

@@
 
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}.
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.
 Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases.
-In practice, these palliatves rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
+Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters, so as to reflect the potential clustering of cases.
+In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
 Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012, king:roberts:2014}.
@@
 We focus on three types of interdependencies that can complicate dyadic analyses.
-First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive events such as conflict.
+First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.
 Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
-We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
+We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
 The multiplicative effects capture dependencies that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.

### 2017-10-28 paper/2_model.tex
get rid of unneces refs
proofread manuscript

@@
 Interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other.
-Thus, in most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook), we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
+In most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook), we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.
@@
 
-% SM: Why are we citing poast????
 For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level.
-More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors \citep{poast:2010}.
+More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors.
 These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.

### 2017-10-28 paper/3_sim.tex
proofread manuscript

@@
 \end{align}
-
 where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.

### 2017-10-28 paper/4_app.tex
proofread manuscript

@@
 \centering
-\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.
-Cases in which the substantive finding is not replicated are highlighted in bold.}
+\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.}
 	\begin{tabular}{l p{7cm} l} \toprule
@@
 Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-%\textcolor{red}{This is inaccurate for directed dyads. I think. MG} 
 Then for each $s \in \{1,\ldots,k\}$, we:
@@
 They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent.
-They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.
+They also discover that dictators are likely to challenge democracies, but that the converse is not true.
 In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
@@
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
-An important takeaway here is that many studies make knowledge claims based on differences in statistical significance between covariates.
-These differences may change dramatically when interdependencies are taken into account directly, as they do in this study.
+An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
+These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
 
@@
 \includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
- \caption{Nodal Random Effects for AME estimation of Rose (2004)}\label{fig:roser}
+ \caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}\label{fig:roser}
 \end{figure}
@@
 First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
-This means that significant testing (for better or worse) is compromised when network effects are ignored.
+This means that significance testing (for better or worse) is compromised when network effects are ignored.
 

### 2017-10-28 paper/netsMatter.tex
recomp
quick fix
proofread manuscript

@@
 % \title{Taking Dependencies Seriously:\\A Latent Network Approach to Dyadic Analysis}
-\title{Taking Dyads Seriously:\\A Latent Network Approach to Dyadic Analysis}
+% \title{Taking Dyads Seriously:\\A Latent Network Approach to Dyadic Analysis}
+\title{Taking Dyads Seriously}
 \vspace{\baselineskip}
@@
 \lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{Networks Matter in IR}}
+\chead{\fancyplain{}{Taking Dyads Seriously}}
 \rhead{\fancyplain{}{\today}}
@@
 	Much international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
-Yet, contemporary research brushes aside the complications that come with analyzing relational data that are interdependent.
+Yet, contemporary research brushes aside the complications that come with analyzing interdependent relational data.
 Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
@@
 These biases have profound consequences for evaluating parametric models.
-We present an alternative, regression-based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data.
-To highlight the utility of this framework for studying relational data, we first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
+We present an alternative, regression-based, approach that accounts for the dependencies complicating this type of analysis.
+We first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
 In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to our alternative.

### 2017-10-29 paper/1_intro.tex
max, you are killing me.

@@
 Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
-Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
+Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
 Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.

### 2017-10-29 paper/2_model.tex
max, you are killing me.

@@
 For example, consider that trade flows both from and to many wealthy, developed countries.
-The bottom-right panel, highlights a second order dependence, specifically, reciprocity.
+The bottom-right panel, highlights a second-order dependence, specifically, reciprocity.
 This is a dependency occurring within dyads involving the same actors whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
@@
 For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level.
-More often we find significant evidence of higher order structures that result from dependencies between multiple groups of actors.
+More often we find significant evidence of higher-order structures that result from dependencies between multiple groups of actors.
 These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
 In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
-To highlight third order dependence patterns, nodes with similar latent attributes are colored similarly.
+To highlight third-order dependence patterns, nodes with similar latent attributes are colored similarly.
 
@@
 
-The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are more rare.
+The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are rarer.
 A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
@@
 The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
-Of course, if one is able to specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
+Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
 However, it is rarely the case that we can do this even for TSCS data, thus we more than often make modeling decisions to account for that structure.
@@
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network based approaches.
+\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network-based approaches.
 Here we provide a brief review and then move to a simulation exercise.
@@
 This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
-The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third order interdependencies \citep{minhas:etal:2016:arxiv}.
+The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third-order interdependencies \citep{minhas:etal:2016:arxiv}.
 The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network.
@@
 
-By integrating the SRRM and LFM into a Bayesian probit framework, we are able to account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw for the exogenous parameters.
+By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw for the exogenous parameters.
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.

### 2017-10-29 paper/3_sim.tex
max, you are killing me.

@@
 We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
-The goal of the simulation is to assess how well AME can provide unbiased and well calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
+The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
 Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable.
@@
 For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated in the presence of unobserved dependencies, $W$.
-In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and are able to include it in our specification for $Y$.
+In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and include it in our specification for $Y$.
 The oracle case provides an important benchmark for the naive and AME approaches.
@@
 
-Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell in those intervals.
+Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals.
 The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the naive approach performs poorly by comparison.
@@
 	\centering
-	\caption{Proportion of times when true value fell in the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
+	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
 	\label{fig:ameCalib}
@@
 
-Last, beyond obtaining less biased and better calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
+Last, beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
 In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.

### 2017-10-29 paper/4_app.tex
max, you are killing me.

@@
 \centering
-\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach.}
+\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network-based approach.}
 	\begin{tabular}{l p{7cm} l} \toprule
@@
 		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
-		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
+		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
 		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
@@
 	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach.
-Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics.
+	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network-based approach.
+Four of the five studies involved a binary dependent variable; for those measures, area under the curve (AUC) statistics are reported.
 The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
-%Cases in which our network based approach outperformed the standard approach are highlighted in bold.
 	}
@@
 Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
-This indicates that conflict is less likely between members of a trade community.
+This indicates that conflict is less likely among members of a trade community.
 Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
@@
 
-When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar.
+When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar.
 The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
@@
 The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model.
-In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone.
+In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone.
 When we examine the AME results, we find that none of these values are distinguishable from zero.
@@
 One explanation for this divergence is the AME model's ability to account for third-order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada and Israel.
+Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel.
 These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.

### 2017-10-29 paper/5_concl.tex
max, you are killing me.

@@
 In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
-This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that can not possibly be valid.
-This in turn leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
+This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that cannot possibly be valid.
+In turn, this leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
 

### 2017-10-29 paper/netsMatter.tex
max, you are killing me.
okay
Summary

@@
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings.
-We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
+John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments.
+We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
 

### 2017-10-29 paper/1_intro.tex
lil bits

@@
 The comparison reveals that in ignoring observational dependence, the standard approach overestimates the effect of key variables in the literature and underestimates their uncertainty.
-Consequently, the latent factor approach (AME) produces results that, at times, contradict those found in these studies in particular, and the broader literatures from which they are drawn.
+Consequently, the latent factor approach (AME) produces results that, at times, contradict those found in these studies in particular as well as the broader literatures from which they are drawn.
 Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.

### 2017-10-29 paper/3_sim.tex
lil bits

@@
 Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable.
-Say that the true data-generating process for a particular $Y$ is given by:
+For instance, assume that the true data-generating process for a particular $Y$ is given by:
 
@@
 This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$.
-To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we just calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors was set to 100 (top pair of panels) and 50.
-Additionally, we calculate the median across the correlations and display that result using a vertical line.
+To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors is set to 100 (top pair of panels) and 50.
+Additionally, we calculate the median across the correlations and display the result using a vertical line.
 For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.

### 2017-10-29 paper/4_app.tex
lil bits

@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
 
 % SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
-Typically, scholars have a small set of independent variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models.
-For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.
+Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model.
+For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of a dispute dyad being in a dispute.
 In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
@@
 Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
+As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
 The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.

### 2017-10-29 paper/5_concl.tex
lil bits

@@
 We then re-analyzed each study using the AME model.
-In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fits, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
+In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
 This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that cannot possibly be valid.
-In turn, this leads to a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
+In turn, this creates a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
 

### 2017-10-30 paper/coverletter/netsMatter_coverLetter.tex
stuff

\documentclass[letterpaper]{article}
\usepackage{graphicx,fullpage}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage[sc,osf]{mathpazo}

\geometry{
  body={6.5in, 8.5in},
  left=1.0in,
  top=1.25in
}
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}
\setlength\parindent{0em}

% Make lists without bullets
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
  }
}{
  \end{list}
}

\begin{document}
\thispagestyle{empty}

  \begin{minipage}{0.64\linewidth}
Shahryar Minhas \\
Department of Political Science \\
Michigan State University University \\
368 Farm Lane \\
East Lansing, MI 48823\\
United States
\end{minipage}
\begin{minipage}{0.45\linewidth}
  \begin{tabular}{lr}
    Phone: & (248) 675-7345 \\
    Email: & \href{mailto:s7.minhas@gmail.com}{\tt s7.minhas@gmail.com}  \\
    Website:& \href{http://s7minhas.com/}{\tt s7minhas.com}
  \end{tabular}
\end{minipage}

\vspace{1.5in}

{Editorial Team of Political Analysis via submission portal}

\vspace{0.5in}

Dear Colleagues:\\[1ex]

This letter accompanies our submission of a manuscript for your consideration. The manuscript ``Taking Dyads Seriously'' introduces the Additive and Multiplicative Effects (AME) framework for conducting inference in the context of the dependencies that we often observe in dyadic data.\\[1ex]

We evaluate the utility of the AME framework using a simulation based exercise and a replication of five recently published works from the field of International Relations. Through the simulation exercise, we show that in the presence of unobserved dependencies the model is able to still provide less biased estimates and better calibrated standard errors than extant approaches in the literature. Next, for almost each of the replications reestimated using the AME framework we find that the case for the key findings from the original works become much less compelling. Additionally, we show through an out-of-sample cross-validation exercise that the AME approach uniformly outperforms each of the replicated approaches in capturing the data generating process of the event of interest.\\[1ex]

We believe that the AME framework is of notable interest to political scientists broadly, and that this study provides not only an introduction but a strong case for its applicability to the questions that we are seeking to address.\\[1ex]

We look forward to your evaluation of this paper.\\[1ex]

Respectfully submitted,

\vspace{.1in}

\includegraphics [scale=.8]{/Users/s7m/Dropbox/Finances/signature.png}

% \vskip 0.5in
% \hrule

\end{document}\bye

### 2017-10-30 paper/netsMatter.tex
lil bits
stuff

--- 2017-10-28 paper/netsMatter.tex
@@
 
-\author[Minhas]{Shahryar Minhas}
-\address{Shahryar Minhas: Department of Political Science}
-\curraddr{Michigan State University}
-\email[Corresponding author]{minhassh@msu.edu}
+% \author[Minhas]{Shahryar Minhas}
+% \address{Shahryar Minhas: Department of Political Science}
+% \curraddr{Michigan State University}
+% \email[Corresponding author]{minhassh@msu.edu}
 
-\author[Dorff]{Cassy L. Dorff}
-\address{Cassy L. Dorff}
-\curraddr{Department of Political Science, University of New Mexico}
-\email{cdorff@unm.edu}
+% \author[Dorff]{Cassy L. Dorff}
+% \address{Cassy L. Dorff}
+% \curraddr{Department of Political Science, University of New Mexico}
+% \email{cdorff@unm.edu}
 
-\author[Foster]{Margaret Foster}
-\address{Margaret Foster}
-\curraddr{Department of Political Science, Duke University}
-\email{margaret.foster@duke.edu}
+% \author[Foster]{Margaret Foster}
+% \address{Margaret Foster}
+% \curraddr{Department of Political Science, Duke University}
+% \email{margaret.foster@duke.edu}
 
-\author[Gallop]{Max Gallop}
-\address{Max Gallop}
-\curraddr{Department of Government and Public Policy, University of Strathclyde}
-\email{max.gallop@strath.ac.uk}
+% \author[Gallop]{Max Gallop}
+% \address{Max Gallop}
+% \curraddr{Department of Government and Public Policy, University of Strathclyde}
+% \email{max.gallop@strath.ac.uk}
 
-\author[Liu]{Howard Liu}
-\address{Howard Liu}
-\curraddr{Department of Political Science, Duke University}
-\email{hao.liu@duke.edu}
+% \author[Liu]{Howard Liu}
+% \address{Howard Liu}
+% \curraddr{Department of Political Science, Duke University}
+% \email{hao.liu@duke.edu}
 
-\author[Tellez]{Juan Tellez}
-\address{Juan Tellez}
-\curraddr{Department of Political Science, Duke University}
-\email{juan.tellez@duke.edu}
+% \author[Tellez]{Juan Tellez}
+% \address{Juan Tellez}
+% \curraddr{Department of Political Science, Duke University}
+% \email{juan.tellez@duke.edu}
 
-\author[Ward]{Michael D. Ward}
-\address{Michael D. Ward}
-\curraddr{Department of Political Science, Duke University}
-\email{michael.d.ward@duke.edu}
+% \author[Ward]{Michael D. Ward}
+% \address{Michael D. Ward}
+% \curraddr{Department of Political Science, Duke University}
+% \email{michael.d.ward@duke.edu}
 
-% \date{\today}
+% % \date{\today}
 
-\thanks{Shahryar Minhas and Michael D.
+% \thanks{Shahryar Minhas and Michael D.
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Robert Franzese provided welcome and helpful comments on the broader project from which this effort emerges at the 2017 Summer Political Methodology Meetings.
-We also thank Douglas Gibler, of the University of Alabama, for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
+John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments.
+We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
 
@@
 \lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{Taking Dyads Seriously}}
-\rhead{\fancyplain{}{\today}}
+% \chead{\fancyplain{}{Taking Dyads Seriously}}
+\chead{\fancyplain{}{}}
+% \rhead{\fancyplain{}{\today}}
+\rhead{\fancyplain{}{}}
 \rfoot{\fancyplain{}{\thepage}}
@@
 
-	Much international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
+	Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
 Yet, contemporary research brushes aside the complications that come with analyzing interdependent relational data.
@@
 % \bibliography{/Users/mdw/Git/whistle/master}
-%\bibliography{/Users/cassydorff/ProjectsGit/master}
+% \bibliography{/Users/cassydorff/ProjectsGit/master}
 % \bibliographystyle{authordate1}

### 2017-11-01 prezzy/TAMU2017.tex
pretty

%!TEX TS-program = xelatex
\documentclass[10pt,xcolor=dvipsnames,compress]{beamer}
\logo{\includegraphics[height=1cm]{Duke_shield.png}}
\usepackage{beamerthemesplit}
\usecolortheme[cmyk={1.0,.69,0.0,.115}]{structure}
\usetheme[height=11mm]{Rochester}
\setbeamertemplate{items}[ball] 
\setbeamerfont{tablefont}{size=\tiny}
\setbeamerfont{quotefont}{size=\small}

\usepackage[export]{adjustbox}
\usepackage{etoolbox}
\usepackage{dcolumn}
\usepackage[scale=2]{ccicons}
\usepackage{color}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{relsize}
\usepackage{mathtools}
\usepackage{bm}

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}

% graphics stuff
\usepackage{subfig}
\usepackage{graphicx}
\usepackage[space]{grffile} 
\usepackage{tikz}

% to change enumeration symbols begin{enumerate}[(a)]
%\usepackage{enumerate}

% square bracket matrices
\let\bbordermatrix\bordermatrix
\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}

% easy command for boldface math symbols
\newcommand{\mbs}[1]{\boldsymbol{#1}}

% command for R package font
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

% approx iid
\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}

% references to graphics
\makeatletter
\def\input@path{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/}, {/Users/mdw/Dropbox/Research/Ongoing/netModels/summResults/}}
\graphicspath{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/},{/Users/mdw/Dropbox/Research/Ongoing/netModels/summResults/}}

\title[AMEN]{Embracing Network Interdependence in International Relations\thanks{Based upon joint works with Shahryar Minhas (Michigan State University), Peter Hoff (Duke University), Cassy L. Dorff (University of New Mexico), Max Gallop (University of Strathclyde), Juan Tellez (Duke University), Howard Liu (Duke University), \& Margaret Foster (Duke University).}
}
\author 
{Michael D. Ward\\ 
Department of Political Science, Duke University} 
%\vspace[*]{5 cm}
\date{Prepared presentation at the Department of Political Science, Texas A\&M University, November 5, 2017. }

\begin{document}
\frame{\titlepage}

\begin{frame}
\frametitle{Abstract}
\framesubtitle{An Abstract for a Talk?}
\begin{itemize}
\item Much international relations scholarship concerns dyads. Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches. As a result, studies often produce results with biased coefficient estimates and poorly calibrated standard errors. These biases have profound consequences for evaluating parametric models.

\item We present an alternative, regression-based, approach that accounts for the dependencies complicating this type of analysis, and present a simulation exercise highlighting the model's ability to account for the dyadic dependencies. We replicate five prominent studies, comparing the standard approach to our alternative.

\item Conventional methods overstate the effect of key variables, underestimate the uncertainties, and often lead researchers to faulty conclusions about the substantive importance of their variables.

\item Our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
\end{itemize}
\end{frame}

\frame{\frametitle{Overview of Seminar}
\framesubtitle{Please, ask questions when you have them.}\tableofcontents}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Setup}
\frame{
  \frametitle{Motivation}

  Relational data consists of 
  \begin{itemize}
    \item a set of units or nodes
    \item a set of measurements, $y_{ij}$, specific to pairs of nodes $(i,j)$ 
  \end{itemize}

  \centering
  \includegraphics[width=1.05\textwidth]{df_adj_net3}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Relational data assumptions}

GLM: $y_{ij} \sim \beta^{T} X_{ij} + e_{ij}$

Networks typically show evidence against independence of {$e_{ij} : i \neq j$}

Not accounting for dependence can lead to:

\begin{itemize}
\item  biased effects estimation
\item  uncalibrated confidence intervals
\item  poor predictive performance
\item  inaccurate description of network phenomena
\end{itemize}

We've been hearing this concern for decades now:

\begin{tabular}{lll}
Thompson \& Walker (1982) & Beck et al. (1998) & Snijders (2011) \\
Frank \& Strauss (1986) & Signorino (1999) & Erikson et al. (2014) \\
Kenny (1996) & Li \& Loken (2002) & Aronow et al. (2015) \\
Krackhardt (1998) & Hoﬀ \& Ward (2004) & Athey et al. (2016) \\
\end{tabular}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What network phenomena? Sender heterogeneity}

  Values across a row, say $\{y_{ij},y_{ik},y_{il}\}$, may be more similar to each other than other values in the adjacency matrix because each of these values has a common sender $i$

  \centering
  \includegraphics[width=.7\textwidth]{adjRowDep}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What network phenomena? Receiver heterogeneity}

  Values across a column, say $\{y_{ji},y_{ki},y_{li}\}$, may be more similar to each other than other values in the adjacency matrix because each of these values has a common receiver $i$

  \centering
  \includegraphics[width=.7\textwidth]{adjColDep}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What network phenomena? Sender-Receiver Covariance}

  Actors who are more likely to send ties in a network may also be more likely to receive them

  \centering
  \includegraphics[width=.7\textwidth]{adjRowColCovar}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What network phenomena? Reciprocity}

  Values of $y_{ij}$ and $y_{ji}$ may be statistically dependent

  \centering
  \includegraphics[width=.7\textwidth]{adjRecip}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Social Relations Model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Social Relations Model (The ``A'' in AME)}

We use this model to form the additive effects portion of AME

\begin{align*}
\begin{aligned}
      y_{ij} &= \color{red}{\mu} + \color{red}{e_{ij}} \\
      e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\Sigma_{ab}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
\end{align*}

\begin{itemize}
\item $\mu$ baseline measure of network activity
\item $e_{ij}$ residual variation that we will use the SRM to decompose
\end{itemize}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Social Relations Model (The ``A'' in AME)}

\begin{align*}
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= {\color{red}{a_{i} + b_{j}}} + \epsilon_{ij} \\
      \color{red}{\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \}} &\sim N(0,\Sigma_{ab}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
\end{align*}

\begin{itemize}
\item  row/sender effect ($a_{i}$) \& column/receiver effect ($b_{j}$)
\item  Modeled jointly to account for correlation in how active an actor is in sending and receiving ties
\end{itemize}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Social Relations Model (The ``A'' in AME)}

\begin{align*}
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= a_{i} + b_{j} + \epsilon_{ij} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\color{red}{\Sigma_{ab}}) \\ 
      \{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\sim N(0,\Sigma_{\epsilon}), \text{ where } \\
      \color{red}{\Sigma_{ab}} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
\end{align*}

\begin{itemize}
\item $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$ capture heterogeneity in the row and column means
\item $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties)
\end{itemize}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Social Relations Model (The ``A'' in AME)}

\begin{align*}
\begin{aligned}
      y_{ij} &= \mu + e_{ij} \\
      e_{ij} &= a_{i} + b_{j} + \color{red}{\epsilon_{ij}} \\
      \{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\sim N(0,\Sigma_{ab}) \\ 
      \color{red}{\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\}} &\sim N(0,\color{red}{\Sigma_{\epsilon}}), \text{ where } \\
      \Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; & \color{red}{\Sigma_{\epsilon}} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
\end{aligned}
\end{align*}

\begin{itemize}
\item  $\epsilon_{ij}$ captures the within dyad effect
\item  Second-order dependencies are described by $\sigma_{\epsilon}^{2}$
\item  Reciprocity, aka within dyad correlation, represented by $\rho$
\end{itemize}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Third-Order Dependencies}

  \vspace{-10mm}

  \begin{table}[ht]
  \begin{tabular}{lcr}
  \scshape{Homophily} & & \scshape{Stochastic Equivalence} \\
  \includegraphics[width=.33\textwidth]{homophNet} & \hspace{2cm} &
  \includegraphics[width=.33\textwidth]{stochEquivNet}  
  \end{tabular}
  \end{table}

  To account for these patterns we can build on what we have so far and find an expression for $\gamma$:

  \vspace{-5mm}
  \begin{align*}
  \centering
  y_{ij} &\approx \beta^{T} X_{ij} + a_{i} + b_{j} + \gamma(u_{i},v_{j})
  \end{align*}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Latent Factor Model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Latent Factor Model: The ``M'' in AME}

Each node $i$ has an unknown latent factor

\vspace{-5mm}
\begin{align*}
{\textbf{u}_{i},\textbf{v}_{j}} \in \mathbb{R}^{k} \;\; i,j \in \{1, \ldots, n \} \\
\end{align*}

\vspace{-5mm}
The probability of a tie from $i$ to $j$ depends on their latent factors

\vspace{-5mm}
\begin{align*}
\begin{aligned}
  \gamma(\textbf{u}_{i}, \textbf{v}_{j}) &= \textbf{u}_{i}^{T} D \textbf{v}_{j} \\
  &= \sum_{k \in K} d_{k} u_{ik} v_{jk} \\
  &D \text{ is a  } K \times K \text{ diagonal matrix}
\end{aligned}
\end{align*}

Can account for both stochastic equivalence and homophily

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of AME}
\frame{
  \frametitle{Simulation Comparing GLM and LFM}
  \framesubtitle{}
We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis. The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies. Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable. For instance, assume that the true data-generating process for a particular $Y$ is given by:

\begin{align}
  y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
  \label{eqn:sim}
\end{align}
where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics. $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
}

\frame{
  \frametitle{Simulation Comparing GLM and LFM}
  \framesubtitle{Three Alternatives:}

\begin{itemize}
  \item the standard international relations approach assuming independent errors (current state-of-the discipline); 
  \item the AME approach with a unidimensional latent factor space ($K=1$);
  \item an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$. 
\end{itemize}

For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1. The value of $\gamma$ is also set to 1, which corresponds to an example where $W$ is associated with homophily. 
}

\frame{
  \frametitle{Regression Parameters from Simulations}
  \framesubtitle{}
\begin{figure}
  \centering
  \caption{Regression parameter estimates for the naive, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
  \label{fig:ameBias}
  \includegraphics[width=\textwidth]{ameSimBias_all.pdf} \\
\end{figure}
}
%
\frame{
  \frametitle{Coverage in the Simulations}
  \framesubtitle{}
\begin{figure}
  \centering
  \caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
  \label{fig:ameCalib}
  \includegraphics[width=\textwidth]{ameSimCover_all.pdf} \\
\end{figure}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Five Easy Pieces: Replication with AME}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
    \frametitle{Does it Really Matter?}
  \framesubtitle{Let's Look at Five Recent Publications in IR}

\begin{table}
\caption{Features of the Studies Re-estimated.}
\scalebox{.7}{
  \begin{tabular}{lcccccc}
    & Model &  Date Range & N. Actors  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
    Reiter \& Stam (JCR; 2003) &Logit &1945--1995 &  193 & $753,456$ & Directed & Robust \\  
    McDonald (JCR; 2004) & Logit &1959--2002 & 198 & $92,354$ & Undirected & Robust\\
    Rose (AER; 2004) & OLS & 1948--1999 & 177 & $234,597$ & Directed & Robust \\  
    Weeks (APSR; 2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
    Gibler (APSR; 2017) & Logit & 1816--2008 &193 &   $650,557$ & Undirected & none \\ \bottomrule
  \end{tabular}
  }
\end{table} 
}

\frame{
    \frametitle{Does it Really Matter?}
  \framesubtitle{Some Deets}
\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network-based approach.}
\scalebox{.7}{
  \begin{tabular}{l p{7cm} l} \toprule
    \multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
      & &  in a Network Model? \\ \toprule
    Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicate} \\ \midrule
    McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
    Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
    Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive; Machines are Not & {Does Not Replicate} \\\midrule
    Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
  \end{tabular}
  }
  \label{tab:modelFindingSumm}
\end{table}
}

\frame{
  \frametitle{Performance Statistics}
  \framesubtitle{In out-of-sample tests}
\vspace{-5mm}
\begin{table}[ht]
\scalebox{.7}{
  \begin{tabular}{ll cc}  \toprule
  Study & Statistic & GLM & AME \\
  \midrule
\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
    ~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\   \midrule
\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
    ~ & AUC-PR & 0.13 &  {0.28} \\    \midrule
\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
    ~ & RMDSE & 2.01 &  {1.06} \\ \midrule
\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
    ~ & AUC-PR & 0.00 &  {0.15} \\    \midrule
\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
    ~ & AUC-PR & 0.00 &  {0.08} \\      \bottomrule
  \end{tabular}
  }
\end{table}
}

\frame{
  \frametitle{Reiter-Stam Performance Comparisons}
  \framesubtitle{ }
\vspace{-5mm}
\begin{figure}
\includegraphics[height=5cm]{rsroc}
\includegraphics[height=5cm]{rspr}
\caption{Reiter-Stam Performance}
\end{figure}
}

\frame{
  \frametitle{McDonald Performance Comparisons }
  \framesubtitle{ }
\vspace{-5mm}
\begin{figure}
\includegraphics[height=5cm]{mcdonaldroc}
\includegraphics[height=5cm]{mcdonaldpr}
\caption{McDonald Performance }
\end{figure}
}

\frame{
  \frametitle{Rose Performance Comparisons }
  \framesubtitle{ }
%\vspace{-7.5mm}
\begin{figure}
\includegraphics[scale=.5]{roseout}\\
\includegraphics[scale=.4]{rose_srmvc}
\caption{Rose Performance Statistics}
\end{figure}
}

\frame{
  \frametitle{Weeks Performance Comparisons }
  \framesubtitle{ }
\vspace{-5mm}
\begin{figure}
\includegraphics[height=5cm]{weeksroc}
\includegraphics[height=5cm]{weekspr}
\caption{Weeks Performance; no single MID predicted}
\end{figure}
}

\frame{
  \frametitle{Gibler Performance Comparisons }
  \framesubtitle{ }
\vspace{-5mm}
\begin{figure}
\includegraphics[height=5cm]{giblerroc}
\includegraphics[height=5cm]{giblerpr}
\caption{Gibler Performance; no single MID predicted}
\end{figure}
}

%%%%%%%%%%%%%%%%%%%%%%%
\section{Lessons Learned}
\frame{
   \frametitle{Some Lessons Learned}
  \framesubtitle{ } 
\begin{itemize}
\item Findings that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME. In GLM approaches to the analysis of dyadic data, coefficents are biased, imprecisely measured, and have poorly calibrated standard errors.  This means that significance testing (for better or worse) is compromised when network effects are ignored.
\item New insights emerge from the additional information derived from AME. In particular, there is actual information about the dependencies so that clusters can be identified,   the extent of reciprocity at the dyad level, and among dyads.  
\item The actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.   
\item The AME approach dominates the GLM approaches in terms of performance. Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$, but it also dominates at correctly identifying occurrences of $1$.

\end{itemize}
}

\begin{frame}{Frame}
  \frametitle{Networks Really Matter}
  \framesubtitle{We miss a lot when we ignore them.}
  \begin{columns}
    \column{0.5\textwidth}
    \centering
      \includegraphics[width=\linewidth]{weeks_circPlot.pdf}
      \column{0.5\textwidth}
      \setlength\arraycolsep{3pt}
      \begin{itemize}
        \item clustering
        \item different patterns for senders and receivers
        \item similar nodal, and dyadic patterns 
        \item gauge reciprocity (and orthogonality)
      \end{itemize}
 \end{columns}
\end{frame}

\frame{\frametitle{Thanks!}
\framesubtitle{There's more, but I'll stop here for your questions and comments.}

More work; copies available upon request:

\begin{itemize}
\item Shahryar Minhas, Peter D. Hoff, and Michael D. Ward. 2017. \textit{Additive and Multiplicative Latent Factor Models for Network Influence}. 
\item Shahryar Minhas, Peter D. Hoff, and Michael D. Ward. 2017. \textit{Influence Networks in International Relations}. 
\item Shahryar Minhas, Cassy L. Dorff, Margaret Foster, Max Gallop, Howard Liu, Juan Tellez, and Michael D. Ward. 2017. \textit{Taking Dyads Seriously}.
\end{itemize}

Thank you!

}
\section{Appendix: Code Example}

\frame{
  \frametitle{Example $\mathcal{R}$ code}
Using the AMEN function requires formatting data into a particular structure. The primary distinction in data formatting is whether the outcome of interest represents a directed or undirected network. Use this link \url{GitHub.com/s7minhas/amen}

If undirected, the AMEN function has three main inputs:

\begin{itemize} 
    \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network.
    \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset. 
    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
\end{itemize}

If directed, AMEN further requires:

\begin{itemize} 
    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
    \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset. 
\end{itemize}

}

\frame{
  \frametitle{AMEN Example}
  \framesubtitle{specification}
Beyond the data inputs, the AMEN function requires additional specification:

\begin{itemize}
    \item model: how to model the outcome variable, e.g., `logit'
    \item symmetric: whether the input network is symmetric
    \item intercept: whether to estimate an intercept
    \item nscan: number of iterations of the Markov chain
    \item burn: burn-in period
    \item odens: thinning interval
    \item R: dimension of the multiplicative effect (referred to as $K$ in the paper)
    \item gof: whether to calculate goodness-of-fit statistics
\end{itemize}

}

\frame{
  \frametitle{Actual Code}
\# running in parallel varying k\\
imps = 10000 ; brn = 25000 ; ods = 10 ; latDims = 0:3\\
\# Run amen in parallel
library(doParallel) ; library(foreach) ; cl=makeCluster(4) \\ registerDoParallel(cl)\\
foreach(ii=1:length(latDims), .packages=c("amen")) \%dopar\% {\\
  \# load previous model run\\
  load(prevModelFiles[ii])\\
  \# extract start vals\\
  startVals0 = ameFit\$'startVals'\\
  \# dump rest\\ 
  rm(ameFit)\\
   ameFit = ame\_repL(\\
    Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL, \\
    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii], \\
    nscan=imps, seed=1, burn=brn, odens=ods, \\
    plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0,\\
    periodicSave=TRUE ) \\    
  save(ameFit, file=paste0('model\_k', latDims[ii],'\_v2.rda') )\\
}
stopCluster(cl)
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Backup Slides}
\frame{\frametitle{backup slides}
Go ahead, ask me about ERGM. }

\frame{
  \frametitle{There is a problem with ERGM.}
  \framesubtitle{It is a bug, not a feature.}
\vspace{-5mm}
It is a bug, not a feature. If you don't believe us, believe these scholars:
\begin{enumerate}
\item Schweinberger, M. (2011). Instability, Sensitivity, and Degeneracy of Discrete Exponential Families. \textbf{Journal of the American Statistical Association}, 106(496):1361–1370.
\item Schweinberger, M. and Handcock, M. S. (2015). Local Dependence in Random Graph Models: Characterization, Properties and Statistical Inference. \textbf{Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, 77(3):647–676.
\item Chatterjee, S. and Diaconis, P. (2013). Estimating and Understanding Exponential Random Graph Models. \textbf{The Annals of Statistics}, 41(5):2428–2461.
\item Rastelli, R., Friel, N., and Raftery, A. E. (2016). Properties of Latent Variable Network Models. \textbf{Network Science}, 4(4):1–26.
\end{enumerate}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What Is Really Wrong? }
\vspace{-5mm}

You may assume this is your likelihood surface.  Courtesy of Chatterjee and Diaconis, who concluded that sufficient statistics are not sufficient to solve this problem.  Estimation of ERGMs contains this flaw, with pseudolikelihood, maximum likelihood, or MCMC. 
\includegraphics[width=1\textwidth]{disconuity.png}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What Does This Mean? }
\vspace{-5mm}
\begin{enumerate}
\item Probabilistic ERGM models place almost all of the probability on networks that
are either nearly empty (degenerate) with no linkages or nearly saturated with
all nodes being interconnected. 
\item The likelihood surface contains steep or discontinuous gradients that render it impossible to solve numerically (or analytically). Even (especially) for very small networks this is problematic.
\item Find a different way to estimate network data.  
\end{enumerate}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Extant Political Science Examples}
\vspace{-5mm}
Extant Political Science Examples Using Latent Distance Models Are Wrong.
\begin{enumerate}
\item There is a problem with the MCMC estimator in \texttt{latentnet}. Don't use it until it is fixed.
\item Latent distance models are hard to correctly interpret, in prominent evaluations in the literature it is done incorrectly.  This is explained in their introduction a decade ago.
\item The advice in Cranmer et al is wrong. Latent network models are not biased and when correctly implemented out-perform ERGM models.
\end{enumerate}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
\frametitle{Compared to ERGM?}
  \framesubtitle{Swiss Climate Change Application}

Cross-sectional network measuring whether an actor indicated that they collaborated with another during the policy design of the Swiss CO$_{2}$ act (Ingold 2008)

\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
  \includegraphics[width=.47\textwidth]{dvNet_outDegree} & 
  \includegraphics[width=.44\textwidth]{dvNet_inDegree}
  \end{tabular}
\end{figure}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Parameter Estimates for Swiss Example}

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Sun Aug 21 03:32:43 2016
\vspace{-15mm}
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{lcccccc}
   & Expected & \multirow{2}{*}{Logit} & \multirow{2}{*}{MRQAP} & \multirow{2}{*}{LSM} & \multirow{2}{*}{ERGM} & \multirow{2}{*}{AME} \\ 
   & Effect & & & & & \\
  \hline
  \textbf{Conflicting policy preferences} &  &  &  &  &  &  \\ 
  $\;\;\;\;$ Business vs. NGO & \color{red}{$-$} & -0.86 & \color{red}{-0.87$^{\ast}$} & \color{red}{-1.37$^{\ast}$} & \color{red}{-1.11$^{\ast}$} & \color{red}{-1.37$^{\ast}$} \\ 
  $\;\;\;\;$ Opposition/alliance & \color{blue}{$+$} &  \color{blue}{1.21$^{\ast}$} & \color{blue}{1.14$^{\ast}$} & 0.00 & \color{blue}{1.22$^{\ast}$} & \color{blue}{1.08$^{\ast}$} \\ 
  $\;\;\;\;$ Preference dissimilarity & \color{red}{$-$} & -0.07 & -0.60 & \color{red}{-1.76$^{\ast}$} & -0.44 & \color{red}{-0.79$^{\ast}$} \\ 
  \textbf{Transaction costs} &  &  &  &  &  &  \\ 
  $\;\;\;\;$ Joint forum participation & \color{blue}{$+$} & \color{blue}{0.88$^{\ast}$} & \color{blue}{0.75$^{\ast}$} & \color{blue}{1.51$^{\ast}$} & \color{blue}{0.90$^{\ast}$} & \color{blue}{0.92$^{\ast}$} \\ 
  \textbf{Influence} & &  &  &  &  &  \\ 
  $\;\;\;\;$ Influence attribution & \color{blue}{$+$} & \color{blue}{1.20$^{\ast}$} & \color{blue}{1.29$^{\ast}$} & 0.08 & \color{blue}{1.00$^{\ast}$} & \color{blue}{1.09$^{\ast}$} \\ 
  $\;\;\;\;$ Alter's influence indegree & \color{blue}{$+$} & \color{blue}{0.10$^{\ast}$} & \color{blue}{0.11$^{\ast}$} & 0.01 & \color{blue}{0.21$^{\ast}$} & \color{blue}{0.11$^{\ast}$} \\ 
  $\;\;\;\;$ Influence absolute diff. & \color{red}{$-$} & \color{red}{-0.03$^{\ast}$} & \color{red}{-0.06$^{\ast}$} & 0.04 & \color{red}{-0.05$^{\ast}$} & \color{red}{-0.07$^{\ast}$} \\ 
  $\;\;\;\;$ Alter = Government actor & \color{blue}{$+$} & \color{blue}{0.63$^{\ast}$} & 0.68 & -0.46 & \color{blue}{1.04$^{\ast}$} & 0.55 \\ 
  \textbf{Functional requirements} &  &  &  &  &  & \\ 
  $\;\;\;\;$ Ego = Environmental NGO & \color{blue}{$+$} & \color{blue}{0.88$^{\ast}$} & 0.99 & -0.60 & \color{blue}{0.79$^{\ast}$} & 0.67 \\ 
  $\;\;\;\;$ Same actor type & \color{blue}{$+$} & \color{blue}{0.74$^{\ast}$} & \color{blue}{1.12$^{\ast}$} & \color{blue}{1.17$^{\ast}$} & \color{blue}{0.99$^{\ast}$} & \color{blue}{1.04$^{\ast}$} \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Latent Factor Visualization for Swiss Example}

  \centering
  \vspace{-10mm}
  \includegraphics[width=.7\textwidth]{ameFitSR_2_UV}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Out-of-Sample Performance Assessment: Strategy}

\begin{itemize}
  \item  Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
  \item  For each $s \in \{1, \ldots, S\}$:
  \begin{itemize}
    \item  Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
    \item  For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
  \end{itemize}
\end{itemize}

This procedure generates a sociomatrix of out-of-sample predictions of the observed data

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Performance Comparison for Swiss Example}

\begin{tabular}{cc}
\includegraphics[width=.5\textwidth]{roc_outSample} & 
\includegraphics[width=.5\textwidth]{rocPr_outSample} 
\end{tabular}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Network Dependencies: Swiss Example}
  \centering
  \includegraphics[width=1\textwidth]{netPerfCoef}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Much current advice is wrong.}
  \textbf{Unfortunately some misleading and incorrect advice is in the current literature. Maybe it will be corrected.}

\begin{enumerate}
\item Cranmer, Leifeld, McClurg, Rolfe (2017 AJPS) stated that latent space models are only preferable if there are few isolates, enough observations, and the interdependence is theoretically uninteresting.

\item All of this is wrong. Not only do they incorrectly assume all latent models to be based on Euclidean distance but their statistical comparisons are based on a program that contains known errors. Moreover, using BIC to choose among different network models is itself questionable.

\item It is not true that ``ERGM offers the most preferable combination of model fit and number of parameters,'' despite their assertions.
\end{enumerate}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Simulation Comparison}

\includegraphics[width=1\textwidth]{ameVergmSim.png}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN v LSM Performance}

  \begin{tabular}{cc}
  \includegraphics[width=.5\textwidth]{roc_latSpace_outSample} & 
  \includegraphics[width=.5\textwidth]{rocPr_latSpace_outSample}
  \end{tabular}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN versus LSM Net Dependence}

  \includegraphics[width=1\textwidth]{netPerfCoef_latSpace}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN varying $K$ Performance}

  \begin{tabular}{cc}
  \includegraphics[width=.5\textwidth]{roc_ameSR_outSample} & 
  \includegraphics[width=.5\textwidth]{rocPr_ameSR_outSample}
  \end{tabular}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN varying $K$ Net Dependence}

  \includegraphics[width=1\textwidth]{netPerfCoef_ameSR}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2017-11-10 prezzy/TAMU2017.tex
post TAMU changes to Slides
This gets at Flo’s issue on uTv and my concerns about estimation.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\frame{
+  \frametitle{What Exactly Is Third-Order Dependency?}
+  \framesubtitle{People Want to Know!}
+\begin{itemize}
+\item<1> If each node has individual characteristics, 
+say $x_i$ and $x_j$ for two nodes.  Then,
+$x_i \times x_j$ is the multiplicative product of their nodal characteristics and may be related to their relationship, 
+$y_{ij}$.  
+Let $x$ be an encoding of whether a node is democratic or not. 
+Then, $x_i \times x_j \equiv 1 \iff$ 
+\textit{both are democracies}.  This approach is often used in standard analyses as a variable often known as 
+\textit{joint democracy}.
+In the context of networks, a positive effect from the interaction of $x_i$ and $x_j$ might affect some dyadic variable, such as not being in a militarized dispute against one another---often denoted the democratic peace.
+This positive effect is known in network terminology as \textit{homophily}.
+Most studies just look at the correlation between aggregated variables, but the proposed approach can more directly estimate the degree of homophily.
+\end{itemize}
+}
+
+\frame{
+  \frametitle{What Exactly Is Third-Order Dependency?}
+  \framesubtitle{People Want to Know!}
+\begin{itemize}
+\item<1> Homophily which stems from nodal characteristics results in certain types of network structures: 1) transitivity, 2) balance, and 3) clustering.
+For example, homophily tends to induce a large number of \textit{triads}, since if $i$ is similar to $j$ and as a result has a link with $j$, for the same reason it will share a link with $k$ if $j$ and $k$ are similar, thus creating a triad.
+Groups of triads tend to result in \textit{clusters}.
+This can result in clusters in which many elements in the cluster are not only related to one another but share similar characteristics.
+This is known as \textit{stochastic equivalence}.
+
+\item To capture this kind of stochastic equivalence---and other types of triadic dependence---we include a latent representation of nodal characteristics in the regression model.
+This allows the network representation in a matrix to be dependent upon how similar these latent characteristics are.
+These latent nodal characteristics are represented by a type of multiplicative effects that are expressed as $\mathbf{u}_i^\intercal \mathbf{v}_i$.
+This captures the third-order dependencies, i.e., the dependencies among triads.
+\end{itemize}
+}
+
 \frame{
@@
 
-Can account for both stochastic equivalence and homophily
+This can account for both stochastic equivalence and homophily, as well as clustering and balance.
+
+}
+
+\frame{
+  \frametitle{How Does Estimation Work?}
+  \framesubtitle{It Is NOT All in the Error Term!}
+ \begin{itemize}
+  \item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+  \item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+  \item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+  \item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
+  \item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
+  \item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
+  \item For each $k \in K$:
+  \begin{itemize}
+    \item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+    \item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+    \item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
+}
+  \end{itemize}
+ \end{itemize}
 

### 2017-11-10 prezzy/netsMatterDeck.tex
fixes

--- 2017-11-10 prezzy/TAMU2017.tex
@@
 \documentclass[10pt,xcolor=dvipsnames,compress]{beamer}
-\logo{\includegraphics[height=1cm]{Duke_shield.png}}
 \usepackage{beamerthemesplit}
@@
 
-\title[AMEN]{Embracing Network Interdependence in International Relations\thanks{Based upon joint works with Shahryar Minhas (Michigan State University), Peter Hoff (Duke University), Cassy L.
-Dorff (University of New Mexico), Max Gallop (University of Strathclyde), Juan Tellez (Duke University), Howard Liu (Duke University), \& Margaret Foster (Duke University).}
+\title[AMEN]{Taking Dyads Seriously\thanks{Based upon joint works with Michael Ward (Duke University), Cassy L.
+Dorff (University of New Mexico), Max Gallop (University of Strathclyde), Juan Tellez (Duke University), Howard Liu (Duke University), \& Margaret Foster (Duke University).
+This is also part of a broader project involving Peter Hoff (Duke University) and Michael Ward.}
 }
+
 \author 
-{Michael D. Ward\\ 
-Department of Political Science, Duke University} 
-%\vspace[*]{5 cm}
-\date{Prepared presentation at the Department of Political Science, Texas A\&M University, November 5, 2017. }
+{Shahryar Minhas\\ minhassh@msu.edu \\ s7minhas.com \\
+Michigan State University}
+
+\date{Beamer is ugly, this needs to be changed.}
 
@@
   \framesubtitle{People Want to Know!}
+Too much to read on a slide.
 \begin{itemize}
@@
 Groups of triads tend to result in \textit{clusters}.
-This can result in clusters in which many elements in the cluster are not only related to one another but share similar characteristics.
-This is known as \textit{stochastic equivalence}.
+\end{itemize}
+}
 
-\item To capture this kind of stochastic equivalence---and other types of triadic dependence---we include a latent representation of nodal characteristics in the regression model.
-This allows the network representation in a matrix to be dependent upon how similar these latent characteristics are.
-These latent nodal characteristics are represented by a type of multiplicative effects that are expressed as $\mathbf{u}_i^\intercal \mathbf{v}_i$.
-This captures the third-order dependencies, i.e., the dependencies among triads.
+\frame{
+  \frametitle{What Exactly Is Third-Order Dependency?}
+  \framesubtitle{People Want to Know!}
+\begin{itemize}
+\item<1> Stochastic equivalence.
 \end{itemize}

### 2017-11-14 prezzy/netsMatterDeck.tex
less ugly, toooooo much text
adding chinese abstract. Korean to come.
pu dekcuf si gnihtemos
sdsfdfs

@@
 %!TEX TS-program = xelatex
-\documentclass[10pt,xcolor=dvipsnames,compress]{beamer}
+% \documentclass[10pt,xcolor=dvipsnames,compress]{beamer}
+% \usepackage{beamerthemesplit}
+% \usecolortheme[cmyk={1.0,.69,0.0,.115}]{structure}
+% \usetheme[height=11mm]{Rochester}
+
+\documentclass[10pt, compress]{beamer}
 \usepackage{beamerthemesplit}
 \usecolortheme[cmyk={1.0,.69,0.0,.115}]{structure}
-\usetheme[height=11mm]{Rochester}
+%\usetheme[usetitleprogressbar]{m}
+\usepackage{booktabs}
+\usepackage[scale=2]{ccicons}
+\usepackage{xeCJK}
+\usepackage{kotex}
+
 \setbeamertemplate{items}[ball] 
@@
 
-\author 
-{Shahryar Minhas\\ minhassh@msu.edu \\ s7minhas.com \\
-Michigan State University}
+\author{Shahryar Minhas\\Cassy L. Dorff\\Margaret Foster\\Max Blau Gallop\\Juan Tellez\\ Michael D. Ward}
+%{Shahryar Minhas\\ minhassh@msu.edu \\ s7minhas.com \\
+%Michigan State University}
 
-\date{Beamer is ugly, this needs to be changed.}
+\date{\today}
 
@@
 \frame{\titlepage}
+
+%\frame{
+%Shahryar Minhas (MSU)\\
+%Cassy L. Dorff (UNM)\\
+%Margaret Foster (Duke)\\
+%Max Blau Gallop (Strathclyde)\\
+%Juan Tellez (Duke)\\
+%Michael D. Ward (Duke)
+%}
 
@@
 \item Our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Abstract}
+\begin{itemize}
+\item 大多数的国际关系研究关注于两两二元关系，而二元关系数据经常无法满足统计意义上对于数据之间条件独立的假设。因此，在忽略数据互赖性之下所建立的统计模型往往估计出有偏差的变数系数与标准误，而这些偏差对于衡量统计参数模型有深远的影响。
+
+\item 在本文中，我们提出另一种立基于回归模型的统计方法，来考查这类数据之间复杂的互赖关系，并将其应用在五篇政治学重要的研究论文中来重置他们的研究结果，进而比较传统的方法与我们的模型之间所呈现的差异。
+
+\item 我们的研究结果发现，传统的方法过度强调主要变数在统计模型中的影响，低估了不确定性，且往往在变数解读上给予研究者有偏差的结论。
+
+\item 我们的模型在样本外的交叉验证上有较佳的表现，赋予研究者更有用的预测应用，以及更有效地模拟应变量数据的生成过程。
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Abstract}
+\begin{itemize}
+\item  많은 국제관계학 연구는 양자 (dyad) 관계를 다룬다.
+그러나 대부분의 경우 양자관계의 관찰은 통계학에서 요구되는 조건부 독립성 기준을 만족시키지 못한다.
+결과적으로, 그러한 연구들은 종종 편향된 계수 값과 잘못 계산된 표준 오차 값을 산출하게 된다.
+이 편차들은 모수적 통계 모형을 이용하여 수치를 구함에 있어 중대한 결과의 차이를 가져온다.
+\item 본 연구는 회귀분석에 근거한 방법론적 대안을 제시하여 양자관계 분석을 까다롭게 하는 종속성을 처리하는 방법을 소개한다.
+또한, 시뮬레이션을 통해 차용된 모형이 어떻게 종속성을 처리하는지를 보여준다.
+본 연구는, 기존의 통상적 접근 방식과 제시된 새로운 방식을 비교하기 위하여 잘 알려진 다섯 연구의 기존 모형을 복제 및 재검토 (replicate) 하였다.
+\item 통상적 방법은 주요 변수의 효과를 과장하고, 통계적 불확실성은 과소 추산함으로써, 연구자들로 하여금 해당 변수의 실질적 중요도에 대해 오판하게 한다.
+\item 반면, 본 연구에서 차용된 방법은 표본 외 예측 (out-of sample) 교차 검증법에서 우세한 결과를 보였다. 따라서, 기존의 방법과 비교하였을 때 차용된 방법은 예측 모형으로의 활용도가 더 높고, 주요 결과 이외의 데이터 생성 과정을 모형화하는 데에 더욱 유용하다.
+
 \end{itemize}

### 2018-01-04 paper/2_model.tex
added discussion of symmetry, changed stuff about R&S/Rose

@@
 Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
-However, it is rarely the case that we can do this even for TSCS data, thus we more than often make modeling decisions to account for that structure.
+However, it is rarely the case that we can do this even for TSCS data, thus we often make modeling decisions to account for that structure.
 Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
@@
 $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
+The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.} Note that this model easily generalizes to the case, common in international relations, where interactions are unidrected (for example the presence of conflict or a bilateral investment treaty).
+In this case, $\rho$ is constrained to be 1, sender and receiver effects are identical for each observation, as are \mathbf{u_{i}}$ and $\mathbf{v_{j}}$.
+In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.
 

### 2018-01-04 paper/4_app.tex
stuff
added discussion of symmetry, changed stuff about R&S/Rose

@@
 			& &  in a Network Model? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicate} \\ \midrule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
@@
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies.
+Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the effects uncovered by the authors remains robust when accounting for dyadic interdependencies.
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
-An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
+%An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
 These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
@@
 The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
-In the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
-However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
+Howevever, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
+In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
+While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
 The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
@@
 When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
-While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.
+While the AME model replicates the main (null) result, there are substantial and substantive improvements gained from moving to an AME estimation.
 
@@
 \begin{figure}[!h]
-	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
+	\includegraphics[width=\textwidth]{weeks_circPlot.png}
 	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012).

### 2018-01-08 paper/2_model.tex
small scale edits

@@
 While the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others, we know this assumption to be false in theory and in practice.
-Relational data comes with an explicit structure that in general leads to particular types of dependencies.
+Relational data comes with an explicit structure that generally leads to particular types of dependencies.
 The importance of accounting for the underlying structure of our data has been a lesson well understood, at least, when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
@@
 
-Using the structure of an adjacency matrix, we visualize in Figure~\ref{fig:adjMatDeps} the types of first- and second-order dependencies that can complicate the analysis of relational data in traditional GLMs.
+Using the structure of an adjacency matrix, Figure~\ref{fig:adjMatDeps} visualizes the types of first- and second-order dependencies that can complicate the analysis of relational data in traditional GLMs.
 The adjacency matrix on the top left highlights a particular row to illustrate that these values may be more similar to each other than other values because each has a common sender.
@@
 For similar reasons one also needs to take into account the shared dependence between observations that share a common receiver.
-The bottom-left panel, illustrates that sender and receiver type dependencies can also blend together.
+The bottom-left panel in Figure~\ref{fig:adjMatDeps} illustrates that sender and receiver type dependencies can also blend together.
 Specifically, actors who are more likely to send ties in a network tend to also be more likely to receive them.
@@
 For example, consider that trade flows both from and to many wealthy, developed countries.
-The bottom-right panel, highlights a second-order dependence, specifically, reciprocity.
+The bottom-right panel highlights a second-order dependence, specifically, reciprocity.
 This is a dependency occurring within dyads involving the same actors whereby values of $y_{ij}$ and $y_{ji}$ are correlated.
@@
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework, and the latter piece also details how it contrasts with alternative network-based approaches.
+\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework and the latter article details how it contrasts with alternative network-based approaches.
 Here we provide a brief review and then move to a simulation exercise.

### 2018-01-08 paper/4_app.tex
small scale edits

@@
 		Weeks (2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
-		Gibler (2017) & Logit & 1816--2008 &193 &   $650,557$ & Undirected & none \\ \bottomrule
+		Gibler (2017) & Logit & 1816--2008 &193 &   $650,557$ & Undirected & None \\ \bottomrule
 	\end{tabular}
@@
 
-Next we discuss, each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.
+Next we discuss each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.
 
@@
 They also discover that dictators are likely to challenge democracies, but that the converse is not true.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.
+In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, rather than the other way around.
 Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
@@
 The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
-Howevever, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
+However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
 In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
@@
 She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case.
+When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case.
 Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
 To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
+One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
 She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.

### 2018-01-08 paper/1_intro.tex
added discussion of symmetry, changed stuff about R&S/Rose
cite

@@
 
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.
-Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, International Studies Quarterly published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
@@
 Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012, king:roberts:2014}.
+%%CD: We need to add the citation for the ISQ thing.
+Couldn't figure out how to cite the special issue and by that I mean I had an easier time finding the symposium than the issue.
 
@@
 The multiplicative effects capture dependencies that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
+%% MG:This point is important and needs more attention
 

### 2018-01-08 paper/4_app.tex
mcd

@@
 \citet{mcdonald:2004} measures free trade in two ways.
-The first captures the idea that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
+The first measurement reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
 This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure captures the score of the state in the dyad that possesses higher barriers to trade.
+This measure thus captures the score of the state in the dyad that possesses higher barriers to trade.
 \citet[p.

### 2018-01-09 paper/2_model.tex
fix source of compilation error

@@
 The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.} Note that this model easily generalizes to the case, common in international relations, where interactions are unidrected (for example the presence of conflict or a bilateral investment treaty).
-In this case, $\rho$ is constrained to be 1, sender and receiver effects are identical for each observation, as are \mathbf{u_{i}}$ and $\mathbf{v_{j}}$.
+In this case, $\rho$ is constrained to be 1, sender and receiver effects are identical for each observation, as are $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$.
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.

### 2018-01-09 paper/1_intro.tex
need to add in third order sentence

@@
 Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012, king:roberts:2014}.
-%%CD: We need to add the citation for the ISQ thing.
-Couldn't figure out how to cite the special issue and by that I mean I had an easier time finding the symposium than the issue.
 
-In this article, we discuss the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.
@@
 The multiplicative effects capture dependencies that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
-%% MG:This point is important and needs more attention
 

### 2018-01-09 paper/1_intro.tex
intro and model edits

@@
 We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-The multiplicative effects capture dependencies that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
+Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we  understand that the United States invaded Iraq in 2003, and we understand that Poland often participates in US led coalitions.
+The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
 

### 2018-01-09 paper/2_model.tex
intro and model edits

@@
 It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework and the latter article details how it contrasts with alternative network-based approaches.
-Here we provide a brief review and then move to a simulation exercise.
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework and the latter article details how it contrasts with alternative network-based approaches.
+Here we provide a brief review and then move to a simulation exercise.}
 

### 2018-01-25 paper/4_app.tex
need to add in third order sentence
some more discussion of false negatives and ame

@@
 
+An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
+These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
 Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy.
@@
 
-% They use the MID data, but note that:
-% \begin{quote}
-% 	We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation.
-This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force.
-We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets.
-\ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).
-% \end{quote}
-
 \citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
@@
 Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
-%An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
-These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
 
@@
 At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
-% Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.
 
@@
 However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
-In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
+The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
 While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
@@
 It is rare for studies in this area to provide performance statistics, but at the same time at least one of the studies is unable to identify a single case in spite of having almost a million observations.
-
-% i thinks that's already pretty explicit
-% Fifth, it is useful to suggest that a lot of what is \textit{known} about international relations based on studies of dyadic data should be taken with a grain of salt, awaiting a reassessment using methods that do not rely on an assumption that all the dyads are iid.

### 2018-01-29 paper/coverletter/netsMatter_coverLetter.tex
sync up

@@
 
-{Editorial Team of Political Analysis via submission portal}
+{Editorial Team of American Political Science Review via submission portal}
 
@@
 We evaluate the utility of the AME framework using a simulation based exercise and a replication of five recently published works from the field of International Relations.
-Through the simulation exercise, we show that in the presence of unobserved dependencies the model is able to still provide less biased estimates and better calibrated standard errors than extant approaches in the literature.
-Next, for almost each of the replications reestimated using the AME framework we find that the case for the key findings from the original works become much less compelling.
+Through the simulation exercise, we show that in the presence of unobserved dependencies the model is able to provide less biased estimates and better calibrated standard errors than extant approaches in the literature.
+Next, for nearly all of the replications reestimated using the AME framework we find that the case for the key findings from the original works become much less compelling.
 Additionally, we show through an out-of-sample cross-validation exercise that the AME approach uniformly outperforms each of the replicated approaches in capturing the data generating process of the event of interest.\\[1ex]

### 2018-02-02 paper/coverletter/netsMatter_coverLetter.tex
letter

@@
 This letter accompanies our submission of a manuscript for your consideration.
-The manuscript ``Taking Dyads Seriously'' introduces the Additive and Multiplicative Effects (AME) framework for conducting inference in the context of the dependencies that we often observe in dyadic data.\\[1ex]
+The manuscript ``Taking Dyads Seriously'' introduces the Additive and Multiplicative Effects (AME) framework for conducting inference in the context of the dependencies that we often observe in dyadic data.
+In submitting this paper, we feel as though it would be helpful to first provide some background information.
+\\[1ex]
 
-We evaluate the utility of the AME framework using a simulation based exercise and a replication of five recently published works from the field of International Relations.
+Among political scientists using statistical models to analyze network data there has emerged a dispute between the advocates of the Exponential Random Graph (ERGM) model and those building on the ``latent space'' approach pioneered by Hoff, Raftery, and Handcock.
+It is worth noting that this fight is not mirrored in the statistical or broader networks literature, as in those communities people recognize these models as having distinct goals.
+
+In political science, however, the dispute has become quite prevalent.
+Although we clearly work in the latent space world, this paper is not about engaging or extending the ERGM-latent space discussion.
+Rather, we evaluate the utility of the AME framework using a simulation based exercise and a replication of five recently published works from the field of International Relations.
 Through the simulation exercise, we show that in the presence of unobserved dependencies the model is able to provide less biased estimates and better calibrated standard errors than extant approaches in the literature.

### 2018-02-03 paper/1_intro.tex
starting up edits to front end

@@
 In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
-Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012, king:roberts:2014}.
+Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012}.
 
@@
 
-We begin with a brief review of these dependencies and the AME model.
-Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of the dependencies that arise in dyadic data.
-Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained using the AME framework.
+We begin with a discussion of these dependencies and an introduction to the AME model.
+Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data.
+Last, to highlight the utility of this approach, we apply the AME model to five recent studies in the international relations (IR) literature.
 The comparison reveals that in ignoring observational dependence, the standard approach overestimates the effect of key variables in the literature and underestimates their uncertainty.
-Consequently, the latent factor approach (AME) produces results that, at times, contradict those found in these studies in particular as well as the broader literatures from which they are drawn.
+Consequently, AME produces results that, at times, contradict those found in these studies in particular as well as the broader literatures from which they are drawn.
 Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
@@
 Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
-Most importantly, it facilitates concentration on the relations aspect of the field of international relations.
+Most importantly, it facilitates concentration on the relational aspect of the field of international relations.

### 2018-02-03 paper/2_model.tex
starting up edits to front end

--- 2018-01-09 paper/2_model.tex
@@
 
-In working with relational data, scholars typically begin by structuring these data as a set of dyadic observations stacked on top of one another.
+In working with relational data, scholars typically begin by structuring it as a set of dyadic observations stacked on top of one another.
 This makes sense if each observation is independent of the others.
@@
 It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.
-\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework and the latter article details how it contrasts with alternative network-based approaches.
-Here we provide a brief review and then move to a simulation exercise.
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.}
 
@@
 $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.} Note that this model easily generalizes to the case, common in international relations, where interactions are unidrected (for example the presence of conflict or a bilateral investment treaty).
+The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
+
+Note that this model easily generalizes to the case, common in international relations, where interactions are unidrected (for example the presence of conflict or a bilateral investment treaty).
 In this case, $\rho$ is constrained to be 1, sender and receiver effects are identical for each observation, as are $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$.

### 2018-02-03 paper/1_intro.tex
some changes and notes

@@
 
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, International Studies Quarterly published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
@@
 In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
-Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012}.
+Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
 

### 2018-02-03 paper/2_model.tex
some changes and notes

@@
 As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
+
 To uncover the underlying structure of relational data, it is helpful to shift towards restructuring dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
@@
 Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flow, that is sent from actor $i$ to actor $j$.
+% add a note about undirected data sharing a similar structure as well
 
@@
 	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
-	\caption{Visualization of network with meso-scopic features.}
+	\caption{Visualization of a network with meso-scopic features.}
+	% need a simpler caption here
 	\label{fig:thirdDeps}
@@
 The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are rarer.
-A prominent example of a network with this type of structure was found by \citet{adamic:glance:2005}, who visualized the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
+A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualizes the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
 \citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs linked to others that were politically similar.

### 2018-02-03 paper/netsMatter.tex
fix source of compilation error
some changes and notes

@@
 \begin{abstract}
-	%International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially dyadic data.
-Yet, much of the work conducted brushes aside the statistical complications that come with analyzing relational data, namely that the observations do not satisfy the conditional independence criterion.
-As a result, these studies often produce results with biased coefficient estimates and poorly calibrated standard errors, factors which have consequences for hypothesis testing.
-We present an alternative regression based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data using a set of additive and multiplicative effects within a Bayesian framework.
-To highlight the utility of this framework for studying relational data, we present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
-Additionally, we replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-For each study, our approach shows that conventional methods overstate the effect of key variables in the study of interstate conflict, underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance of their variables.
-Further, the new approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
-
 	Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
@@
 We first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
-In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to our alternative.
+In addition, we replicate five recent studies in recent international relations scholarship, comparing the standard approach to our alternative.
 For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables.

### 2018-02-04 paper/1_intro.tex
edits to intro to make app section more positive, edits to Reiter and Stam section

@@
 Last, to highlight the utility of this approach, we apply the AME model to five recent studies in the international relations (IR) literature.
-The comparison reveals that in ignoring observational dependence, the standard approach overestimates the effect of key variables in the literature and underestimates their uncertainty.
+The comparison reveals that in accounting for observational dependence AME produces more precise estimates and better calibrated confidence intervals for key variables in the literature.
 Consequently, AME produces results that, at times, contradict those found in these studies in particular as well as the broader literatures from which they are drawn.

### 2018-02-04 paper/4_app.tex
edits to intro to make app section more positive, edits to Reiter and Stam section

@@
 We choose five prominent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
-These studies are recent and have been cited over 100 times.
+These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.
+It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable.
+The AME is not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).
+
+These studies were also selected to be recent and prominent (cited over 100 times).
 Each of these pieces was published in a prominent journal and is well-known in the literature.
@@
 \citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent.
-They also discover that dictators are likely to challenge democracies, but that the converse is not true.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, rather than the other way around.
+Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state.
+Using directed dyads, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.
+In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true.
 Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
@@
 
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
+In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
 In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.

### 2018-02-05 paper/1_intro.tex
front sec edits, remake sim plots with nicer labs

@@
 
-In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to better reflect the interdependencies underlying the data generating process of dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.
-First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.
-Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
+First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still important to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
 We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we  understand that the United States invaded Iraq in 2003, and we understand that Poland often participates in US led coalitions.
+Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003, if we  understand that the United States invaded Iraq in 2003 and we understand that Poland often participates in US led coalitions.
 The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
@@
 The comparison reveals that in accounting for observational dependence AME produces more precise estimates and better calibrated confidence intervals for key variables in the literature.
-Consequently, AME produces results that, at times, contradict those found in these studies in particular as well as the broader literatures from which they are drawn.
+Consequently, AME produces results that, at times, differ from those found in these studies in particular as well as the broader literatures from which they are drawn.
 Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
@@
 Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
-Most importantly, it facilitates concentration on the relational aspect of the field of international relations.
+Most importantly, it facilitates concentration on the relational aspect of the field of international relations through a statistical framework that will be familiar to most scholars.

### 2018-02-05 paper/2_model.tex
front sec edits, remake sim plots with nicer labs

@@
 Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flow, that is sent from actor $i$ to actor $j$.
-% add a note about undirected data sharing a similar structure as well
+In many applications, scholars are interested in studying undirected (i.e., symmetric) outcomes in which there is no clear sender or receiver, these type of outcomes still can, and we would argue should, be studied using the type of framework we discuss below.
 
@@
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.
+In an undirected setting instead of studying sender and receiver heterogeneity, we would just be concerned with actor heterogeneity in general.
 
@@
 
-The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
-Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
-However, it is rarely the case that we can do this even for TSCS data, thus we often make modeling decisions to account for that structure.
-Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem.
-The study of international relations is founded on the relations among actors.
-Why ignore the interdependencies that led to the study of IR in the first place?
-
 \section{\textbf{Additive and Multiplicative Effect Models for Networks}}
@@
 It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.}
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.} The AME model accounts for higher-order dependencies using an approach similar to that of the recommender systems underlying the recommendation algorithms of companies such as Amazon \citep{resnick:varian:1997}.
 
 The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
-For details on the SRRM see \citet{li:loken:2002,dorff:minhas:2017}.
-An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
-The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure underlying the LFM.
-The AME model is specified as follows:
+For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
+The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure underlying the LFM.} The AME model is specified as follows:
 
@@
 $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both the presence and absence of homophily.}
+The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both positive and negative homophily.}
 
@@
  \end{itemize}
+
+The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
+Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
+However, it is rarely the case that we can do this even for TSCS data, thus we often make modeling decisions to account for that structure.
+Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
+Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem.
+The study of international relations is founded on the relations among actors.
+Why ignore the interdependencies that led to the study of IR in the first place?

### 2018-02-05 paper/3_sim.tex
some changes and notes
front sec edits, remake sim plots with nicer labs

@@
 
-The first model corresponds to the ``naive'' approach in which little is explicitly done to account for latent dependencies in dyadic data.
+The first model corresponds to the ``standard'' approach in which little is explicitly done to account for latent dependencies in dyadic data.
 In the second model, we account for dependencies in dyadic data using the AME framework described in the previous section.
@@
 In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and include it in our specification for $Y$.
-The oracle case provides an important benchmark for the naive and AME approaches.
+The oracle case provides an important benchmark for the standard and AME approaches.
 
 For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.\footnote{The value of $\gamma$ is also set to 1, which corresponds to a an example where the $W$ character is associated with homophily.} We conduct two sets of simulations, one in which the number of actors in the network is set to 50 and the other at 100.
-In total, we ran 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimated a naive, AME, and oracle model.
+In total, we ran 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimated a standard, AME, and oracle model.
 
@@
 The panels in the left show the results for when the number of actors is set to 50 and those on the right for 100; and the top pair of panels represents the estimates for $\mu$ while the bottom pair do the same for $\beta$.
-In each case, we find that the estimates for $\mu$ and $\beta$ produced by the naive approach are notably off from their true values.
+In each case, we find that the estimates for $\mu$ and $\beta$ produced by the standard approach are notably off from their true values.
 On the other hand, the AME model performs just as well as the oracle case in estimating the true values.
@@
 	\centering
-	\caption{Regression parameter estimates for the naive, AME, and oracle models from 1,000 simulations.
+	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
 Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
@@
 Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals.
-The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the naive approach performs poorly by comparison.
+The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the standard approach performs poorly by comparison.
 
@@
 	\centering
-	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
+	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
 	\label{fig:ameCalib}
@@
 
-Last, beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
 In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.
@@
 \end{figure}
+
+What this simulation has shown is that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
+
+% - in the presence of deendencies, ame can be used to easily perform accurate inference on exogenous parameters.
+the simulation here shows that the typial glm estimator has issues with bias and calibration
+% - apart from just allowing inference on xbeta, the random effects are not just capturing random noise but underlying structure that we can further explore.
+we illustrate some strategies for doing so in teh application section.

### 2018-02-05 code/1_simulation/2_ameBias.R

@@ -26,7 +26,7 @@ for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
 
 #
 modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
-modKey$clean = c('Naive', 'AME', 'Oracle')
+modKey$clean = c('Standard', 'AME', 'Oracle')
 ##############################
 
 ##############################
@@ -57,7 +57,7 @@ ameSimBias = ameSimBias %>% group_by(model, var, n) %>%
 # fix mod labels
 ameSimBias$model = modKey$clean[match(ameSimBias$model, modKey$dirty)]
 ameSimBias$model = factor(ameSimBias$model, levels=modKey$clean)
-modCols = c(Naive='#d6604d', AME='#4393c3', Oracle='#4daf4a')
+modCols = c(Standard='#d6604d', AME='#4393c3', Oracle='#4daf4a')
 
 # fix var labels
 ameSimBias$var[ameSimBias$var=='X1'] = '$\\beta$'

### 2018-02-05 paper/4_app.tex
small edit to top of app

@@
 
-We choose five prominent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+We choose five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
 These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.
-It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable.
+It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable  -- for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
 The AME is not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).
 
-These studies were also selected to be recent and prominent (cited over 100 times).
+These studies were also selected based on how recent they were and whether they had more then 100 citations.
 Each of these pieces was published in a prominent journal and is well-known in the literature.

### 2018-02-05 paper/2_model.tex
modeling + sim section edits, recompile

@@
 It is flexible and easy to use for analyzing the kind of relational data often found in social science.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.} The AME model accounts for higher-order dependencies using an approach similar to that of the recommender systems underlying the recommendation algorithms of companies such as Amazon \citep{resnick:varian:1997}.
-
-The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
 For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
-The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure underlying the LFM.} The AME model is specified as follows:
+The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.} The AME model is specified as follows:
 
@@
 $K$ denotes the dimensions of the latent space.
+The construction of the LFM here is actually quite similar to the recommender systems that companies like Amazon and Netflix have used to model customer behavior \citep{resnick:varian:1997,bennett:etal:2007}.
 This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
-The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third-order interdependencies \citep{minhas:etal:2016:arxiv}.
+The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third-order interdependencies.
 The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network.
@@
 
-Note that this model easily generalizes to the case, common in international relations, where interactions are unidrected (for example the presence of conflict or a bilateral investment treaty).
-In this case, $\rho$ is constrained to be 1, sender and receiver effects are identical for each observation, as are $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$.
+Note that this model easily generalizes to the case, common in IR, where interactions are undirected (for example the presence of conflict or a bilateral investment treaty).
+In the case of the SRRM, $\rho$ is constrained to be one and instead of separate sender and receiver random effects a single actor random effect is utilized.
+For the LFM, an eigendecomposition scheme is used to capture higher order dependence patterns.
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.

### 2018-02-05 paper/3_sim.tex
modeling + sim section edits, recompile

@@
 We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
+Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
 The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
@@
 $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
-We compare inference for $\mu$ and $\beta$---the latter parameter would be of primarily theoretical concern for applied scholars---using three models :
+We compare inference for $\mu$ and $\beta$---the latter parameter would be of primarily theoretical concern for applied scholars---using three models:
 
 \begin{itemize}
-	\item the standard international relations approach assuming independent errors; 
-	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);
+	\item the ``standard'' international relations approach assuming independent errors; 
+	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);\footnote{Results with higher values of $K$ are similar.}
 	\item and an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$. 
@@
 In each case, we find that the estimates for $\mu$ and $\beta$ produced by the standard approach are notably off from their true values.
-On the other hand, the AME model performs just as well as the oracle case in estimating the true values.
+On the other hand, the AME model performs just as well as the oracle case in estimating the true parameter values.
 
@@
 The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the standard approach performs poorly by comparison.
+The implication of the results presented in Figures~\ref{fig:ameBias} and ~\ref{fig:ameCalib} is that standard approaches can often fail at estimating parameter values and conducting inferential tasks in the presence of unobserved dependencies.
+The AME approach by comparison can be used as a tool for scholars working with dyadic data to still estimate the true effects of their main variables of interest, while accounting for dependencies that do often emerge in dyadic data.
 
@@
 
-In the case of the data generating process for $Y$, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.
+Moreover, the AME approach allows scholars to better understand what parameters their model may be missing.
+In the case of the simulation here, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.
 Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$.
@@
 What this simulation has shown is that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
-
-% - in the presence of deendencies, ame can be used to easily perform accurate inference on exogenous parameters.
-the simulation here shows that the typial glm estimator has issues with bias and calibration
-% - apart from just allowing inference on xbeta, the random effects are not just capturing random noise but underlying structure that we can further explore.
-we illustrate some strategies for doing so in teh application section.
+Scholars can use this framework in an iterative fashion where once they have estimated a model, they can empirically study the structure of the random effects portion of the model to assess whether there are unobserved covariates that they want to account for in their models.
+At the very least, what this simulation section shows is that the AME model can help in estimating the true effect and conducting inference on the independent variables that scholars have operationalized to test their theoretical propositions.

### 2018-02-05 paper/4_app.tex
modeling + sim section edits, recompile

@@
 These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.
-It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable  -- for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
+It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
 The AME is not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).

### 2018-02-05 paper/4_app.tex
edits to rose

@@
 It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
-The AME is not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).
+The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).
 
@@
 While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
+Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error.
+\citep{tomz:etal:2007}.
+While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
 The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies.
-Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
+Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.}  Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
 When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.

### 2018-02-07 paper/1_intro.tex
incorp some of mf edits

@@
 \section{\textbf{Introduction}}
-
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
-However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters, so as to reflect the potential clustering of cases.
-In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
-Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
 
@@
 
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
+However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
+Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters, so as to reflect the potential clustering of cases.
+In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
+Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
+
 We begin with a discussion of these dependencies and an introduction to the AME model.
@@
 Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
-Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.
-Most importantly, it facilitates concentration on the relational aspect of the field of international relations through a statistical framework that will be familiar to most scholars.
+
+The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while easily testing the substantive effect of variables of interest.
+Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but to better model the data generating process behind events of interest.
+Most importantly, it concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-02-07 paper/2_model.tex
incorp some of mf edits

@@
 
-In working with relational data, scholars typically begin by structuring it as a set of dyadic observations stacked on top of one another.
+Scholars typically begin with dyadic data by stacking observations associated with each dyad on top of one another.
 This makes sense if each observation is independent of the others.
@@
 The other major type of meso-scopic feature that emerges in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
-This concept simply refers to the idea that groups of nodes that act similarly in the network are stochastically equivalent.
+Stochastic equivalence refers to a type of pattern in which actors can be divided into groups such that members of the same group have similar patterns of relationships.
 In the example we have laid out above each of the left leaning blogs would be considered stochastically equivalent to one another.

### 2018-02-08 paper/4_app.tex
a few changes to the aps section

@@
 These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.
-It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
-The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).
+The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
 
-These studies were also selected based on how recent they were and whether they had more then 100 citations.
-Each of these pieces was published in a prominent journal and is well-known in the literature.
+These studies were also selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+Of course, by the criteria of influence many other papers could have been considered for the replication.
+However, as decribed in the following sections, the AME model has specific data requirements.}  Each of these pieces was published in a prominent journal and is well-known in the literature.
 Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.}  It is worth noting that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
+The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.
 

### 2018-02-09 paper/1_intro.tex
intro changes

@@
 \section{\textbf{Introduction}}
-
-In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to better reflect the interdependencies underlying the data generating process of dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
-We focus on three types of interdependencies that can complicate dyadic analyses.
-First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still important to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
-We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003, if we  understand that the United States invaded Iraq in 2003 and we understand that Poland often participates in US led coalitions.
-The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
 
@@
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters, so as to reflect the potential clustering of cases.
+Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters, to reflect the potential clustering of cases.
 In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
 Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
+
+In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect better the interdependencies underlying the data generating process of dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+We focus on three types of interdependencies that can complicate dyadic analyses.
+First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still essential to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
+We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
+Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
+The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
 
@@
 Last, to highlight the utility of this approach, we apply the AME model to five recent studies in the international relations (IR) literature.
-The comparison reveals that in accounting for observational dependence AME produces more precise estimates and better calibrated confidence intervals for key variables in the literature.
-Consequently, AME produces results that, at times, differ from those found in these studies in particular as well as the broader literatures from which they are drawn.
-Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with.
+The comparison reveals that in accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature.
+Consequently, AME produces results that, at times, differ from those found in these studies in particular as well as the broader literature from which they are drawn.
+Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies.
 Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
@@
 The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while easily testing the substantive effect of variables of interest.
-Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but to better model the data generating process behind events of interest.
+Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but to sufficiently model the data generating process behind political events of interest.
 Most importantly, it concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-02-09 paper/2_model.tex
note about example
incorp example from juan
model edits 1

@@
 This makes sense if each observation is independent of the others.
-For example, a conflict sent from the United States to Japan, is assumed to be independent of any action that Japan may send to the United States.
+For example, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.
 Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
-While the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others, we know this assumption to be false in theory and in practice.
+While most scholars begin with the assumption that each dyadic interaction is taking place in isolation of the others, we know this assumption to be false both in theory and in practice.
 Relational data comes with an explicit structure that generally leads to particular types of dependencies.
@@
 
-To uncover the underlying structure of relational data, it is helpful to shift towards restructuring dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
+To uncover the underlying structure of relational data, it is helpful to restructure dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
 Rows designate the senders of an event and columns the receivers.
@@
 In most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook), we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
+For example, in an analysis of international trade certain countries (e.g., China) export in much larger volumes than other countries for a variety of structural, contextual, and idiosyncratic reasons.
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.
@@
 As a result, the rows and columns in an adjacency matrix are often correlated.
-For example, consider that trade flows both from and to many wealthy, developed countries.
+For example, consider trade flows both from and to many wealthy, developed countries.
 The bottom-right panel highlights a second-order dependence, specifically, reciprocity.
@@
 More often we find significant evidence of higher-order structures that result from dependencies between multiple groups of actors.
-These dependencies arise because there may be a or some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
+These dependencies arise because there may be some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
 In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
@@
 
-The visualization illustrates that the actors belonging to the same group have a higher likelihood of having an interaction with each other, whereas interactions across groups are rarer.
-A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualizes the ways in which right and left leaning political blogs linked to one another preceding the 2004 United States Presidential Election.
-\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs linked to others that were politically similar.
+% hmmm should use an actual IR example here .....
+The visualization illustrates that the actors belonging to the same group have a higher likelihood of interacting with each other, whereas interactions across groups are rarer.
+A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualizes linkages between right and left leaning political blogs preceding the 2004 United States Presidential Election.
+\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs were linked to others that were politically similar.
 This showcases the types of higher-order dependencies that can emerge in relational data.
@@
 Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
-The other major type of meso-scopic feature that emerges in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
+
+The other major type of meso-scopic features that emerge in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
 Stochastic equivalence refers to a type of pattern in which actors can be divided into groups such that members of the same group have similar patterns of relationships.
 In the example we have laid out above each of the left leaning blogs would be considered stochastically equivalent to one another.
+We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another.
+In this case, countries with (formal or informal) alliances would be stochastically equivalent to each other.
 

### 2018-02-09 paper/2_model.tex
model2

@@
 We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another.
-In this case, countries with (formal or informal) alliances would be stochastically equivalent to each other.
+In this case, countries with (formal or informal) alliances are a member of the same group due to similar conflictual ties to other actors in the system.
 
@@
 
-The LFM contribution to the AME comes in the multiplicative term: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$.
+The LFM contribution to the AME is in the multiplicative term: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$.
 $K$ denotes the dimensions of the latent space.
@@
 The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network.
-When taking the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$.
+When computing the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$.
 This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network or, for example, how similar actors are in terms of who they initiate conflict with.
@@
 In the case of the SRRM, $\rho$ is constrained to be one and instead of separate sender and receiver random effects a single actor random effect is utilized.
-For the LFM, an eigendecomposition scheme is used to capture higher order dependence patterns.
+For the LFM, an eigen-decomposition scheme is used to capture higher order dependence patterns.
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.
 
-By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the underlying structure in dyadic data that if left unestimated would complicate any inferences we might wish to draw for the exogenous parameters.
+By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the underlying structure in dyadic data that, if left un-estimated, would complicate any inferences we might wish to draw for the exogenous parameters.
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
@@
 The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
-Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then one can be assured that the conditional independence assumption underlying standard approaches will be satisfied.
-However, it is rarely the case that we can do this even for TSCS data, thus we often make modeling decisions to account for that structure.
+Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.
+However, it is rarely the case that this is possible even for TSCS data and thus modeling decisions must account for underlying structure.
 Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-Further by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem.
+Additionally, by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem.
 The study of international relations is founded on the relations among actors.

### 2018-02-09 paper/4_app.tex
some updates to the start of the applications section

@@
 \section{\textbf{Re-estimation with AME}}
+
+%% What is the point:
 
@@
 
-We choose five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+In order to showcase the insights the AME approach permits scholars to bring into their analysis, we illustrate the application of AME on five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
 These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.
-The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
+We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the method.
+Firstly, the results of AME estimation are interpretable alongside GLM results.
+Secondly, although the data requirements of AME are, admittedly, more intensive than needed for GLMs, this rich data is increasingly coming online in both new and existing data.
+Thirdly, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
 
@@
 Of course, by the criteria of influence many other papers could have been considered for the replication.
-However, as decribed in the following sections, the AME model has specific data requirements.}  Each of these pieces was published in a prominent journal and is well-known in the literature.
-Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables.
-In addition, most of these studies adjust the standard errors in an attempt to account for the clustering of observations.
-We follow the strategy of the original study in our re-estimation of the original models.
+However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
+However, reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression.
+These GLM models ignore dyadic interdependencies, except as they may reveal themselves in included variables, and then adjusting the standard errors in an attempt to account for the clustering of observations.
+
+In the following section, we present two sets of results for each study, one that follows the original study in our re-estimation of the original models and another that tests the central hypothesis within an AME implementation.
+We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependences in the data.
 

### 2018-02-09 paper/4_app.tex
more changes to ap section

@@
 
-In order to showcase the insights the AME approach permits scholars to bring into their analysis, we illustrate the application of AME on five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
-These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.
-We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the method.
-Firstly, the results of AME estimation are interpretable alongside GLM results.
-Secondly, although the data requirements of AME are, admittedly, more intensive than needed for GLMs, this rich data is increasingly coming online in both new and existing data.
-Thirdly, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
+In order to showcase the insights the AME approach permits scholars to
+bring into their analysis, we illustrate the application of AME on
+five recent studies from the broad field of international relations
+and international political economy that utilize relational data
+\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012,
+  gibler:2017}.  These articles demonstrate the broad range of
+application for the AME framework, which include testing the effect of
+both dyadic and monadic variables on varied conflict behaviors and
+trade.  We chose to demonstrate the capabilities of AME with reference
+to existing studies in order to highlight several features of the
+method. Firstly, the results of AME estimation are interpretable
+alongside GLM results. Secondly, although the data requirements of AME
+are, admittedly, more intensive than needed for GLMs, this rich data
+is increasingly coming online in both new and existing data. Thirdly,
+through AME, we can examine the extent to which interdependencies
+within typical dyadic data alter previous conclusions about
+international relations from recent empirical studies using a
+framework that a priori assumes conditional independence.
 
-These studies were also selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
-Of course, by the criteria of influence many other papers could have been considered for the replication.
-However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-However, reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression.
-These GLM models ignore dyadic interdependencies, except as they may reveal themselves in included variables, and then adjusting the standard errors in an attempt to account for the clustering of observations.
+These studies were also selected based on how recent they were and
+whether they had more then 100 citations.\footnote{Note that we chose
+  papers with at least 100 citations as an indicator of influence
+  within the discipline. Of course, by the criteria of influence many
+  other papers could have been considered for the
+  replication. However, as described in the following sections, the
+  AME model has specific data requirements, which limited the scope of
+  potential papers.}  Each of these pieces, published in prominent
+journals well-known in their respective literatures, posited a theory
+in which interdependencies are consequential. However, reflecting the
+dominant approach in the literature, each of the authors tested their
+hypothesis by employing some form of a general linearized
+regression.
 
-In the following section, we present two sets of results for each study, one that follows the original study in our re-estimation of the original models and another that tests the central hypothesis within an AME implementation.
-We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependences in the data.
+In the following section, we present two sets of results for each
+study, one that follows the original study in our re-estimation of the
+original models and another that tests the central hypothesis within
+an AME implementation. We demonstrate that AME allows scholars the
+ability to test their hypotheses of interest, while directly
+accounting for both theorized and observed interdependencies in the
+data. Directly accounting for interdependences can change the outcome
+of regression models, both in substance and in estimates. In part, this outcome
+follows from AME's ability to precisely capture higher-order
+interdependencies, whereas GLM approaches implicitly ignore dyadic
+interdependencies. As this is a widely-known limitation of GLM
+approaches, scholars often attempt to account for clustering of
+observations by including additional variables and adjusting the
+standard errors of the resulting estimates. At best, this method
+introduces noise and imprecision into results, and at worst can
+produce misleading outcomes.
 
@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.}  It is worth noting that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
-The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.
+We obtained the data for each of these studies from their replication
+archives and replicated the main results of each
+article.\footnote{Without exception this was straightforward to
+  accomplish, thanks to the authors'  and an increasing norm in the social sciences of
+  open data sharing.}  It is worth noting that the AME is applicable
+only where a full set of dyads for an outcome of interest is
+observable---for example it is unsuitable for studying onsets, where
+dyad-years representing conflict continuations are removed from the
+sample. The AME is also not applicable to analyses of case-level data,
+for example, studies that examine the decision to go to war by
+particular states.
 
-% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
-Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model.
+% SM: maybe too on the nose (Whether putative control variables, for
+% example, are precisely estimated is less important.) 
+%% MJF: 2/9-- took this part out, because I don't think that it's necessar.
+% Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model.
 For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of a dispute dyad being in a dispute.
-In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework.
-Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
+In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framewo
+%rk.
+Each of the five studies has a crucial finding that we hone in on to
+further draw into focus the analytical power of the AME estimation
+procedure.  In Table~\ref{tab:modelFindingSumm}, we present the
+overall results; the term \textit{Remains after accounting for
+  dependencies?} indicates only that the sign and/or significance of
+the putatively crucial finding in the original study is not found to
+hold in the AME estimation.\footnote{Full tabular results for each of
+  the original and reestimated models are presented in the Appendix.}
 
@@
 \centering
-\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network-based approach.}
+\caption{Here we provide a brief summary of the key variable in each
+  of the five replications and a note about whether or not the
+  highlighted finding remains when using our network-based approach.}
 	\begin{tabular}{l p{7cm} l} \toprule
-		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
-			& &  in a Network Model? \\ \toprule
+		\multirow{2}{*}{Study} & \multirow{2}{*}{Central
+                  Finding} &  Remains after \\ 
+			& &  accounting for dependencies? \\ \toprule
 		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
@@
 
-An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
-These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
-Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy.
-Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
+An important takeaway here is that as a result of the incompatibility
+between GLM models and the underlying processes that shape
+international events, many scholars are forced to make knowledge claims
+based on the statistical significance of a small set of covariates, or
+the differences between these covariates. These differences may change
+dramatically when interdependencies are taken into account directly,
+as they do in the case of this study. Beyond just comparing parameter
+estimates, we examine how well each approach can represent the data
+generating process using an out-of-sample cross validation
+strategy. Specifically, for each study, we randomly divide the data
+into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$
+is assigned.
+
 Then for each $s \in \{1,\ldots,k\}$, we:

### 2018-02-09 paper/4_app.tex
even more changes to ap section

@@
+\documentclass{article}[12pt]
+\usepackage{setspace}
+\usepackage{listings}
+\doublespacing
+\usepackage[letterpaper, portrait, margin=1in]{geometry}
+\setlength{\parskip}{.5em}
+\setlength{\parindent}{0cm}
+\usepackage{multirow}
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{framed}
+\usepackage[sort,round]{natbib}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\usepackage{hyperref}
+
+\begin{document}
+
 \section{\textbf{Re-estimation with AME}}
@@
 In order to showcase the insights the AME approach permits scholars to
-bring into their analysis, we illustrate the application of AME on
-five recent studies from the broad field of international relations
-and international political economy that utilize relational data
-\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012,
-  gibler:2017}.  These articles demonstrate the broad range of
-application for the AME framework, which include testing the effect of
-both dyadic and monadic variables on varied conflict behaviors and
-trade.  We chose to demonstrate the capabilities of AME with reference
+bring into their analysis, we apply AME to five recent studies from the broad field of international relations
+and international political economy~\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012,
+  gibler:2017}. Each of these studies are representative of broader trends in the field in that they use
+relational data of state interactions and propose both dyadic,
+monadic, and structural explanations for behavior of actors in the system.
+
+We chose to demonstrate the capabilities of AME with reference
 to existing studies in order to highlight several features of the
-method. Firstly, the results of AME estimation are interpretable
-alongside GLM results. Secondly, although the data requirements of AME
-are, admittedly, more intensive than needed for GLMs, this rich data
-is increasingly coming online in both new and existing data. Thirdly,
+AME approach. Firstly, the results of AME estimation are interpretable
+alongside GLM results but have the additional benefit of being able to
+quantify higher-order features of interest to scsholars, such as
+``sender'' and ``receiver'' effects. Secondly, although the data
+requirements of AME are, admittedly, more intensive than needed for GLMs, ever richer data
+is increasingly prevalent in both new and existing data. Thirdly,
 through AME, we can examine the extent to which interdependencies
@@
 
-These studies were also selected based on how recent they were and
+The five studies replicated below were selected based on how recent they were and
 whether they had more then 100 citations.\footnote{Note that we chose
@@
 In the following section, we present two sets of results for each
-study, one that follows the original study in our re-estimation of the
-original models and another that tests the central hypothesis within
-an AME implementation. We demonstrate that AME allows scholars the
-ability to test their hypotheses of interest, while directly
-accounting for both theorized and observed interdependencies in the
-data. Directly accounting for interdependences can change the outcome
-of regression models, both in substance and in estimates. In part, this outcome
-follows from AME's ability to precisely capture higher-order
-interdependencies, whereas GLM approaches implicitly ignore dyadic
-interdependencies. As this is a widely-known limitation of GLM
-approaches, scholars often attempt to account for clustering of
-observations by including additional variables and adjusting the
-standard errors of the resulting estimates. At best, this method
-introduces noise and imprecision into results, and at worst can
-produce misleading outcomes.
+study, one that replicates the original study and another that tests
+the central hypothesis within an AME implementation. We demonstrate
+that AME allows scholars the ability to test their hypotheses of
+interest, while directly accounting for both theorized and observed interdependencies in the
+data. As accounting for interdependences can change the outcome
+of regression models, both in substance and in estimate, we also summarize the
+predictive performance of the GLM and AME implementations. This
+predictive view allows us to better compare the models across
+specifications. Finally, for several of the studies, we report the
+AME-derived sender and receiver effects in order to highlight
+additional value of the AME framework.
 
@@
 based on the statistical significance of a small set of covariates, or
-the differences between these covariates. These differences may change
-dramatically when interdependencies are taken into account directly,
-as they do in the case of this study. Beyond just comparing parameter
+the differences between these covariates.  These differences may
+change dramatically when interdependencies are taken into account
+directly, as they do in the case of this study. %MJF: Keeping this here for now, but it reads more like a conclusion
+This outcome follows from AME's ability to precisely capture higher-order
+interdependencies, whereas GLM approaches implicitly ignore dyadic
+interdependencies. As this is a widely-known limitation of GLM
+approaches, scholars often attempt to account for clustering of
+observations by including additional variables and adjusting the
+standard errors of the resulting estimates. At best, this method
+introduces noise and imprecision into results, and at worst can
+produce misleading outcomes.   Beyond just comparing parameter
 estimates, we examine how well each approach can represent the data
@@
 
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-The link between conflict and trade is perplexing, with many persistent, yet competing explanations.
-\citet[p.
-547]{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''.
-Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade.
-\citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measures free trade in two ways.
-The first measurement reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
-This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-This measure thus captures the score of the state in the dyad that possesses higher barriers to trade.
-\citet[p.
-560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
-The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables.
-The models include splines to correct for temporal dependence, and robust standard errors clustered on each dyad.
+\citet{mcdonald:2004} studied whether trade promotes peace between
+nations.  \citet[p. 547]{mcdonald:2004} introduced the argument that
+free trade between states ``makes conflict less likely
+because of its efficiency over conquest in acquiring
+resources\ldots''. Accordingly, he provided evidence challenging the
+generalized linkage between peace and trade and refined the
+measurement of the standard ``trade'' variable, arguing that
+\textit{free} trade, rather than trade alone, reduces the likelihood
+of conflict between states. His key hypothesis is that greater levels
+of protection increase the probability of interstate conflict, an
+argument that builds on the work of classic liberalism and connects
+free trade to the power of domestic audiences. \citet{mcdonald:2004}
+measured free trade in two ways. The first, the ``protection''
+variable measures the proportion of customs revenue divided by total
+imports in the state that possesses the greater such ratio in each
+dyad. reflects the notion that larger protected sectors generate
+greater societal pressures resulting in pockets of support for war.
+This measure thus captures the score of the state in the dyad that
+possesses higher barriers to trade. \citet[p. 560]{mcdonald:2004} also
+includes a measure of economic integration  calculated as ``the lower
+proportion of total dyadic trade (imports plus exports) divided by
+state $i$'s GDP or total dyadic trade divided by state $j$'s
+GDP''. The (binary) dependent variable is the onset of a new
+militarized interstate dispute within a given dyad.
+ \citet{mcdonald:2004} employs logistic regression to examine the
+ putative statistical significance of these variables. The models
+ include splines to correct for temporal dependence, and robust
+ standard errors clustered on each dyad.
 
@@
 
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.
-These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.
-She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
+\citet{weeks:2012} examines the influence of domestic institutions on
+the initiation of military conflicts by autocratic leaders.  She
+argues that in some circumstances autocrats are held accountable for
+their foreign policy decisions. She adds the nuance that autocratic
+audiences are not homogeneous. When the autocratic regime is
+nonmilitary, domestic audiences do not favor military actions, but in
+military autocracies this is not the case. Further she argues that in
+personalistic regimes without a military or civilian domestic
+audience, the leaders tend to be more likely to employ military force
+in their foreign policy.  To study this question, she uses a dyadic
+design in which the dependent variable is ``whether country A in a
+directed dyad initiated military conflict against country B during
+year t'' (page 337).  One major innovation in her study resides in the
+nuanced way in which she conceptualizes and codes regime type into
+four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also
+includes a variety of putative control variables focusing on
+capabilities for both sides of the dyad, alliances, geography, trade
+dependence, regime instability, and the regime type of ``side B.''
+She uses a logistic regression, but follows \citet{beck:etal:1998} and
+includes splines to capture temporal covariation in the dependent
+variable along with fixed, unit effects. The analysis is done for
+dyads, but is considered to be from the perspective of the actor that
+initiated the dispute. The basic finding is that a) juntas, boss type,
+and strongmen type regimes are more likely to initiate conflict than
+machine-type regimes (and maybe democracies) and that b) machine-type
+regimes are no more belligerent than democracies.  
+These insights are mainly emphasized in the paper by the parameter
+estimates depicted in Tables 1 and 2 (pages 339--340) from the
+paper. She argued that ignoring important nuances between
+different types of autocracies hinders our understanding of the
+initiation of military conflict by autocracies.
 
@@
 
-The replications we have undertaken are all from articles over the past fifteen years.
 A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
@@
 It is rare for studies in this area to provide performance statistics, but at the same time at least one of the studies is unable to identify a single case in spite of having almost a million observations.
+
+\end{document}

### 2018-02-10 paper/1_intro.tex
edited entire manuscript

@@
 We focus on three types of interdependencies that can complicate dyadic analyses.
-First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still essential to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior.
+First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still essential to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict against a government, the government will likely reciprocate that behavior.
 We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
+Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads--for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
 The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.

### 2018-02-10 paper/2_model.tex
edited entire manuscript

@@
 
-Scholars typically begin with dyadic data by stacking observations associated with each dyad on top of one another.
+Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another.
 This makes sense if each observation is independent of the others.
@@
 Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
+For example in International Trade, we might observe homophily wherein states with open economies and an absence of Non-Tariff Barriers are more likely to trade with each-other all else equal; alternatively we might observe heterophily in the sense that states that have greater endowments of capital are more likely to trade with states with greater endowments of land or labor.
 

### 2018-02-11 prezzy/netsMatterDeck.tex
prezzy

@@
 Dorff (University of New Mexico), Max Gallop (University of Strathclyde), Juan Tellez (Duke University), Howard Liu (Duke University), \& Margaret Foster (Duke University).
-This is also part of a broader project involving Peter Hoff (Duke University) and Michael Ward.}
-}
+}}
 
-\author{Shahryar Minhas\\Cassy L. Dorff\\Margaret Foster\\Max Blau Gallop\\Juan Tellez\\ Michael D. Ward}
+\author{Michael D. Ward}
 %{Shahryar Minhas\\ minhassh@msu.edu \\ s7minhas.com \\
@@
 
-\date{\today}
+\date{\today\\ 2018 Asian Political Methodology Meeting\\Asia Center\\Seoul National University\\Seoul, Korea}
 
@@
 
-Thank you!
+Thank you!  Thanks especially to Jong Hee Park and the Organizing Committee.
 
@@
   \frametitle{Actual Code}
+  {\footnotesize
 \# running in parallel varying k\\
 imps = 10000 ; brn = 25000 ; ods = 10 ; latDims = 0:3\\
-\# Run amen in parallel
 library(doParallel) ; library(foreach) ; cl=makeCluster(4) \\ registerDoParallel(cl)\\
@@
 stopCluster(cl)
+}
 }

### 2018-02-11 paper/3_sim.tex

@@
 
-We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
-Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
+We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM).
+While both these models have led to numerous contributions to a variety of literatures, the applicability of these approaches may be limited to certain types of networks and individual level characteristics.
+Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
 The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable.
+Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an unaccounted for higher order dependence pattern.
 For instance, assume that the true data-generating process for a particular $Y$ is given by:
@@
 \begin{itemize}
-	\item the ``standard'' international relations approach assuming independent errors; 
+	\item the ``standard'' international relations approach estimated through a typical generalized linear model; 
 	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);\footnote{Results with higher values of $K$ are similar.}

### 2018-02-11 paper/4_app.tex

@@
 alongside GLM results but have the additional benefit of being able to
-quantify higher-order features of interest to scsholars, such as
+quantify higher-order features of interest to scholars, such as
 ``sender'' and ``receiver'' effects. Secondly, although the data
@@
 journals well-known in their respective literatures, posited a theory
-in which interdependencies are consequential. However, reflecting the
+in which interdependencies are consequential. Reflecting the
 dominant approach in the literature, each of the authors tested their
@@
 predictive view allows us to better compare the models across
-specifications. Finally, for several of the studies, we report the
+specifications (and where they differ in their coefficients, see which specification does better at modeling the underlying data-generating process). Finally, for several of the studies, we report the
 AME-derived sender and receiver effects in order to highlight
@@
 article.\footnote{Without exception this was straightforward to
-  accomplish, thanks to the authors'  and an increasing norm in the social sciences of
+  accomplish, thanks to the authors' transparency  and an increasing norm in the social sciences of
   open data sharing.}  It is worth noting that the AME is applicable
@@
 		\multirow{2}{*}{Study} & \multirow{2}{*}{Central
-                  Finding} &  Remains after \\ 
+                  Finding} &  Confirmed \\ 
 			& &  accounting for dependencies? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
-		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
-		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
-		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Confirmed} \\ \midrule
+		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Unconfirmed}\\ \midrule
+		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Confirmed}\\ \midrule
+		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
+		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Unconfirmed}\\ \bottomrule
 	\end{tabular}
@@
 
-An important takeaway here is that as a result of the incompatibility
-between GLM models and the underlying processes that shape
-international events, many scholars are forced to make knowledge claims
-based on the statistical significance of a small set of covariates, or
+An important takeaway here is that %as a result of the incompatibility between GLM models and the underlying processes that shape international events, 
+many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or
 the differences between these covariates.  These differences may
@@
 \noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.\footnote{For more details on this type of cross-validation strategy see \citet{minhas:etal:2016:arxiv}.} We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
-For the binary models we just provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
+For the binary models we  provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
 Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
@@
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the effects uncovered by the authors remains robust when accounting for dyadic interdependencies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.
+Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies.
+Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
 

### 2018-02-12 paper/1_intro.tex
intro edits

@@
 
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to test hypothesized relationships.
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
@@
 
-In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect better the interdependencies underlying the data generating process of dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to better reflect the interdependencies underlying the data generating process of dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.
@@
 
-The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while easily testing the substantive effect of variables of interest.
-Thus, the AME approach can be used by scholars in the field to not only generate substantive insights, but to sufficiently model the data generating process behind political events of interest.
+The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while enabling scholars to still focus on testing the substantive effect of variables of interest.
+Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while having the benefit of being able to better account for the data generating process behind political events of interest.
 Most importantly, it concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-02-12 paper/4_app.tex
intro edits

@@
-\documentclass{article}[12pt]
-\usepackage{setspace}
-\usepackage{listings}
-\doublespacing
-\usepackage[letterpaper, portrait, margin=1in]{geometry}
-\setlength{\parskip}{.5em}
-\setlength{\parindent}{0cm}
-\usepackage{multirow}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage[sort,round]{natbib}
-\usepackage{fancyhdr}
-\pagestyle{fancy}
-\usepackage{hyperref}
-
-\begin{document}
-
 \section{\textbf{Re-estimation with AME}}
-
-%% What is the point:
 
@@
 
-In order to showcase the insights the AME approach permits scholars to
-bring into their analysis, we apply AME to five recent studies from the broad field of international relations
-and international political economy~\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012,
-  gibler:2017}. Each of these studies are representative of broader trends in the field in that they use
-relational data of state interactions and propose both dyadic,
-monadic, and structural explanations for behavior of actors in the system.
+In order to showcase the insights the AME approach permits scholars to bring into their analysis, we apply AME to five recent studies from the broad field of international relations and international political economy~\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
 
-We chose to demonstrate the capabilities of AME with reference
-to existing studies in order to highlight several features of the
-AME approach. Firstly, the results of AME estimation are interpretable
-alongside GLM results but have the additional benefit of being able to
-quantify higher-order features of interest to scholars, such as
-``sender'' and ``receiver'' effects. Secondly, although the data
-requirements of AME are, admittedly, more intensive than needed for GLMs, ever richer data
-is increasingly prevalent in both new and existing data. Thirdly,
-through AME, we can examine the extent to which interdependencies
-within typical dyadic data alter previous conclusions about
-international relations from recent empirical studies using a
-framework that a priori assumes conditional independence.
+We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
+Firstly, the results of AME estimation are interpretable alongside GLM results but have the additional benefit of being able to quantify higher-order features of interest to scholars, such as ``sender'' and ``receiver'' effects.
+Secondly, although the data requirements of AME are, admittedly, more intensive than needed for GLMs, ever richer data is increasingly prevalent in both new and existing data.
+Thirdly, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
 
-The five studies replicated below were selected based on how recent they were and
-whether they had more then 100 citations.\footnote{Note that we chose
-  papers with at least 100 citations as an indicator of influence
-  within the discipline. Of course, by the criteria of influence many
-  other papers could have been considered for the
-  replication. However, as described in the following sections, the
-  AME model has specific data requirements, which limited the scope of
-  potential papers.}  Each of these pieces, published in prominent
-journals well-known in their respective literatures, posited a theory
-in which interdependencies are consequential. Reflecting the
-dominant approach in the literature, each of the authors tested their
-hypothesis by employing some form of a general linearized
-regression.
+The five studies replicated below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+Of course, by the criteria of influence many other papers could have been considered for the replication.
+However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
+Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression.
 
-In the following section, we present two sets of results for each
-study, one that replicates the original study and another that tests
-the central hypothesis within an AME implementation. We demonstrate
-that AME allows scholars the ability to test their hypotheses of
-interest, while directly accounting for both theorized and observed interdependencies in the
-data. As accounting for interdependences can change the outcome
-of regression models, both in substance and in estimate, we also summarize the
-predictive performance of the GLM and AME implementations. This
-predictive view allows us to better compare the models across
-specifications (and where they differ in their coefficients, see which specification does better at modeling the underlying data-generating process). Finally, for several of the studies, we report the
-AME-derived sender and receiver effects in order to highlight
-additional value of the AME framework.
+In the following section, we present two sets of results for each study, one that replicates the original study and another that tests the central hypothesis within an AME implementation.
+We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependencies in the data.
+As accounting for interdependences can change the outcome of regression models, both in substance and in estimate, we also summarize the predictive performance of the GLM and AME implementations.
+This predictive view allows us to better compare the models across specifications (and where they differ in their coefficients, see which specification does better at modeling the underlying data-generating process).
+Finally, for several of the studies, we report the AME-derived sender and receiver effects in order to highlight additional value of the AME framework.
 
@@
 
-We obtained the data for each of these studies from their replication
-archives and replicated the main results of each
-article.\footnote{Without exception this was straightforward to
-  accomplish, thanks to the authors' transparency  and an increasing norm in the social sciences of
-  open data sharing.}  It is worth noting that the AME is applicable
-only where a full set of dyads for an outcome of interest is
-observable---for example it is unsuitable for studying onsets, where
-dyad-years representing conflict continuations are removed from the
-sample. The AME is also not applicable to analyses of case-level data,
-for example, studies that examine the decision to go to war by
-particular states.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency  and an increasing norm in the social sciences of open data sharing.}  It is worth noting that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
+The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.
 
@@
 %rk.
-Each of the five studies has a crucial finding that we hone in on to
-further draw into focus the analytical power of the AME estimation
-procedure.  In Table~\ref{tab:modelFindingSumm}, we present the
-overall results; the term \textit{Remains after accounting for
-  dependencies?} indicates only that the sign and/or significance of
-the putatively crucial finding in the original study is not found to
-hold in the AME estimation.\footnote{Full tabular results for each of
-  the original and reestimated models are presented in the Appendix.}
+
+Each of the five studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Remains after accounting for dependencies?} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
 
@@
 
-An important takeaway here is that %as a result of the incompatibility between GLM models and the underlying processes that shape international events, 
-many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or
-the differences between these covariates.  These differences may
-change dramatically when interdependencies are taken into account
-directly, as they do in the case of this study. %MJF: Keeping this here for now, but it reads more like a conclusion
-This outcome follows from AME's ability to precisely capture higher-order
-interdependencies, whereas GLM approaches implicitly ignore dyadic
-interdependencies. As this is a widely-known limitation of GLM
-approaches, scholars often attempt to account for clustering of
-observations by including additional variables and adjusting the
-standard errors of the resulting estimates. At best, this method
-introduces noise and imprecision into results, and at worst can
-produce misleading outcomes.   Beyond just comparing parameter
-estimates, we examine how well each approach can represent the data
-generating process using an out-of-sample cross validation
-strategy. Specifically, for each study, we randomly divide the data
-into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$
-is assigned.
+% as a result of the incompatibility between GLM models and the underlying processes that shape international events, 
+%MJF: Keeping this here for now, but it reads more like a conclusion
+
+An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
+These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
+This outcome follows from AME's ability to precisely capture higher-order interdependencies, whereas GLM approaches implicitly ignore dyadic interdependencies.
+As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates.
+At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes.
+Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy.
+Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
 
@@
 
-\citet{mcdonald:2004} studied whether trade promotes peace between
-nations.  \citet[p. 547]{mcdonald:2004} introduced the argument that
-free trade between states ``makes conflict less likely
-because of its efficiency over conquest in acquiring
-resources\ldots''. Accordingly, he provided evidence challenging the
-generalized linkage between peace and trade and refined the
-measurement of the standard ``trade'' variable, arguing that
-\textit{free} trade, rather than trade alone, reduces the likelihood
-of conflict between states. His key hypothesis is that greater levels
-of protection increase the probability of interstate conflict, an
-argument that builds on the work of classic liberalism and connects
-free trade to the power of domestic audiences. \citet{mcdonald:2004}
-measured free trade in two ways. The first, the ``protection''
-variable measures the proportion of customs revenue divided by total
-imports in the state that possesses the greater such ratio in each
-dyad. reflects the notion that larger protected sectors generate
-greater societal pressures resulting in pockets of support for war.
-This measure thus captures the score of the state in the dyad that
-possesses higher barriers to trade. \citet[p. 560]{mcdonald:2004} also
-includes a measure of economic integration  calculated as ``the lower
-proportion of total dyadic trade (imports plus exports) divided by
-state $i$'s GDP or total dyadic trade divided by state $j$'s
-GDP''. The (binary) dependent variable is the onset of a new
-militarized interstate dispute within a given dyad.
- \citet{mcdonald:2004} employs logistic regression to examine the
- putative statistical significance of these variables. The models
- include splines to correct for temporal dependence, and robust
- standard errors clustered on each dyad.
+\citet{mcdonald:2004} studied whether trade promotes peace between nations.
+\citet[p.
+547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''.
+Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
+His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
+\citet{mcdonald:2004} measured free trade in two ways.
+The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
+reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
+This measure thus captures the score of the state in the dyad that possesses higher barriers to trade.
+\citet[p.
+560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
+The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad.
+\citet{mcdonald:2004} employs logistic regression to examine the  putative statistical significance of these variables.
+The models  include splines to correct for temporal dependence, and robust  standard errors clustered on each dyad.
 
@@
 
-\citet{weeks:2012} examines the influence of domestic institutions on
-the initiation of military conflicts by autocratic leaders.  She
-argues that in some circumstances autocrats are held accountable for
-their foreign policy decisions. She adds the nuance that autocratic
-audiences are not homogeneous. When the autocratic regime is
-nonmilitary, domestic audiences do not favor military actions, but in
-military autocracies this is not the case. Further she argues that in
-personalistic regimes without a military or civilian domestic
-audience, the leaders tend to be more likely to employ military force
-in their foreign policy.  To study this question, she uses a dyadic
-design in which the dependent variable is ``whether country A in a
-directed dyad initiated military conflict against country B during
-year t'' (page 337).  One major innovation in her study resides in the
-nuanced way in which she conceptualizes and codes regime type into
-four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also
-includes a variety of putative control variables focusing on
-capabilities for both sides of the dyad, alliances, geography, trade
-dependence, regime instability, and the regime type of ``side B.''
-She uses a logistic regression, but follows \citet{beck:etal:1998} and
-includes splines to capture temporal covariation in the dependent
-variable along with fixed, unit effects. The analysis is done for
-dyads, but is considered to be from the perspective of the actor that
-initiated the dispute. The basic finding is that a) juntas, boss type,
-and strongmen type regimes are more likely to initiate conflict than
-machine-type regimes (and maybe democracies) and that b) machine-type
-regimes are no more belligerent than democracies.  
-These insights are mainly emphasized in the paper by the parameter
-estimates depicted in Tables 1 and 2 (pages 339--340) from the
-paper. She argued that ignoring important nuances between
-different types of autocracies hinders our understanding of the
-initiation of military conflict by autocracies.
+\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
+She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
+She adds the nuance that autocratic audiences are not homogeneous.
+When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case.
+Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
+To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
+One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
+She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
+The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
+The basic finding is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes (and maybe democracies) and that b) machine-type regimes are no more belligerent than democracies.
+These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.
+She argued that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
 
@@
 It is rare for studies in this area to provide performance statistics, but at the same time at least one of the studies is unable to identify a single case in spite of having almost a million observations.
-
-\end{document}

### 2018-02-12 paper/2_model.tex
mod, sim edits

--- 2018-02-09 paper/2_model.tex
@@
 For example, in an analysis of international trade certain countries (e.g., China) export in much larger volumes than other countries for a variety of structural, contextual, and idiosyncratic reasons.
-Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.
-In an undirected setting instead of studying sender and receiver heterogeneity, we would just be concerned with actor heterogeneity in general.
+Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.\footnote{In an undirected setting instead of studying sender and receiver heterogeneity, we would just be concerned with actor heterogeneity in general.}
 
@@
 	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
-	\caption{Visualization of a network with meso-scopic features.}
+	\caption{Visualization of a network with higher-order dependence patterns.}
 	% need a simpler caption here
@@
 
-% hmmm should use an actual IR example here .....
 The visualization illustrates that the actors belonging to the same group have a higher likelihood of interacting with each other, whereas interactions across groups are rarer.
-A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualizes linkages between right and left leaning political blogs preceding the 2004 United States Presidential Election.
+A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualize linkages between right and left leaning political blogs preceding the 2004 United States Presidential Election.
 \citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs were linked to others that were politically similar.
@@
 Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
-
 The other major type of meso-scopic features that emerge in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}.
 Stochastic equivalence refers to a type of pattern in which actors can be divided into groups such that members of the same group have similar patterns of relationships.
-In the example we have laid out above each of the left leaning blogs would be considered stochastically equivalent to one another.
-We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another.
+In the example we have laid out above, each of the left leaning blogs would be considered stochastically equivalent to one another because any given left-leaning blog is more likely to interact with a blog of a similar political position and less likely to interact with one with a divergent political position.
+
+These types of patterns frequently emerge in IR contexts as well.\footnote{For example, see: \citet{manger:etal:2012, kinne:2013, chyzh:2016}.} For example, a perennial finding in the interstate trade literature emphasizes the role that geography plays in determining trade flows.
+Geographic proximity in the network context is an example of a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place.
+Alternatively, in the interstate conflict literature, we may find that actors who are each a member of a particular (formal or informal) alliance are likely to act similarly in the conflict network.
+Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance.
+In both these examples, we are able to explicitly parameterize the attribute that might explain the emergence of higher order dependence patterns, but it is likely that at times we will not able to account for the variety of reasons why higher order dependence patterns in networks may develop.
+
+% International Trade, we might observe homophily wherein states with open economies and an absence of Non-Tariff Barriers are more likely to trade with each-other all else equal; alternatively we might observe heterophily in the sense that states that have greater endowments of capital are more likely to trade with states with greater endowments of land or labor.
+
+%  We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another.
 In this case, countries with (formal or informal) alliances are a member of the same group due to similar conflictual ties to other actors in the system.

### 2018-02-12 paper/3_sim.tex
mod, sim edits

@@
 
-The first model corresponds to the ``standard'' approach in which little is explicitly done to account for latent dependencies in dyadic data.
-In the second model, we account for dependencies in dyadic data using the AME framework described in the previous section.
+The first model corresponds to the ``standard'' approach in which little is explicitly done to account for dependencies in dyadic data.
+In the second model, we use the AME framework described in the previous section.
 For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated in the presence of unobserved dependencies, $W$.

### 2018-02-12 paper/4_app.tex
app section

@@
-\section{\textbf{Re-estimation with AME}}
+\section{\textbf{Applications with AME}}
 
@@
 
-In order to showcase the insights the AME approach permits scholars to bring into their analysis, we apply AME to five recent studies from the broad field of international relations and international political economy~\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+In order to showcase the insights the AME approach permits scholars to bring into their analysis, we apply AME to five recent studies from international relations and international political economy: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
 Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
+We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
+First, the results of AME estimation are interpretable alongside GLM results but have the additional benefit of being able to quantify higher-order features of interest to scholars, such as ``sender'' and ``receiver'' effects.
+Second, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
+Third, we will show that by using the AME framework scholars can better model the data generating process behind their events of interest.
 
-We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
-Firstly, the results of AME estimation are interpretable alongside GLM results but have the additional benefit of being able to quantify higher-order features of interest to scholars, such as ``sender'' and ``receiver'' effects.
-Secondly, although the data requirements of AME are, admittedly, more intensive than needed for GLMs, ever richer data is increasingly prevalent in both new and existing data.
-Thirdly, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
-
-The five studies replicated below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+The five studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.
 However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression.
+Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.
 
-In the following section, we present two sets of results for each study, one that replicates the original study and another that tests the central hypothesis within an AME implementation.
+% In the following section, we present two sets of results for each study, one that replicates the original study and another that tests the central hypothesis within an AME implementation.
 We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependencies in the data.
@@
 
-% SM: maybe too on the nose (Whether putative control variables, for
-% example, are precisely estimated is less important.) 
-%% MJF: 2/9-- took this part out, because I don't think that it's necessar.
-% Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model.
-For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of a dispute dyad being in a dispute.
-In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framewo
-%rk.
-
 Each of the five studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
@@
 \centering
-\caption{Here we provide a brief summary of the key variable in each
-  of the five replications and a note about whether or not the
-  highlighted finding remains when using our network-based approach.}
+\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
 	\begin{tabular}{l p{7cm} l} \toprule
-		\multirow{2}{*}{Study} & \multirow{2}{*}{Central
-                  Finding} &  Confirmed \\ 
+		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed \\
 		& &  accounting for dependencies? \\ \toprule
@@
 \end{table}
-
-% as a result of the incompatibility between GLM models and the underlying processes that shape international events, 
-%MJF: Keeping this here for now, but it reads more like a conclusion
 
@@
 In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true.
-Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures.
+Independent variables focus on various encodings of regime types, contiguity, alliance, and capability measures.
 As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.

### 2018-02-12 paper/4_app.tex
starting to bring in cassy's changes

@@
 Each of the five studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Remains after accounting for dependencies?} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
 

### 2018-02-12 paper/2_model.tex
more cleanup

@@
 These types of patterns frequently emerge in IR contexts as well.\footnote{For example, see: \citet{manger:etal:2012, kinne:2013, chyzh:2016}.} For example, a perennial finding in the interstate trade literature emphasizes the role that geography plays in determining trade flows.
-Geographic proximity in the network context is an example of a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place.
+Geographic proximity in the network context is an example of homophily --- a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place.
 Alternatively, in the interstate conflict literature, we may find that actors who are each a member of a particular (formal or informal) alliance are likely to act similarly in the conflict network.
-Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance.
+Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance --- an example of stochastic equivalence.
 In both these examples, we are able to explicitly parameterize the attribute that might explain the emergence of higher order dependence patterns, but it is likely that at times we will not able to account for the variety of reasons why higher order dependence patterns in networks may develop.

### 2018-02-12 paper/4_app.tex
more cleanup

@@
 
-In order to showcase the insights the AME approach permits scholars to bring into their analysis, we apply AME to five recent studies from international relations and international political economy: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+To showcase how the AME generates new insights for researchers, we apply AME to five recent studies from the broad field of international relations and international political economy: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
 Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
@@
 
-First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news---since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
+First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.
+Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
 This means that significance testing (for better or worse) is compromised when network effects are ignored.
@@
 In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.
-It is rare for studies in this area to provide performance statistics, but at the same time at least one of the studies is unable to identify a single case in spite of having almost a million observations.

### 2018-02-13 paper/4_app.tex
app edits

@@
 
-To showcase how the AME generates new insights for researchers, we apply AME to five recent studies from the broad field of international relations and international political economy: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+We apply AME to five recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
 Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
-We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
-First, the results of AME estimation are interpretable alongside GLM results but have the additional benefit of being able to quantify higher-order features of interest to scholars, such as ``sender'' and ``receiver'' effects.
-Second, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.
-Third, we will show that by using the AME framework scholars can better model the data generating process behind their events of interest.
+We choose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
+First, the results of AME estimation are interpretable alongside results using standard approaches, but, as shown in the simulation section, have the additional benefit of being able to take into account dependencies that may complicate inference.
+Second, through using this approach, we can also quantify the degree to which first, second, and third order dependencies are present in events of interest.
+Third, we show that by using the AME framework scholars can better model the data generating process behind their events of interest.
 
-The five studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The five studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.
 However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.
-
-% In the following section, we present two sets of results for each study, one that replicates the original study and another that tests the central hypothesis within an AME implementation.
-We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependencies in the data.
-As accounting for interdependences can change the outcome of regression models, both in substance and in estimate, we also summarize the predictive performance of the GLM and AME implementations.
-This predictive view allows us to better compare the models across specifications (and where they differ in their coefficients, see which specification does better at modeling the underlying data-generating process).
-Finally, for several of the studies, we report the AME-derived sender and receiver effects in order to highlight additional value of the AME framework.
+Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.\footnote{It is important to note that the AME framework is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
+The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.}
 
@@
 \end{table}
-
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency  and an increasing norm in the social sciences of open data sharing.}  It is worth noting that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
-The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.
 
@@
 These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
-This outcome follows from AME's ability to precisely capture higher-order interdependencies, whereas GLM approaches implicitly ignore dyadic interdependencies.
+This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates.
 As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates.
@@
 
-\noindent The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.\footnote{For more details on this type of cross-validation strategy see \citet{minhas:etal:2016:arxiv}.} We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
+The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.\footnote{For more details on this type of cross-validation strategy see \citet{minhas:etal:2016:arxiv}.} We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
 For the binary models we  provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
@@
 
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero.
-In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero.
+In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) has a positive effect while the variable ``Democ/Personalist Directed Dyad'' is too imprecisely measured to indicate a direction.
+In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect while Democ-Pers directed dyad is still too imprecisely measured to indicate a direction.
 Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
 Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies.
-Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
+%Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
 
@@
 In 2004, Andrew Rose published a study in the \textit{American Economic Review} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007,ward:etal:2013}.
+It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007}.
 Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
@@
 While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.}  Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
-While the AME model replicates the main (null) result, there are substantial and substantive improvements gained from moving to an AME estimation.
 
@@
 \FloatBarrier
+
+The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
+Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.}  Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
+When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
+While the AME model replicates the main (null) result, there are substantial and substantive improvements gained from moving to an AME estimation.
 
@@
 
+\begin{figure}
+	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
+	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
+ 	\label{fig:gibmargeff}
+ \end{figure}
+
 We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario focusing on the variables measuring rivalry.
@@
 Thus, you get quite substantially different expected values from these two models.
-
-\begin{figure}
-	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
-	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
- 	\label{fig:gibmargeff}
- \end{figure}
 

### 2018-02-13 paper/5_concl.tex
more cleanup
app edits

@@
 International relations is generally about the interactions and dependencies among a set of countries or other important actors such as international governmental organizations (IGOs).
-This is particularly true of those scholars who work in the tradition of the Correlates of War Project, but is by no means limited to them.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
+This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
 
@@
 
-% These approaches have been developed for a while, but are not yet widely used in international relations scholarship.
-The latent network model captures first-order dependencies (the tendencies of certain actors to send and receive more of a certain action), second-order dependencies (the tendency of actors to reciprocate actions), and third-order dependencies (the ability for actions within a dyad to effect a third party) in dyadic data using a familiar regression framework that has been adapted for relational data---such as dyadic data---which are not independent nor identically distributed.
-
 To explore this approach in the context of international relations we have presented two broad analyses.
 The first is a simulation where the characteristics of the network are known.
-This shows that, when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
-The second is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
+This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
+The second analysis is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
 These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
 We then re-analyzed each study using the AME model.
-In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
-This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that cannot possibly be valid.
-In turn, this creates a certain arbitrariness in some research findings, which might lead to puzzles that are more apparent than real \citep{zinnes:1980}.
+In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, and c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
+This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that are not valid.
 
@@
 Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences, better out-of-sample predictive performance, and new substantive insights.
-
-% We also examine each of the models using the AME framework described above.
-Our goal is to ascertain whether accounting for the statistical complications that arise in dyadic data result in different model estimates when addressed in an AME framework, and more importantly we highlight the substantive opportunities presented by our alternative approach.
-
-% Based in part on our results as well as the statistical characteristics of the AME framework, we believe that it does.
-The dynamic latent factor model provides a step forward in the modeling of international relations.

### 2018-02-13 code/reiter_stam_2003/4 - posterior analysis.R

@@ -1,5 +1,6 @@
 devtools::install_github('s7minhas/amen') ; library(amen)
 resultsPath = '/Users/juanftellez/Dropbox/netsMatter/replications/Reiter_Stam_2003/output/'
+resultsPath = '~/s7m/Research/Dropbox/netsMatter/replications/Reiter_Stam_2003/output/'
 #
 library(magrittr)
 

### 2018-02-14 paper/4_app.tex
some tests

@@
 An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
-These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study.
+These differences may change dramatically when interdependencies are taken into account directly.
 This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates.
@@
 At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes.
+
 Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy.
@@
 
-The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.\footnote{For more details on this type of cross-validation strategy see \citet{minhas:etal:2016:arxiv}.} We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
+The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
+We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
 For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
 Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
+This is important as it indicates that switching to the AME framework---even when using the exact same specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
+The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, rather the dependence parameters we include are capturing underlying structure previously missed by the exogenous covariates in the models.
 
@@
 In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect while Democ-Pers directed dyad is still too imprecisely measured to indicate a direction.
-Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.
-Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies.
+Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.\footnote{Comparing } Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies.
 %Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
@@
 
-\citet{mcdonald:2004} studied whether trade promotes peace between nations.
+\citet{mcdonald:2004} studies whether trade promotes peace between nations.
 \citet[p.
@@
 560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
-The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad.
-\citet{mcdonald:2004} employs logistic regression to examine the  putative statistical significance of these variables.
-The models  include splines to correct for temporal dependence, and robust  standard errors clustered on each dyad.
+The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad, and \citet{mcdonald:2004} includes splines to correct for temporal dependence with robust standard errors clustered on each dyad.
 
@@
 The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
-While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
 Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error.
@@
 	\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
-	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}\label{fig:roser}
+	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}
+	\label{fig:roser}
 \end{figure}
@@
 To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
+One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
 She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
@@
 	\includegraphics[width=\textwidth]{weeks_circPlot.png}
-	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012).
+	\caption{ Visualization of multiplicative effects for Weeks (2012).
 Each circle designates a country and the color corresponds to the legend at the center of the visualization.
@@
 	}
+	\label{fig:weekscirc}
 \end{figure}
@@
 Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel.
-These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
+These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.
+Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
 In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries.
 In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category).
+They tend to initiative and be targets of conflict by similar actors but they are not more likely to initiate
+
+Thus she is kinda right but wrong in that they are more likely to intiate conflict in general ... but they do have a similar pattern of relationshsips
+
 In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type.
@@
 Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).
-In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model.
+In the case of the study with a continuous dependent variable, the AME approach has average error statistics that are about one-half that found in the original model.

### 2018-02-14 code/2_codeForPaperGraphics/weeks_posterior.R

@@ -72,11 +72,11 @@ load(paste0(resultsPath,'weeks_glmfit.rda'))
 # ggPerfCurves(predDfs, 'weeks')
 # ############################################
 
-# plot mult eff ###########################################
-# load AME model data
+# # mult eff analysis ###########################################
+# # load AME model data
 load(paste0(resultsPath, 'WeeksamenData.rda')) ## xDyadList, xNodeList.R, xNodeList.s, Ylist
 
-# subset data
+# # subset data
 yArr = listToArray(
 	actors=sort(unique(unlist(lapply(yList,rownames)))),
 	Y=yList[42:47], Xdyad=NULL, Xrow=NULL, Xcol=NULL)$Y
@@ -94,66 +94,152 @@ rownames(yArrSumm)=colnames(yArrSumm)=cntryKey$cowc
 uData = ameFit$U[cntryKey$code,] ; rownames(uData) = cntryKey$cowc
 vData = ameFit$V[cntryKey$code,] ; rownames(vData) = cntryKey$cowc
 
-# geo colors for nodes
-loadPkg('cshapes')
-cmap = wmap = cshp(date=as.Date('2016-1-1'))
-wmap@data$cowc = countrycode(wmap@data$COWCODE, 'cown', 'cowc')
-wmap@data$cowc[wmap@data$COWCODE==731]='PRK'
-wmap = wmap[which(as.character(wmap$cowc) %in% cntryKey$cowc),]
-coords=coordinates(wmap) ; rownames(coords)=wmap$cowc
-coords=coords[cntryKey$cowc,]
+# # geo colors for nodes
+# loadPkg('cshapes')
+# cmap = wmap = cshp(date=as.Date('2016-1-1'))
+# wmap@data$cowc = countrycode(wmap@data$COWCODE, 'cown', 'cowc')
+# wmap@data$cowc[wmap@data$COWCODE==731]='PRK'
+# wmap = wmap[which(as.character(wmap$cowc) %in% cntryKey$cowc),]
+# coords=coordinates(wmap) ; rownames(coords)=wmap$cowc
+# coords=coords[cntryKey$cowc,]
 
-# Create colors
-rlon = pi*coords[,1]/180 ; rlat = pi*coords[,2]/180
[…]

### 2018-02-14 paper/4_app.tex
rebuilding app section

@@
 
-We apply AME to five recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}.
+We apply AME to four recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  weeks:2012, gibler:2017}.
 Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The five studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The four studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.
@@
 		McDonald (2004) & Logit &1959--2002 & 198 & $92,354$ & Undirected & Robust\\
-		Rose (2004) & OLS & 1948--1999 & 177 & $234,597$ & Directed & Robust \\	 
+		% Rose (2004) & OLS & 1948--1999 & 177 & $234,597$ & Directed & Robust \\	 
 		Weeks (2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
@@
 
-Each of the five studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
+Each of the four studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
 In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
@@
 \centering
-\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
+\caption{Here we provide a brief summary of the key variable in each of the four replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
 	\begin{tabular}{l p{7cm} l} \toprule
-		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed \\
+		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
 		& &  accounting for dependencies? \\ \toprule
 		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Confirmed} \\ \midrule
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Unconfirmed}\\ \midrule
-		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Confirmed}\\ \midrule
+		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Confirmed}\\ \midrule
+		% Rose (2004) & WTO Membership Does not Affect Trade & {Partially Confirmed}\\ \midrule
 		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
-		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Unconfirmed}\\ \bottomrule
+		Gibler (2017) & Power Parity at Time of Entry to International System Increases Conflict & {Unconfirmed}\\ \bottomrule
 	\end{tabular}
@@
 				~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
-		\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
-				~ & RMDSE & 2.01 &  {1.06} \\	\midrule
+		% \multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
+		% 		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
 		\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
@@
 	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network-based approach.
-Four of the five studies involved a binary dependent variable; for those measures, area under the curve (AUC) statistics are reported.
-The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE).
+	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the four replications when using the standard dyadic approach and our network-based approach.
+Four of the four studies involved a binary dependent variable; for those measures, area under the curve (AUC) statistics are reported.\tabularnewline
+	%The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). 
 	}
@@
 \FloatBarrier
-
-Next we discuss each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.
-
-\subsection{Re-estimation of Reiter \& Stam (2003)}
-
-\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state.
-Using directed dyads, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.
-In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true.
-Independent variables focus on various encodings of regime types, contiguity, alliance, and capability measures.
-As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
-The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads.
-Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
-
-In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) has a positive effect while the variable ``Democ/Personalist Directed Dyad'' is too imprecisely measured to indicate a direction.
-In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect while Democ-Pers directed dyad is still too imprecisely measured to indicate a direction.
-Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.\footnote{Comparing } Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies.
-%Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
-
-\subsection{Re-estimation of McDonald (2004)}
-
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-\citet[p.
-547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''.
-Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measured free trade in two ways.
-The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad.
-reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
-This measure thus captures the score of the state in the dyad that possesses higher barriers to trade.
-\citet[p.
-560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
-The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad, and \citet{mcdonald:2004} includes splines to correct for temporal dependence with robust standard errors clustered on each dyad.
-
-Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.
-Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$.
-This indicates that conflict is less likely among members of a trade community.
-Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.
-The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$).
-At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
-
-\subsection{Re-estimation of Rose (2004)}
-
-\nocite{rose:2004}
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007}.
-Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
-He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion is that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
-
-When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar.
-The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
-However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
-The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
-Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error.
-\citep{tomz:etal:2007}.
-While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
-	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}
-	\label{fig:roser}
-\end{figure}
-\FloatBarrier
-
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.}  Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
-While the AME model replicates the main (null) result, there are substantial and substantive improvements gained from moving to an AME estimation.
-
-\subsection{Re-estimation of Weeks (2012)}
-
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
-She argues that in some circumstances autocrats are held accountable for their foreign policy decisions.
-She adds the nuance that autocratic audiences are not homogeneous.
-When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
-To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).
-One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.
-She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
-The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute.
-The basic finding is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes (and maybe democracies) and that b) machine-type regimes are no more belligerent than democracies.
-These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.
-She argued that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.
-
-\begin{figure}[!h]
-	\includegraphics[width=\textwidth]{weeks_circPlot.png}
-	\caption{ Visualization of multiplicative effects for Weeks (2012).
-Each circle designates a country and the color corresponds to the legend at the center of the visualization.
-Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries.
-The inner ring clusters countries by the similarity of who they receive conflict from.
-	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
-	}
-	\label{fig:weekscirc}
-\end{figure}
-\FloatBarrier
-
-The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model.
-In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone.
-When we examine the AME results, we find that none of these values are distinguishable from zero.
-Similarly, we find less pronounced effects for military capabilities.
-One explanation for this divergence is the AME model's ability to account for third-order effects.
-Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel.
-These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.
-Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
-In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries.
-In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category).
-They tend to initiative and be targets of conflict by similar actors but they are not more likely to initiate
-
-Thus she is kinda right but wrong in that they are more likely to intiate conflict in general ... but they do have a similar pattern of relationshsips
-
-In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type.
-Structural equivalence is present even when accounting for nodal characteristics like regime type.
-The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.
-
-\subsection{Re-estimation of Gibler (2017)}
-
-A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors.
-In addition, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
-
-We re-estimated model 6 from Table 6 (2017, 34).
-The results are presented in the Appendix.
-The results obtained with AME stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).
-Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.
-Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$).
-In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.
-As might be expected, the strong geographic clustering in the original study is about one-quarter as strong in the AME estimations.
-Similarly, rivalry coefficients are about one-third the size in the AME formulation, but a great deal more precisely measured ($z=18.116$).
-
-\begin{figure}
-	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
-	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
- 	\label{fig:gibmargeff}
- \end{figure}
-
-We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario focusing on the variables measuring rivalry.
-We employed mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.
-The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.
-As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notably.
-First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
-These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
-Thus, you get quite substantially different expected values from these two models.
-
-\subsection{Lessons Learned from Re-estimating Five Prominent Studies}
-
-First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.
-Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
-This means that significance testing (for better or worse) is compromised when network effects are ignored.
-
-Second, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights emerge from the additional information derived.
-In particular, there is actual information about the dependencies so that clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.
-This kind of information is absent in standard approaches and adds to our ability to explain specific as well as general results.
-
-Third, it is evident that the actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.
-This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.
-
-Fourth, it is clear that the AME approach dominates the GLM approaches in terms of performance.
-Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).
-In the case of the study with a continuous dependent variable, the AME approach has average error statistics that are about one-half that found in the original model.

### 2018-02-14 paper/5_concl.tex
rebuilding app section

@@
 This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
-The second analysis is a replication of five prominent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
-These five studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
+The second analysis is a replication of four recent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
+These four studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
 We then re-analyzed each study using the AME model.

### 2018-02-14 paper/6_appendix.tex
rebuilding app section

@@
 The GLM estimation with a probit link function is provided so as to ease comparison between the AME model, which is also based on the probit link.
-For Rose (2004) we just provide a table of coefficient estimates that includes the original OLS estimation and the AME model.
+%For Rose (2004) we just provide a table of coefficient estimates that includes the original OLS estimation and the AME model.
 
@@
 
-\subsubsection*{Rose (2004)}
+% \subsubsection*{Rose (2004)}
 
-Additional information for the Rose (2004) re-estimation.
+% Additional information for the Rose (2004) re-estimation.
 
-\input{rose_coeftable}
-\FloatBarrier
+% \input{rose_coeftable}
+% \FloatBarrier
 
-\begin{figure}
-	\centering   
-	\includegraphics[width=1\textwidth]{rose_outSample.pdf}
-	\caption{Assessments of out-of-sample predictive performance for Rose (2004) using root mean squared error (RMSE), root median squared error (RMDSE), and $R^{2}$.}
-\end{figure}
-\FloatBarrier
-\clearpage
+% \begin{figure}
+% 	\centering   
+% 	\includegraphics[width=1\textwidth]{rose_outSample.pdf}
+% 	\caption{Assessments of out-of-sample predictive performance for Rose (2004) using root mean squared error (RMSE), root median squared error (RMDSE), and $R^{2}$.}
+% \end{figure}
+% \FloatBarrier
+% \clearpage
 

### 2018-02-14 paper/app_deprc.tex
rebuilding app section

Next we discuss each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.

\subsection{Re-estimation of Reiter \& Stam (2003)}

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes. Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state. Using directed dyads, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.  In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true. Independent variables focus on various encodings of regime types, contiguity, alliance, and capability measures. As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.

In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) has a positive effect while the variable ``Democ/Personalist Directed Dyad'' is too imprecisely measured to indicate a direction. In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect while Democ-Pers directed dyad is still too imprecisely measured to indicate a direction. Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one.\footnote{Comparing } Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies, though many of the major effects  uncovered by the authors remains robust when accounting for dyadic interdependencies. %Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.

\subsection{Re-estimation of McDonald (2004)}

\citet{mcdonald:2004} studies whether trade promotes peace between nations. \citet[p. 547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''. Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states. His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences. \citet{mcdonald:2004} measured free trade in two ways. The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad; this reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war. This measure thus captures the score of the state in the dyad that possesses higher barriers to trade. \citet[p. 560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''. The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad, and \citet{mcdonald:2004} includes splines to correct for temporal dependence with robust standard errors clustered on each dyad.

\subsection{Re-estimation of Rose (2004)}

\nocite{rose:2004}
In 2004, Andrew Rose published a study in the \textit{American Economic Review} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations. It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007}.  Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims. He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades. Estimating this model using OLS within many differing contexts, his conclusion is that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated. The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade. Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error. \citep{tomz:etal:2007}. While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.

\begin{figure}
	\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}
	\label{fig:roser}
\end{figure}
\FloatBarrier

The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.}  Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the AME model replicates the main (null) result, there are substantial and substantive improvements gained from moving to an AME estimation.

\subsection{Re-estimation of Weeks (2012)}

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  One major innovation in her study resides in the nuanced way in which she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. The basic finding is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes (and maybe democracies) and that b) machine-type regimes are no more belligerent than democracies.   These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper. She argued that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{weeks_circPlot.png}
	\caption{ Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from. 
	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
	}
	\label{fig:weekscirc}
\end{figure}
\FloatBarrier

The re-estimation of \citet{weeks:2012} has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone. When we examine the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third-order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see the US, the UK, Germany, Canada, and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets. Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China. In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries. In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category). They tend to initiative and be targets of conflict by similar actors but they are not more likely to initiate

Thus she is kinda right but wrong in that they are more likely to intiate conflict in general ... but they do have a similar pattern of relationshsips

In the GLM, which ignores these third-order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.

\subsection{Re-estimation of Gibler (2017)}

A more recent example is \citet{gibler:2017} which examines the onset of militarized disputes using capabilities, joint democracy, alliances, and power parity in a undirected dyadic study using logistic regression and dyad clustered standard errors. In addition, Gibler shows that the long-standing relationship between the relative parity of capabilities and initiation of international conflict is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members. This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.

We re-estimated model 6 from Table 6 (2017, 34). The results are presented in the Appendix. The results obtained with AME stand in stark contrast to those found with a logistic regression (with dyad clustered, robust standard errors).  Most importantly, the primary variable from the Gibler study, parity of the members of a dyad at the year in which they entered the international system, is shown to be unimportant in the AME results.  Not only is the value of this parameter small, but it has a very large relative standard error---over a magnitude larger than the parameter itself ($z= 0.038$). In addition, the variable indicating whether both members of the dyad were coded as democracies (joint democracy) follows the same pattern: important and strong in the logistic results, but this disappears once interdependencies are modeled.  As might be expected, the strong geographic clustering in the original study is about one-quarter as strong in the AME estimations. Similarly, rivalry coefficients are about one-third the size in the AME formulation, but a great deal more precisely measured ($z=18.116$).

\begin{figure}
	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
	\includegraphics[width=\textwidth]{gibler_margeff.pdf}
 	\label{fig:gibmargeff}
 \end{figure}

We utilized the original and the AME results from Gibler's model 6 in Table 6 to calculate the expected values for one scenario focusing on the variables measuring rivalry.  We employed mean or modal values for all independent variables, except we changed the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.  The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: Gibler's GLM estimation and our AME approach.  As this Figure~\ref{fig:gibmargeff} illustrates, the AME results differ notably. First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.  These are rare events, so the probabilities are low, but the difference is a factor of almost 2. Thus, you get quite substantially different expected values from these two models.

\subsection{Lessons Learned from Re-estimating Five Prominent Studies}

First and foremost, many findings that emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME. Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.  This means that significance testing (for better or worse) is compromised when network effects are ignored.

Second, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights emerge from the additional information derived. In particular, there is actual information about the dependencies so that clusters can be identified, and the extent of reciprocity at the dyad level, as well as among senders and receivers.  This kind of information is absent in standard approaches and adds to our ability to explain specific as well as general results.

Third, it is evident that the actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.  This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.

Fourth, it is clear that the AME approach dominates the GLM approaches in terms of performance. Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics). In the case of the study with a continuous dependent variable, the AME approach has average error statistics that are about one-half that found in the original model.

### 2018-02-14 code/2_codeForPaperGraphics/reiter_stam_posterior.R

@@ -4,20 +4,18 @@ rm(list=ls())
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/output/'
 	inputPath= '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/input/'
-	plotPath = '~/Research/netsMatter/paper/' }
+	plotPath = '~/Research/netsMatter/paper/'
+  source('~/Research/netsMatter/code/helpers/functions.R')
+  source('~/Research/netsMatter/code/helpers/ameHelpers.R')
+  source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
+  source('~/Research/netsMatter/code/helpers/stargazerHelpers.R') }
 
 if(Sys.info()['user'] == 'juanftellez'){
   resultsPath = '/Users/juanftellez/Dropbox/netsMatter/replications/Reiter_Stam_2003/output/'
   plotPath = '~/OneDrive/netsMatter/paper/'
   source('~/OneDrive/netsMatter/code/helpers/ameHelpers.R')
   source('helperEx.R')
-  #
 }
-
-source('~/Research/netsMatter/code/helpers/functions.R')
-source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
-source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
 ############################################
 
 # modData ###########################################
@@ -25,34 +23,34 @@ load( paste0(resultsPath,'reiter_stam_glmfit.rda') )
 load( paste0(resultsPath,'model_k2_v12.rda') )
 ############################################
 
-# coefSumm ###########################################
-ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
[…]

### 2018-02-14 code/2_codeForPaperGraphics/weeks_posterior.R

@@ -16,64 +16,64 @@ load( paste0(resultsPath,'model_k2.rda') )
 load(paste0(resultsPath,'weeks_glmfit.rda'))
 ############################################
 
-# # coefSumm ###########################################
-# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-# rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-# glmSumm = modSumm[,]
-# probSumm = proSumm[,]
+# coefSumm ###########################################
+ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+glmSumm = modSumm[,]
+probSumm = proSumm[,]
 
-# modList = list(glmSumm, probSumm, ameSumm)
-# modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+modList = list(glmSumm, probSumm, ameSumm)
+modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
 
-# varKey = data.frame(
-# 	dirty=names(coef(mod)),
-# 	clean=c( '(Intercept)', 
[…]

### 2018-02-14 paper/1_intro.tex
spacetest
rebuilding app section
intro edits

@@
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters, to reflect the potential clustering of cases.
-In practice, these palliatives rarely work because they fail to address the fundamental data generating process that remains a threat to inference.
-Namely, it is not just the diagonals of the variance-covariance matrix that are affected \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
+Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
+In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
+This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix.
+\citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
 
@@
 Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads--for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
-The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.
+The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.
 
@@
 Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data.
-Last, to highlight the utility of this approach, we apply the AME model to five recent studies in the international relations (IR) literature.
+Last, to highlight the utility of this approach, we apply the AME model to four recent studies in the international relations (IR) literature.
 The comparison reveals that in accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature.
-Consequently, AME produces results that, at times, differ from those found in these studies in particular as well as the broader literature from which they are drawn.
+Consequently, AME produces results that, at times, differ from those found in the original study as well as from the broader literature.
 Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies.
@@
 
-The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while enabling scholars to still focus on testing the substantive effect of variables of interest.
+The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
 Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while having the benefit of being able to better account for the data generating process behind political events of interest.
-Most importantly, it concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.
+Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-02-16 paper/1_intro.tex
intro and model

@@
 The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
-Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while having the benefit of being able to better account for the data generating process behind political events of interest.
+Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while being able to account for the data generating process behind political events of interest.
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-02-16 paper/2_model.tex
intro and model

@@
 Relational data comes with an explicit structure that generally leads to particular types of dependencies.
-The importance of accounting for the underlying structure of our data has been a lesson well understood, at least, when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
+The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
 As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
@@
 In most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook), we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}.
-For example, in an analysis of international trade certain countries (e.g., China) export in much larger volumes than other countries for a variety of structural, contextual, and idiosyncratic reasons.
+For example, in an analysis of international trade certain countries (e.g., China) export much larger volumes than other countries for a variety of structural, contextual, and idiosyncratic reasons.
 Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.\footnote{In an undirected setting instead of studying sender and receiver heterogeneity, we would just be concerned with actor heterogeneity in general.}
@@
 	\begin{tabular}{cc}
-		\textsc{Sender heterogeneity} & \textsc{Receiver Heterogeneity} \\
+		\textsc{Sender Heterogeneity} & \textsc{Receiver Heterogeneity} \\
 		\includegraphics[width=.4\textwidth]{adjRowDep.png} & \includegraphics[width=.4\textwidth]{adjColDep.png} \\
@@
 
-For similar reasons one also needs to take into account the shared dependence between observations that share a common receiver.
+For similar reasons one also needs to take into account the dependence between observations that share a common receiver.
 The bottom-left panel in Figure~\ref{fig:adjMatDeps} illustrates that sender and receiver type dependencies can also blend together.
@@
 More often we find significant evidence of higher-order structures that result from dependencies between multiple groups of actors.
-These dependencies arise because there may be some set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
+These dependencies arise because there may be a set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}.
 In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place.
@@
 	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
-	\caption{Visualization of a network with higher-order dependence patterns.}
+	\caption{Higher-order dependence patterns in a network.}
 	% need a simpler caption here
@@
 A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualize linkages between right and left leaning political blogs preceding the 2004 United States Presidential Election.
-\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs was minimal, and that most blogs were linked to others that were politically similar.
+\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs is minimal and that most blogs are linked to others that are politically similar.
 This showcases the types of higher-order dependencies that can emerge in relational data.
-First, the fact that interactions were determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}.
+First, the fact that interactions are determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}.
 Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations.
@@
 Stochastic equivalence refers to a type of pattern in which actors can be divided into groups such that members of the same group have similar patterns of relationships.
-In the example we have laid out above, each of the left leaning blogs would be considered stochastically equivalent to one another because any given left-leaning blog is more likely to interact with a blog of a similar political position and less likely to interact with one with a divergent political position.
+In the example above, each of the left leaning blogs would be considered stochastically equivalent to one another because any given left-leaning blog is more likely to interact with a blog of a similar political position and less likely to interact with one with a divergent political position.
 
@@
 Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance --- an example of stochastic equivalence.
-In both these examples, we are able to explicitly parameterize the attribute that might explain the emergence of higher order dependence patterns, but it is likely that at times we will not able to account for the variety of reasons why higher order dependence patterns in networks may develop.
+In both these examples, we are able to explicitly parameterize the attribute that might explain the emergence of higher order dependence patterns.
+While sometimes the conditions driving these patterns, such as geography, are easy to identify, at other times it can be difficult to describe exactly why higher order dependence patterns in networks may develop.
+%CD to SM ^
 
@@
 
-where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver).
-We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
-Next, to account for the dependencies that emerge in dyadic data and that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM.
+where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver) at time $t$.
+We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
+Next, to account for the dependencies that emerge in dyadic data that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM.
 $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:
@@
 The sender and receiver random effects are modeled jointly from a multivariate normal distribution to account for correlation in how active an actor is in sending and receiving ties.
-Heterogeneity in the the sender and receiver effects is captured by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively, and $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties).
+Heterogeneity in the sender and receiver effects is captured by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively, and $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties).
 Beyond these first-order dependencies, second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a within dyad correlation, or reciprocity, parameter $\rho$.
@@
 $K$ denotes the dimensions of the latent space.
-The construction of the LFM here is actually quite similar to the recommender systems that companies like Amazon and Netflix have used to model customer behavior \citep{resnick:varian:1997,bennett:etal:2007}.
+The construction of the LFM here is actually quite similar to the recommender systems that companies like Amazon and Netflix use to model customer behavior \citep{resnick:varian:1997,bennett:etal:2007}.
 This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
@@
 $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model, the singular values are not restricted to be positive, thus allowing us to account for both positive and negative homophily.}
+The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model the singular values are not restricted to be positive, thus allowing us to account for both positive and negative homophily.}
 
@@
 However, it is rarely the case that this is possible even for TSCS data and thus modeling decisions must account for underlying structure.
-Failing to account for the underlying structure in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
-Additionally, by ignoring these potential interdependencies, we often ignore substantively interesting features of the problem.
+Failing to do so in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.
+Additionally, by ignoring these potential interdependencies, we often ignore substantively interesting features of the phenomena we investigate.
 The study of international relations is founded on the relations among actors.

### 2018-02-19 paper/1_intro.tex
intro edit

@@
 Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data.
-Last, to highlight the utility of this approach, we apply the AME model to four recent studies in the international relations (IR) literature.
+Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature.
 The comparison reveals that in accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature.
@@
 The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
-Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while being able to account for the data generating process behind political events of interest.
+Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while accounting for the data generating process behind political events of interest.
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-02-19 paper/2_model.tex
model edit

@@
 Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flow, that is sent from actor $i$ to actor $j$.
-In many applications, scholars are interested in studying undirected (i.e., symmetric) outcomes in which there is no clear sender or receiver, these type of outcomes still can, and we would argue should, be studied using the type of framework we discuss below.
+In many applications, scholars are interested in studying undirected (i.e., symmetric) outcomes in which there is no clear sender or receiver, these type of outcomes still can, and we argue should, be studied using the type of framework we discuss below.
 
@@
 
-The visualization illustrates that the actors belonging to the same group have a higher likelihood of interacting with each other, whereas interactions across groups are rarer.
+The visualization illustrates that actors belonging to the same group have a higher likelihood of interacting with each other, whereas interactions across groups are rarer.
 A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualize linkages between right and left leaning political blogs preceding the 2004 United States Presidential Election.
@@
 The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages.
-It is flexible and easy to use for analyzing the kind of relational data often found in social science.
+It is flexible and easy to use for analyzing the kind of relational data often found in the social sciences.
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.

### 2018-02-19 paper/3_sim.tex
sim  edit

@@
 
-For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.\footnote{The value of $\gamma$ is also set to 1, which corresponds to a an example where the $W$ character is associated with homophily.} We conduct two sets of simulations, one in which the number of actors in the network is set to 50 and the other at 100.
+For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.\footnote{The value of $\gamma$ is also set to 1, which corresponds to an example where the $W$ character is associated with homophily.} We conduct two sets of simulations, one in which the number of actors in the network is set to 50 and the other at 100.
 In total, we ran 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimated a standard, AME, and oracle model.
@@
 This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$.
-To estimate how well the model performs in doing this we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors is set to 100 (top pair of panels) and 50.
+To estimate how well the model performs, we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors is set to 100 (top pair of panels) and 50.
 Additionally, we calculate the median across the correlations and display the result using a vertical line.
-For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the structure provided by this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.
+For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.
 
@@
 
-What this simulation has shown is that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is that we can also estimate unobserved dependencies through the random effects structure of the model.
-Scholars can use this framework in an iterative fashion where once they have estimated a model, they can empirically study the structure of the random effects portion of the model to assess whether there are unobserved covariates that they want to account for in their models.
+What this simulation has shown is that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is to directly estimate unobserved dependencies through the random effects structure of the model.
+Scholars can use this framework in an iterative fashion: beginning with an estimated a model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their models.
 At the very least, what this simulation section shows is that the AME model can help in estimating the true effect and conducting inference on the independent variables that scholars have operationalized to test their theoretical propositions.

### 2018-02-20 paper/4_app.tex
ab for rose
sync

@@
 
-We apply AME to four recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  weeks:2012, gibler:2017}.
+We apply AME to three recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  weeks:2012, gibler:2017}.
 Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
@@
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The four studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.
 However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.\footnote{It is important to note that the AME framework is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample.
-The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.}
+Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.\footnote{It is important to note that the AME framework is not well suited for studies where the focus may be on studying onsets and dyad-years representing conflict continuations are removed from the sample.
+In studies with this type of focus, network models are of limited utility since any interdependencies underlying how events may spread from one actor to another are explicitly masked from the estimation process.}
 
@@
 \caption{Features of the Studies Re-estimated.}
-	\begin{tabular}{lcccccc}
-		& Model &  Date Range & N. Actors  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
-		Reiter \& Stam (2003) &Logit &1945--1995 &  193 & $753,456$ & Directed & Robust \\	
-		McDonald (2004) & Logit &1959--2002 & 198 & $92,354$ & Undirected & Robust\\
-		% Rose (2004) & OLS & 1948--1999 & 177 & $234,597$ & Directed & Robust \\	 
-		Weeks (2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
-		Gibler (2017) & Logit & 1816--2008 &193 &   $650,557$ & Undirected & None \\ \bottomrule
+	\begin{tabular}{lccccc}
+		& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
+		Reiter \& Stam (2003) &Logit &1945--1995 &  193 & Directed & Robust \\	
+		McDonald (2004) & Logit &1959--2002 & 198 & Undirected & Robust\\
+		Weeks (2012) & Logit & 1946--1999 &197 & Directed & Robust \\
+		Gibler (2017) & Logit & 1816--2008 &193 & Undirected & None \\ \bottomrule
 	\end{tabular}
@@
 
-Each of the four studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
+Each of the studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
 In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
@@
 \centering
-\caption{Here we provide a brief summary of the key variable in each of the four replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
+\caption{Here we provide a brief summary of the key variable in each of the replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
 	\begin{tabular}{l p{7cm} l} \toprule
@@
 		& &  accounting for dependencies? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Confirmed} \\ \midrule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Confirmed} \\ \midrule
 		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Confirmed}\\ \midrule
-		% Rose (2004) & WTO Membership Does not Affect Trade & {Partially Confirmed}\\ \midrule
 		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
@@
 We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
-For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
-Only one of the studies here had a continuous dependent variable and for this we provide the root mean squared error (RMSE) and root median squared error (RMDSE).\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
+For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
 This is important as it indicates that switching to the AME framework---even when using the exact same specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
-The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, rather the dependence parameters we include are capturing underlying structure previously missed by the exogenous covariates in the models.
+The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, rather the additive and multiplicative effects we include are capturing underlying structure previously missed by the exogenous covariates in the models.
 
@@
 				~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
-		\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
-				~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
-		% \multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
-		% 		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
 		\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
@@
 	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the four replications when using the standard dyadic approach and our network-based approach.
-Four of the four studies involved a binary dependent variable; for those measures, area under the curve (AUC) statistics are reported.\tabularnewline
+	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the replications when using the standard dyadic approach and our network-based approach.
+Each of the studies involved a binary dependent variable and area under the curve (AUC) statistics are reported.\tabularnewline
 	%The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). 
@@
 \FloatBarrier
+
+\subsection{....}
+
+Apart from providing stronger inferential guarantees and predictive performance than conventional approaches, the AME framework
+
+reiter and stam take a look at the srm effeccts to understand where the model is still falling short
+
+mcdonald we recover support fro trade dependenicence and also find eveidence for their highest barrier to trade variable
+
+for weeks even though we dont find support for her hypothesis we do se esomething interesting in the multiplicative effects
+
+for gibler

### 2018-02-20 paper/6_appendix.tex
sync

@@
 
-\subsubsection*{McDonald (2004)}
+% \subsubsection*{McDonald (2004)}
 
-Additional information for the McDonald (2004) re-estimation.
+% Additional information for the McDonald (2004) re-estimation.
 
-\input{mcdonald_coeftable}
-\FloatBarrier
+% \input{mcdonald_coeftable}
+% \FloatBarrier
 
-\begin{figure}
-	\centering   
-	\subfigure[AUC]{\label{fig:mcdonaldroc}\includegraphics[width=.45\textwidth]{mcdonald_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:mcdonaldpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves, PR curves, and separation plots.}
-\end{figure}
-\FloatBarrier
-\clearpage
+% \begin{figure}
+% 	\centering   
+% 	\subfigure[AUC]{\label{fig:mcdonaldroc}\includegraphics[width=.45\textwidth]{mcdonald_roc_outSample.pdf}}
+% 	\subfigure[Precision and Recall]{\label{fig:mcdonaldpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
+% 	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves, PR curves, and separation plots.}
+% \end{figure}
+% \FloatBarrier
+% \clearpage
 

### 2018-02-20 code/2_codeForPaperGraphics/mcdonald_posterior.R

@@ -17,50 +17,54 @@ load(paste0(resultsPath, 'mcdonald_glmfit.rda'))
 load( paste0(resultsPath,'ameFitIntercepts_k2.rda') )
 ############################################
 
-# coefSumm ###########################################
-ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-glmSumm = modSumm[,]
-probSumm = proSumm[,]
+# # coefSumm ###########################################
+# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+# rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+# glmSumm = modSumm[,]
+# probSumm = proSumm[,]
 
-modList = list(glmSumm, probSumm, ameSumm)
-modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+# modList = list(glmSumm, probSumm, ameSumm)
+# modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
 
[…]

### 2018-02-20 code/2_codeForPaperGraphics/reiter_stam_posterior.R

@@ -55,29 +55,65 @@ load( paste0(resultsPath,'model_k2_v12.rda') )
 # addEffdata ###########################################
 # sender effects
 effdat = getAddEffData(fit = ameFit) ##This function is in helperEx.R
-effdat = effdat[which(
-  effdat$addEff>=quantile(effdat$addEff,.9) |
-  effdat$addEff<=quantile(effdat$addEff,.1)
-  ),]
+# effdat = effdat[which(
+#   effdat$addEff>=quantile(effdat$addEff,.9) |
+#   effdat$addEff<=quantile(effdat$addEff,.1)
+#   ),]
+effdat$cown = effdat$actor
 effdat$actor = countrycode::countrycode(effdat$actor, 'cown', 'country.name')
-effdat = na.omit(effdat)
+# effdat = na.omit(effdat)
 effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor'])
 
-## subset of countries
-addEffPlot(fit = effdat, addEffData = effdat, row = T) + coord_flip()
+# ## subset of countries
+# addEffPlot(fit = effdat, addEffData = effdat, row = T) +
+#   coord_flip() + 
+#   theme(
+#     axis.text.x=element_text(angle=0)
+#     )
+
+## spatial visualization
+# fix actor names
+cntryKey = data.frame(code=effdat$cown, cname=char(effdat$actor),stringsAsFactors = FALSE)
+cntryKey$cname[cntryKey$code==731]='North Korea'
+cntryKey$cowc = countrycode(cntryKey$cname, 'country.name', 'cowc')
+
+# geo colors for nodes
+loadPkg('cshapes')
+cmap = cshp(date=as.Date('2016-1-1'))
+cmapDF=fortify(cmap,region='FEATUREID') ; names(cmapDF)[6]='FEATUREID' ; cmapDF=join(cmapDF, cmap@data)
+cmapDF$addEff = effdat$addEff[match(cmapDF$COWCODE, effdat$cown)]
+
+ggMap = ggplot() +
+  geom_polygon(data=cmapDF, aes(x=long, y=lat,group=group,fill=addEff),color='grey30',size=.05) +
+  scale_fill_gradient2() +
+  coord_equal() + xlab('') + ylab('') +
+  labs(
+    x='', y='', fill='Sender Effects'
+    ) +
+  theme(
+    legend.position = 'top',
+    panel.border = element_blank(), panel.grid=element_blank(),
[…]

### 2018-02-21 paper/4_app.tex
moving stuff around

@@
 We apply AME to three recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  weeks:2012, gibler:2017}.
-Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
+Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
 We choose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
-First, the results of AME estimation are interpretable alongside results using standard approaches, but, as shown in the simulation section, have the additional benefit of being able to take into account dependencies that may complicate inference.
-Second, through using this approach, we can also quantify the degree to which first, second, and third order dependencies are present in events of interest.
-Third, we show that by using the AME framework scholars can better model the data generating process behind their events of interest.
+First, we show that simply by using the AME framework scholars can better model the data generating process behind their events of interest.
+Second, the results of AME estimation are interpretable alongside results using standard approaches, but, as shown in the simulation section, have the additional benefit of being able to take into account dependencies that may otherwise complicate inference.
+Third, through using this approach, we can also quantify the degree to which first, second, and third order dependencies are present in events of interest.
 
-We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
-Of course, by the criteria of influence many other papers could have been considered for the replication.
-However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.\footnote{It is important to note that the AME framework is not well suited for studies where the focus may be on studying onsets and dyad-years representing conflict continuations are removed from the sample.
-In studies with this type of focus, network models are of limited utility since any interdependencies underlying how events may spread from one actor to another are explicitly masked from the estimation process.}
+We obtained the data for these studies from their replication archives and replicated their main results.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies below were selected based on how recently they were published and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
+Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a general linearized model.\footnote{It is important to note that the AME framework is not well suited for studies where the focus may be on studying onsets and dyad-years representing conflict continuations are removed from the sample.
+In studies with this type of focus, network models in general are of limited utility since any interdependencies underlying how events may spread from one actor to another are explicitly masked from the estimation process.} Table~\ref{tab:modelDesign} provides descriptive information for the studies that we replicated.
 
 \begin{table}
-\caption{Features of the Studies Re-estimated.}
+\caption{Descriptive information about the replicated studies. }
 	\begin{tabular}{lccccc}
@@
 		Reiter \& Stam (2003) &Logit &1945--1995 &  193 & Directed & Robust \\	
-		McDonald (2004) & Logit &1959--2002 & 198 & Undirected & Robust\\
 		Weeks (2012) & Logit & 1946--1999 &197 & Directed & Robust \\
@@
 	\end{tabular}
+	\label{tab:modelDesign}
 \end{table}
-
-Each of the studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
-
-\begin{table}[ht]
-\centering
-\caption{Here we provide a brief summary of the key variable in each of the replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
-	\begin{tabular}{l p{7cm} l} \toprule
-		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
-		& &  accounting for dependencies? \\ \toprule
-		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Confirmed} \\ \midrule
-		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Confirmed}\\ \midrule
-		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
-		Gibler (2017) & Power Parity at Time of Entry to International System Increases Conflict & {Unconfirmed}\\ \bottomrule
-	\end{tabular}
-	\label{tab:modelFindingSumm}
-\end{table}
-
-An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
-These differences may change dramatically when interdependencies are taken into account directly.
-This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates.
-As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates.
-At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes.
 
@@
 We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
-For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.\footnote{More details on the performance of each of these models can be found in the Appendix.} For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
+For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
+For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
 This is important as it indicates that switching to the AME framework---even when using the exact same specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
@@
 
+\begin{figure}
+	\centering   
+	\begin{tabular}{cc}
+		\multicolumn{2}{l}{\textbf{Reiter \& Stam (2003)}} \\
+		\includegraphics[width=.4\textwidth]{reiter_stam_roc_outSample.pdf} & 
+		\includegraphics[width=.4\textwidth]{reiter_stam_pr_outSample.pdf} \\
+		\multicolumn{2}{l}{\textbf{Weeks (2012)}} \\
+		\includegraphics[width=.4\textwidth]{weeks_roc_outSample.pdf} & 
+		\includegraphics[width=.4\textwidth]{weeks_pr_outSample.pdf} \\
+		\multicolumn{2}{l}{\textbf{Gibler (2017)}} \\
+		\includegraphics[width=.4\textwidth]{gibler_roc_outSample.pdf} & 
+		\includegraphics[width=.4\textwidth]{gibler_pr_outSample.pdf} \\
+	\end{tabular}
+	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
+	\label{fig:perf}	
+\end{figure}
+
+Each of the studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
+
 \begin{table}[ht]
 \centering
-	\begin{tabular}{l|l cc}
-	~ & ~ & GLM & AME \\
-	\toprule
-		\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
-				~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
-		\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
-				~ & AUC-PR & 0.00 &  {0.15} \\		\midrule
-		\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
-				~ & AUC-PR & 0.00 &  {0.08} \\			\bottomrule
+\caption{Here we provide a brief summary of the key variable in each of the replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
+	\begin{tabular}{l p{7cm} l} \toprule
+		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
+		& &  accounting for dependencies? \\ \toprule
+		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Confirmed} \\ \midrule
+		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
+		Gibler (2017) & Power Parity at Time of Entry to International System Increases Conflict & {Unconfirmed}\\ \bottomrule
 	\end{tabular}
-	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the replications when using the standard dyadic approach and our network-based approach.
-Each of the studies involved a binary dependent variable and area under the curve (AUC) statistics are reported.\tabularnewline
-	%The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). 
-	}
-	\label{tab:modelPerfSumm}
+	\label{tab:modelFindingSumm}
 \end{table}
-\FloatBarrier
+
+An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
+These differences may change dramatically when interdependencies are taken into account directly.
+This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates.
+As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates.
+At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes.
 

### 2018-02-21 paper/6_appendix.tex
moving stuff around

--- 2018-02-14 paper/6_appendix.tex
@@
 
-\begin{figure}
-	\centering   
-	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves, PR curves, and separation plots.}
-\end{figure}
+% \begin{figure}
+% 	\centering   
+% 	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
+% 	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
+% 	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves, PR curves, and separation plots.}
+% \end{figure}
 \FloatBarrier
 \clearpage
-
-\subsubsection*{McDonald (2004)}
-
-Additional information for the McDonald (2004) re-estimation.
-
-\input{mcdonald_coeftable}
-\FloatBarrier
-
-\begin{figure}
-	\centering   
-	\subfigure[AUC]{\label{fig:mcdonaldroc}\includegraphics[width=.45\textwidth]{mcdonald_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:mcdonaldpr}\includegraphics[width=.45\textwidth]{mcdonald_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for McDonald (2004) using ROC curves, PR curves, and separation plots.}
-\end{figure}
-\FloatBarrier
-\clearpage
-
-% \subsubsection*{Rose (2004)}
-
-% Additional information for the Rose (2004) re-estimation.
-
-% \input{rose_coeftable}
-% \FloatBarrier
-
-% \begin{figure}
-% 	\centering   
-% 	\includegraphics[width=1\textwidth]{rose_outSample.pdf}
-% 	\caption{Assessments of out-of-sample predictive performance for Rose (2004) using root mean squared error (RMSE), root median squared error (RMDSE), and $R^{2}$.}
-% \end{figure}
-% \FloatBarrier
-% \clearpage
 
@@
 
-% already show roc-pr in main text
-\begin{figure}
-\centering   
-  \subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
-  \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
-  \caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
-\end{figure}
+% \begin{figure}
+% \centering   
+%   \subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
+%   \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
+%   \caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
+% \end{figure}
 \FloatBarrier
@@
 
-\begin{figure}
-	\centering   
-	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
-	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
-	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
-\end{figure}
+% \begin{figure}
+% 	\centering   
+% 	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
+% 	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
+% 	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
+% \end{figure}
 \FloatBarrier

### 2018-02-21 code/2_codeForPaperGraphics/gibler_posterior.R

@@ -17,38 +17,38 @@ load(paste0(resultsPath,'mdwameFit_k2_v4.rda'))
 load( paste0(resultsPath,'glmFit.rda') )
 ############################################
 
-# coefSumm ###########################################
-ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+# # coefSumm ###########################################
+# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+# rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
 
-glmSumm = modSumm[,]
+# glmSumm = modSumm[,]
 
-dat = cbind(y=mod$y, mod$data[,names(coef(mod)[-1])], dyad=mod$data$dyad)
-form = paste0('y~',paste(names(coef(mod)[-1]), collapse='+'))
-probMod = glm( form, data=dat, family=binomial(link='probit') )
-loadPkg('lmtest')
-probSumm = coeftest(probMod, vcov = vcovCluster(mod, cluster = dat$dyad))[,]
[…]

### 2018-02-21 code/2_codeForPaperGraphics/mcdonald_posterior.R

@@ -1,82 +1,0 @@
-# setup ###########################################
-rm(list=ls())
-
-if(Sys.info()['user'] %in% c('s7m', 'janus829')){
-	resultsPath = '~/Dropbox/Research/netsMatter/replications/McDonald_2004/data/'
-	inputPath= '~/Dropbox/Research/netsMatter/replications/McDonald_2004/data/'
-	plotPath = '~/Research/netsMatter/paper/' }
-
-source('~/Research/netsMatter/code/helpers/functions.R')
-source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
-source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
-############################################
-
-# modData ###########################################
-load(paste0(resultsPath, 'mcdonald_glmfit.rda'))
-load( paste0(resultsPath,'ameFitIntercepts_k2.rda') )
-############################################
-
-# # coefSumm ###########################################
-# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-# rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-# glmSumm = modSumm[,]
-# probSumm = proSumm[,]
-
-# modList = list(glmSumm, probSumm, ameSumm)
-# modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
-
-# varKey = data.frame(
-# 	dirty=names(coef(mod)),
-# 	clean=c( '(Intercept)', 
-# 		"Spline0","Spline1","Spline2","Spline3",
-# 		"Shared Alliance","Contiguous","Log Capabilities Ratio",
-# 		"Trade Dependence","Preconflict GDP Change","Lowest Dyadic Polity Score",
[…]

### 2018-02-21 code/2_codeForPaperGraphics/rose_posterior.R

@@ -1,126 +1,0 @@
-# setup ###########################################
-rm(list=ls())
-
-if(Sys.info()['user'] %in% c('s7m', 'janus829')){
-	resultsPath = '~/Dropbox/Research/netsMatter/replications/rose2004/'
-	plotPath = '~/Research/netsMatter/paper/' }
-
-source('~/Research/netsMatter/code/helpers/functions.R')
-source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
-############################################
-
-# modData ###########################################
-load(paste0(resultsPath,'outputData/ameFit_k2_v2_imps_50000_intercept.rda'))
-load( paste0(resultsPath,'glmResults_rose.rda') )
-############################################
-
-# # coefSumm ###########################################
-# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-# lmSumm = baseModelSumm[,]
-
-# modList = list(lmSumm, ameSumm)
-# modNames = c('LM', 'AME')
-
-# varKey = data.frame(
-# 	dirty=rownames(ameSumm),
-# 	clean=c('Intercept', 'Both in GATT/WTO', 'One in GATT/WTO', 'GSP',
-# 		'Log Distance',  'Log Product Real GDP', 'Log Product Real GDPpc',
-# 		'Regional FTA', 'Currency Union', 'Common language', 'Land Border',
-# 		'Number Landlocked' ,'Number Islands', 'Log Product Land Area',
-# 		'Common Colonizer','Currently Colonized', 'Ever Colony','Common Country' ),
-# 	stringsAsFactors = FALSE )
-
-# #
-# getCoefTable(varKey, modList, modNames, 'rose', 'Rose (2004)', plotPath, 3)
-# ############################################
-
[…]

### 2018-02-21 code/2_codeForPaperGraphics/weeks_posterior.R

@@ -16,46 +16,46 @@ load( paste0(resultsPath,'model_k2.rda') )
 load(paste0(resultsPath,'weeks_glmfit.rda'))
 ############################################
 
-# coefSumm ###########################################
-ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
-	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
-	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
-rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
-rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
-glmSumm = modSumm[,]
-probSumm = proSumm[,]
+# # coefSumm ###########################################
+# ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+# 	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+# 	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+# rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+# rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+# glmSumm = modSumm[,]
+# probSumm = proSumm[,]
 
-modList = list(glmSumm, probSumm, ameSumm)
-modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
+# modList = list(glmSumm, probSumm, ameSumm)
+# modNames = c('GLM (Logit)','GLM (Probit)', 'AME')
 
-varKey = data.frame(
-	dirty=names(coef(mod)),
[…]

### 2018-02-22 paper/4_app.tex
reorg by mdw, tiny changes

@@
 		\multicolumn{2}{l}{\textbf{Reiter \& Stam (2003)}} \\
-		\includegraphics[width=.4\textwidth]{reiter_stam_roc_outSample.pdf} & 
-		\includegraphics[width=.4\textwidth]{reiter_stam_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_roc_outSample.pdf} & 
+		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{Weeks (2012)}} \\
-		\includegraphics[width=.4\textwidth]{weeks_roc_outSample.pdf} & 
-		\includegraphics[width=.4\textwidth]{weeks_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/weeks_roc_outSample.pdf} & 
+		\includegraphics[width=.4\textwidth]{graphics/weeks_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{Gibler (2017)}} \\
-		\includegraphics[width=.4\textwidth]{gibler_roc_outSample.pdf} & 
-		\includegraphics[width=.4\textwidth]{gibler_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/gibler_roc_outSample.pdf} & 
+		\includegraphics[width=.4\textwidth]{graphics/gibler_pr_outSample.pdf} \\
 	\end{tabular}

### 2018-02-22 paper/4_app.tex
changes
basically adding a new writeup of gibler to the app section.

@@
 
-\subsection{....}
+\subsection{Gibler}
+The allied variable along with contiguity and joint-democracy comprise the so-called ``control variables'' that are required by the literature.
+The coefficient for the allied variable is negative and imprecisely measured (non-significant in Gibler's terminology).
+It has about the same size when estimated with AME, but is very precisely measured and would be adjudged as very important in Gibler's initial assessment.
+The joint-democracy variable is negative and has a small standard error ($\sigma_{\hat{\beta}}$) in Gibler's Table 6, but when estimated with AME it has virtually zero magnitude and a standard error larger than this, indicating that it has no strong independent effect.
+Both Gibler's estimate and that derived with AME for contiguity show that it is an important independent stimulant to conflict.
+
+Turning away from the less interesting control variables, both approaches show that, consistent with Gibler's argument, parity does not exhibit a strong, precise impact on conflict.
+Rivalry is positive and precisely measured as an influence on conflict with both approaches, though it is about $65\%$ smaller and an order of magnitude more precisely measured by AME.
+Turning to Gibler's main argument, while the logit reported by Gibler confirms his hypothesis that parity at the entry year is very important to the onset of conflicts (positive and precisely measured), this effect is competely eliminated when the dependencies in the data are taken into account by the AME estimation.
+
+Gibler's argument that parity is unimportant because it is static is borne out in both estimations.
+But his argument that parity at the time at which a country enters the international system determines---in part---its propensity toward militarized dispute onsets is not borne out at all.
+
+Finally, we might ask about the performance of Gibler's model when estimated with standard (logit) approaches, compared to its performance using the AME framework.
+As shown above, the AME dominates in a predictive sense.
+Gibler's model is slightly better than a coin-toss in terms of correctly predicting non-onsets of MIDS, and always wrong in predicting actual onsets.
+The AME approach is much more credible.
+
+\subsection{...}
 
@@
 
-mcdonald we recover support fro trade dependenicence and also find eveidence for their highest barrier to trade variable
+%mcdonald we recover support fro trade dependenicence and also find eveidence for their highest barrier to trade variable
 
 for weeks even though we dont find support for her hypothesis we do se esomething interesting in the multiplicative effects
-
-for gibler

### 2018-02-22 paper/app_deprc.tex
sync up

@@
 %Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
-
-\subsection{Re-estimation of McDonald (2004)}
-
-\citet{mcdonald:2004} studies whether trade promotes peace between nations.
-\citet[p.
-547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''.
-Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
-His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
-\citet{mcdonald:2004} measured free trade in two ways.
-The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad; this reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
-This measure thus captures the score of the state in the dyad that possesses higher barriers to trade.
-\citet[p.
-560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
-The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad, and \citet{mcdonald:2004} includes splines to correct for temporal dependence with robust standard errors clustered on each dyad.
-
-\subsection{Re-estimation of Rose (2004)}
-
-\nocite{rose:2004}
-In 2004, Andrew Rose published a study in the \textit{American Economic Review} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations.
-It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007}.
-Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims.
-He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades.
-Estimating this model using OLS within many differing contexts, his conclusion is that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.
-
-When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are similar.
-The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model.
-However, in a sense the AME result provides even stronger evidence against a positive effect of WTO membership: in the original model, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated.
-The fact that this variable attains conventional statistical significance under an AME estimation shows an additional value of this approach to applied researchers -- while accounting for interdependencies and removing that source of bias will remove false positive findings, it can also avoid false negative findings where the bias prevented a researcher from uncovering a consistent relationship.\footnote{This benefit comes up less often in these replications because most published studies are positive, rather than null findings.} In both cases we find  a rejection of the conventional wisdom, that WTO membership increases trade.
-Of course, the controversial result of the paper, strengthened in the AME analysis might be an artifact of measurement error.
-\citep{tomz:etal:2007}.
-While in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model.
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{rose_aeff_top10.pdf}
-	\caption{Nodal Random Effects for AME estimation of Rose (2004), for the countries with the highest and lowest Random Effects.}
-	\label{fig:roser}
-\end{figure}
-\FloatBarrier
-
-The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence.
-Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.}  Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies.
-When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model.
-While the AME model replicates the main (null) result, there are substantial and substantive improvements gained from moving to an AME estimation.
 

### 2018-02-26 paper/4_app.tex
sync up
off to cd for proofing

--- 2018-02-22 paper/4_app.tex
@@
 
-\subsection{Design}
-
-We apply AME to three recent IR studies: \citet{reiter:stam:2003, mcdonald:2004,  weeks:2012, gibler:2017}.
+We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017}.
 Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
-We choose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach.
+We choose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach.
 First, we show that simply by using the AME framework scholars can better model the data generating process behind their events of interest.
@@
 
-We obtained the data for these studies from their replication archives and replicated their main results.\footnote{Without exception this was straightforward to accomplish, thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies below were selected based on how recently they were published and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retreive thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies below were selected based on how recently they were published and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a general linearized model.\footnote{It is important to note that the AME framework is not well suited for studies where the focus may be on studying onsets and dyad-years representing conflict continuations are removed from the sample.
-In studies with this type of focus, network models in general are of limited utility since any interdependencies underlying how events may spread from one actor to another are explicitly masked from the estimation process.} Table~\ref{tab:modelDesign} provides descriptive information for the studies that we replicated.
+Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a general linearized model.
+Table~\ref{tab:modelDesign} provides descriptive information for the studies that we replicated.
+% \footnote{It is important to note that the AME framework is not well suited for studies where the focus may be on studying onsets as in these type of studies dyad-years representing conflict continuations are removed from the sample.
+In studies with this type of focus, network models in general are of limited utility since any interdependencies underlying how events may spread from one actor to another are explicitly masked from the estimation process due to the removal of observations.} ...........
+gibler uses a study onsets.
+in some ways this is problematic in other it is not, i'm (sm) suggesting that we ignore for now.
 
@@
 		& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
-		Reiter \& Stam (2003) &Logit &1945--1995 &  193 & Directed & Robust \\	
-		Weeks (2012) & Logit & 1946--1999 &197 & Directed & Robust \\
-		Gibler (2017) & Logit & 1816--2008 &193 & Undirected & None \\ \bottomrule
+		Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed & Robust \\	
+		Weeks (2012) & Logit & 1946--1999 & 197 & Directed & Robust \\
+		Gibler (2017) & Logit & 1816--2008 & 193 & Undirected & Robust \\ \bottomrule
 	\end{tabular}
@@
 
-Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy.
-Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
+From each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' neatly assembled replication scripts.} Next, utilizing the same covariate specifications, we estimated the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
+Results with alternative values of $K$ are similar.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
+In the directed formulation, separate random effects are used for senders and receivers in both the additive and multiplicative portions of the model.
 
+A key claim that we have made is that by accounting for dependencies inherent in relational data we can better capture the data generating process behind events of interest.
+To assess whether or not the AME approach successfully does this, we turn to an out-of-sample cross validation strategy.
+An out-of-sample approach here is essential as relying on in-sample procedures would enable models with more parameters, such as AME, to simply overfit the data.
+The cross validation procedure works as follows, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
 Then for each $s \in \{1,\ldots,k\}$, we:
@@
 The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-We summarize the performance of the various models in Table~\ref{tab:modelPerfSumm} below.
-For the binary models we provide the area under the Receiver Operator Characteristic (ROC) and Precision Recall (PR) curves.
+For each study we next conducted a series of analyses to discern whether or not the AME model provides any benefit.
+These analyses are summarized in Figure~\ref{fig:perf}.
+The left-most plot in each of the panels evaluates performance using Receiver Operating Characteristic (ROC) curves.
+Models that have a better fit according to this metric will follow the left-hand and then top borders of the ROC space.
+In addition, to ROC curves we also use separation plots, these visualize each of the observations in the dataset according to their predicted probability.
+Observations in which there is an actual link are colored in by the modeling approach.
+Models with a better fit should have the majority of colored links towards the right end of the plot.
+The right-most plot in each of the panels evaluates performance using precision-recall (PR) curves.
+PR curves are useful in situations where correctly predicting events is more interesting than simply predicting non-events \citep{davis:goadrich:2006}.
+This is especially relevant in the context of our applications here, as they each are trying to model conflict in a dyadic setting, which is an infrequent occurrence.
+
 For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance.
-This is important as it indicates that switching to the AME framework---even when using the exact same specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
+This indicates that switching to the AME framework---even when using the same covariate specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
 The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, rather the additive and multiplicative effects we include are capturing underlying structure previously missed by the exogenous covariates in the models.
@@
 	\begin{tabular}{cc}
-		\multicolumn{2}{l}{\textbf{Reiter \& Stam (2003)}} \\
+		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
 		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_roc_outSample.pdf} & 
 		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_pr_outSample.pdf} \\
-		\multicolumn{2}{l}{\textbf{Weeks (2012)}} \\
+		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
 		\includegraphics[width=.4\textwidth]{graphics/weeks_roc_outSample.pdf} & 
 		\includegraphics[width=.4\textwidth]{graphics/weeks_pr_outSample.pdf} \\
-		\multicolumn{2}{l}{\textbf{Gibler (2017)}} \\
+		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
 		\includegraphics[width=.4\textwidth]{graphics/gibler_roc_outSample.pdf} & 
@@
 	\end{tabular}
-	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
+	\caption{Assessments of out-of-sample predictive performance using ROC curves, PR curves, and separation plots.}
 	\label{fig:perf}	
@@
 
-Each of the studies has a crucial finding that we hone in on to further draw into focus the analytical power of the AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
+Ignoring this underlying structure has implications when trying to draw inferences about the effects of covariates.\footnote{As shown in the simulation section, inference in a dyadic setting can become complicated when there are unobserved dependencies.} The fact that there is such a divergence in the performance of AME versus the original estimation procedures highlights that there are unobserved sources of bias in each of these studies.
+We hone in on the main finding of the studies to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
 
@@
 \caption{Here we provide a brief summary of the key variable in each of the replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
-	\begin{tabular}{l p{7cm} l} \toprule
+	\begin{tabular}{l p{6cm} l} \toprule
 		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
@@
 
-\subsection{....}
+Next we discuss each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.
 
-Apart from providing stronger inferential guarantees and predictive performance than conventional approaches, the AME framework
+\subsection{Reiter \& Stam (2003)}
 
-reiter and stam take a look at the srm effeccts to understand where the model is still falling short
+\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
+Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state \citep{peceny:etal:2002}.
+Using a directed dyadic dataset of almost a million observations, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.
+In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true.\footnote{Independent variables focus on various encodings of regime types, contiguity, alliance, and capability measures.
+For a full tabular display of the results see the Appendix.} As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines.
+Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
 
-mcdonald we recover support fro trade dependenicence and also find eveidence for their highest barrier to trade variable
+The key variables in their original model were indicator for whether or not the sender in the directed dyad is a personalist regime and the target a democratic (``Pers/Democ directed dyad"), and whether the sender was a democratic regime and the target personalist (``Democ/Personalist directed dyad'').
+They had found that the Pers/Democ Directed Dyad indicator was positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
+In our re-estimation using the AME framework, we confirm these results, indicating that dictators are more likely to initiate or be in conflict with democratic regimes but not vice versa.
 
-for weeks even though we dont find support for her hypothesis we do se esomething interesting in the multiplicative effects
+Though we are able to confirm their results it is not the case that employing the AME model does not still come with benefits.
+As already shown in Figure~\ref{fig:perf}, our approach performs notably better in reflecting the data generating process.
+The reason for this is that there is still underlying structure within this conflict system that the Reiter \& Stam model does not fully capture.
+To highlight this we visualize the estimated sender random effects ($a_{i}$) from the SRRM portion of the AME framework in Figure~\ref{fig:reiter_stam_aEff}.
 
-for gibler
+\begin{figure}[ht]
+	\begin{tabular}{cc}
+	\includegraphics[width=.7\textwidth]{graphics/reiter_stam_aEff_map.pdf} &
+	\includegraphics[width=.3\textwidth]{graphics/reiter_stam_aEff_line.pdf} \\
+	\end{tabular}
+	\caption{Estimates of sender random effects ($a_{i}$) from AME for the Reiter \& Stam (2003) model.
+Positive values indicate that the particular country is more likely to be involved in conflict than predicted by the covariates in the model, and negative values indicate that the country is less likely to be involved in conflict.
+Left-panel shows $a_{i}$ estimates for all countries and right-panel highlights the top ten countries in terms of positive and values of $a_{i}$.}
+	\label{fig:reiter_stam_aEff}
+\end{figure}
+\FloatBarrier
+
+The visualization of the sender random effects highlights that the behavior of a number of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
+Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than what the covariates specified in the model would predict.
+Further other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than what the fixed effects in the model would predict.
+None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.
+
+\subsection{Weeks (2012)}
+
+\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
+She argues that in some circumstances autocrats are held accountable for their foreign policy decisions, and that this is dependent on the audiences of autocrats.
+When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case.
+Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
+To study this question, Weeks employs a directed dyad design of conflict similar to that used by Reiter \& Stam.
+
+The major innovation in her study resides in the nuanced way in which she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
+The key finding of this work is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes and that b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone.
+When analyzing the results using AME, however, we find that the parameters on each of her autocratic regime type variables are too imprecisely measured to draw any inferences.
+The implication being that none of the findings from her original analysis are confirmed when accounting for dependencies.
+
+There is also a striking difference in terms of the ability of Weeks' original model and our estimation using AME in terms of capturing the data generating process.
+Specifically, The area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models are employing the same fixed effects.
+As with Reiter \& Stam, the divergence is a result of the GLM framework not being able to account for underlying structure in the event of interest.
+To uncover some of this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$.
+These random effects account for higher order dependencies in relational data that manifest as a result of the presence of homophily and stochastic equivalence in dyadic data.
+
+\begin{figure}[!h]
+	\includegraphics[width=\textwidth]{graphics/weeks_circPlot.png}
+	\caption{ Visualization of multiplicative effects for Weeks (2012).
+Each circle designates a country and the color corresponds to the legend at the center of the visualization.
+Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries.
+The inner ring clusters countries by the similarity of who they receive conflict from.
+	}
+	\label{fig:weekscirc}
+\end{figure}
+\FloatBarrier
+
+To visualize the multiplicative effects, we develop the circular diagram shown in Figure~\ref{fig:weekscirc}.
+The nodes throughout the diagram represent countries and they are colored according to their geographic position -- a legend is provided in the center.
+The outer ring visualizes higher order dependence patterns that countries have in terms of their sending relationships -- country positions here are estimated in the $\mathbf{u}_{i}$ random effects described in Equation~\ref{eqn:ame}.
+Countries that are more proximate to each other in this outer ring are more likely to send or initiate conflictual events with similar targets.
+The inner ring, based on estimates of $\mathbf{v}_{i}$, is constructed such that countries more proximate to each other are more likely to receive conflict from the same countries.
+Last, the distance between countries in the inner and outer rings proportionately reflects who a country is more likely to be in conflict with.
+
+Interestingly, when examining Figure~\ref{fig:weekscirc} we quickly are able to reveal a number of clusters.
+For example, in the top left corner we see the US, UK, Germany, Canada, and Israel.
+These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.
+Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
+We see similar clusters develop in the inner ring, specifically, we again see the US, UK, Germany, Canada, and Israel clustering together indicating that they are more likely to receive conflict from the same countries.
+Additionally, we again see authoritarian regimes like North Korea, Iraq, and Syria clustering.
+Further, the cluster of democratic and authoritarian countries are facing each other in the inner and outer rings, indicating that they are more likely to be in conflict with one another.
+
+Most important, in evaluating this visualization we are actually able to see some of Weeks' logic play out in terms of how states behave in the conflict system.
+Specifically, Iraq, Syria, Libya, and North Korea all fell under Weeks' ``boss" category, and each of these states tends to cluster together in the inner and outer rings.
+This indicates that even though we do not find support for Weeks' assertion that certain authoritarian regime types are more likely to initiate conflict, we do find that these regimes are more likely to behave similarly in terms of who they are more likely to target and receive conflict from.
+
+\subsection{Gibler (2017)}
+
+The last re-estimation we conduct with the AME model involves a study by \citet{gibler:2017}.
+Gibler argues that the long-standing relationship between the relative parity of capabilities and initiation of militarized interstate disputes (MIDs) is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
+This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
+
+To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for his primary variable, parity of the members of a dyad at the year in which they entered the international system.
+However, when we reestimate using AME, this measure becomes too imprecisely measured for us to draw any inferences about its effect on the initiation of MIDs.
+Not only is our estimate of the effect of Gibler's parity variable small, but it has a very large relative standard error---over a magnitude larger than the parameter itself.
+
+In the two previous replications, we showed how to parse apart underlying structure using the SRM and LFM portions of the AME framework.
+Here we turn our focus to parameter interpretation in a substantive context.
+Specifically, even when the GLM and AME frameworks produce results that may seem to be in accordance with one another, the substantive interpretations of the effects of covariates can differ notably as the AME model uses a set of random effects to account for unobserved factors.
+To do this, we focus on the effect of rivalry in the onset of a MID.
+Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects the two models provide are quite stark.
+
+To clarify the difference, we turn to a simulation based approach.
+We employ mean or modal values for all independent variables, except we change the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none.
+This provides us with two scenarios, one in which rivalry is set to one and the other zero, while in both scenarios all other parameters are set to their measure of central tendency.
+The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: the GLM estimation and our AME approach.
+
+\begin{figure}
+	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
+	\includegraphics[width=\textwidth]{graphics/gibler_margeff.pdf}
+ 	\label{fig:gibmargeff}
+ \end{figure}
+
+As Figure~\ref{fig:gibmargeff} illustrates, the substantive results differ notably.
+First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
+These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
+Thus, you get quite substantially different expected values from these two models.
+
+\subsection{Lessons Learned}
+
+First, utilizing the AME framework enables scholars to better model the underlying structure that dyadic data often comes with.
+In each of the replications, the predictive performance in an out-of-sample context of the AME framework is notably superior.
+Specifically, not only does AME perform better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves).
+
+Second, in both the simulation section and these replications we have shown that results can change notably when interdependencies are not taken into account.
+Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
+This means that significance testing (for better or worse) is compromised when network effects are ignored.
+
+Third, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights can emerge from the structure that the random effects of the AME framework uncover.
+In particular, there is actual information about the dependencies so that clusters can be identified, and this information can also be used to generate new hypotheses.
+This kind of information is absent in standard approaches and adds to our ability to explain specific as well as general results.
+
+Fourth, it is evident that the actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.
+This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.

### 2018-02-26 paper/5_concl.tex
off to cd for proofing

@@
 This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
-The second analysis is a replication of four recent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
-These four studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
+The second analysis is a replication of recent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
+These studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.
 We then re-analyzed each study using the AME model.
 In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, and c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
-This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that are not valid.
+
+% sm: i dont feel like pissing everyone off
+%This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that are not valid.
 

### 2018-02-26 paper/6_appendix.tex
off to cd for proofing

@@
 
-Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis.
+% Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis.
 %For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.
@@
 
-There is often little theoretical reason to choose a particular value of $R$ (above).
-One strategy is to estimate models at different values of $R$ and compare goodness of fit statistics across models.
+There is often little theoretical reason to choose a particular value of $\sf{R}$ (above).
+One strategy is to estimate models at different values of $\sf{R}$ and compare goodness of fit statistics across models.
 
@@
 
-The code below presents an example of an AME model running in parallel across 4 different levels of $R$.
+The code below presents an example of an AME model running in parallel across 4 different levels of $\sf{R}$.
 Note also that the model is using starting values from a previous run, defined in \textit{startVals0}.

### 2018-02-26 code/2_codeForPaperGraphics/gibler_posterior.R

@@ -3,7 +3,7 @@ rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/gibler_2017/outputData/'
-	plotPath = '~/Research/netsMatter/paper/' }
+	plotPath = '~/Research/netsMatter/paper/graphics/' }
 
 source('~/Research/netsMatter/code/helpers/functions.R')
 source('~/Research/netsMatter/code/helpers/ameHelpers.R')

### 2018-02-26 code/helpers/binPerfHelpers.R

@@ -222,7 +222,7 @@ ggPerfCurves = function(predDfs, suffix, cutBy=0.001){
     annotate('text', hjust=0, x=.6, y=c(0.65,.75), 
       label=rev(rownames(aucSumm))) + 
     annotate('text', hjust=0, x=.78, y=c(0.65,.75), 
-      label=rev(apply(aucSumm, 1, function(x){paste(x, collapse='     ')})), size=5 )
+      label=rev(apply(aucSumm, 1, function(x){paste(x, collapse='    ')})), size=4 )
   ggsave(tmp, file=paste0(plotPath, suffix, '_pr_outSample.pdf'), width=5, height=5)  
 }  
 ####################################################################

### 2018-03-01 paper/netsMatter.tex
ab for rose
moving stuff around
off to cd for proofing
abstract edits

@@
 \rfoot{\fancyplain{}{\thepage}}
+
 % references to graphics
 \makeatletter
-% let's keep them in the same directory \def\input@path{{}
+% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
+% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
 
@@
 Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
-As a result, many studies often produce results with biased coefficient estimates and poorly calibrated standard errors.
+As a result, studies often produce results with biased coefficient estimates and poorly calibrated standard errors.
 These biases have profound consequences for evaluating parametric models.
-We present an alternative, regression-based, approach that accounts for the dependencies complicating this type of analysis.
+We present an alternative, regression-based approach that accounts for the dependencies complicating this type of analysis.
 We first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
-In addition, we replicate five recent studies in recent international relations scholarship, comparing the standard approach to our alternative.
+In addition, we replicate three recent studies in recent international relations scholarship, and compare  the standard approach to our alternative.
 For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables.
-Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
+Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and for  modeling the data generating process behind outcomes of interest.
 \end{abstract}
@@
 \clearpage
+
+%\bibliography{/Users/maxgallop/documents/whistle/master}
 \bibliography{/Users/s7m/whistle/master}

### 2018-03-01 paper/1_intro.tex
intro edits

@@
 In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
-This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix.
-\citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
+This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
 
@@
 We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads--for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
+Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads---for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
 The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.
@@
 The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
-Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while accounting for the data generating process behind political events of interest.
+Thus, the AME approach allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena and yet also account for the data generating process behind such events of interest.
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-03-01 paper/2_model.tex
reorg by mdw, tiny changes
model & sim

@@
 Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
-While most scholars begin with the assumption that each dyadic interaction is taking place in isolation of the others, we know this assumption to be false both in theory and in practice.
+While most scholars begin with the assumption that each dyadic interaction takes place in isolation of the others, we know this assumption to be false both in theory and in practice.
 Relational data comes with an explicit structure that generally leads to particular types of dependencies.
@@
 		\textsc{Sender Heterogeneity} & \textsc{Receiver Heterogeneity} \\
-		\includegraphics[width=.4\textwidth]{adjRowDep.png} & \includegraphics[width=.4\textwidth]{adjColDep.png} \\
+		\includegraphics[width=.4\textwidth]{graphics/adjRowDep.png} & 
+		\includegraphics[width=.4\textwidth]{graphics/adjColDep.png} \\
 		\textsc{Sender-Receiver Covariance} & \textsc{Reciprocity} \\
-		\includegraphics[width=.4\textwidth]{adjRowColCovar.png} & \includegraphics[width=.4\textwidth]{adjRecip.png} \\
+		\includegraphics[width=.4\textwidth]{graphics/adjRowColCovar.png} & 
+		\includegraphics[width=.4\textwidth]{graphics/adjRecip.png} \\
 	\end{tabular}
@@
 \begin{figure}[ht]
-	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
+	\includegraphics[width=.6\textwidth]{graphics/stochEquiv_v2.pdf}
 	\caption{Higher-order dependence patterns in a network.}
@@
 Stochastic equivalence refers to a type of pattern in which actors can be divided into groups such that members of the same group have similar patterns of relationships.
-In the example above, each of the left leaning blogs would be considered stochastically equivalent to one another because any given left-leaning blog is more likely to interact with a blog of a similar political position and less likely to interact with one with a divergent political position.
+In the example above, each of the left leaning blogs would be considered stochastically equivalent to one another because any given left-leaning blog is more likely to interact with a blog of a similar political position and less likely to interact with one of a divergent political position.
 

### 2018-03-01 paper/3_sim.tex
reorg by mdw, tiny changes
model & sim

@@
 	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{ameSimBias_all.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimBias_all.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{ameSimCover_all.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimCover_all.pdf} \\
 \end{figure}
@@
 Moreover, the AME approach allows scholars to better understand what parameters their model may be missing.
-In the case of the simulation here, $W$ is set as an unobserved dyadic covariate that had a homophilous effect on $Y$.
-Homophilous because $W$ within this framework is simply an example of a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$.
+In the case of the simulation here, $W$ is set as an unobserved dyadic covariate that has a homophilous effect on $Y$.
+The effect of $W$ is homophilous within this framework because it is a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$.
 This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$.
@@
 	\label{fig:ameCorr}
-	\includegraphics[width=1\textwidth]{ameSimCorr.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimCorr.pdf} \\
 \end{figure}
 
-What this simulation has shown is that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is to directly estimate unobserved dependencies through the random effects structure of the model.
-Scholars can use this framework in an iterative fashion: beginning with an estimated a model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their models.
-At the very least, what this simulation section shows is that the AME model can help in estimating the true effect and conducting inference on the independent variables that scholars have operationalized to test their theoretical propositions.
+The  simulation shows that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is to directly estimate unobserved dependencies through the random effects structure of the model.
+Scholars can use this framework in an iterative fashion: beginning with an estimated model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their model.
+At the very least,  this simulation  demonstrates that the AME model can help conduct inference and estimate the true effect of independent variables that scholars operationalize to test theoretical propositions.

### 2018-03-01 paper/4_app.tex
app edits

@@
 We choose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach.
-First, we show that simply by using the AME framework scholars can better model the data generating process behind their events of interest.
+First, we show that simply by using the AME framework scholars can better model the data generating process behind events of interest.
 Second, the results of AME estimation are interpretable alongside results using standard approaches, but, as shown in the simulation section, have the additional benefit of being able to take into account dependencies that may otherwise complicate inference.
-Third, through using this approach, we can also quantify the degree to which first, second, and third order dependencies are present in events of interest.
+Third, through using this approach we can also quantify the degree to which first, second, and third order dependencies are present.
 
-We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retreive thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies below were selected based on how recently they were published and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies below were selected based on how recently they were published and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
-Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a general linearized model.
+Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model.
 Table~\ref{tab:modelDesign} provides descriptive information for the studies that we replicated.
@@
 To assess whether or not the AME approach successfully does this, we turn to an out-of-sample cross validation strategy.
-An out-of-sample approach here is essential as relying on in-sample procedures would enable models with more parameters, such as AME, to simply overfit the data.
-The cross validation procedure works as follows, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
+An out-of-sample approach here is essential since relying on in-sample procedures would enable models with more parameters, such as AME, to simply overfit the data.
+The cross validation procedure works as follows.
+For each study  we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
 Then for each $s \in \{1,\ldots,k\}$, we:
@@
 
-The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from using a subset of the data that does not include $y_{ij,t}$.
-For each study we next conducted a series of analyses to discern whether or not the AME model provides any benefit.
+The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from a subset of the data that does not include $y_{ij,t}$.
+Next we conducted a series of analyses to discern whether or not the AME model provides any benefit for each study.
 These analyses are summarized in Figure~\ref{fig:perf}.
 The left-most plot in each of the panels evaluates performance using Receiver Operating Characteristic (ROC) curves.
-Models that have a better fit according to this metric will follow the left-hand and then top borders of the ROC space.
-In addition, to ROC curves we also use separation plots, these visualize each of the observations in the dataset according to their predicted probability.
-Observations in which there is an actual link are colored in by the modeling approach.
-Models with a better fit should have the majority of colored links towards the right end of the plot.
+Models that have a better fit according to this metric will follow the upper-left corner border of the ROC space.
+In addition to ROC curves we also use separation plots.
+Separation plots plots visualize each of the observations in the dataset according to their predicted probability.
+In this graph, the shaded panels correspond to the the occurrence of an event.
+Darker panels are events, lighter panels are non events.
+If a model performs well, then the events that actually occur would stack to the far right of the graph, where the predicted probabilities generated by the model are highest.
 The right-most plot in each of the panels evaluates performance using precision-recall (PR) curves.
 PR curves are useful in situations where correctly predicting events is more interesting than simply predicting non-events \citep{davis:goadrich:2006}.
-This is especially relevant in the context of our applications here, as they each are trying to model conflict in a dyadic setting, which is an infrequent occurrence.
+This is especially relevant in the context of our applications here, as they each are trying to model conflict in a dyadic setting which is an infrequent occurrence.
 
@@
 This indicates that switching to the AME framework---even when using the same covariate specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
-The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, rather the additive and multiplicative effects we include are capturing underlying structure previously missed by the exogenous covariates in the models.
+The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, instead it suggests that the additive and multiplicative effects we include are capturing underlying structure previously missed by the exogenous covariates in the models.
 
@@
 Ignoring this underlying structure has implications when trying to draw inferences about the effects of covariates.\footnote{As shown in the simulation section, inference in a dyadic setting can become complicated when there are unobserved dependencies.} The fact that there is such a divergence in the performance of AME versus the original estimation procedures highlights that there are unobserved sources of bias in each of these studies.
-We hone in on the main finding of the studies to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.
+We hone in on the main finding of each study to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.
 In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
@@
 An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
-These differences may change dramatically when interdependencies are taken into account directly.
+These differences may change dramatically once interdependencies are taken into account.
 This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates.
 As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates.
-At best, this method introduces noise and imprecision into results, and at worst can produce misleading outcomes.
+At best, this method introduces noise and imprecision into results and at worst can produce misleading outcomes.
 
@@
 \citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state \citep{peceny:etal:2002}.
+Their work contests prior scholarship that claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state \citep{peceny:etal:2002}.
 Using a directed dyadic dataset of almost a million observations, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.
@@
 
-The key variables in their original model were indicator for whether or not the sender in the directed dyad is a personalist regime and the target a democratic (``Pers/Democ directed dyad"), and whether the sender was a democratic regime and the target personalist (``Democ/Personalist directed dyad'').
-They had found that the Pers/Democ Directed Dyad indicator was positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
+The key variables in the original model measure  whether or not the sender in the directed dyad is a personalist regime and the target a democratic (``Pers/Democ directed dyad"), and whether the sender is a democratic regime and the target personalist (``Democ/Personalist directed dyad'').
+The authors find that the Pers/Democ Directed Dyad indicator is positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
 In our re-estimation using the AME framework, we confirm these results, indicating that dictators are more likely to initiate or be in conflict with democratic regimes but not vice versa.
 
-Though we are able to confirm their results it is not the case that employing the AME model does not still come with benefits.
+Though we are able to confirm their results it is not the case that employing the AME model offers few benefits.
 As already shown in Figure~\ref{fig:perf}, our approach performs notably better in reflecting the data generating process.
@@
 The visualization of the sender random effects highlights that the behavior of a number of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
-Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than what the covariates specified in the model would predict.
+Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than what the covariates would predict.
 Further other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than what the fixed effects in the model would predict.
@@
 
-The major innovation in her study resides in the nuanced way in which she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
-The key finding of this work is that a) juntas, boss type, and strongmen type regimes are more likely to initiate conflict than machine-type regimes and that b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strong-men are more conflict-prone.
+The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
+The key finding of this work is that a) juntas, boss, and strongmen type regimes are more likely to initiate conflict than machine-type regimes and that b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strongmen are more conflict-prone.
 When analyzing the results using AME, however, we find that the parameters on each of her autocratic regime type variables are too imprecisely measured to draw any inferences.
-The implication being that none of the findings from her original analysis are confirmed when accounting for dependencies.
+Consequentially, none of the findings from her original analysis are confirmed once accounting for dependencies.
 
-There is also a striking difference in terms of the ability of Weeks' original model and our estimation using AME in terms of capturing the data generating process.
-Specifically, The area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models are employing the same fixed effects.
-As with Reiter \& Stam, the divergence is a result of the GLM framework not being able to account for underlying structure in the event of interest.
+There is also a striking difference between  Weeks' original model and our estimation using AME in terms of capturing the data generating process.
+Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same fixed effects.
+As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure in the event of interest.
 To uncover some of this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$.
-These random effects account for higher order dependencies in relational data that manifest as a result of the presence of homophily and stochastic equivalence in dyadic data.
+These random effects account for higher order dependencies  that manifest as a result of homophily and stochastic equivalence in dyadic data.
 
@@
 
-Interestingly, when examining Figure~\ref{fig:weekscirc} we quickly are able to reveal a number of clusters.
+Interestingly, Figure~\ref{fig:weekscirc} reveals a number of interesting clusters.
 For example, in the top left corner we see the US, UK, Germany, Canada, and Israel.
@@
 Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
-We see similar clusters develop in the inner ring, specifically, we again see the US, UK, Germany, Canada, and Israel clustering together indicating that they are more likely to receive conflict from the same countries.
-Additionally, we again see authoritarian regimes like North Korea, Iraq, and Syria clustering.
-Further, the cluster of democratic and authoritarian countries are facing each other in the inner and outer rings, indicating that they are more likely to be in conflict with one another.
+We observe similar clusters in the inner ring.
+Specifically, we again see the US, UK, Germany, Canada, and Israel clustering together indicating that they are more likely to receive conflict from the same countries.
+The  cluster of democratic and authoritarian countries facing each other in the inner and outer rings indicate that they are more likely to engage in conflict with one another.
 
-Most important, in evaluating this visualization we are actually able to see some of Weeks' logic play out in terms of how states behave in the conflict system.
+Perhaps most  importantly, an evaluation of this visualization highlights Weeks' expectation of how states behave in the conflict system.
 Specifically, Iraq, Syria, Libya, and North Korea all fell under Weeks' ``boss" category, and each of these states tends to cluster together in the inner and outer rings.
-This indicates that even though we do not find support for Weeks' assertion that certain authoritarian regime types are more likely to initiate conflict, we do find that these regimes are more likely to behave similarly in terms of who they are more likely to target and receive conflict from.
+This indicates that even though we do not find support for Weeks' assertion that certain authoritarian regime types are more likely to initiate conflict, we do find that these regimes are more likely to behave similarly in terms of who they are more likely to both target  and receive conflict from.
 
@@
 
-To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for his primary variable, parity of the members of a dyad at the year in which they entered the international system.
+To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for his primary variable: parity of the members of a dyad at the year in which they entered the international system.
 However, when we reestimate using AME, this measure becomes too imprecisely measured for us to draw any inferences about its effect on the initiation of MIDs.
@@
 Specifically, even when the GLM and AME frameworks produce results that may seem to be in accordance with one another, the substantive interpretations of the effects of covariates can differ notably as the AME model uses a set of random effects to account for unobserved factors.
-To do this, we focus on the effect of rivalry in the onset of a MID.
-Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects the two models provide are quite stark.
+To do this, we focus on the effect of rivalry on  MID initiation.
+Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects the two models provide  starkly contrast.
 
@@
 These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
-Thus, you get quite substantially different expected values from these two models.
+Thus, we find substantially different expected values from these two models.
 
@@
 
-First, utilizing the AME framework enables scholars to better model the underlying structure that dyadic data often comes with.
-In each of the replications, the predictive performance in an out-of-sample context of the AME framework is notably superior.
+First, utilizing the AME framework enables scholars to better model the underlying structure often inherent to dyadic data.
+In each of the replications, the AME's predictive performance out-of-sample is notably superior.
 Specifically, not only does AME perform better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves).
 
-Second, in both the simulation section and these replications we have shown that results can change notably when interdependencies are not taken into account.
-Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with poorly calibrated standard errors.
+Second, in both the simulation  and replication sections  we have shown that results can change notably when interdependencies are not taken into account.
+Not only are coefficients biased in the GLM approaches, but they are often imprecisely measured with poorly calibrated standard errors.
 This means that significance testing (for better or worse) is compromised when network effects are ignored.
 
-Third, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights can emerge from the structure that the random effects of the AME framework uncover.
+Third, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights can emerge from the structure uncovered by the random effects of the AME framework.
 In particular, there is actual information about the dependencies so that clusters can be identified, and this information can also be used to generate new hypotheses.
@@
 
-Fourth, it is evident that the actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.
+Fourth, it is evident that the actual results---not the estimated coefficients and their covariances---generated by the models differ greatly in expectations.
 This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.

### 2018-03-01 paper/5_concl.tex
the end

@@
 The first is a simulation where the characteristics of the network are known.
-This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared with standard approach employed in international relations to study dyadic data (i.e., GLM models).
+This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared to the standard approach employed in international relations to study dyadic data (i.e., GLM models).
 The second analysis is a replication of recent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
@@
 
-% sm: i dont feel like pissing everyone off
-%This leads us to speculate that many of the findings in the international relations literature may be fragile in the sense that they can only be obtained under stringent assumptions that are not valid.
-
 It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.

### 2018-03-02 paper/netsMatter.tex
app edits
abstract

@@
 Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
-Yet, contemporary research brushes aside the complications that come with analyzing interdependent relational data.
-Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
-As a result, studies often produce results with biased coefficient estimates and poorly calibrated standard errors.
-These biases have profound consequences for evaluating parametric models.
-We present an alternative, regression-based approach that accounts for the dependencies complicating this type of analysis.
-We first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
-In addition, we replicate three recent studies in recent international relations scholarship, and compare  the standard approach to our alternative.
-For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables.
-Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and for  modeling the data generating process behind outcomes of interest.
+These types of data do not typically satisfy the conditional independence criterion required for most statistical approaches.
+Consequently, modeling approaches that fail to adequately capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors.
+We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that more fully accounts for the dependencies in dyadic data.
+First, we conduct a simulation that reveals how the model captures dependencies that emerge in relational data and demonstrate why accounting for these processes improves inference.
+Second, through replication, we compare the AME model to approaches used in three studies from recent international relations scholarship.
+For each study, we find that the modeling approach used in the original study performs notably worse at capturing the data generating process compared to the AME.
+Further, these conventional methods overstate the effect of key variables and underestimate the uncertainty in these effects.
+In sum, our work shows the consequences of failing to seriously consider dependencies inherent to dyadic data.
+Perhaps most importantly, by precisely modeling the data generating process underlying political phenomena we improve scholars’ ability to conduct inferential analyses.
+
 \end{abstract}
@@
 %\bibliography{/Users/maxgallop/documents/whistle/master}
-\bibliography{/Users/s7m/whistle/master}
+%\bibliography{/Users/s7m/whistle/master}
  % \bibliography{/Users/mdw/Git/whistle/master}
-% \bibliography{/Users/cassydorff/ProjectsGit/master}
+ \bibliography{/Users/cassydorff/ProjectsGit/master}
 % \bibliographystyle{authordate1}

### 2018-03-02 paper/netsMatter.tex
minor changes to abstract

@@
 	Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
-These types of data do not typically satisfy the conditional independence criterion required for most statistical approaches.
-Consequently, modeling approaches that fail to adequately capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors.
-We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that more fully accounts for the dependencies in dyadic data.
-First, we conduct a simulation that reveals how the model captures dependencies that emerge in relational data and demonstrate why accounting for these processes improves inference.
-Second, through replication, we compare the AME model to approaches used in three studies from recent international relations scholarship.
+Yet, modeling approaches that fail to properly capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors.
+We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the dependencies in dyadic data.
+First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference.
+Second, we compare the AME model to approaches used in three studies from recent international relations scholarship.
 For each study, we find that the modeling approach used in the original study performs notably worse at capturing the data generating process compared to the AME.
-Further, these conventional methods overstate the effect of key variables and underestimate the uncertainty in these effects.
-In sum, our work shows the consequences of failing to seriously consider dependencies inherent to dyadic data.
-Perhaps most importantly, by precisely modeling the data generating process underlying political phenomena we improve scholars’ ability to conduct inferential analyses.
-
+Further, conventional methods misstate the effect of key variables and the uncertainty in these effects.
+In sum, our work shows the consequences of failing to take the dependencies inherent to dyadic data seriously.
+Most importantly, by better modeling the data generating process underlying political phenomena, the AME framework improves scholars' ability to conduct inferential analyses.
 \end{abstract}

### 2018-03-03 paper/1_intro.tex
again

@@
 
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM) to test hypothesized relationships.
+\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
-Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
+Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
 In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
@@
 
-In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to better reflect the interdependencies underlying the data generating process of dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2018-03-03 paper/1_intro.tex
moreIntro

@@
 We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads---for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
+Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads.
+For example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
 The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.
@@
 Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature.
-The comparison reveals that in accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature.
-Consequently, AME produces results that, at times, differ from those found in the original study as well as from the broader literature.
-Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies.
+The comparison reveals that by accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature.
+Consequently, AME generates results that, at times, differ from those found in the original study as well as from the broader literature.
+Moreover, we demonstrate that the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies.
 Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
 
-The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
-Thus, the AME approach allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena and yet also account for the data generating process behind such events of interest.
+Our AME approach advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
+Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena and account for the data generating process behind such events of interest.
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-03-03 paper/5_concl.tex
conlc

@@
 
-International relations is generally about the interactions and dependencies among a set of countries or other important actors such as international governmental organizations (IGOs).
-This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} It is clear from a survey of the literature and from work in this area published as recently as 2017 that many find dyadic data to be an important touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
+International relations is about the interactions, relationships, and dependencies among countries or other important international actors.
+This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} Research published as recently as 2017 and a broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
 
@@
 Virtually all of the standard statistical models---ordinary least squares and logistic regressions, to name a few---fail if the data are not conditionally independent.
-This fact has been accepted when it comes to temporal dependencies, but adoption of methods to account for network dependencies have seen less progress.
-By definition dyadic data are not iid and thus the standard approaches can not be used cavalierly to analyze these data.
+This fact has been accepted as it relates to temporal dependencies, but adoption of methods to account for network dependencies have seen less progress.
+By definition dyadic data are not iid and thus the standard approaches cannot be used cavalierly to analyze these data.
 \citet{signorino:1999} showed why this is true of models of strategic interaction, but it is more broadly true of models that employ dyadic data.
@@
 
-To explore this approach in the context of international relations we have presented two broad analyses.
+To explore this approach in the context of international relations we have presented two analyses.
 The first is a simulation where the characteristics of the network are known.
 This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared to the standard approach employed in international relations to study dyadic data (i.e., GLM models).
-The second analysis is a replication of recent studies that have been published recently using a broad range of dyadic data to draw inferences about international relations.
+The second analysis is a replication of recent studies that use a broad range of dyadic data to draw inferences about international relations.
 These studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.

### 2018-03-03 paper/3_sim.tex
sim

@@
 
-We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM).
+Through a simulation study, we highlight the utility of AME as an inferential tool for dyadic analysis.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM).
 While both these models have led to numerous contributions to a variety of literatures, the applicability of these approaches may be limited to certain types of networks and individual level characteristics.
 Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
-The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an unaccounted for higher order dependence pattern.
-For instance, assume that the true data-generating process for a particular $Y$ is given by:
+The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
+Assume that the true data-generating process for a particular $Y$ is given by:
 
@@
 $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
-We compare inference for $\mu$ and $\beta$---the latter parameter would be of primarily theoretical concern for applied scholars---using three models:
+We compare inference for $\mu$ and $\beta$---the latter parameter would be of primary concern for applied scholars---using three models:
 
@@
 In the second model, we use the AME framework described in the previous section.
-For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated in the presence of unobserved dependencies, $W$.
+For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated by the presence of unobserved dependencies, $W$.
 In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and include it in our specification for $Y$.
@@
 
-We compare the performance of the models first in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias} by depicting the average $\mu$ and $\beta$ estimates from the simulations for the three models.
+We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias} by depicting the average $\mu$ and $\beta$ estimates from the simulations for the three models.
 The panels in the left show the results for when the number of actors is set to 50 and those on the right for 100; and the top pair of panels represents the estimates for $\mu$ while the bottom pair do the same for $\beta$.

### 2018-03-03 paper/3_sim.tex
more sim

@@
 For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.\footnote{The value of $\gamma$ is also set to 1, which corresponds to an example where the $W$ character is associated with homophily.} We conduct two sets of simulations, one in which the number of actors in the network is set to 50 and the other at 100.
-In total, we ran 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimated a standard, AME, and oracle model.
+In total, we run 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimate a standard, AME, and oracle model.
 
-We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias} by depicting the average $\mu$ and $\beta$ estimates from the simulations for the three models.
-The panels in the left show the results for when the number of actors is set to 50 and those on the right for 100; and the top pair of panels represents the estimates for $\mu$ while the bottom pair do the same for $\beta$.
+We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias} $\mu$ and $\beta$ estimates from the simulations for the three models.
+The panels on the left show the results for when the number of actors is set to 50 and on the  right for 100 actors.
+The top pair of panels represents the estimates for $\mu$ while the bottom pair do the same for $\beta$.
 In each case, we find that the estimates for $\mu$ and $\beta$ produced by the standard approach are notably off from their true values.
-On the other hand, the AME model performs just as well as the oracle case in estimating the true parameter values.
+On the other hand, the AME model performs just as well as the oracle at estimating the true parameter values.
 
@@
 In the case of the simulation here, $W$ is set as an unobserved dyadic covariate that has a homophilous effect on $Y$.
-The effect of $W$ is homophilous within this framework because it is a dyadic attribute involving $i$ and $j$ that positively affects the degree to which they will interact with one another, i.e., $y_{ij}$.
+The effect of $W$ is homophilous within this framework because it is a dyadic attribute involving $i$ and $j$ that positively affects the degree to which actors interact with one another, i.e., $y_{ij}$.
 This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$.
@@
 Additionally, we calculate the median across the correlations and display the result using a vertical line.
-For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that this framework is not simply capturing noise but can be used as a tool to estimate unobserved structure.
+For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the AME does not simply capture noise but also works as a tool to estimate unobserved structure.
 

### 2018-03-04 paper/4_app.tex
appAgain

@@
 We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017}.
-Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.
-We choose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach.
+Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for the behavior of actors in the system.
+We demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach.
 First, we show that simply by using the AME framework scholars can better model the data generating process behind events of interest.
-Second, the results of AME estimation are interpretable alongside results using standard approaches, but, as shown in the simulation section, have the additional benefit of being able to take into account dependencies that may otherwise complicate inference.
+Second, the results of AME estimation are interpretable alongside results using standard approaches.
 Third, through using this approach we can also quantify the degree to which first, second, and third order dependencies are present.
 
-We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The studies below were selected based on how recently they were published and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
+We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
 Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
@@
 
-From each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' neatly assembled replication scripts.} Next, utilizing the same covariate specifications, we estimated the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
+For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' neatly assembled replication scripts.} Next, utilizing the same covariate specifications, we estimated the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
 Results with alternative values of $K$ are similar.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
@@
 
-A key claim that we have made is that by accounting for dependencies inherent in relational data we can better capture the data generating process behind events of interest.
+A key claim that we have made is that by accounting for dependencies inherent to relational data we can better capture the data generating process behind events of interest.
 To assess whether or not the AME approach successfully does this, we turn to an out-of-sample cross validation strategy.
-An out-of-sample approach here is essential since relying on in-sample procedures would enable models with more parameters, such as AME, to simply overfit the data.
-The cross validation procedure works as follows.
+An out-of-sample approach is essential since relying on in-sample procedures would enable models with more parameters, such as AME, to simply overfit the data.
+The cross validation procedure is executed as follows.
 For each study  we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
@@
 Separation plots plots visualize each of the observations in the dataset according to their predicted probability.
-In this graph, the shaded panels correspond to the the occurrence of an event.
+In this graph, the shaded panels correspond to the occurrence of an event.
 Darker panels are events, lighter panels are non events.
@@
 PR curves are useful in situations where correctly predicting events is more interesting than simply predicting non-events \citep{davis:goadrich:2006}.
-This is especially relevant in the context of our applications here, as they each are trying to model conflict in a dyadic setting which is an infrequent occurrence.
+This is especially relevant in the context of our applications here, as they each are trying to model conflict within dyads, which is an infrequent occurrence.
 
@@
 This indicates that switching to the AME framework---even when using the same covariate specification as the original studies---enables scholars to better represent the data generating process of their events of interest.
-The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters, instead it suggests that the additive and multiplicative effects we include are capturing underlying structure previously missed by the exogenous covariates in the models.
+The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters.
+Instead, it suggests that the additive and multiplicative effects are capturing underlying structure previously missed by the exogenous covariates in the models.
 
@@
 
-Ignoring this underlying structure has implications when trying to draw inferences about the effects of covariates.\footnote{As shown in the simulation section, inference in a dyadic setting can become complicated when there are unobserved dependencies.} The fact that there is such a divergence in the performance of AME versus the original estimation procedures highlights that there are unobserved sources of bias in each of these studies.
+Ignoring this underlying structure has consequences for inferential analysis.\footnote{As shown in the simulation section, inference in a dyadic setting can become complicated when there are unobserved dependencies.} The fact that there is such a divergence in the performance of AME versus the original estimation procedures highlights that there are unobserved sources of bias in each of these studies.
 We hone in on the main finding of each study to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and re-estimated models are presented in the Appendix.}
 

### 2018-03-04 paper/4_app.tex
appAgain2

@@
 
-An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates.
+An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates or the differences between these covariates.
 These differences may change dramatically once interdependencies are taken into account.
@@
 
-Next we discuss each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.
+Next we discuss each of the replications in more detail and highlight the substantive insights drawn from the AME framework.
 
@@
 \citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.
-Their work contests prior scholarship that claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state \citep{peceny:etal:2002}.
+Their work contests prior scholarship claiming that interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state \citep{peceny:etal:2002}.
 Using a directed dyadic dataset of almost a million observations, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.
@@
 
-The key variables in the original model measure  whether or not the sender in the directed dyad is a personalist regime and the target a democratic (``Pers/Democ directed dyad"), and whether the sender is a democratic regime and the target personalist (``Democ/Personalist directed dyad'').
+The key variables in the original model measure whether or not the sender in the directed dyad is a personalist regime and the target a democratic regime (``Pers/Democ directed dyad") or whether the opposite is true (``Democ/Personalist directed dyad'').
 The authors find that the Pers/Democ Directed Dyad indicator is positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
-In our re-estimation using the AME framework, we confirm these results, indicating that dictators are more likely to initiate or be in conflict with democratic regimes but not vice versa.
+In our re-estimation using the AME framework, we confirm these results, indicating that dictators are more likely to initiate or engage in conflict with democratic regimes but not vice versa.
 
-Though we are able to confirm their results it is not the case that employing the AME model offers few benefits.
+Even though we are able to confirm the original results, employing the AME model offers clear benefits in this case.
 As already shown in Figure~\ref{fig:perf}, our approach performs notably better in reflecting the data generating process.
@@
 	\caption{Estimates of sender random effects ($a_{i}$) from AME for the Reiter \& Stam (2003) model.
-Positive values indicate that the particular country is more likely to be involved in conflict than predicted by the covariates in the model, and negative values indicate that the country is less likely to be involved in conflict.
-Left-panel shows $a_{i}$ estimates for all countries and right-panel highlights the top ten countries in terms of positive and values of $a_{i}$.}
+Positive values indicate that the particular country is more likely to be involved in conflict than predicted by the covariates in the model.
+Negative values indicate that the country is less likely to be involved in conflict.
+The left-panel shows $a_{i}$ estimates for all countries and right-panel highlights the top ten countries in terms of positive and values of $a_{i}$.}
 	\label{fig:reiter_stam_aEff}
@@
 
-The visualization of the sender random effects highlights that the behavior of a number of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
+The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
 Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than what the covariates would predict.
-Further other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than what the fixed effects in the model would predict.
+Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than what the fixed effects in the model would suggest.
 None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.
@@
 When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case.
-Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.
+Further, she argues that in personalistic regimes without a military or civilian domestic audience, the leaders are more likely to employ military force in their foreign policy.
 To study this question, Weeks employs a directed dyad design of conflict similar to that used by Reiter \& Stam.
@@
 The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
-The key finding of this work is that a) juntas, boss, and strongmen type regimes are more likely to initiate conflict than machine-type regimes and that b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strongmen are more conflict-prone.
+The key findings of her work are that a) juntas, boss, and strongmen regimes are more likely to initiate conflict than machine-type regimes and b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strongmen are more conflict-prone.
 When analyzing the results using AME, however, we find that the parameters on each of her autocratic regime type variables are too imprecisely measured to draw any inferences.
@@
 Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same fixed effects.
-As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure in the event of interest.
-To uncover some of this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$.
+As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest.
+To uncover this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$.
 These random effects account for higher order dependencies that manifest as a result of homophily and stochastic equivalence in dyadic data.
@@
 To visualize the multiplicative effects, we develop the circular diagram shown in Figure~\ref{fig:weekscirc}.
-The nodes throughout the diagram represent countries and they are colored according to their geographic position -- a legend is provided in the center.
-The outer ring visualizes higher order dependence patterns that countries have in terms of their sending relationships -- country positions here are estimated in the $\mathbf{u}_{i}$ random effects described in Equation~\ref{eqn:ame}.
-Countries that are more proximate to each other in this outer ring are more likely to send or initiate conflictual events with similar targets.
-The inner ring, based on estimates of $\mathbf{v}_{i}$, is constructed such that countries more proximate to each other are more likely to receive conflict from the same countries.
+The nodes throughout the diagram represent countries and are colored according to their geographic position -- a legend is provided in the center.
+The outer ring visualizes higher order dependence patterns through countries' sender relationships -- country positions here are estimated in the $\mathbf{u}_{i}$ random effects described in Equation~\ref{eqn:ame}.
+Countries that are more proximate to each other in this outer ring are more likely to send or initiate conflicts with similar targets.
+The inner ring, based on estimates of $\mathbf{v}_{i}$, is constructed such that countries closer together are more likely to receive conflict from the same sender countries.
 Last, the distance between countries in the inner and outer rings proportionately reflects who a country is more likely to be in conflict with.
 
-Interestingly, Figure~\ref{fig:weekscirc} reveals a number of interesting clusters.
+Interestingly, Figure~\ref{fig:weekscirc} reveals a number of clusters.
 For example, in the top left corner we see the US, UK, Germany, Canada, and Israel.
@@
 
-Perhaps most  importantly, an evaluation of this visualization highlights Weeks' expectation of how states behave in the conflict system.
+Perhaps most critically, an evaluation of this visualization highlights Weeks' original expectation of how states behave in the conflict system.
 Specifically, Iraq, Syria, Libya, and North Korea all fell under Weeks' ``boss" category, and each of these states tends to cluster together in the inner and outer rings.
-This indicates that even though we do not find support for Weeks' assertion that certain authoritarian regime types are more likely to initiate conflict, we do find that these regimes are more likely to behave similarly in terms of who they are more likely to both target  and receive conflict from.
+This indicates that even though we do not find support for Weeks' assertion that certain authoritarian regime types are more likely to initiate conflict, we do find that these regimes are more likely to behave similarly in terms of who they target and receive conflict from.
 
@@
 
-The last re-estimation we conduct with the AME model involves a study by \citet{gibler:2017}.
+The last replication we conduct with the AME model considers a study by \citet{gibler:2017}.
 Gibler argues that the long-standing relationship between the relative parity of capabilities and initiation of militarized interstate disputes (MIDs) is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-This finding calls into question many IR theories about the role of balance in terms of generating international conflict \citep{organski:1958}.
+This finding calls into question many IR theories about the role of balance in generating international conflict \citep{organski:1958}.
 
-To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for his primary variable: parity of the members of a dyad at the year in which they entered the international system.
-However, when we reestimate using AME, this measure becomes too imprecisely measured for us to draw any inferences about its effect on the initiation of MIDs.
+To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for his primary variable: parity of dyad members at the year in which they enter the international system.
+However, when we reestimate using AME, this measure becomes too imprecisely measured for using to draw any inferences about its effect on the initiation of MIDs.
 Not only is our estimate of the effect of Gibler's parity variable small, but it has a very large relative standard error---over a magnitude larger than the parameter itself.
@@
 Here we turn our focus to parameter interpretation in a substantive context.
-Specifically, even when the GLM and AME frameworks produce results that may seem to be in accordance with one another, the substantive interpretations of the effects of covariates can differ notably as the AME model uses a set of random effects to account for unobserved factors.
-To do this, we focus on the effect of rivalry on  MID initiation.
-Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects the two models provide  starkly contrast.
+Specifically, even when the GLM and AME frameworks produce results that may seem to be in accordance with one another, the substantive interpretations of the effects of covariates can differ notably since the AME model uses a set of random effects to account for unobserved factors.
+To explore this, we focus on the effect of rivalry on  MID initiation.
+Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects the two models contrast greatly.
 
@@
 As Figure~\ref{fig:gibmargeff} illustrates, the substantive results differ notably.
-First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower when taking interdependencies into account with the AME model.
+First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.
 These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
@@
 
-First, utilizing the AME framework enables scholars to better model the underlying structure often inherent to dyadic data.
+First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data.
 In each of the replications, the AME's predictive performance out-of-sample is notably superior.
@@
 Third, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights can emerge from the structure uncovered by the random effects of the AME framework.
-In particular, there is actual information about the dependencies so that clusters can be identified, and this information can also be used to generate new hypotheses.
-This kind of information is absent in standard approaches and adds to our ability to explain specific as well as general results.
+In particular, the AME provides information about the dependencies so that clusters can be identified.
+This information can then be used to generate new hypotheses.
+%These kinds of insights are absent in standard approaches and add to our ability to explain specific as well as general results.
 
 Fourth, it is evident that the actual results---not the estimated coefficients and their covariances---generated by the models differ greatly in expectations.
-This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to often give misleading results compared to the AME approach.
+This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to give misleading results compared to the AME approach.

### 2018-03-05 paper/1_intro.tex
getting closer
SM there are some issues you need to address, noted herein.  Let me know when you are finished, please.

@@
 We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model.
-Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads.
+Third-order dependencies capture relationships of transitivity, balance, and clusterability between different dyads.
 For example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
@@
 Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data.
-Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature.
-The comparison reveals that by accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature.
+Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature (published in this journal).
+Our comparison reveals that by accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables.
 Consequently, AME generates results that, at times, differ from those found in the original study as well as from the broader literature.
-Moreover, we demonstrate that the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies.
-Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.
+Moreover, we demonstrate that the AME latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data widely employed in IR.
+Finally, we show that for each replication our network-based approach provides substantially more accurate out-of-sample predictions than the models used in the original studies.
 
 Our AME approach advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
-Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena and account for the data generating process behind such events of interest.
+Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena without abandoning a regression based approach, but at the same time account for the data generating processes behind such events of interest.
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-03-05 paper/2_model.tex
getting closer
SM there are some issues you need to address, noted herein.  Let me know when you are finished, please.

@@
 The visualization illustrates that actors belonging to the same group have a higher likelihood of interacting with each other, whereas interactions across groups are rarer.
-A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualize linkages between right and left leaning political blogs preceding the 2004 United States Presidential Election.
-\citeauthor{adamic:glance:2005} find that the degree of interaction between right and left leaning blogs is minimal and that most blogs are linked to others that are politically similar.
+A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualize linkages between political blogs preceding the 2004 United States Presidential Election.
+\citeauthor{adamic:glance:2005} find that the degree of interaction between right- and left-leaning blogs is minimal and that most blogs are linked to others that are politically similar.
 This showcases the types of higher-order dependencies that can emerge in relational data.
@@
 These types of patterns frequently emerge in IR contexts as well.\footnote{For example, see: \citet{manger:etal:2012, kinne:2013, chyzh:2016}.} For example, a perennial finding in the interstate trade literature emphasizes the role that geography plays in determining trade flows.
-Geographic proximity in the network context is an example of homophily --- a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place.
+Geographic proximity in the network context is an example of homophily---a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place.
 Alternatively, in the interstate conflict literature, we may find that actors who are each a member of a particular (formal or informal) alliance are likely to act similarly in the conflict network.
-Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance --- an example of stochastic equivalence.
+Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance---an example of stochastic equivalence.
 In both these examples, we are able to explicitly parameterize the attribute that might explain the emergence of higher order dependence patterns.
@@
 where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver) at time $t$.
-We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
+We use a \textbf{Bayesian probit framework}, SHAHRYAR, WHY DO WE NEED A PROBIT FRAMEWORK AT THIS POINT?
+in which we model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
+$f$ is typically a mapping function, and can be one that applies to dichotomous, ordinal, or continuous distribitions.
+
 Next, to account for the dependencies that emerge in dyadic data that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM.
@@
 In the case of the SRRM, $\rho$ is constrained to be one and instead of separate sender and receiver random effects a single actor random effect is utilized.
-For the LFM, an eigen-decomposition scheme is used to capture higher order dependence patterns.
+For the LFM, an eigen-decomposition scheme is used to capture higher-order dependence patterns.
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.
@@
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
-Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
+\textbf{PUT here to end of paragraph IN APPENDIX}Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
 
@@
 
-The presence of these dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
-Of course, if one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.
+The presence of the kinds of dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events or processes that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
+If one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.
 However, it is rarely the case that this is possible even for TSCS data and thus modeling decisions must account for underlying structure.

### 2018-03-05 paper/3_sim.tex
getting closer
SM there are some issues you need to address, noted herein.  Let me know when you are finished, please.

@@
 The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Assume that the true data-generating process for a particular $Y$ is given by:
+Assume that the true data-generating process for a particular $Y$ is given by: \textbf{Shahryar: What distribution does y follow here?
+Needs to be specified.}
 
@@
 
+\textbf{SM: Color is not needed for this graphic.}
+
 \begin{figure}
@@
 
+\textbf{SM: Color is not needed for this graphic. Make the vertical line a shade of gray that is visible.}
+
 \begin{figure}

### 2018-03-05 paper/4_app.tex
getting closer
SM there are some issues you need to address, noted herein.  Let me know when you are finished, please.

@@
 
-We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017}.
+We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017} published in the \textit{American Political Science Review}.
 Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for the behavior of actors in the system.
@@
 Second, the results of AME estimation are interpretable alongside results using standard approaches.
-Third, through using this approach we can also quantify the degree to which first, second, and third order dependencies are present.
+Third, through using this approach we can also quantify the degree to which first-, second-, and third-order dependencies are present.
 
 We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
-Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential.
+Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, is well-known and posited an hypothesis in which interdependencies are consequential.
 Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model.
@@
 
-For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' neatly assembled replication scripts.} Next, utilizing the same covariate specifications, we estimated the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
+For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimated the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
 Results with alternative values of $K$ are similar.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
@@
 For each study  we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
-Then for each $s \in \{1,\ldots,k\}$, we:
-
-\begin{enumerate}
-	\item estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$,
-	\item and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters. 
-\end{enumerate}
+Then for each $s \in \{1,\ldots,k\}$, we estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$, and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters.
 
@@
 Models that have a better fit according to this metric will follow the upper-left corner border of the ROC space.
-In addition to ROC curves we also use separation plots.
+In addition to ROC curves we also use separation plots \citep{greenhill:etal:2011}.
 Separation plots plots visualize each of the observations in the dataset according to their predicted probability.
@@
 We hone in on the main finding of each study to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates only that the sign and/or significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and re-estimated models are presented in the Appendix.}
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates that the putative statistical significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and re-estimated models are presented in the Appendix.}
 
@@
 The key variables in the original model measure whether or not the sender in the directed dyad is a personalist regime and the target a democratic regime (``Pers/Democ directed dyad") or whether the opposite is true (``Democ/Personalist directed dyad'').
-The authors find that the Pers/Democ Directed Dyad indicator is positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
+The authors find that coefficient of the Pers/Democ Directed Dyad indicator is positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
 In our re-estimation using the AME framework, we confirm these results, indicating that dictators are more likely to initiate or engage in conflict with democratic regimes but not vice versa.
@@
 The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
-Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than what the covariates would predict.
-Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than what the fixed effects in the model would suggest.
+Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
+Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict the fixed effects in the model would suggest.
 None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.
@@
 
-\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
+\nocite{weeks:2012} Weeks examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.
 She argues that in some circumstances autocrats are held accountable for their foreign policy decisions, and that this is dependent on the audiences of autocrats.
@@
 
-The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
-The key findings of her work are that a) juntas, boss, and strongmen regimes are more likely to initiate conflict than machine-type regimes and b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses, and Strongmen are more conflict-prone.
-When analyzing the results using AME, however, we find that the parameters on each of her autocratic regime type variables are too imprecisely measured to draw any inferences.
-Consequentially, none of the findings from her original analysis are confirmed once accounting for dependencies.
+The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, following \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
+The key findings of her work are that a) juntas, boss, and strongmen regimes are more likely to initiate conflict than machine-type regimes and b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas juntas, bosses, and strongmen are more conflict-prone.
+When analyzing the results using AME, however, we find that the parameters on each of her autocratic regime type variables are too imprecisely measured to draw any inference about their putative causal effects.
+Consequentially, none of the findings from her original analysis are confirmed once known dependencies among the data are taken into account via AME.
 
@@
 
-To visualize the multiplicative effects, we develop the circular diagram shown in Figure~\ref{fig:weekscirc}.
+To visualize the multiplicative effects, we display the circular diagram shown in Figure~\ref{fig:weekscirc}.
 The nodes throughout the diagram represent countries and are colored according to their geographic position -- a legend is provided in the center.
@@
 First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.
-These are rare events, so the probabilities are low, but the difference is a factor of almost 2.
+These are rare events, so the probabilities are low, but the difference is a factor of almost two.
 Thus, we find substantially different expected values from these two models.
@@
 
-First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data.
+First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data..
 In each of the replications, the AME's predictive performance out-of-sample is notably superior.

### 2018-03-05 paper/netsMatter.tex
getting closer
SM there are some issues you need to address, noted herein.  Let me know when you are finished, please.

@@
 \begin{abstract}
-	Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
-Yet, modeling approaches that fail to properly capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors.
-We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the dependencies in dyadic data.
-First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference.
-Second, we compare the AME model to approaches used in three studies from recent international relations scholarship.
-For each study, we find that the modeling approach used in the original study performs notably worse at capturing the data generating process compared to the AME.
+	Much international relations scholarship concerns dyads.
+Dyadic hypotheses and especially dyadic data abound.
+Standard mnodeling approaches fail to adequately capture the data generating process behind dyadic events and processes.
+As a result, they suffer from biased coefficients and poorly calibrated standard errors.
+We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the inherent dependencies in dyadic data.
+First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference about dyadic hypotheses.
+Second, we compare the AME model to approaches used in three prominent studies from recent international relations scholarship.
+For each study, we find that compared to AME, the modeling approach used in each original study performs notably worse at capturing the data generating process.
 Further, conventional methods misstate the effect of key variables and the uncertainty in these effects.
+Finally, AME dominates standard approaches in terms of out-of-sample fit.
 In sum, our work shows the consequences of failing to take the dependencies inherent to dyadic data seriously.
@@
 %\bibliography{/Users/s7m/whistle/master}
- % \bibliography{/Users/mdw/Git/whistle/master}
- \bibliography{/Users/cassydorff/ProjectsGit/master}
+ \bibliography{/Users/mdw/Git/whistle/master}
+% \bibliography{/Users/cassydorff/ProjectsGit/master}
 % \bibliographystyle{authordate1}

### 2018-03-06 paper/2_model.tex
mikes qs

@@
 
+% We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
 where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver) at time $t$.
-We use a \textbf{Bayesian probit framework}, SHAHRYAR, WHY DO WE NEED A PROBIT FRAMEWORK AT THIS POINT?
-in which we model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
+We model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
 $f$ is typically a mapping function, and can be one that applies to dichotomous, ordinal, or continuous distribitions.

### 2018-03-06 paper/3_sim.tex
mikes qs

--- 2018-03-03 paper/3_sim.tex
@@
 The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Assume that the true data-generating process for a particular $Y$ is given by:
+Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}
 
@@
 
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
+where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix generated, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
 $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.

### 2018-03-06 paper/4_app.tex
mikes qs

@@
 
-The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, following \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects.
+The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, following \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with dyad clustered standard errors.
 The key findings of her work are that a) juntas, boss, and strongmen regimes are more likely to initiate conflict than machine-type regimes and b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas juntas, bosses, and strongmen are more conflict-prone.

### 2018-03-07 paper/2_model.tex
sync up

@@
 
+% \textbf{PUT here to end of paragraph IN APPENDIX}
 By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the underlying structure in dyadic data that, if left un-estimated, would complicate any inferences we might wish to draw for the exogenous parameters.
 Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
-\textbf{PUT here to end of paragraph IN APPENDIX}Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
+Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
 
@@
 
-The presence of the kinds of dependencies in relational data points to the fact that there is a complex structure underlying the dyadic events or processes that we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
+Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
 If one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.

### 2018-03-08 paper/2_model.tex
incorp mike suggestions

@@
 $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers.
-The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model the singular values are not restricted to be positive, thus allowing us to account for both positive and negative homophily.}
-
-Note that this model easily generalizes to the case, common in IR, where interactions are undirected (for example the presence of conflict or a bilateral investment treaty).
+The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model the singular values are not restricted to be positive, thus allowing us to account for both positive and negative homophily.} Note that this model easily generalizes to the case, common in IR, where interactions are undirected (for example the presence of conflict or a bilateral investment treaty).
 In the case of the SRRM, $\rho$ is constrained to be one and instead of separate sender and receiver random effects a single actor random effect is utilized.
@@
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.
+\footnote{Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
+The algorithm is provided in the Appendix.}
 
-% \textbf{PUT here to end of paragraph IN APPENDIX}
-By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the underlying structure in dyadic data that, if left un-estimated, would complicate any inferences we might wish to draw for the exogenous parameters.
-Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
-Specifically, given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
-
- \begin{itemize}
- 	\item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 	\item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 	\item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-	\item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
- 	\item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
- 	\item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
- 	\item For each $k \in K$:
- 	\begin{itemize}
- 		\item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
- 		\item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
-}
- 	\end{itemize}
- \end{itemize}
+ % in dyadic data that, if left un-estimated, would complicate any inferences we might wish to draw for the exogenous parameters.
+By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the types of underlying structure discussed in this section.
 

### 2018-03-08 paper/6_appendix.tex
sync up
incorp mike suggestions

@@
 
+\subsection*{AME Monte Carlo Markov Chain Algorithm}
+
+Given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
+
+ \begin{itemize}
+  \item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+  \item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+  \item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+  \item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
+  \item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
+  \item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
+  \item For each $k \in K$:
+  \begin{itemize}
+    \item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+    \item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
+    \item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
+}
+  \end{itemize}
+ \end{itemize}
+
 \subsection*{Additional Replication Information}
 
-For each of the replications involving a binary dependent variable we provide a table of coefficient estimates that includes the original GLM estimation with a logit link, a GLM estimation with a probit link, and the AME model.
-The GLM estimation with a probit link function is provided so as to ease comparison between the AME model, which is also based on the probit link.
-%For Rose (2004) we just provide a table of coefficient estimates that includes the original OLS estimation and the AME model.
+For each of the replications involving a binary dependent variable we provide a table of coefficient estimates that includes the original GLM estimation and the results from the AME model.
 

### 2018-03-10 paper/5_concl.tex
prepping for the end
possible conclusion section added, still incomplete

@@
 
+\subsection*{Are There Other Alternatives?}
+
+Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014}, which proposes a variational estimation approach to latent network models.}
+None of these approaches has as yet been widely employed nor even cited in the substantive literature, and each is mainly known inside of the network methodology community.
+We discuss each of these approaches and briefly compare them to the AME approach.
+
+\subsubsection*{Cluster-Robust Variance Estimation}
+
+\cite{aronow:etal:2015} discussed the problem of estimation of general linear models in the absence of $\simiid$ validity.
+In particular, they look at the so-called clustering in dyadic data.
+This is the clustering of errors that occur because multiple dyads share individual members.
+They develop a sandwich style estimator which leaves untouched the parameters, opting instead for a nonparametric approach which attempts to correct the variance estimates across dyads.
+They do not compare their approach to AME or other strategies to explicitly estimate parametric effects in networks.
+Instead, they examine a spatial lag approach and general mixed effects models that abide by the $\simiid$ assumption.
+Their basic equation is
+\begin{eqnarray*}
+Y_{ij} = \mathbf{X}_{ij}\intercal \beta + \varepsilon_{ij}
+\end{eqnarray*}
+and the important parameter is $\beta$, defined in the standard way, via an appropriate estimator (they use OLS).
+They assert that a random sampling of dyads that do not share a member will be uncorrelated, i.e., $\text{Cov}(\varepsilon_d,\varepsilon_{d'}) \equiv 0$, where $d'$ is the set of dyads that do not share members.
+Following this leads to the conclusion that the variance in the full sample of the dyadic, independent data and the dyadic error contains a known bias, part of which is a dyad specific variance component and another part is due to repeated, shared memberships in dyads.
+$\hat{\beta}$ is shown to be consistent, i.e., that as the number of \emph{dyads} increase to infinity, it converges in probability to its true value.
+
+The variance calculations used by Aronow et al basically decompose the variance into three parts, the cluster robust variance (all dyads with unit $i$, assuming all others to be 
+$\simiid$), the separate cluster robust variance of the repeated dyads, and clustered robust variance of all dyads, assuming they are all independent.
+Using simulation, they demonstrate that as sample size increases, this formula for calculating the variance-covariance matrix will decline as the number of nodes in the network increase and approach its true value.
+
+This approach focuses on the standard errors, needed for significance testing, but ignores the bias introduced into the regression coefficients, as discussed above.
+Nor does it provide any way to access the higher-order network effects.
+It simply provides a fix to the estimated standard errors.
+One of the main problems with dependent data, is that it takes more data to calculate basic quantities accurately.
+This is true even for a simple t-test \citep{ward:gleditsch:2008}.
+
+However, more generally how does cluster-robust vriance estimation perform compared for example to AME?  XYZ.
+
+%Isn't brandon stewart overview a better place to start?
+
+\subsubsection*{Choice Based Sampling}
+
+Poast \citeyear{poast:2010,poast:2016} doesn't really offer a way to estimate dyadic data in the presence of independence,
+``I am not presenting an alternative procedure for modeling such spatial, strategic, temporal,
+or monadic interdependencies in the data.'' Instead, he argues that multilateral events are poorly represented or misrepresented by dyadic data.
+Thus, he is arguing against dyadic data.
+He creates an artificial data that is triadic and generates a process that combines the capabilities of the three actors in each triad into a ratio that drives whether an alliance is formed among all three actors.
+Then these triadic data are converted into a dyadic format.
+He then uses a logit regression framework to show that with the omniscient data in the triadic case logit regression can recover the true value of the data generating process, but logit, as applied to the dyadic construction of the triadic data, produces estimates that are inflated.
+
+Poast's solution to this is called choice-based sampling, ``[t]he sampling method entails constructing a data set containing all observations for which the dependent variable is coded with a positive value, along with a random sample of observations for which the dependent variable is coded 0'' (page 410).
+This new sample is then estimated with a rare-events logit regression.
+His results (page 412) show that their is a small bias (about $1\%$) using the dyadic approach compared to the alternatives he proposes.
+As one moves from triadic to \textit{5-adic}, the divergence from the analysis of the dyadic data becomes substantially larger.
+But Poast doesn't address the underlying issue of how to properly estimate dependent data, but raises the question whether dyadic data are capable of representing multi-adic data generating processes.
+Unfortunately, adjudicating this with an approach (logit regression) that it itself known to not be capable of capturing the higher-order aspects of the data generating process is, itself, problematic on first principles.
+
+However, how does Poast's approach compare to AME?  XX, YY, ZZ.
+
+%see also Poast, Paul. ``Dyads are Dead, Long Live Dyads: The Limits (but not Rejection) of Dyadic Designs in International #Relations Research.'' International Studies Quarterly. June 2016: 369-374
+%and
+%Fordham, Benjamin and Poast, Paul. ``All Alliances are Multilateral: Rethinking Alliance Formation in the International #System.'' Journal of Conflict Resolution. August 2016: 840-865.
+%but note: Below are files for replicating or extending results in the paper ``All Alliances are Multilateral'' (w/ Benjamin #Fordham).
+We thank Daina Chiba (Essex), whose effort to replicate our work uncovered the difficulties scholars face (myself #included!) when creating k-adic data and, therefore, led to this webpage as a help to future scholarship.
+% also used by Brandon Kinne
+
+%maybe used by Peaceful Uncertainty: When Power Shocks Do Not Create Commitment Problems
+%Muhammet A Bas  Robert Schub ISQ 2017
+
+\subsubsection*{Multiple Regression Quadratic Assignment Procedure}
+
+\citet{erikson:etal:2014} base their approach on MRQAP developed by Krackhardt (1988), \nocite{krackhart:1988} an approach from network studies that draws upon the fact that networks have repeated observations, which induces some correlation of observations that in turn lead to bias in standard errors of general linear models of these data.
+MRQAP randomly permutes the rows and columns of the data, creating multiple datasets which are then each analyzed in the standard way to generate a sampling distribution of all possible (but not observed) networks from the observed data.
+Herein, the data are randomly permuted on the assumption that the errors are exchangeable.
+Many permutations are created and models are run on all of them, then the distributions of tests statistics (not parameters) are examined.
+They find in their reanalysis that ``the randomization results show that conventional t-tests are inappropriate for
+testing hypotheses on dyad-year data because they rely on highly overconfident standard errors'' (page 462).
+This shows that in dyadic data analysis with general linear models, especially with hundreds of thousands of observations, the standard t-test is polyannaish in finding significant relations.
+One simple fix is MRQAP, which address---again---the standard errors.
+But this approach doesn't really get at the underlying data generating process, and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
+It is a fix of a leaky window, when the roof is collapsed.
+(okay, too much, but hey.)
+
+How does MRQAP compare to AME?  XXX, YYY, ZZZ
+
+\subsection*{Summary}
 International relations is about the interactions, relationships, and dependencies among countries or other important international actors.

### 2018-03-16 paper/5_concl.tex
some changes to concl

@@
 We discuss each of these approaches and briefly compare them to the AME approach.
+The takeaway is that the AME approach enables users to more accurately conduct inference on dependent data than these other approaches which have been proposed.
 
@@
 This approach focuses on the standard errors, needed for significance testing, but ignores the bias introduced into the regression coefficients, as discussed above.
-Nor does it provide any way to access the higher-order network effects.
-It simply provides a fix to the estimated standard errors.
-One of the main problems with dependent data, is that it takes more data to calculate basic quantities accurately.
-This is true even for a simple t-test \citep{ward:gleditsch:2008}.
+Nor does it provide any way to access the higher-order network effects, a point that Aronow et alia make in their article, noting that it is only targeted at dyadic (not higher level) dependencies.
+This approach simply provides a fix to the estimated standard errors.
+%One of the main problems with dependent data is that it takes more data to calculate basic quantities accurately. This is true even for a simple t-test \citep{ward:gleditsch:2008}.
 
-However, more generally how does cluster-robust vriance estimation perform compared for example to AME?  XYZ.
+Thus, cluster-robust vriance estimation is not designed to answer the same questions as AME.
+As shown in the original article, it does provide a way to  estimate standard errors (in a sitution with only dyadic dependence) that substantially improves over the standard logistic regression approach.
+However, it ignores the higher-order dependencies which are the focus of the AME model.
 
@@
 
-However, how does Poast's approach compare to AME?  XX, YY, ZZ.
-
-%see also Poast, Paul. ``Dyads are Dead, Long Live Dyads: The Limits (but not Rejection) of Dyadic Designs in International #Relations Research.'' International Studies Quarterly. June 2016: 369-374
-%and
-%Fordham, Benjamin and Poast, Paul. ``All Alliances are Multilateral: Rethinking Alliance Formation in the International #System.'' Journal of Conflict Resolution. August 2016: 840-865.
-%but note: Below are files for replicating or extending results in the paper ``All Alliances are Multilateral'' (w/ Benjamin #Fordham).
-We thank Daina Chiba (Essex), whose effort to replicate our work uncovered the difficulties scholars face (myself #included!) when creating k-adic data and, therefore, led to this webpage as a help to future scholarship.
-% also used by Brandon Kinne
-
-%maybe used by Peaceful Uncertainty: When Power Shocks Do Not Create Commitment Problems
-%Muhammet A Bas  Robert Schub ISQ 2017
+However, how does Poast's approach compare to AME?
+We found in simulations structured like those reported herein run with the k-adic model that the k-adic model was not able to improve over AME, and in fact were dramatically worse at recovering the real data generating process than AME models.
 
@@
 This shows that in dyadic data analysis with general linear models, especially with hundreds of thousands of observations, the standard t-test is polyannaish in finding significant relations.
-One simple fix is MRQAP, which address---again---the standard errors.
+One simple fix is MRQAP, which addresses---again---the standard errors.
 But this approach doesn't really get at the underlying data generating process, and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
-It is a fix of a leaky window, when the roof is collapsed.
-(okay, too much, but hey.)
-
-How does MRQAP compare to AME?  XXX, YYY, ZZZ
+In our simulations when applied to a MRQAP model, we find that it does not recover the data generating process as well as AME.
+This is quite due in part to the fact that it treats the dependencies as nuisance to be eliminated rather than focused on explicitly as with AME.
 

### 2018-03-18 paper/5_concl.tex
what I hope are the FFC
Final Frikking Changes.

(I should know better than use the word "final"!)

@@
 
-\subsection*{Are There Other Alternatives?}
-
+We have shown above that the AME framework for modeling dependent data dominates the standard approaches which use some form of a general linear model---typically a logit---to analyze dyadic data, especially in international relations.
+But are there any alternatives that have been developed?
 Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014}, which proposes a variational estimation approach to latent network models.}
@@
 We discuss each of these approaches and briefly compare them to the AME approach.
-The takeaway is that the AME approach enables users to more accurately conduct inference on dependent data than these other approaches which have been proposed.
+The takeaway is that based on simulation evidence, the AME approach also dominates these approaches in terms of recovering the data generating process and thus AME enables users to more accurately conduct inference on dependent data.\footnote{The detailed results are not presented herein.}
 
-\subsubsection*{Cluster-Robust Variance Estimation}
+\subsection*{Cluster-Robust Variance Estimation}
 
@@
 \begin{eqnarray*}
-Y_{ij} = \mathbf{X}_{ij}\intercal \beta + \varepsilon_{ij}
+Y_{ij} = \mathbf{X}_{ij}^\intercal \beta + \varepsilon_{ij}
 \end{eqnarray*}
@@
 This approach focuses on the standard errors, needed for significance testing, but ignores the bias introduced into the regression coefficients, as discussed above.
-Nor does it provide any way to access the higher-order network effects, a point that Aronow et alia make in their article, noting that it is only targeted at dyadic (not higher level) dependencies.
-This approach simply provides a fix to the estimated standard errors.
+Nor does it provide any way to access the higher-order network effects, a point that Aronow et alia make in their article, noting that it is only targeted at first-order, dyadic (not higher level) dependencies.
+The cluster robust variance approach simply provides a fix to the estimated standard errors.
+In this sense, it treats the dependent data as having dependencies that are a nuisance, in the statistical sense.
 %One of the main problems with dependent data is that it takes more data to calculate basic quantities accurately. This is true even for a simple t-test \citep{ward:gleditsch:2008}.
 
-Thus, cluster-robust vriance estimation is not designed to answer the same questions as AME.
+Cluster-robust vriance estimation is not designed to answer the same questions as AME.
 As shown in the original article, it does provide a way to  estimate standard errors (in a sitution with only dyadic dependence) that substantially improves over the standard logistic regression approach.
@@
 
-\subsubsection*{Choice Based Sampling}
+\subsection*{Choice Based Sampling}
 
@@
 or monadic interdependencies in the data.'' Instead, he argues that multilateral events are poorly represented or misrepresented by dyadic data.
-Thus, he is arguing against dyadic data.
-He creates an artificial data that is triadic and generates a process that combines the capabilities of the three actors in each triad into a ratio that drives whether an alliance is formed among all three actors.
+In a sense, he is arguing against dyadic data.
+He creates an artificial database that is triadic and from it generates a process that combines the capabilities of the three actors in each triad into a ratio that drives whether an alliance is formed among all three actors.
 Then these triadic data are converted into a dyadic format.
-He then uses a logit regression framework to show that with the omniscient data in the triadic case logit regression can recover the true value of the data generating process, but logit, as applied to the dyadic construction of the triadic data, produces estimates that are inflated.
+He then uses a logit regression framework to show that with the omniscient data in the triadic case, logit regression can recover the true value of the data generating process, but logit, as applied to the dyadic construction of the triadic data, produces estimates that are inflated.
 
@@
 As one moves from triadic to \textit{5-adic}, the divergence from the analysis of the dyadic data becomes substantially larger.
-But Poast doesn't address the underlying issue of how to properly estimate dependent data, but raises the question whether dyadic data are capable of representing multi-adic data generating processes.
-Unfortunately, adjudicating this with an approach (logit regression) that it itself known to not be capable of capturing the higher-order aspects of the data generating process is, itself, problematic on first principles.
+Poast doesn't address the underlying issue of how to properly estimate dependent data, but raises the question whether dyadic data are capable of representing multi-adic data generating processes.
+Unfortunately, adjudicating this with an approach (logit regression) that itself is known to be incapable of capturing the higher-order aspects of the data generating process is, itself, problematic on first principles.
 
 However, how does Poast's approach compare to AME?
-We found in simulations structured like those reported herein run with the k-adic model that the k-adic model was not able to improve over AME, and in fact were dramatically worse at recovering the real data generating process than AME models.
+We found in simulations structured like those reported herein run with the k-adic model that the k-adic model was dramatically worse at recovering the real data generating process than AME.
 
-\subsubsection*{Multiple Regression Quadratic Assignment Procedure}
+\subsection*{Multiple Regression Quadratic Assignment Procedure}
 
-\citet{erikson:etal:2014} base their approach on MRQAP developed by Krackhardt (1988), \nocite{krackhart:1988} an approach from network studies that draws upon the fact that networks have repeated observations, which induces some correlation of observations that in turn lead to bias in standard errors of general linear models of these data.
+\citet{erikson:pinto:2014} base their approach on MRQAP developed by 
+Krackhardt (1988), \nocite{krackhart:1988} an approach from network studies that draws upon the fact that networks have repeated observations, which induces some correlation of observations leading to bias in standard errors of general linear models of these data.
 MRQAP randomly permutes the rows and columns of the data, creating multiple datasets which are then each analyzed in the standard way to generate a sampling distribution of all possible (but not observed) networks from the observed data.
@@
 One simple fix is MRQAP, which addresses---again---the standard errors.
-But this approach doesn't really get at the underlying data generating process, and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
-In our simulations when applied to a MRQAP model, we find that it does not recover the data generating process as well as AME.
+This approach does not focus on the underlying data generating process, and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
+In our simulations of the MRQAP model, we find that it does not recover the data generating process as well as AME.
 This is quite due in part to the fact that it treats the dependencies as nuisance to be eliminated rather than focused on explicitly as with AME.

### 2018-03-20 paper/6_appendix.tex
I am fucking tired of this paper. Please put it on an editor's desk.
for real

@@
 
-\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\thefigure}{\arabic{figure}}
 \setcounter{figure}{0}
-\renewcommand{\thetable}{A.\arabic{table}}
+\renewcommand{\thetable}{\thesection \arabic{table}}
 \setcounter{table}{0}
-\renewcommand{\thesection}{A.\arabic{section}}
+%\renewcommand{\thesection}{A.\arabic{section}}
 \setcounter{section}{0}
 
-\section*{\textbf{Appendix}}
-
-\subsection*{AME Monte Carlo Markov Chain Algorithm}
+\appendix
+\section{AME Monte Carlo Markov Chain Algorithm}
 
@@
  \end{itemize}
-
-\subsection*{Additional Replication Information}
+\newpage
+\section{Additional Replication Information}
 
 For each of the replications involving a binary dependent variable we provide a table of coefficient estimates that includes the original GLM estimation and the results from the AME model.
+We have adopted for presentation purposes only the use of significance testing on these observational data for facile comparison with the replicated studies.
+However, note that we do provide predictive heuristics for these models as well.
 
@@
 \clearpage
-\subsubsection*{Reiter \& Stam (2003)}
+\subsection*{Reiter \& Stam (2003)}
 
@@
 
-\subsubsection*{Weeks (2012)}
+\subsection*{Weeks (2012)}
 
@@
 
-\subsubsection*{Gibler (2017)}
+\subsection*{Gibler (2017)}
 
@@
 
-\subsection*{AME Tutorial}
+\section{AME Tutorial}
 

### 2018-03-20 paper/netsMatter.tex
mikes qs
typo fix
prepping for the end
possible conclusion section added, still incomplete
I am fucking tired of this paper. Please put it on an editor's desk.
for real

@@
 % % add watermark
-% \usepackage{draftwatermark}
-% \SetWatermarkText{Version 1.0}
-% \SetWatermarkScale{.5}
-% \SetWatermarkColor[gray]{0.9}
+\usepackage{draftwatermark}
+\SetWatermarkText{Submission Copy}
+\SetWatermarkScale{.5}
+\SetWatermarkColor[gray]{0.9}
 
@@
 	pdfauthor={Shahryar Minhas, et alia.},%
-	pdftitle={Networks Matter},%
+	pdftitle={Taking Dyads Seriously},%
 	colorlinks,%
@@
 
-% \thanks{Shahryar Minhas and Michael D.
+%\thanks{Shahryar Minhas and Michael D.
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
 John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments.
-We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
+We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.
+We benefited from presentations of this material at the 2017 Political Networks Conference, the 2017 Political Methodology Conference, and our home institutions.}
 
@@
 Dyadic hypotheses and especially dyadic data abound.
-Standard mnodeling approaches fail to adequately capture the data generating process behind dyadic events and processes.
+Standard modeling approaches fail to adequately capture the data generating process behind dyadic events and processes.
 As a result, they suffer from biased coefficients and poorly calibrated standard errors.
@@
 \input{5_concl}
+\appendix
 \input{6_appendix}

### 2018-03-20 paper/5_concl.tex
I am fucking tired of this paper. Please put it on an editor's desk.
for real

@@
 But are there any alternatives that have been developed?
-Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014}, which proposes a variational estimation approach to latent network models.}
+Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014} who proposes a variational estimation approach to latent network models.}
 None of these approaches has as yet been widely employed nor even cited in the substantive literature, and each is mainly known inside of the network methodology community.
@@
 \cite{aronow:etal:2015} discussed the problem of estimation of general linear models in the absence of $\simiid$ validity.
-In particular, they look at the so-called clustering in dyadic data.
-This is the clustering of errors that occur because multiple dyads share individual members.
-They develop a sandwich style estimator which leaves untouched the parameters, opting instead for a nonparametric approach which attempts to correct the variance estimates across dyads.
+In particular, they look at the clustering of errors that occur in dyadic data because multiple dyads share individual members.
+They develop a sandwich style estimator which leaves untouched the parameters, opting instead for a nonparametric approach which attempts to correct the variance-covariance estimates of the estimated parameters, leaving the parameters untouched.
 They do not compare their approach to AME or other strategies to explicitly estimate parametric effects in networks.
@@
 \end{eqnarray*}
-and the important parameter is $\beta$, defined in the standard way, via an appropriate estimator (they use OLS).
+and the important parameter is $\beta$, defined in the standard way, via an appropriate estimator---they use OLS.
 They assert that a random sampling of dyads that do not share a member will be uncorrelated, i.e., $\text{Cov}(\varepsilon_d,\varepsilon_{d'}) \equiv 0$, where $d'$ is the set of dyads that do not share members.
-Following this leads to the conclusion that the variance in the full sample of the dyadic, independent data and the dyadic error contains a known bias, part of which is a dyad specific variance component and another part is due to repeated, shared memberships in dyads.
+This leads to the conclusion that the variance in the full sample of the dyadic, independent data and the dyadic error contains a known bias, part of which is a dyad specific variance component and another part is due to repeated, shared memberships in dyads.
 $\hat{\beta}$ is shown to be consistent, i.e., that as the number of \emph{dyads} increase to infinity, it converges in probability to its true value.
 
-The variance calculations used by Aronow et al basically decompose the variance into three parts, the cluster robust variance (all dyads with unit $i$, assuming all others to be 
-$\simiid$), the separate cluster robust variance of the repeated dyads, and clustered robust variance of all dyads, assuming they are all independent.
+The variance calculations used by Aronow \textit{et alia} basically decompose the variance into three parts, the cluster-robust variance (all dyads with unit $i$, assuming all others to be $\simiid$), the separate cluster-robust variance of the repeated dyads, and clustered-robust variance of all dyads, assuming they are all independent.
 Using simulation, they demonstrate that as sample size increases, this formula for calculating the variance-covariance matrix will decline as the number of nodes in the network increase and approach its true value.
 
-This approach focuses on the standard errors, needed for significance testing, but ignores the bias introduced into the regression coefficients, as discussed above.
+This approach focuses on the standard errors (the diagonal of the variance-covariance matrix) needed for significance testing, but ignores the bias introduced into the regression coefficients by ignoring dependence among the data, as discussed above.
 Nor does it provide any way to access the higher-order network effects, a point that Aronow et alia make in their article, noting that it is only targeted at first-order, dyadic (not higher level) dependencies.
-The cluster robust variance approach simply provides a fix to the estimated standard errors.
-In this sense, it treats the dependent data as having dependencies that are a nuisance, in the statistical sense.
+The cluster-robust variance approach simply provides a fix to the estimated standard errors, much like is presumed to work for many sandwich estimators of variance-covariance matrices.
+In this sense, it treats the dependent data as having dependencies that are a nuisance that only show up in the variance-covariance matrix, in the statistical sense.
 %One of the main problems with dependent data is that it takes more data to calculate basic quantities accurately. This is true even for a simple t-test \citep{ward:gleditsch:2008}.
 
-Cluster-robust vriance estimation is not designed to answer the same questions as AME.
-As shown in the original article, it does provide a way to  estimate standard errors (in a sitution with only dyadic dependence) that substantially improves over the standard logistic regression approach.
-However, it ignores the higher-order dependencies which are the focus of the AME model.
+Cluster-robust variance estimation is not designed to answer the same questions as AME.
+As shown in the original article, it does provide a way to  estimate standard errors (in a sitution with only dyadic dependence) that substantially improves over the standard regression approach.
+However, it ignores the higher-order dependencies which are the focus of the AME model and many network analyses.
 
@@
 
-\subsection*{Choice Based Sampling}
+\subsection*{Choice-Based Sampling}
 
@@
 ``I am not presenting an alternative procedure for modeling such spatial, strategic, temporal,
-or monadic interdependencies in the data.'' Instead, he argues that multilateral events are poorly represented or misrepresented by dyadic data.
+or monadic interdependencies in the data'' (2010, page 404, ironically).
+Instead, he argues that multilateral events are poorly represented or misrepresented by dyadic data.
 In a sense, he is arguing against dyadic data.
 He creates an artificial database that is triadic and from it generates a process that combines the capabilities of the three actors in each triad into a ratio that drives whether an alliance is formed among all three actors.
-Then these triadic data are converted into a dyadic format.
+These triadic data are converted into a dyadic format.
 He then uses a logit regression framework to show that with the omniscient data in the triadic case, logit regression can recover the true value of the data generating process, but logit, as applied to the dyadic construction of the triadic data, produces estimates that are inflated.
+Thus, Poast's approach is about how the data are constructed in the first instance.
+However, we generally do not have multiadic data, so some strategy is needed to transform the data we do have, i.e., dyadic data.
 
-Poast's solution to this is called choice-based sampling, ``[t]he sampling method entails constructing a data set containing all observations for which the dependent variable is coded with a positive value, along with a random sample of observations for which the dependent variable is coded 0'' (page 410).
+Poast's solution is called choice-based sampling, ``[t]he sampling method entails constructing a data set containing all observations for which the dependent variable is coded with a positive value, along with a random sample of observations for which the dependent variable is coded 0'' (2010, page 410).
 This new sample is then estimated with a rare-events logit regression.
-His results (page 412) show that their is a small bias (about $1\%$) using the dyadic approach compared to the alternatives he proposes.
+His results (2010, page 412) show that there is a small bias (about $1\%$) using the dyadic approach compared to the alternatives he proposes.
 As one moves from triadic to \textit{5-adic}, the divergence from the analysis of the dyadic data becomes substantially larger.
-Poast doesn't address the underlying issue of how to properly estimate dependent data, but raises the question whether dyadic data are capable of representing multi-adic data generating processes.
+Poast does not address the underlying issue of how to properly estimate dependent data, but raises the question whether dyadic data are capable of representing multi-adic data generating processes.
 Unfortunately, adjudicating this with an approach (logit regression) that itself is known to be incapable of capturing the higher-order aspects of the data generating process is, itself, problematic on first principles.
@@
 However, how does Poast's approach compare to AME?
-We found in simulations structured like those reported herein run with the k-adic model that the k-adic model was dramatically worse at recovering the real data generating process than AME.
+We found in simulations structured like those reported herein run with the k-adic model that the k-adic model was dramatically worse at recovering the real data generating process compared to AME.
+This is because the AME approach captures the higher-order dependencies in dyadic data.
+It is these dependencies that Poast is trying to get at with his k-adic sampling approach.
 
@@
 
-\citet{erikson:pinto:2014} base their approach on MRQAP developed by 
-Krackhardt (1988), \nocite{krackhart:1988} an approach from network studies that draws upon the fact that networks have repeated observations, which induces some correlation of observations leading to bias in standard errors of general linear models of these data.
+\citet{erikson:pinto:2014} base their approach on the MRQAP model developed by 
+Krackhardt (1988), \nocite{krackhart:1988} an approach from network studies that draws upon the fact that networks have repeated observations.
+This induces correlation of observations thereby leading to bias in standard errors in general linear models.
 MRQAP randomly permutes the rows and columns of the data, creating multiple datasets which are then each analyzed in the standard way to generate a sampling distribution of all possible (but not observed) networks from the observed data.
@@
 Many permutations are created and models are run on all of them, then the distributions of tests statistics (not parameters) are examined.
-They find in their reanalysis that ``the randomization results show that conventional t-tests are inappropriate for
-testing hypotheses on dyad-year data because they rely on highly overconfident standard errors'' (page 462).
-This shows that in dyadic data analysis with general linear models, especially with hundreds of thousands of observations, the standard t-test is polyannaish in finding significant relations.
-One simple fix is MRQAP, which addresses---again---the standard errors.
-This approach does not focus on the underlying data generating process, and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
-In our simulations of the MRQAP model, we find that it does not recover the data generating process as well as AME.
-This is quite due in part to the fact that it treats the dependencies as nuisance to be eliminated rather than focused on explicitly as with AME.
+\citet{erikson:pinto:2014} find  that ``the randomization results show that conventional t-tests are inappropriate for
+testing hypotheses on dyad-year data because they rely on highly overconfident standard errors'' (2014, page 462).
+This shows that in dyadic data analysis with general linear models, especially with hundreds of thousands of observations, the standard t-test is polyannaish in finding significant relations---a situation which characterizes much international relations literature with an abundance of statistically significant coefficients.
+MRQAP addresses the calculation of the standard errors, and is a kind of windowing approach \citep{heagerty:lumley:2000,heagerty:etal:2002}.
+This approach does not focus on the underlying data generating process and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
+MRQAP is only about the variance-covariance of the parameters, not the parameters themselves.
+In our simulations of the MRQAP model, we find that it does not recover the data generating process as well as AME---nor would we expect it to since its focus is different.
+This is due in part to the fact that it treats the dependencies as nuisance to be eliminated rather than focused on explicitly as with AME.
 

### 2018-04-11 paper/5_concl.tex
tiny fix to conclusion refs.

@@
 
-Poast \citeyear{poast:2010,poast:2016} doesn't really offer a way to estimate dyadic data in the presence of independence,
+Poast (\citeyear{poast:2010,poast:2016}) doesn't really offer a way to estimate dyadic data in the presence of independence,
 ``I am not presenting an alternative procedure for modeling such spatial, strategic, temporal,
-or monadic interdependencies in the data'' (2010, page 404, ironically).
+or monadic interdependencies in the data'' (Poast,2010, page 404, ironically).
 Instead, he argues that multilateral events are poorly represented or misrepresented by dyadic data.
@@
 
-Poast's solution is called choice-based sampling, ``[t]he sampling method entails constructing a data set containing all observations for which the dependent variable is coded with a positive value, along with a random sample of observations for which the dependent variable is coded 0'' (2010, page 410).
+Poast's solution is called choice-based sampling, ``[t]he sampling method entails constructing a data set containing all observations for which the dependent variable is coded with a positive value, along with a random sample of observations for which the dependent variable is coded 0'' (Poast, 2010, page 410).
 This new sample is then estimated with a rare-events logit regression.
-His results (2010, page 412) show that there is a small bias (about $1\%$) using the dyadic approach compared to the alternatives he proposes.
+His results (Poast, 2010, page 412) show that there is a small bias (about $1\%$) using the dyadic approach compared to the alternatives he proposes.
 As one moves from triadic to \textit{5-adic}, the divergence from the analysis of the dyadic data becomes substantially larger.

### 2018-04-11 paper/4_app.tex
final edits

@@
 
-For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimated the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
+For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimate the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
 Results with alternative values of $K$ are similar.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
@@
 The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from a subset of the data that does not include $y_{ij,t}$.
-Next we conducted a series of analyses to discern whether or not the AME model provides any benefit for each study.
+Next we conduct a series of analyses to discern whether or not the AME model provides any benefit for each study.
 These analyses are summarized in Figure~\ref{fig:perf}.
@@
 We hone in on the main finding of each study to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.
-In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates that the putative statistical significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and re-estimated models are presented in the Appendix.}
+In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates that the statistical significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and re-estimated models are presented in the Appendix.}
 
@@
 Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
-Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict the fixed effects in the model would suggest.
+Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the fixed effects in the model would suggest.
 None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.
@@
 
-Interestingly, Figure~\ref{fig:weekscirc} reveals a number of clusters.
+Figure~\ref{fig:weekscirc} reveals a number of notable clusters.
 For example, in the top left corner we see the US, UK, Germany, Canada, and Israel.
@@
 To explore this, we focus on the effect of rivalry on  MID initiation.
-Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects the two models contrast greatly.
+Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects between the two models differs greatly.
 

### 2018-04-11 paper/5_concl.tex
final edits

@@
 But are there any alternatives that have been developed?
-Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014} who proposes a variational estimation approach to latent network models.}
-None of these approaches has as yet been widely employed nor even cited in the substantive literature, and each is mainly known inside of the network methodology community.
+Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014} who proposes a variational estimation approach to latent network models.} None of these approaches has as yet been widely employed nor even cited in the substantive literature, and each is mainly known inside of the network methodology community.
 We discuss each of these approaches and briefly compare them to the AME approach.
-The takeaway is that based on simulation evidence, the AME approach also dominates these approaches in terms of recovering the data generating process and thus AME enables users to more accurately conduct inference on dependent data.\footnote{The detailed results are not presented herein.}
 
@@
 Their basic equation is
+
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+
 and the important parameter is $\beta$, defined in the standard way, via an appropriate estimator---they use OLS.
@@
 
-\citet{erikson:pinto:2014} base their approach on the MRQAP model developed by 
-Krackhardt (1988), \nocite{krackhart:1988} an approach from network studies that draws upon the fact that networks have repeated observations.
+\citet{erikson:pinto:2014} base their approach on the MRQAP model developed by Krackhardt (1988), \nocite{krackhardt:1988} an approach from network studies that draws upon the fact that networks have repeated observations.
 This induces correlation of observations thereby leading to bias in standard errors in general linear models.
@@
 Many permutations are created and models are run on all of them, then the distributions of tests statistics (not parameters) are examined.
-\citet{erikson:pinto:2014} find  that ``the randomization results show that conventional t-tests are inappropriate for
-testing hypotheses on dyad-year data because they rely on highly overconfident standard errors'' (2014, page 462).
+\citet{erikson:pinto:2014} find  that ``the randomization results show that conventional t-tests are inappropriate for testing hypotheses on dyad-year data because they rely on highly overconfident standard errors'' (2014, page 462).
 This shows that in dyadic data analysis with general linear models, especially with hundreds of thousands of observations, the standard t-test is polyannaish in finding significant relations---a situation which characterizes much international relations literature with an abundance of statistically significant coefficients.
@@
 \subsection*{Summary}
+
 International relations is about the interactions, relationships, and dependencies among countries or other important international actors.
@@
 It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.
-Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citep{king:roberts:2014}).
+Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citealp{king:roberts:2014}).
 Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences, better out-of-sample predictive performance, and new substantive insights.

### 2018-04-11 paper/netsMatter.tex
final edits

@@
 \begin{abstract}
-	Much international relations scholarship concerns dyads.
+	Much of international relations scholarship concerns dyads.
 Dyadic hypotheses and especially dyadic data abound.
@@
 Second, we compare the AME model to approaches used in three prominent studies from recent international relations scholarship.
-For each study, we find that compared to AME, the modeling approach used in each original study performs notably worse at capturing the data generating process.
+For each study, we find that compared to AME, the modeling approach used performs notably worse at capturing the data generating process.
 Further, conventional methods misstate the effect of key variables and the uncertainty in these effects.
@@
 %\bibliography{/Users/maxgallop/documents/whistle/master}
-%\bibliography{/Users/s7m/whistle/master}
-\bibliography{/Users/mdw/Git/whistle/master}
+\bibliography{/Users/s7m/whistle/master}
+% \bibliography{/Users/mdw/Git/whistle/master}
 % \bibliography{/Users/cassydorff/ProjectsGit/master}

### 2018-04-11 paper/5_concl.tex
cleaned up concl

@@
 \section{\textbf{Conclusion}}
-
-We have shown above that the AME framework for modeling dependent data dominates the standard approaches which use some form of a general linear model---typically a logit---to analyze dyadic data, especially in international relations.
-But are there any alternatives that have been developed?
-Three basic approaches have appeared as alternatives in the literature.\footnote{A good overview may be found in \cite{stewart:2014} who proposes a variational estimation approach to latent network models.} None of these approaches has as yet been widely employed nor even cited in the substantive literature, and each is mainly known inside of the network methodology community.
-We discuss each of these approaches and briefly compare them to the AME approach.
-
-\subsection*{Cluster-Robust Variance Estimation}
-
-\cite{aronow:etal:2015} discussed the problem of estimation of general linear models in the absence of $\simiid$ validity.
-In particular, they look at the clustering of errors that occur in dyadic data because multiple dyads share individual members.
-They develop a sandwich style estimator which leaves untouched the parameters, opting instead for a nonparametric approach which attempts to correct the variance-covariance estimates of the estimated parameters, leaving the parameters untouched.
-They do not compare their approach to AME or other strategies to explicitly estimate parametric effects in networks.
-Instead, they examine a spatial lag approach and general mixed effects models that abide by the $\simiid$ assumption.
-Their basic equation is
-
-\begin{eqnarray*}
-Y_{ij} = \mathbf{X}_{ij}^\intercal \beta + \varepsilon_{ij}
-\end{eqnarray*}
-
-and the important parameter is $\beta$, defined in the standard way, via an appropriate estimator---they use OLS.
-They assert that a random sampling of dyads that do not share a member will be uncorrelated, i.e., $\text{Cov}(\varepsilon_d,\varepsilon_{d'}) \equiv 0$, where $d'$ is the set of dyads that do not share members.
-This leads to the conclusion that the variance in the full sample of the dyadic, independent data and the dyadic error contains a known bias, part of which is a dyad specific variance component and another part is due to repeated, shared memberships in dyads.
-$\hat{\beta}$ is shown to be consistent, i.e., that as the number of \emph{dyads} increase to infinity, it converges in probability to its true value.
-
-The variance calculations used by Aronow \textit{et alia} basically decompose the variance into three parts, the cluster-robust variance (all dyads with unit $i$, assuming all others to be $\simiid$), the separate cluster-robust variance of the repeated dyads, and clustered-robust variance of all dyads, assuming they are all independent.
-Using simulation, they demonstrate that as sample size increases, this formula for calculating the variance-covariance matrix will decline as the number of nodes in the network increase and approach its true value.
-
-This approach focuses on the standard errors (the diagonal of the variance-covariance matrix) needed for significance testing, but ignores the bias introduced into the regression coefficients by ignoring dependence among the data, as discussed above.
-Nor does it provide any way to access the higher-order network effects, a point that Aronow et alia make in their article, noting that it is only targeted at first-order, dyadic (not higher level) dependencies.
-The cluster-robust variance approach simply provides a fix to the estimated standard errors, much like is presumed to work for many sandwich estimators of variance-covariance matrices.
-In this sense, it treats the dependent data as having dependencies that are a nuisance that only show up in the variance-covariance matrix, in the statistical sense.
-%One of the main problems with dependent data is that it takes more data to calculate basic quantities accurately. This is true even for a simple t-test \citep{ward:gleditsch:2008}.
-
-Cluster-robust variance estimation is not designed to answer the same questions as AME.
-As shown in the original article, it does provide a way to  estimate standard errors (in a sitution with only dyadic dependence) that substantially improves over the standard regression approach.
-However, it ignores the higher-order dependencies which are the focus of the AME model and many network analyses.
-
-%Isn't brandon stewart overview a better place to start?
-
-\subsection*{Choice-Based Sampling}
-
-Poast (\citeyear{poast:2010,poast:2016}) doesn't really offer a way to estimate dyadic data in the presence of independence,
-``I am not presenting an alternative procedure for modeling such spatial, strategic, temporal,
-or monadic interdependencies in the data'' (Poast,2010, page 404, ironically).
-Instead, he argues that multilateral events are poorly represented or misrepresented by dyadic data.
-In a sense, he is arguing against dyadic data.
-He creates an artificial database that is triadic and from it generates a process that combines the capabilities of the three actors in each triad into a ratio that drives whether an alliance is formed among all three actors.
-These triadic data are converted into a dyadic format.
-He then uses a logit regression framework to show that with the omniscient data in the triadic case, logit regression can recover the true value of the data generating process, but logit, as applied to the dyadic construction of the triadic data, produces estimates that are inflated.
-Thus, Poast's approach is about how the data are constructed in the first instance.
-However, we generally do not have multiadic data, so some strategy is needed to transform the data we do have, i.e., dyadic data.
-
-Poast's solution is called choice-based sampling, ``[t]he sampling method entails constructing a data set containing all observations for which the dependent variable is coded with a positive value, along with a random sample of observations for which the dependent variable is coded 0'' (Poast, 2010, page 410).
-This new sample is then estimated with a rare-events logit regression.
-His results (Poast, 2010, page 412) show that there is a small bias (about $1\%$) using the dyadic approach compared to the alternatives he proposes.
-As one moves from triadic to \textit{5-adic}, the divergence from the analysis of the dyadic data becomes substantially larger.
-Poast does not address the underlying issue of how to properly estimate dependent data, but raises the question whether dyadic data are capable of representing multi-adic data generating processes.
-Unfortunately, adjudicating this with an approach (logit regression) that itself is known to be incapable of capturing the higher-order aspects of the data generating process is, itself, problematic on first principles.
-
-However, how does Poast's approach compare to AME?
-We found in simulations structured like those reported herein run with the k-adic model that the k-adic model was dramatically worse at recovering the real data generating process compared to AME.
-This is because the AME approach captures the higher-order dependencies in dyadic data.
-It is these dependencies that Poast is trying to get at with his k-adic sampling approach.
-
-\subsection*{Multiple Regression Quadratic Assignment Procedure}
-
-\citet{erikson:pinto:2014} base their approach on the MRQAP model developed by Krackhardt (1988), \nocite{krackhardt:1988} an approach from network studies that draws upon the fact that networks have repeated observations.
-This induces correlation of observations thereby leading to bias in standard errors in general linear models.
-MRQAP randomly permutes the rows and columns of the data, creating multiple datasets which are then each analyzed in the standard way to generate a sampling distribution of all possible (but not observed) networks from the observed data.
-Herein, the data are randomly permuted on the assumption that the errors are exchangeable.
-Many permutations are created and models are run on all of them, then the distributions of tests statistics (not parameters) are examined.
-\citet{erikson:pinto:2014} find  that ``the randomization results show that conventional t-tests are inappropriate for testing hypotheses on dyad-year data because they rely on highly overconfident standard errors'' (2014, page 462).
-This shows that in dyadic data analysis with general linear models, especially with hundreds of thousands of observations, the standard t-test is polyannaish in finding significant relations---a situation which characterizes much international relations literature with an abundance of statistically significant coefficients.
-MRQAP addresses the calculation of the standard errors, and is a kind of windowing approach \citep{heagerty:lumley:2000,heagerty:etal:2002}.
-This approach does not focus on the underlying data generating process and is based on an approach to modeling dependent data that is known to produce bias and ignore dependencies.
-MRQAP is only about the variance-covariance of the parameters, not the parameters themselves.
-In our simulations of the MRQAP model, we find that it does not recover the data generating process as well as AME---nor would we expect it to since its focus is different.
-This is due in part to the fact that it treats the dependencies as nuisance to be eliminated rather than focused on explicitly as with AME.
-
-\subsection*{Summary}
 

### 2018-04-12 paper/coverletter/netsMatter_coverLetter.tex
wrappig up; ; git push

@@
 
-\vspace{1.5in}
+\vspace{.5in}
 
@@
 
-\vspace{0.5in}
+\vspace{0.25in}
 
@@
 Among political scientists using statistical models to analyze network data there has emerged a dispute between the advocates of the Exponential Random Graph (ERGM) model and those building on the ``latent space'' approach pioneered by Hoff, Raftery, and Handcock.
-It is worth noting that this fight is not mirrored in the statistical or broader networks literature, as in those communities people recognize these models as having distinct goals.
+This fight is not mirrored in the statistical or broader networks literature, as in those communities people recognize these models as having distinct goals.
+\\[1ex]
 
@@
 Although we clearly work in the latent space world, this paper is not about engaging or extending the ERGM-latent space discussion.
-Rather, we evaluate the utility of the AME framework using a simulation based exercise and a replication of five recently published works from the field of International Relations.
+Rather, we evaluate the utility of the AME framework using a simulation based exercise and a replication of three recently published works from the field of International Relations.
 Through the simulation exercise, we show that in the presence of unobserved dependencies the model is able to provide less biased estimates and better calibrated standard errors than extant approaches in the literature.
@@
 
-We believe that the AME framework is of notable interest to political scientists broadly, and that this study provides not only an introduction but a strong case for its applicability to the questions that we are seeking to address.\\[1ex]
-
+We believe that the AME framework is of notable interest to political scientists broadly, and that this study provides not only an introduction but a strong case for its applicability to the questions that we are seeking to address.
 We look forward to your evaluation of this paper.\\[1ex]

### 2018-04-12 paper/netsMatter.tex
wrappig up; ; git push

@@
 % % add watermark
-\usepackage{draftwatermark}
-\SetWatermarkText{Submission Copy}
-\SetWatermarkScale{.5}
-\SetWatermarkColor[gray]{0.9}
+% \usepackage{draftwatermark}
+% \SetWatermarkText{Submission Copy}
+% \SetWatermarkScale{.5}
+% \SetWatermarkColor[gray]{0.9}
 
@@
 We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.
-We benefited from presentations of this material at the 2017 Political Networks Conference, the 2017 Political Methodology Conference, and our home institutions.}
+We benefited from presentations of this material at the 2017 Political Methodology Conference, and our home institutions.}
 
@@
 We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the inherent dependencies in dyadic data.
-First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference about dyadic hypotheses.
+First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference on dyadic data.
 Second, we compare the AME model to approaches used in three prominent studies from recent international relations scholarship.

### 2018-04-17 paper/netsMatterSS.tex
making sharable copy
making a distributable single-space copy

--- 2017-10-29 paper/netsMatter.tex
@@
 % \usepackage{draftwatermark}
-% \SetWatermarkText{Version 1.0}
+% \SetWatermarkText{Submission Copy}
 % \SetWatermarkScale{.5}
@@
 
-% \date{\today}
+\date{\today}
 
-\thanks{Shahryar Minhas and Michael D.
+\thanks{\singlespacing Shahryar Minhas and Michael D.
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
 John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments.
-We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.}
+We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.
+We benefited from presentations of this material at the 2017 Political Methodology Conference, and our home institutions.}
 
@@
 \lhead{\fancyplain{}{}}
-\chead{\fancyplain{}{Taking Dyads Seriously}}
-\rhead{\fancyplain{}{\today}}
+% \chead{\fancyplain{}{Taking Dyads Seriously}}
+\chead{\fancyplain{}{}}
+% \rhead{\fancyplain{}{\today}}
+\rhead{\fancyplain{}{}}
 \rfoot{\fancyplain{}{\thepage}}
+
 % references to graphics
 \makeatletter
-% let's keep them in the same directory \def\input@path{{}
+% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
+% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
 
@@
 \begin{abstract}
-	%International relations scholarship is filled with dyads.
-This includes dyadic hypotheses and propositions, but especially dyadic data.
-Yet, much of the work conducted brushes aside the statistical complications that come with analyzing relational data, namely that the observations do not satisfy the conditional independence criterion.
-As a result, these studies often produce results with biased coefficient estimates and poorly calibrated standard errors, factors which have consequences for hypothesis testing.
-We present an alternative regression based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data using a set of additive and multiplicative effects within a Bayesian framework.
-To highlight the utility of this framework for studying relational data, we present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
-Additionally, we replicate five prominent studies in recent IR scholarship and compare the standard approach to the latent factor approach.
-For each study, our approach shows that conventional methods overstate the effect of key variables in the study of interstate conflict, underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance of their variables.
-Further, the new approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
-
-	Much international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data.
-Yet, contemporary research brushes aside the complications that come with analyzing interdependent relational data.
-Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
-As a result, many studies often produce results with biased coefficient estimates and poorly calibrated standard errors.
-These biases have profound consequences for evaluating parametric models.
-We present an alternative, regression-based, approach that accounts for the dependencies complicating this type of analysis.
-We first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data.
-In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to our alternative.
-For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables.
-Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
+Much of international relations scholarship concerns dyads.
+Dyadic hypotheses and especially dyadic data abound.
+Standard modeling approaches fail to adequately capture the data generating process behind dyadic events and processes.
+As a result, they suffer from biased coefficients and poorly calibrated standard errors.
+We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the inherent dependencies in dyadic data.
+First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference on dyadic data.
+Second, we compare the AME model to approaches used in three prominent studies from recent international relations scholarship.
+For each study, we find that compared to AME, the modeling approach used performs notably worse at capturing the data generating process.
+Further, conventional methods misstate the effect of key variables and the uncertainty in these effects.
+Finally, AME dominates standard approaches in terms of out-of-sample fit.
+In sum, our work shows the consequences of failing to take the dependencies inherent to dyadic data seriously.
+Most importantly, by better modeling the data generating process underlying political phenomena, the AME framework improves scholars' ability to conduct inferential analyses.
 \end{abstract}
@@
 \newpage\setcounter{page}{1} 
-
+\singlespacing
 \input{1_intro}
@@
 \clearpage
-%\bibliography{/Users/s7m/whistle/master}
+
+%\bibliography{/Users/maxgallop/documents/whistle/master}
+% \bibliography{/Users/s7m/whistle/master}
  \bibliography{/Users/mdw/Git/whistle/master}
-%\bibliography{/Users/cassydorff/ProjectsGit/master}
+% \bibliography{/Users/cassydorff/ProjectsGit/master}
 % \bibliographystyle{authordate1}

### 2018-05-22 code/1_simulation/1_ameSim.R
minor edits

@@ -10,14 +10,26 @@ if(Sys.info()['user']=='s7m'){
 toLoad = c(
 	'devtools', 
 	'foreach', 'doParallel',
-	'magrittr', 'dplyr', 'ggplot2' )
+	'magrittr', 'dplyr', 'ggplot2',
+	'sna', # mrqap
+	'sandwich', 'Matrix', 'lmtest', 'arm', 'reshape2' # asa
+	)
 loadPkg(toLoad)
-devtools::install_github('s7minhas/amen', ref='dev')
+# devtools::install_github('s7minhas/amen', ref='dev')
+library(amen)
 ##############################
 
 ##############################
 ### Homophily
-simRun = function(seed, n, mu, beta, gamma){
+seed = 1
+n = 30
+intEff=-2 ; x1Eff=1 ; x2Eff=1
+mu = intEff
+beta = x1Eff
+gamma = x2Eff
+imps = 1000
+imp=1
+# simRun = function(seed, n, mu, beta, gamma){
 	set.seed(seed) 
 	xw = matrix(rnorm(n*2),n,2)
 	X = tcrossprod(xw[,1])  ; W = tcrossprod(xw[,2])
@@ -34,13 +46,64 @@ simRun = function(seed, n, mu, beta, gamma){
 	fitO = ame(Y,XW,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
 		print=FALSE,plot=FALSE)
 
+	# MRQAP
+	loadPkg('sna')
+	fitQAP = netlm(y=Y, x=X, 
+		intercept=TRUE, mode='digraph', diag=FALSE,
+		nullhyp='qap', reps=200)
+
+	# aronow et al
+	robust.se.nodfc <- function(model, cluster){
+		M <- length(unique(cluster))
+		N <- length(cluster)
+		K <- model$rank
+		dfc <- 1
+		uj <- apply(estfun(model), 2, function(x) tapply(x, cluster, sum))
+		rcse.cov <- dfc * sandwich(model, meat = crossprod(uj)/N)
+		rcse.se <- coeftest(model, rcse.cov)
+		return(list(rcse.cov, rcse.se)) }
+	robust.se <- function(model, cluster){
+		M <- length(unique(cluster))
+		N <- length(cluster)
+		K <- model$rank
[…]

### 2018-05-22 paper/1_intro.tex
minor edits

@@
 In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
-This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hayes:2007,king:roberts:2014}.
+This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.
 

### 2018-05-22 code/1_simulation/epr_shuffle_repl/DyadReplicationPA.R

@@ -2,7 +2,7 @@
 ### Robert S. Erikson, Pablo M. Pinto, Kelly T. Rader
 
 ### Set working directory
-#setwd( )
+setwd('~/Research/netsMatter/code/1_simulation/epr_shuffle_repl/')
 
 library(foreign)
 library(plm)

### 2018-06-04 paper/2_model.tex
recompile some refs

@@
 $K$ denotes the dimensions of the latent space.
-The construction of the LFM here is actually quite similar to the recommender systems that companies like Amazon and Netflix use to model customer behavior \citep{resnick:varian:1997,bennett:etal:2007}.
+The construction of the LFM here is actually quite similar to the recommender systems that companies like Amazon and Netflix use to model customer behavior \citep{resnick:varian:1997,bennett:lanning:2007}.
 This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.

### 2018-07-15 prezzy/netsMatter.tex
use netmodels for base

--- 2017-11-10 prezzy/netsMatterDeck.tex
@@
 %!TEX TS-program = xelatex
-\documentclass[10pt,xcolor=dvipsnames,compress]{beamer}
-\usepackage{beamerthemesplit}
-\usecolortheme[cmyk={1.0,.69,0.0,.115}]{structure}
-\usetheme[height=11mm]{Rochester}
-\setbeamertemplate{items}[ball] 
-\setbeamerfont{tablefont}{size=\tiny}
-\setbeamerfont{quotefont}{size=\small}
+\documentclass[10pt, compress]{beamer}
+
+\usetheme[usetitleprogressbar]{m}
 
@@
 \usepackage{etoolbox}
+\usepackage{booktabs}
 \usepackage{dcolumn}
@@
 \usepackage{graphicx}
-\usepackage[space]{grffile} 
+\usepackage[space]{grffile} % allows us to specify directories that have spaces
 \usepackage{tikz}
@@
 % to change enumeration symbols begin{enumerate}[(a)]
-%\usepackage{enumerate}
+\usepackage{enumerate}
+
+% Add some colors
+\definecolor{red1}{RGB}{253,219,199}
+\definecolor{red2}{RGB}{244,165,130}
+\definecolor{red3}{RGB}{178,24,43}
+
+\definecolor{green1}{RGB}{229,245,224}
+\definecolor{green2}{RGB}{161,217,155}
+\definecolor{green3}{RGB}{49,163,84}
+
+\definecolor{blue0}{RGB}{255,247,251}
+\definecolor{blue1}{RGB}{222,235,247}
+\definecolor{blue2}{RGB}{158,202,225}
+\definecolor{blue3}{RGB}{49,130,189}
+\definecolor{blue4}{RGB}{4,90,141}
+
+\definecolor{purple1}{RGB}{191,211,230}
+\definecolor{purple2}{RGB}{140,150,198}
+\definecolor{purple3}{RGB}{140,107,177}
+
+\definecolor{brown1}{RGB}{246,232,195}
+\definecolor{brown2}{RGB}{223,194,125}
+\definecolor{brown3}{RGB}{191,129,45}
 
@@
 \makeatletter
-\def\input@path{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/}, {/Users/mdw/Dropbox/Research/Ongoing/netModels/summResults/}}
-\graphicspath{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/},{/Users/mdw/Dropbox/Research/Ongoing/netModels/summResults/}}
+\def\input@path{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/}, {/Users/mdw/Dropbox/Research/netModels/summResults/}}
+\graphicspath{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/},{/Users/mdw/Dropbox/Research/netModels/summResults/}}
 
-\title[AMEN]{Taking Dyads Seriously\thanks{Based upon joint works with Michael Ward (Duke University), Cassy L.
-Dorff (University of New Mexico), Max Gallop (University of Strathclyde), Juan Tellez (Duke University), Howard Liu (Duke University), \& Margaret Foster (Duke University).
-This is also part of a broader project involving Peter Hoff (Duke University) and Michael Ward.}
-}
+\title[AMEN]{\textsc{Taking Dyads Seriously}}
+\author[Minhas et al.]{Shahryar Minhas$^\dag$, and all the others \\ Duke University \\ $^\dag$ Department of Political Science \\ $^\ddag$ Department of Statistical Science}
 
-\author 
-{Shahryar Minhas\\ minhassh@msu.edu \\ s7minhas.com \\
-Michigan State University}
-
-\date{Beamer is ugly, this needs to be changed.}
+\date{\today}
 
@@
 
-\begin{frame}
-\frametitle{Abstract}
-\framesubtitle{An Abstract for a Talk?}
-\begin{itemize}
-\item Much international relations scholarship concerns dyads.
-Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches.
-As a result, studies often produce results with biased coefficient estimates and poorly calibrated standard errors.
-These biases have profound consequences for evaluating parametric models.
-
-\item We present an alternative, regression-based, approach that accounts for the dependencies complicating this type of analysis, and present a simulation exercise highlighting the model's ability to account for the dyadic dependencies.
-We replicate five prominent studies, comparing the standard approach to our alternative.
-
-\item Conventional methods overstate the effect of key variables, underestimate the uncertainties, and often lead researchers to faulty conclusions about the substantive importance of their variables.
-
-\item Our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
-\end{itemize}
-\end{frame}
-
-\frame{\frametitle{Overview of Seminar}
-\framesubtitle{Please, ask questions when you have them.}\tableofcontents}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Setup}
 \frame{
@@
 
+  \vspace{-10mm}
   Relational data consists of 
   \begin{itemize}
-    \item a set of units or nodes
-    \item a set of measurements, $y_{ij}$, specific to pairs of nodes $(i,j)$ 
+    \item - a set of units or nodes
+    \item - a set of measurements, $y_{ij}$, specific to pairs of nodes $(i,j)$ 
   \end{itemize}
@@
 \begin{itemize}
-\item  biased effects estimation
-\item  uncalibrated confidence intervals
-\item  poor predictive performance
-\item  inaccurate description of network phenomena
+\item - biased effects estimation
+\item - uncalibrated confidence intervals
+\item - poor predictive performance
+\item - inaccurate description of network phenomena
 \end{itemize}
@@
 \end{tabular}
+
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{Outline}
+
+\begin{itemize}
+  \item - Nodal and dyadic dependencies in networks
+  \item \qquad -  Can model using the ``A'' in AME 
+  \item - Third order dependencies
+  \item \qquad - Can model using the ``M'' in AME
+  \item - Application
+\end{itemize}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Social Relations Model}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{itemize}
-\item $\mu$ baseline measure of network activity
-\item $e_{ij}$ residual variation that we will use the SRM to decompose
+\item - $\mu$ baseline measure of network activity
+\item - $e_{ij}$ residual variation that we will use the SRM to decompose
 \end{itemize}
@@
       y_{ij} &= \mu + e_{ij} \\
-      e_{ij} &= {\color{red}{a_{i} + b_{j}}} + \epsilon_{ij} \\
+      e_{ij} &= \color{red}{a_{i} + b_{j}} + \epsilon_{ij} \\
       \color{red}{\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \}} &\sim N(0,\Sigma_{ab}) \\ 
@@
 \begin{itemize}
-\item  row/sender effect ($a_{i}$) \& column/receiver effect ($b_{j}$)
-\item  Modeled jointly to account for correlation in how active an actor is in sending and receiving ties
+\item - row/sender effect ($a_{i}$) \& column/receiver effect ($b_{j}$)
+\item - Modeled jointly to account for correlation in how active an actor is in sending and receiving ties
 \end{itemize}
@@
 \begin{itemize}
-\item $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$ capture heterogeneity in the row and column means
-\item $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties)
+\item - $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$ capture heterogeneity in the row and column means
+\item - $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties)
 \end{itemize}
@@
 \begin{itemize}
-\item  $\epsilon_{ij}$ captures the within dyad effect
-\item  Second-order dependencies are described by $\sigma_{\epsilon}^{2}$
-\item  Reciprocity, aka within dyad correlation, represented by $\rho$
+\item - $\epsilon_{ij}$ captures the within dyad effect
+\item - Second-order dependencies are described by $\sigma_{\epsilon}^{2}$
+\item - Reciprocity, aka within dyad correlation, represented by $\rho$
 \end{itemize}
@@
 \frame{
-  \frametitle{Third-Order Dependencies}
+  \frametitle{Third Order Dependencies}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Latent Factor Model}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\frame{
-  \frametitle{What Exactly Is Third-Order Dependency?}
-  \framesubtitle{People Want to Know!}
-Too much to read on a slide.
-\begin{itemize}
-\item<1> If each node has individual characteristics, 
-say $x_i$ and $x_j$ for two nodes.  Then,
-$x_i \times x_j$ is the multiplicative product of their nodal characteristics and may be related to their relationship, 
-$y_{ij}$.  
-Let $x$ be an encoding of whether a node is democratic or not. 
-Then, $x_i \times x_j \equiv 1 \iff$ 
-\textit{both are democracies}.  This approach is often used in standard analyses as a variable often known as 
-\textit{joint democracy}.
-In the context of networks, a positive effect from the interaction of $x_i$ and $x_j$ might affect some dyadic variable, such as not being in a militarized dispute against one another---often denoted the democratic peace.
-This positive effect is known in network terminology as \textit{homophily}.
-Most studies just look at the correlation between aggregated variables, but the proposed approach can more directly estimate the degree of homophily.
-\end{itemize}
-}
-
-\frame{
-  \frametitle{What Exactly Is Third-Order Dependency?}
-  \framesubtitle{People Want to Know!}
-\begin{itemize}
-\item<1> Homophily which stems from nodal characteristics results in certain types of network structures: 1) transitivity, 2) balance, and 3) clustering.
-For example, homophily tends to induce a large number of \textit{triads}, since if $i$ is similar to $j$ and as a result has a link with $j$, for the same reason it will share a link with $k$ if $j$ and $k$ are similar, thus creating a triad.
-Groups of triads tend to result in \textit{clusters}.
-\end{itemize}
-}
-
-\frame{
-  \frametitle{What Exactly Is Third-Order Dependency?}
-  \framesubtitle{People Want to Know!}
-\begin{itemize}
-\item<1> Stochastic equivalence.
-\end{itemize}
-}
-
 \frame{
@@
 
-This can account for both stochastic equivalence and homophily, as well as clustering and balance.
+Can account for both stochastic equivalence and homophily
 
-}
-
-\frame{
-  \frametitle{How Does Estimation Work?}
-  \framesubtitle{It Is NOT All in the Error Term!}
- \begin{itemize}
-  \item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-  \item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-  \item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-  \item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
-  \item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
-  \item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
-  \item For each $k \in K$:
-  \begin{itemize}
-    \item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-    \item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
-    \item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$.
-}
-  \end{itemize}
- \end{itemize}
-
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulation of AME}
-\frame{
-  \frametitle{Simulation Comparing GLM and LFM}
-  \framesubtitle{}
-We utilize a simulation study to highlight the utility of AME as an inferential tool for dyadic analysis.
-The goal of the simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Specifically, we are concerned with conducting inference on regression parameters of a linear model for a network in the case where there is an omitted variable.
-For instance, assume that the true data-generating process for a particular $Y$ is given by:
-
-\begin{align}
-  y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
-  \label{eqn:sim}
-\end{align}
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
-$Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
-}
-
-\frame{
-  \frametitle{Simulation Comparing GLM and LFM}
-  \framesubtitle{Three Alternatives:}
-
-\begin{itemize}
-  \item the standard international relations approach assuming independent errors (current state-of-the discipline); 
-  \item the AME approach with a unidimensional latent factor space ($K=1$);
-  \item an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$. 
-\end{itemize}
-
-For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.
-The value of $\gamma$ is also set to 1, which corresponds to an example where $W$ is associated with homophily.
-}
-
-\frame{
-  \frametitle{Regression Parameters from Simulations}
-  \framesubtitle{}
-\begin{figure}
-  \centering
-  \caption{Regression parameter estimates for the naive, AME, and oracle models from 1,000 simulations.
-Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-  \label{fig:ameBias}
-  \includegraphics[width=\textwidth]{ameSimBias_all.pdf} \\
-\end{figure}
-}
-%
-\frame{
-  \frametitle{Coverage in the Simulations}
-  \framesubtitle{}
-\begin{figure}
-  \centering
-  \caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the naive, AME, and oracle models from 1,000 simulations.}
-  \label{fig:ameCalib}
-  \includegraphics[width=\textwidth]{ameSimCover_all.pdf} \\
-\end{figure}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Five Easy Pieces: Replication with AME}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-    \frametitle{Does it Really Matter?}
-  \framesubtitle{Let's Look at Five Recent Publications in IR}
-
-\begin{table}
-\caption{Features of the Studies Re-estimated.}
-\scalebox{.7}{
-  \begin{tabular}{lcccccc}
-    & Model &  Date Range & N. Actors  & N. Dyads & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
-    Reiter \& Stam (JCR; 2003) &Logit &1945--1995 &  193 & $753,456$ & Directed & Robust \\  
-    McDonald (JCR; 2004) & Logit &1959--2002 & 198 & $92,354$ & Undirected & Robust\\
-    Rose (AER; 2004) & OLS & 1948--1999 & 177 & $234,597$ & Directed & Robust \\  
-    Weeks (APSR; 2012) & Logit & 1946--1999 &197 &  $901,540$ & Directed & Robust \\
-    Gibler (APSR; 2017) & Logit & 1816--2008 &193 &   $650,557$ & Undirected & none \\ \bottomrule
-  \end{tabular}
-  }
-\end{table} 
-}
-
-\frame{
-    \frametitle{Does it Really Matter?}
-  \framesubtitle{Some Deets}
-\begin{table}[ht]
-\centering
-\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network-based approach.}
-\scalebox{.7}{
-  \begin{tabular}{l p{7cm} l} \toprule
-    \multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
-      & &  in a Network Model? \\ \toprule
-    Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Does Not Replicate} \\ \midrule
-    McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
-    Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
-    Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive; Machines are Not & {Does Not Replicate} \\\midrule
-    Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
-  \end{tabular}
-  }
-  \label{tab:modelFindingSumm}
-\end{table}
-}
-
-\frame{
-  \frametitle{Performance Statistics}
-  \framesubtitle{In out-of-sample tests}
-\vspace{-5mm}
-\begin{table}[ht]
-\scalebox{.7}{
-  \begin{tabular}{ll cc}  \toprule
-  Study & Statistic & GLM & AME \\
-  \midrule
-\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
-    ~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\   \midrule
-\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
-    ~ & AUC-PR & 0.13 &  {0.28} \\    \midrule
-\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
-    ~ & RMDSE & 2.01 &  {1.06} \\ \midrule
-\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
-    ~ & AUC-PR & 0.00 &  {0.15} \\    \midrule
-\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
-    ~ & AUC-PR & 0.00 &  {0.08} \\      \bottomrule
-  \end{tabular}
-  }
-\end{table}
-}
-
-\frame{
-  \frametitle{Reiter-Stam Performance Comparisons}
-  \framesubtitle{ }
-\vspace{-5mm}
-\begin{figure}
-\includegraphics[height=5cm]{rsroc}
-\includegraphics[height=5cm]{rspr}
-\caption{Reiter-Stam Performance}
-\end{figure}
-}
-
-\frame{
-  \frametitle{McDonald Performance Comparisons }
-  \framesubtitle{ }
-\vspace{-5mm}
-\begin{figure}
-\includegraphics[height=5cm]{mcdonaldroc}
-\includegraphics[height=5cm]{mcdonaldpr}
-\caption{McDonald Performance }
-\end{figure}
-}
-
-\frame{
-  \frametitle{Rose Performance Comparisons }
-  \framesubtitle{ }
-%\vspace{-7.5mm}
-\begin{figure}
-\includegraphics[scale=.5]{roseout}\\
-\includegraphics[scale=.4]{rose_srmvc}
-\caption{Rose Performance Statistics}
-\end{figure}
-}
-
-\frame{
-  \frametitle{Weeks Performance Comparisons }
-  \framesubtitle{ }
-\vspace{-5mm}
-\begin{figure}
-\includegraphics[height=5cm]{weeksroc}
-\includegraphics[height=5cm]{weekspr}
-\caption{Weeks Performance; no single MID predicted}
-\end{figure}
-}
-
-\frame{
-  \frametitle{Gibler Performance Comparisons }
-  \framesubtitle{ }
-\vspace{-5mm}
-\begin{figure}
-\includegraphics[height=5cm]{giblerroc}
-\includegraphics[height=5cm]{giblerpr}
-\caption{Gibler Performance; no single MID predicted}
-\end{figure}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-\section{Lessons Learned}
-\frame{
-   \frametitle{Some Lessons Learned}
-  \framesubtitle{ } 
-\begin{itemize}
-\item Findings that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.
-In GLM approaches to the analysis of dyadic data, coefficents are biased, imprecisely measured, and have poorly calibrated standard errors.
-This means that significance testing (for better or worse) is compromised when network effects are ignored.
-\item New insights emerge from the additional information derived from AME.
-In particular, there is actual information about the dependencies so that clusters can be identified,   the extent of reciprocity at the dyad level, and among dyads.
-\item The actual results---not the estimated coefficients and their covariances---which are generated by the models differ greatly in expectations.   
-\item The AME approach dominates the GLM approaches in terms of performance.
-Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$, but it also dominates at correctly identifying occurrences of $1$.
-
-\end{itemize}
-}
-
-\begin{frame}{Frame}
-  \frametitle{Networks Really Matter}
-  \framesubtitle{We miss a lot when we ignore them.}
-  \begin{columns}
-    \column{0.5\textwidth}
-    \centering
-      \includegraphics[width=\linewidth]{weeks_circPlot.pdf}
-      \column{0.5\textwidth}
-      \setlength\arraycolsep{3pt}
-      \begin{itemize}
-        \item clustering
-        \item different patterns for senders and receivers
-        \item similar nodal, and dyadic patterns 
-        \item gauge reciprocity (and orthogonality)
-      \end{itemize}
- \end{columns}
-\end{frame}
-
-\frame{\frametitle{Thanks!}
-\framesubtitle{There's more, but I'll stop here for your questions and comments.}
-
-More work; copies available upon request:
-
-\begin{itemize}
-\item Shahryar Minhas, Peter D. Hoff, and Michael D. Ward. 2017. \textit{Additive and Multiplicative Latent Factor Models for Network Influence}. 
-\item Shahryar Minhas, Peter D. Hoff, and Michael D. Ward. 2017. \textit{Influence Networks in International Relations}. 
-\item Shahryar Minhas, Cassy L. Dorff, Margaret Foster, Max Gallop, Howard Liu, Juan Tellez, and Michael D. Ward. 2017. \textit{Taking Dyads Seriously}.
-\end{itemize}
-
-Thank you!
-
-}
-\section{Appendix: Code Example}
-
-\frame{
-  \frametitle{Example $\mathcal{R}$ code}
-Using the AMEN function requires formatting data into a particular structure.
-The primary distinction in data formatting is whether the outcome of interest represents a directed or undirected network.
-Use this link \url{GitHub.com/s7minhas/amen}
-
-If undirected, the AMEN function has three main inputs:
-
-\begin{itemize} 
-    \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network.
-    \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset. 
-    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
-\end{itemize}
-
-If directed, AMEN further requires:
-
-\begin{itemize} 
-    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
-    \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset. 
-\end{itemize}
-
-}
-
-\frame{
-  \frametitle{AMEN Example}
-  \framesubtitle{specification}
-Beyond the data inputs, the AMEN function requires additional specification:
-
-\begin{itemize}
-    \item model: how to model the outcome variable, e.g., `logit'
-    \item symmetric: whether the input network is symmetric
-    \item intercept: whether to estimate an intercept
-    \item nscan: number of iterations of the Markov chain
-    \item burn: burn-in period
-    \item odens: thinning interval
-    \item R: dimension of the multiplicative effect (referred to as $K$ in the paper)
-    \item gof: whether to calculate goodness-of-fit statistics
-\end{itemize}
-
-}
-
-\frame{
-  \frametitle{Actual Code}
-\# running in parallel varying k\\
-imps = 10000 ; brn = 25000 ; ods = 10 ; latDims = 0:3\\
-\# Run amen in parallel
-library(doParallel) ; library(foreach) ; cl=makeCluster(4) \\ registerDoParallel(cl)\\
-foreach(ii=1:length(latDims), .packages=c("amen")) \%dopar\% {\\
-  \# load previous model run\\
-  load(prevModelFiles[ii])\\
-  \# extract start vals\\
-  startVals0 = ameFit\$'startVals'\\
-  \# dump rest\\ 
-  rm(ameFit)\\
-   ameFit = ame\_repL(\\
-    Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL, \\
-    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii], \\
-    nscan=imps, seed=1, burn=brn, odens=ods, \\
-    plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0,\\
-    periodicSave=TRUE ) \\    
-  save(ameFit, file=paste0('model\_k', latDims[ii],'\_v2.rda') )\\
-}
-stopCluster(cl)
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Backup Slides}
-\frame{\frametitle{backup slides}
-Go ahead, ask me about ERGM. }
-
-\frame{
-  \frametitle{There is a problem with ERGM.}
-  \framesubtitle{It is a bug, not a feature.}
-\vspace{-5mm}
-It is a bug, not a feature. If you don't believe us, believe these scholars:
-\begin{enumerate}
-\item Schweinberger, M. (2011). Instability, Sensitivity, and Degeneracy of Discrete Exponential Families. \textbf{Journal of the American Statistical Association}, 106(496):1361–1370.
-\item Schweinberger, M.
-and Handcock, M.
-S.
-(2015).
-Local Dependence in Random Graph Models: Characterization, Properties and Statistical Inference.
-\textbf{Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, 77(3):647–676.
-\item Chatterjee, S. and Diaconis, P. (2013). Estimating and Understanding Exponential Random Graph Models. \textbf{The Annals of Statistics}, 41(5):2428–2461.
-\item Rastelli, R., Friel, N., and Raftery, A. E. (2016). Properties of Latent Variable Network Models. \textbf{Network Science}, 4(4):1–26.
-\end{enumerate}
 }
@@
 \frame{
-  \frametitle{What Is Really Wrong? }
-\vspace{-5mm}
-
-You may assume this is your likelihood surface.
-Courtesy of Chatterjee and Diaconis, who concluded that sufficient statistics are not sufficient to solve this problem.
-Estimation of ERGMs contains this flaw, with pseudolikelihood, maximum likelihood, or MCMC.
-\includegraphics[width=1\textwidth]{disconuity.png}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-  \frametitle{What Does This Mean? }
-\vspace{-5mm}
-\begin{enumerate}
-\item Probabilistic ERGM models place almost all of the probability on networks that
-are either nearly empty (degenerate) with no linkages or nearly saturated with
-all nodes being interconnected. 
-\item The likelihood surface contains steep or discontinuous gradients that render it impossible to solve numerically (or analytically). Even (especially) for very small networks this is problematic.
-\item Find a different way to estimate network data.  
-\end{enumerate}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-  \frametitle{Extant Political Science Examples}
-\vspace{-5mm}
-Extant Political Science Examples Using Latent Distance Models Are Wrong.
-\begin{enumerate}
-\item There is a problem with the MCMC estimator in \texttt{latentnet}. Don't use it until it is fixed.
-\item Latent distance models are hard to correctly interpret, in prominent evaluations in the literature it is done incorrectly.  This is explained in their introduction a decade ago.
-\item The advice in Cranmer et al is wrong. Latent network models are not biased and when correctly implemented out-perform ERGM models.
-\end{enumerate}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-\frametitle{Compared to ERGM?}
-  \framesubtitle{Swiss Climate Change Application}
+  \frametitle{Swiss Climate Change Application}
 
@@
 \frame{
-  \frametitle{Parameter Estimates for Swiss Example}
+  \frametitle{Parameter Estimates}
 
@@
 \frame{
-  \frametitle{Latent Factor Visualization for Swiss Example}
+  \frametitle{Latent Factor Visualization}
 
@@
 \frame{
-  \frametitle{Out-of-Sample Performance Assessment: Strategy}
+  \frametitle{Out of Sample Performance Assessment}
 
 \begin{itemize}
-  \item  Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
-  \item  For each $s \in \{1, \ldots, S\}$:
+  \item - Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
+  \item - For each $s \in \{1, \ldots, S\}$:
   \begin{itemize}
-    \item  Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
-    \item  For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
+    \item - Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
+    \item - For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
   \end{itemize}
@@
 \frame{
-  \frametitle{Performance Comparison for Swiss Example}
+  \frametitle{Performance Comparison}
 
@@
 \frame{
-  \frametitle{Network Dependencies: Swiss Example}
+  \frametitle{Network Dependencies}
+
   \centering
@@
 }
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \frame{
-  \frametitle{Much current advice is wrong.}
-  \textbf{Unfortunately some misleading and incorrect advice is in the current literature. Maybe it will be corrected.}
+  \frametitle{What's Next?}
 
-\begin{enumerate}
-\item Cranmer, Leifeld, McClurg, Rolfe (2017 AJPS) stated that latent space models are only preferable if there are few isolates, enough observations, and the interdependence is theoretically uninteresting.
+  \begin{itemize}
+    \item - Extending AMEN to handle changing actor compositions
+    \item - Generalize multiplicative effects estimation over tensors
+  \end{itemize}
 
-\item All of this is wrong.
-Not only do they incorrectly assume all latent models to be based on Euclidean distance but their statistical comparisons are based on a program that contains known errors.
-Moreover, using BIC to choose among different network models is itself questionable.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\item It is not true that ``ERGM offers the most preferable combination of model fit and number of parameters,'' despite their assertions.
-\end{enumerate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\plain{Thanks.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{Standard Network Dependence Measures}
+\vspace{-5mm}
+\includegraphics[width=1\textwidth]{ggGofAll_preeze}
 }

### 2018-08-01 paper/netsMatter.tex
add word count to title
non anon version for apsa

@@
 
-% \title{Taking Dependencies Seriously:\\A Latent Network Approach to Dyadic Analysis}
-% \title{Taking Dyads Seriously:\\A Latent Network Approach to Dyadic Analysis}
 \title{Taking Dyads Seriously}
@@
 
-% \author[Minhas]{Shahryar Minhas}
-% \address{Shahryar Minhas: Department of Political Science}
-% \curraddr{Michigan State University}
-% \email[Corresponding author]{minhassh@msu.edu}
+\author[Minhas]{Shahryar Minhas}
+\address{Shahryar Minhas: Department of Political Science}
+\curraddr{Michigan State University}
+\email[Corresponding author]{minhassh@msu.edu}
 
-% \author[Dorff]{Cassy L. Dorff}
-% \address{Cassy L. Dorff}
-% \curraddr{Department of Political Science, University of New Mexico}
-% \email{cdorff@unm.edu}
+\author[Dorff]{Cassy L. Dorff}
+\address{Cassy L. Dorff}
+\curraddr{Department of Political Science, University of New Mexico}
+\email{cdorff@unm.edu}
 
-% \author[Foster]{Margaret Foster}
-% \address{Margaret Foster}
-% \curraddr{Department of Political Science, Duke University}
-% \email{margaret.foster@duke.edu}
+\author[Foster]{Margaret Foster}
+\address{Margaret Foster}
+\curraddr{Department of Political Science, Duke University}
+\email{margaret.foster@duke.edu}
 
-% \author[Gallop]{Max Gallop}
-% \address{Max Gallop}
-% \curraddr{Department of Government and Public Policy, University of Strathclyde}
-% \email{max.gallop@strath.ac.uk}
+\author[Gallop]{Max Gallop}
+\address{Max Gallop}
+\curraddr{Department of Government and Public Policy, University of Strathclyde}
+\email{max.gallop@strath.ac.uk}
 
-% \author[Liu]{Howard Liu}
-% \address{Howard Liu}
-% \curraddr{Department of Political Science, Duke University}
-% \email{hao.liu@duke.edu}
+\author[Liu]{Howard Liu}
+\address{Howard Liu}
+\curraddr{Department of Political Science, Duke University}
+\email{hao.liu@duke.edu}
 
-% \author[Tellez]{Juan Tellez}
-% \address{Juan Tellez}
-% \curraddr{Department of Political Science, Duke University}
-% \email{juan.tellez@duke.edu}
+\author[Tellez]{Juan Tellez}
+\address{Juan Tellez}
+\curraddr{Department of Political Science, Duke University}
+\email{juan.tellez@duke.edu}
 
-% \author[Ward]{Michael D. Ward}
-% \address{Michael D. Ward}
-% \curraddr{Department of Political Science, Duke University}
-% \email{michael.d.ward@duke.edu}
+\author[Ward]{Michael D. Ward}
+\address{Michael D. Ward}
+\curraddr{Department of Political Science, Duke University}
+\email{michael.d.ward@duke.edu}
 
-% % \date{\today}
+% \date{\today}
 
-%\thanks{Shahryar Minhas and Michael D.
+\thanks{Shahryar Minhas and Michael D.
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments.
+Peter Hoff, John Ahlquist, Robert Franzese, Arturas Rozenas, and Benjamin Appel provided welcome and helpful comments.
 We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.
@@
 
-\maketitle\thispagestyle{empty}
+\maketitle
+
+\begin{center}
+	\textbf{Word count}: 9,177
+\end{center}
+
+\thispagestyle{empty}
 \newpage\setcounter{page}{1}

### 2018-08-02 paper/2_model.tex
fix citations

@@
 It is flexible and easy to use for analyzing the kind of relational data often found in the social sciences.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2016:arxiv} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2018} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
 For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.

### 2018-08-29 prezzy/netsMatter.tex
preeze deck

@@
 \usepackage{multirow}
+\usepackage{colortbl}
 
@@
 \makeatletter
-\def\input@path{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/}, {/Users/mdw/Dropbox/Research/netModels/summResults/}}
-\graphicspath{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/},{/Users/mdw/Dropbox/Research/netModels/summResults/}}
+\def\input@path{{/Users/janus829/Research/netsMatter/prezzy/graphics/}, {/Users/s7m/Research/netsMatter/prezzy/graphics/}}
+\graphicspath{{/Users/janus829/Research/netsMatter/prezzy/graphics/}, {/Users/s7m/Research/netsMatter/prezzy/graphics/}}
 
 \title[AMEN]{\textsc{Taking Dyads Seriously}}
-\author[Minhas et al.]{Shahryar Minhas$^\dag$, and all the others \\ Duke University \\ $^\dag$ Department of Political Science \\ $^\ddag$ Department of Statistical Science}
+\author[Minhas et al.]{Shahryar Minhas, Cassy Dorff, Max Gallop, Margaret Foster, Juan Tellez, Howard Liu, Michael Ward}
 
@@
 \begin{itemize}
-\item - biased effects estimation
-\item - uncalibrated confidence intervals
-\item - poor predictive performance
-\item - inaccurate description of network phenomena
+\item biased effects estimation
+\item uncalibrated confidence intervals
+\item poor predictive performance
+\item inaccurate description of network phenomena
 \end{itemize}
@@
 Kenny (1996) & Li \& Loken (2002) & Aronow et al. (2015) \\
-Krackhardt (1998) & Hoﬀ \& Ward (2004) & Athey et al. (2016) \\
 \end{tabular}
@@
 \begin{itemize}
-  \item - Nodal and dyadic dependencies in networks
-  \item \qquad -  Can model using the ``A'' in AME 
-  \item - Third order dependencies
-  \item \qquad - Can model using the ``M'' in AME
-  \item - Application
+  \item Nodal and dyadic dependencies in networks
+  \begin{itemize}
+	  \item Can model using the ``A'' in AME 
+  \end{itemize}
+  \item Third order dependencies
+  \begin{itemize}
+	  \item Can model using the ``M'' in AME
+  \end{itemize}	  
+  \item Simulation	  
+  \item Empirical Application
 \end{itemize}
@@
 
-We use this model to form the additive effects portion of AME
+We use this model to form the additive effects portion of AME (Warner et al. 1979, Li \& Loken 2002, Hoff 2005):
 
@@
 
-Each node $i$ has an unknown latent factor
+Each node $i$ has an unknown latent factor:
 
@@
 \vspace{-5mm}
-The probability of a tie from $i$ to $j$ depends on their latent factors
+The probability of a tie from $i$ to $j$ depends on their latent factors (Hoff 2008, Hoff 2015, Minhas, Hoff \& Ward 2018):
 
@@
 \frame{
-  \frametitle{Swiss Climate Change Application}
+  \frametitle{AME}
 
-Cross-sectional network measuring whether an actor indicated that they collaborated with another during the policy design of the Swiss CO$_{2}$ act (Ingold 2008)
+The full AME framework enables us to not only account for the types of dependencies that often arise in dyadic data but also estimate them:
 
-\begin{figure}[ht]
-  \centering
-  \begin{tabular}{cc}
-  \includegraphics[width=.47\textwidth]{dvNet_outDegree} & 
-  \includegraphics[width=.44\textwidth]{dvNet_inDegree}
-  \end{tabular}
-\end{figure}
+\begin{align}
+  \begin{aligned}
+    y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
+    \theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} \text{\qquad(Exogenous parameters)} \\
+    & + a_{i} + b_{j} + \epsilon_{ij} \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
+    & + \mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
+  \label{eqn:ame}
+  \end{aligned}
+\end{align}
 
@@
 \frame{
-  \frametitle{Parameter Estimates}
+  \frametitle{Simulation}
 
-% latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Sun Aug 21 03:32:43 2016
-\vspace{-15mm}
-\begin{table}[ht]
-\centering
-\begingroup\scriptsize
-\begin{tabular}{lcccccc}
-   & Expected & \multirow{2}{*}{Logit} & \multirow{2}{*}{MRQAP} & \multirow{2}{*}{LSM} & \multirow{2}{*}{ERGM} & \multirow{2}{*}{AME} \\ 
-   & Effect & & & & & \\
-  \hline
-  \textbf{Conflicting policy preferences} &  &  &  &  &  &  \\ 
-  $\;\;\;\;$ Business vs. NGO & \color{red}{$-$} & -0.86 & \color{red}{-0.87$^{\ast}$} & \color{red}{-1.37$^{\ast}$} & \color{red}{-1.11$^{\ast}$} & \color{red}{-1.37$^{\ast}$} \\ 
-  $\;\;\;\;$ Opposition/alliance & \color{blue}{$+$} &  \color{blue}{1.21$^{\ast}$} & \color{blue}{1.14$^{\ast}$} & 0.00 & \color{blue}{1.22$^{\ast}$} & \color{blue}{1.08$^{\ast}$} \\ 
-  $\;\;\;\;$ Preference dissimilarity & \color{red}{$-$} & -0.07 & -0.60 & \color{red}{-1.76$^{\ast}$} & -0.44 & \color{red}{-0.79$^{\ast}$} \\ 
-  \textbf{Transaction costs} &  &  &  &  &  &  \\ 
-  $\;\;\;\;$ Joint forum participation & \color{blue}{$+$} & \color{blue}{0.88$^{\ast}$} & \color{blue}{0.75$^{\ast}$} & \color{blue}{1.51$^{\ast}$} & \color{blue}{0.90$^{\ast}$} & \color{blue}{0.92$^{\ast}$} \\
-  \textbf{Influence} & &  &  &  &  &  \\ 
-  $\;\;\;\;$ Influence attribution & \color{blue}{$+$} & \color{blue}{1.20$^{\ast}$} & \color{blue}{1.29$^{\ast}$} & 0.08 & \color{blue}{1.00$^{\ast}$} & \color{blue}{1.09$^{\ast}$} \\ 
-  $\;\;\;\;$ Alter's influence indegree & \color{blue}{$+$} & \color{blue}{0.10$^{\ast}$} & \color{blue}{0.11$^{\ast}$} & 0.01 & \color{blue}{0.21$^{\ast}$} & \color{blue}{0.11$^{\ast}$} \\ 
-  $\;\;\;\;$ Influence absolute diff. & \color{red}{$-$} & \color{red}{-0.03$^{\ast}$} & \color{red}{-0.06$^{\ast}$} & 0.04 & \color{red}{-0.05$^{\ast}$} & \color{red}{-0.07$^{\ast}$} \\ 
-  $\;\;\;\;$ Alter = Government actor & \color{blue}{$+$} & \color{blue}{0.63$^{\ast}$} & 0.68 & -0.46 & \color{blue}{1.04$^{\ast}$} & 0.55 \\ 
-  \textbf{Functional requirements} &  &  &  &  &  & \\ 
-  $\;\;\;\;$ Ego = Environmental NGO & \color{blue}{$+$} & \color{blue}{0.88$^{\ast}$} & 0.99 & -0.60 & \color{blue}{0.79$^{\ast}$} & 0.67 \\ 
-  $\;\;\;\;$ Same actor type & \color{blue}{$+$} & \color{blue}{0.74$^{\ast}$} & \color{blue}{1.12$^{\ast}$} & \color{blue}{1.17$^{\ast}$} & \color{blue}{0.99$^{\ast}$} & \color{blue}{1.04$^{\ast}$} \\
-   \hline
-\end{tabular}
-\endgroup
-\end{table}
+Assume that the true data-generating process for a particular $Y$ is given by:
+
+\vspace{-8mm}
+\begin{align}
+  y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
+  \label{eqn:sim}
+\end{align}
+
+We compare inference for $\mu$ and $\beta$---the latter parameter would be of primary concern for applied scholars---using three models:
+
+\begin{itemize}
+  \item the ``standard'' international relations approach estimated through a typical generalized linear model; 
+  \item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);
+  \item and an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$. 
+\end{itemize}
 
@@
 \frame{
-  \frametitle{Latent Factor Visualization}
+  \frametitle{Bias Comparison}
+
+We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$:
 
 \centering
-  \vspace{-10mm}
-  \includegraphics[width=.7\textwidth]{ameFitSR_2_UV}
+\includegraphics[width=1\textwidth]{ameSimBias_all.pdf}
 
@@
 \frame{
-  \frametitle{Out of Sample Performance Assessment}
+  \frametitle{Coverage Comparison}
 
-\begin{itemize}
-  \item - Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
-  \item - For each $s \in \{1, \ldots, S\}$:
-  \begin{itemize}
-    \item - Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
-    \item - For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
-  \end{itemize}
-\end{itemize}
+Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals:
 
-This procedure generates a sociomatrix of out-of-sample predictions of the observed data
+\centering
+\includegraphics[width=1\textwidth]{ameSimCover_all.pdf}
 
@@
 \frame{
-  \frametitle{Performance Comparison}
+  \frametitle{What about the missing variable?}
 
-\begin{tabular}{cc}
-\includegraphics[width=.5\textwidth]{roc_outSample} & 
-\includegraphics[width=.5\textwidth]{rocPr_outSample} 
+Unobserved dependencies should be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$:
+
+\centering
+\includegraphics[width=1\textwidth]{ameSimCorr.pdf}
+
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{Empirical Applications}
+
+Descriptive information about the replicated studies:
+
+\footnotesize{
+\begin{tabular}{lccccc}
+	& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
+	Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed & Robust \\ 
+	Weeks (2012) & Logit & 1946--1999 & 197 & Directed & Robust \\
+	Gibler (2017) & Logit & 1816--2008 & 193 & Undirected & Robust \\ \bottomrule
+\end{tabular}
+}
+
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{Predictive Comparison}
+
+\centering   
+\begin{tabular}{lccc}
+~ &  Model & AUC (ROC) & AUC (PR) \\ \toprule
+\multirow{2}{*}{Reiter \& Stam (2003)} & \cellcolor{blue!10}AME & \cellcolor{blue!10}0.96 & \cellcolor{blue!10}0.15 \\
+~ & GLM & 0.92 & 0.08 \\
+\multirow{2}{*}{Weeks (2012)} & \cellcolor{blue!10}AME & \cellcolor{blue!10}0.97 & \cellcolor{blue!10}0.15 \\
+~ & GLM & 0.64 & 0.00 \\
+\multirow{2}{*}{Gibler (2017)} & \cellcolor{blue!10}AME & \cellcolor{blue!10}0.91 & \cellcolor{blue!10}0.08 \\
+~ & GLM & 0.52 & 0.00 \\ \bottomrule
 \end{tabular}
@@
 \frame{
-  \frametitle{Network Dependencies}
+  \frametitle{Results robust?}
 
+\footnotesize{
+\begin{table}[ht]
 \centering
-  \includegraphics[width=1\textwidth]{netPerfCoef}
+  \begin{tabular}{l p{4cm} l} \toprule
+    \multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
+    & &  accounting for dependencies? \\ \toprule
+    Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Confirmed} \\ \midrule
+    Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
+    Gibler (2017) & Power Parity at Time of Entry to International System Increases Conflict & {Unconfirmed}\\ \bottomrule
+  \end{tabular}
+\end{table}
+}
 
@@
 \frame{
-  \frametitle{What's Next?}
+  \frametitle{What else did we learn? Reiter \& Stam (2003)}
 
-  \begin{itemize}
-    \item - Extending AMEN to handle changing actor compositions
-    \item - Generalize multiplicative effects estimation over tensors
-  \end{itemize}
+\begin{tabular}{cc}
+  \includegraphics[width=.7\textwidth]{reiter_stam_aEff_map.pdf} &
+  \includegraphics[width=.3\textwidth]{reiter_stam_aEff_line.pdf} \\
+\end{tabular}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{What else did we learn? Weeks (2012)}
+
+\centering
+\includegraphics[width=.7\textwidth]{weeks_circPlot.png}
+
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{What else did we learn? Gibler (2017)}
+
+\centering
+\includegraphics[width=\textwidth]{gibler_margeff.pdf}
+
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \frame{
+%   \frametitle{What's Next?}
+%
+%   \begin{itemize}
+%     \item
+%   \end{itemize}
+%
+% }
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \frame{
-  \frametitle{Standard Network Dependence Measures}
-\vspace{-5mm}
-\includegraphics[width=1\textwidth]{ggGofAll_preeze}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \frametitle{Predictive Comparison: Reiter \& Stam (2003)}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-  \frametitle{Simulation Comparison}
-
-\includegraphics[width=1\textwidth]{ameVergmSim.png}
-
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-  \frametitle{AMEN v LSM Performance}
-
-  \begin{tabular}{cc}
-  \includegraphics[width=.5\textwidth]{roc_latSpace_outSample} & 
-  \includegraphics[width=.5\textwidth]{rocPr_latSpace_outSample}
+\centering
+\begin{tabular}{lr}
+  \includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf} & 
+  \includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf} \\
 \end{tabular}
@@
 \frame{
-  \frametitle{AMEN versus LSM Net Dependence}
+  \frametitle{Predictive Comparison: Weeks (2012)}
 
-  \includegraphics[width=1\textwidth]{netPerfCoef_latSpace}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\frame{
-  \frametitle{AMEN varying $K$ Performance}
-
-  \begin{tabular}{cc}
-  \includegraphics[width=.5\textwidth]{roc_ameSR_outSample} & 
-  \includegraphics[width=.5\textwidth]{rocPr_ameSR_outSample}
+\centering
+\begin{tabular}{lr}
+  \includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf} & 
+  \includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf} \\
 \end{tabular}
@@
 \frame{
-  \frametitle{AMEN varying $K$ Net Dependence}
+  \frametitle{Predictive Comparison: Gibler (2017)}
 
-  \includegraphics[width=1\textwidth]{netPerfCoef_ameSR}
+\centering
+\begin{tabular}{lr}
+  \includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf} & 
+  \includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf} \\
+\end{tabular}
+
 }

### 2018-08-30 prezzy/netsMatter.tex
need a concl slide
minor edits and reshuffling

@@
   \vspace{-10mm}
-  Relational data consists of 
+  Much of international relations data consists of
   \begin{itemize}
-    \item - a set of units or nodes
-    \item - a set of measurements, $y_{ij}$, specific to pairs of nodes $(i,j)$ 
+    \item a set of units or nodes
+    \item a set of measurements, $y_{ij}$, specific to pairs of nodes $(i,j)$ 
   \end{itemize}
@@
   \centering
-  \includegraphics[width=1.05\textwidth]{df_adj_net3}
+  \includegraphics[width=.5\textwidth]{dyadDF}
 }
@@
 
+The approach that we will discuss here to deal with dependencies is the Additive and Multiplicative Effects (AME) model (Hoff 2015; Minhas, Hoff \& Ward 2018)
+
 \begin{itemize}
@@
 
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+  \frametitle{Dependencies in relational data}
+
+  \vspace{-10mm}
+  To start to think about dependencies that arise in relational data ...
+  \vspace{5mm}
+
+  \centering
+  \includegraphics[width=1.05\textwidth]{df_adj_net3}
 }
@@
 
-The full AME framework enables us to not only account for the types of dependencies that often arise in dyadic data but also estimate them:
+The full AME framework enables us to not only account for the types of dependencies that often arise in dyadic data but also estimate them in a fully Bayesian framework:
 
@@
 
+\vspace{8mm}
 \centering
@@
 
-\footnotesize{
-\begin{tabular}{lccccc}
-	& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
-	Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed & Robust \\ 
-	Weeks (2012) & Logit & 1946--1999 & 197 & Directed & Robust \\
-	Gibler (2017) & Logit & 1816--2008 & 193 & Undirected & Robust \\ \bottomrule
+\vspace{5mm}
+% \footnotesize{
+% \begin{tabular}{lccccc}
+% 	& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
+% 	Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed & Robust \\
+% 	Weeks (2012) & Logit & 1946--1999 & 197 & Directed & Robust \\
+% 	Gibler (2017) & Logit & 1816--2008 & 193 & Undirected & Robust \\ \bottomrule
+% \end{tabular}
+% }
+\begin{tabular}{lcccc}
+	& Model &  Date Range & N. Actors  & Dyads Type \\ \toprule
+	Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed \\ 
+	Weeks (2012) & Logit & 1946--1999 & 197 & Directed \\
+	Gibler (2017) & Logit & 1816--2008 & 193 & Undirected \\ \bottomrule
 \end{tabular}
-}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\frame{
+\frametitle{Next steps}
+
+\begin{itemize}
+	\item Dealing with time varying dependence structure
+	\item Parsing out formation, persistence and dissolution
+	\item Many other extensions (bipartite, endogenous multilayer networks, ...)
+	\item Computational issues ... Bayesian models take time to converge 
+\end{itemize}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-10-02 paper/1_intro.tex
minor mods, need to fix gibler

@@
 
-In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2016:arxiv}.
+In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2018-10-02 paper/2_model.tex
minor mods, need to fix gibler

@@
 It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2018} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
-For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:ward:2004}.
+For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:2005}.
 The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.} The AME model is specified as follows:

### 2018-10-02 paper/4_app.tex
minor mods, need to fix gibler

@@
 
-We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017} published in the \textit{American Political Science Review}.
+We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017}.
 Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for the behavior of actors in the system.
@@
 Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
-Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the fixed effects in the model would suggest.
+Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
 None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.
@@
 There is also a striking difference between Weeks' original model and our estimation using AME in terms of capturing the data generating process.
-Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same fixed effects.
+Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same exogenous covariates.
 As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest.

### 2018-10-02 paper/4_app.tex
reframed gibler rep

@@
 		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
-		Gibler (2017) & Power Parity at Time of Entry to International System Increases Conflict & {Unconfirmed}\\ \bottomrule
+		Gibler (2017) & Entry to the International System, Not Power Parity, Determines Conflict & {Partially Confirmed}\\ \bottomrule
 	\end{tabular}
@@
 Gibler argues that the long-standing relationship between the relative parity of capabilities and initiation of militarized interstate disputes (MIDs) is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
+In most specifications, when you take into account the initial conditions for dyadic entry, the statistical significance of power parity vanishes.
 This finding calls into question many IR theories about the role of balance in generating international conflict \citep{organski:1958}.
 
-To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for his primary variable: parity of dyad members at the year in which they enter the international system.
-However, when we reestimate using AME, this measure becomes too imprecisely measured for using to draw any inferences about its effect on the initiation of MIDs.
-Not only is our estimate of the effect of Gibler's parity variable small, but it has a very large relative standard error---over a magnitude larger than the parameter itself.
+To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for both the insignificance of contemporary power parity, and the effect of the initial conditions for entry on driving conflict behavior.
+When we reestimate using AME power parity still has a small, imprecisely measured effect, supporting Gibler's argument that contemporary power parity is an artifact of other aspects of state relations, and does not drive contemporary conflict behavior.
+At the same time, the AME approach finds that the effect of power parity at entry into the international system becomes too imprecisely measured for using to draw any inferences about its effect on the initiation of MIDs.
+Not only is our estimate of the effect of this variable small, but it has a very large relative standard error---over a magnitude larger than the parameter itself.
+Thus, while we can support the claims that, when accounting for both initial conditions and network dependencies, the effect of power parity vanishes, we are able to find less support for the claims about the effects of initial conditions.
 

### 2018-10-02 paper/1_intro.tex
CD edits

@@
 Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data.
-Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature (published in this journal).
+Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature.
 Our comparison reveals that by accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables.
@@
 
-Our AME approach advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
-Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena without abandoning a regression based approach, but at the same time account for the data generating processes behind such events of interest.
+The AME approach advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest.
+Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena without abandoning a regression based approach, while at the same time accounting for the data generating processes behind such events of interest.
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

### 2018-10-02 paper/2_model.tex
CD edits

@@
 This makes sense if each observation is independent of the others.
-For example, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.
+For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.
 Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
@@
 
-These types of patterns frequently emerge in IR contexts as well.\footnote{For example, see: \citet{manger:etal:2012, kinne:2013, chyzh:2016}.} For example, a perennial finding in the interstate trade literature emphasizes the role that geography plays in determining trade flows.
+These types of patterns frequently emerge in IR contexts.\footnote{For example, see: \citet{manger:etal:2012, kinne:2013, chyzh:2016}.} For example, a perennial finding in the interstate trade literature emphasizes the role that geography plays in determining trade flows.
 Geographic proximity in the network context is an example of homophily---a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place.
@@
 While sometimes the conditions driving these patterns, such as geography, are easy to identify, at other times it can be difficult to describe exactly why higher order dependence patterns in networks may develop.
-%CD to SM ^
 
@@
 We model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
-$f$ is typically a mapping function, and can be one that applies to dichotomous, ordinal, or continuous distribitions.
+$f$ is typically a mapping function, and can be one that applies to dichotomous, ordinal, or continuous distributions.
 

### 2018-10-02 paper/4_app.tex
CD edits

@@
 We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
-Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces, is well-known and posited an hypothesis in which interdependencies are consequential.
+Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.
 Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model.
@@
 The authors find that coefficient of the Pers/Democ Directed Dyad indicator is positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction.
-In our re-estimation using the AME framework, we confirm these results, indicating that dictators are more likely to initiate or engage in conflict with democratic regimes but not vice versa.
+In our re-estimation using the AME framework, we confirm these results, indicating that dictators are likely to initiate or engage in conflict with democratic regimes but not vice versa.
 
@@
 When we reestimate using AME power parity still has a small, imprecisely measured effect, supporting Gibler's argument that contemporary power parity is an artifact of other aspects of state relations, and does not drive contemporary conflict behavior.
-At the same time, the AME approach finds that the effect of power parity at entry into the international system becomes too imprecisely measured for using to draw any inferences about its effect on the initiation of MIDs.
+At the same time, the AME approach finds that the effect of power parity when a country enters into the international system is too imprecisely measured to draw any inferences about its effect on the initiation of MIDs.
 Not only is our estimate of the effect of this variable small, but it has a very large relative standard error---over a magnitude larger than the parameter itself.
-Thus, while we can support the claims that, when accounting for both initial conditions and network dependencies, the effect of power parity vanishes, we are able to find less support for the claims about the effects of initial conditions.
+Thus, while we can confirm the argument that the effect of power parity vanishes when accounting for both initial conditions and network dependencies, we find less support for the argument that initial conditions drive the initiation of MIDs.
 
@@
 
-First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data..
+First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data.
 In each of the replications, the AME's predictive performance out-of-sample is notably superior.

### 2018-10-02 paper/netsMatter.tex
CD edits

@@
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-Peter Hoff, John Ahlquist, Robert Franzese, Arturas Rozenas, and Benjamin Appel provided welcome and helpful comments.
+Peter Hoff, John Ahlquist, Robert Franzese, Arturas Rozenas, Betsy Sinclair, Jacob Montgomery and Benjamin Appel provided welcome and helpful comments.
 We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.
@@
 \begin{abstract}
-Much of international relations scholarship concerns dyads.
-Dyadic hypotheses and especially dyadic data abound.
-Standard modeling approaches fail to adequately capture the data generating process behind dyadic events and processes.
+International relations scholarship concerns dyads, yet standard modeling approaches fail to adequately capture the data generating process behind dyadic events and processes.
 As a result, they suffer from biased coefficients and poorly calibrated standard errors.

### 2019-07-02 paper/netsMatter.tex
sync up

--- 2018-04-12 paper/netsMatter.tex
@@
 
-% % \date{\today}
+% \date{\today}
 
-%\thanks{Shahryar Minhas and Michael D.
+\thanks{Shahryar Minhas and Michael D.
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
-John Ahlquist, Robert Franzese, and Arturas Rozenas provided welcome and helpful comments.
+Peter Hoff, John Ahlquist, Robert Franzese, Arturas Rozenas, Betsy Sinclair, Jacob Montgomery and Benjamin Appel provided welcome and helpful comments.
 We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}.

### 2019-07-12 paper/coverletter/netsMatter_coverLetter.tex
syncing

@@
 
-{Editorial Team of American Political Science Review via submission portal}
+{Editorial Team of Political Science Research \& Methods via submission portal}
 
@@
 Although we clearly work in the latent space world, this paper is not about engaging or extending the ERGM-latent space discussion.
-Rather, we evaluate the utility of the AME framework using a simulation based exercise and a replication of three recently published works from the field of International Relations.
+A piece by Minhas, Hoff, and Ward published this year already delves into the differences between various network approaches.
+This piece, however, is meant to evaluate the utility of the AME framework using a novel simulation based exercise and a replication of three recently published works from the field of International Relations.
 Through the simulation exercise, we show that in the presence of unobserved dependencies the model is able to provide less biased estimates and better calibrated standard errors than extant approaches in the literature.
@@
 
-\includegraphics [scale=.8]{/Users/s7m/Dropbox/Finances/signature.png}
+\includegraphics [scale=.8]{/Volumes/Samsung_X5/Dropbox/Finances/signature.png}
 

### 2019-07-12 paper/netsMatter.tex
syncing

@@
 
-\thanks{Shahryar Minhas and Michael D.
+% \thanks{Shahryar Minhas and Michael D.
 Ward acknowledge support from National Science Foundation (NSF) Award 1259266.
@@
 Further, conventional methods misstate the effect of key variables and the uncertainty in these effects.
-Finally, AME dominates standard approaches in terms of out-of-sample fit.
+Finally, AME outperforms standard approaches in terms of out-of-sample fit.
 In sum, our work shows the consequences of failing to take the dependencies inherent to dyadic data seriously.
-Most importantly, by better modeling the data generating process underlying political phenomena, the AME framework improves scholars' ability to conduct inferential analyses.
+Most importantly, by better modeling the data generating process underlying political phenomena, the AME framework improves scholars' ability to conduct inferential analyses on dyadic data.
 \end{abstract}

### 2019-07-12 paper/coverletter/netsMatter_coverLetter.tex
cleaning for new submission

@@
 The manuscript ``Taking Dyads Seriously'' introduces the Additive and Multiplicative Effects (AME) framework for conducting inference in the context of the dependencies that we often observe in dyadic data.
-In submitting this paper, we feel as though it would be helpful to first provide some background information.
+In submitting this paper, we feel as though it is necessary to first provide some background information.
 \\[1ex]
 
-Among political scientists using statistical models to analyze network data there has emerged a dispute between the advocates of the Exponential Random Graph (ERGM) model and those building on the ``latent space'' approach pioneered by Hoff, Raftery, and Handcock.
-This fight is not mirrored in the statistical or broader networks literature, as in those communities people recognize these models as having distinct goals.
+Among some political scientists using statistical models to analyze network data there has emerged a dispute between the advocates of the Exponential Random Graph (ERGM) model and those building on the ``latent space'' approach pioneered by Hoff, Raftery, and Handcock.
+This unproductive fight is not mirrored in the statistical or broader networks literature, as in those communities people recognize these models as having quite distinct goals.
+These models having differing goals is something that I also believe and which is why in a recent ERGM related article I reviewed for PSRM, I did not express any need for the authors to engage in this dispute.
 \\[1ex]
@@
 Although we clearly work in the latent space world, this paper is not about engaging or extending the ERGM-latent space discussion.
-A piece by Minhas, Hoff, and Ward published this year already delves into the differences between various network approaches.
+A Political Analysis piece by Minhas, Hoff, and Ward published this year already delves into the differences between various network approaches.
 This piece, however, is meant to evaluate the utility of the AME framework using a novel simulation based exercise and a replication of three recently published works from the field of International Relations.
@@
 Next, for nearly all of the replications reestimated using the AME framework we find that the case for the key findings from the original works become much less compelling.
-Additionally, we show through an out-of-sample cross-validation exercise that the AME approach uniformly outperforms each of the replicated approaches in capturing the data generating process of the event of interest.\\[1ex]
+Additionally, we show how the AME can actually be used by IR scholars to glain meaningful insights from the dependencies that estimated by the model.\\[1ex]
 
@@
 
-\includegraphics [scale=.8]{/Volumes/Samsung_X5/Dropbox/Finances/signature.png}
+\includegraphics [scale=.6]{/Volumes/Samsung_X5/Dropbox/Finances/signature.png}
 

### 2019-07-12 paper/netsMatter.tex
cleaning for new submission

@@
 As a result, they suffer from biased coefficients and poorly calibrated standard errors.
-We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the inherent dependencies in dyadic data.
+We show how a regression-based approach, the Additive and Multiplicative Effects (AME) model, can be used to account for the inherent dependencies in dyadic data and glean substantive sights in the interrelations between actors.
 First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference on dyadic data.

### 2019-07-13 paper/1_intro.tex
sync

@@
 
-In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
+In this article, we discuss how the Additive and Multiplicative Effects (AME) model can be used to model the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2019-07-15 paper/1_intro.tex
submit

@@
 
-In this article, we discuss how the Additive and Multiplicative Effects (AME) model can be used to model the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
+In this article, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
 We focus on three types of interdependencies that can complicate dyadic analyses.
@@
 For example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions.
-The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.
+The multiplicative effects capture these sorts of dependencies, specifically, those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.
 

### 2019-07-15 paper/5_concl.tex
submit

@@
 We then re-analyzed each study using the AME model.
-In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, and c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.} In several cases, the new approach overturns the basic findings of the original research.
+In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, and c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.}
 
-It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.
+% It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.
 Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citealp{king:roberts:2014}).

### 2020-01-17 code/1_simulation/3_ameCalib.R
move repl code to my cluster, org revision memo, start to assign tasks

@@ -7,8 +7,16 @@ if(Sys.info()['user']=='s7m'){
 	graphicsPath = paste('~/Research/netsMatter/paper/')
 	source(paste0(fPath, 'functions.R')) }
 
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R')) }
+
 toLoad = c(
-	'devtools', 
+	'devtools',
 	'foreach', 'doParallel',
 	'magrittr', 'dplyr', 'ggplot2',
 	'latex2exp', 'Cairo'
@@ -39,7 +47,7 @@ getCoverage = function(ameSim, model, n, varName, varNum, actual){
 	intSumm$n = n
 	intSumm$varName = varName
 	intSumm$actual = actual
-	return(intSumm)	
+	return(intSumm)
 }
 
 #
@@ -88,13 +96,13 @@ ggCoverPlot = function(var,h=3, w=8){
 	if(var=='all'){ g=ggplot(coverSumm, aes(x=model, y=coverage, fill=model,color=model)) }
 	g = g +
 		geom_hline(aes(yintercept=.95), color='grey60', size=4, alpha=.5) +
-		geom_linerange(aes(ymin=0, ymax=coverage),size=1.25) + 
-		geom_point(size=2) + 
+		geom_linerange(aes(ymin=0, ymax=coverage),size=1.25) +
+		geom_point(size=2) +
 		facet_grid(varName~n, scales='free_y',
-			labeller=as_labeller(facet_labeller, default = label_parsed)) + 			
-		xlab('') + 
+			labeller=as_labeller(facet_labeller, default = label_parsed)) +
+		xlab('') +
 		scale_color_manual(values=modCols) +
-		scale_fill_manual(values=modCols) +		
+		scale_fill_manual(values=modCols) +
 		scale_y_continuous('', breaks=seq(0,1.2,.2), labels=seq(0,1.2,.2), limits=c(0,1)) +
 		annotate('text', x=1, y=.95, label='95% CI', color='black', size=2.5, fontface='bold') +
 		geom_text(aes(label=covLab, y=covLabY), hjust=-.3, size=2.5, fontface='bold', color='black') +
[…]

### 2020-01-17 code/2_codeForPaperGraphics/gibler_posterior.R

@@ -3,13 +3,26 @@ rm(list=ls())
 
 if(Sys.info()['user'] %in% c('s7m', 'janus829')){
 	resultsPath = '~/Dropbox/Research/netsMatter/replications/gibler_2017/outputData/'
-	plotPath = '~/Research/netsMatter/paper/graphics/' }
+	plotPath = '~/Research/netsMatter/paper/graphics/'
+	source('~/Research/netsMatter/code/helpers/functions.R')
+	source('~/Research/netsMatter/code/helpers/ameHelpers.R')
+	source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
+	source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
+	source('~/Research/netsMatter/code/helpers/clusteredSE.R')
+}
 
-source('~/Research/netsMatter/code/helpers/functions.R')
-source('~/Research/netsMatter/code/helpers/ameHelpers.R')
-source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
-source('~/Research/netsMatter/code/helpers/stargazerHelpers.R')
-source('~/Research/netsMatter/code/helpers/clusteredSE.R')
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R'))
+	source(paste0(fPath, 'ameHelpers.R'))
+	source(paste0(fPath, 'binPerfHelpers.R'))
+	source(paste0(fPath, 'stargazerHelpers.R'))
+	source(paste0(fPath, 'clusteredSE.R'))
+}
 ############################################
 
 # modData ###########################################
@@ -39,9 +52,9 @@ modList = list(glmSumm, ameSumm)
 modNames = c('GLM (Logit)', 'AME')
 varKey = data.frame(
 	dirty=names(coef(mod)),
-	clean=c( '(Intercept)', 
+	clean=c( '(Intercept)',
 		'Allied', 'Joint Democracy',
-		'Peace Years', 'Spline 1', 'Spline 2', 
[…]

### 2020-01-17 revisionMemo/rev1.tex
move repl code to my cluster, org revision memo, start to assign tasks

\section*{Reviewer 1}

\begin{enumerate}
	\item P1: ``Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases. In practice, such strategies rarely work because they do not directly address the fundamental data generating process''.
	\begin{itemize}
		\item The term “rarely work” is too casual.  What problem exactly is motivating this paper?  E.g., the Aronow et al. paper cited at the start of the same paragraph looks at consistent variance estimators for semiparametric regression estimators on dyadic data. By the standards set out by those authors in that paper, the strategies “work” just fine.  I suppose then that the current paper has a different sense of what it means for something “to work.” For example, I am supposing that the current paper is talking about people misconstruing the structural interpretation of a reduced form parameter estimated in a GLM fit to dyadic data.  That would be wholly orthogonal to what Aronow et al. were considering, but then this needs explication.
	\end{itemize}
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item P2: “Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another. This makes sense if each observation is independent of the others.”
	\begin{itemize}
		\item Revise this opening, since it is perfectly fine to stack observations and then incorporate auxiliary information (such as adjacency matrix) to characterize dependencies between rows.
	\end{itemize}
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item P3: “Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.”
	\begin{itemize}
		\item This is another example of language that is too causal and imprecise, like the opening paragraph. What are “parameter estimates from standard statistical models”?  And “biased” with respect to what target of inference? Again, I think the authors have in mind “bias” with respect to parameters that are meant to have a particular structural interpretation, but this needs to be clarified.
		\item Generally, section 2 should be clearer in motivating the problem.  Provide us with a toy example that illustrates the kind of “bias” that the paper has in mind from “standard models”.  For me at least, after reading section 2 I still wasn’t sure what problem the paper is actually trying to address.
	\end{itemize}
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item P7: Regarding the specification, this augments a dyadic random effects model, which is quite common, with  the LFM component, which resembles an interactive fixed effects specification (cf. Bai 2009, Ecta) and is based on an SVD factorization of the observed network.
	\begin{itemize}
		\item It is useful to think of the infeasible estimator and the implications for interpreting the LFM component.  The infeasible estimator takes the transformation f(.), fixed effects parameters, and SRRM random effects as know, in which case the LFM amounts to predictions from a factor model fit to the adjacency matrix of residuals on the scale of the latent outcome.  Would this be a reasonable interpretation? If so it would be useful to draw the connection to factor models commonly in use these days (such as the interactive fixed effects model of Bai, which is the basis of, e.g., generalized synthetic control methods).
	\end{itemize}
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item P8: “accounting for this structure is necessary if we are to adequately represent the data generating process.” Continuing with the problems I have with the language above, what does “adequately represent” mean?. P9: “a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.”
	\begin{itemize}
		\item Okay here is where the paper are starting to clarify goals.  So a) is about causal identification.  But as I understand, the solution here is parametric identification of causal effects.  This means that the issue of misspecification and robustness ought to be taken up.  Similarly, b) is addressed in this paper under a fairly simple parametric data generating process. But again, how robust is the proposed model to situations where the real DGP does not abide by the parametric assumptions (linearity, normally distributed random effects, and a factor model for the residuals of dimension K)? For c), I think most would agree that the “proof” is simply in comparison to whatever other good models are available.  Not however, that in social science applications, forecasting and therefor c) are rarely of interest (at least not in the types of research that appear in political science journals).  As such, from the perspective of what the researchers in discipline mostly cares about, a) and b) are of primary importance.
	\end{itemize}
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item P9-10: we need to know exactly how and from what distributions X and W are being drawn to interpret the results here.  Are they independent/orthogonal or no?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		SM will deal with this.
		}}
	\end{itemize}
	\item Section 4 needs to incorporate an assessment of robustness to misspecification.  See the Aronow et al. paper for an example of how to do this, where they study the performance of a dyadic random effects model.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		SM will deal with this.
		}}
	\end{itemize}
\end{enumerate}

### 2020-01-17 code/2_codeForPaperGraphics/reiter_stam_posterior.R

@@ -16,6 +16,19 @@ if(Sys.info()['user'] == 'juanftellez'){
   source('~/OneDrive/netsMatter/code/helpers/ameHelpers.R')
   source('helperEx.R')
 }
+
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R'))
+	source(paste0(fPath, 'ameHelpers.R'))
+	source(paste0(fPath, 'binPerfHelpers.R'))
+	source(paste0(fPath, 'stargazerHelpers.R'))
+	source(paste0(fPath, 'clusteredSE.R'))
+}
 ############################################
 
 # modData ###########################################
@@ -42,9 +55,9 @@ varKey = data.frame(
 	dirty=names(coef(mod)),
 	clean=c('Intercept', 'Pers/Democ Directed Dyad',
 		'Democ/Pers Directed Dyad',
-        'Personal', 'Military', 'Single', 'Democracy', 
-        'Contiguous', 'Major Power', 'Ally', 'Higher/Lower Power Ratio', 
-        'Economically Advanced', 'Years Since Last Dispute', 'Cubic Spline 1', 
+        'Personal', 'Military', 'Single', 'Democracy',
+        'Contiguous', 'Major Power', 'Ally', 'Higher/Lower Power Ratio',
+        'Economically Advanced', 'Years Since Last Dispute', 'Cubic Spline 1',
         'Cubic Spline 2', 'Cubic Spline 3'),
 	stringsAsFactors = FALSE )
 
@@ -102,8 +115,8 @@ ggMap = ggplot() +
     axis.text = element_blank() )
 ggMap
 ggsave(ggMap, file=paste0(plotPath, 'reiter_stam_aEff_map.pdf'), width=6, height=3)
-system(paste('pdfcrop', 
-  paste0(plotPath, 'reiter_stam_aEff_map.pdf'), 
+system(paste('pdfcrop',
+  paste0(plotPath, 'reiter_stam_aEff_map.pdf'),
   paste0(plotPath, 'reiter_stam_aEff_map.pdf')))
 
 # pick out top countries
[…]

### 2020-01-17 revisionMemo/rev2.tex
move repl code to my cluster, org revision memo, start to assign tasks

\section*{Reviewer 2}

\begin{enumerate}
	\item The additive and multiplicative effects (AME) model presented in the paper allows for a better treatment of first, second and third order dependencies in the data, and account for unobserved sources of bias in the data. The presentation of the AME estimator is clear, and the simulations provide clear evidence that the model outperforms naive OLS/GLM models fitted to the same data. The paper will be of great interest to the PSRM audience and is likely to have a strong impact on future research in international relations. I do, however, have a few suggestions for revisions that will hopefully strengthen the paper’s contributions.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank the reviewer for their comments and suggestions.
		}}
	\end{itemize}
	\item The paper shows that the AME model performs well in comparison to the “oracle” model where all sources of dependencies in the data are accounted for, which is reassuring. Yet this comparison feels incomplete. First, the naive model is clearly misspecified given the underlying data generating process. A large number of dependencies in international relations are linked to geography. What would happen if the dependencies were modeled using spatial econometric techniques?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item It would help if the author discussed in more depth the properties of the AME estimator in the presence of unobserved dependencies. One would imagine that the oracle model, where all the unobserved factors and dependencies are accounted for should always outperform a parametric or semi-parametric estimator. Relatedly, it also would help to have a more technical discussion of the properties of AME models, particularly in terms of the bias-variance trade off.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item The author could also underscore a related issue, which usually escapes empirical IR scholars: the importance of developing theories that account for the direct and indirect effects of interactions in a setting where the units are interconnected, and where the effects on one unit have the potential to impact other units in the system. A traditional approach is to throw as many observable variables as possible as controls, without much consideration on where they fit in the model. As the comparison of the AME and oracle models suggests, a careful modeling of spillovers and dependencies can result in better estimates of the underlying relationships.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item The paper emphasizes on the importance of estimating the random effects for theory building and hypothesis test. A longer elaboration and illustration of this point would be very helpful.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item I would like to see a more detailed discussion of the setup and choices made for the Monte Carlo simulations. What type of network do the simulated data model? How does the network in the simulation relate to typical networks in IR studies, and to the ones in the replicated papers.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		SM will deal with this.
		}}
	\end{itemize}
	\item The criteria for choosing the studies for replication is unclear and needs a better justification. I would have liked to see a larger set of replications particularly in areas and topics where the data generating process, the dependencies among units varies, or the relationship between the units is hierarchical, cross-classified or multilevel.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
	\item Given the importance of promoting the use of the AME method I would encourage the author to revise the tutorial presented in Appendix C to make it more accessible to a larger audience.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		Insert great response.
		}}
	\end{itemize}
\end{enumerate}

### 2020-01-17 revisionMemo/revisionLetter.tex
move repl code to my cluster, org revision memo, start to assign tasks

\documentclass[12pt,onesided,fullpage]{amsart}

%%% PAGE DIMENSIONS
\usepackage{multirow}
\usepackage{placeins}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
%\pdfpagewidth=8.5in % for pdflatex
%\pdfpageheight=11in % for pdflatex

\usepackage{graphicx} % support the \includegraphics command and options

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array,amsmath,graphicx,psfrag,amssymb,subfigure,tabularx,booktabs}
\usepackage{mparhack}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{multicol}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{placeins}
\setcitestyle{authordate,round,semicolon,aysep={,},yysep={,}}
\bibpunct[:]{(}{)}{;}{a}{,}{,}

% \doublespacing
\def\input@path{
	{C:/Users/Owner/Research/netsMatter/paper/graphics/},
	{C:/Users/herme/Research/netsMatter/paper/graphics/},
	{/Users/s7m/Research/netsMatter/paper/graphics/}
}

\graphicspath{
	{C:/Users/Owner/Research/netsMatter/paper/graphics/},
	{C:/Users/herme/Research/netsMatter/paper/graphics/},
	{/Users/s7m/Research/netsMatter/paper/graphics/},
}

\begin{document}
\singlespacing

\title[PSRM-OA-2019-0126]{Ms. No. PSRM-OA-2019-0126 entitled ``Taking Dyads Seriously''}

\date{\today~~Version 0.01}
\maketitle

$\hspace{-5mm}$Dear Editors, \\ [1ex]

Thank you for the opportunity to revise and resubmit our manuscript. We believe the manuscript has benefited from the Reviewers' helpful and thoughtful comments. We have revised the manuscript, taking seriously each individual point raised by the Reviewers. Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each of the points made by the Reviewers.

We hope you agree that the manuscript has improved through this process and we are looking forward to your response.\\ [1ex]

% Sincerely, \\ [1ex]

% The Authors

\include{rev1}
\include{rev2}

\newpage\tiny
\end{document}\bye

### 2020-01-17 code/1_simulation/2_ameBias.R

@@ -7,8 +7,16 @@ if(Sys.info()['user']=='s7m'){
 	graphicsPath = paste('~/Research/netsMatter/paper/')
 	source(paste0(fPath, 'functions.R')) }
 
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R')) }
+
 toLoad = c(
-	'devtools', 
+	'devtools',
 	'foreach', 'doParallel',
 	'magrittr', 'dplyr', 'ggplot2',
 	'latex2exp', 'Cairo'
@@ -72,14 +80,14 @@ ggBiasPlot = function(varName, h=4, w=8){
 			aes(x=model, y=value, fill=model,color=model)) }
 	if(varName=='all'){ g=ggplot(ameSimBias, aes(x=model,y=value,fill=model,color=model)) }
 	g = g +
-		geom_hline(aes(yintercept=act), color='grey60', size=2) +	
+		geom_hline(aes(yintercept=act), color='grey60', size=2) +
 		geom_jitter(alpha=.1) +
-		geom_boxplot( color='black') + 	
+		geom_boxplot( color='black') +
 		# geom_violin(
 		# 	draw_quantiles=c(0.025,0.5,0.975),
-		# 	trim=TRUE, color='black', alpha=.4) +	
+		# 	trim=TRUE, color='black', alpha=.4) +
 		facet_grid(var ~ n, scales='free_y',
-			labeller=as_labeller(facet_labeller, default = label_parsed)) + 
+			labeller=as_labeller(facet_labeller, default = label_parsed)) +
 		xlab('') + ylab('') +
 		scale_color_manual(values=modCols) +
 		scale_fill_manual(values=modCols) +
@@ -95,11 +103,11 @@ ggBiasPlot = function(varName, h=4, w=8){
 			strip.text.x = element_text(size=9, color='white'
 				# ,family="Source Code Pro Semibold"
 				),
-			strip.text.y = element_text(size=9, color='white' 
+			strip.text.y = element_text(size=9, color='white'
 				# ,family="Source Code Pro Semibold",
 				,angle=0
-				),		
[…]

### 2020-01-17 code/1_simulation/1_ameSim.R
move repl code to my cluster, org revision memo, start to assign tasks

@@ -7,8 +7,16 @@ if(Sys.info()['user']=='s7m'){
 	simResPath = paste0(dPath, 'simulation/')
 	source(paste0(fPath, 'functions.R')) }
 
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R')) }
+
 toLoad = c(
-	'devtools', 
+	'devtools',
 	'foreach', 'doParallel',
 	'magrittr', 'dplyr', 'ggplot2',
 	'sna', # mrqap
@@ -30,7 +38,7 @@ gamma = x2Eff
 imps = 1000
 imp=1
 # simRun = function(seed, n, mu, beta, gamma){
-	set.seed(seed) 
+	set.seed(seed)
 	xw = matrix(rnorm(n*2),n,2)
 	X = tcrossprod(xw[,1])  ; W = tcrossprod(xw[,2])
 	XW = array(dim=c(n,n,2)) ; XW[,,1] = X ; XW[,,2] = W
@@ -48,7 +56,7 @@ imp=1
 
 	# MRQAP
 	loadPkg('sna')
-	fitQAP = netlm(y=Y, x=X, 
+	fitQAP = netlm(y=Y, x=X,
 		intercept=TRUE, mode='digraph', diag=FALSE,
 		nullhyp='qap', reps=200)
 
@@ -98,7 +106,7 @@ imp=1
 	qapBeta = matrix(fitQAP$coefficients,ncol=2,nrow=1)
 	uv = list(naive=fit0$UVPM, ame=fit1$UVPM, oracle=fitO$UVPM)
 	u = list(naive=fit0$U, ame=fit1$U, oracle=fitO$U)
-	v = list(naive=fit0$V, ame=fit1$V, oracle=fitO$V) 
+	v = list(naive=fit0$V, ame=fit1$V, oracle=fitO$V)
 	gof = list(naive=fit0$GOF, ame=fit1$GOF, oracle=fitO$GOF)
 	out = list(
 		beta=beta, uv=uv, u=u, v=v, W=W, gof=gof,
@@ -115,7 +123,7 @@ intEff=-2 ; x1Eff=1 ; x2Eff=1
 #
[…]

### 2020-01-23 paper/6_appendix.tex
started response to reviewer comment on appendix c tutorial

@@
     \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network. 
+    \begin{itemize}
+      \item An adjacency matrix describes relationships between nodes in a particular year of data.
+For example, in an adjacency matrix of interstate MIDs, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had a MID in that year, and '0' otherwise.
+The diagonal in the adjacency matrix is typically missing.
+Y is a list of these adjacency matrices for the outcome variable, where each element in the list is a different year of data.
+    \end{itemize}
     \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset. 
+    \begin{itemize}
+      \item An array is a data object in R that contains a series of 'stacked' matrices for each year of data.
+An array of dimension (2, 3, 4), for example, contains 4 rectangular matrices each with 2 rows and 3 columns.
+Each matrix in the array describes the relationship between nodes with respect to some covariate.
+For example, for interstate alliances, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had an alliance in that year, and '0' otherwise.
+An array contains a matrix of this kind for each covariate going into the model.
+Xdyad is a list of these arrays, where each element in the list is a different year of data.
+    \end{itemize}
     \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
+    \begin{itemize}
+      \item Each matrix in Xrow has the nodes running down the rows and the covariates in the dataset running along the columns.
+An entry in the matrix captures the value that a node takes on for each covariate in a given year.
+For example, if column $j$ is GDP per capita and column $k$ is population, then row $i$, column $j$ measures country $i$'s GDP per capita and row $i$, column $k$ measures country $i$'s population in a given year.
+Xrow is a list of these matrices, where each element in the list is a different year of data.
+    \end{itemize}
 \end{itemize}
@@
     \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
+    \begin{itemize}
+      \item Xrow here is nearly identical to the undirected case.
+The only difference is that rather than each matrix containing covariate data on all nodes, in the directed context each matrix only contains data on the nodes that are acting as senders in that particular year.
+Using interstate MIDs as an example, if in 1983 only two countries initiated MIDs, Xrow would only contain data for those two countries (i.e., Xrow would only have two rows).
+    \end{itemize}
     \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset. 
+    \begin{itemize}
+      \item Xcol here is nearly identical to Xrow, except each matrix contains data on receiver nodes, not sender nodes.
+Using interstate MIDs as an example, if in 1983 only two countries had MIDs initiated against them, Xcol would only contain data for those two countries (i.e., Xcol would only have two rows).
+    \end{itemize}
 \end{itemize}

### 2020-01-29 revisionMemo/rev1.tex
CD try

@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		TO DO: The reviewer is correct in that our paper addresses the majority of IR scholarship that misuses GLM approaches for dyadic data.
+We have rewritten these sentences in the following way: In practice, such strategies remain problematic because they do not directly address the fundamental data generating process but instead seek to correct for dependence while using a baseline model that makes strong assumptions about the independence of observations in the DGP.
+While corrections such as X and Y are beneficial at times for A and B, these approaches remain only partially useful, and do not sufficiently capture third-order dependencies that organically arise from dyadic data.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		TO DO: We have revised this opening statement to clarify our position. 
 		}}

### 2020-01-29 revisionMemo/rev1.tex
CD attempts

@@
 		\item \textcolor{blue}{ \emph{
-		TO DO: The reviewer is correct in that our paper addresses the majority of IR scholarship that misuses GLM approaches for dyadic data.
+		TO DO/CD: The reviewer is correct in that our paper addresses the majority of IR scholarship that misuses GLM approaches for dyadic data.
 We have rewritten these sentences in the following way: In practice, such strategies remain problematic because they do not directly address the fundamental data generating process but instead seek to correct for dependence while using a baseline model that makes strong assumptions about the independence of observations in the DGP.
@@
 		\item \textcolor{blue}{ \emph{
-		TO DO: We have revised this opening statement to clarify our position. 
+		TO DO/CD: We have revised this opening statement to clarify our position. Max: I'm wondering if we should combine this with the Q above or is R1's question here distinct enough?
 		}}

### 2020-01-29 revisionMemo/rev2.tex
assign one more task
started response to reviewer comment on appendix c tutorial
CD attempts

@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		TODO/CD: We agree with this suggestion.
+We have now emphasized the importance of developing theory in our discussion of the AME and Oracle comparison (p.
+X), as well as in the conclusion of our paper (p.Y).
 		}}
 	\end{itemize}
-	\item The paper emphasizes on the importance of estimating the random effects for theory building and hypothesis test. A longer elaboration and illustration of this point would be very helpful.
+	\item The paper emphasizes the importance of estimating the random effects for theory building and hypothesis test. A longer elaboration and illustration of this point would be very helpful.
 	\begin{itemize}
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		TODO/CD: I think this comment and the above comment are essentially the same. We could even cite our JOP as an example of incorporating theory?  Perhaps other new-ish net paper as well?
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		TO DO/CD: We selected our cases based on a few criteria.
+We focus on studies that are explicitly about International Relations, were published in the last 10 years, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we also decided to focus on one journal, the APSR).
+We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.
+In addition, we selected studies that could help illustrate different key components of the AME approach (Max, more here on each?).
+While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, they are beyond the scope of this paper at this time.
+(CD: Should we also mention something like 'in consultation with the editor, we believe they are beyond the scope....")
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		The reviewer raises a valid point.
+We have expanded our tutorial in Appendix C, particularly in giving more detail and concrete examples when it comes to formatting the data, which may represent the biggest hurdle for some users.
+We also...
 		}}

### 2020-01-30 revisionMemo/rev1.tex
reviewer 1 issue

@@
 		\item \textcolor{blue}{ \emph{
-		TO DO/CD: We have revised this opening statement to clarify our position. Max: I'm wondering if we should combine this with the Q above or is R1's question here distinct enough?
+		TO DO/CD: We have revised this opening statement to clarify that the problem is treating dyads as independent, not stacking them.
 		}}

### 2020-01-30 paper/2_model.tex
submit
reviewer 1 issue

@@
 
-Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another.
-This makes sense if each observation is independent of the others.
+Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another, and treat each observation as independent of the others.
 For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.

### 2020-01-30 revisionMemo/rev1.tex
more efforts from CD and MG

--- 2020-01-29 revisionMemo/rev1.tex
@@
 		\item \textcolor{blue}{ \emph{
-		TO DO: The reviewer is correct in that our paper addresses the majority of IR scholarship that misuses GLM approaches for dyadic data.
-We have rewritten these sentences in the following way: In practice, such strategies remain problematic because they do not directly address the fundamental data generating process but instead seek to correct for dependence while using a baseline model that makes strong assumptions about the independence of observations in the DGP.
-While corrections such as X and Y are beneficial at times for A and B, these approaches remain only partially useful, and do not sufficiently capture third-order dependencies that organically arise from dyadic data.
+		TO DO in paper: Strategies like those employed by Aronow et al do not directly model all relevant dependencies in the data \footnote{As Aronow et al explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data." p.
+16.} but instead adjust the variance estimator.
+While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).
+The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		TO DO: We have revised this opening statement to clarify our position. 
+		 The point is well taken, we have clarified this section (on page 2) to point to the assumption of independence as the primary problem, not the stacking of observations in and of itself.
 		}}
@@
 For me at least, after reading section 2 I still wasn’t sure what problem the paper is actually trying to address.
+		\item \textcolor{blue}{ \emph{ MG/CD: We think the reviewer makes a good suggestion here to clarify, mathematically or in some other illustrative way, exactly how we think that dependencies bias the estimator.
+For example we can say that our target of inference is the estimation of beta (with some variance).
+If a coefficient is correlated with a dependency, then we could obtain an estimate for beta that is artificially high (and the same is true for the reverse).}}
+	\begin{itemize}
+	\item  \textcolor{blue}{ \emph{ What we think: True DGP: $Y = B0 + XB1 + a + b +$ 3rd order stuff + e
+	\item What most do: $Y = \hat{B0} + X\hat{B1} + e$
+	\item Something like if we have dependencies then $E(\hat{B1}) = E(B1) + cov(B1,$ other stuff)}}
 	\end{itemize}
-	\begin{itemize}
-		\item \textcolor{blue}{ \emph{
-		Insert great response.
-		}}
 \end{itemize}

### 2020-01-30 revisionMemo/rev1.tex
MG and CD get shit done

@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		SM insert great response. 
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		We have edited P8 to read: Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes we observe.
+We have also clarified the goals of the paper in the earlier sections.
+		\item Here we have clarified that our study does not address causal identification but instead focuses on providing unbiased estimates of covariates, calibrated confidence intervals that reduce inferential error, and improved predictive performance.
+Though we understand these goals are not always central to every political scientists aim, we do believe they relate closely to the field of IR (as demonstrated by the studies we replicate) and a growing interest in forecasting and machine learning in Political Science.
+(CITE?)
+		\item For part (a), we have clarified that our study does not address causal identification.
+We instead focuses on providing unbiased estimates of covariates, which improves inference for observational data.
+		\item As for (b), the reviewer raises an important point. How well does AME work out when the DGP does not abide by parametric assumptions...relates to SM's to-do below.  
+		\item Finally, we understand these goals are not always central political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in \citet{people}.
 		}}

### 2020-01-30 revisionMemo/rev2.tex
MG and CD get shit done

@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here? 
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		Insert great response.
+		CDMG: Flag for discussion.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		TODO/CD: We agree with this suggestion.
+		TO DO (CD): We agree with this suggestion.
 We have now emphasized the importance of developing theory in our discussion of the AME and Oracle comparison (p.
 X), as well as in the conclusion of our paper (p.Y).
+We agree that innovations in dependence modeling can and should speak to theoretical advancements that, like our model, take into account the important relational and interdependent nature of socio-political processes.
+JOP CITE.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		TODO/CD: I think this comment and the above comment are essentially the same. We could even cite our JOP as an example of incorporating theory?  Perhaps other new-ish net paper as well?
+		TO DO (MG): We have added an additional paragraph on pg.
+18 in our discussion of Reiter and Stam which reveals...For example in this case, the finding that countries in the middle east experience more violence might lead one to more carefully examine the effects of geography or natural resources on the intiation of conflict.
 		}}
@@
 We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.
-In addition, we selected studies that could help illustrate different key components of the AME approach (Max, more here on each?).
+In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam), third-order dependencies (Weeks), precision of fixed effects estimates (Gibler).
 While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, they are beyond the scope of this paper at this time.
-(CD: Should we also mention something like 'in consultation with the editor, we believe they are beyond the scope....")
+(CD to SM: Should we also mention something like ``in consultation with the editor, we believe they are beyond the scope....")
 		}}

### 2020-02-04 revisionMemo/rev1.tex
sync up notes from meeting

@@
 The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
+		}}
+		\item \textcolor{blue}{ \emph{
+		Aronow et al. also explicitly note that their approach does not deal with issues of third order dependence.
+		}}
+		\item \textcolor{blue}{ \emph{
+		Aronow et al. approach does not deal with bias in parameter estimates.
 		}}
@@
 	\item What most do: $Y = \hat{B0} + X\hat{B1} + e$
-	\item Something like if we have dependencies then $E(\hat{B1}) = E(B1) + cov(B1,$ other stuff)}}
+	\item Something like if we have dependencies then $E(\hat{B1}) = E(B1) + cov(B1,$ other stuff)
+	\item SM: needs to add more here to indicate how the existing simulation section is meant to deal with this issue ...
+onus is on SM to bring in the problems that we highlight in the simulation section to section 2, so readers are clearer on the problem that we are trying to address
+	}}
 	\end{itemize}
@@
 		SM insert great response.
+		paper by Bai: https://sci-hub.tw/https://onlinelibrary.wiley.com/doi/10.3982/ECTA6135
+		connection between LFM and interactive fixed effects?
 		}}
@@
 Though we understand these goals are not always central to every political scientists aim, we do believe they relate closely to the field of IR (as demonstrated by the studies we replicate) and a growing interest in forecasting and machine learning in Political Science.
-(CITE?)
+(CITE?) [add reference to DGM JOP in which AME beat random forest]
 		\item For part (a), we have clarified that our study does not address causal identification.

### 2020-02-04 revisionMemo/rev2.tex
sync up notes from meeting

@@
 		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
+		Reference papers that have thought about what happens when we get the wrong W
+		general gist ..
+spatial models of dyadic dvs are great if we know how to characterize the diffusion amtrices, but the point of our approach is that we may not know how to actually characterize the ways in which actors are related to one another on some diffusion space.
+		citations ... flo ... and others based on something that we had researched in the past
+		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
 		}}
@@
 		CDMG: Flag for discussion.
+		in the appendix add more technical description of the estimation process
+		SM: figuring out a simulation case where ame fails to show i guess the limitations of it as an estimator ...
+		get stats from simulation
 		}}
@@
 JOP CITE.
+		maoz has stuff on linking theories with dependencies
+		brandon kinne's paper saom at apsr in which he models the formation of international treaties
+		siegel, larson
 		}}
@@
 While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, they are beyond the scope of this paper at this time.
-(CD to SM: Should we also mention something like ``in consultation with the editor, we believe they are beyond the scope....")
 		}}

### 2020-02-04 code/1_simulation/1_ameSim.R

@@ -7,37 +7,17 @@ if(Sys.info()['user']=='s7m'){
 	simResPath = paste0(dPath, 'simulation/')
 	source(paste0(fPath, 'functions.R')) }
 
-if(Sys.info()['user']=='herme' | 'Owner'){
-	user=Sys.info()['user']
-	base = paste0('C:/Users/',user,'/')
-	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
-	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
-	simResPath = paste0(dPath, 'simulation/')
-	source(paste0(fPath, 'functions.R')) }
-
 toLoad = c(
 	'devtools',
 	'foreach', 'doParallel',
-	'magrittr', 'dplyr', 'ggplot2',
-	'sna', # mrqap
-	'sandwich', 'Matrix', 'lmtest', 'arm', 'reshape2' # asa
-	)
+	'magrittr', 'dplyr', 'ggplot2' )
 loadPkg(toLoad)
-# devtools::install_github('s7minhas/amen', ref='dev')
-library(amen)
+devtools::install_github('s7minhas/amen', ref='dev')
 ##############################
 
 ##############################
 ### Homophily
-seed = 1
-n = 30
-intEff=-2 ; x1Eff=1 ; x2Eff=1
-mu = intEff
-beta = x1Eff
-gamma = x2Eff
-imps = 1000
-imp=1
-# simRun = function(seed, n, mu, beta, gamma){
+simRun = function(seed, n, mu, beta, gamma){
 	set.seed(seed)
 	xw = matrix(rnorm(n*2),n,2)
 	X = tcrossprod(xw[,1])  ; W = tcrossprod(xw[,2])
@@ -54,64 +34,13 @@ imp=1
 	fitO = ame(Y,XW,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
 		print=FALSE,plot=FALSE)
 
-	# MRQAP
-	loadPkg('sna')
-	fitQAP = netlm(y=Y, x=X,
-		intercept=TRUE, mode='digraph', diag=FALSE,
-		nullhyp='qap', reps=200)
-
-	# aronow et al
-	robust.se.nodfc <- function(model, cluster){
-		M <- length(unique(cluster))
-		N <- length(cluster)
-		K <- model$rank
-		dfc <- 1
[…]

### 2020-02-04 paper/2_model.tex
sync up notes from meeting

@@
 
-Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another, and treat each observation as independent of the others.
+Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another and not passing any auxiliary information to the model about how observations may be interdependent.
 For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.

### 2020-02-13 paper/4_app.tex
quick footnote

@@
 
-We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline.
-Of course, by the criteria of influence many other papers could have been considered for the replication.} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.
+We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{\textcolor{blue}{We selected our cases based on a few criteria.
+We focus on studies that are explicitly about International Relations, were published in the last 10 years, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we focus on one journal (American Political Science Review).
+We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.}} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.
 Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model.

### 2020-02-13 revisionMemo/rev2.tex
quick footnote

@@
 In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam), third-order dependencies (Weeks), precision of fixed effects estimates (Gibler).
-While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, they are beyond the scope of this paper at this time.
+While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, we believe that they are beyond the scope of this paper at this time.
 		}}

### 2020-02-13 revisionMemo/rev2.tex
2000s catch

@@
 		\item \textcolor{blue}{ \emph{
-		TO DO/CD: We selected our cases based on a few criteria.
+		Done: We selected our cases based on a few criteria.
 We focus on studies that are explicitly about International Relations, were published in the last 10 years, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we also decided to focus on one journal, the APSR).
 We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.
-In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam), third-order dependencies (Weeks), precision of fixed effects estimates (Gibler).
+In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam (2003), third-order dependencies (Weeks), precision of fixed effects estimates (Gibler).
 While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, we believe that they are beyond the scope of this paper at this time.

### 2020-02-13 paper/4_app.tex
2000s catch

@@
 We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{\textcolor{blue}{We selected our cases based on a few criteria.
-We focus on studies that are explicitly about International Relations, were published in the last 10 years, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we focus on one journal (American Political Science Review).
+We focus on studies that are explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we focus on one journal (American Political Science Review).
 We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.}} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.

### 2020-02-13 paper/5_concl.tex
update JOP

@@
 International relations is about the interactions, relationships, and dependencies among countries or other important international actors.
-This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} Research published as recently as 2017 and a broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
+This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} A broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
+\textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence plays an important role not just in terms of statistical estimations, but in the ways that scholars theorize about international level processes.
+Scholars have demonstrate this through research on conflict \cite{dorff:etal:2020}, economic interdependence \cite{maoz:2009a}, and
 

### 2020-02-13 paper/3_sim.tex
minor edits
update JOP

--- 2018-03-03 paper/3_sim.tex
@@
 Scholars can use this framework in an iterative fashion: beginning with an estimated model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their model.
-At the very least,  this simulation  demonstrates that the AME model can help conduct inference and estimate the true effect of independent variables that scholars operationalize to test theoretical propositions.
+\textcolor{blue}{Importantly, this simulation underscores how a careful consideration of a systems' interconnectedness, both through theoretical approaches and empirical models, can result in more precise estimates of direct and indirect effects across the system.}
+
+%older: This simulation also demonstrates that the AME model can help conduct inference and estimate the true effect of independent variables that scholars operationalize to test theoretical propositions.

### 2020-02-13 revisionMemo/rev2.tex
Cd to dos

@@
 		\item \textcolor{blue}{ \emph{
-		TO DO (CD): We agree with this suggestion.
+		Done: We agree with this suggestion.
 We have now emphasized the importance of developing theory in our discussion of the AME and Oracle comparison (p.
 X), as well as in the conclusion of our paper (p.Y).
-We agree that innovations in dependence modeling can and should speak to theoretical advancements that, like our model, take into account the important relational and interdependent nature of socio-political processes.
-JOP CITE.
-		maoz has stuff on linking theories with dependencies
-		brandon kinne's paper saom at apsr in which he models the formation of international treaties
-		siegel, larson
+We agree that innovations in dependence modeling can and should speak to theoretical advancements that, like our model, take into account the important relational and interdependent nature of political processes.
 		}}

### 2020-02-13 paper/5_concl.tex
Cd to dos

@@
 This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} A broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
-\textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence plays an important role not just in terms of statistical estimations, but in the ways that scholars theorize about international level processes.
-Scholars have demonstrate this through research on conflict \cite{dorff:etal:2020}, economic interdependence \cite{maoz:2009a}, and
+\textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence plays an important role not just in terms of statistical estimations, but in the ways that scholars theorize about internationally relevant politics.
+Scholars have demonstrated the theoretical importance of interdependence through research on conflict \citep{dorff:etal:2020}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.}
 

### 2020-02-14 revisionMemo/rev2.tex
talk about spatial stuff

@@
 		\item \textcolor{blue}{ \emph{
-		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
-		Reference papers that have thought about what happens when we get the wrong W
-		general gist ..
+%		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
+%		Reference papers that have thought about what happens when we get the wrong W
+%		general gist ..
 spatial models of dyadic dvs are great if we know how to characterize the diffusion amtrices, but the point of our approach is that we may not know how to actually characterize the ways in which actors are related to one another on some diffusion space.
-		citations ... flo ... and others based on something that we had researched in the past
-		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
+%		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
+In the appendix, we compare our models approach in simulations to a number of other methods of dealing with dependencies, notably LIST THEM.
+In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern.
+Further, spatial lags are likely to be particularly complex in the cases of dyadic data because we would need an appropriate measure of dyadic similarity (see \citet{plumper:neumayer:2010}).
+
+NOTE TO US: Reading this Q again, do they want us to redo the simulation such that the variance is coming from spatial stuff?
 		}}

### 2020-02-14 paper/5_concl.tex
edits with MG

@@
 \textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence plays an important role not just in terms of statistical estimations, but in the ways that scholars theorize about internationally relevant politics.
-Scholars have demonstrated the theoretical importance of interdependence through research on conflict \citep{dorff:etal:2020}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.}
+Scholars have demonstrated the theoretical importance of interdependence through research on intrastate conflict \citep{dorff:etal:2020}, interstate bargaining \citep{gallop:2017}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.}
 

### 2020-02-14 revisionMemo/rev2.tex
edits with MG

@@
 \item \textcolor{blue}{ \emph{
+The reviewer raises an important connection to the spatial literature.
+The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria.
+With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l).
+In an effort to speak to other known approaches in the literature, we compare our model via simulations to a number of other methods of dealing with dependencies, notably LIST THEM.
+In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).
+}}
+
 %		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
@@
 %		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
-In the appendix, we compare our models approach in simulations to a number of other methods of dealing with dependencies, notably LIST THEM.
-In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern.
-Further, spatial lags are likely to be particularly complex in the cases of dyadic data because we would need an appropriate measure of dyadic similarity (see \citet{plumper:neumayer:2010}).
 
-NOTE TO US: Reading this Q again, do they want us to redo the simulation such that the variance is coming from spatial stuff?
-		}}
 	\end{itemize}

### 2020-02-20 revisionMemo/rev1.tex
cites

@@
 		\item Here we have clarified that our study does not address causal identification but instead focuses on providing unbiased estimates of covariates, calibrated confidence intervals that reduce inferential error, and improved predictive performance.
-Though we understand these goals are not always central to every political scientists aim, we do believe they relate closely to the field of IR (as demonstrated by the studies we replicate) and a growing interest in forecasting and machine learning in Political Science.
-(CITE?) [add reference to DGM JOP in which AME beat random forest]
 		\item For part (a), we have clarified that our study does not address causal identification.
@@
 		\item As for (b), the reviewer raises an important point. How well does AME work out when the DGP does not abide by parametric assumptions...relates to SM's to-do below.
-		\item Finally, we understand these goals are not always central political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in \citet{people}.
+		\item Finally, we understand these goals are not always central political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Muchlinski et al.
+2016; Colaresi and Mahmood 2017; Mueller and Rauh 2018.
 		}}

### 2020-02-27 revisionMemo/rev1.tex
R1 done CD

@@
 	\begin{itemize}
-		\item \textcolor{blue}{ \emph{
-		TO DO in paper: Strategies like those employed by Aronow et al do not directly model all relevant dependencies in the data \footnote{As Aronow et al explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data." p.
-16.} but instead adjust the variance estimator.
-While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).
+		\item  \emph{
+		We have clarified our language in the paper (on page 1) to address this question.
+Strategies like those employed by Aronow et al.
+do not directly model all relevant dependencies in the data but instead adjust the variance estimator.\footnote{As Aronow et al.
+explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data." p.
+16.}   While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).
 The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
-		}}
-		\item \textcolor{blue}{ \emph{
-		Aronow et al. also explicitly note that their approach does not deal with issues of third order dependence.
-		}}
-		\item \textcolor{blue}{ \emph{
-		Aronow et al. approach does not deal with bias in parameter estimates.
-		}}
+		}
+		%\item \textcolor{blue}{ \emph{
+		%Aronow et al. also explicitly note that their approach does not deal with issues of third order dependence.
+		%}}
+		%\item \textcolor{blue}{ \emph{
+		%Aronow et al. approach does not deal with bias in parameter estimates.
+		%}}
 	\end{itemize}
@@
 	\begin{itemize}
-		\item \textcolor{blue}{ \emph{
+		\item {\emph{
 		 The point is well taken, we have clarified this section (on page 2) to point to the assumption of independence as the primary problem, not the stacking of observations in and of itself.
@@
 	\begin{itemize}
-		\item \textcolor{blue}{ \emph{
-		We have edited P8 to read: Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes we observe.
-We have also clarified the goals of the paper in the earlier sections.
-		\item Here we have clarified that our study does not address causal identification but instead focuses on providing unbiased estimates of covariates, calibrated confidence intervals that reduce inferential error, and improved predictive performance.
-		\item For part (a), we have clarified that our study does not address causal identification.
-We instead focuses on providing unbiased estimates of covariates, which improves inference for observational data.
-		\item As for (b), the reviewer raises an important point. How well does AME work out when the DGP does not abide by parametric assumptions...relates to SM's to-do below.
-		\item Finally, we understand these goals are not always central political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Muchlinski et al.
-2016; Colaresi and Mahmood 2017; Mueller and Rauh 2018.
-		}}
+		\item  \emph{
+For part (a), we have clarified that our study does not address causal identification.
+We instead focus on providing unbiased estimates of covariates, which improves inference for observational data.
+We have also edited P8 to read: Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.
+We have also clarified the goals of the paper in the earlier sections.}
+		\item \emph{\textcolor{blue}{As for (b), the reviewer raises an important point. How well does AME work out when the DGP does not abide by parametric assumptions...relates to SM's to-do below.}}
+		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; Mueller and Rauh 2018.
+CD: Do I add this to text somewhere or leave as memo only?}
+		}
 	\end{itemize}

### 2020-02-27 paper/2_model.tex
r1 edits

@@
 
-Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes we observe, and that accounting for this structure is necessary if we are to adequately represent the data generating process.
-If one can specify each of the nodal, dyadic, and triadic set of attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.
+\textcolor{blue}{Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.} Accounting for this structure is necessary if we are to adequately represent the data generating process.
+If one can specify each of the nodal, dyadic, and triadic attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.
 However, it is rarely the case that this is possible even for TSCS data and thus modeling decisions must account for underlying structure.

### 2020-02-27 paper/5_concl.tex
r1 edits

@@
 This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} A broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
-\textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence plays an important role not just in terms of statistical estimations, but in the ways that scholars theorize about internationally relevant politics.
+\textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence influences not only statistical estimations, but how scholars theorize about internationally relevant politics.
 Scholars have demonstrated the theoretical importance of interdependence through research on intrastate conflict \citep{dorff:etal:2020}, interstate bargaining \citep{gallop:2017}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.}

### 2020-02-27 paper/1_intro.tex
r1 edits

@@
 Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
-In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
-This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.
+\textcolor{blue}{In practice, such strategies remain insufficient because they do not directly model all relevant dependencies in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data."} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null)} The main goal of our study is to directly account for the relevant dependency structures in the data, and thus provide an accurate estimate while addressing both types of possible errors.} This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.
 
-In this article, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
+In this article, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2019}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2020-02-27 paper/3_sim.tex
started R2

@@
 Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
-The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.
-Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}
+The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.\footnote{\textcolor{blue}{The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria.
+With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l).
+In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).}} Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}
 

### 2020-02-27 revisionMemo/rev2.tex
started R2 in memo

@@
 With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l).
-In an effort to speak to other known approaches in the literature, we compare our model via simulations to a number of other methods of dealing with dependencies, notably LIST THEM.
-In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).
+In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see Plumper and Neumayer 2010 for a more in-depth discuss on spatial lags).
+CD: I added this as a footnote on p.10.
+I moved this footnote around a LOT but I think it needs to live somewhere in the simulation.
+It felt awkward within the text.
 }}

### 2020-02-27 revisionMemo/rev2.tex
memo edits again

@@
 In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see Plumper and Neumayer 2010 for a more in-depth discuss on spatial lags).
-CD: I added this as a footnote on p.10.
+\textbf{CD}: I added this as a footnote on p.10.
 I moved this footnote around a LOT but I think it needs to live somewhere in the simulation.
@@
 		\item \textcolor{blue}{ \emph{
-		Done: We agree with this suggestion.
+		We agree with this suggestion.
 We have now emphasized the importance of developing theory in our discussion of the AME and Oracle comparison (p.
@@
 		\item \textcolor{blue}{ \emph{
-		TO DO (MG): We have added an additional paragraph on pg.
-18 in our discussion of Reiter and Stam which reveals...For example in this case, the finding that countries in the middle east experience more violence might lead one to more carefully examine the effects of geography or natural resources on the intiation of conflict.
+		We have added an additional interpretation of the random effects on pg.
+19 in our discussion of Reiter and Stam.
+(``Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
+Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
+In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography or natural resources on the initiation of conflict.") \textbf{CD}: Likely need to re-write to link to THEORY specifically?
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		Done: We selected our cases based on a few criteria.
-We focus on studies that are explicitly about International Relations, were published in the last 10 years, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we also decided to focus on one journal, the APSR).
+		We selected our cases based on a few criteria.
+We focus on studies that are explicitly about International Relations, were published since the year 2000, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we also decided to focus on one journal, the APSR).
 We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.
-In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam (2003), third-order dependencies (Weeks), precision of fixed effects estimates (Gibler).
+In addition, we selected studies that could help illustrate different key components of the AME approach such as random effects (Reiter and Stam 2003), third-order dependencies (Weeks 2012), precision of fixed effects estimates (Gibler 2017).
 While we agree with the reviewer that further replications would effectively demonstrate the benefits of the AME, we believe that they are beyond the scope of this paper at this time.

### 2020-02-28 code/reiter_stam_2003/amen-tutorial.Rmd
added tutorial

---
title: "AMEN tutorial"
author: "Juan Tellez"
date: "2/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE,
                      tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Introduction

This tutorial walks you through using the `AMEN` package with data on interstate conflict (MIDs) from Reiter and Stam (2003).

```{r load-libraries}
library(tidyverse)
library(haven)
```

Let's load the data below and look at it.

```{r load-data, cache=TRUE}
## read data and add variables
data = read_dta('/Users/JuamnTellez/Dropbox/papers/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
head(data)
```

For this tutorial, we only want the subset of the variables that were in the original analysis, and then we want to drop missing observations.

```{r drop-miss}
## get variables needed
base_vars = c('sideaa', "pdemdtar", "pdemdin", "personal", "military", 
              "single", "democ", "contig", "majpow", "ally", "loglsrat",
              "advanced", "dispyrs", "dspline1", "dspline2", "dspline3",
              "statea", "stateb", "year")

## subset the data by variables needed and years
dataComp = na.omit(data[, base_vars])
```

# Formatting the Data

Using the AMEN function requires formatting data into a particular structure. The primary distinction in data formatting is whether the outcome of interest represents a directed or undirected network.

If undirected, the AMEN function has three main inputs:

\begin{itemize}
    \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network. 
    \begin{itemize}
      \item An adjacency matrix describes relationships between nodes in a particular year of data. For example, in an adjacency matrix of interstate MIDs, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had a MID in that year, and '0' otherwise. The diagonal in the adjacency matrix is typically missing. Y is a list of these adjacency matrices for the outcome variable, where each element in the list is a different year of data. 
    \end{itemize}
    \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset. 
    \begin{itemize}
      \item An array is a data object in R that contains a series of 'stacked' matrices for each year of data. An array of dimension (2, 3, 4), for example, contains 4 rectangular matrices each with 2 rows and 3 columns. Each matrix in the array describes the relationship between nodes with respect to some covariate. For example, for interstate alliances, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had an alliance in that year, and '0' otherwise. An array contains a matrix of this kind for each covariate going into the model. Xdyad is a list of these arrays, where each element in the list is a different year of data. 
    \end{itemize}
    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
    \begin{itemize}
      \item Each matrix in Xrow has the nodes running down the rows and the covariates in the dataset running along the columns. An entry in the matrix captures the value that a node takes on for each covariate in a given year. For example, if column $j$ is GDP per capita and column $k$ is population, then row $i$, column $j$ measures country $i$'s GDP per capita and row $i$, column $k$ measures country $i$'s population in a given year. Xrow is a list of these matrices, where each element in the list is a different year of data. 
    \end{itemize}
\end{itemize}

If directed, AMEN further requires:

\begin{itemize}
    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
    \begin{itemize}
      \item Xrow here is nearly identical to the undirected case. The only difference is that rather than each matrix containing covariate data on all nodes, in the directed context each matrix only contains data on the nodes that are acting as senders in that particular year. Using interstate MIDs as an example, if in 1983 only two countries initiated MIDs, Xrow would only contain data for those two countries (i.e., Xrow would only have two rows). 
    \end{itemize}
    \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset. 
    \begin{itemize}
      \item Xcol here is nearly identical to Xrow, except each matrix contains data on receiver nodes, not sender nodes. Using interstate MIDs as an example, if in 1983 only two countries had MIDs initiated against them, Xcol would only contain data for those two countries (i.e., Xcol would only have two rows). 
    \end{itemize}
\end{itemize}

The Reiter and Stam (2003) example is undirected.

## Helper Functions

We provide some helper functions to make formatting easier.

First we need a *vector* of the years in the dataset, and then a *list* that tells us, *for each year*, which countries are present in the data (here represented by COW codes).

```{r get-state-years}
yrs = sort(unique(dataComp$year)) 
cntriesT = lapply(yrs, function(t){ as.character( unique( dataComp$statea[dataComp$year==t] ) ) })
```

### Format Y (The Outcome)

Now we will format `Y`, the outcome variable, which is a list of adjacency matrices as described above. Roughly, the code below does the following: for each year of data, subset to the $N$ countries present in that year, and then recast the data frame into an  $N \times N$ adjacency matrix, before finally storing it as an element into a list. Notice that Y is a list with 47 elements (one for each year in the data).

```{r format-y,cache=TRUE}
## format dependent variable
yVar = 'sideaa' # name of outcome in dataset
yList = lapply(1:length(yrs), function(ii){
  slice = data[ which( 
    data$year==yrs[ii] & 
      data$statea %in% cntriesT[[ii]] & 
      data$stateb %in% cntriesT[[ii]]
  ), c('statea', 'stateb', yVar) ]
  adj = reshape2::acast(slice, statea ~ stateb, value.var=yVar)
  return( adj[ cntriesT[[ii]], cntriesT[[ii]] ] )
})
```

Here's an example of one year of data (only 5 countries shown for space purposes):

```{r y-example}
yList[[1]][1:5, 1:5]
```

### Format Xdyad, the Dyadic Covariates

Now we will format `Xdyad` from above, which includes the dyadic covariates. `Xdyad` is a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset.

Roughly, the code below does the following: for each year in the dataset, subset down to the countries present in that year and all of the dyadic variables in `dyad_vars`; then cast the data into an array that is  $n\times n\times p$, before storing it into the list.

```{r format-xDyad, cache=TRUE}
# get dyadic variables
dyad_vars = c('personal', 'military', 'single', 'democ', 'contig', 
              'ally', 'majpow', 'loglsrat', 'advanced', 
              'dispyrs', 'dspline1', 'dspline2', 'dspline3', 
              'pdemdtar', 'pdemdin')

## format dyadic covariates
xDyadList = lapply(1:length(yrs), function(ii){
  slice = data[ which( 
    data$year==yrs[ii] & 
      data$statea %in% cntriesT[[ii]] & 
      data$stateb %in% cntriesT[[ii]]
  ), c('statea', 'stateb', dyad_vars) ]
  sliceL = reshape2::melt(slice, id=c('statea','stateb'))
  adj = reshape2::acast(sliceL, statea ~ stateb ~ variable, value.var='value')
  return( adj[ cntriesT[[ii]], cntriesT[[ii]],  ] )
})

```

Here's an example of what the first year of data looks like for the first covariate, named `personal` (only 5 countries shown for space purposes):

```{r xdyad-example}
xDyadList[[1]][1:5,1:5,1]
```

We are done formatting the data and are ready for fitting the model.

# Running AMEN

Beyond the data inputs, the AMEN function requires additional specification:

\begin{itemize}
    \item model: how to model the outcome variable, e.g., `logit'
    \item symmetric: whether the input network is symmetric
    \item intercept: whether to estimate an intercept
    \item nscan: number of iterations of the Markov chain
    \item burn: burn-in period
    \item odens: thinning interval
    \item R: dimension of the multiplicative effect (referred to as $K$ in the paper)
    \item gof: whether to calculate goodness of fit statistics
\end{itemize}

There is often little theoretical reason to choose a particular value of $\sf{R}$ (above). One strategy is to estimate models at different values of $\sf{R}$ and compare goodness of fit statistics across models.

Given the computational intensity needed for parameter estimates to converge, parallelization strategies are recommended to speed up analysis. In addition, providing AMEN function with starting values, either dictated by theory, previous research, or previous runs can also help speed up convergence time.

Below, we load the library, and in `resultsPath` we set a path to store results.

```{r load-amen, eval=FALSE}
## load most updated amen package
devtools::install_github('s7minhas/amen') ; library(amen)

## Run amen in parallel
library(doParallel) ; library(foreach)

## setup result output directory
resultsPath = '/Users/JuamnTellez/Dropbox/papers/netsMatter/replications/Reiter_Stam_2003/output/'
```

Now we set the parameters described above. Note that `latDims` here is the same as $\sf{R}$ above, and we have set it to 4.

```{r amen-params, eval=FALSE}
## running in parallel varying k
imps = 10000  # number of interations in MCMC
brn =  2500   # number of burn-in
ods = 1 # thinning parameter 
## (1 is suggested unless there is some evidence of autocorrelation in the chain)
latDims = 0:3 # dimension of latent space (search up to 3 dimension  is suggested)
seed=6886 # set your model seed
```

Now we run the model, in parallel. We fit a separate model for each element in `latDims` (so a model with a zero dimensional latent space, 1 dimensional latent space, etc.).

The key function here is `ame_repL` which fits the model. We wrap the whole function in a `foreach` call so we can run separate models for each element in `latDims`.

```{r run-amen, eval=FALSE}
## select number of cores to locate
cl=makeCluster(4) ; registerDoParallel(cl) # number of clusters depends 
## on how many cores your computer or cluster has

foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {

  ameFit = ame_repL(
    Y=yList,Xdyad=xDyadList, 
    Xrow=NULL,Xcol=NULL, 
    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii], 
    nscan=imps, seed=seed, burn=brn, odens=ods, 
    plot=FALSE, print=FALSE, gof=TRUE,
    periodicSave=TRUE)

  save(ameFit, file=paste0(resultsPath, 'model_k', latDims[ii],'_v1.rda') ) # save the output of your data
}

## terminate parallel sockets
stopCluster(cl)
```

In cases when the model fails to converge you may want to run the model again or with a longer chain. In this case, it is often useful to set starting values which can help with convergence issues.

Below, we demonstrate how to take starting values from a prior run and then fit the model again.

```{r rerun-amen, eval=FALSE}
## load resuls from previous model run if needed
prevModelFiles = paste0(resultsPath, 'model_k', latDims,'_v1.rda')

## select number of cores to locate
cl=makeCluster(4) ; registerDoParallel(cl)

foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {

  load(prevModelFiles[ii]) # load previous model run

  startVals0 = ameFit$'startVals' # extract start vals

  rm(ameFit)  # dump the rest of the model run

  ameFit = ame_repL(
    Y=yList,Xdyad=xDyadList, 
    Xrow=NULL,Xcol=NULL, 
    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii], 
    nscan=imps, seed=seed, burn=brn, odens=ods, 
    plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0,
    periodicSave=TRUE)

  # save the output of your data
  save(ameFit, file=paste0(resultsPath, 'model_k', latDims[ii],'_v2.rda') ) 
}

## terminate parallel sockets
stopCluster(cl)
```

# Posterior Analysis

Once we are satisfied that the model has converged, we can do posterior analysis to examine our parameter estimates. Below we load the results from the `AMEN` run. Simplicity, we only load results from $R = 0$.

```{r load-posterior, eval=FALSE}
## load model fits from 0 to 3 dimensions
load( paste0(resultsPath,'ameFit_k0.rda') ) ; ameFit_k0=ameFit
# load( paste0(resultsPath,'ameFit_k1.rda') ) ; ameFit_k1=ameFit
# load( paste0(resultsPath,'ameFit_k2.rda') ) ; ameFit_k2=ameFit
# load( paste0(resultsPath,'ameFit_k3.rda') ) ; ameFit_k3=ameFit
```

We can look at goodness of fit (GOF) statistics with the following:

```{r amen-gof, eval=FALSE}
## check out goodness of fit stats in posterior estimate for each dimension
amen::gofPlot(ameFit_k0$GOF, symmetric = TRUE)
```

We can also summarize the distribution of our beta parameters, below, using the `summStats` function which returns the mean, median, standard deviation, 95\%, and 99\% credible intervals:

```{r beta-params, eval=FALSE}
## check out beta parameters (posterior estimates for coefficients) 
## and convergence for each dimension
summStats = function(x){
  # credible intervals
  res=c(mu=mean(x),med=median(x),sd=sd(x),quantile(x,probs=c(0.025,0.05,0.95,0.975))) 
  round(res, 3)
}
apply(ameFit_k0$BETA, 2, summStats) %>% t() %>% nrow()
```

We can also verify that parameters have converged for sender and receiver effects with the code below:

```{r beta-converge, eval=FALSE}
betaIndices<-split(1:ncol(ameFit_k0$BETA), ceiling(seq_along(1:ncol(ameFit_k0$BETA))/5))
```

And finally make plots to assess the converge of parameters:

```{r plot-converge, eval=FALSE}
## plot the parameters convergence using dimension k=0
for(bIndex in betaIndices){ amen::paramPlot( ameFit_k0$BETA[,bIndex,drop=FALSE] ) }
amen::paramPlot(ameFit_k0$VC)
```

### 2020-02-28 code/reiter_stam_2003/amen-tutorial.R
added tutorial

@@ -9,7 +9,7 @@
 rm(list=ls())
 
 ## setup working directory
-setwd('yourPath')
+setwd('code/reiter_stam_2003')
 
 ## load libraries
 packs = c('dplyr', 'ggplot2', 'foreign', 'readr', 'lmtest','magrittr')

### 2020-03-02 revisionMemo/rev1.tex
memo edits again
slight memo changes

@@
 	\begin{itemize}
-		\item  \emph{
+		\item  \emph{ \textcolor{blue}{
 		We have clarified our language in the paper (on page 1) to address this question.
@@
 The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
-		}
+		}}
 		%\item \textcolor{blue}{ \emph{
@@
 	\begin{itemize}
-		\item {\emph{
+		\item {\emph{\textcolor{blue}{
 		 The point is well taken, we have clarified this section (on page 2) to point to the assumption of independence as the primary problem, not the stacking of observations in and of itself.
-		}}
+		}}}
 	\end{itemize}
@@
 	\begin{itemize}
-		\item  \emph{
+		\item  \emph{\textcolor{blue}{
 For part (a), we have clarified that our study does not address causal identification.
@@
 We have also edited P8 to read: Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.
-We have also clarified the goals of the paper in the earlier sections.}
-		\item \emph{\textcolor{blue}{As for (b), the reviewer raises an important point. How well does AME work out when the DGP does not abide by parametric assumptions...relates to SM's to-do below.}}
-		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; Mueller and Rauh 2018.
-CD: Do I add this to text somewhere or leave as memo only?}
+We have also clarified the goals of the paper in the earlier sections.}}
+		\item \emph{\textcolor{blue}{As for (b), the reviewer raises an important point.
+How well does AME work out when the DGP does not abide by parametric assumptions we address this in the section concerning the simulation and Aronow's approach.}}
+		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; Mueller and Rauh 2018.}
 		}

### 2020-03-02 revisionMemo/rev2.tex
theory for random effects

@@
 Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
-In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography or natural resources on the initiation of conflict.") \textbf{CD}: Likely need to re-write to link to THEORY specifically?
+In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for \citet{colgan:2010}'s theory that revolutionary petrostates are more aggressive.")
 				}}

### 2020-03-02 code/reiter_stam_2003/amen-tutorial.Rmd
small change to tutorial
small changes

@@
 ## read data and add variables
-data = read_dta('/Users/JuamnTellez/Dropbox/papers/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
+data = read_dta('/Users/dorffc/Dropbox/Research/nothingbutnet/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
 head(data)
@@
 For example, in an adjacency matrix of interstate MIDs, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had a MID in that year, and '0' otherwise.
-The diagonal in the adjacency matrix is typically missing.
+The diagonal in the adjacency matrix is typically set to `NA`.
 Y is a list of these adjacency matrices for the outcome variable, where each element in the list is a different year of data.
@@
     \end{itemize}
-    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
+    \item `Xrow`: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset. 
     \begin{itemize}
-      \item Each matrix in Xrow has the nodes running down the rows and the covariates in the dataset running along the columns.
+      \item Each matrix in `Xrow` has the nodes running down the rows and the covariates in the dataset running along the columns.
 An entry in the matrix captures the value that a node takes on for each covariate in a given year.
 For example, if column $j$ is GDP per capita and column $k$ is population, then row $i$, column $j$ measures country $i$'s GDP per capita and row $i$, column $k$ measures country $i$'s population in a given year.
-Xrow is a list of these matrices, where each element in the list is a different year of data.
+`Xrow` is a list of these matrices, where each element in the list is a different year of data.
     \end{itemize}
@@
 \begin{itemize}
-    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
+    \item `Xrow`: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset. 
     \begin{itemize}
-      \item Xrow here is nearly identical to the undirected case.
+      \item `Xrow` here is nearly identical to the undirected case.
 The only difference is that rather than each matrix containing covariate data on all nodes, in the directed context each matrix only contains data on the nodes that are acting as senders in that particular year.
-Using interstate MIDs as an example, if in 1983 only two countries initiated MIDs, Xrow would only contain data for those two countries (i.e., Xrow would only have two rows).
+Using interstate MIDs as an example, if in 1983 only two countries initiated MIDs, `Xrow` would only contain data for those two countries (i.e., `Xrow` would only have two rows).
     \end{itemize}
@@
     \begin{itemize}
-      \item Xcol here is nearly identical to Xrow, except each matrix contains data on receiver nodes, not sender nodes.
+      \item Xcol here is nearly identical to `Xrow`, except each matrix contains data on receiver nodes, not sender nodes.
 Using interstate MIDs as an example, if in 1983 only two countries had MIDs initiated against them, Xcol would only contain data for those two countries (i.e., Xcol would only have two rows).
@@
 
-  save(ameFit, file=paste0(resultsPath, 'model_k', latDims[ii],'_v1.rda') ) # save the output of your data
+  save(ameFit, file=paste0(resultsPath, 'model_k', latDims[ii],'_v1.rda') ) 
+  # save the output of your data
 }

### 2020-03-06 revisionMemo/rev1.tex
memo edits

@@
 	\item P1: ``Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
-In practice, such strategies rarely work because they do not directly address the fundamental data generating process''.
+In practice, such strategies rarely work because they do not directly address the fundamental data generating process.''
 	\begin{itemize}
@@
 explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data." p.
-16.}   While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).
+16.}   While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will lead us to falsely accept the null).
 The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
@@
 For c), I think most would agree that the “proof” is simply in comparison to whatever other good models are available.
-Not however, that in social science applications, forecasting and therefor c) are rarely of interest (at least not in the types of research that appear in political science journals).
+Not however, that in social science applications, forecasting and therefore c) are rarely of interest (at least not in the types of research that appear in political science journals).
 As such, from the perspective of what the researchers in discipline mostly cares about, a) and b) are of primary importance.
@@
 For part (a), we have clarified that our study does not address causal identification.
-We instead focus on providing unbiased estimates of covariates, which improves inference for observational data.
-We have also edited P8 to read: Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.
-We have also clarified the goals of the paper in the earlier sections.}}
+We instead focus on providing unbiased estimates of covariates, which improve inference for observational data.
+We have clarified the goals of the paper in the earlier sections (such as the opening paragraph of page 1) and have also edited P8 to read: ``Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe." }}
 		\item \emph{\textcolor{blue}{As for (b), the reviewer raises an important point.
-How well does AME work out when the DGP does not abide by parametric assumptions we address this in the section concerning the simulation and Aronow's approach.}}
-		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; Mueller and Rauh 2018.}
+How well does AME work out when the DGP does not abide by parametric assumptions?
+We address this in the section concerning the simulation and Aronow's approach.
+\textbf{SM we need to add more here}}}
+		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others.}
 		}

### 2020-03-06 code/reiter_stam_2003/amen-tutorial.Rmd
CD RMD checks

@@
 data = read_dta('/Users/dorffc/Dropbox/Research/nothingbutnet/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
-head(data)
+#data = read_dta('/Users/JuamnTellez/Dropbox/papers/netsMatter/replications/Reiter_Stam_2003/input/doublereduce.dta')
 ```
@@
 ## setup result output directory
-resultsPath = '/Users/JuamnTellez/Dropbox/papers/netsMatter/replications/Reiter_Stam_2003/output/'
+#resultsPath = '/Users/JuamnTellez/Dropbox/papers/netsMatter/replications/Reiter_Stam_2003/output/'
+resultsPath = '/Users/dorffc/Dropbox/Research/nothingbutnet/netsMatter/replications/Reiter_Stam_2003/output/'
 ```
@@
 ## select number of cores to locate
-cl=makeCluster(4) ; registerDoParallel(cl) # number of clusters depends 
+cl=makeCluster(6) ; registerDoParallel(cl) # number of clusters depends 
 ## on how many cores your computer or cluster has

### 2020-03-06 code/reiter_stam_2003/amen-tutorial.Rmd
CD RMD code review

@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 
-```{r load-amen, eval=FALSE}
+```{r load-amen, eval=TRUE}
 ## load most updated amen package
@@
 
-```{r load-posterior, eval=FALSE}
+```{r load-posterior, eval=TRUE}
 ## load model fits from 0 to 3 dimensions
-load( paste0(resultsPath,'ameFit_k0.rda') ) ; ameFit_k0=ameFit
-# load( paste0(resultsPath,'ameFit_k1.rda') ) ; ameFit_k1=ameFit
-# load( paste0(resultsPath,'ameFit_k2.rda') ) ; ameFit_k2=ameFit
-# load( paste0(resultsPath,'ameFit_k3.rda') ) ; ameFit_k3=ameFit
+load( paste0(resultsPath,'model_k0_v12.rda') ) ; ameFit_k0=ameFit
+# load( paste0(resultsPath,'model_k1_v12.rda') ) ; ameFit_k1=ameFit
+# load( paste0(resultsPath,'model_k2_v12.rda') ) ; ameFit_k2=ameFit
+# load( paste0(resultsPath,'model_k3_v12.rda') ) ; ameFit_k3=ameFit
 ```
@@
 
-```{r amen-gof, eval=FALSE}
+```{r amen-gof, eval=TRUE}
 ## check out goodness of fit stats in posterior estimate for each dimension
@@
 
-```{r beta-params, eval=FALSE}
+```{r beta-params, eval=TRUE}
 ## check out beta parameters (posterior estimates for coefficients) 
@@
 }
-apply(ameFit_k0$BETA, 2, summStats) %>% t() %>% nrow()
+apply(ameFit_k0$BETA, 2, summStats) %>% t() 
 ```
@@
 
-```{r beta-converge, eval=FALSE}
+```{r beta-converge, eval=TRUE}
 betaIndices<-split(1:ncol(ameFit_k0$BETA), ceiling(seq_along(1:ncol(ameFit_k0$BETA))/5))
@@
 
-```{r plot-converge, eval=FALSE}
+```{r plot-converge, eval=TRUE}
 ## plot the parameters convergence using dimension k=0

### 2020-05-18 revisionMemo/rev1.tex
cleanup

@@
 For me at least, after reading section 2 I still wasn’t sure what problem the paper is actually trying to address.
-		\item \textcolor{blue}{ \emph{ MG/CD: We think the reviewer makes a good suggestion here to clarify, mathematically or in some other illustrative way, exactly how we think that dependencies bias the estimator.
+		\item \textcolor{blue}{ \emph{ We think the reviewer makes a good suggestion here to clarify, mathematically or in some other illustrative way, exactly how we think that dependencies bias the estimator.
 For example we can say that our target of inference is the estimation of beta (with some variance).
-If a coefficient is correlated with a dependency, then we could obtain an estimate for beta that is artificially high (and the same is true for the reverse).}}
-	\begin{itemize}
-	\item  \textcolor{blue}{ \emph{ What we think: True DGP: $Y = B0 + XB1 + a + b +$ 3rd order stuff + e
-	\item What most do: $Y = \hat{B0} + X\hat{B1} + e$
-	\item Something like if we have dependencies then $E(\hat{B1}) = E(B1) + cov(B1,$ other stuff)
-	\item SM: needs to add more here to indicate how the existing simulation section is meant to deal with this issue ...
+If a coefficient is correlated with a dependency, then we could obtain an estimate for beta that is artificially high (and the same is true for the reverse).
+		}}
+	\item \textcolor{blue}{ \emph{ Additionally, we have SM: needs to add more here to indicate how the existing simulation section is meant to deal with this issue ...
 onus is on SM to bring in the problems that we highlight in the simulation section to section 2, so readers are clearer on the problem that we are trying to address
@@
 		\item It is useful to think of the infeasible estimator and the implications for interpreting the LFM component.
-The infeasible estimator takes the transformation f(.), fixed effects parameters, and SRRM random effects as know, in which case the LFM amounts to predictions from a factor model fit to the adjacency matrix of residuals on the scale of the latent outcome.
+The infeasible estimator takes the transformation f(.), fixed effects parameters, and SRRM random effects as known, in which case the LFM amounts to predictions from a factor model fit to the adjacency matrix of residuals on the scale of the latent outcome.
 Would this be a reasonable interpretation?
@@
 We have clarified the goals of the paper in the earlier sections (such as the opening paragraph of page 1) and have also edited P8 to read: ``Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe." }}
-		\item \emph{\textcolor{blue}{As for (b), the reviewer raises an important point.
+		\item \emph{\textcolor{blue}{
+			As for (b), the reviewer raises an important point.
 How well does AME work out when the DGP does not abide by parametric assumptions?
 We address this in the section concerning the simulation and Aronow's approach.
-\textbf{SM we need to add more here}}}
+}}
 		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others.}

### 2020-06-04 code/1_simulation/4_corUVandMiss.R
sync up sim code

@@ -7,12 +7,13 @@ if(Sys.info()['user']=='s7m'){
 	graphicsPath = paste('~/Research/netsMatter/paper/')
 	source(paste0(fPath, 'functions.R')) }
 
-if(Sys.info()['user']=='herme' | 'Owner'){
-	user=Sys.info()['user']
-	base = paste0('C:/Users/',user,'/')
-	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
+	base=paste0('C:/Users/',Sys.info()['user'],'/')
+	gPath = paste0(base, 'Research/netsMatter/')
+	fPath = paste0(gPath, 'code/helpers/')
 	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
 	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste0(gPath, 'paper/')
 	source(paste0(fPath, 'functions.R')) }
 
 toLoad = c(
@@ -30,7 +31,8 @@ facet_labeller = function(string){ TeX(string) }
 NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 # load sim results
-for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
+# for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'_asa.rda')) }
 ##############################
 
 ##############################
@@ -99,5 +101,6 @@ g = ggplot() +
 
 #
 ggsave(g, height=3, width=8,
-	file=paste0(graphicsPath, 'ameSimCorr.pdf'))
+	# file=paste0(graphicsPath, 'ameSimCorr.pdf'))	
+	file=paste0(graphicsPath, 'ameSimCorr_asa.pdf'))
 ##############################

### 2020-06-04 paper/2_model.tex
sync up sim code

@@
 It is flexible and easy to use for analyzing the kind of relational data often found in the social sciences.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2018} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2019} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
 For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:2005}.

### 2020-06-04 code/1_simulation/1_ameSim_asa.R
sync up sim code

@@ -7,31 +7,40 @@ if(Sys.info()['user']=='s7m'){
 	simResPath = paste0(dPath, 'simulation/')
 	source(paste0(fPath, 'functions.R')) }
 
+if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
+	base=paste0('C:/Users/',Sys.info()['user'],'/')
+	gPath = paste0(base, 'Research/netsMatter/')
+	fPath = paste0(gPath, 'code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste(gPath, 'paper/')
+	source(paste0(fPath, 'functions.R')) }
+
 toLoad = c(
 	'devtools',
 	'foreach', 'doParallel',
 	'magrittr', 'dplyr', 'ggplot2' )
 loadPkg(toLoad)
-devtools::install_github('s7minhas/amen', ref='dev')
+devtools::install_github('s7minhas/amen')
+library(amen)
 ##############################
 
 ##############################
 ### Homophily
 simRun = function(seed, n, mu, beta, gamma){
 	set.seed(seed)
-	xw = matrix(rnorm(n*2),n,2)
-	X = tcrossprod(xw[,1])  ; W = tcrossprod(xw[,2])
+	x=matrix(rnorm(n),n,1); X=tcrossprod(x[,1]); W = X^2
 	XW = array(dim=c(n,n,2)) ; XW[,,1] = X ; XW[,,2] = W
 
 	# create DV
-	Y<- 1*(mu + beta*X + gamma*W + matrix(rnorm(n*n),n,n) >0)
+	Y<- mu + beta*X + gamma*W + matrix(rnorm(n*n),n,n)
 
 	# run AME models
-	fit0 = ame(Y,X,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
+	fit0 = ame(Y,X,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='nrm',
 		print=FALSE,plot=FALSE)
-	fit1 = ame(Y,X,R=1,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
+	fit1 = ame(Y,X,R=1,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='nrm',
 		print=FALSE,plot=FALSE)
[…]

### 2020-06-04 revisionMemo/rev1.tex
sync up sim code

@@
 		\item  \emph{ \textcolor{blue}{
+		We thank the reviewer for taking the time to provide these valuable comments.
 We have clarified our language in the paper (on page 1) to address this question.
 Strategies like those employed by Aronow et al.
-do not directly model all relevant dependencies in the data but instead adjust the variance estimator.\footnote{As Aronow et al.
-explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data." p.
-16.}   While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will lead us to falsely accept the null).
+do not directly model all relevant dependencies in the data but instead adjust the variance estimator.
+As Aronow et al.
+explain, "Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data" (p.
+16).
+While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will lead us to falsely accept the null).
 The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
@@
 		}}
-	\item \textcolor{blue}{ \emph{ Additionally, we have SM: needs to add more here to indicate how the existing simulation section is meant to deal with this issue ...
-onus is on SM to bring in the problems that we highlight in the simulation section to section 2, so readers are clearer on the problem that we are trying to address
+		\item \textcolor{blue}{ \emph{ Additionally, we have added language motivating whatn is to come in the simulation section.
 	}}
@@
 		\item \textcolor{blue}{ \emph{
+		We thank the reviewer for providing this reference and have added discussion of this alternative to the manuscript.
 		SM insert great response.

### 2020-06-08 revisionMemo/rev2.tex
memo edits
add additional sim case

@@
 In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see Plumper and Neumayer 2010 for a more in-depth discuss on spatial lags).
-\textbf{CD}: I added this as a footnote on p.10.
+\textbf{CD}: I added this as a footnote on p.9.
 I moved this footnote around a LOT but I think it needs to live somewhere in the simulation.
@@
 			}}
-
 %		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
@@
 %		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
-
 	\end{itemize}
@@
 		\item \textcolor{blue}{ \emph{
-		CDMG: Flag for discussion.
-		in the appendix add more technical description of the estimation process
-		SM: figuring out a simulation case where ame fails to show i guess the limitations of it as an estimator ...
-		get stats from simulation
+		add example of case where omitted variable is correlated
+		tech discussion bias-variance
 		}}
@@
 We have now emphasized the importance of developing theory in our discussion of the AME and Oracle comparison (p.
-X), as well as in the conclusion of our paper (p.Y).
+13), as well as in the conclusion of our paper.
 We agree that innovations in dependence modeling can and should speak to theoretical advancements that, like our model, take into account the important relational and interdependent nature of political processes.

### 2020-06-08 code/1_ameSim_rev2_correl.R

@@ -1,0 +1,151 @@
+##############################
+rm(list=ls())
+
+if(Sys.info()['user']=='s7m'){
+	fPath = '~/Research/netsMatter/code/helpers/'
+	dPath = '~/Dropbox/Research/netsMatter/'
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R')) }
+
+if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
+	base=paste0('C:/Users/',Sys.info()['user'],'/')
+	gPath = paste0(base, 'Research/netsMatter/')
+	fPath = paste0(gPath, 'code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste(gPath, 'paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+  'amen' )
+loadPkg(toLoad)
+##############################
+
+##############################
+### Homophily
+simRun = function(seed, n, mu, beta, gamma, rho){
+
+  # cor mat
+  theta = asin(rho)/2
+  C = matrix( c(cos(theta),sin(theta),sin(theta),cos(theta)),2,2)
+
+  # seed
+  set.seed(seed)
+
+  # gen covars
+  xw = matrix(rnorm(n*2),n,2) %*% C
+  X = tcrossprod(xw[,1])
+  W = tcrossprod(xw[,2])
+  XW = array(dim=c(n,n,2)) ; XW[,,1]<-X ; XW[,,2]<-W
+
+	# create DV
+	Y = mu + beta*X + gamma*W + matrix(rnorm(n*n),n,n)
+
+	# run AME models
+	fit0 = ame(Y,X,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,
+    model='nrm',
+		print=FALSE,plot=FALSE)
+	fit1 = ame(Y,X,R=1,rvar=FALSE,cvar=FALSE,dcor=FALSE,
+    model='nrm',
+		print=FALSE,plot=FALSE)
+	fitO = ame(Y,XW,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,
+    model='nrm',
+		print=FALSE,plot=FALSE)
+
+	# gather together results
[…]

### 2020-06-08 code/1_simulation/2_ameBias.R
add additional sim case

@@ -7,12 +7,6 @@ if(Sys.info()['user']=='s7m'){
 	graphicsPath = paste('~/Research/netsMatter/paper/')
 	source(paste0(fPath, 'functions.R')) }
 
-if(Sys.info()['user']=='maxgallop'){
-  fPath = '~/Documents/netsMatter/code/helpers/'
-  dPath = '~/Dropbox/netsMatter/'
-  simResPath = paste0(dPath, 'simulation/')
-  source(paste0(fPath, 'functions.R')) }
-
 if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
 	base=paste0('C:/Users/',Sys.info()['user'],'/')
 	gPath = paste0(base, 'Research/netsMatter/')

### 2020-06-08 code/1_simulation/1_ameSim.R
add additional sim case

@@ -7,6 +7,15 @@ if(Sys.info()['user']=='s7m'){
 	simResPath = paste0(dPath, 'simulation/')
 	source(paste0(fPath, 'functions.R')) }
 
+if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
+	base=paste0('C:/Users/',Sys.info()['user'],'/')
+	gPath = paste0(base, 'Research/netsMatter/')
+	fPath = paste0(gPath, 'code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste(gPath, 'paper/')
+	source(paste0(fPath, 'functions.R')) }
+
 toLoad = c(
 	'devtools',
 	'foreach', 'doParallel',

### 2020-06-09 revisionMemo/rev1.tex
additional sim cases

@@
 As Aronow et al.
-explain, "Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data" (p.
+explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data'' (p.
 16).
@@
 		\item \textcolor{blue}{ \emph{
-		SM will deal with this.
+		We've added additional discussion about how each of the variables in our simulation were generated.
 		}}

### 2020-06-09 paper/3_sim.tex
additional sim cases

@@
 \end{align}
+
+x is a vector of observed node-specific characteristics and w is a vector of unobserved node specific characteristics.
+We draw n values for both from a standard normal.
+For both x and we we then take the cross-product of their transpose to generate n x n matrices.
+Values for mu, beta, and gamma are set to -2, 1, and 1, respectively.
 

### 2020-06-09 paper/6_appendix.tex
additional sim cases

@@
 \clearpage
+
+\section{Correlation with Omitted Variable}
+
+Things get more challenging when the omitted variable is correlated with the observed variable.
+In the case of normal linear regression you can show that in general a regression coefficient is unbiased for its "unconditional effect", which is the direct effect plus something related to the direct effects of any omitted variables that it is correlated with.
+For example, suppose $w_{i,j}$ is correlated with $x_{i,j}$ via the relation $w_{i,j} = \alpha x_{i,j} + z_{i,j}$.
+Then:
+
+\begin{align*}
+y_{i,j} & = \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j} \\
+&= \mu + (\gamma \alpha+\beta ) x_{i,j} + (\gamma z_{i,j} + \epsilon_{i,j})  \\
+&= \mu + \tilde \beta x_{i,j} + \tilde \epsilon_{i,j}
+\end{align*}
+
+The naive model is in some sense still "correct" as it is an unbiased estimator of the unconditional (direct plus indirect) effect $\gamma\alpha + \beta$ of $x$ on $y$.
+
+As far as I know there is no way to eliminate this extra biased without making further assumptions about the nature of the omitted variable.
+However, an interesting question is whether or not we can reduce the bias in the network setting by assuming the omitted variable $w_{i,j}$ is the product of latent node-specific factors.
+An idea related to this is discussed in \citet{minhas:etal:2017:arxiv}.
 

### 2020-06-09 code/1_simulation/asa_robust_repl/2-dyad-sims-misspec-shocks-2015-02-23.R
additional sim cases

@@ -1,111 +1,0 @@
-library(sandwich)
-library(lmtest)
-library(arm)
-
-rm(list=ls())
-
-
-# Cluster robust variance estimation function
-robust.se.nodfc <- function(model, cluster){
-	require(sandwich)
-	require(lmtest)
-	M <- length(unique(cluster))
-	N <- length(cluster)
-	K <- model$rank
-	dfc <- 1
-	uj <- apply(estfun(model), 2, function(x) tapply(x, cluster, sum))
-	rcse.cov <- dfc * sandwich(model, meat = crossprod(uj)/N)
-	rcse.se <- coeftest(model, rcse.cov)
-	return(list(rcse.cov, rcse.se))
-}
-
-
-#################
-# Global settings
-#################
-
-nsim <- 500
-
-#################
-# Cross-sectional: misspecification
-#################
-
-#N <- 150
-
-for(N in c(20,50,100,150)){
-
-index <- 1:N
-P <- choose(N,2)
-
-# Dyads
-
-dyads <-  as.matrix(apply(t(combn(index, 2)), 1,
-					function(x)paste(x[1],
-					x[2],sep="-")))
-
-dyad.mat <- t(apply(dyads, 1, function(x)unlist(strsplit(x,"-"))))
-
-# Simulation
-
-bre.irreg <- seRE.irreg <- seHC.irreg <- sehat.irreg <- bhat.irreg <- matrix(NA, ncol=2, nrow=nsim)
-
-for(sim in 1:nsim){
-cat(paste("Sim #",sim,":",sep="")); flush.console()
-# Generate the data
-
-a <- rnorm(N)
-X <- rnorm(N)
-
-# Misspec
-
-dX <- da <- NA
-for(i in 1:P){
-da[i] <- sum(a[as.numeric(dyad.mat[i,])])
-dX[i] <- abs(diff(X[as.numeric(dyad.mat[i,])]))
-}
-
-dY <- da + dX + .25*dX^2 + rnorm(P)
-
-dataUp <- data.frame(dyads,
-					 dY, dX,
-					 dyad1 = as.numeric(dyad.mat[,1]),
-					 dyad2 = as.numeric(dyad.mat[,2]))
-
-
-fit <- lm(dY~dX, data=dataUp)
-bhat.irreg[sim,] <- coef(fit)
-
[…]

### 2020-06-09 code/1_simulation/rev2_correl/4_corUVandMiss.r

@@ -1,0 +1,106 @@
+##############################
+rm(list=ls())
+if(Sys.info()['user']=='s7m'){
+	fPath = '~/Research/netsMatter/code/helpers/'
+	dPath = '~/Dropbox/Research/netsMatter/'
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste('~/Research/netsMatter/paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
+	base=paste0('C:/Users/',Sys.info()['user'],'/')
+	gPath = paste0(base, 'Research/netsMatter/')
+	fPath = paste0(gPath, 'code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste0(gPath, 'paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+	'latex2exp', 'Cairo'
+	)
+loadPkg(toLoad)
+facet_labeller = function(string){ TeX(string) }
+##############################
+
+##############################
+# params
+NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# load sim results
+# for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'_asa.rda')) }
+##############################
+
+##############################
+# check that uv covers missvar
+# missing var
+genMissVar = function(n, imps=NSIM){
+	lapply(1:imps, function(seed){ n=n ;
+		set.seed(seed) ; xw = matrix(rnorm(n*2),n,2)
+		W = tcrossprod(xw[,2]) ; return(W) }) }
+
+#
+getCor = function(ameSim, W){
+	cor = lapply(1:length(ameSim), function(i){
+		uv = c(ameSim[[i]]$uv$ame)
+		w = c(W[[i]])
+		cor(uv, w, use='pairwise.complete.obs') })
+	return(unlist(cor)) }
+
+#
+ameSimCor = rbind(
[…]

### 2020-06-09 code/1_simulation/rev2_correl/5_gof.R

@@ -1,0 +1,45 @@
+##############################
+rm(list=ls())
+if(Sys.info()['user']=='s7m'){
+	fPath = '~/Research/netsMatter/code/helpers/'
+	dPath = '~/Dropbox/Research/netsMatter/'
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste('~/Research/netsMatter/paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R')) }
+
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+	'latex2exp', 'Cairo'
+	)
+loadPkg(toLoad)
+facet_labeller = function(string){ TeX(string) }
+##############################
+
+##############################
+# params
+NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# load sim results
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
+
+#
+modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
+modKey$clean = c('Naive', 'AME', 'Oracle')
+##############################
+
+##############################
+loadPkg('amen')
+gofPlot(ameSim50[[1]]$gof$naive, FALSE)
+gofPlot(ameSim50[[1]]$gof$ame, FALSE)
+gofPlot(ameSim50[[1]]$gof$oracle, FALSE)
+##############################

### 2020-06-09 code/1_simulation/asa_robust_repl/5-russett-oneal-reanalysis.R
additional sim cases

@@ -1,57 +1,0 @@
-library(foreign)
-library(sandwich)
-library(Matrix)
-rm(list=ls())
-
-# Cluster robust variance estimation function
-robust.se.nodfc <- function(model, cluster){
-	require(sandwich)
-	require(lmtest)
-	M <- length(unique(cluster))
-	N <- length(cluster)
-	K <- model$rank
-	dfc <- 1
-	uj <- apply(estfun(model), 2, function(x) tapply(x, cluster, sum))
-	rcse.cov <- dfc * sandwich(model, meat = crossprod(uj)/N)
-	rcse.se <- coeftest(model, rcse.cov)
-	return(list(rcse.cov, rcse.se))
-}
-
-triangle <- read.dta("~/Dropbox/dyadic-variance/pa-submission/replication-files/TRIANGLE-out.dta")
-regSpec <- paste("dispute1~", 
-				paste("allies","lcaprat2","smldmat","smldep",
-					"smigoabi","noncontg","logdstab","minrpwrs",
-				sep="+"), sep="")
-
-regForm <- as.formula(regSpec)
-fit <- glm(regForm, data=triangle, x=T, family="binomial")
-
-index.sc <- unique(c(triangle$statea, triangle$stateb))
-index <- index.sc[order(index.sc)]
-dyad.mat <- triangle[,c("statea","stateb")]
-
-# Dyadic cluster robust via multiway decomposition
-for(i in 1:length(index)){
-	iUp <- index[i]
-	clusUp <- apply(dyad.mat, 
-					1,
-					function(x)as.numeric(iUp %in% x))
-	clusIndexUp <- clusUp*-99 + (1-clusUp)*1:nrow(dyad.mat)
-	if(i==1){dcrUp <- robust.se.nodfc(fit, clusIndexUp)[[1]]}
-	if(i>1){dcrUp <- dcrUp + robust.se.nodfc(fit, clusIndexUp)[[1]]}	
-	cat(paste(iUp, " ")); flush.console()
-}
-# substract naive CR:
-dcrUp2 <- dcrUp - robust.se.nodfc(fit, triangle$dyadid)[[1]]
-# substract HR:
-Vhat <- dcrUp2 - (length(index)-2)*vcovHC(fit,type="HC0")
-sqrt(diag(Vhat))
-
-abs(coef(fit)/sqrt(diag(Vhat))) < 1.64
-
-
-
-
-
-
-

### 2020-06-09 code/1_simulation/asa_robust_repl/1-dyad-sims-2015-02-23.R
additional sim cases

@@ -1,196 +1,0 @@
-library(sandwich)
-library(lmtest)
-library(arm)
-
-rm(list=ls())
-
-
-# Cluster robust variance estimation function
-robust.se.nodfc <- function(model, cluster){
-	require(sandwich)
-	require(lmtest)
-	M <- length(unique(cluster))
-	N <- length(cluster)
-	K <- model$rank
-	dfc <- 1
-	uj <- apply(estfun(model), 2, function(x) tapply(x, cluster, sum))
-	rcse.cov <- dfc * sandwich(model, meat = crossprod(uj)/N)
-	rcse.se <- coeftest(model, rcse.cov)
-	return(list(rcse.cov, rcse.se))
-}
-
-robust.se <- function(model, cluster){
-	require(sandwich)
-	require(lmtest)
-	M <- length(unique(cluster))
-	N <- length(cluster)
-	K <- model$rank
-	dfc <- (M/(M - 1)) * ((N - 1)/(N - K))
-	uj <- apply(estfun(model), 2, function(x) tapply(x, cluster, sum))
-	rcse.cov <- dfc * sandwich(model, meat = crossprod(uj)/N)
-	rcse.se <- coeftest(model, rcse.cov)
-	return(list(rcse.cov, rcse.se))
-}
-
-#################
-# Global settings
-#################
-
-nsim <- 500  # To fix the number of simuluation runs
-
-
-#################
-# Cross-sectional
-#################
-
-# for(N in c(20,50,100,150)){
-N = 50
-
-index <- 1:N
-P <- choose(N,2)
-
-# Dyads
-
-dyads <-  as.matrix(apply(t(combn(index, 2)), 1,
-					function(x)paste(x[1],
-					x[2],sep="-")))
-
-dyad.mat <- t(apply(dyads, 1, function(x)unlist(strsplit(x,"-"))))
-
-# Simulation
-
-seRE <- bRE <- seHC <- sehat <- bhat <- matrix(NA, ncol=2, nrow=nsim)
-
-# for(sim in 1:nsim){
-sim=1
-cat(paste("Sim #",sim,":",sep="")); flush.console()
-# Generate the data
-
-a <- rnorm(N) 
-X <- rnorm(N)
-
[…]

### 2020-06-09 code/1_simulation/rev2_correl/2_ameBias.R

@@ -1,0 +1,123 @@
+##############################
+rm(list=ls())
+if(Sys.info()['user']=='s7m'){
+	fPath = '~/Research/netsMatter/code/helpers/'
+	dPath = '~/Dropbox/Research/netsMatter/'
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste('~/Research/netsMatter/paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
+	base=paste0('C:/Users/',Sys.info()['user'],'/')
+	gPath = paste0(base, 'Research/netsMatter/')
+	fPath = paste0(gPath, 'code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	graphicsPath = paste0(gPath, 'paper/')
+	source(paste0(fPath, 'functions.R')) }
+
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+	'latex2exp', 'Cairo'
+	)
+loadPkg(toLoad)
+facet_labeller = function(string){ TeX(string) }
+##############################
+
+##############################
+# params
+NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# iterate over correlation levels
+corrLev = '_corrHi'
+
+# load sim data
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,corrLev,'.rda')) }
+
+#
+modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
+modKey$clean = c('Standard', 'AME', 'Oracle')
+##############################
+
+##############################
+# check bias of x1
+getBiasDF = function(ameSim){
+	biasDF = lapply(ameSim, function(x){
+		betaMod=lapply(x$beta, function(z){ apply(z, 2, median)[1:2] })
+		betaMod = betaMod %>% reshape2::melt() %>%
+			dplyr::mutate(
+				var=rep(c('Intercept','X1'), length(x$beta)),
+				act = rep(c(intEff, x1Eff), length(x$beta)),
+				bias = value - act ) %>%
+			dplyr::rename(model = L1)
+		return(betaMod) })
+	return( suppressMessages( reshape2::melt(biasDF,id=names(biasDF[[1]])) ) ) }
+
+#
[…]

### 2020-06-12 revisionMemo/rev1.tex
add graphs

@@
 		SM will deal with this.
+		\begin{figure}
+			\centering
+			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+			\label{fig:ameBias}
+			\includegraphics[width=1\textwidth]{ameSimBias_all_asaProbit.pdf} \\
+		\end{figure}
+
+		\begin{figure}
+			\centering
+			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+			\label{fig:ameBias}
+			\includegraphics[width=1\textwidth]{ameSimBias_all_asa.pdf} \\
+		\end{figure}
+
+		\begin{figure}
+			\centering
+			\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
+			\label{fig:ameCalib}
+			\includegraphics[width=1\textwidth]{ameSimCover_all_asaProbit.pdf} \\
+		\end{figure}
+
+		\begin{figure}
+			\centering
+			\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
+			\label{fig:ameCalib}
+			\includegraphics[width=1\textwidth]{ameSimCover_all_asa.pdf} \\
+		\end{figure}		
 		}}

### 2020-06-12 revisionMemo/rev2.tex
additional sim cases
add graphs

@@
 		tech discussion bias-variance
+
+		\begin{figure}[ht]
+		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+		\label{fig:ameBias}
+		\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitMed.pdf} \\
+		\end{figure}
+
+		\begin{figure}[ht]
+		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+		\label{fig:ameBias}
+		\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitHi.pdf} \\
+		\end{figure}
+
+		\begin{figure}[ht]
+		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+		\label{fig:ameBias}
+		\includegraphics[width=1\textwidth]{ameSimBias_all_corrMed.pdf} \\
+		\end{figure}
+
+		\begin{figure}[ht]
+		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+		\label{fig:ameBias}
+		\includegraphics[width=1\textwidth]{ameSimBias_all_corrHi.pdf} \\
+		\end{figure}		
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		SM will deal with this.
+		We've added additional discussion about how each of the variables in our simulation were generated.
+To make the model more applicable to the types of networks we study in IR we have provided a probit example.
 		}}

### 2020-06-12 revisionMemo/rev1.tex
formatting

@@
 	\begin{itemize}
-		\item  \emph{ \textcolor{blue}{
+		\item \textcolor{blue}{ \emph{
 		We thank the reviewer for taking the time to provide these valuable comments.
@@
 		}}
-		%\item \textcolor{blue}{ \emph{
-		%Aronow et al. also explicitly note that their approach does not deal with issues of third order dependence.
-		%}}
-		%\item \textcolor{blue}{ \emph{
-		%Aronow et al. approach does not deal with bias in parameter estimates.
-		%}}
 	\end{itemize}
@@
 	\begin{itemize}
-		\item {\emph{\textcolor{blue}{
+		\item \textcolor{blue}{ \emph{
 		 The point is well taken, we have clarified this section (on page 2) to point to the assumption of independence as the primary problem, not the stacking of observations in and of itself.
-		}}}
+		}}
 	\end{itemize}
@@
 For me at least, after reading section 2 I still wasn’t sure what problem the paper is actually trying to address.
-		\item \textcolor{blue}{ \emph{ We think the reviewer makes a good suggestion here to clarify, mathematically or in some other illustrative way, exactly how we think that dependencies bias the estimator.
+		\item \textcolor{blue}{ \emph{
+		We think the reviewer makes a good suggestion here to clarify, mathematically or in some other illustrative way, exactly how we think that dependencies bias the estimator.
 For example we can say that our target of inference is the estimation of beta (with some variance).
@@
 	\begin{itemize}
-		\item  \emph{\textcolor{blue}{
+		\item \textcolor{blue}{ \emph{
 			For part (a), we have clarified that our study does not address causal identification.
@@
 We have clarified the goals of the paper in the earlier sections (such as the opening paragraph of page 1) and have also edited P8 to read: ``Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe." }}
-		\item \emph{\textcolor{blue}{
+		\item \textcolor{blue}{ \emph{
 			As for (b), the reviewer raises an important point.
@@
 }}
-		\item \emph{\textcolor{blue}{Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others.}
+		\item \textcolor{blue}{ \emph{
+			Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others.}
 		}

### 2020-06-12 revisionMemo/rev2.tex
formatting

@@
 Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
-In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for \citet{colgan:2010}'s theory that revolutionary petrostates are more aggressive.")
+In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for Colgan's (2010) theory that revolutionary petrostates are more aggressive.")
 				}}

### 2020-06-16 paper/3_sim.tex
wrapping up

@@
 Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
-The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.\footnote{\textcolor{blue}{The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria.
+The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.\footnote{The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria.
 With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l).
-In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).}} Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}
+In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).} Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}
 
@@
 Scholars can use this framework in an iterative fashion: beginning with an estimated model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their model.
-\textcolor{blue}{Importantly, this simulation underscores how a careful consideration of a systems' interconnectedness, both through theoretical approaches and empirical models, can result in more precise estimates of direct and indirect effects across the system.}
+Importantly, this simulation underscores how a careful consideration of a systems' interconnectedness, both through theoretical approaches and empirical models, can result in more precise estimates of direct and indirect effects across the system.
 

### 2020-06-16 paper/1_intro.tex
wrapping up

--- 2019-07-15 paper/1_intro.tex
@@
 Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
-In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
+In practice, such strategies remain insufficient because they do not directly model all relevant dependencies in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, “Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data."} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null)} The main goal of our study is to directly account for the relevant dependency structures in the data, and thus provide an accurate estimate while addressing both types of possible errors.
 This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.
 
-In this article, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
+In this article, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2019}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2020-06-16 paper/2_model.tex
wrapping up

@@
 
-\textcolor{blue}{Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.} Accounting for this structure is necessary if we are to adequately represent the data generating process.
+Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.
+Accounting for this structure is necessary if we are to adequately represent the data generating process.
 If one can specify each of the nodal, dyadic, and triadic attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied.

### 2020-06-16 paper/4_app.tex
memo edits again
theory for random effects
wrapping up

@@
 
-We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{\textcolor{blue}{We selected our cases based on a few criteria.
+We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{We selected our cases based on a few criteria.
 We focus on studies that are explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we focus on one journal (American Political Science Review).
-We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.}} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.
+We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.
 Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model.
@@
 Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
+In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for \citet{colgan:2010}'s theory that revolutionary petrostates are more aggressive.
 None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.

### 2020-07-18 paper/4_app.tex
sync up

@@
 Table~\ref{tab:modelDesign} provides descriptive information for the studies that we replicated.
-% \footnote{It is important to note that the AME framework is not well suited for studies where the focus may be on studying onsets as in these type of studies dyad-years representing conflict continuations are removed from the sample.
-In studies with this type of focus, network models in general are of limited utility since any interdependencies underlying how events may spread from one actor to another are explicitly masked from the estimation process due to the removal of observations.} ...........
-gibler uses a study onsets.
-in some ways this is problematic in other it is not, i'm (sm) suggesting that we ignore for now.
 
@@
 For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimate the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
-Results with alternative values of $K$ are similar.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
+Results with alternative values of $K$ are similar.
+We recommend that authors try out a range of $K$s} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
 In the directed formulation, separate random effects are used for senders and receivers in both the additive and multiplicative portions of the model.

### 2020-07-18 code/1_simulation/2_ameBias.R

@@ -20,9 +20,10 @@ toLoad = c(
 	'devtools',
 	'foreach', 'doParallel',
 	'magrittr', 'dplyr', 'ggplot2',
-	'latex2exp', 'Cairo'
+	'latex2exp', 'Cairo', 'extrafont'
 	)
 loadPkg(toLoad)
+suppressMessages(loadfonts())
 facet_labeller = function(string){ TeX(string) }
 ##############################
 
@@ -31,8 +32,7 @@ facet_labeller = function(string){ TeX(string) }
 NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 # load sim results
-# for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
-for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'_asa.rda')) }
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
 
 #
 modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
@@ -85,9 +85,6 @@ ggBiasPlot = function(varName, h=4, w=8){
 		geom_hline(aes(yintercept=act), color='grey60', size=2) +
 		geom_jitter(alpha=.1) +
 		geom_boxplot( color='black') +
-		# geom_violin(
-		# 	draw_quantiles=c(0.025,0.5,0.975),
-		# 	trim=TRUE, color='black', alpha=.4) +
 		facet_grid(var ~ n, scales='free_y',
 			labeller=as_labeller(facet_labeller, default = label_parsed)) +
 		xlab('') + ylab('') +
@@ -98,28 +95,26 @@ ggBiasPlot = function(varName, h=4, w=8){
 			legend.title=element_blank(),
 			axis.ticks=element_blank(),
 			panel.border=element_blank(),
-			axis.text.y=element_text(size=8
-				# , family="Source Code Pro Light")
+			axis.text.y=element_text(
+				size=8, family="Source Code Pro Light"
 			),
 			axis.text.x=element_text(size=10, face='bold'),
-			strip.text.x = element_text(size=9, color='white'
-				# ,family="Source Code Pro Semibold"
+			strip.text.x = element_text(
+				size=9, color='white',family="Source Code Pro Semibold"
 				),
-			strip.text.y = element_text(size=9, color='white'
-				# ,family="Source Code Pro Semibold",
-				,angle=0
[…]

### 2020-07-18 paper/2_model.tex
sync up

@@
 
-Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another and not passing any auxiliary information to the model about how observations may be interdependent.
+When modeling dyadic data scholars typically employ a generalized linear model (GLM) estimated via maximum-likelihood.
+This type of model is expressed via a stochastic and systematic component \citep{pawitan:2013}.
+The stochastic component reflects our assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function, such as the normal, binomial, or Poisson, and we assume that each dyad in the sample is independently drawn from that particular distribution, given $\theta_{ij}$.
+The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariatse, $\mathbf{X_{ij}}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$.
+A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of our distribution, each of the dyadic observations are conditionally independent.
+The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood.
+After having chosen a set of covariates and specifying a distribution, we construct the joint density function over all dyads, for example:
+
+\vspace{-8mm}
+\begin{align}
+\begin{aligned}
+	P(y_{ij}, y_{ik}, \ldots | \theta_{ij}, \theta_{ik}, \ldots ) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots  \\
+	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
+\end{aligned}
+\end{align}
+
+\noindent We next convert the joint probability into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.
+
+We can then estimate the parameters by maximizing the likelihood, or, more typically, the log-likelihood.
+However, the likelihood as defined above is only valid if we are able to make the assumption that, for example, $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$.
+Without the assumption of conditional observational independence the joint density function cannot be written in the way described above and a valid likelihood does not exist.
+Accordingly, inferences drawn from misspecified models that ignore potential interdependencies between dyadic observations are likely to have a number of issues including biased estimates of the effect of independent variables, uncalibrated confidence intervals, and poor predictive performance.
+The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
+As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
+
+% Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another and not passing any auxiliary information to the model about how observations may be interdependent.
 For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.
@@
 It is flexible and easy to use for analyzing the kind of relational data often found in the social sciences.
-It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2019} detail how this framework contrasts with alternative network-based approaches.} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.
-For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:2005}.
-The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.} The AME model is specified as follows:
+It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2019} detail how this framework contrasts with alternative network-based approaches.
+} The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:2005}.
+The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.}  For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}.
+The AME model is specified as follows:
 
@@
 $K$ denotes the dimensions of the latent space.
-The construction of the LFM here is actually quite similar to the recommender systems that companies like Amazon and Netflix use to model customer behavior \citep{resnick:varian:1997,bennett:lanning:2007}.
-This model posits a latent vector of characteristics $\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
+The construction of the LFM here is actually quite similar to work on low rank approximations in computer science and has been applied to the development of recommender systems that companies like Amazon and Netflix use to model customer behavior \citep{resnick:varian:1997,bennett:lanning:2007}.\footnote{The LFM also shares similarities with work in the econometric literature on interactive fixed effects \citep{bai:2009,pang:2014}.
+In this stream of work, interactive fixed effects are used to deal with cross-sectional dependence in TSCS data, in such a way that a latent factor for time can be used to capture common shocks to actors and a latent factor on actors can capture varying responses to those shocks.} This model posits a latent vector of characteristics
+$\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$.
 The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third-order interdependencies.

### 2020-07-18 paper/6_appendix.tex
sync up

--- 2020-01-23 paper/6_appendix.tex
@@
 \newpage
+
 \section{Additional Replication Information}
@@
 % \end{figure}
+\FloatBarrier
+\clearpage
+
+\section{Additional Simulations: Non-Linear Data}
+
+Aronow et al note that ``fitting a linear approximation to mildly non-linear data'' can be problematic for dyadic random effect models.
+We agree on the importance of testing this in the context of the simulation that we posed in the paper.
+Aronow et al incorporate non-linear misspecification by adding in squared version of the key dyadic homophily variable to the DGP for their simulations: $Y_{i,j} = \beta_{0} + \beta_{1} X_{i,j} + \beta_{2} X_{i,j}^{2} + \epsilon_{i,j}$.
+Here $X$ is their representation of homophily and is considered observed, while $X^{2}$ is their check for robustness to misspecification and is unobserved.
+We incorporate the same type of misspecification into our probit simulation setup: $Z_{i,j} =  \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j}$, where $W$ now instead of being drawn from a separate normal distribution is just the squared version of $X$.
+For the simulation we set values of $\mu$, $\beta$, and $\gamma$ to -2, 1, and 1, respectively.
+Results are shown below in Figures~\ref{fig:ameBias_asa} and \ref{fig:ameCalib_asa} below.
+As in the paper, along the x-axis we implement a standard model that does not take any steps to account for dependencies, AME, and the Oracle model, and as in the paper we find that the AME approach provides a number of benefits in terms of bias and coverage over the standard approach of not dealing with dependencies.
+
+\begin{figure}[ht]
+	\centering
+	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
+	\label{fig:ameBias_asa}
+	\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_asaProbit.pdf} \\
+\end{figure}
+
+\begin{figure}[ht]
+	\centering
+	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
+	\label{fig:ameCalib_asa}
+	\includegraphics[width=1\textwidth]{graphics/ameSimCover_all_asaProbit.pdf} \\
+\end{figure}
+
+\FloatBarrier
+\clearpage
+
+\section{Additional Simulations: Correlation with Omitted Variable}
+
+Estimation certainly gets more challenging when the omitted variable is correlated with the observed variable.
+In the case of normal linear regression, we can show that in general a regression coefficient is unbiased for its ``unconditional effect'', which is the direct effect plus something related to the direct effects of any omitted variables that it is correlated with.
+For example, suppose in our simulation setup $X_{i,j}$ is correlated with $W_{i,j}$ via the relation $W_{i,j} = \alpha X_{i,j} + e_{i,j}$.
+Then:
+
+\begin{align*}
+  Y_{i,j} & = \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j} \\
+  &= \mu + (\gamma \alpha+\beta ) X_{i,j} + (\gamma Z_{i,j} + \epsilon_{i,j})  \\
+  &= \mu + \tilde \beta X_{i,j} + \tilde \epsilon_{i,j}
+\end{align*}
+
+The naive model is in some sense still ``correct'' as it is an unbiased estimator of the unconditional (direct plus indirect) effect $\gamma\alpha + \beta$ of $X$ on $Y$.
+However, there is no way to eliminate this extra bias without making further assumptions about the nature of the omitted variable.
+Additonally, in the binary case with the probit setup, other sources of bias get introduced as well because of the nonlinearity of the population moments as a function of the parameters.
+An interesting question is whether or not we can reduce the bias in the network setting by assuming the omitted variable $w_{i,j}$ is the product of latent node-specific factors.
+An idea related to this is discussed in \citet{minhas:etal:2017:arxiv}.
+
+For now though to show the amount of bias that gets introduced into our setup when $X$ and $W$ are correlated we repeat our simulation exercise but induce varying levels of correlation between $X$ and $W$.
+Results with $X$ and $W$ correlated at 0.4 and then 0.7 are shown below, respectively.
+
+\begin{figure}[ht]
+\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.
+$X$ and $W$ are correlated at 0.4 in this case.}
+\label{fig:ameBias_corrMed}
+\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_corrProbitMed.pdf} \\
+\end{figure}
+
+\begin{figure}[ht]
+\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.
+$X$ and $W$ are correlated at 0.7 in this case.}
+\label{fig:ameBias_corrHi}
+\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_corrProbitHi.pdf} \\
+\end{figure}
+
 \FloatBarrier
@@
 One strategy is to estimate models at different values of $\sf{R}$ and compare goodness of fit statistics across models.
-
-Given the computational intensity needed for parameter estimates to converge, parallelization strategies are recommended to speed up analysis.
+Given the computational time needed for parameter estimates to converge, parallelization strategies are recommended to speed up analysis.
 In addition, providing AMEN function with starting values, either dictated by theory, previous research, or previous runs can also help speed up convergence time.
@@
 The code below presents an example of an AME model running in parallel across 4 different levels of $\sf{R}$.
-Note also that the model is using starting values from a previous run, defined in \textit{startVals0}.
+Note also that the model is using starting values from a previous run, defined in \textit{startVals}.
 

### 2020-07-18 code/1_simulation/1_ameSim.R

@@ -19,9 +19,9 @@ if(Sys.info()['user'] %in% c('herme','Owner','S7M')){
 toLoad = c(
 	'devtools',
 	'foreach', 'doParallel',
-	'magrittr', 'dplyr', 'ggplot2' )
+	'magrittr', 'dplyr' )
 loadPkg(toLoad)
-devtools::install_github('s7minhas/amen', ref='dev')
+library(amen)
 ##############################
 
 ##############################
@@ -33,7 +33,7 @@ simRun = function(seed, n, mu, beta, gamma){
 	XW = array(dim=c(n,n,2)) ; XW[,,1] = X ; XW[,,2] = W
 
 	# create DV
-	Y<- 1*(mu + beta*X + gamma*W + matrix(rnorm(n*n),n,n) >0)
+	Y = 1*(mu + beta*X + gamma*W + matrix(rnorm(n*n),n,n) >0)
 
 	# run AME models
 	fit0 = ame(Y,X,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
@@ -55,40 +55,20 @@ simRun = function(seed, n, mu, beta, gamma){
 
 ##############################
 # params
-imps = 1000
+imps = 100
 intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 #
-getTime = function(){format(Sys.time(), '%X')}
-writeLines(c('n=30 using 8 cores\n'), paste0(simResPath,'ameSim30Log.txt'))
-cl=makeCluster(8) ; registerDoParallel(cl)
-ameSim30 = foreach(imp = 1:imps, .packages=c('amen')) %dopar% {
-	out=simRun(seed=imp, n=30, mu=intEff, beta=x1Eff, gamma=x2Eff)
-	sink(paste0(simResPath,'ameSim30Log.txt'), append=TRUE)
-	cat(paste0(imp, ' out of ', imps, ' completed: ', getTime(),'\n'))
-	return(out) }  ; stopCluster(cl)
-save( ameSim30, file=paste0(simResPath, 'ameSim30.rda') )
-
-#
-writeLines(c('n=50 using 8 cores\n'), paste0(simResPath,'ameSim50Log.txt'))
-cl=makeCluster(8) ; registerDoParallel(cl)
+cl=makeCluster(20) ; registerDoParallel(cl)
 ameSim50 = foreach(imp = 1:imps, .packages=c('amen')) %dopar% {
[…]

### 2020-07-18 paper/1_intro.tex
sync up

--- 2019-07-15 paper/1_intro.tex
@@
 
+The aim of this work is to address how to estimate the effect of exogenous covariates in a GLM context when there are network dependencies between observations in dyadic data.
 \citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).
@@
 Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
-In practice, such strategies rarely work because they do not directly address the fundamental data generating process.
-This is important to consider since the inferential problems caused by non-iid  observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.
+In practice, such strategies remain insufficient because they do not directly model underlying structural features in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data.''} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).} The aims of our study are twofold.
+First, we provide a framework that accounts for dependencies often present in dyadic data.\footnote{This is important to consider since the inferential problems caused by non-iid observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.} Second, the framework we discuss here enables applied scholars to understand the structure underlying dyadic data rather than simply controlling for them.
 
-In this article, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2008,minhas:etal:2018}.
+Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019}.
 We focus on three types of interdependencies that can complicate dyadic analyses.

### 2020-07-18 paper/3_sim.tex
sync up

@@
 
-Through a simulation study, we highlight the utility of AME as an inferential tool for dyadic analysis.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM).
+We explore the utility of AME as an inferential tool for dyadic analysis via a simulation exercise.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM).
 While both these models have led to numerous contributions to a variety of literatures, the applicability of these approaches may be limited to certain types of networks and individual level characteristics.
 Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable.
-The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of coefficient parameters in the presence of unobserved dependencies.\footnote{The types of geographic dependencies in international relations are generally monadic in nature -- civil war spreads from Rwanda to Uganda, or protest from Lebanon to Syria.
-With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l).
-In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see \citet{plumper:neumayer:2010} for a more in-depth discuss on spatial lags).} Assume that the true data-generating process for a particular $Y$ is given by:\footnote{Note that with the data generating process specified in Equation~\ref{eqn:sim} edge values in $Y$ will be normally distributed.}
+The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of regression coefficients in the presence of unobserved dependencies, specifically, homophily.
+As discussed in the previous section, homophily is the idea that actors are more likely to have a tie if they have similar values on a particular variable, and in networks the presence of homophily can lead to third order dependencies such as transitivity.
+Homophily can be operationalized by creating a dyadic covariate via the multiplication of a nodal covariate with its transpose.
+For example, if the nodal covariate was a binary indicator for democracy, multiplying it by its transpose would give us a dyadic covariate that represents whether any dyad is jointly democratic or not.\footnote{This process of operationalizing homophily is equivalent to the `nodematch` function in the `ergm` and `latentnet` packages.
+There are many other options of operationalizing homophily including calculating the difference in scores that a pair of actors may have on a particular nodal variable.}
+
+Assume that the true data-generating process for a binary variable, $Y$, is given by:
 
 \begin{align}
-	y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
+	Z_{i,j} &=  \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j}, \; \epsilon \sim normal(0,1) \nonumber \\
+	Y_{i,j} &= I(Z_{i,j} > 0)
 	\label{eqn:sim}
@@
 
-x is a vector of observed node-specific characteristics and w is a vector of unobserved node specific characteristics.
-We draw n values for both from a standard normal.
-For both x and we we then take the cross-product of their transpose to generate n x n matrices.
-Values for mu, beta, and gamma are set to -2, 1, and 1, respectively.
-
-where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics.
-$Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
+$X_{i,j} = x \times x^{T}$, where $x$ is a nodal covariate that is drawn from a standard normal distribution.
+Similarly,  $W_{i,j} = w \times w^{T}$, where $w$ is also a nodal covariate that is independently drawn from a standard normal distribution.
+We generate our binary dependent variable, $Y$, within a probit framework with $Z$ serving as the latent variable.
+$X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed.
 We compare inference for $\mu$ and $\beta$---the latter parameter would be of primary concern for applied scholars---using three models:

### 2020-07-18 code/1_simulation/3_ameCalib.R

@@ -20,9 +20,10 @@ toLoad = c(
 	'devtools',
 	'foreach', 'doParallel',
 	'magrittr', 'dplyr', 'ggplot2',
-	'latex2exp', 'Cairo'
+	'latex2exp', 'Cairo', 'extrafont'
 	)
 loadPkg(toLoad)
+suppressMessages(loadfonts())
 facet_labeller = function(string){ TeX(string) }
 ##############################
 
@@ -31,8 +32,7 @@ facet_labeller = function(string){ TeX(string) }
 NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
 
 # load sim results
-# for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
-for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'_asa.rda')) }
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'.rda')) }
 
 #
 modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
@@ -113,27 +113,25 @@ ggCoverPlot = function(var,h=3, w=8){
 			legend.title=element_blank(),
 			axis.ticks=element_blank(),
 			panel.border=element_blank(),
-			axis.text.y=element_text(size=8
-				# , family="Source Code Pro Light")
+			axis.text.y=element_text(
+				size=8, family="Source Code Pro Light")
 			),
 			axis.text.x=element_text(size=10, face='bold'),
-			strip.text.x = element_text(size=9, color='white'
-				# ,family="Source Code Pro Semibold"
+			strip.text.x = element_text(
+				size=9, color='white',family="Source Code Pro Semibold"
 				),
-			strip.text.y = element_text(size=9, color='white'
-				# ,family="Source Code Pro Semibold",
+			strip.text.y = element_text(
+				size=9, color='white',family="Source Code Pro Semibold",
 				,angle=0
 				),
 			strip.background = element_rect(fill = "#525252", color='#525252')
 			)
 	ggsave(g, height=3, width=8,
-		# file=paste0(graphicsPath, 'ameSimCover_',var,'.pdf')
-		file=paste0(graphicsPath, 'ameSimCover_',var,'_asa.pdf')
-		# , device=cairo_pdf
+		file=paste0(graphicsPath, 'ameSimCover_',var,'_asa.pdf'),
+		device=cairo_pdf
 		)
[…]

### 2021-01-04 revisionMemo2/revisionLetter.tex
start on rr2

--- 2020-01-17 revisionMemo/revisionLetter.tex
@@
 	{C:/Users/herme/Research/netsMatter/paper/graphics/},
-	{/Users/s7m/Research/netsMatter/paper/graphics/}
+	{/Users/s7m/Research/netsMatter/paper/graphics/},
+	{/Users/dorffc/ProjectsGit/netsMatter/paper/graphics}
 }
@@
 	{C:/Users/herme/Research/netsMatter/paper/graphics/},
-	{/Users/s7m/Research/netsMatter/paper/graphics/},
+	{/Users/s7m/Research/netsMatter/paper/graphics/},å
+	{/Users/dorffc/ProjectsGit/netsMatter/paper/graphics}
 }

### 2021-01-04 code/1_simulation/asa/1_ameSim_asaProbit.R
start on rr2

@@ -69,7 +69,7 @@ simRun = function(seed, n, mu, beta, gamma, alpha){
 
 ##############################
 # params
-imps = 100 ; cores = 20
+imps = 100 ; cores = 30
 
 #
 cl=makeCluster(cores) ; registerDoParallel(cl)

### 2021-01-04 paper/netsMatter.tex
syncing
move appendix to end
more efforts from CD and MG
sync up notes from meeting
quick footnote
edits with MG
cleanup
sync up sim code

--- 2019-07-12 paper/netsMatter.tex
@@
 % at some point figure out bolding ...
-\usepackage[default,osfigures,scale=0.95]{opensans}
+\usepackage[default,oldstyle,scale=0.95]{opensans}
 %\usepackage[sfdefault,scaled=.85]{FiraSans}
@@
 As a result, they suffer from biased coefficients and poorly calibrated standard errors.
-We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the inherent dependencies in dyadic data.
+We show how a regression-based approach, the Additive and Multiplicative Effects (AME) model, can be used to account for the inherent dependencies in dyadic data and glean substantive insights in the interrelations between actors.
 First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference on dyadic data.
@@
 \begin{center}
-	\textbf{Word count}: 9,177
+	\textbf{Word count}: 9,542
 \end{center}
@@
 \input{5_concl}
-\appendix
-\input{6_appendix}
 
@@
 
+\bibliography{C:/Users/Owner/whistle/master}
+% \bibliography{C:/Users/herme/whistle/master}
 %\bibliography{/Users/maxgallop/documents/whistle/master}
-\bibliography{/Users/s7m/whistle/master}
+%\bibliography{/Users/s7m/whistle/master}
 % \bibliography{/Users/mdw/Git/whistle/master}
-% \bibliography{/Users/cassydorff/ProjectsGit/master}
-% \bibliographystyle{authordate1}
-\bibliographystyle{sm}
+ % \bibliography{/Users/dorffc/ProjectsGit/whistle/master}
+\bibliographystyle{authordate1}
+% \bibliographystyle{sm}
 % \clearpage
@@
 
+\appendix
+\input{6_appendix}
+
 \end{document}

### 2021-01-04 revisionMemo/rev1.tex

@@
 	\item P1: ``Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
-In practice, such strategies rarely work because they do not directly address the fundamental data generating process.''
-	\begin{itemize}
-		\item The term “rarely work” is too casual.
+In practice, such strategies rarely work because they do not directly address the fundamental data generating process.'' The term “rarely work” is too casual.
 What problem exactly is motivating this paper?
@@
 were considering, but then this needs explication.
-	\end{itemize}
 	\begin{itemize}
@@
 		We thank the reviewer for taking the time to provide these valuable comments.
-We have clarified our language in the paper (on page 1) to address this question.
+We have clarified our language in the paper to address this question.
 Strategies like those employed by Aronow et al.
@@
 While such an approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will lead us to falsely accept the null).
-The main aim of our research is to directly account for the relevant dependency structures in the data, thus providing an accurate estimate and addressing both types of possible errors.
+The main aims of our research are to account for the relevant dependency structures in the data thereby enabling more accurate inference for covariates and to enable applied scholars using this framework to understand how to better understand the types of dependencies that exist in their data.
 		}}
@@
 	\item P2: “Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another.
-This makes sense if each observation is independent of the others.”
-	\begin{itemize}
-		\item Revise this opening, since it is perfectly fine to stack observations and then incorporate auxiliary information (such as adjacency matrix) to characterize dependencies between rows.
-	\end{itemize}
+This makes sense if each observation is independent of the others.” Revise this opening, since it is perfectly fine to stack observations and then incorporate auxiliary information (such as adjacency matrix) to characterize dependencies between rows.
 	\begin{itemize}
@@
 	\end{itemize}
-	\item P3: “Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.”
-	\begin{itemize}
-		\item This is another example of language that is too causal and imprecise, like the opening paragraph.
+	\item P3: “Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.” This is another example of language that is too causal and imprecise, like the opening paragraph.
 What are “parameter estimates from standard statistical models”?
@@
 Again, I think the authors have in mind “bias” with respect to parameters that are meant to have a particular structural interpretation, but this needs to be clarified.
-		\item Generally, section 2 should be clearer in motivating the problem.
+Generally, section 2 should be clearer in motivating the problem.
 Provide us with a toy example that illustrates the kind of “bias” that the paper has in mind from “standard models”.
 For me at least, after reading section 2 I still wasn’t sure what problem the paper is actually trying to address.
+	\begin{itemize}
 		\item \textcolor{blue}{ \emph{
-		We think the reviewer makes a good suggestion here to clarify, mathematically or in some other illustrative way, exactly how we think that dependencies bias the estimator.
-For example we can say that our target of inference is the estimation of beta (with some variance).
-If a coefficient is correlated with a dependency, then we could obtain an estimate for beta that is artificially high (and the same is true for the reverse).
+		We thank the reviewer for this suggestion. We have thoroughly revised the beginning of section two, so that the paper is clearer on the problem that it is seeking out.
 	}}
-		\item \textcolor{blue}{ \emph{ Additionally, we have added language motivating whatn is to come in the simulation section.
-	}}
-	\end{itemize}
 	\end{itemize}
@@
 Bai 2009, Ecta) and is based on an SVD factorization of the observed network.
-	\begin{itemize}
-		\item It is useful to think of the infeasible estimator and the implications for interpreting the LFM component.
+It is useful to think of the infeasible estimator and the implications for interpreting the LFM component.
 The infeasible estimator takes the transformation f(.), fixed effects parameters, and SRRM random effects as known, in which case the LFM amounts to predictions from a factor model fit to the adjacency matrix of residuals on the scale of the latent outcome.
@@
 If so it would be useful to draw the connection to factor models commonly in use these days (such as the interactive fixed effects model of Bai, which is the basis of, e.g., generalized synthetic control methods).
-	\end{itemize}
 	\begin{itemize}
 		\item \textcolor{blue}{ \emph{
-		We thank the reviewer for providing this reference and have added discussion of this alternative to the manuscript.
-		SM insert great response.
-		paper by Bai: https://sci-hub.tw/https://onlinelibrary.wiley.com/doi/10.3982/ECTA6135
-		connection between LFM and interactive fixed effects?
+		We thank the reviewer for providing this reference and have added discussion of Bai's work on interactive fixed effects to the manuscript.
+Within both Bai's work and the model we discuss here the goal is to find ways to use latent factor approaches to model heterogeneity.
+We'd also note that the recommender system literature from computer science also has its roots in using latent factor models to find low rank approximations.
+We had noted that already in the manuscript but have tried to make the link clearer in the revised ``Additive and Multiplicative Effect Models for Networks'' section.
 		}}
@@
 	\item P8: “accounting for this structure is necessary if we are to adequately represent the data generating process.” Continuing with the problems I have with the language above, what does “adequately represent” mean?.
-P9: “a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.”
-	\begin{itemize}
-		\item Okay here is where the paper are starting to clarify goals.
+P9: “a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance.” Okay here is where the paper are starting to clarify goals.
 So a) is about causal identification.
@@
 As such, from the perspective of what the researchers in discipline mostly cares about, a) and b) are of primary importance.
-	\end{itemize}
 	\begin{itemize}
@@
 How well does AME work out when the DGP does not abide by parametric assumptions?
-We address this in the section concerning the simulation and Aronow's approach.
+We have added an additional simulation section to the appendix that shows how well AME does when parametric assumptions of linearity are violated as discussed in Aronow et al.
+We also provide simulation evidence to this effect in the end of this section of the revision memo, when responding to another question from the reviewer.
 }}
 		\item \textcolor{blue}{ \emph{
-			Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others.}
-		}
+			Finally, we understand these goals are not always central to political science research, but we believe that prediction is both a valuable metric to evaluate model performance as well as a goal for scholars in and of itself, as in Gleditsch and Ward 2013; Grimmer 2015; Neunhoeffer and Sternberg 2019; Colaresi and Mahmood 2017; and Mueller and Rauh 2018 among others.
+}}
 	\end{itemize}
@@
 		\item \textcolor{blue}{ \emph{
-		We've added additional discussion about how each of the variables in our simulation were generated.
+		We thank the reviewer for pointing out that the simulation section was unclear.
+$X$ and $W$ are independent in the simulation that we present in the paper.
+We have thoroughly revised this section and our explication for how each of the variables were generated in the simulation study.
+Additionally, we have added in additional sections to the Appendix with simulation setups where $X$ and $W$ are related in alternative ways.
+These additonal sections come in response to requests from both Reviewer 1 and 2, and we provide more detail on them later in the memo.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		SM will deal with this.
-		\begin{figure}
+		We have added a simulation section to the Appendix in which we have incorporated an additional assessment of robustness to misspecification such as the one considered in Aronow et al.
+Specifically, Aronow et al note that ``fitting a linear approximation to mildly non-linear data'' can be problematic for dyadic random effect models.
+We agree on the importance of testing this in the context of our simulation.
+Aronow et al incorporate non-linear misspecification by adding in squared version of the key dyadic homophily variable to the DGP for their simulations: $Y_{i,j} = \beta_{0} + \beta_{1} X_{i,j} + \beta_{2} X_{i,j}^{2} + \epsilon_{i,j}$.
+Here $X$ is their representation of homophily and is considered observed, while $X^{2}$ is their check for robustness to misspecification and is unobserved.
+We incorporate the same type of misspecification into our probit simulation setup: $Z_{i,j} =  \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j}$, where $W$ now instead of being drawn from a separate normal distribution is just the squared version of $X$.
+For the simulation we set values of $\mu$, $\beta$, and $\gamma$ to -2, 1, and 1, respectively.
+Results are shown below in Figures~\ref{fig:ameBias_asa} and \ref{fig:ameCalib_asa} below.
+As in the paper, along the x-axis we implement a standard model that does not take any steps to account for dependencies, AME, and the Oracle model, and as in the paper we find that the AME approach provides a number of benefits in terms of bias and coverage over the standard approach of not dealing with dependencies.
+	}}
+
+	\begin{figure}[ht]
 		\centering
@@
 Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-			\label{fig:ameBias}
+		\label{fig:ameBias_asa}
 		\includegraphics[width=1\textwidth]{ameSimBias_all_asaProbit.pdf} \\
@@
 
-		\begin{figure}
-			\centering
-			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
-Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-			\label{fig:ameBias}
-			\includegraphics[width=1\textwidth]{ameSimBias_all_asa.pdf} \\
-		\end{figure}
-
-		\begin{figure}
+	\begin{figure}[ht]
 		\centering
 		\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
-			\label{fig:ameCalib}
+		\label{fig:ameCalib_asa}
 		\includegraphics[width=1\textwidth]{ameSimCover_all_asaProbit.pdf} \\
@@
 
-		\begin{figure}
-			\centering
-			\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
-			\label{fig:ameCalib}
-			\includegraphics[width=1\textwidth]{ameSimCover_all_asa.pdf} \\
-		\end{figure}
-		}}
 	\end{itemize}
 \end{enumerate}
+
+\FloatBarrier
+\clearpage

### 2021-01-04 revisionMemo/revisionLetter.tex

@@
 	{/Users/s7m/Research/netsMatter/paper/graphics/},
+	{/Users/dorffc/ProjectsGit/netsMatter/paper/graphics}
 }

### 2021-01-04 revisionMemo/rev2.tex

@@
 With dyadic data, however, we cannot use the common geographic adjacency matrices used for normal spatial econometrics, because in dyadic data a W matrix would need to specify how dyad i,j relates to dyad k,l (for every i,j,k,l).
-In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (see Plumper and Neumayer 2010 for a more in-depth discuss on spatial lags).
-\textbf{CD}: I added this as a footnote on p.9.
-I moved this footnote around a LOT but I think it needs to live somewhere in the simulation.
-It felt awkward within the text.
+In the simulation, we do not include a spatial lag model because this would require specifying a spatial dependence pattern (Plumper and Neumayer (2010) provide an in-depth discussion on varying operationalizations of spatial lags).
+If there is a particular type of spatial dependence that the reviewer would like us to simulate, however, we are happy to do so.
 			}}
-%		CDMG: Flag for discussion. We brainstormed how to fit this into the simulation, but wouldn't simulating a W matrix be a bit arbitrary here?
-%		Reference papers that have thought about what happens when we get the wrong W
-%		general gist ..
-spatial models of dyadic dvs are great if we know how to characterize the diffusion amtrices, but the point of our approach is that we may not know how to actually characterize the ways in which actors are related to one another on some diffusion space.
-%		reference franzesse et al mstar framework as a way to operationalize multiple pathways of diffusion in a spatial econometric approach
 	\end{itemize}
@@
 		\item \textcolor{blue}{ \emph{
-		add example of case where omitted variable is correlated
-		tech discussion bias-variance
+			Estimation certainly gets more challenging when the omitted variable is correlated with the observed variable.
+In the case of normal linear regression, we can show that in general a regression coefficient is unbiased for its ``unconditional effect'', which is the direct effect plus something related to the direct effects of any omitted variables that it is correlated with.
+For example, suppose in our simulation setup $X_{i,j}$ is correlated with $W_{i,j}$ via the relation $W_{i,j} = \alpha X_{i,j} + e_{i,j}$.
+Then: }}
+
+			\begin{align*}
+			  Y_{i,j} & = \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j} \\
+			  &= \mu + (\gamma \alpha+\beta ) X_{i,j} + (\gamma Z_{i,j} + \epsilon_{i,j})  \\
+			  &= \mu + \tilde \beta X_{i,j} + \tilde \epsilon_{i,j}
+			\end{align*}
+
+		\item \textcolor{blue}{ \emph{
+			The naive model is in some sense still ``correct'' as it is an unbiased estimator of the unconditional (direct plus indirect) effect $\gamma\alpha + \beta$ of $X$ on $Y$.
+However, there is no way to eliminate this extra bias without making further assumptions about the nature of the omitted variable.
+Additonally, in the binary case with the probit setup, other sources of bias get introduced as well because of the nonlinearity of the population moments as a function of the parameters.
+An interesting question is whether or not we can reduce the bias in the network setting by assuming the omitted variable $w_{i,j}$ is the product of latent node-specific factors.
+An idea related to this is discussed in \citet{minhas:etal:2017:arxiv}.}}
+
+		\item \textcolor{blue}{ \emph{
+			For now though to show the amount of bias that gets introduced into our setup when $X$ and $W$ are correlated we repeat our simulation exercise but induce varying levels of correlation between $X$ and $W$.
+Results with $X$ and $W$ correlated at 0.4 and then 0.7 are shown below, respectively.}}
 
@@
 			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
-Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-		\label{fig:ameBias}
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.
+$X$ and $W$ are correlated at 0.4 in this case.}
+			\label{fig:ameBias_corrMed}
 			\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitMed.pdf} \\
@@
 			\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
-Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-		\label{fig:ameBias}
+Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.
+$X$ and $W$ are correlated at 0.7 in this case.}
+			\label{fig:ameBias_corrHi}
 			\includegraphics[width=1\textwidth]{ameSimBias_all_corrProbitHi.pdf} \\
 			\end{figure}
+			\FloatBarrier
 
-		\begin{figure}[ht]
-		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
-Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-		\label{fig:ameBias}
-		\includegraphics[width=1\textwidth]{ameSimBias_all_corrMed.pdf} \\
-		\end{figure}
-
-		\begin{figure}[ht]
-		\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations.
-Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
-		\label{fig:ameBias}
-		\includegraphics[width=1\textwidth]{ameSimBias_all_corrHi.pdf} \\
-		\end{figure}
+		\item \textcolor{blue}{ \emph{
+			In terms of the bias-variance tradeoff discussion, this enters most prominently as a consideration for AME when choosing the dimensions of the latent factor space, specifically $K$ in the model.
+Everytime we add another dimension to the latent factor space the number of parameters in the model increases notably.
+For example, in a directed network with a choice of $K=1$ there are $2 \times n$ parameters introduced from the LFM portion of the model.
+With $K=2$ this number increases by another $2 \times n$.
+Increasing K certainly comes with a cost in terms of the number of extra parameters that are introduced into the model.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		We've added additional discussion about how each of the variables in our simulation were generated.
-To make the model more applicable to the types of networks we study in IR we have provided a probit example.
+		We thank the reviewer for this comment and have revised our simulation setup as a result.
+Most IR studies into networks have been interested in modeling binary outcomes (e.g., indicators for international agreements between countries or conflicts between countries), thus we we have revised our simulation to have a binary dependent variable.
+Key results remain the same.
+		}}
+		\item \textcolor{blue}{ \emph{
+		Additionally, we have thoroughly revised our description to make it clearer how each of the variables in the simulations were generated.
 		}}
@@
 		\item \textcolor{blue}{ \emph{
-		The reviewer raises a valid point.
-We have expanded our tutorial in Appendix C, particularly in giving more detail and concrete examples when it comes to formatting the data, which may represent the biggest hurdle for some users.
-We also...
+		The reviewer raises a valid point. We have made a number of revisions and extensions to the tutorial at the end of the Appendix so that applied scholars can better take advantage of AME.
 		}}

### 2021-01-04 paper/5_concl.tex

--- 2020-02-13 paper/5_concl.tex
@@
 This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} A broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}.
-\textcolor{blue}{Our findings bolster a growing recognition in the field of International Relations that interdependence plays an important role not just in terms of statistical estimations, but in the ways that scholars theorize about internationally relevant politics.
-Scholars have demonstrated the theoretical importance of interdependence through research on conflict \citep{dorff:etal:2020}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.}
+Our findings bolster a growing recognition in the field of International Relations that interdependence influences not only statistical estimations, but how scholars theorize about internationally relevant politics.
+Scholars have demonstrated the theoretical importance of interdependence through research on intrastate conflict \citep{dorff:etal:2020}, interstate bargaining \citep{gallop:2017}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.
 

### 2021-01-04 revisionMemo2/rev2.tex
revision memo for second round

\section*{Reviewer 2}

\begin{enumerate}
	\item I would have liked to see a justification for the choice of papers used in the replications. It is still unclear. Are these cases where the theory fail to characterize the direct and indirect effects, and implicitly assume that the latent dependencies are nuisances?
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		something, something
		}}
	\end{itemize}
	\item The presentation of the AME model clearly states that the Xs are exogenous; it is likely that they are not in replications. Still, the AME estimates (which in the simulations do not seem to outperform the standard method for estimating the betas when correlation between X and W is high) suggest that the models in the original papers are misspecified, and that these differences are likely driven by a better modeling of the DGP. This is a very important take-home point.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			something, something
			}}
	\end{itemize}
\end{enumerate}

### 2021-01-04 revisionMemo2/rev1.tex
revision memo for second round

\section*{Reviewer 1}

\begin{enumerate}
	\item In my previous review I pointed out some problems with vague language, mostly in the opening sections.   Honestly, I still find the introduction to be problematic. Maybe it could be cut and we just open with the part of section 2 where the paper begins to discuss processes that lead to dyadic and higher order dependencies?  The reason is that the intro tries to strike a very general tone.  But in reality, the paper proposes a specific estimation strategy for parameters in a specific parametric model of networked interactions.  I think the paper would be more useful to readers if it started with the specific applications, and then on the basis of that, motivated the statistical specification on substantive grounds.  There are some swipes at other literature are neither necessary nor clear enough to be very compelling.
	\begin{itemize}
		\item \emph{ \textcolor{blue}{
		something, something
		}}
	\end{itemize}
	\item ``Most of these studies use a generalized linear model (GLM). However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).'' It should really read, ``Most of these studies use a generalized linear model (GLM). However, this approach to studying dyadic data increases the chance of faulty inferences **if the estimation and inference assume** data are conditionally independent and identically distributed (iid).''  But then if this is all just about mistakenly assuming iid, just say it quickly and move on.
	\begin{itemize}
		\item  \emph{ \textcolor{blue}{
		something, something.
		}}
	\end{itemize}
	\item Why would the mean estimate be biased downward?  And, **biased with respect to what target of inference**?  A problem with the opening discussion is that it does not establish a target of inference.  Without doing so, "bias" is undefined.  It seems like they mean the coefficient for the linear relationship between some regressor and the latent outcome variable. But what if one is using a nonparametric model or a semiparametric model that operates on a different latent space---is bias even defined?  It seems that for bias to be defined we have to be restricting ourselves to a class of models that propose the same latent space, but that makes this much less general that the discussion leads on.  Moreover, that paper is not committing to talking about bias relative to causal effects under effect heterogeneity, but isn't that usually what people are after?  Or, maybe the paper is really concerned with bias with respect to a forecast prediction? If so, then we need to set up a forecast prediction error analysis.
	\begin{itemize}
		\item  \emph{ \textcolor{blue}{
		something, something.
		}}
	\end{itemize}
	\item Moving to section 2, the extended discussion of specifying a likelihood under iid is overkill.  Again, I would just move on to the discussion of the application. The point that iid is inappropriate will receive due attention in the course of developing the networked interaction model.
	\begin{itemize}
		\item  \emph{ \textcolor{blue}{
		something, something.
		}}
	\end{itemize}
	\item Explain how inference is done for the parameters from the AME model. How are intervals produced exactly? (Need this because you examine these in the simulation and make many statements about intervals throughout the text.)  What are the operating characteristics, in terms of convergence properties, etc.?
	\begin{itemize}
		\item  \emph{ \textcolor{blue}{
		something, something.
		}}
	\end{itemize}
	\item The DGP is simple enough that the paper needs to include the analytical calculation of the bias from the simple probit regression of Y on X.  We get no insight by just having monte carlo simulation results. For the simulations, clarify the "standard international relations approach" as just Y regressed on X, when you introduce it. Then explain how confidence intervals were produced. Are these just normal-approximation Wald intervals with non-robust standard errors? Explain exactly how the confidence intervals are produced for the AME model.  They were never characterized in the section that discussion the model.
	\begin{itemize}
		\item  \emph{ \textcolor{blue}{
		something, something.
		}}
	\end{itemize}		
\end{enumerate}

### 2021-01-04 revisionMemo2/revisionLetter.tex
revision memo for second round

--- 2020-01-17 revisionMemo/revisionLetter.tex
@@
 	{C:/Users/Owner/Research/netsMatter/paper/graphics/},
+	{C:/Users/S7M/Research/netsMatter/paper/graphics/},
 	{C:/Users/herme/Research/netsMatter/paper/graphics/},
-	{/Users/s7m/Research/netsMatter/paper/graphics/}
+	{/Users/s7m/Research/netsMatter/paper/graphics/},
+	{/Users/dorffc/ProjectsGit/netsMatter/paper/graphics}
 }
@@
 	{C:/Users/Owner/Research/netsMatter/paper/graphics/},
+	{C:/Users/S7M/Research/netsMatter/paper/graphics/},
 	{C:/Users/herme/Research/netsMatter/paper/graphics/},
-	{/Users/s7m/Research/netsMatter/paper/graphics/},
+	{/Users/s7m/Research/netsMatter/paper/graphics/},å
+	{/Users/dorffc/ProjectsGit/netsMatter/paper/graphics}
 }
@@
 
-\title[PSRM-OA-2019-0126]{Ms. No. PSRM-OA-2019-0126 entitled ``Taking Dyads Seriously''}
+\title[PSRM-OA-2019-0126.R1]{Ms. No. PSRM-OA-2019-0126.R1 entitled ``Taking Dyads Seriously''}
 
-\date{\today~~Version 0.01}
+\date{\today~~Version 0.02}
 \maketitle
@@
 
-Thank you for the opportunity to revise and resubmit our manuscript.
+Thank you for the opportunity to revise and resubmit our manuscript once more.
 We believe the manuscript has benefited from the Reviewers' helpful and thoughtful comments.

### 2021-01-06 paper/1_intro.tex
rev1 notes

@@
 
-The aim of this work is to address how to estimate the effect of exogenous covariates in a GLM context when there are network dependencies between observations in dyadic data.
+The aim of this work is to address how to estimate regression coefficients in a GLM context when there are network dependencies between observations in dyadic data.
 \citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).

### 2021-01-06 revisionMemo2/rev1.tex
rev1 notes

@@
 		\item \emph{ \textcolor{blue}{
-		something, something
+		We thank the reviewer for taking the time to provide their comments and helping us to make the manuscript more focused.
+We have made a number of changes to the framing of the article per the reviewer's suggestions.
 		}}
@@
 However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).'' It should really read, ``Most of these studies use a generalized linear model (GLM).
-However, this approach to studying dyadic data increases the chance of faulty inferences **if the estimation and inference assume** data are conditionally independent and identically distributed (iid).''  But then if this is all just about mistakenly assuming iid, just say it quickly and move on.
+However, this approach to studying dyadic data increases the chance of faulty inferences if the estimation and inference assume data are conditionally independent and identically distributed (iid).''  But then if this is all just about mistakenly assuming iid, just say it quickly and move on.
 	\begin{itemize}
@@
 	\item Why would the mean estimate be biased downward?
-And, **biased with respect to what target of inference**?
+And, biased with respect to what target of inference?
 A problem with the opening discussion is that it does not establish a target of inference.
-Without doing so, "bias" is undefined.
+Without doing so, ``bias'' is undefined.
 It seems like they mean the coefficient for the linear relationship between some regressor and the latent outcome variable.
@@
 		\item  \emph{ \textcolor{blue}{
-		something, something.
+		agree that in the case of nonparametric or semiparametric models ... but this is not what most
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		something, something.
+		speaking specifically about glms so ...
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		something, something.
+		intervals are produced via a gibbs sampler ...
 		}}
@@
 We get no insight by just having monte carlo simulation results.
-For the simulations, clarify the "standard international relations approach" as just Y regressed on X, when you introduce it.
+For the simulations, clarify the ``standard international relations approach'' as just Y regressed on X, when you introduce it.
 Then explain how confidence intervals were produced.
@@
 		\item  \emph{ \textcolor{blue}{
-		something, something.
+		we compare with a vanilla ols ...
+		and then the other set of results in the appendix re aranow use their dyad clustered approach ... per your request from last time
 		}}

### 2021-01-07 paper/4_app.tex
little change from cassy and max

@@
 First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data.
-In each of the replications, the AME's predictive performance out-of-sample is notably superior.
-Specifically, not only does AME perform better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves).
+In each of the models the AME substantially outperforms the original model out of sample.
+Not only does AME perform better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves).
+This may be because by ignoring dependencies, the original models are misspecifying the DGP, and the AME better accounts for it.
 

### 2021-01-07 revisionMemo2/rev2.tex
r2

@@
 		\item \textcolor{blue}{ \emph{
-		something, something
-		}}
+		In footnote 20 we state the primary criteria for our choice of papers used: we focus on studies explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science journal (for consistency in editorial standards and reviews, we focus on one journal, the American Political Science Review).
+In addition to these criteria, we searched for studies we understood to have hypotheses in which interdependencies are theoretically consequential due to the dyadic setting in which their theories operate (p.14) but where-in dependencies are not directly modeled (beyond the use of standard error corrections).
+We chose papers that, in spirit, seem to treat interdependence as technical problem rather than a threat to inference.}}
 	\end{itemize}
@@
 		\item \textcolor{blue}{ \emph{
-			something, something
+			Thank you.
+We hope that this take home point is made clear throughout the paper, but especially in sections 5.4 `lessons learned' and in our conclusion.
+We have added the following change to section 5.4 ``In each of the models the AME substantially outperforms the original model out of sample.
+This may be because by ignoring dependencies, the original models are misspecifying the DGP, and the AME better accounts for it."
 			}}

### 2021-01-08 revisionMemo2/rev1.tex
rev 1

@@
 		We thank the reviewer for taking the time to provide their comments and helping us to make the manuscript more focused.
-We have made a number of changes to the framing of the article per the reviewer's suggestions.
+We have made a number of changes to the framing of the article per the reviewer's suggestions, these appear throughout the manuscript, but especially on pages X.
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		something, something.
+		Thank you for the suggestion, we have made this change on page X.
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		speaking specifically about glms so ...
+		We understand the position that a detailed discussion of specifying a likelihood under iid seems laborious and potentially unnecessary.
+After a lengthy discussion among our co-authors, we have decided to retain this section of the paper.
+We believe that since our approach is directly aimed at applied scholarship in International Relations, and because we focus on applications that have appeared in top journals during the last 10 years, this section enables our study to serve a broad audience of readers.
+By keeping it, we ensure that a reader can directly follow our main points about GLM specifications and how the AME addresses these concerns.
 		}}

### 2021-01-08 revisionMemo2/rev1.tex
rev 1

@@
 After a lengthy discussion among our co-authors, we have decided to retain this section of the paper.
-We believe that since our approach is directly aimed at applied scholarship in International Relations, and because we focus on applications that have appeared in top journals during the last 10 years, this section enables our study to serve a broad audience of readers.
+We believe that since our approach is directly aimed at applied scholarship in International Relations, and because we focus on applications using GLM that have recently appeared in top journals, this section enables our study to serve a broad audience of readers.
 By keeping it, we ensure that a reader can directly follow our main points about GLM specifications and how the AME addresses these concerns.

### 2021-02-03 paper/3_sim.tex
push up memo revisions

@@
 \begin{itemize}
-	\item the ``standard'' international relations approach estimated through a typical generalized linear model;
+	\item the ``standard'' international relations approach estimated through a generalized linear model;\footnote{Specifically, here we are just regressing $Y$ on $X$ assuming independent errors.}
 	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);\footnote{Results with higher values of $K$ are similar.}
@@
 
-We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias} $\mu$ and $\beta$ estimates from the simulations for the three models.
+We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias}.
 The panels on the left show the results for when the number of actors is set to 50 and on the  right for 100 actors.
@@
 
-Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals.
+Next, we estimate the 95\% credible interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals.
 The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the standard approach performs poorly by comparison.

### 2021-02-03 paper/6_appendix.tex
push up memo revisions

@@
 
-Given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:
+Given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:\footnote{Further details on the full conditional distributions and estimation approach underlying AME can be found in \citet{hoff:2021}.}
 

### 2021-02-03 revisionMemo2/editor.tex
push up memo revisions

\section*{Editor}

\begin{enumerate}
	\item `effect'' is not precisely defined, so you might substitute ``how to estimate regression coefficients in a GLM context''
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
		We thank you for this suggestion and have changed the manuscript accordingly.}}
	\end{itemize}
	\item Also, my understanding is that modeling the ``underlying structural features in the data'' means including these ``structural features'' in the conditional mean equation of the GLM rather than including them exclusively in the variance equation (or ignoring them altogether). Or, more precisely, including SRRM and LFM parameters in the conditional mean equation. The advances you make come from this specific modeling choice. Augmenting the GLM in this particular way leads to 1) unbiased (or nearly unbiased) estimates of regression coefficients under homopholic dependence, when these coefficients exist and are known (the Monte Carlo simulations) and 2) superior out-of-sample forecast predictions for real data, in which case the regression model is just an approximation with unknown parameters (the replication exercises). Including LFM parameters in the conditional mean equation also makes it more likely that IR scholars will treat dyadic dependence as substantively important.
	\begin{itemize}
		\item \textcolor{blue}{ \emph{
			Thank you for these suggestions. We have made significant modifications to the introduction in line with the recommendations that you have posted and those from reviewer 1. We have removed references that reviewer 1 perceived as swipes at previous work. In addition, we have adopted your helpful language about how the AME approach essentially calls for including structural features in the conditional mean equation of the GLM. 
			}}
	\end{itemize}
\end{enumerate}

### 2021-02-03 revisionMemo2/rev1.tex
push up memo revisions

--- 2021-01-08 revisionMemo2/rev1.tex
@@
 		We thank the reviewer for taking the time to provide their comments and helping us to make the manuscript more focused.
-We have made a number of changes to the framing of the article per the reviewer's suggestions, these appear throughout the manuscript, but especially on pages X.
+We have made a number of changes to the framing of the article per the reviewer's suggestions.
+Additionally, we have significantly modified the introduction and removed what could be considered as, but certainly not intended, swipes at other literatures.
+The editor also suggested a number of changes to be made in the introduction and we have adopted those as well.
+We disagree with the choice of removing the introduction completely and starting with section 2 because we feel some relatively non-technical context will be useful to more general readers.
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		Thank you for the suggestion, we have made this change on page X.
+		We thank R1 for this suggestion, per R1's first comment, we have actually removed this from the introduction and clarified the goal of the paper along the lines of what R1 and the editor suggested.
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		agree that in the case of nonparametric or semiparametric models ... but this is not what most
+		These are fair questions especially in terms of the complications that arise when using nonparametric or semiparametric models.
+Instead of trying to resolve all this in the introduction, we have removed the portion of the introduction that led the reviewer to ask this question.
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		intervals are produced via a gibbs sampler ...
+		Bayesian inference for the parameters from the AME model proceed via Gibbs sampling algorithms where we iteratively simulate values of the model parameters from their full conditional distributions.
+We have added additional discussion about how the parameters are estimated to both the simulation section and the applications.
+In Appendix A we provide a brief outline of the MCMC algorithm used.
+We do not re-derive the full conditional distributions for the AME models in this paper as it has been done in previous work that we have cited: Hoff 2005, 2008; Minhas et al.
+2019.
+Additionally, Hoff (2021), which we now cite in the paper, more fully details the history of the AME model and how it has evolved since its original formulation in Hoff et al.
+(2002).
 		}}
@@
 		\item  \emph{ \textcolor{blue}{
-		we compare with a vanilla ols ...
-		and then the other set of results in the appendix re aranow use their dyad clustered approach ... per your request from last time
+		Given the various forms of network dependence that can arise it's not clear to us how to actually construct an analytical calculation like what R1 is suggesting.
+If the purpose of this analytical calculation would be to highlight the consequences that dyadic dependence would have in the context of inference, then that is exactly why we open Section 2 in the paper with a discussion of the assumptiong underlying GLMs.
 }}
+		\item  \emph{ \textcolor{blue}{
+		In terms of the simulation section, we thank R1 for pointing out the parts that are still unclear and have added in their suggested language with respect to clarifying the standard IR approach.
+		}}
+		\item  \emph{ \textcolor{blue}{
+		We thank R1 for asking a question about how confidence intervals are produced because that was an error on our part as we meant to say credible interval.
+Within our framework These are generated directly from the posterior distribution generated from the three models.
+Specifically, for every simulation and model we calculate the 95\% credible interval and then across all simulations we estimate the proportion of times that the true value fell within those intervals.
+We have clarified this in the simulation section of the manuscript.}}
 	\end{itemize}

### 2021-02-03 paper/2_model.tex
push up memo revisions

@@
 
-When modeling dyadic data scholars typically employ a generalized linear model (GLM) estimated via maximum-likelihood.
+When modeling dyadic data scholars typically employ a GLM estimated via maximum-likelihood.
 This type of model is expressed via a stochastic and systematic component \citep{pawitan:2013}.
@@
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.
-\footnote{Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.
-The algorithm is provided in the Appendix.}
+Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.\footnote{Additional details on the sampling procedure and the full conditional distributions can be found in: \citet{hoff:2005,hoff:2008,minhas:etal:2019,hoff:2021}.}
 

### 2021-02-03 paper/1_intro.tex
push up memo revisions

@@
 
-The aim of this work is to address how to estimate regression coefficients in a GLM context when there are network dependencies between observations in dyadic data.
-\citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).
-However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
-Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
-In practice, such strategies remain insufficient because they do not directly model underlying structural features in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data.''} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).} The aims of our study are twofold.
-First, we provide a framework that accounts for dependencies often present in dyadic data.\footnote{This is important to consider since the inferential problems caused by non-iid observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.} Second, the framework we discuss here enables applied scholars to understand the structure underlying dyadic data rather than simply controlling for them.
+The aim of this work is to address how to estimate regression coefficients in a generalized linear model (GLM) context when there are network dependencies between observations in dyadic data.
+Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019}.
+The AME works by including a set of parameters meant to capture network effects in the conditional mean equation of the GLM.
 
-Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019}.
-We focus on three types of interdependencies that can complicate dyadic analyses.
+We focus on three types of network effects that can complicate dyadic analyses.
 First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still essential to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict against a government, the government will likely reciprocate that behavior.
@@
 We begin with a discussion of these dependencies and an introduction to the AME model.
-Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data.
+Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network based dependence.
 Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature.
-Our comparison reveals that by accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables.
-Consequently, AME generates results that, at times, differ from those found in the original study as well as from the broader literature.
-Moreover, we demonstrate that the AME latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data widely employed in IR.
+Our comparison reveals that by accounting for observational dependence, leads to results that, at times, differ from those found in the original study as well as from the broader literature.
+Moreover, we demonstrate that the additional parameters included by AME in the conditional mean equation of a typical GLM can offer substantive insights that are often occluded by ignoring the interdependencies found in relational data.
 Finally, we show that for each replication our network-based approach provides substantially more accurate out-of-sample predictions than the models used in the original studies.
@@
 Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.
+
+% \citet{aronow:etal:2015} estimate that during the period 2010 to 2015, over sixty articles were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization} using dyadic data.\footnote{In 2017, \textit{International Studies Quarterly} published a special issue on Dyadic Research Designs along with an online symposium to discuss the papers.} Most of these studies use a generalized linear model (GLM).
+However, this approach to studying dyadic data increases the chance of faulty inferences by assuming data are conditionally independent and identically distributed (iid).
+Most standard approaches assume that the problems raised by non-iid relational data can be addressed by recalculating the standard errors of estimated parameters to reflect the potential clustering of cases.
+In practice, such strategies remain insufficient because they do not directly model underlying structural features in the data\footnote{As \citet[p.15]{aronow:etal:2015} explain, ``Of course, accounting for dynamic dyadic clustering may fail to fully account for all relevant dependencies in the data.''} but instead adjust the variance estimator.\footnote{While this approach decreases the risk of a type 1 error, it increases risk of type 2 error (if the mean estimate is biased downwards, increasing the variance estimate will cause us to falsely accept the null).} The aims of our study are twofold.
+First, we provide a framework that accounts for dependencies often present in dyadic data.\footnote{This is important to consider since the inferential problems caused by non-iid observations affect more than just diagonals of the variance co-variance matrix \citep{beck:2012,franzese:hays:2007,king:roberts:2014}.} Second, the framework we discuss here enables applied scholars to understand the structure underlying dyadic data rather than simply controlling for them.

### 2021-02-04 paper/1_intro.tex

@@
 
-The aim of this work is to address how to estimate regression coefficients in a generalized linear model (GLM) context when there are network dependencies between observations in dyadic data.
+The aim of this study is to address how to estimate regression coefficients in a generalized linear model (GLM) context when there are network dependencies in dyadic data.
 Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019}.

### 2021-02-04 revisionMemo2/rev1.tex

@@
 We have made a number of changes to the framing of the article per the reviewer's suggestions.
-Additionally, we have significantly modified the introduction and removed what could be considered as, but certainly not intended, swipes at other literatures.
+Additionally, we have significantly modified the introduction and removed what could be considered as, but certainly not intended, swipes at other research.
 The editor also suggested a number of changes to be made in the introduction and we have adopted those as well.
@@
 		Given the various forms of network dependence that can arise it's not clear to us how to actually construct an analytical calculation like what R1 is suggesting.
-If the purpose of this analytical calculation would be to highlight the consequences that dyadic dependence would have in the context of inference, then that is exactly why we open Section 2 in the paper with a discussion of the assumptiong underlying GLMs.
+If the purpose of this analytical calculation would be to highlight the consequences that dyadic dependence would have in the context of inference, then that is exactly why we open Section 2 in the paper with a discussion of the assumption underlying GLMs.
 }}
@@
 		We thank R1 for asking a question about how confidence intervals are produced because that was an error on our part as we meant to say credible interval.
-Within our framework These are generated directly from the posterior distribution generated from the three models.
+Within our framework, these are generated directly from the posterior distribution generated from the three models.
 Specifically, for every simulation and model we calculate the 95\% credible interval and then across all simulations we estimate the proportion of times that the true value fell within those intervals.

### 2021-02-04 revisionMemo2/rev2.tex

--- 2021-01-04 revisionMemo2/rev2.tex
@@
 		\item \textcolor{blue}{ \emph{
-		something, something
-		}}
+		Thank you for raising this concern.
+In footnote 20 we state the primary criteria for our choice of papers used: we focus on studies explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science journal (for consistency in editorial standards and reviews, we focus on one journal, the American Political Science Review).
+In addition to these criteria, we searched for studies we understood to have hypotheses in which interdependencies are theoretically consequential due to the dyadic setting in which their theories operate (p.14) but where-in dependencies are not directly modeled (beyond the use of standard error corrections).
+We chose papers that, in spirit, seem to treat interdependence as technical problem rather than a threat to inference within the GLM context.
+In this round of revisions, we note clearly our focus on the GLM context starting with the first sentence of the MS.}}
 	\end{itemize}
@@
 		\item \textcolor{blue}{ \emph{
-			something, something
+			Thank you for highlighting this point.
+We hope that this take home point is now made clear throughout the paper, but especially in sections 5.4 `lessons learned' and in our conclusion.
+We have also added the following change to section 5.4 ``In each of the models the AME substantially outperforms the original model out of sample.
+This may be because by ignoring dependencies, the original models are misspecifying the DGP, and the AME better accounts for it."
 			}}

### 2021-02-04 revisionMemo2/revisionLetter.tex

@@
 We believe the manuscript has benefited from the Reviewers' helpful and thoughtful comments.
-We have revised the manuscript, taking seriously each individual point raised by the Reviewers.
-Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each of the points made by the Reviewers.
+We have revised the manuscript, taking seriously the points raised by you and the Reviewers.
+Our comments and responses are shown in \textcolor{blue}{\emph{BLUE}} below each of the points from your letter.
 
@@
 
+\include{editor}
 \include{rev1}

### 2021-02-04 paper/1_intro.tex
revisions

@@
 Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature.
-Our comparison reveals that by accounting for observational dependence, leads to results that, at times, differ from those found in the original study as well as from the broader literature.
+Our comparison reveals that accounting for observational dependence leads to results that, at times, differ from those found in the original study as well as from the broader literature.
 Moreover, we demonstrate that the additional parameters included by AME in the conditional mean equation of a typical GLM can offer substantive insights that are often occluded by ignoring the interdependencies found in relational data.
@@
 Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena without abandoning a regression based approach, while at the same time accounting for the data generating processes behind such events of interest.
-Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.
+Perhaps most importantly, the AME approach concentrates on the relational aspect of international relations through a statistical framework that is familiar to most scholars.
 

### 2021-02-04 revisionMemo2/rev2.tex
revisions

@@
 In footnote 20 we state the primary criteria for our choice of papers used: we focus on studies explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science journal (for consistency in editorial standards and reviews, we focus on one journal, the American Political Science Review).
-In addition to these criteria, we searched for studies we understood to have hypotheses in which interdependencies are theoretically consequential due to the dyadic setting in which their theories operate (p.14) but where-in dependencies are not directly modeled (beyond the use of standard error corrections).
+In addition to these criteria, we searched for studies we understood to have hypotheses in which interdependencies are theoretically consequential due to the dyadic setting in which their theories operate (p.14) but wherein dependencies are not directly modeled (beyond the use of standard error corrections).
 We chose papers that, in spirit, seem to treat interdependence as technical problem rather than a threat to inference within the GLM context.

### 2021-02-04 paper/4_app.tex
revisions

@@
 We focus on studies that are explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we focus on one journal (American Political Science Review).
-We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.} Each of these pieces is well-known and posited an hypothesis in which interdependencies are consequential.
+We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.} Each of these pieces is well-known and posited a hypothesis in which interdependencies are consequential.
 Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model.
@@
 Results with alternative values of $K$ are similar.
-We recommend that authors try out a range of $K$s} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
+We recommend that authors try out a range of $K$s.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version.
 In the directed formulation, separate random effects are used for senders and receivers in both the additive and multiplicative portions of the model.
@@
 Gibler argues that the long-standing relationship between the relative parity of capabilities and initiation of militarized interstate disputes (MIDs) is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members.
-In most specifications, when you take into account the initial conditions for dyadic entry, the statistical significance of power parity vanishes.
+In most specifications, after taking into account the initial conditions for dyadic entry, the statistical significance of power parity vanishes.
 This finding calls into question many IR theories about the role of balance in generating international conflict \citep{organski:1958}.
@@
 To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for both the insignificance of contemporary power parity, and the effect of the initial conditions for entry on driving conflict behavior.
-When we reestimate using AME power parity still has a small, imprecisely measured effect, supporting Gibler's argument that contemporary power parity is an artifact of other aspects of state relations, and does not drive contemporary conflict behavior.
+When we re-estimate using AME power parity still has a small, imprecisely measured effect, supporting Gibler's argument that contemporary power parity is an artifact of other aspects of state relations, and does not drive contemporary conflict behavior.
 At the same time, the AME approach finds that the effect of power parity when a country enters into the international system is too imprecisely measured to draw any inferences about its effect on the initiation of MIDs.

### 2021-02-04 revisionMemo2/editor.tex
revisions

@@
 \begin{enumerate}
-	\item `effect'' is not precisely defined, so you might substitute ``how to estimate regression coefficients in a GLM context''
+	\item `effect' is not precisely defined, so you might substitute ``how to estimate regression coefficients in a GLM context''
 	\begin{itemize}
 		\item \textcolor{blue}{ \emph{
-		We thank you for this suggestion and have changed the manuscript accordingly.}}
+		We thank you for this suggestion and have changed the manuscript accordingly, particularly in the first line of the manuscript.}}
 	\end{itemize}

### 2021-02-04 paper/1_intro.tex
small intro edit

@@
 The aim of this study is to address how to estimate regression coefficients in a generalized linear model (GLM) context when there are network dependencies in dyadic data.
-Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019}.
+Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019} in International Relations data.
 The AME works by including a set of parameters meant to capture network effects in the conditional mean equation of the GLM.

### 2021-02-05 paper/4_app.tex
minor edits

@@
 
-For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward to do because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimate the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
+For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimate the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$.
 Results with alternative values of $K$ are similar.
@@
 As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest.
-To uncover this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model -- $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$.
+To uncover this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$.
 These random effects account for higher order dependencies that manifest as a result of homophily and stochastic equivalence in dyadic data.
@@
 As Figure~\ref{fig:gibmargeff} illustrates, the substantive results differ notably.
-First, the expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.
+The expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.
 These are rare events, so the probabilities are low, but the difference is a factor of almost two.

### 2021-02-05 paper/2_model.tex
minor edits

@@
 
-When modeling dyadic data scholars typically employ a GLM estimated via maximum-likelihood.
+When modeling dyadic data, scholars typically employ a GLM estimated via maximum-likelihood.
 This type of model is expressed via a stochastic and systematic component \citep{pawitan:2013}.
@@
 Without the assumption of conditional observational independence the joint density function cannot be written in the way described above and a valid likelihood does not exist.
-Accordingly, inferences drawn from misspecified models that ignore potential interdependencies between dyadic observations are likely to have a number of issues including biased estimates of the effect of independent variables, uncalibrated confidence intervals, and poor predictive performance.
+Accordingly, inferences drawn from misspecified models that ignore potential interdependencies between dyadic observations are likely to have a number of issues, including biased estimates of the effect of independent variables, uncalibrated confidence intervals, and poor predictive performance.
 The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.

### 2021-02-05 revisionMemo2/editor.tex
minor edits

@@
 		\item \textcolor{blue}{ \emph{
-		We thank you for this suggestion and have changed the manuscript accordingly, particularly in the first line of the manuscript.}}
+		We thank you for this suggestion and have changed the manuscript accordingly, particularly the first line of the manuscript.}}
 	\end{itemize}

### 2021-02-09 revisionMemo2/editor.tex
minor rev memo edits

@@
 			Thank you for these suggestions.
-We have made significant modifications to the introduction in line with the recommendations that you have posted and those from reviewer 1.
-We have removed references that reviewer 1 perceived as swipes at previous work.
+We have made significant modifications to the introduction in line with the recommendations that you have posted and those from Reviwer 1 (R1).
+We have also removed references that R1 perceived as swipes at previous work.
 In addition, we have adopted your helpful language about how the AME approach essentially calls for including structural features in the conditional mean equation of the GLM.

### 2021-02-09 revisionMemo2/rev1.tex
minor rev memo edits

@@
 After a lengthy discussion among our co-authors, we have decided to retain this section of the paper.
-We believe that since our approach is directly aimed at applied scholarship in International Relations, and because we focus on applications using GLM that have recently appeared in top journals, this section enables our study to serve a broad audience of readers.
+We believe that since our approach is directly aimed at applied scholarship in International Relations (IR), and because we focus on applications using GLM that have recently appeared in top journals, this section enables our study to serve a broad audience of readers.
 By keeping it, we ensure that a reader can directly follow our main points about GLM specifications and how the AME addresses these concerns.

### 2021-02-09 paper/2_model.tex
fix refs in manu

@@
 In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data.
-Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.\footnote{Additional details on the sampling procedure and the full conditional distributions can be found in: \citet{hoff:2005,hoff:2008,minhas:etal:2019,hoff:2021}.}
+Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.\footnote{Details on the sampling procedure and the full conditional distributions can be found in: \citet{hoff:2005,hoff:2008,minhas:etal:2019,hoff:2021}.}
 

### 2021-03-13 replArchive/helpers/coefHelpers.R
most repls done

@@ -1,0 +1,15 @@
+getAmeCoef = function(fit, addTypeLabel=FALSE){
+	tmp<-cbind(
+		apply(fit$BETA,2,mean), apply(fit$BETA,2,sd),
+		apply(fit$BETA,2,mean)/apply(fit$BETA,2,sd),
+		2*(1-pnorm( abs(apply(fit$BETA,2,mean)/apply(fit$BETA,2,sd)))),
+		t( apply(fit$BETA, 2, function(x){ quantile(x, c(0.025,0.5,0.95,0.975))  }) )
+		)
+	colnames(tmp)<-c("pmean","psd","z-stat","p-val", 'lo95','lo90','hi90','hi95') 
+	if(!addTypeLabel){
+		rownames(tmp) = gsub('.dyad', '', rownames(tmp))
+		rownames(tmp) = gsub('.row', '', rownames(tmp))
+	}
+	rownames(tmp)[rownames(tmp)=='intercept'] = '(Intercept)'
+	return(tmp)
+}

### 2021-03-13 replArchive/1_simulation/appendixC/appendix_figureC1.R
most repls done

@@ -1,0 +1,116 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+graphicsPath=simResPath = pth
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+
+# source in helper functions
+source(paste0(pth, '/helpers/functions.R'))
+
+# load in additional pkgs
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+	'latex2exp', 'extrafont', 'Cairo'
+	)
+loadPkg(toLoad)
+suppressMessages(loadfonts(device='win'))
+facet_labeller = function(string){ TeX(string) }
+##############################
+
+##############################
+# params
+intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# iterate over sims
+
+# sim labs
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,'_asaProbit.rda')) }
+
+#
+modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
+modKey$clean = c('Standard', 'AME', 'Oracle')
+##############################
+
+##############################
+# check bias of x1
+getBiasDF = function(ameSim){
+	biasDF = lapply(ameSim, function(x){
+		betaMod=lapply(x$beta, function(z){ apply(z, 2, median)[1:2] })
+		betaMod = betaMod %>% reshape2::melt() %>%
+			dplyr::mutate(
+				var=rep(c('Intercept','X1'), length(x$beta)),
+				act = rep(c(intEff, x1Eff), length(x$beta)),
+				bias = value - act ) %>%
+			dplyr::rename(model = L1)
+		return(betaMod) })
+	return( suppressMessages( reshape2::melt(biasDF,id=names(biasDF[[1]])) ) ) }
+
+#
+ameSimBias = rbind(
+	cbind(getBiasDF(ameSim50), n=50),
+	cbind(getBiasDF(ameSim100), n=100) )
+##############################
+
+##############################
+# clean
+ameSimBias = ameSimBias %>% group_by(model, var, n) %>%
+	dplyr::mutate( mse = mean(bias^2) ) %>% data.frame()
+
+# fix mod labels
+ameSimBias$model = modKey$clean[match(ameSimBias$model, modKey$dirty)]
[…]

### 2021-03-13 replArchive/1_simulation/1_simManuscript.R
most repls done

@@ -1,0 +1,75 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+simResPath = pth
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+
+# source in helper functions
+source(paste0(pth, '/helpers/functions.R'))
+
+# load in additional pkgs
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr' )
+loadPkg(toLoad)
+##############################
+
+##############################
+### Homophily
+simRun = function(seed, n, mu, beta, gamma){
+	set.seed(seed)
+	xw = matrix(rnorm(n*2),n,2)
+	X = tcrossprod(xw[,1])  ; W = tcrossprod(xw[,2])
+	XW = array(dim=c(n,n,2)) ; XW[,,1] = X ; XW[,,2] = W
+
+	# create DV
+	Y = 1*(mu + beta*X + gamma*W + matrix(rnorm(n*n),n,n) >0)
+
+	# run AME models
+	fit0 = ame(Y,X,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
+		print=FALSE,plot=FALSE)
+	fit1 = ame(Y,X,R=1,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
+		print=FALSE,plot=FALSE)
+	fitO = ame(Y,XW,R=0,rvar=FALSE,cvar=FALSE,dcor=FALSE,model='bin',
+		print=FALSE,plot=FALSE)
+
+	# gather together results
+	beta = list(naive=fit0$BETA, ame=fit1$BETA, oracle=fitO$BETA)
+	uv = list(naive=fit0$UVPM, ame=fit1$UVPM, oracle=fitO$UVPM)
+	u = list(naive=fit0$U, ame=fit1$U, oracle=fitO$U)
+	v = list(naive=fit0$V, ame=fit1$V, oracle=fitO$V)
+	gof = list(naive=fit0$GOF, ame=fit1$GOF, oracle=fitO$GOF)
+	out = list(beta=beta, uv=uv, u=u, v=v, W=W, gof=gof)
+	return(out) }
+##############################
+
+##############################
+# params
+startSimTime = Sys.time()
+imps = 1000
[…]

### 2021-03-13 replArchive/2_applications/gibler_crossval.R
most repls done

@@ -1,0 +1,123 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+fPth = paste0(pth, 'helpers/')
+gPth = paste0(pth, '2_applications/application_data/gibler/')
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+
+# some necessary libs
+source(paste0(fPth, 'loadPkg.R'))
+loadPkg(c(
+  'reshape2', # data management
+  'doParallel', 'foreach' # parallel
+  ))
+
+# misc
+char = function(x){ as.character(x) }
+num = function(x){ as.numeric(char(x)) }
+trim = function (x) { gsub("^\\s+|\\s+$", "", x) }
+
+#
+loadPkg(c('ROCR', 'RColorBrewer', 'caTools'))
+source(paste0(fPth, 'binPerfHelpers.R'))
+#############################
+
+#############################
+# load data
+load(paste0(gPth, 'giblerData.rda'))
+load(paste0(gPth, 'ameFitGibler.rda'))
+#############################
+
+#############################
+# function to run k-fold cross validation analysis using ame
+ameOutSamp = function(
+  yList, xDyadL=NULL, xRowL=NULL, xColL=NULL, startVals,
+  seed=6886, folds=4, cores=folds,
+  R=2, model='bin', burn=1000, nscan=2000, odens=10,
+  intercept=TRUE, rvar=TRUE, cvar=TRUE, symmetric=TRUE
+){
+
+  ################
+  # divide dataset into folds
+  set.seed(seed)
+  yListFolds = lapply(yList, function(y){ # assign each y-matrix to a fold
+    yFold=matrix(sample(1:folds, length(y), replace=TRUE),
+                 nrow=nrow(y),ncol=ncol(y), dimnames=dimnames(y))
+    diag(yFold) = NA
+    return(yFold) })
+  ################
+
+  ################
+  # run models by fold
+  yCrossValTrain = lapply(1:folds, function(f){
+    yListMiss = lapply(1:length(yList), function(t){
+      foldID = yListFolds[[t]] ; y = yList[[t]] # find all values for given fold
+      foldID[foldID==f]=NA ;
[…]

### 2021-03-13 replArchive/2_applications/gibler_posterior.R
most repls done

@@ -1,0 +1,144 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+gPth = paste0(pth, '2_applications/application_data/gibler/')
+
+#
+source(paste0(pth, 'helpers/clusteredSE.R'))
+source(paste0(pth, 'helpers/functions.R'))
+source(paste0(pth, 'helpers/ameHelpers.R'))
+source(paste0(pth, 'helpers/binPerfHelpers.R'))
+source(paste0(pth, 'helpers/stargazerHelpers.R'))
+##############################
+
+##############################
+# load models
+load(paste0(gPth,'ameFitGibler.rda'))
+load( paste0(gPth,'glmFitGibler.rda') )
+##############################
+
+##############################
+ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
+	'Pr(>|z|)'=2*(1-pnorm( abs(mean(x)/sd(x)))) )}))
+rownames(ameSumm) = gsub('.row','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.col','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('.dyad','',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('_s','s',rownames(ameSumm),fixed=TRUE)
+rownames(ameSumm) = gsub('intercept','(Intercept)',rownames(ameSumm),fixed=TRUE)
+
+glmSumm = modSumm[,]
+
+dat = cbind(y=mod$y, mod$data[,names(coef(mod)[-1])], dyad=mod$data$dyad)
+form = paste0('y~',paste(names(coef(mod)[-1]), collapse='+'))
+probMod = glm( form, data=dat, family=binomial(link='probit') )
+loadPkg('lmtest')
+probSumm = coeftest(probMod, vcov = vcovCluster(mod, cluster = dat$dyad))[,]
+
+modList = list(glmSumm, ameSumm)
+modNames = c('GLM (Logit)', 'AME')
+varKey = data.frame(
+	dirty=names(coef(mod)),
+	clean=c( '(Intercept)',
+		'Allied', 'Joint Democracy',
+		'Peace Years', 'Spline 1', 'Spline 2',
+		'Spline 3', 'Contiguity', 'Parity',
[…]

### 2021-03-13 replArchive/1_simulation/appendixD/appendix_figureD1.R
most repls done

@@ -1,0 +1,127 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+graphicsPath=simResPath = pth
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+
+# source in helper functions
+source(paste0(pth, '/helpers/functions.R'))
+
+# load in additional pkgs
+toLoad = c(
+	'devtools',
+	'foreach', 'doParallel',
+	'magrittr', 'dplyr', 'ggplot2',
+	'latex2exp', 'extrafont', 'Cairo'
+	)
+loadPkg(toLoad)
+suppressMessages(loadfonts(device='win'))
+facet_labeller = function(string){ TeX(string) }
+##############################
+
+##############################
+# params
+NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# iterate over correlation levels
+figNames = c('a', 'b')
+types = c('_corrProbit')
+levs = c('Med', 'Hi')
+corrLevs = expand.grid(types, levs) %>% apply(.,1,paste,collapse='')
+
+# iterate over corrSims
+for(ii in 1:length(corrLevs)){
+
+#
+corrLev = corrLevs[ii]
+
+# load sim data
+for(n in c( 50,100)){ load(paste0(simResPath,'ameSim',n,corrLev,'.rda')) }
+
+#
+modKey = data.frame(dirty=names(ameSim50[[1]]$beta))
+modKey$clean = c('Standard', 'AME', 'Oracle')
+##############################
+
+##############################
+# check bias of x1
+getBiasDF = function(ameSim){
+	biasDF = lapply(ameSim, function(x){
+		betaMod=lapply(x$beta, function(z){ apply(z, 2, median)[1:2] })
+		betaMod = betaMod %>% reshape2::melt() %>%
+			dplyr::mutate(
+				var=rep(c('Intercept','X1'), length(x$beta)),
+				act = rep(c(intEff, x1Eff), length(x$beta)),
+				bias = value - act ) %>%
+			dplyr::rename(model = L1)
+		return(betaMod) })
+	return( suppressMessages( reshape2::melt(biasDF,id=names(biasDF[[1]])) ) ) }
+
+#
+ameSimBias = rbind(
+	cbind(getBiasDF(ameSim50), n=50),
+	cbind(getBiasDF(ameSim100), n=100) )
[…]

### 2021-03-13 replArchive/2_applications/figure7.R

@@ -1,0 +1,77 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+rsPth = paste0(pth, '2_applications/application_data/reiter_stam/')
+
+#
+source(paste0(pth, 'helpers/functions.R'))
+source(paste0(pth, 'helpers/ameHelpers.R'))
+
+#
+library(amen)
+##############################
+
+##############################
+# load models
+load(paste0(rsPth,'ameFitReiterStam.rda'))
+##############################
+
+# addEffdata ###########################################
+# sender effects
+effdat = getAddEffData(fit = ameFit) ##This function is in helperEx.R
+effdat$cown = effdat$actor
+effdat$actor = countrycode::countrycode(effdat$actor, 'cown', 'country.name')
+effdat$actor = factor(effdat$actor,  levels=effdat[order(effdat$addEff),'actor'])
+
+## spatial visualization
+# fix actor names
+cntryKey = data.frame(code=effdat$cown, cname=char(effdat$actor),stringsAsFactors = FALSE)
+cntryKey$cname[cntryKey$code==731]='North Korea'
+cntryKey$cowc = countrycode(cntryKey$cname, 'country.name', 'cowc')
+
+# geo colors for nodes
+loadPkg('cshapes')
+cmap = cshp(date=as.Date('2016-1-1'))
+cmapDF=fortify(cmap,region='FEATUREID') ; names(cmapDF)[6]='FEATUREID' ; cmapDF=join(cmapDF, cmap@data)
+cmapDF$addEff = effdat$addEff[match(cmapDF$COWCODE, effdat$cown)]
+
+ggMap = ggplot() +
+  geom_polygon(data=cmapDF, aes(x=long, y=lat,group=group,fill=addEff),color='grey30',size=.05) +
+  scale_fill_gradient2() +
+  coord_equal() + xlab('') + ylab('') +
+  labs(
+    x='', y='', fill='Sender Effects'
+    ) +
+  theme(
+    legend.position = 'bottom',
+    legend.key.width = unit(1.25,"cm"),
+    panel.border = element_blank(), panel.grid=element_blank(),
+    axis.ticks = element_blank(), axis.line=element_blank(),
+    axis.text = element_blank() )
+ggsave(ggMap, file=paste0(pth, 'figure7a.pdf'), width=6, height=3)
+system(paste('pdfcrop',
[…]

### 2021-03-13 replArchive/2_applications/reiter_stam_posterior.R

@@ -1,0 +1,250 @@
+# setup ###########################################
+rm(list=ls())
+
+if(Sys.info()['user'] %in% c('s7m', 'janus829')){
+	resultsPath = '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/output/'
+	inputPath= '~/Dropbox/Research/netsMatter/replications/Reiter_Stam_2003/input/'
+	plotPath = '~/Research/netsMatter/paper/graphics/'
+  source('~/Research/netsMatter/code/helpers/functions.R')
+  source('~/Research/netsMatter/code/helpers/ameHelpers.R')
+  source('~/Research/netsMatter/code/helpers/binPerfHelpers.R')
+  source('~/Research/netsMatter/code/helpers/stargazerHelpers.R') }
+
+if(Sys.info()['user'] == 'juanftellez'){
+  resultsPath = '/Users/juanftellez/Dropbox/netsMatter/replications/Reiter_Stam_2003/output/'
+  plotPath = '~/OneDrive/netsMatter/paper/'
+  source('~/OneDrive/netsMatter/code/helpers/ameHelpers.R')
+  source('helperEx.R')
+}
+
+if(Sys.info()['user']=='herme' | 'Owner'){
+	user=Sys.info()['user']
+	base = paste0('C:/Users/',user,'/')
+	fPath = paste0(base, 'Research/netsMatter/code/helpers/')
+	dPath = paste0(base, 'Dropbox/Research/netsMatter/')
+	simResPath = paste0(dPath, 'simulation/')
+	source(paste0(fPath, 'functions.R'))
+	source(paste0(fPath, 'ameHelpers.R'))
+	source(paste0(fPath, 'binPerfHelpers.R'))
+	source(paste0(fPath, 'stargazerHelpers.R'))
+	source(paste0(fPath, 'clusteredSE.R'))
+}
+############################################
+
+# modData ###########################################
+load( paste0(resultsPath,'reiter_stam_glmfit.rda') )
+load( paste0(resultsPath,'model_k2_v12.rda') )
+############################################
+
+# coefSumm ###########################################
+ameSumm = t(apply(ameFit$BETA, 2, function(x){ c(
+	'Estimate'=mean(x), 'Std. Error'=sd(x), 'z value'=mean(x)/sd(x),
[…]

### 2021-03-13 replArchive/2_applications/gibler_ameRun.R
most repls done

@@ -1,0 +1,29 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+gPth = paste0(pth, '2_applications/application_data/gibler/')
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+##############################
+
+##############################
+# load ame version of data
+load(paste0(gPth, 'giblerData.rda'))
+
+# run ame and save
+startTime = Sys.time()
+ameFit = ame_repL(
+	Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL,
+	model="bin",symmetric=TRUE, # MID DV: undirected data
+	R=2,
+	nscan=50000, seed=6886, burn=25000, odens=25,
+	plot=FALSE, print=FALSE,
+	startVals = startVals0$startVals )
+endTime = Sys.time()
+print( endTime-startTime )
+save(ameFit, startTime, endTime, file=paste0(gPth, 'ameFitGibler.rda'))
+##############################

### 2021-03-13 replArchive/2_applications/gibler_glmRun.R

@@ -1,0 +1,41 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+gPth = paste0(pth, '2_applications/application_data/gibler/')
+
+#
+source(paste0(pth, 'helpers/clusteredSE.R'))
+library(magrittr)
+library(lmtest)
+##############################
+
+##############################
+# load data
+load(paste0(gPth,"gibler2017_replication.rda"))
+data <- gilberData
+names(data)[which( names(data) %in% paste0('_spline',1:3) )] = paste0('spline',1:3)
+##############################
+
+##############################
+# set up and run glm model
+# modspec # table 6 # column 6
+ids = c('ccode1', 'ccode2', 'year', 'dyad')
+dv = 'cwmid'
+ivs = c('allied','jointdem', 'cwpceyrs_bkt',
+	'spline1', 'spline2', 'spline3', 'contig', 'parity4',
+	'firstparity', 'riv1' )
+data = na.omit( data[,c(ids, dv, ivs)] )
+
+# set up model formula
+modForm = paste(ivs, collapse=' + ') %>% paste0(dv, ' ~ ', .) %>% formula()
+
+# run
+mod = glm(modForm, family = binomial(link = 'logit'), data = data)
+
+# get clustered SEs
+modSumm = coeftest(mod, vcov = vcovCluster(mod, cluster = data$dyad))
+
+# save
+save(mod, modSumm, file=paste0(gPth, 'glmFitGibler.rda'))
+##############################

### 2021-03-13 replArchive/2_applications/reiter_stam_ameRun.R

@@ -1,0 +1,33 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+rsPth = paste0(pth, '2_applications/application_data/reiter_stam/')
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+##############################
+
+##############################
+# load data
+load(paste0(rsPth, 'reiterStamData.rda'))
+##############################
+
+##############################
+# run and save
+# ~5-6 hours
+startTime = Sys.time()
+ameFit = ame_repL(
+  Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL,
+  model="bin",symmetric=FALSE,intercept=TRUE,
+  R=2,
+  nscan=10000, seed=6886, burn=25000, odens=10,
+  plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0$startVals,
+  periodicSave=TRUE
+)
+endTime = Sys.time()
+print( endTime-startTime )
+save(ameFit, startTime, endTime, file=paste0(rsPth, 'ameFitReiterStam.rda'))
+##############################

### 2021-03-13 replArchive/helpers/binPerfHelpers.R

@@ -1,0 +1,228 @@
+loadPkg(c('ROCR', 'caTools','RColorBrewer'))
+
+# Roc curve, depends ROCR
+roc = function(prediction, actual){
+      pred = prediction(prediction, actual)
+      perf = performance(pred,"tpr","fpr")
+      rocData = data.frame(attributes(perf)$x.values[[1]], attributes(perf)$y.values[[1]])
+      names(rocData) = c('FPR', 'TPR')
+      return(rocData)
+}
+
+# Auc, depends ROCR
+getAUC = function(prediction, actual){
+	pred = prediction(prediction, actual)	
+	attributes(performance(pred,"auc"))$y.values[[1]]
+}
+
+# Plot roc curves, depends RColorBrewer
+# plot_type is "roc" or "pr"
+rocPlot = function(
+  rocData, type='roc', legPos=c(.56,.25), colorPal = 'Set1', 
+  colorManual=NULL, linetypes, legText=6, legSpace=3){
+
+	if(type=='roc'){ 
+    tmp=ggplot(rocData, aes(x=FPR, y=TPR, color=model, linetype=model)) + 
+      geom_abline(intercept=0, slope=1, color='darkgrey') + 
+      ylab('True Positive Rate (Sensitivity)') + xlab('False Positive Rate (1-Specificity)')
+
+  }
+
+  if(type=='pr'){ 
+    tmp=ggplot(rocData, aes(x=rec, y=prec, color=model, linetype=model)) + 
+      ylab('Precision') + xlab('Recall (True Positive Rate)')
+  }
+
+  if(is.null(colorManual)){
+    tmp = tmp + scale_color_brewer(palette=colorPal)
+  } else {
+    tmp = tmp + scale_color_manual(values=colorManual)
+  }
+
+	tmp=tmp + 
+    geom_line(lwd=1) +
+    ylim(0,1) + 
+    scale_linetype_manual(values=linetypes) + 
+    # theme_light(base_family="Source Sans Pro") + 
+    theme(
+  		legend.position=legPos, legend.title=element_blank(),
+      legend.background=element_blank(), 
+      legend.text.align = 0, legend.text=element_text(size=legText),
+      legend.key=element_rect(colour = NA, fill = NA), legend.key.size=unit(legSpace,'lines'),
+      # axis.text.x=element_text(family="Source Sans Pro Light"),
[…]

### 2021-03-13 replArchive/helpers/functions.R

@@ -1,0 +1,25 @@
+# install/load libraries
+loadPkg=function(toLoad){
+	for(lib in toLoad){
+	  if(!(lib %in% installed.packages()[,1])){
+	    install.packages(lib, repos='http://cran.rstudio.com/') }
+	  suppressMessages( library(lib, character.only=TRUE) )
+	}
+}
+
+# some necessary libs
+loadPkg(c(
+	'ggplot2', 'latex2exp', 'Cairo', 'RColorBrewer',	# plotting
+	'xtable', # tables
+	'countrycode',
+	'reshape2', 'tidyr', 'stringr',	'dplyr'
+	))
+
+# Set a theme for gg
+theme_set(theme_bw())
+
+# misc
+pasteVec = function(x,y){ as.vector( outer( x, y, paste0 ) ) }
+char = function(x){ as.character(x) }
+num = function(x){ as.numeric(char(x)) }
+trim = function (x) { gsub("^\\s+|\\s+$", "", x) }

### 2021-03-14 replArchive/2_applications/gibler_ameRun.R
fixing gibler, getting crossval code cleaned

@@ -13,17 +13,19 @@ library(amen)
 ##############################
 # load ame version of data
 load(paste0(gPth, 'giblerData.rda'))
+##############################
 
-# run ame and save
+##############################
+# run ame and save ~2 days
 startTime = Sys.time()
 ameFit = ame_repL(
 	Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL,
 	model="bin",symmetric=TRUE, # MID DV: undirected data
 	R=2,
-	nscan=50000, seed=6886, burn=25000, odens=25,
+	nscan=100000, seed=6886, burn=25000, odens=25,
 	plot=FALSE, print=FALSE,
-	startVals = startVals0$startVals )
+	startVals = startVals0 )
 endTime = Sys.time()
 print( endTime-startTime )
-save(ameFit, startTime, endTime, file=paste0(gPth, 'ameFitGibler.rda'))
+save(ameFit, startTime, endTime, file=paste0(gPth, 'ameFitGibler_v2.rda'))
 ##############################

### 2021-03-14 replArchive/2_applications/reiter_stam_crossval.R
fixing gibler, getting crossval code cleaned

@@ -1,0 +1,16 @@
+#############################
+# set a path
+require(here)
+pth = paste0(here::here(), '/replArchive/')
+rsPth = paste0(pth, '2_applications/application_data/reiter_stam/')
+
+# install github from specific repo
+# to get same results
+# devtools::install_github('s7minhas/amen', ref='pa2018_version')
+library(amen)
+##############################
+
+##############################
+# load data
+load(paste0(rsPth, 'reiterStamData.rda'))
+##############################

### 2021-07-01 replArchive/2_applications/appendix_tableB3.R
paper revisions per chat

@@ -23,6 +23,8 @@ ameFit$BETA = do.call('rbind', lapply(pthFiles, function(pthFile){
 	load(pthFile) ; return( ameFit$BETA ) } ) )
 
 # load glm data
+if(!file.exists(paste0(gPth,'glmFitGibler.rda'))){
+	source(paste0(pth, '2_applications/gibler_glmRun.R')) }
 load( paste0(gPth,'glmFitGibler.rda') )
 ##############################
 

### 2021-07-01 paper_v2/legacy/app_deprc.tex
paper revisions per chat

--- 2018-02-22 paper/app_deprc.tex
@@
 %Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework suggesting that some of their effects were actually the products of interdependencies.
+
+\subsection{Re-estimation of McDonald (2004)}
+
+\citet{mcdonald:2004} studies whether trade promotes peace between nations.
+\citet[p.
+547]{mcdonald:2004} introduced the argument that free trade between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''.
+Accordingly, he provided evidence challenging the generalized linkage between peace and trade and refined the measurement of the standard ``trade'' variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states.
+His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences.
+\citet{mcdonald:2004} measured free trade in two ways.
+The first, the ``protection'' variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad; this reflects the notion that larger protected sectors generate greater societal pressures resulting in pockets of support for war.
+This measure thus captures the score of the state in the dyad that possesses higher barriers to trade.
+\citet[p.
+560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''.
+The (binary) dependent variable is the onset of a new militarized interstate dispute within a given dyad, and \citet{mcdonald:2004} includes splines to correct for temporal dependence with robust standard errors clustered on each dyad.
 

### 2021-07-01 replArchive/1_simulation/appendixD/appendix_figureD1.R

@@ -2,6 +2,7 @@
 # set a path
 require(here)
 pth = paste0(here::here(), '/')
+pth = 'C:/Users/S7M/Research/netsMatter/replArchive/'
 graphicsPath=simResPath = pth
 
 # install github from specific repo
@@ -27,6 +28,18 @@ facet_labeller = function(string){ TeX(string) }
 ##############################
 # params
 NSIM = 1000 ; intEff=-2 ; x1Eff=1 ; x2Eff=1
+
+# check if sim results are in path
+files = paste0(simResPath, c(
+	'ameSim50_corrProbitMed.rda',
+	'ameSim100_corrProbitMed.rda',
+	'ameSim50_corrProbitHi.rda',
+	'ameSim100_corrProbitHi.rda' ))
+fileLogical = !all( file.exists( files ) )
+
+# if they arent in path then run sim script
+if( fileLogical ){
+	source(paste0(pth, '1_simulation/appendixD/1_ameSim_rev2_probitCorrel.R')) }
 
 # iterate over correlation levels
 figNames = c('a', 'b')

### 2021-07-01 paper_v2/coverletter/netsMatter_coverLetter.tex
paper revisions per chat

--- 2019-07-12 paper/coverletter/netsMatter_coverLetter.tex
@@
 This letter accompanies our submission of a manuscript for your consideration.
-The manuscript ``Taking Dyads Seriously'' introduces the Additive and Multiplicative Effects (AME) framework for conducting inference in the context of the dependencies that we often observe in dyadic data.
+The manuscript ``Taking Dyads Seriously'' discusses and tests the utility of the Additive and Multiplicative Effects (AME) framework for conducting inference in the context of the dependencies that often occur in dyadic data.
 In submitting this paper, we feel as though it is necessary to first provide some background information.
@@
 This unproductive fight is not mirrored in the statistical or broader networks literature, as in those communities people recognize these models as having quite distinct goals.
-These models having differing goals is something that I also believe and which is why in a recent ERGM related article I reviewed for PSRM, I did not express any need for the authors to engage in this dispute.
+These models having differing goals is something that I also believe and which is why in a recent ERGM related article I reviewed for PSRM, I did not express any need for the authors to engage in a lengthy contrast of their approach with latent variable models.
 \\[1ex]

### 2021-07-01 paper_v2/4_app.tex
paper revisions per chat

--- 2021-02-05 paper/4_app.tex
@@
 		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/weeks_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/weeks_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/gibler_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/gibler_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_pr_outSample.pdf} \\
 	\end{tabular}
@@
 	\begin{tabular}{cc}
-	\includegraphics[width=.7\textwidth]{graphics/reiter_stam_aEff_map.pdf} &
-	\includegraphics[width=.3\textwidth]{graphics/reiter_stam_aEff_line.pdf} \\
+	\includegraphics[width=.7\textwidth]{graphics/figure7a.pdf} &
+	\includegraphics[width=.3\textwidth]{graphics/figure7b.pdf} \\
 	\end{tabular}
@@
 \begin{figure}[!h]
-	\includegraphics[width=\textwidth]{graphics/weeks_circPlot.png}
+	\includegraphics[width=\textwidth]{graphics/figure8.png}
 	\caption{ Visualization of multiplicative effects for Weeks (2012).
@@
 	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
-	\includegraphics[width=\textwidth]{graphics/gibler_margeff.pdf}
+	\includegraphics[width=\textwidth]{graphics/figure9.pdf}
  	\label{fig:gibmargeff}

### 2021-07-02 replArchive/1_simulation/appendixD/appendix_figureD1.R
cleanup

@@ -2,7 +2,6 @@
 # set a path
 require(here)
 pth = paste0(here::here(), '/')
-pth = 'C:/Users/S7M/Research/netsMatter/replArchive/'
 graphicsPath=simResPath = pth
 
 # install github from specific repo
@@ -118,23 +117,26 @@ ggBiasPlot = function(varName, figLabel, h=4, w=8){
 			legend.title=element_blank(),
 			axis.ticks=element_blank(),
 			panel.border=element_blank(),
-			axis.text.y=element_text(size=8,
-				family="Source Code Pro Light"),
+			axis.text.y=element_text(size=8
+				#, family="Source Code Pro Light"
+				),
 			axis.text.x=element_text(size=10, face='bold'),
 			strip.text.x = element_text(size=9, color='white'
-				, family="Source Code Pro Semibold"),
+				#, family="Source Code Pro Semibold"
+				),
 			strip.text.y = element_text(size=9, color='white',
-				family="Source Code Pro Semibold", angle=0),
+				# family="Source Code Pro Semibold",
+				angle=0),
 			strip.background = element_rect(fill = "#525252", color='#525252')
 			)
 	ggsave(g, height=4, width=8,
-		file=paste0(graphicsPath, 'appendix_figureD1',figLabel,'.pdf'),
-		device=cairo_pdf
+		file=paste0(graphicsPath, 'appendix_figureD1',figLabel,'.pdf')
+		#, device=cairo_pdf
 		)
 	return(g)
 }
 
 #
-print(ggBiasPlot('all', figLabel=figNames[ii], h=12, w=8))
+ggBiasPlot('all', figLabel=figNames[ii], h=12, w=8)
 }
 ##############################

### 2021-07-02 replArchive/2_applications/figure8.R
cleanup

@@ -16,14 +16,14 @@ library(amen)
 # load models
 
 # if ame model for weeks is not present run script
-if(!file.exists(paste0(rsPth, 'weeksStam.rda'))){
+if(!file.exists(paste0(wPth, 'ameFitWeeks.rda'))){
   source(paste0(pth, '2_applications/weeks_ameRun.R')) }
 
 # load in ame results
 load(paste0(wPth,'ameFitWeeks.rda'))
 
 # load in weeks data
-load(paste0(wPth, 'WeeksamenData.rda'))
+load(paste0(wPth, 'weeksData.rda'))
 ##############################
 
 ##############################

### 2021-07-02 replArchive/1_simulation/figure4.R
cleanup

@@ -112,20 +112,21 @@ ggCoverPlot = function(var,h=3, w=8){
 			axis.ticks=element_blank(),
 			panel.border=element_blank(),
 			axis.text.y=element_text(
-				size=8, family="Source Code Pro Light"),
+				size=8#, family="Source Code Pro Light"
+				),
 			axis.text.x=element_text(size=10, face='bold'),
 			strip.text.x = element_text(
-				size=9, color='white',family="Source Code Pro Semibold"
+				size=9, color='white'#,family="Source Code Pro Semibold"
 				),
 			strip.text.y = element_text(
-				size=9, color='white',family="Source Code Pro Semibold",
+				size=9, color='white'#,family="Source Code Pro Semibold",
 				,angle=0
 				),
 			strip.background = element_rect(fill = "#525252", color='#525252')
 			)
 	ggsave(g, height=3, width=8,
-		file=paste0(graphicsPath, 'figure4.pdf'),
-		device=cairo_pdf
+		file=paste0(graphicsPath, 'figure4.pdf')
+		#,device=cairo_pdf
 		)
 	return(g)
 }

### 2021-07-02 replArchive/2_applications/figure6.R

@@ -53,12 +53,12 @@ wFiles = c(
 if(!file.exists(wFiles[1])){
   source(
     paste0(pth, '2_applications/weeks_ameCrossval.R')) }
-load(rsFiles[1])
+load(wFiles[1])
 
 if(!file.exists(wFiles[2])){
   source(
     paste0(pth, '2_applications/weeks_glmCrossval.R')) }
-load(rsFiles[2])
+load(wFiles[2])
 
 # org
 predDfs = list(

### 2021-07-02 replArchive/1_simulation/appendixC/appendix_figureC2.R
cleanup

@@ -120,18 +120,21 @@ ggCoverPlot = function(var,h=3, w=8){
 			legend.title=element_blank(),
 			axis.ticks=element_blank(),
 			panel.border=element_blank(),
-			axis.text.y=element_text(size=8,
-				family="Source Code Pro Light"),
+			axis.text.y=element_text(size=8
+				#, family="Source Code Pro Light"
+				),
 			axis.text.x=element_text(size=10, face='bold'),
-			strip.text.x = element_text(size=9, color='white',
-				family="Source Code Pro Semibold"),
+			strip.text.x = element_text(size=9, color='white'
+				#, family="Source Code Pro Semibold"
+				),
 			strip.text.y = element_text(size=9, color='white',
-				family="Source Code Pro Semibold", angle=0),
+				# family="Source Code Pro Semibold",
+				angle=0),
 			strip.background = element_rect(fill = "#525252", color='#525252')
 			)
 	ggsave(g, height=3, width=8,
-		file=paste0(graphicsPath, 'appendix_figureC2.pdf'),
-		device=cairo_pdf
+		file=paste0(graphicsPath, 'appendix_figureC2.pdf')
+		#, device=cairo_pdf
 		)
 	return(g)
 }

### 2021-07-02 replArchive/1_simulation/appendixC/appendix_figureC1.R
cleanup

@@ -31,6 +31,7 @@ intEff=-2 ; x1Eff=1 ; x2Eff=1
 # check if sim results are in path
 fileLogical = !all( file.exists(
 	paste0(simResPath,'ameSim',c(50,100),'_asaProbit.rda') ) )
+
 # if they arent in path then run sim script
 if( fileLogical ){
 	source(paste0(pth, '1_simulation/appendixC/1_ameSim_asaProbit.R')) }
@@ -99,18 +100,21 @@ ggBiasPlot = function(varName, h=4, w=8){
 			legend.title=element_blank(),
 			axis.ticks=element_blank(),
 			panel.border=element_blank(),
-			axis.text.y=element_text(size=8,
-				family="Source Code Pro Light"),
+			axis.text.y=element_text(size=8
+				#, family="Source Code Pro Light"
+				),
 			axis.text.x=element_text(size=10, face='bold'),
-			strip.text.x = element_text(size=9, color='white',
-				family="Source Code Pro Semibold"),
+			strip.text.x = element_text(size=9, color='white'
+				#, family="Source Code Pro Semibold"
+				),
 			strip.text.y = element_text(size=9, color='white',
-				family="Source Code Pro Semibold", angle=0),
+				#family="Source Code Pro Semibold",
+				angle=0),
 			strip.background = element_rect(fill = "#525252", color='#525252')
 			)
 	ggsave(g, height=4, width=8,
-		file=paste0(graphicsPath, 'appendix_figureC1.pdf'),
-		device=cairo_pdf
+		file=paste0(graphicsPath, 'appendix_figureC1.pdf')
+		#, device=cairo_pdf
 		)
 	return(g)
 }

### 2021-07-05 paper_v2/texTOdoc.sh
final run before dataverse submission

@@ -1,1 +1,0 @@
- pandoc netsMatter.tex --bibliography=/Users/s7m/whistle/master.bib -o netsmatter.docx

### 2021-07-05 replArchive/helpers/ame_repL.R

@@ -1,0 +1,470 @@
+#' AME model fitting routine for replicated relational data
+#'
+#' An MCMC routine providing a fit to an additive and multiplicative effects
+#' (AME) regression model to replicated relational data of
+#' various types.
+#'
+#' This command provides posterior inference for parameters in AME models of
+#' independent replicated relational data, assuming one of six possible data
+#' types/models:
+#'
+#' "nrm": A normal AME model.
+#'
+#' "bin": A binary probit AME model.
+#'
+#' "ord": An ordinal probit AME model. An intercept is not identifiable in this
+#' model.
+#'
+#' "cbin": An AME model for censored binary data.  The value of 'odmax'
+#' specifies the maximum number of links each row may have.
+#'
+#' "frn": An AME model for fixed rank nomination networks. A higher value of
+#' the rank indicates a stronger relationship. The value of 'odmax' specifies
+#' the maximum number of links each row may have.
+#'
+#' "rrl": An AME model based on the row ranks. This is appropriate if the
+#' relationships across rows are not directly comparable in terms of scale. An
+#' intercept, row random effects and row regression effects are not estimable
+#' for this model.
+#'
+#' @usage ame_repL(Y,Xdyad=NULL, Xrow=NULL, Xcol=NULL, rvar = !(model=="rrl")
+#' , cvar = TRUE, dcor = !symmetric, nvar=TRUE,  R = 0, model="nrm",
+#' intercept=!is.element(model,c("rrl","ord")),
+#' symmetric=FALSE,
+#' odmax=NULL, seed = 1,
+#' nscan = 10000, burn = 500, odens = 25, plot=TRUE, print = TRUE, gof=TRUE)
+#' @param Y a T length list of n x n relational matrices, where T
+#' corresponds to the number of replicates (over time, for example).
+#' See model below for various data types.
+#' @param Xdyad a T length list of n x n x pd arrays of covariates
[…]

### 2021-07-19 diffNotes/netsMatter_v2.tex
highlight diffs for psrm

\documentclass[12pt]{amsart}
      % \newif\ifpreview
      % \previewfalse

% approx iid
\newcommand\simiid{\stackrel{\mathclap{\normalfont\mbox{\tiny{iid}}}}{\sim}}
\usepackage{natbib}
\usepackage{titlecaps}
\Addlcwords{a about above after again against all am an and any are aren't as at be because been before being below between both but by can't cannot could couldn't did didn't do does doesn't doing don't down during each few for from further had hadn't has hasn't have haven't having he he'd he'll he's her here here's hers herself him himself his how how's i i'd i'll i'm i've if in into is isn't it it's its itself let's me more most mustn't my myself no nor not of off on once only or other ought our ours ourselves out over own same shan't she she'd she'll she's should shouldn't so some such than that that's the their theirs them themselves then there there's these they they'd they'll they're they've this those through to too under until up very was wasn't we we'd we'll we're we've were weren't what what's when when's where where's which while who who's whom why why's with won't would wouldn't you you'd you'll you're you've your yours yourself yourselves}

% Fonts
% at some point figure out bolding ...
\usepackage[default,oldstyle,scale=0.95]{opensans}
%\usepackage[sfdefault,scaled=.85]{FiraSans}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[varqu,varl]{zi4}% inconsolata typewriter
\usepackage{amsmath,amsthm}
\usepackage[cmintegrals]{newtxsf}
\usepackage[T1]{fontenc}
\usepackage{ae}

% easy commands for number propers
\newcommand{\first}{1$^{\text{st}}$}
\newcommand{\second}{2$^{\text{nd}}$}
\newcommand{\third}{3$^{\text{rd}}$}
\newcommand{\nth}[1]{${#1}^{\text{th}}$}

% Generate some fake text
\usepackage{blindtext}

% % add watermark
% \usepackage{draftwatermark}
% \SetWatermarkText{Submission Copy}
% \SetWatermarkScale{.5}
% \SetWatermarkColor[gray]{0.9}

% layout control
\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1.25in,lmargin=1in,rmargin=1in}
\usepackage{parallel}
\usepackage{parcolumns}
\usepackage{fancyhdr}
\usepackage[export]{adjustbox}
\usepackage{etoolbox}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{relsize}
\usepackage{mathtools}
\usepackage{bm}
%\usepackage[%decimalsymbol=.,digitsep=fullstop]{siunitx}

% restricts float objects to be inserted before end of section
% creates float barriers
\usepackage[section]{placeins}

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{longtable}

% to adapt caption style
\usepackage[font={small},labelfont=bf]{caption}

% footnotes at bottom
%\usepackage[bottom]{footmisc}

% to change enumeration symbols begin{enumerate}[(a)]
\usepackage{enumerate}
\usepackage{paralist}

% to colorize links in document. See color specification below
\usepackage[pdftex,hyperref,x11names]{xcolor}

% for multiple references and insertion of the word "figure" or "table"
% \usepackage{cleveref}

% load the hyper-references package and set document info
\usepackage[pdftex]{hyperref}

% graphics stuff
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage[space]{grffile} % allows us to specify directories that have spaces
\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
\usepackage{tikz}

% sideway figures
\usepackage{rotating}
\usepackage{lscape}

%\usepackage[figuresright]{rotating}
% \newenvironment{amssidewaysfigure}  {\begin{sidewaysfigure}\vspace*{.8\textwidth}\begin{minipage}{\textheight}\centering}
%   {\end{minipage}\end{sidewaysfigure}}

% Spacing
\usepackage[doublespacing]{setspace}

% outside code
\usepackage{listings}
\usepackage{enumitem}

% define clickable links and their colors
\hypersetup{%
	unicode=false,          % non-Latin characters in Acrobat's bookmarks
	pdftoolbar=true,        % show Acrobat's toolbar?
	pdfmenubar=true,        % show Acrobat's menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdfnewwindow=true,%
	%pagebackref=false,%
	pdfauthor={Shahryar Minhas, et alia.},%
	pdftitle={Taking Dyads Seriously},%
	colorlinks,%
	citecolor=black,%
	filecolor=black,%
	linkcolor=black,%
	urlcolor=RoyalBlue4
}

% -------------------- title -------------------- %

\title{Taking Dyads Seriously}
\vspace{\baselineskip}

% \author[Minhas]{Shahryar Minhas}
% \address{Shahryar Minhas: Department of Political Science}
% \curraddr{Michigan State University}
% \email[Corresponding author]{minhassh@msu.edu}

% \author[Dorff]{Cassy L. Dorff}
% \address{Cassy L. Dorff}
% \curraddr{Department of Political Science, University of New Mexico}
% \email{cdorff@unm.edu}

% \author[Foster]{Margaret Foster}
% \address{Margaret Foster}
% \curraddr{Department of Political Science, Duke University}
% \email{margaret.foster@duke.edu}

% \author[Gallop]{Max Gallop}
% \address{Max Gallop}
% \curraddr{Department of Government and Public Policy, University of Strathclyde}
% \email{max.gallop@strath.ac.uk}

% \author[Liu]{Howard Liu}
% \address{Howard Liu}
% \curraddr{Department of Political Science, Duke University}
% \email{hao.liu@duke.edu}

% \author[Tellez]{Juan Tellez}
% \address{Juan Tellez}
% \curraddr{Department of Political Science, Duke University}
% \email{juan.tellez@duke.edu}

% \author[Ward]{Michael D. Ward}
% \address{Michael D. Ward}
% \curraddr{Department of Political Science, Duke University}
% \email{michael.d.ward@duke.edu}

% \date{\today}

% \thanks{Shahryar Minhas and Michael D. Ward acknowledge support from National Science Foundation (NSF) Award 1259266. Peter Hoff, John Ahlquist, Robert Franzese, Arturas Rozenas, Betsy Sinclair, Jacob Montgomery and Benjamin Appel provided welcome and helpful comments. We also thank Douglas Gibler for kindly providing us with a copy of the data and programs he used in \citet{gibler:2017}. We benefited from presentations of this material at the 2017 Political Methodology Conference, and our home institutions.}

\setlength{\headheight}{15pt}
\setlength{\headsep}{20pt}
\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{}}
% \chead{\fancyplain{}{Taking Dyads Seriously}}
\chead{\fancyplain{}{}}
% \rhead{\fancyplain{}{\today}}
\rhead{\fancyplain{}{}}
\rfoot{\fancyplain{}{\thepage}}

% references to graphics
\makeatletter
% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}

% easy command for boldface math symbols
\newcommand{\mbs}[1]{\boldsymbol{#1}}

% command for R package font
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

% Add some colors
\definecolor{red1}{RGB}{253,219,199}
\definecolor{red2}{RGB}{244,165,130}
\definecolor{red3}{RGB}{178,24,43}

\definecolor{green1}{RGB}{229,245,224}
\definecolor{green2}{RGB}{161,217,155}
\definecolor{green3}{RGB}{49,163,84}

\definecolor{blue0}{RGB}{255,247,251}
\definecolor{blue1}{RGB}{222,235,247}
\definecolor{blue2}{RGB}{158,202,225}
\definecolor{blue3}{RGB}{49,130,189}
\definecolor{blue4}{RGB}{4,90,141}

\definecolor{purple1}{RGB}{191,211,230}
\definecolor{purple2}{RGB}{140,150,198}
\definecolor{purple3}{RGB}{140,107,177}

\definecolor{brown1}{RGB}{246,232,195}
\definecolor{brown2}{RGB}{223,194,125}
\definecolor{brown3}{RGB}{191,129,45}

% square bracket matrices
\let\bbordermatrix\bordermatrix
\patchcmd{\bbordermatrix}{8.75}{4.75}{}{}
\patchcmd{\bbordermatrix}{\left(}{\left[}{}{}
\patchcmd{\bbordermatrix}{\right)}{\right]}{}{}

\begin{document}
% \ifpreview {
%    \definecolor{personal}{rgb}{1.0, 1.0, 0.0}
%    \pagecolor{personal}
%    }
% \fi

\small{\singlespacing{
\begin{abstract}
	International relations scholarship concerns dyads, yet standard modeling approaches fail to adequately capture the data generating process behind dyadic events and processes. As a result, they suffer from biased coefficients and poorly calibrated standard errors. We show how a regression-based approach, the Additive and Multiplicative Effects (AME) model, can be used to account for the inherent dependencies in dyadic data and glean substantive insights in the interrelations between actors. First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference on dyadic data. Second, we compare the AME model to approaches used in three prominent studies from recent international relations scholarship. For each study, we find that compared to AME, the modeling approach used performs notably worse at capturing the data generating process. Further, conventional methods misstate the effect of key variables and the uncertainty in these effects. Finally, AME outperforms standard approaches in terms of out-of-sample fit. In sum, our work shows the consequences of failing to take the dependencies inherent to dyadic data seriously. Most importantly, by better modeling the data generating process underlying political phenomena, the AME framework improves scholars' ability to conduct inferential analyses on dyadic data.
\end{abstract}
}}

\maketitle

\begin{center}
	\textbf{Word count}: 9,542
\end{center}

\thispagestyle{empty}
\newpage\setcounter{page}{1}

\section{\textbf{Introduction}}

The aim of this study is to address how to estimate regression coefficients in a generalized linear model (GLM) context when there are network dependencies in dyadic data. Specifically, we discuss and evaluate how well the Additive and Multiplicative Effects (AME) model can be used to account for the interdependencies underlying the data generating process of dyadic structures \citep{hoff:2005,hoff:2008,minhas:etal:2019} in International Relations data. The AME works by including a set of parameters meant to capture network effects in the conditional mean equation of the GLM.

We focus on three types of network effects that can complicate dyadic analyses. First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict.\footnote{In the case of undirected data where there is no clear sender or receiver, it is still essential to take into account the variance in how active actors are in the system.} Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict against a government, the government will likely reciprocate that behavior. We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model. Third-order dependencies capture relationships of transitivity, balance, and clusterability between different dyads. For example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we understand that the United States invaded Iraq in 2003 and that Poland often participates in US-led coalitions. The multiplicative effects capture these sorts of dependencies, specifically, those that result because the specified model has not accounted for a latent set of shared attributes that affect actors' probability of interacting with one another.

We begin with a discussion of these dependencies and an introduction to the AME model. Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network based dependence. Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature. Our comparison reveals that accounting for observational dependence leads to results that, at times, differ from those found in the original study as well as from the broader literature. Moreover, we demonstrate that the additional parameters included by AME in the conditional mean equation of a typical GLM can offer substantive insights that are often occluded by ignoring the interdependencies found in relational data. Finally, we show that for each replication our network-based approach provides substantially more accurate out-of-sample predictions than the models used in the original studies.

The AME approach advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest. Thus, the AME allows scholars to achieve a two-fold goal: to continue to generate meaningful, substantive insights about political phenomena without abandoning a regression based approach, while at the same time accounting for the data generating processes behind such events of interest. Perhaps most importantly, the AME approach concentrates on the relational aspect of international relations through a statistical framework that is familiar to most scholars.

\section{\textbf{Dependencies in Dyadic Data}}

When modeling dyadic data, scholars typically employ a GLM estimated via maximum-likelihood. This type of model is expressed via a stochastic and systematic component \citep{pawitan:2013}. The stochastic component reflects our assumptions about the probability distribution from which the data are generated: $y_{ij} \sim P(Y | \theta_{ij})$, with a probability density or mass function, such as the normal, binomial, or Poisson, and we assume that each dyad in the sample is independently drawn from that particular distribution, given $\theta_{ij}$. The systematic component characterizes the model for the parameters of that distribution and describes how $\theta_{ij}$ varies as a function of a set of nodal and dyadic covariatse, $\mathbf{X_{ij}}$: $\theta_{ij} = \bm\beta^{T} \mathbf{X}_{ij}$. A fundamental assumption we make when applying this modeling technique is that given $\mathbf{X}_{ij}$ and the parameters of our distribution, each of the dyadic observations are conditionally independent. The importance of this assumption becomes clearer in the process of estimating a GLM via maximum likelihood. After having chosen a set of covariates and specifying a distribution, we construct the joint density function over all dyads, for example:

\vspace{-8mm}
\begin{align}
\begin{aligned}
	P(y_{ij}, y_{ik}, \ldots | \theta_{ij}, \theta_{ik}, \ldots ) &= P(y_{ij} | \theta_{ij}) \times P(y_{ik} | \theta_{ik}) \times \ldots  \\
	P(\mathbf{Y} \; | \; \bm{\theta}) &= \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})  \\
\end{aligned}
\end{align}

\noindent We next convert the joint probability into a likelihood: $\displaystyle \mathcal{L} (\bm{\theta} | \mathbf{Y}) = \prod_{\alpha=1}^{n \times (n-1)} P(y_{\alpha} | \theta_{\alpha})$.

We can then estimate the parameters by maximizing the likelihood, or, more typically, the log-likelihood. However, the likelihood as defined above is only valid if we are able to make the assumption that, for example, $y_{ij}$ is independent of $y_{ji}$ and $y_{ik}$ given the set of covariates we specified, or the values of $\theta_{ij}$. Without the assumption of conditional observational independence the joint density function cannot be written in the way described above and a valid likelihood does not exist. Accordingly, inferences drawn from misspecified models that ignore potential interdependencies between dyadic observations are likely to have a number of issues, including biased estimates of the effect of independent variables, uncalibrated confidence intervals, and poor predictive performance. The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}. As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.

% Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another and not passing any auxiliary information to the model about how observations may be interdependent. For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States. Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan. While most scholars begin with the assumption that each dyadic interaction takes place in isolation of the others, we know this assumption to be false both in theory and in practice. Relational data comes with an explicit structure that generally leads to particular types of dependencies. The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}. As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.

To uncover the underlying structure of relational data, it is helpful to restructure dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}. Rows designate the senders of an event and columns the receivers. The cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns. Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flow, that is sent from actor $i$ to actor $j$. In many applications, scholars are interested in studying undirected (i.e., symmetric) outcomes in which there is no clear sender or receiver, these type of outcomes still can, and we argue should, be studied using the type of framework we discuss below.

Using the structure of an adjacency matrix, Figure~\ref{fig:adjMatDeps} visualizes the types of first- and second-order dependencies that can complicate the analysis of relational data in traditional GLMs. The adjacency matrix on the top left highlights a particular row to illustrate that these values may be more similar to each other than other values because each has a common sender. Interactions involving a common sender also manifest heterogeneity in how active actors are across the network when compared to each other. In most relational datasets (e.g., trade flows, conflict, participation in international organizations, even networks derived from Twitter or Facebook), we often find that there are some actors that are much more active than others \citep{barabasi:reka:1999}. For example, in an analysis of international trade certain countries (e.g., China) export much larger volumes than other countries for a variety of structural, contextual, and idiosyncratic reasons. Unless one is able to develop a model that can account for the variety of explanations that may play a role in determining why a particular actor is more active than others, parameter estimates from standard statistical models will be biased.\footnote{In an undirected setting instead of studying sender and receiver heterogeneity, we would just be concerned with actor heterogeneity in general.}

\begin{figure}
	\begin{tabular}{cc}
		\textsc{Sender Heterogeneity} & \textsc{Receiver Heterogeneity} \\
		\includegraphics[width=.4\textwidth]{graphics/adjRowDep.png} &
		\includegraphics[width=.4\textwidth]{graphics/adjColDep.png} \\
		\textsc{Sender-Receiver Covariance} & \textsc{Reciprocity} \\
		\includegraphics[width=.4\textwidth]{graphics/adjRowColCovar.png} &
		\includegraphics[width=.4\textwidth]{graphics/adjRecip.png} \\
	\end{tabular}
	\caption{Nodal and dyadic dependencies in relational data.}
	\label{fig:adjMatDeps}
\end{figure}

For similar reasons one also needs to take into account the dependence between observations that share a common receiver. The bottom-left panel in Figure~\ref{fig:adjMatDeps} illustrates that sender and receiver type dependencies can also blend together. Specifically, actors who are more likely to send ties in a network tend to also be more likely to receive them. As a result, the rows and columns in an adjacency matrix are often correlated. For example, consider trade flows both from and to many wealthy, developed countries. The bottom-right panel highlights a second-order dependence, specifically, reciprocity. This is a dependency occurring within dyads involving the same actors whereby values of $y_{ij}$ and $y_{ji}$ are correlated. The concept of reciprocity has deep roots in the study of relations between states \citep{richardson:1960,keohane:1989}.

For most relational data, however, dependencies do not simply manifest at the nodal or dyadic level. More often we find significant evidence of higher-order structures that result from dependencies between multiple groups of actors. These dependencies arise because there may be a set of latent attributes between actors that affects their probability of interacting with one another \citep{zinnes:1967,wasserman:faust:1994}. In Figure~\ref{fig:thirdDeps} we provide a visualization of a simulated relational dataset wherein the nodes designate actors and edges between the nodes indicate that an interaction between the two took place. To highlight third-order dependence patterns, nodes with similar latent attributes are colored similarly.

\begin{figure}[ht]
	\includegraphics[width=.6\textwidth]{graphics/stochEquiv_v2.pdf}
	\caption{Higher-order dependence patterns in a network.}
	% need a simpler caption here
	\label{fig:thirdDeps}
\end{figure}

The visualization illustrates that actors belonging to the same group have a higher likelihood of interacting with each other, whereas interactions across groups are rarer. A prominent example of a network with this type of structure is discussed by \citet{adamic:glance:2005}, who visualize linkages between political blogs preceding the 2004 United States Presidential Election. \citeauthor{adamic:glance:2005} find that the degree of interaction between right- and left-leaning blogs is minimal and that most blogs are linked to others that are politically similar. This showcases the types of higher-order dependencies that can emerge in relational data. First, the fact that interactions are determined by a shared attribute, in this case political ideology, is an example of \textit{homophily}. Homophily explains the emergence of patterns such as transitivity (``a friend of a friend is a friend'') and balance (``an enemy of a friend is an enemy''), which also have a long history in international relations. The other major type of meso-scopic features that emerge in relational data is community structure \citep{mucha:etal:2010}, which is often formalized through the concept of stochastic equivalence \citep{anderson:etal:1992}. Stochastic equivalence refers to a type of pattern in which actors can be divided into groups such that members of the same group have similar patterns of relationships. In the example above, each of the left leaning blogs would be considered stochastically equivalent to one another because any given left-leaning blog is more likely to interact with a blog of a similar political position and less likely to interact with one of a divergent political position.

These types of patterns frequently emerge in IR contexts.\footnote{For example, see: \citet{manger:etal:2012, kinne:2013, chyzh:2016}.} For example, a perennial finding in the interstate trade literature emphasizes the role that geography plays in determining trade flows. Geographic proximity in the network context is an example of homophily---a shared attribute between actors that corresponds to a greater likelihood of the event of interest taking place. Alternatively, in the interstate conflict literature, we may find that actors who are each a member of a particular (formal or informal) alliance are likely to act similarly in the conflict network. Specifically, they will tend to initiate conflictual events with actors that their fellow alliance members initiate conflict with, and they will be unlikely to initiate conflict with members of their alliance---an example of stochastic equivalence. In both these examples, we are able to explicitly parameterize the attribute that might explain the emergence of higher order dependence patterns. While sometimes the conditions driving these patterns, such as geography, are easy to identify, at other times it can be difficult to describe exactly why higher order dependence patterns in networks may develop.

% International Trade, we might observe homophily wherein states with open economies and an absence of Non-Tariff Barriers are more likely to trade with each-other all else equal; alternatively we might observe heterophily in the sense that states that have greater endowments of capital are more likely to trade with states with greater endowments of land or labor.

%  We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another. In this case, countries with (formal or informal) alliances are a member of the same group due to similar conflictual ties to other actors in the system.

\section{\textbf{Additive and Multiplicative Effect Models for Networks}}

To account for the dependencies that are prevalent in dyadic data, we turn to the AME model. The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages. It is flexible and easy to use for analyzing the kind of relational data often found in the social sciences. It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{minhas:etal:2019} detail how this framework contrasts with alternative network-based approaches. } The AME model combines the social relations regression model (SRRM) to account for nodal and dyadic dependencies and the latent factor model (LFM) for third-order dependencies.\footnote{An earlier version of the LFM  used in AME is presented as the general bilinear mixed effects (GBME) model in \citet{hoff:2005}. The GBME model is more limited in the types of dependence patterns that it can capture due to the formulation of the matrix decomposition procedure.}  For details on the SRRM see \citet{li:loken:2002,hoff:2005,dorff:minhas:2017}. The AME model is specified as follows:

\begin{align}
	\begin{aligned}
		y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
		\theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} \text{\qquad(Exogenous parameters)} \\
		& + a_{i} + b_{j} + \epsilon_{ij} \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
		& + \mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\
	\label{eqn:ame}
	\end{aligned}
\end{align}

% We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$).
where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver) at time $t$. We model a latent variable, $\theta_{ij}$, first using a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$). $f$ is typically a mapping function, and can be one that applies to dichotomous, ordinal, or continuous distributions.

Next, to account for the dependencies that emerge in dyadic data that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM. $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:

\begin{align}
	\begin{aligned}
		\{ (a_{1}, b_{1}), \ldots, (a_{n}, b_{n}) \} &\simiid N(0,\Sigma_{ab}) \\
		\{ (\epsilon_{ij}, \epsilon_{ji}) : \; i \neq j\} &\simiid N(0,\Sigma_{\epsilon}), \text{ where } \\
		\Sigma_{ab} = \begin{pmatrix} \sigma_{a}^{2} & \sigma_{ab} \\ \sigma_{ab} & \sigma_{b}^2   \end{pmatrix} \;\;\;\;\; &\Sigma_{\epsilon} = \sigma_{\epsilon}^{2} \begin{pmatrix} 1 & \rho \\ \rho & 1  \end{pmatrix}
	\label{eqn:srm}
	\end{aligned}
\end{align}

The sender and receiver random effects are modeled jointly from a multivariate normal distribution to account for correlation in how active an actor is in sending and receiving ties. Heterogeneity in the sender and receiver effects is captured by $\sigma_{a}^{2}$ and $\sigma_{b}^{2}$, respectively, and $\sigma_{ab}$ describes the linear relationship between these two effects (i.e., whether actors who send [receive] a lot of ties also receive [send] a lot of ties). Beyond these first-order dependencies, second-order dependencies are described by $\sigma_{\epsilon}^{2}$ and a within dyad correlation, or reciprocity, parameter $\rho$.

The LFM contribution to the AME is in the multiplicative term: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}=\sum_{k \in K} d_{k} u_{ik} v_{jk}$. $K$ denotes the dimensions of the latent space. The construction of the LFM here is actually quite similar to work on low rank approximations in computer science and has been applied to the development of recommender systems that companies like Amazon and Netflix use to model customer behavior \citep{resnick:varian:1997,bennett:lanning:2007}.\footnote{The LFM also shares similarities with work in the econometric literature on interactive fixed effects \citep{bai:2009,pang:2014}. In this stream of work, interactive fixed effects are used to deal with cross-sectional dependence in TSCS data, in such a way that a latent factor for time can be used to capture common shocks to actors and a latent factor on actors can capture varying responses to those shocks.} This model posits a latent vector of characteristics
$\mathbf{u_{i}}$ and $\mathbf{v_{j}}$ for each sender $i$ and receiver $j$. The similarity or dissimilarity of these vectors will then influence the likelihood of activity, and provides a representation of third-order interdependencies. The LFM parameters are estimated by a process similar to computing the singular value decomposition (SVD) of the observed network. When computing the SVD we factorize our observed network into the product of three matrices: $\mathbf{U}, \mathbf{D}, \text{ and }, \mathbf{V}$. This provides us with a low-dimensional representation of our original network.\footnote{The dimensions of $\mathbf{U}$ and $\mathbf{V}$ are $n \times K$ and $\mathbf{D}$ is a $K \times K$ diagonal matrix.} Values in $\mathbf{U}$ provide a representation of how stochastically equivalent actors are as senders in a network or, for example, how similar actors are in terms of who they initiate conflict with. $\hat{\mathbf{u}}_{i} \approx \hat{\mathbf{u}}_{j}$ would indicate that actor $i$ and $j$ initiate battles with similar third actors. $\mathbf{V}$ provide a similar representation but from the perspective of how similar actors are as receivers. The values in $\mathbf{D}$, a diagonal matrix, represent levels of homophily in the network.\footnote{Unlike traditional SVD, in the latent factor model the singular values are not restricted to be positive, thus allowing us to account for both positive and negative homophily.} Note that this model easily generalizes to the case, common in IR, where interactions are undirected (for example the presence of conflict or a bilateral investment treaty). In the case of the SRRM, $\rho$ is constrained to be one and instead of separate sender and receiver random effects a single actor random effect is utilized. For the LFM, an eigen-decomposition scheme is used to capture higher-order dependence patterns. In the application section, we show the applicability of the AME approach to both directed and undirected dyadic data. Parameter estimation in the AME takes place within the context of a Gibbs sampler in which we iteratively sample from the posterior distribution of the full conditionals for each parameter.\footnote{Details on the sampling procedure and the full conditional distributions can be found in: \citet{hoff:2005,hoff:2008,minhas:etal:2019,hoff:2021}.}

Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe. Accounting for this structure is necessary if we are to adequately represent the data generating process. If one can specify each of the nodal, dyadic, and triadic attributes that influence interactions then the conditional independence assumption underlying standard approaches will be satisfied. However, it is rarely the case that this is possible even for TSCS data and thus modeling decisions must account for underlying structure. Failing to do so in either TSCS or dyadic data leads to a number of well-known challenges: a) biased estimates of the effect of independent variables, b) uncalibrated confidence intervals, and c) poor predictive performance. Additionally, by ignoring these potential interdependencies, we often ignore substantively interesting features of the phenomena we investigate. The study of international relations is founded on the relations among actors. Why ignore the interdependencies that led to the study of IR in the first place?

\section{\textbf{Simulation study}}

We explore the utility of AME as an inferential tool for dyadic analysis via a simulation exercise.\footnote{Alternative network based approaches for dyadic data are exponential random graph models (ERGMs) and the related stochastic actor oriented model (SAOM). While both these models have led to numerous contributions to a variety of literatures, the applicability of these approaches may be limited to certain types of networks and individual level characteristics. Specifically, \citet{block:etal:2017} note that these types of  models may not be appropriate in situations where network and behavioral data depend on unobserved latent variables, which is explicitly the focus of our analysis here.} Most scholars working with dyadic data are primarily concerned with understanding the effect of a particular independent variable on a dyadic dependent variable. The goal of our simulation is to assess how well AME can provide unbiased and well-calibrated estimates of regression coefficients in the presence of unobserved dependencies, specifically, homophily. As discussed in the previous section, homophily is the idea that actors are more likely to have a tie if they have similar values on a particular variable, and in networks the presence of homophily can lead to third order dependencies such as transitivity. Homophily can be operationalized by creating a dyadic covariate via the multiplication of a nodal covariate with its transpose. For example, if the nodal covariate was a binary indicator for democracy, multiplying it by its transpose would give us a dyadic covariate that represents whether any dyad is jointly democratic or not.\footnote{This process of operationalizing homophily is equivalent to the `nodematch` function in the `ergm` and `latentnet` packages. There are many other options of operationalizing homophily including calculating the difference in scores that a pair of actors may have on a particular nodal variable.}

Assume that the true data-generating process for a binary variable, $Y$, is given by:

\begin{align}
	Z_{i,j} &=  \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j}, \; \epsilon \sim normal(0,1) \nonumber \\
	Y_{i,j} &= I(Z_{i,j} > 0)
	\label{eqn:sim}
\end{align}

$X_{i,j} = x \times x^{T}$, where $x$ is a nodal covariate that is drawn from a standard normal distribution. Similarly,  $W_{i,j} = w \times w^{T}$, where $w$ is also a nodal covariate that is independently drawn from a standard normal distribution. We generate our binary dependent variable, $Y$, within a probit framework with $Z$ serving as the latent variable. $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is not observed. We compare inference for $\mu$ and $\beta$---the latter parameter would be of primary concern for applied scholars---using three models:

\begin{itemize}
	\item the ``standard'' international relations approach estimated through a generalized linear model;\footnote{Specifically, here we are just regressing $Y$ on $X$ assuming independent errors.}
	\item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);\footnote{Results with higher values of $K$ are similar.}
	\item and an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$.
\end{itemize}

The first model corresponds to the ``standard'' approach in which little is explicitly done to account for dependencies in dyadic data. In the second model, we use the AME framework described in the previous section. For both the first and second models, we are simply estimating a linear model of $X$ on $Y$, and assessing the extent to which inference on the regression parameters are complicated by the presence of unobserved dependencies, $W$. In the last model, we provide an illustration of the ideal case in which we have observed and measured $W$ and include it in our specification for $Y$. The oracle case provides an important benchmark for the standard and AME approaches.

For the simulation we set the true value of $\mu$ (the intercept term) to -2 and $\beta$ (the effect of $X$ on $Y$) to 1.\footnote{The value of $\gamma$ is also set to 1, which corresponds to an example where the $W$ character is associated with homophily.} We conduct two sets of simulations, one in which the number of actors in the network is set to 50 and the other at 100. In total, we run 1,000 simulations where we begin by simulating $Y$ from the specification given in Equation~\ref{eqn:sim} and then for each simulated $Y$ we estimate a standard, AME, and oracle model.

We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$ in Figure~\ref{fig:ameBias}. The panels on the left show the results for when the number of actors is set to 50 and on the  right for 100 actors. The top pair of panels represents the estimates for $\mu$ while the bottom pair do the same for $\beta$. In each case, we find that the estimates for $\mu$ and $\beta$ produced by the standard approach are notably off from their true values. On the other hand, the AME model performs just as well as the oracle at estimating the true parameter values.

\begin{figure}
	\centering
	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
	\label{fig:ameBias}
	\includegraphics[width=1\textwidth]{graphics/figure3.pdf} \\
\end{figure}

Next, we estimate the 95\% credible interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals. The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle, while the standard approach performs poorly by comparison. The implication of the results presented in Figures~\ref{fig:ameBias} and ~\ref{fig:ameCalib} is that standard approaches can often fail at estimating parameter values and conducting inferential tasks in the presence of unobserved dependencies. The AME approach by comparison can be used as a tool for scholars working with dyadic data to still estimate the true effects of their main variables of interest, while accounting for dependencies that do often emerge in dyadic data.

\begin{figure}
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib}
	\includegraphics[width=1\textwidth]{graphics/figure4.pdf} \\
\end{figure}

Moreover, the AME approach allows scholars to better understand what parameters their model may be missing. In the case of the simulation here, $W$ is set as an unobserved dyadic covariate that has a homophilous effect on $Y$. The effect of $W$ is homophilous within this framework because it is a dyadic attribute involving $i$ and $j$ that positively affects the degree to which actors interact with one another, i.e., $y_{ij}$. This type of unobserved dependency will be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$. To estimate how well the model performs, we recover the multiplicative effects term for each simulation and calculate the correlation between it and the unobserved dependency, $W$.\footnote{Specifically, since both the multiplicative effects term and $W$ are continuous dyadic variables, we calculate the Pearson correlation coefficient.} We visualize the distribution of the correlations from each of the 1,000 simulations in Figure~\ref{fig:ameCorr} for the case where the number of actors is set to 100 (top pair of panels) and 50. Additionally, we calculate the median across the correlations and display the result using a vertical line. For both $n=50$ and $n=100$, we find that the multiplicative effects perform very well in capturing the unobserved dependency, which indicates that the AME does not simply capture noise but also works as a tool to estimate unobserved structure.

\begin{figure}
	\centering
	\caption{Distribution of correlation between missing variable and multiplicative random effect in AME across the 1,000 simulations. Vertical line through the distribution represents the median value across the simulations.}
	\label{fig:ameCorr}
	\includegraphics[width=1\textwidth]{graphics/figure5.pdf} \\
\end{figure}

The simulation shows that beyond obtaining less biased and better-calibrated parameter estimates, a key benefit of the AME framework is to directly estimate unobserved dependencies through the random effects structure of the model. Scholars can use this framework in an iterative fashion: beginning with an estimated model, they can then empirically study the structure of the random effects to assess whether there are unobserved covariates that they want to include in their model. Importantly, this simulation underscores how a careful consideration of a systems' interconnectedness, both through theoretical approaches and empirical models, can result in more precise estimates of direct and indirect effects across the system.

\section{\textbf{Applications with AME}}

We apply AME to three recent IR studies: \citet{reiter:stam:2003, weeks:2012, gibler:2017}. Each of these studies use relational data of state interactions and propose both dyadic, monadic, and structural explanations for the behavior of actors in the system. We demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of our network based approach. First, we show that simply by using the AME framework scholars can better model the data generating process behind events of interest. Second, the results of AME estimation are interpretable alongside results using standard approaches. Third, through using this approach we can also quantify the degree to which first-, second-, and third-order dependencies are present.

We obtained the data for these studies from their replication archives.\footnote{Without exception the data was easy to retrieve thanks to the authors' transparency and an increasing norm in the social sciences of open data sharing.} The chosen studies have each gained over 100 or more citations and were recently published.\footnote{We selected our cases based on a few criteria. We focus on studies that are explicitly about International Relations, were published after the year 2000, and were published in a top ranking general political science outlet (for consistency in editorial standards and reviews, we focus on one journal (American Political Science Review). We hope that these three criteria ensure that our paper is readable and interpretable to an applied audience.} Each of these pieces is well-known and posited a hypothesis in which interdependencies are consequential. Reflecting the dominant approach in the literature, the studies tested their hypothesis by employing some form of a generalized linear model. Table~\ref{tab:modelDesign} provides descriptive information for the studies that we replicated.

\begin{table}
\caption{Descriptive information about the replicated studies. }
	\begin{tabular}{lccccc}
		& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
		Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed & Robust \\
		Weeks (2012) & Logit & 1946--1999 & 197 & Directed & Robust \\
		Gibler (2017) & Logit & 1816--2008 & 193 & Undirected & Robust \\ \bottomrule
	\end{tabular}
	\label{tab:modelDesign}
\end{table}

For each of the studies listed above we replicated the authors' key model using their original estimation procedure, a generalized linear model.\footnote{Replicating the key models from each study was straightforward because of the authors' assembled replication scripts.} Next, utilizing the same covariate specifications, we estimate the models with the AME framework.\footnote{In estimating AME, we show results when setting $K=2$. Results with alternative values of $K$ are similar. We recommend that authors try out a range of $K$s.} For \citet{reiter:stam:2003, weeks:2012}, we utilized a version of AME that accounts for the directed nature of the data and for \citet{gibler:2017} we used an undirected version. In the directed formulation, separate random effects are used for senders and receivers in both the additive and multiplicative portions of the model.

A key claim that we have made is that by accounting for dependencies inherent to relational data we can better capture the data generating process behind events of interest. To assess whether or not the AME approach successfully does this, we turn to an out-of-sample cross validation strategy. An out-of-sample approach is essential since relying on in-sample procedures would enable models with more parameters, such as AME, to simply overfit the data. The cross validation procedure is executed as follows. For each study  we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned. Then for each $s \in \{1,\ldots,k\}$, we estimate model parameters with $\{y_{ij,t}: s_{ij,t} \neq s\}$, the data not in set $s$, and predict $\{\hat{y}_{ij,t}: s_{ij,t} = s\}$ from these estimated parameters.

The result of this procedure is a set of sociomatrices $\bm \hat Y$, in which each entry $\hat y_{ij,t}$ is a predicted value obtained from a subset of the data that does not include $y_{ij,t}$. Next we conduct a series of analyses to discern whether or not the AME model provides any benefit for each study. These analyses are summarized in Figure~\ref{fig:perf}. The left-most plot in each of the panels evaluates performance using Receiver Operating Characteristic (ROC) curves. Models that have a better fit according to this metric will follow the upper-left corner border of the ROC space. In addition to ROC curves we also use separation plots \citep{greenhill:etal:2011}.  Separation plots plots visualize each of the observations in the dataset according to their predicted probability. In this graph, the shaded panels correspond to the occurrence of an event. Darker panels are events, lighter panels are non events. If a model performs well, then the events that actually occur would stack to the far right of the graph, where the predicted probabilities generated by the model are highest. The right-most plot in each of the panels evaluates performance using precision-recall (PR) curves. PR curves are useful in situations where correctly predicting events is more interesting than simply predicting non-events \citep{davis:goadrich:2006}. This is especially relevant in the context of our applications here, as they each are trying to model conflict within dyads, which is an infrequent occurrence.

For each of the replications, we find that the AME approach substantially outperforms the original models in terms of out-of-sample predictive performance. This indicates that switching to the AME framework---even when using the same covariate specification as the original studies---enables scholars to better represent the data generating process of their events of interest. The fact that this analysis is done in an out-of-sample context ensures that the AME framework is not simply overfitting with more parameters. Instead, it suggests that the additive and multiplicative effects are capturing underlying structure previously missed by the exogenous covariates in the models.

\begin{figure}
	\centering
	\begin{tabular}{cc}
		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_pr_outSample.pdf} \\
		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_pr_outSample.pdf} \\
		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_roc_outSample.pdf} &
		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_pr_outSample.pdf} \\
	\end{tabular}
	\caption{Assessments of out-of-sample predictive performance using ROC curves, PR curves, and separation plots.}
	\label{fig:perf}
\end{figure}

Ignoring this underlying structure has consequences for inferential analysis.\footnote{As shown in the simulation section, inference in a dyadic setting can become complicated when there are unobserved dependencies.} The fact that there is such a divergence in the performance of AME versus the original estimation procedures highlights that there are unobserved sources of bias in each of these studies. We hone in on the main finding of each study to draw into focus the potential consequences for ignoring these sources of bias and the inferential benefits of the AME estimation procedure.  In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{Unconfirmed} indicates that the statistical significance of the crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and re-estimated models are presented in the Appendix.}

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the replications and a note about whether or not the highlighted finding remains when using our network-based approach.}
	\begin{tabular}{l p{6cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
		& &  accounting for dependencies? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Confirmed} \\ \midrule
		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
		Gibler (2017) & Entry to the International System, Not Power Parity, Determines Conflict & {Partially Confirmed}\\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

An important takeaway here is that many scholars are forced to make knowledge claims based on the statistical significance of a small set of covariates or the differences between these covariates. These differences may change dramatically once interdependencies are taken into account. This outcome follows from AME's ability to better account for the dependencies discussed in the previous section, whereas GLM approaches explicitly assume observational independence conditional on the specified covariates. As this is a widely-known limitation of GLM approaches, scholars often attempt to account for clustering of observations by including additional variables and adjusting the standard errors of the resulting estimates. At best, this method introduces noise and imprecision into results and at worst can produce misleading outcomes.

Next we discuss each of the replications in more detail and highlight the substantive insights drawn from the AME framework.

\subsection{Reiter \& Stam (2003)}

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes. Their work contests prior scholarship claiming that interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state \citep{peceny:etal:2002}. Using a directed dyadic dataset of almost a million observations, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around. In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true.\footnote{Independent variables focus on various encodings of regime types, contiguity, alliance, and capability measures. For a full tabular display of the results see the Appendix.} As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.

The key variables in the original model measure whether or not the sender in the directed dyad is a personalist regime and the target a democratic regime (``Pers/Democ directed dyad") or whether the opposite is true (``Democ/Personalist directed dyad''). The authors find that coefficient of the Pers/Democ Directed Dyad indicator is positive, while the Democ/Personalist directed dyad is too imprecisely measured to indicate a direction. In our re-estimation using the AME framework, we confirm these results, indicating that dictators are likely to initiate or engage in conflict with democratic regimes but not vice versa.

Even though we are able to confirm the original results, employing the AME model offers clear benefits in this case. As already shown in Figure~\ref{fig:perf}, our approach performs notably better in reflecting the data generating process. The reason for this is that there is still underlying structure within this conflict system that the Reiter \& Stam model does not fully capture. To highlight this we visualize the estimated sender random effects ($a_{i}$) from the SRRM portion of the AME framework in Figure~\ref{fig:reiter_stam_aEff}.

\begin{figure}[ht]
	\begin{tabular}{cc}
	\includegraphics[width=.7\textwidth]{graphics/figure7a.pdf} &
	\includegraphics[width=.3\textwidth]{graphics/figure7b.pdf} \\
	\end{tabular}
	\caption{Estimates of sender random effects ($a_{i}$) from AME for the Reiter \& Stam (2003) model. Positive values indicate that the particular country is more likely to be involved in conflict than predicted by the covariates in the model. Negative values indicate that the country is less likely to be involved in conflict. The left-panel shows $a_{i}$ estimates for all countries and right-panel highlights the top ten countries in terms of positive and values of $a_{i}$.}
	\label{fig:reiter_stam_aEff}
\end{figure}
\FloatBarrier

The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam. Specifically, we can see that countries such as Iraq, Israel, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict. Further, other countries such as Sweden, Finland, and Swaziland are less likely to engage in conflict than the exogenous covariates in the model would suggest. In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for \citet{colgan:2010}'s theory that revolutionary petrostates are more aggressive. None of these findings change the key conclusion from Reiter \& Stam's work, but by using the AME framework we are able to better understand the limitations of their model.

\subsection{Weeks (2012)}

\nocite{weeks:2012} Weeks examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders. She argues that in some circumstances autocrats are held accountable for their foreign policy decisions, and that this is dependent on the audiences of autocrats. When the autocratic regime is nonmilitary, domestic audiences do not favor military actions, but in military autocracies this is not the case. Further, she argues that in personalistic regimes without a military or civilian domestic audience, the leaders are more likely to employ military force in their foreign policy. To study this question, Weeks employs a directed dyad design of conflict similar to that used by Reiter \& Stam.

The major innovation in her study resides in the nuanced way she conceptualizes and codes regimes into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen.\footnote{Weeks also includes a variety of control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.'' For a full tabular display of the results see the Appendix.} She uses a logistic regression, following \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with dyad clustered standard errors. The key findings of her work are that a) juntas, boss, and strongmen regimes are more likely to initiate conflict than machine-type regimes and b) machine-type regimes are no more belligerent than democracies.\footnote{These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper.} In the empirical analysis, Weeks finds that machines are less prone to initiate conflict than the reference category, whereas juntas, bosses, and strongmen are more conflict-prone. When analyzing the results using AME, however, we find that the parameters on each of her autocratic regime type variables are too imprecisely measured to draw any inference about their putative causal effects. Consequentially, none of the findings from her original analysis are confirmed once known dependencies among the data are taken into account via AME.

There is also a striking difference between Weeks' original model and our estimation using AME in terms of capturing the data generating process. As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest. To uncover this structure, we illustrate another beneficial aspect of the AME framework, which is the multiplicative random effects estimated through the LFM portion of the model: $\mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}$. These random effects account for higher order dependencies that manifest as a result of homophily and stochastic equivalence in dyadic data.

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{graphics/figure8.pdf}
	\caption{ Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from.
	}
	\label{fig:weekscirc}
\end{figure}
\FloatBarrier

To visualize the multiplicative effects, we display the circular diagram shown in Figure~\ref{fig:weekscirc}. The nodes throughout the diagram represent countries and are colored according to their geographic position -- a legend is provided in the center. The outer ring visualizes higher order dependence patterns through countries' sender relationships -- country positions here are estimated in the $\mathbf{u}_{i}$ random effects described in Equation~\ref{eqn:ame}. Countries that are more proximate to each other in this outer ring are more likely to send or initiate conflicts with similar targets. The inner ring, based on estimates of $\mathbf{v}_{i}$, is constructed such that countries closer together are more likely to receive conflict from the same sender countries. Last, the distance between countries in the inner and outer rings proportionately reflects who a country is more likely to be in conflict with.

Figure~\ref{fig:weekscirc} reveals a number of notable clusters. For example, in the bottom right corner we see the US, UK, Germany, Canada, and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets. Conversely, in the top left of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China. We observe similar clusters in the inner ring. Specifically, we again see the US, UK, Germany, Canada, and Israel clustering together indicating that they are more likely to receive conflict from the same countries.  The  cluster of democratic and authoritarian countries facing each other in the inner and outer rings indicate that they are more likely to engage in conflict with one another.

Perhaps most critically, an evaluation of this visualization highlights Weeks' original expectation of how states behave in the conflict system.  Specifically, Iraq, Syria, Libya, and North Korea all fell under Weeks' ``boss" category, and each of these states tends to cluster together in the inner and outer rings. This indicates that even though we do not find support for Weeks' assertion that certain authoritarian regime types are more likely to initiate conflict, we do find that these regimes are more likely to behave similarly in terms of who they target and receive conflict from.

\subsection{Gibler (2017)}

The last replication we conduct with the AME model considers a study by \citet{gibler:2017}. Gibler argues that the long-standing relationship between the relative parity of capabilities and initiation of militarized interstate disputes (MIDs) is almost completely mediated by the initial conditions for the members of the dyad when they joined the international system as sovereign members. In most specifications, after taking into account the initial conditions for dyadic entry, the statistical significance of power parity vanishes. This finding calls into question many IR theories about the role of balance in generating international conflict \citep{organski:1958}.

To test this hypothesis, Gibler employs an undirected dyadic design and estimates his model using a GLM with dyad-clustered standard errors.\footnote{Full tabular results are shown in the appendix, we focus on model 6 from Table 6 (2017, 34).} With this design, Gibler finds support for both the insignificance of contemporary power parity, and the effect of the initial conditions for entry on driving conflict behavior. When we re-estimate using AME power parity still has a small, imprecisely measured effect, supporting Gibler's argument that contemporary power parity is an artifact of other aspects of state relations, and does not drive contemporary conflict behavior. At the same time, the AME approach finds that the effect of power parity when a country enters into the international system is too imprecisely measured to draw any inferences about its effect on the initiation of MIDs. Not only is our estimate of the effect of this variable small, but it has a very large relative standard error---over a magnitude larger than the parameter itself. Thus, while we can confirm the argument that the effect of power parity vanishes when accounting for both initial conditions and network dependencies, we find less support for the argument that initial conditions drive the initiation of MIDs.

In the two previous replications, we showed how to parse apart underlying structure using the SRM and LFM portions of the AME framework. Here we turn our focus to parameter interpretation in a substantive context. Specifically, even when the GLM and AME frameworks produce results that may seem to be in accordance with one another, the substantive interpretations of the effects of covariates can differ notably since the AME model uses a set of random effects to account for unobserved factors. To explore this, we focus on the effect of rivalry on MID initiation. Both the GLM and AME estimations find that rivalry has a positive effect on MID initiation, but the expected effects between the two models differs greatly.

To clarify the difference, we turn to a simulation based approach. We employ mean or modal values for all independent variables, except we change the rivalry variable to indicate that there was a rivalry when the actual data suggest there is none. This provides us with two scenarios, one in which rivalry is set to one and the other zero, while in both scenarios all other parameters are set to their measure of central tendency. The expected values of this scenario are essentially a first difference plot comparing results with the model when estimated in two different ways: the GLM estimation and our AME approach.

\begin{figure}
	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
	\includegraphics[width=\textwidth]{graphics/figure9.pdf}
 	\label{fig:gibmargeff}
 \end{figure}

As Figure~\ref{fig:gibmargeff} illustrates, the substantive results differ notably. The expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.  These are rare events, so the probabilities are low, but the difference in predictions is notable.

\subsection{Lessons Learned}

First, utilizing the AME framework enables scholars to better model the underlying structure inherent to dyadic data. In each of the models the AME substantially outperforms the original model out of sample. Not only does AME perform better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves). This may be because by ignoring dependencies, the original models are misspecifying the DGP, and the AME better accounts for it.

Second, in both the simulation  and replication sections  we have shown that results can change notably when interdependencies are not taken into account. Not only are coefficients biased in the GLM approaches, but they are often imprecisely measured with poorly calibrated standard errors.  This means that significance testing (for better or worse) is compromised when network effects are ignored.

Third, even when the results from the AME estimation conform with those found in an OLS or logistic regression, new insights can emerge from the structure uncovered by the random effects of the AME framework. In particular, the AME provides information about the dependencies so that clusters can be identified.  This information can then be used to generate new hypotheses.
%These kinds of insights are absent in standard approaches and add to our ability to explain specific as well as general results.

Fourth, it is evident that the actual results---not the estimated coefficients and their covariances---generated by the models differ greatly in expectations.  This implies that policy experimentations with the models, as well as scenario-based simulations and forecasting of GLM models are likely to give misleading results compared to the AME approach.

\section{\textbf{Conclusion}}

International relations is about the interactions, relationships, and dependencies among countries or other important international actors. This is particularly true of scholarship in the tradition of the Correlates of War Project, but it is by no means limited to it.\footnote{See \cite{singer:1972} for an early description of the project and also see the project's Web site for an history and more recent efforts \url{http://www.correlatesofwar.org/}.} Many scholars have debated the use and abuse of dyadic data.\footnote{One recent on-line symposium can be found at \url{http://bit.ly/2wB2hab}.} A broad survey of the IR literature makes it clear that scholars find dyadic data to be an essential touchstone in the study of international relations \citep{erikson:pinto:2014,aronow:etal:2015}. Our findings bolster a growing recognition in the field of International Relations that interdependence influences not only statistical estimations, but how scholars theorize about internationally relevant politics. Scholars have demonstrated the theoretical importance of interdependence through research on intrastate conflict \citep{dorff:etal:2020}, interstate bargaining \citep{gallop:2017}, economic interdependence \citep{maoz:2009a}, and international treaties \citep{kinne:2013} among other topics.

At the same time, we know that research designs focusing on the statistical analysis of dyadic data quickly go astray if the dyadic data are assumed to be iid.  Virtually all of the standard statistical models---ordinary least squares and logistic regressions, to name a few---fail if the data are not conditionally independent. This fact has been accepted as it relates to temporal dependencies, but adoption of methods to account for network dependencies have seen less progress. By definition dyadic data are not iid and thus the standard approaches cannot be used cavalierly to analyze these data. \citet{signorino:1999} showed why this is true of models of strategic interaction, but it is more broadly true of models that employ dyadic data.  We show that the AME framework can be employed to account for the statistical issues that arise when studying dyadic data.

To explore this approach in the context of international relations we have presented two analyses. The first is a simulation where the characteristics of the network are known. This shows that when there are unobserved dependencies, the AME approach is less biased in terms of parameter estimation compared to the standard approach employed in international relations to study dyadic data (i.e., GLM models). The second analysis is a replication of recent studies that use a broad range of dyadic data to draw inferences about international relations.  These studies have been replicated with the original research designs, each of which used a statistical method that assumes the dyadic data are all independent from one another.  We then re-analyzed each study using the AME model.  In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, and c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.}

% It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored. Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citealp{king:roberts:2014}). Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences, better out-of-sample predictive performance, and new substantive insights.

\newpage
% Bib stuff
\clearpage

% \bibliography{C:/Users/Owner/whistle/master}
\bibliography{C:/Users/S7M/whistle/master}
% \bibliography{C:/Users/herme/whistle/master}
%\bibliography{/Users/s7m/whistle/master}
\bibliographystyle{authordate1}
% \bibliographystyle{sm}
% \clearpage
% \section*{All About Us}

\appendix

\clearpage

\renewcommand{\thefigure}{\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{\thesection \arabic{table}}
\setcounter{table}{0}
%\renewcommand{\thesection}{A.\arabic{section}}
\setcounter{section}{0}

\appendix
\section{AME Monte Carlo Markov Chain Algorithm}

Given initial values of $\{\bm\beta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}\}$, the algorithm proceeds as follows until convergence:\footnote{Further details on the full conditional distributions and estimation approach underlying AME can be found in \citet{hoff:2021}.}

 \begin{itemize}
  \item sample $\bm\theta \; | \;  \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\bm\beta \; | \;  \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\mathbf{a}, \mathbf{b} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
  \item sample $\Sigma_{ab} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Inverse-Wishart)
  \item update $\rho$ using a Metropolis-Hastings step with proposal $p^{*} | p  \sim$ truncated normal$_{[-1,1]}(\rho, \sigma_{\epsilon}^{2})$
  \item sample $\sigma_{\epsilon}^{2} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \text{ and } \rho$ (Inverse-Gamma)
  \item For each $k \in K$:
  \begin{itemize}
    \item Sample $\mathbf{U}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}_{[,-k]}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
    \item Sample $\mathbf{V}_{[,k]} \; | \; \bm\beta, \mathbf{X}, \bm\theta, \mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}_{[,-k]}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)
    \item Sample $\mathbf{D}_{[k,k]}  \; | \; \bm\beta, \mathbf{X}, \bm\theta,\mathbf{a}, \mathbf{b}, \mathbf{U}, \mathbf{V}, \Sigma_{ab}, \rho, \text{ and } \sigma_{\epsilon}^{2}$ (Normal)\footnote{Subsequent to estimation, $\mathbf{D}$ matrix is absorbed into the calculation for $\mathbf{V}$ as we iterate through $K$. }
  \end{itemize}
 \end{itemize}
\newpage

\section{Additional Replication Information}

For each of the replications involving a binary dependent variable we provide a table of coefficient estimates that includes the original GLM estimation and the results from the AME model.  We have adopted for presentation purposes only the use of significance testing on these observational data for facile comparison with the replicated studies.  However, note that we do provide predictive heuristics for these models as well.

% Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis. %For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.

\clearpage
\subsection*{Reiter \& Stam (2003)}

Additional information for the Reiter \& Stam (2003) re-estimation.

\input{graphics/appendix_tableB1}
\FloatBarrier

% \begin{figure}
% 	\centering
% 	\subfigure[AUC]{\label{fig:reiterstamroc}\includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf}}
% 	\subfigure[Precision and Recall]{\label{fig:reiterstampr}\includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf}}
% 	\caption{Assessments of out-of-sample predictive performance for Reiter \& Stam (2003) using ROC curves, PR curves, and separation plots.}
% \end{figure}
\FloatBarrier
\clearpage

\subsection*{Weeks (2012)}

Additional information for the Weeks (2012) re-estimation.

\input{graphics/appendix_tableB2}
\FloatBarrier

% \begin{figure}
% \centering
%   \subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
%   \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
%   \caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
% \end{figure}
\FloatBarrier
\clearpage

\subsection*{Gibler (2017)}

Additional information for the Gibler (2017) re-estimation.

\input{graphics/appendix_tableB3}
\FloatBarrier

% \begin{figure}
% 	\centering
% 	\subfigure[AUC]{\label{fig:giblerroc}\includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf}}
% 	\subfigure[Precision and Recall]{\label{fig:giblerpr}\includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf}}
% 	\caption{Assessments of out-of-sample predictive performance for Gibler (2017) using ROC curves, PR curves, and separation plots.}
% \end{figure}
\FloatBarrier
\clearpage

\section{Additional Simulations: Non-Linear Data}

Aronow et al note that ``fitting a linear approximation to mildly non-linear data'' can be problematic for dyadic random effect models. We agree on the importance of testing this in the context of the simulation that we posed in the paper. Aronow et al incorporate non-linear misspecification by adding in squared version of the key dyadic homophily variable to the DGP for their simulations: $Y_{i,j} = \beta_{0} + \beta_{1} X_{i,j} + \beta_{2} X_{i,j}^{2} + \epsilon_{i,j}$. Here $X$ is their representation of homophily and is considered observed, while $X^{2}$ is their check for robustness to misspecification and is unobserved. We incorporate the same type of misspecification into our probit simulation setup: $Z_{i,j} =  \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j}$, where $W$ now instead of being drawn from a separate normal distribution is just the squared version of $X$. For the simulation we set values of $\mu$, $\beta$, and $\gamma$ to -2, 1, and 1, respectively. Results are shown below in Figures~\ref{fig:ameBias_asa} and \ref{fig:ameCalib_asa} below. As in the paper, along the x-axis we implement a standard model that does not take any steps to account for dependencies, AME, and the Oracle model, and as in the paper we find that the AME approach provides a number of benefits in terms of bias and coverage over the standard approach of not dealing with dependencies.

\begin{figure}[ht]
	\centering
	\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points.}
	\label{fig:ameBias_asa}
	\includegraphics[width=1\textwidth]{graphics/appendix_figureC1.pdf} \\
\end{figure}

\begin{figure}[ht]
	\centering
	\caption{Proportion of times the true value fell within the estimated 95\% confidence interval for the standard, AME, and oracle models from 1,000 simulations.}
	\label{fig:ameCalib_asa}
	\includegraphics[width=1\textwidth]{graphics/appendix_figureC2.pdf} \\
\end{figure}

\FloatBarrier
\clearpage

\section{Additional Simulations: Correlation with Omitted Variable}

Estimation certainly gets more challenging when the omitted variable is correlated with the observed variable. In the case of normal linear regression, we can show that in general a regression coefficient is unbiased for its ``unconditional effect'', which is the direct effect plus something related to the direct effects of any omitted variables that it is correlated with. For example, suppose in our simulation setup $X_{i,j}$ is correlated with $W_{i,j}$ via the relation $W_{i,j} = \alpha X_{i,j} + e_{i,j}$. Then:

\begin{align*}
  Y_{i,j} & = \mu + \beta X_{i,j} + \gamma W_{i,j} + \epsilon_{i,j} \\
  &= \mu + (\gamma \alpha+\beta ) X_{i,j} + (\gamma Z_{i,j} + \epsilon_{i,j})  \\
  &= \mu + \tilde \beta X_{i,j} + \tilde \epsilon_{i,j}
\end{align*}

The naive model is in some sense still ``correct'' as it is an unbiased estimator of the unconditional (direct plus indirect) effect $\gamma\alpha + \beta$ of $X$ on $Y$. However, there is no way to eliminate this extra bias without making further assumptions about the nature of the omitted variable. Additonally, in the binary case with the probit setup, other sources of bias get introduced as well because of the nonlinearity of the population moments as a function of the parameters. An interesting question is whether or not we can reduce the bias in the network setting by assuming the omitted variable $w_{i,j}$ is the product of latent node-specific factors. An idea related to this is discussed in \citet{minhas:etal:2017:arxiv}.

For now though to show the amount of bias that gets introduced into our setup when $X$ and $W$ are correlated we repeat our simulation exercise but induce varying levels of correlation between $X$ and $W$. Results with $X$ and $W$ correlated at 0.4 and then 0.7 are shown below, respectively.

\begin{figure}[ht]
\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.4 in this case.}
\label{fig:ameBias_corrMed}
\includegraphics[width=1\textwidth]{graphics/appendix_figureD1a.pdf} \\
\end{figure}

\begin{figure}[ht]
\caption{Regression parameter estimates for the standard, AME, and oracle models from 1,000 simulations. Summary statistics are presented through a traditional box plot, and the estimates from each simulation are visualized as well as points. $X$ and $W$ are correlated at 0.7 in this case.}
\label{fig:ameBias_corrHi}
\includegraphics[width=1\textwidth]{graphics/appendix_figureD1b.pdf} \\
\end{figure}

\FloatBarrier
\clearpage

\section{AME Tutorial}

Using the AMEN function requires formatting data into a particular structure. The primary distinction in data formatting is whether the outcome of interest represents a directed or undirected network.

If undirected, the AMEN function has three main inputs:

\begin{itemize}[noitemsep,nolistsep]
    \item Y: a $T$ length \textbf{list} of $n\times n$ adjacency matrices, where $T$ = number of years in the dataset and $n$ = number of nodes in the network.
    \begin{itemize}
      \item An adjacency matrix describes relationships between nodes in a particular year of data. For example, in an adjacency matrix of interstate MIDs, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had a MID in that year, and '0' otherwise. The diagonal in the adjacency matrix is typically missing. Y is a list of these adjacency matrices for the outcome variable, where each element in the list is a different year of data.
    \end{itemize}
    \item Xdyad: a $T$ length \textbf{list} of $n\times n\times p$ arrays, where $p$ = number of dyadic covariates in dataset.
    \begin{itemize}
      \item An array is a data object in R that contains a series of 'stacked' matrices for each year of data. An array of dimension (2, 3, 4), for example, contains 4 rectangular matrices each with 2 rows and 3 columns. Each matrix in the array describes the relationship between nodes with respect to some covariate. For example, for interstate alliances, row $i$ and column $j$ takes a value of '1' if country $i$ and country $j$ had an alliance in that year, and '0' otherwise. An array contains a matrix of this kind for each covariate going into the model. Xdyad is a list of these arrays, where each element in the list is a different year of data.
    \end{itemize}
    \item Xrow: a $T$ length \textbf{list} of $n\times p$ matrices, where $p$ = number of monadic (nodal) covariates in dataset.
    \begin{itemize}
      \item Each matrix in Xrow has the nodes running down the rows and the covariates in the dataset running along the columns. An entry in the matrix captures the value that a node takes on for each covariate in a given year. For example, if column $j$ is GDP per capita and column $k$ is population, then row $i$, column $j$ measures country $i$'s GDP per capita and row $i$, column $k$ measures country $i$'s population in a given year. Xrow is a list of these matrices, where each element in the list is a different year of data.
    \end{itemize}
\end{itemize}

If directed, AMEN further requires:

\begin{itemize}[noitemsep,nolistsep]
    \item Xrow: a $T$ length list of $n\times p$ matrices, where $p$ = number of sender (nodal) covariates in dataset.
    \begin{itemize}
      \item Xrow here is nearly identical to the undirected case. The only difference is that rather than each matrix containing covariate data on all nodes, in the directed context each matrix only contains data on the nodes that are acting as senders in that particular year. Using interstate MIDs as an example, if in 1983 only two countries initiated MIDs, Xrow would only contain data for those two countries (i.e., Xrow would only have two rows).
    \end{itemize}
    \item Xcol: a $T$ length list of $n\times p$ matrices, where $p$ = number of receiver (nodal) covariates in dataset.
    \begin{itemize}
      \item Xcol here is nearly identical to Xrow, except each matrix contains data on receiver nodes, not sender nodes. Using interstate MIDs as an example, if in 1983 only two countries had MIDs initiated against them, Xcol would only contain data for those two countries (i.e., Xcol would only have two rows).
    \end{itemize}
\end{itemize}

Beyond the data inputs, the AMEN function requires additional specification:

\begin{itemize}
    \item model: how to model the outcome variable, e.g., `logit'
    \item symmetric: whether the input network is symmetric
    \item intercept: whether to estimate an intercept
    \item nscan: number of iterations of the Markov chain
    \item burn: burn-in period
    \item odens: thinning interval
    \item R: dimension of the multiplicative effect (referred to as $K$ in the paper)
    \item gof: whether to calculate goodness of fit statistics
\end{itemize}

There is often little theoretical reason to choose a particular value of $\sf{R}$ (above). One strategy is to estimate models at different values of $\sf{R}$ and compare goodness of fit statistics across models. Given the computational time needed for parameter estimates to converge, parallelization strategies are recommended to speed up analysis. In addition, providing AMEN function with starting values, either dictated by theory, previous research, or previous runs can also help speed up convergence time.

The code below presents an example of an AME model running in parallel across 4 different levels of $\sf{R}$. Note also that the model is using starting values from a previous run, defined in \textit{startVals}.

\begin{lstlisting}[language=R]
# running in parallel varying k
imps = 10000 ; brn = 25000 ; ods = 10 ; latDims = 0:3

# Run amen in parallel
library(doParallel) ; library(foreach) ; cl=makeCluster(4) ; registerDoParallel(cl)
foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {

  # load previous model run
  load(prevModelFiles[ii])
  # extract start vals
  startVals0 = ameFit$'startVals'
  # dump rest
  rm(ameFit)

  ameFit = ame_repL(
    Y=yList,Xdyad=xDyadList,Xrow=NULL,Xcol=NULL,
    model="bin",symmetric=FALSE,intercept=TRUE,R=latDims[ii],
    nscan=imps, seed=1, burn=brn, odens=ods,
    plot=FALSE, print=FALSE, gof=TRUE, startVals=startVals0,
    periodicSave=TRUE )
  save(ameFit, file=paste0('model_k', latDims[ii],'_v2.rda') )
}
stopCluster(cl)
\end{lstlisting}

\end{document}

### 2021-07-19 diffNotes/netsMatter_v1.tex
highlight diffs for psrm

--- 2021-07-19 diffNotes/netsMatter_v2.tex
@@
 
+ % in dyadic data that, if left un-estimated, would complicate any inferences we might wish to draw for the exogenous parameters.
+By integrating the SRRM and LFM into a Bayesian probit framework, we can account for the types of underlying structure discussed in this section.
+
 Non-iid observations in relational data result from the fact that there is a complex structure underlying the dyadic events or processes that we observe.
@@
 	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{graphics/figure3.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimBias_all.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{graphics/figure4.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimCover_all.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCorr}
-	\includegraphics[width=1\textwidth]{graphics/figure5.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimCorr.pdf} \\
 \end{figure}
@@
 		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{graphics/reiter_stam_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/weeks_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{graphics/weeks_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{graphics/gibler_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{graphics/gibler_pr_outSample.pdf} \\
 	\end{tabular}
@@
 	\begin{tabular}{cc}
-	\includegraphics[width=.7\textwidth]{graphics/figure7a.pdf} &
-	\includegraphics[width=.3\textwidth]{graphics/figure7b.pdf} \\
+	\includegraphics[width=.7\textwidth]{graphics/reiter_stam_aEff_map.pdf} &
+	\includegraphics[width=.3\textwidth]{graphics/reiter_stam_aEff_line.pdf} \\
 	\end{tabular}
@@
 The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
-Specifically, we can see that countries such as Iraq, Israel, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
-Further, other countries such as Sweden, Finland, and Swaziland are less likely to engage in conflict than the exogenous covariates in the model would suggest.
+Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
+Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
 In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for \citet{colgan:2010}'s theory that revolutionary petrostates are more aggressive.
@@
 There is also a striking difference between Weeks' original model and our estimation using AME in terms of capturing the data generating process.
+Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same exogenous covariates.
 As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest.
@@
 \begin{figure}[!h]
-	\includegraphics[width=\textwidth]{graphics/figure8.pdf}
+	\includegraphics[width=\textwidth]{graphics/weeks_circPlot.png}
 	\caption{ Visualization of multiplicative effects for Weeks (2012).
@@
 Figure~\ref{fig:weekscirc} reveals a number of notable clusters.
-For example, in the bottom right corner we see the US, UK, Germany, Canada, and Israel.
+For example, in the top left corner we see the US, UK, Germany, Canada, and Israel.
 These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.
-Conversely, in the top left of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
+Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
 We observe similar clusters in the inner ring.
@@
 	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
-	\includegraphics[width=\textwidth]{graphics/figure9.pdf}
+	\includegraphics[width=\textwidth]{graphics/gibler_margeff.pdf}
  	\label{fig:gibmargeff}
@@
 The expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.
-These are rare events, so the probabilities are low, but the difference in predictions is notable.
+These are rare events, so the probabilities are low, but the difference is a factor of almost two.
+Thus, we find substantially different expected values from these two models.
 
@@
 
-% It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.
-Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citealp{king:roberts:2014}).
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences, better out-of-sample predictive performance, and new substantive insights.
-
 \newpage
@@
 % \bibliography{C:/Users/herme/whistle/master}
+% \bibliography{/Users/maxgallop/documents/whistle/master}
+% \bibliography{/Users/howardliu/documents/whistle/master}
 %\bibliography{/Users/s7m/whistle/master}
+% \bibliography{/Users/mdw/Git/whistle/master}
+  %\bibliography{/Users/dorffc/ProjectsGit/whistle/master}
 \bibliographystyle{authordate1}
@@
 
-\input{graphics/appendix_tableB1}
+\input{reiter_stam_coeftable}
 \FloatBarrier
@@
 
-\input{graphics/appendix_tableB2}
+\input{weeks_coeftable}
 \FloatBarrier
@@
 
-\input{graphics/appendix_tableB3}
+\input{gibler_coeftable}
 \FloatBarrier
@@
 	\label{fig:ameBias_asa}
-	\includegraphics[width=1\textwidth]{graphics/appendix_figureC1.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_asaProbit.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCalib_asa}
-	\includegraphics[width=1\textwidth]{graphics/appendix_figureC2.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/ameSimCover_all_asaProbit.pdf} \\
 \end{figure}
@@
 \label{fig:ameBias_corrMed}
-\includegraphics[width=1\textwidth]{graphics/appendix_figureD1a.pdf} \\
+\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_corrProbitMed.pdf} \\
 \end{figure}
@@
 \label{fig:ameBias_corrHi}
-\includegraphics[width=1\textwidth]{graphics/appendix_figureD1b.pdf} \\
+\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_corrProbitHi.pdf} \\
 \end{figure}

### 2022-08-03 paper/4_app.tex
cleanup

@@
 The visualization of the sender random effects highlights that the behavior of countries does not fully accord with predictions from the covariates specified by Reiter \& Stam.
-Specifically, we can see that countries such as Iraq, Israel, Egypt, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
-Further, other countries such as Bhutan, Finland, and Turkmenistan are less likely to engage in conflict than the exogenous covariates in the model would suggest.
+Specifically, we can see that countries such as Iraq, Israel, and Iran are more likely to be involved in initiating or continuing conflict with other countries than the model would predict.
+Further, other countries such as Sweden, Finland, and Swaziland are less likely to engage in conflict than the exogenous covariates in the model would suggest.
 In this case, the finding that countries in the middle east experience more conflict with other countries might lead one to more carefully examine the effects of geography on conflict initiation or to account for \citet{colgan:2010}'s theory that revolutionary petrostates are more aggressive.
@@
 There is also a striking difference between Weeks' original model and our estimation using AME in terms of capturing the data generating process.
-Specifically, the area under the ROC curve for AME and GLM models are 0.97 and 0.64, respectively, which is a dramatic difference given that both models use the same exogenous covariates.
 As with Reiter \& Stam, the divergence is a result of the GLM framework's inability to account for underlying structure generating the event of interest.
@@
 \begin{figure}[!h]
-	\includegraphics[width=\textwidth]{graphics/weeks_circPlot.png}
+	\includegraphics[width=\textwidth]{graphics/figure8.pdf}
 	\caption{ Visualization of multiplicative effects for Weeks (2012).
@@
 Figure~\ref{fig:weekscirc} reveals a number of notable clusters.
-For example, in the top left corner we see the US, UK, Germany, Canada, and Israel.
+For example, in the bottom right corner we see the US, UK, Germany, Canada, and Israel.
 These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.
-Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
+Conversely, in the top left of the outer ring, we see a cluster of authoritarian countries: Iraq, Russia, Syria, North Korea, and China.
 We observe similar clusters in the inner ring.
@@
 The expected value of the dependent variable---the probability of the onset of a militarized interstate dispute, is considerably lower once interdependencies are taken into account with the AME model.
-These are rare events, so the probabilities are low, but the difference is a factor of almost two.
-Thus, we find substantially different expected values from these two models.
+These are rare events, so the probabilities are low, but the difference in predictions is notable.
 

### 2022-08-03 paper/6_appendix.tex
cleanup

@@
 
-\input{reiter_stam_coeftable}
+\input{graphics/appendix_tableB1}
 \FloatBarrier
@@
 
-\input{weeks_coeftable}
+\input{graphics/appendix_tableB2}
 \FloatBarrier
@@
 
-\input{gibler_coeftable}
+\input{graphics/appendix_tableB3}
 \FloatBarrier
@@
 	\label{fig:ameBias_asa}
-	\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_asaProbit.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/appendix_figureC1.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCalib_asa}
-	\includegraphics[width=1\textwidth]{graphics/ameSimCover_all_asaProbit.pdf} \\
+	\includegraphics[width=1\textwidth]{graphics/appendix_figureC2.pdf} \\
 \end{figure}
@@
 \label{fig:ameBias_corrMed}
-\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_corrProbitMed.pdf} \\
+\includegraphics[width=1\textwidth]{graphics/appendix_figureD1a.pdf} \\
 \end{figure}
@@
 \label{fig:ameBias_corrHi}
-\includegraphics[width=1\textwidth]{graphics/ameSimBias_all_corrProbitHi.pdf} \\
+\includegraphics[width=1\textwidth]{graphics/appendix_figureD1b.pdf} \\
 \end{figure}

### 2022-08-03 paper/toSubmit/netsMatter.tex
cleanup

--- 2021-07-19 diffNotes/netsMatter_v2.tex
@@
 \makeatletter
-% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
-% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/Graphics/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/Graphics/}}
+% \def\input@path{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/}}
+% \graphicspath{{/Users/janus829/Dropbox/toDo_notes/nsfSumm/}, {/Users/s7m/Dropbox/toDo_notes/nsfSumm/}}
 
@@
 
-% Scholars working with dyadic data typically begin by stacking observations associated with each dyad on top of one another and not passing any auxiliary information to the model about how observations may be interdependent.
-For example, in this design, a conflict initiated from the United States against Japan is assumed to be independent of any conflictual action that Japan may send to the United States.
-Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan.
-While most scholars begin with the assumption that each dyadic interaction takes place in isolation of the others, we know this assumption to be false both in theory and in practice.
-Relational data comes with an explicit structure that generally leads to particular types of dependencies.
-The importance of accounting for the underlying structure of our data has been a lesson well understood, at least when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}.
-As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause.
-
 To uncover the underlying structure of relational data, it is helpful to restructure dyadic data in the form of a matrix---often referred to as an adjacency matrix---as shown in Figure~\ref{fig:adjMatDeps}.
@@
 		\textsc{Sender Heterogeneity} & \textsc{Receiver Heterogeneity} \\
-		\includegraphics[width=.4\textwidth]{graphics/adjRowDep.png} &
-		\includegraphics[width=.4\textwidth]{graphics/adjColDep.png} \\
+		\includegraphics[width=.4\textwidth]{adjRowDep.png} &
+		\includegraphics[width=.4\textwidth]{adjColDep.png} \\
 		\textsc{Sender-Receiver Covariance} & \textsc{Reciprocity} \\
-		\includegraphics[width=.4\textwidth]{graphics/adjRowColCovar.png} &
-		\includegraphics[width=.4\textwidth]{graphics/adjRecip.png} \\
+		\includegraphics[width=.4\textwidth]{adjRowColCovar.png} &
+		\includegraphics[width=.4\textwidth]{adjRecip.png} \\
 	\end{tabular}
@@
 \begin{figure}[ht]
-	\includegraphics[width=.6\textwidth]{graphics/stochEquiv_v2.pdf}
+	\includegraphics[width=.6\textwidth]{stochEquiv_v2.pdf}
 	\caption{Higher-order dependence patterns in a network.}
@@
 While sometimes the conditions driving these patterns, such as geography, are easy to identify, at other times it can be difficult to describe exactly why higher order dependence patterns in networks may develop.
-
-% International Trade, we might observe homophily wherein states with open economies and an absence of Non-Tariff Barriers are more likely to trade with each-other all else equal; alternatively we might observe heterophily in the sense that states that have greater endowments of capital are more likely to trade with states with greater endowments of land or labor.
-
-%  We would also likely observe stochastic equivalence in an interstate war network wherein countries who are in conflict with the same target country are unlikely to fight one another.
-In this case, countries with (formal or informal) alliances are a member of the same group due to similar conflictual ties to other actors in the system.
 
@@
 	\label{fig:ameBias}
-	\includegraphics[width=1\textwidth]{graphics/figure3.pdf} \\
+	\includegraphics[width=1\textwidth]{figure3.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCalib}
-	\includegraphics[width=1\textwidth]{graphics/figure4.pdf} \\
+	\includegraphics[width=1\textwidth]{figure4.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCorr}
-	\includegraphics[width=1\textwidth]{graphics/figure5.pdf} \\
+	\includegraphics[width=1\textwidth]{figure5.pdf} \\
 \end{figure}
@@
 		\multicolumn{2}{l}{\textbf{\tiny{Reiter \& Stam (2003)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/figure6_reiter_stam_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{figure6_reiter_stam_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{figure6_reiter_stam_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{\tiny{Weeks (2012)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/figure6_weeks_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{figure6_weeks_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{figure6_weeks_pr_outSample.pdf} \\
 		\multicolumn{2}{l}{\textbf{\tiny{Gibler (2017)}}} \\
-		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_roc_outSample.pdf} &
-		\includegraphics[width=.4\textwidth]{graphics/figure6_gibler_pr_outSample.pdf} \\
+		\includegraphics[width=.4\textwidth]{figure6_gibler_roc_outSample.pdf} &
+		\includegraphics[width=.4\textwidth]{figure6_gibler_pr_outSample.pdf} \\
 	\end{tabular}
@@
 	\begin{tabular}{cc}
-	\includegraphics[width=.7\textwidth]{graphics/figure7a.pdf} &
-	\includegraphics[width=.3\textwidth]{graphics/figure7b.pdf} \\
+	\includegraphics[width=.7\textwidth]{figure7a.pdf} &
+	\includegraphics[width=.3\textwidth]{figure7b.pdf} \\
 	\end{tabular}
@@
 \begin{figure}[!h]
-	\includegraphics[width=\textwidth]{graphics/figure8.pdf}
+	\includegraphics[width=\textwidth]{figure8.pdf}
 	\caption{ Visualization of multiplicative effects for Weeks (2012).
@@
 	\caption{Marginal effects of a change in the Rivalry variable for both the AME and the Gibler estimation.  \label{fig:gibmargeff}}
-	\includegraphics[width=\textwidth]{graphics/figure9.pdf}
+	\includegraphics[width=\textwidth]{figure9.pdf}
  	\label{fig:gibmargeff}
@@
 In every case, we found that the AME approach provided a) increased precision of estimation, b) better out-of-sample fit, and c) evidence of 1st-, 2nd-, and 3rd-order dependencies that were overlooked in the original studies.\footnote{The Appendix contains performance data on all of these replications, as well as sample code illustrating how to undertake AME analysis using \texttt{amen}.}
-
-% It is no longer necessary to assume that the interesting, innate interdependencies in relational data about international relations can be ignored.
-Nor do they have to be approximated with \textit{ad hoc}, incomplete solutions that purport to control for dependencies (such as modifying the post-estimation standard errors of the estimated coefficients \citealp{king:roberts:2014}).
-Instead, the interdependencies may be addressed directly with additive and multiplicative effects in the context of a generalized linear model that provides more reliable inferences, better out-of-sample predictive performance, and new substantive insights.
 
@@
 
-\input{graphics/appendix_tableB1}
+\input{appendix_tableB1}
 \FloatBarrier
@@
 
-\input{graphics/appendix_tableB2}
+\input{appendix_tableB2}
 \FloatBarrier
@@
 
-\input{graphics/appendix_tableB3}
+\input{appendix_tableB3}
 \FloatBarrier
@@
 	\label{fig:ameBias_asa}
-	\includegraphics[width=1\textwidth]{graphics/appendix_figureC1.pdf} \\
+	\includegraphics[width=1\textwidth]{appendix_figureC1.pdf} \\
 \end{figure}
@@
 	\label{fig:ameCalib_asa}
-	\includegraphics[width=1\textwidth]{graphics/appendix_figureC2.pdf} \\
+	\includegraphics[width=1\textwidth]{appendix_figureC2.pdf} \\
 \end{figure}
@@
 \label{fig:ameBias_corrMed}
-\includegraphics[width=1\textwidth]{graphics/appendix_figureD1a.pdf} \\
+\includegraphics[width=1\textwidth]{appendix_figureD1a.pdf} \\
 \end{figure}
@@
 \label{fig:ameBias_corrHi}
-\includegraphics[width=1\textwidth]{graphics/appendix_figureD1b.pdf} \\
+\includegraphics[width=1\textwidth]{appendix_figureD1b.pdf} \\
 \end{figure}
